### 2023-04-20 Experiment_1/analysis.Rmd
initialize repository

---
title: "Analysis experiment 1"
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-09"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
```

```{r, message=FALSE}
d <- read_csv("./data/cleaned.csv")
```

### Some themes and functions

```{r, echo=FALSE}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)

#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

### Hypothesis 1

To assess the effect of convergence on participants' confidence about their guesses, we run a paired t-test.

```{r}
# summary stats
d %>%
  group_by(convergence) %>%
  get_summary_stats(confidence, type = "mean_sd")
```

```{r}
# T-test
t <- t.test(confidence ~ convergence, data = d, paired = TRUE)
t
```

The paired t-test shows a significant difference in confidence ratings between *no convergence* and *convergence*. In convergent conditions, participants were on average more confident by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7).

For a robustness check, we run a mixed model regression analysis of convergence on confidence ratings. The first model (`m1`) includes random intercepts for participants. The second model (`m2`) additionally controls for a fixed effect of our second experimental factor, the number of estimates. Including the factor number as a control will also address RQ1. The third model (`m3`) additionally includes random slopes for participants.

For reference, we also include fixed-only effects regression models.

```{r, echo=FALSE}
# Note that in this table we will include models for both outcome variables

# fixed effects only

# baseline
lm_accuracy <- lm(confidence ~ convergence, data= d)
lm_competence <- lm(competence ~ convergence, data= d)
# control
lm_control_accuracy <- lm(confidence ~ convergence + number, data= d)
lm_control_competence <- lm(competence ~ convergence + number, data= d)

# mixed effects

# first model with random intercept per participant
m1_accuracy = lmer(confidence ~ convergence + (1|ID), data = d)
m1_competence = lmer(competence ~ convergence + (1|ID), data = d)

# second model which controls for second experimental factor "number"
m2_accuracy = lmer(confidence ~ convergence + number + (1|ID), data = d)
m2_competence = lmer(competence ~ convergence + number + (1|ID), data = d)

# third model which adds a random slope (for convergence effect) per participant
m3_accuracy = lmer(confidence ~ convergence + number + (1 + convergence | ID), data = d)
m3_competence = lmer(confidence ~ convergence + number + (1 + convergence | ID), data = d)
```

```{r, echo=FALSE}
# select models with different outcome variables
panels_accuracy <- list(
  "Baseline" = lm_accuracy,
  "Baseline + control" = lm_control_accuracy, 
  "Random intercept" = m1_accuracy, 
  "Random intercept + control" = m2_accuracy, 
  "Random intercept & slope + control" = m3_accuracy
)

panels_competence <- list(
  "Baseline" = lm_competence,
  "Baseline + control" = lm_control_competence, 
  "Random intercept" = m1_competence, 
  "Random intercept + control" = m2_competence, 
  "Random intercept & slope + control" = m3_competence
)

panels <- list(
  "Outcome: Accuracy" = panels_accuracy,
  "Outcome: Competence" = panels_competence
)

modelsummary(panels, 
             shape = "rbind",
             stars = TRUE, 
             statistic = 'p.value',
             title = 'Models estimating effect of convergence on accuracy and competence', 
             gof_omit = "IC|Log|Adj|p\\.value|statistic|se_type") %>%
    add_header_above(c(" " = 1, "Fixed" = 2, "Mixed" = 3))
```

The first linear regression model (`Baseline`) yields the same estimate for the effect of `convergence` as the paired t-test. The second model (`Baseline+Control`) controls for `number`. The effect of convergence remains the same, but there is an additional effect of `number`, statistically significant at the 5%-level: Conditions with a number of estimates increased participants' confidence ratings by 0.18 (on a scale from 1 to 7) on average, holding `convergence` constant.

Next, we calculate the mixed effects models. There will be the "fixed", average effects of `convergence` and `number` that we just found in the standard regression models (also called population-level effects, averages across *all* participants). In addition, there will also be random effects by participants.

We have two extra random parameters:

1.  The estimated standard deviation of the by-participant random intercept, represented by the `SD (Intercept ID)`.
2.  The residual variation (i.e. unexplained variation in `competence` once we account for all effects specified in the model), represented by the `SD (Observations)` term in the Residual group.

The first parameter tells how much `competence` ratings bounce around as we move from participant to participant. We can predict `competence` based on `convergence`, but each participant has their own unique offset (i.e. difference from the fixed intercept). We can think about the participant random effects in terms of percentage of the total residual variance in the model (i.e. the variance that's left given the fixed effects in our model). We can calculate that percentage as $\text{intercept_ID}/ (SD_{\text{intercept_ID}} + SD_{observations})$

The fixed effect of `convergence` remains the same. It, again, remains unchanged when adding `number` as a control variable to the mixed model. The fixed effect of `number` is also the same as in the standard regression model.

In the third mixed model we let both the intercept and the effect of `convergence` vary by participant. This is similar to an *interaction*, where the effect one variable (`convergence`) is allowed to vary by the levels/values of another (`ID`).

We have a new term for participant-specific variation in the `convergence`, which is `SD (convergencenoconv ID)` effect.

The term `Cor (Intercept~convergencenoconv ID)` indicates the correlation between the random terms. A negative correlation means that if the random intercept for a specific participant is higher, the slope tends to be lower and vice versa.

In the third model, too, the fixed effect of `convergence` remains unchanged.

### Hypothesis 2

We will proceed just as with hypothesis 1, but for `competence`.

```{r}
# summary stats
d %>%
  group_by(convergence) %>%
  get_summary_stats(competence, type = "mean_sd")

```

```{r}
# T-test
t <- t.test(competence ~ convergence, data = d, paired = TRUE)
t
```

The paired t-test shows a significant difference in competence ratings between *no convergence* and *convergence*. In convergent conditions, participants were on average more confident by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7).

For a robustness check, we ran the same mixed models as described for confidence.

## RQ2 - Effect of number on accuracy (i.e. 'confidence')

```{r}
# get summary stats by number
d %>% 
  group_by(number) %>% 
  summarise(across(c(confidence, competence),  
                   list(mean = mean, sd = sd)))
```

```{r, echo=FALSE}
# T-tests
# accuracy
t <- t.test(confidence ~ number, data = d, paired = TRUE)
t
# competence
t <- t.test(competence ~ number, data = d, paired = TRUE)
t
```

There is a significant effect of `number` on `accuracy`. However, this effect is small compared to the effect of `convergence`. There is no effect on `competence`.

We find the same results using a mixed model.

```{r, echo=FALSE}
# as mixed model
accuracy <- lmer(confidence ~ number + (1 | ID), data= d)
competence <- lmer(competence ~ number + (1 | ID), data= d)

modelsummary(list("Accuracy" = accuracy , "Competence" = competence), 
             title = 'Models estimating effect of number', 
             stars = TRUE, 
             statistic = 'p.value')
```

## RQ3 - Interaction between number and convergence on competence

In our test for an interaction, we use effect-coded versions of our variables.

```{r}
# effect-coded ('_eff') interaction with random intercepts
d <- d %>% 
  mutate(convergence_eff = recode(convergence, "noconv" = -0.5, "conv" = +0.5),
         number_eff = recode(number, "small" = -0.5, "large" = +0.5))
```

```{r}
model_interaction = lmer(competence ~ convergence_eff + number_eff + convergence_eff*number_eff +
            (1  | ID), data = d)
tidy(model_interaction)
```

There seems to be only just a significant interaction in a mixed model where we let the intercept vary across participants. It suggests that the positive effect convergence is enhanced when the number of estimates is `large` (compared to `small`).

```{r}
# quick visualization
ggplot(d, aes(x = convergence, y = competence, 
                     shape = number,
                     group = number,
                     color = number)) +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_color_manual(values = c("#8c510a", "#80cdc1")) +
  theme_classic()
```

## Visualization

### Main plot

```{r summary-plot}
confidence_plot <- ggplot(data=d, aes(x=confidence, fill=convergence)) + 
  geom_density(adjust=3, alpha=.4)+
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
    scale_fill_manual(
    name = "Convergence: ",
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=d, aes(x=competence, fill=convergence)) + 
  geom_density(adjust=3, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
    scale_fill_manual(
    name = "Convergence: ",
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
    guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, competence_plot, common.legend = T) +
  theme(legend.position="top") 
figure
```

### Other plots

#### Confidence/Accuracy

```{r, include=FALSE}
# make plot data
d_means <- d %>%
  group_by(convergence) %>%
  summarize(mean_confidence=mean(confidence))
```

```{r, echo=FALSE}
# Smoothed density plot
density_plot_smoothed <- ggplot(data=d, aes(x=confidence, fill=convergence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Confidence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  geom_vline(data=d_means, aes(xintercept=mean_confidence, 
                                       color=convergence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text(data = d_means, 
            aes(x = mean_confidence, y = 0.3, 
                label =  paste("mean", round(mean_confidence, digits = 1),
                               sep = " = ")),
            nudge_x = -0.5) +
  scale_fill_manual(
    name = "Convergence: ",
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr() + 
  # larger axis titles
  theme(axis.title = element_text(size = 20)) +
  theme(legend.text = element_text(size = 15))
density_plot_smoothed
```

```{r, echo=FALSE}
# Bar plot
bar_plot <- ggplot(data=d, aes(x=confidence, fill=convergence)) + 
 geom_bar(alpha=0.4, 
                position="identity")+ 
  scale_x_continuous(name = "Confidence ratings", breaks = seq(1, 7)) +
  scale_y_continuous(name = "Frequency") +
  geom_vline(data=d_means, aes(xintercept=mean_confidence, 
                                       color=convergence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text(data = d_means, 
            aes(x = mean_confidence, y = 280, 
                label =  paste("mean", round(mean_confidence, digits = 1),
                               sep = " = ")),
            nudge_x = -0.6) +
  scale_fill_manual(
    name = "Convergence: ",
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr() +
    # larger axis titles
  theme(axis.title = element_text(size = 20)) +
  theme(legend.text = element_text(size = 15))  
bar_plot
```

#### Competence

We start by visualizing the difference in `competence`ratings between *convergence* and *no convergence* conditions.

```{r, include=FALSE}
# make plot data
d_means <- d %>%
  group_by(convergence) %>%
  summarize(mean_competence = mean(competence))
```

```{r, echo=FALSE}
# Smoothed density plot
density_plot_smoothed <- ggplot(data=d, aes(x=competence, fill=convergence)) + 
  geom_density(adjust=3, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  geom_vline(data=d_means, aes(xintercept=mean_competence, 
                               color=convergence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text(data = d_means, 
            aes(x = mean_competence, y = 0.3, 
                label =  paste("mean", round(mean_competence, digits = 1),
                               sep = " = ")),
            nudge_x = -0.5) +
  scale_fill_manual(
    name = "Convergence: ",
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr() + 
  # larger axis titles
  theme(axis.title = element_text(size = 20)) +
  theme(legend.text = element_text(size = 15))
density_plot_smoothed
```

### 2023-04-20 Experiment_1/cleaning.Rmd
initialize repository

---
title: "Data cleaning experiment 1"
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-09"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
```

## Import data.
```{r import, message=FALSE}
# import data
d <- read_csv("./data/qualtrics.csv")

# delete first two rows
d <- d[3:202,]
```

## Attention checks. 
```{r attention}
# attention check
# to see different answers given (i.e.levels), transform into factor
d$attention <- as.factor(d$attention)
# check levels to see different answer types
levels(d$attention) 
```
All levels are some variation of "I pay attention", so we do not exclude any participants.

## Re-shaping data

For now, the data is in a wide format where each combination of stimulus item and dependent variable has its own column. We need to bring that data into long format.

```{r, message=FALSE, warning=FALSE}
d <- d %>% 
  # build an easy-to-read ID variable for subjects
  mutate(ID = as.factor(1:nrow(.))) %>% 
  # order data by subjects
  arrange(ID) %>% 
  # from wide to long format
  gather(item, score, conv_large_guess_a_4:noconv_small_comp_b...91,
               factor_key=TRUE) %>% 
  # seperate variables 
  separate(item, into = c("convergence", "number", "dependent_variable",
                          "variant"), convert = T, sep = "\\_")  %>% 
  # remove NA's
  filter(!is.na(score)) %>% 
  # make sure data is still ordered by participants
  arrange(ID) %>% 
  # The variable "variant" that distinguishes between the two items of the same
  # condition did not come as intended. Without changing it, we won't be able
  # to "spread" our dependent variables (DV)
  mutate(variant = rep(c("a", "b"), each = 3, length.out = nrow(.))) %>% 
  # now for each subject, each condition and each variant the three DVs
  # can be assigned
  pivot_wider(names_from = dependent_variable, values_from = score) %>% 
  # so far, all variables are coded characters (see e.g. str(d))
  # we want our dependent variables to be numeric
  mutate(competence = as.numeric(comp),
         guess = as.numeric(guess), 
         confidence = as.numeric(conf))
```

## Recoding some values

To see if there are any weird values, we inspect the distributions of the dependent variables.
```{r, message = FALSE, echo=FALSE}
# competence
ggplot(d, aes(x = competence)) +
  geom_histogram()# seems like 7s have been coded as 8s

# confidence
ggplot(d, aes(x = confidence)) +
  geom_histogram()# same, 7s have been coded as 8s

# guess
ggplot(d, aes(x = guess)) +
  geom_histogram() # all guesses are coded withing range, looks good
```

For the two Likert scale measures (`confidence` and `competence`) The maximum of the scale has been coded as `8` instead of `7`.

```{r}
# change values `8` to `7`
d <- d %>% 
  mutate(competence = ifelse(competence == 8, 7, competence), 
         confidence = ifelse(confidence ==8, 7, confidence))

# check again
ggplot(d, aes(x = competence)) +
  geom_histogram()# looks good

ggplot(d, aes(x = confidence)) +
  geom_histogram()# looks good
```

## Recoding factors

Lastly, we recode the `convergence` variable. For now, the baseline category is _convergence_, which will yield a negative estimate for the effect of _no convergence_. We want to have the positive effect of _convergence_ instead. Likewise, we reconde the `number` variable so that we get the effect of _large_ (i.e. put _small_ as baseline category).

```{r}
# recode convergence
d <- d %>%
  mutate(convergence = as.factor(convergence),
         convergence = fct_relevel(convergence, "noconv", "conv"),
         number = as.factor(number),
         number = fct_relevel(number, "small", "large"))
```

## Export data

```{r}
write_csv(d, "data/cleaned.csv")
```

### 2023-04-20 Experiment_2/analysis.Rmd
initialize repository

---
title: "Analysis experiment 2"
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-09"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
library(ggrepel)       # annotations in plots
```

```{r, message=FALSE}
d <- read_csv("./data/cleaned.csv")
```

### Some themes and functions

```{r, echo=FALSE}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)

#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

## Data analysis

### Hypothesis 1

To assess the effect of independence on participants’ confidence about their guesses, we run a paired t-test.

```{r}
# get summary stats by number
d %>% 
  group_by(independence) %>% 
  summarise(across(c(confidence, competence),  
                   list(mean = mean, sd = sd)))
```

```{r}
# T-test
t <- t.test(confidence ~ independence, data = d, paired = TRUE)
t
```
The paired t-test shows a significant difference in confidence ratings between _independence_ and _dependence_. In the independence condition, participants were on average more confident by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7).

For a robustness check, we run a mixed model regression analysis of `independence` on `confidence` ratings. The first model (`m1`) includes random intercepts for participants.

For reference, we start with standard, fixed-only effects regression models. 
```{r}
# for confidence 
# fixed linear regression models for reference
fixed_confidence <- lm(confidence ~ independence, data= d)
# model with random intercept per participant
mixed_confidence = lmer(confidence ~ independence + (1 |ID), data = d)

# for competence 
# fixed linear regression models for reference
fixed_competence <- lm(competence ~ independence, data= d)
# model with random intercept per participant
mixed_competence= lmer(competence ~ independence + (1 |ID), data = d)
# store in list
models = list("fixed" = fixed_confidence,
              "mixed" = mixed_confidence, 
              "fixed" = fixed_competence, 
              "mixed" = mixed_competence)
```

```{r}
modelsummary(models,
             stars = TRUE,
             title = 'Models estimating effect of convergence on accuracy and competence') %>%
    add_header_above(c(" " = 1, "Accuracy" = 2, "Competence" = 2))
```

The fixed-effect-only linear regression model yields the same estimate for the effect of `independence` as the paired t-test, with the difference that this is not statistically significant at the 5%-level anymore. In the mixed model, the main effect of `independence` remains the same, but is significant again, like in the paired t-test.

### Hypothesis 2

We will proceed just as with hypothesis 1, but for `competence`.

```{r}
# T-test
t <- t.test(competence ~ independence, data = d, paired = TRUE)
t
```
The paired t-test does not show a significant difference in confidence ratings between _no convergence_ and _convergence_.

## Visualization

### Main plot

```{r}
# make plots
confidence_plot <- ggplot(data=d, aes(x=confidence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.45)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1, 
                    name = "Independence: ",
                    limits = c("dep", "indep"), 
                    labels = c("discussion", "independence")) + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=d, aes(x=competence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.45)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1, 
                    name = "Independence: ",
                    limits = c("dep", "indep"), 
                    labels = c("discussion", "independence")) + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, competence_plot, common.legend = T) +
  theme(legend.position="top") 
figure
```

### Other plots

Here's a visualization of the difference in `confidence` ratings between the _independence_ and _dependence_ conditions. 
```{r}
# create objects for means to include in graphics
d_means <- d %>%
  group_by(independence) %>%
  summarize(mean_confidence=mean(confidence))

# Vizualization
density_plot <- ggplot(data=d, aes(x=confidence, fill=independence)) +
geom_density(alpha=.4) 
density_plot

density_plot_smoothed <- ggplot(data=d, aes(x=confidence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Confidence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_fill_brewer(palette = "Dark2") +
  theme_pubr()
density_plot_smoothed

# prettier
density_plot_smoothed <- ggplot(data=d, aes(x=confidence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Confidence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  geom_vline(data=d_means, aes(xintercept=mean_confidence, 
                                       color=independence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text_repel(data = d_means, 
            aes(x = mean_confidence, y = 0.3, 
                label =  paste("mean", round(mean_confidence, digits = 1),
                               sep = " = "))) +
  scale_fill_brewer(
    name = "Independence: ",
    labels = c("discussion", "independence"),
    palette = "Accent") + 
  scale_color_brewer(palette = "Accent") +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr() + 
  # larger axis titles
  theme(axis.title = element_text(size = 20)) +
  theme(legend.text = element_text(size = 15))
density_plot_smoothed

```

Here's a visualization of the difference in `competence` ratings between the _independence_ and _dependence_ conditions. 
```{r}
# create objects for means to include in graphics
d_means <- d %>%
  group_by(independence) %>%
  summarize(mean_competence=mean(competence))

# Vizualization
density_plot <- ggplot(data=d, aes(x=competence, fill=independence)) +
geom_density(alpha=.4) 
density_plot

# prettier
density_plot_smoothed <- ggplot(data=d, aes(x=competence, fill=independence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.5)) +
  geom_vline(data=d_means, aes(xintercept=mean_competence, 
                                       color=independence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text_repel(data = d_means, 
            aes(x = mean_competence, y = 0.47, 
                label =  paste("mean", round(mean_competence, digits = 1),
                               sep = " = "))) +
  scale_fill_brewer(
    name = "Independence: ",
    labels = c("discussion", "independence"),
    palette = "Accent") + 
  scale_color_brewer(palette = "Accent") +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr() + 
  # larger axis titles
  theme(axis.title = element_text(size = 20)) +
  theme(legend.text = element_text(size = 15))
density_plot_smoothed

```

### 2023-04-20 Experiment_2/cleaning.Rmd
initialize repository

--- 2023-04-20 Experiment_1/cleaning.Rmd
@@
 ---
-title: "Data cleaning experiment 1"
+title: "Data cleaning experiment 2"
 author: "Jan Pfänder, Hugo Mercier"
@@
 
-## Import data.
-```{r import, message=FALSE}
-# import data
+## Import data
+
+```{r import}
 d <- read_csv("./data/qualtrics.csv")
+head(d)
 
@@
 
-## Attention checks. 
+## Attention checks
+
 ```{r attention}
@@
 
+## Variable name correction
+
+There was a coding problem in the data - a variable that start with "dep_conf_b_III" appears twice.
+Looking at the qualtrics, one of them should in fact identify "dep_guess_b_III", i.e.
+the numeric guess instead of a confidence rating.
+We identify this column looking at the values.
+Large values in the thousands indicate it's a guess and not a confidence rating.
+
+```{r}
+d %>% 
+  select(starts_with("dep_conf_b_III"))
+```
+
+The output indicates that "dep_conf_b_III_4" captures the guess and should be named "dep_guess_b_III".
+
+```{r}
+d <- d %>% 
+  rename(dep_guess_b_III = dep_conf_b_III_4)
+```
+
 ## Re-shaping data
 
-For now, the data is in a wide format where each combination of stimulus item and dependent variable has its own column. We need to bring that data into long format.
+For now, the data is in a wide format where each combination of stimulus item and dependent variable has its own column. We need to "tidy" the data.
 
-```{r, message=FALSE, warning=FALSE}
+```{r}
 d <- d %>% 
@@
   # from wide to long format
-  gather(item, score, conv_large_guess_a_4:noconv_small_comp_b...91,
+  gather(item, score, dep_guess_b_I_4:dep_comp_a_III,
                factor_key=TRUE) %>% 
   # seperate variables 
-  separate(item, into = c("convergence", "number", "dependent_variable",
-                          "variant"), convert = T, sep = "\\_")  %>% 
+  separate(item, into = c("independence", "dependent_variable", 
+                          "stimulus_variant", "sample"), 
+           convert = T, sep = "\\_")  %>% 
   # remove NA's
@@
   arrange(ID) %>%
-  # The variable "variant" that distinguishes between the two items of the same
-  # condition did not come as intended. Without changing it, we won't be able
-  # to "spread" our dependent variables (DV)
-  mutate(variant = rep(c("a", "b"), each = 3, length.out = nrow(.))) %>% 
   # now for each subject, each condition and each variant the three DVs
@@
 To see if there are any weird values, we inspect the distributions of the dependent variables.
-```{r, message = FALSE, echo=FALSE}
+```{r}
 # competence
@@
 
-For the two Likert scale measures (`confidence` and `competence`) The maximum of the scale has been coded as `8` instead of `7`.
+For the two Likert scale measures (`confidence` and `competence`) the maximum of the scale has been coded as `8` instead of `7`.
 
 ```{r}
-# change values `8` to `7`
 d <- d %>% 
@@
 
-## Recoding factors
-
-Lastly, we recode the `convergence` variable.
-For now, the baseline category is _convergence_, which will yield a negative estimate for the effect of _no convergence_.
-We want to have the positive effect of _convergence_ instead.
-Likewise, we reconde the `number` variable so that we get the effect of _large_ (i.e.
-put _small_ as baseline category).
-
-```{r}
-# recode convergence
-d <- d %>%
-  mutate(convergence = as.factor(convergence),
-         convergence = fct_relevel(convergence, "noconv", "conv"),
-         number = as.factor(number),
-         number = fct_relevel(number, "small", "large"))
-```
-
 ## Export data

### 2023-04-20 Experiment_3/analysis.Rmd
initialize repository

---
title: "Analysis experiment 3"
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-09"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE, warning=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(sjPlot)        # plot the interaction model
library(gghalves)      # Special half geoms
library(ggbeeswarm)    # Special distribution-shaped point jittering
library(ggrepel)       # for non-overlapping labels in graphs
library(kableExtra)   # for tables
```

```{r, message=FALSE}
d <- read_csv("./data/cleaned.csv")
```

### Some themes and functions

```{r, echo=FALSE}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)

#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

## Data analysis

### Hypothesis 1a

We will perform a paired t-test to assess the effect of convergence on participants’ accuracy ratings for the experts' predictions. This test accounts for the dependency between the two conditions of _convergence_ issued by the fact that in our design, this was a within-participant factor.

We will limit the sample to those participants that were assigned to the indepdendence condition of our second, between-participants factor _independence_.

```{r}
# filter data
independence_sample <- d %>% filter(independence == "indep")
```

```{r}
# combined summary stats
independence_sample %>%
  group_by(convergence) %>%
  summarize(across(c(accuracy, competence), 
                   list(mean = mean, sd = sd)))
```

```{r}
# T-test
t <- t.test(accuracy ~ convergence, data = independence_sample, paired = TRUE)
t
```
The paired t-test shows that *given independence*, there is a significant difference in accuracy ratings between _convergence_ and _divergence_. In the convergence condition, participants rated experts' accuracy higher by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7) on average.

For a robustness check, we run a mixed model, including random intercepts and slopes by participants. For reference, we also report a fixed-only effects regression. 
```{r}
# accuracy
# fixed linear regression models for reference
fixed_accuracy = lm(accuracy ~ convergence, data= independence_sample)
# mixed model
mixed_accuracy = lmer(accuracy ~ convergence + (1 + convergence |ID), data = independence_sample)

# competence
# fixed linear regression models for reference
fixed_competence = lm(competence ~ convergence, data= independence_sample)
# mixed model
mixed_competence = lmer(competence ~ convergence + (1 + convergence |ID), data = independence_sample)

# store in list
models = list("fixed" = fixed_accuracy,
              "mixed" = mixed_accuracy, 
              "fixed" = fixed_competence, 
              "mixed" = mixed_competence)
```

```{r}
modelsummary(models,
             stars = TRUE,
             title = 'Models estimating effect of convergence on accuracy and competence') %>%
    add_header_above(c(" " = 1, "Accuracy" = 2, "Competence" = 2))
```

### Hypothesis 1b
We proceed just as with hypothesis 1a, but for `competence` ratings, given _independence_.

```{r}
# T-test
t <- t.test(competence ~ convergence, data = independence_sample, paired = TRUE)
t
```
The paired t-test shows that *given independence*, there is a significant difference in accuracy ratings between _convergence_ and _divergence_. In the convergence condition, participants rated experts' accuracy higher by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7) on average.

For a robustness check, we run a mixed model (see above).

### Hypothesis 2a

We want to know how the between factor `independence` alters the effect of the within factor `convergence` . We run an OLS regression with both convergence and independence and their interaction as independent variables, and `accuracy` ratings as dependent variables. We have

accuracy = a + b1 convergence + b2 independence + b3 independence*convergence

where b1 is the effect of convergence, conditional on dependence, b2 is the effect of independence, conditional on divergence, and b3 the interaction term, indicating the difference of the effect of convergence between the independence and dependence condition. Based on these coefficients, we can calculate the expected means of the accuracy ratings in the four conditions:

    E(competence | convergence = 1 , independence = 0) = a + b1
    E(competence | convergence = 0 , independence = 0) = a 
    E(competence | convergence = 1 , independence = 1) = a + b1 + b2 + b3
    E(competence | convergence = 0 , independence = 1) = a + b2

If the interaction term b3 is significant, then there is a difference in the effect of convergence between independent and dependent contexts. If the term is positive, the effect is more positive in the independent context.

```{r}
# effect-code convergence and independence
d <- d %>% 
  mutate(convergence_effect = recode(convergence, "div" = -0.5, "conv" = +0.5),
         independence_effect = recode(independence, "confl" = -0.5,
                                     "indep" = +0.5))
# accuracy
fixed_interaction_accuracy = lm(accuracy ~ convergence*independence, data = d)
mixed_interaction_accuracy = lmer(accuracy ~ convergence_effect + 
                            independence_effect + 
                            convergence_effect*independence_effect + 
                            (1 + convergence_effect | ID), data = d)

# competence
fixed_interaction_competence = lm(competence ~ convergence*independence, data = d)
mixed_interaction_competence = lmer(competence ~ convergence_effect + 
                            independence_effect + 
                            convergence_effect*independence_effect + 
                            (1 + convergence_effect | ID), data = d)
# store in list
models = list("fixed" = fixed_interaction_accuracy,
              "mixed" = mixed_interaction_accuracy, 
              "fixed" = fixed_interaction_competence, 
              "mixed" = mixed_interaction_competence)

```

```{r}
modelsummary(models,
             stars = TRUE,
             title = 'Models estimating interaction of convergence and independence on accuracy and competence') %>%
    add_header_above(c(" " = 1, "Accuracy" = 2, "Competence" = 2))
```

As predicted, we find a significant and positive interaction term (0.99), suggesting that the effect of `convergence` is more positive in the _independence_ condition. Meanwhile, there is also a significant positive effect, in the _conflict of interest_ condition (0.89). The positive effect of `convergence` on `accuracy` is thus more than twice as big in the _independence_ condition, but generally positive.

As a robustness check, we run a mixed model OLS regression that will additionally include random intercepts and random slopes for participants. This is to control for the dependency of our observations, because each participant gives several _convergence_, but also several _divergence_ ratings. Since we have one between (`independence`) and one within (`convergence`) factor, random slopes will only be included for the within factor, i.e. `convergence`. In this model, in order to assign meaningful random effects, we will change the dummy coding (0,1) for our factors to effect coding (-0.5, 0.5).

With the effect coding, the interpretation of the regression coefficients changes, except for the interaction term:

    a = grand mean, i.e. the mean of competence ratings across all conditions
    b1 = main effect of convergence, i.e. average difference for within conditions convergence - divergence across levels of the factor independence
    b2 = main effect of independence, i.e. average difference for between conditions independence - conflict of interest across levels of the factor convergence
    b3 = interaction term, i.e. difference in effect of convergence between the two levels of independence (independence and conflict of interest)

In the mixed effect model, we find the same positive and significant interaction term as in the simple linear model before (0.99).

What is new is that we additionally find a significant main effect of `independence` (0.33). In other words, across _convergent_ and _divergent_ conditions, people rated _independent_ experts as more competent by 0.33 points. This effect is small compared to the main effect of `convergence` (1.38).

### Hypothesis 2b

We proceed just as with H2a, but this time with `competence` as dependent variable (see above).

As predicted, we find a significant and positive interaction term (0.80), suggesting that the effect of `convergence` is more positive in the _independence_ condition. Meanwhile, there is also a significant positive effect, in the _conflict of interest_ condition (0.82). The positive effect of `convergence` on `competence` is thus twice as big in the _independence_ condition, but generally positive.

In the mixed effect model, we find the same positive and significant interaction term as in the simple linear model before (0.80).

What is new is that we additionally find a significant main effect of `independence` (0.28). In other words, across _convergent_ and _divergent_ conditions, people rated experts as more competent by 0.28 points. This effect is small compared to the main effect of `convergence` (1.22).

### Research question 1

We will perform a paired t-test to assess the effect of convergence on participants’ accuracy ratings for the experts' predictions.

We will limit the sample to those participants that were assigned to the _conflict of interest_ condition of our second, between-participants factor `independence`.

```{r}
# reduce data to independence conditions
conflict_sample <- d %>% 
  filter(independence == "confl")
```

```{r}
# descriptive stats
conflict_sample %>% 
  group_by(convergence) %>% 
  summarise(across(c(accuracy,competence), 
                   list(mean = mean, sd = sd)
                   )
            )
```

```{r}
# T-test
t <- t.test(accuracy ~ convergence, data = conflict_sample, paired = TRUE)
t
```

The paired t-test shows that *given conflict of interest*, there is a significant difference in `accuracy` ratings between _convergence_ and _divergence_. However, this difference goes into the same direction as given _independence_ (see H1a) and is not inverse as we had speculated: In the convergence condition, participants rated experts' accuracy higher by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7) on average.

### Research question 2

We proceed just as with RQ1. We perform a paired t-test to assess the effect of `convergence`, but this time on participants’ `competence` ratings, given _conflict of interest_.

```{r}
# T-test
t <- t.test(competence ~ convergence, data = conflict_sample, paired = TRUE)
t
```

The paired t-test shows that *given conflict of interest*, there is a significant difference in `competence` ratings between _convergence_ and _divergence_. However, again, this difference goes into the same direction as given _independence_ (see H1b) and is not inversed as we had speculated: In the convergence condition, participants rated experts' competence higher by `r abs(t$estimate[[1]])` points (on a scale from 1 to 7) on average.

## Visualization

### Main plot

```{r main-plot}
interaction_accuracy <- ggplot(d, aes(x=independence, y=accuracy, fill = convergence, 
                     shape = convergence,
                     group = convergence,
                     color = convergence)) +
  scale_x_discrete(limits = c("confl", "indep"), 
                    labels = c("conflict of interest", "independence")) +
  geom_half_violin(data = d %>% filter(independence=="confl"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = d %>% filter(independence=="indep"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Accuracy") +
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7)) +
  #ggtitle('Effect of convergence by independence/conflict of interest') +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none",
         fill = guide_legend(title = NULL)) +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

interaction_competence <- ggplot(d, aes(x=independence, y=competence, 
                                        fill = convergence, 
                                        shape = convergence, 
                                        group = convergence, 
                                        color = convergence)) +
  scale_x_discrete(limits = c("confl", "indep"), 
                   labels = c("conflict of interest", "independence")) +
  geom_half_violin(data = d %>% filter(independence=="confl"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = d %>% filter(independence=="indep"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Competence")+
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7))+
  #ggtitle('Effect of convergence by independence/conflict of interest') +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none") +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

figure <- ggarrange(interaction_accuracy, interaction_competence, 
                    common.legend = T) +
  theme(legend.position="top") 
figure

```

### Other plots

```{r}
# create objects for means to include in graphics
d_means <- independence_sample %>%
  group_by(convergence) %>%
  summarize(mean_accuracy = mean(accuracy), 
            mean_competence = mean(competence))

accuracy_plot <- ggplot(data=independence_sample, 
                        aes(x=accuracy, fill=convergence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Accuracy ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  geom_vline(data=d_means, aes(xintercept=mean_accuracy, 
                                       color=convergence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text_repel(data = d_means, 
            aes(x = mean_accuracy, y = 0.35, 
                label =  paste("mean", round(mean_accuracy, digits = 1),
                               sep = " = "))) +
  scale_fill_manual(
    name = "Convergence: ",
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr()

accuracy_plot

competence_plot <- ggplot(data=independence_sample, 
                          aes(x=competence, fill=convergence)) + 
  geom_density(adjust=2, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  geom_vline(data=d_means, aes(xintercept=mean_competence, 
                               color=convergence),
             show.legend = FALSE,
             linetype="dashed", size=1.5) +
  geom_text_repel(data = d_means, 
            aes(x = mean_competence, y = 0.35, 
                label =  paste("mean", round(mean_competence, digits = 1),
                               sep = " = "))) +
  scale_fill_manual(
    name = "Convergence: ",
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) + 
  scale_color_manual(values = c("#E69F00", "#56B4E9")) +
  guides(fill = guide_legend(title = NULL)) +
  theme_pubr()

figure <- ggarrange(accuracy_plot, competence_plot, common.legend = T) +
  theme(legend.position="top") 
figure
```

Here's a visualization of the difference in `accuracy` ratings between _convergence_ and _divergence_ conditions. 
```{r}
# reduce data to independence conditions
independence_sample <- d %>% 
  filter(independence == "indep")

# Vizualization
density_plot <- ggplot(data=independence_sample, aes(x=accuracy, fill=convergence)) +
geom_density(alpha=.4) 
density_plot

density_plot_smoothed <- ggplot(data=independence_sample, aes(x=accuracy, fill=convergence)) + 
  geom_density(adjust=3, alpha=.4)+
  scale_x_continuous(name = "Accuracy ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  theme_pubr()
density_plot_smoothed
```

Here's a visualization of the difference in `competence` ratings between _convergence_ and _divergence_ conditions. 
```{r}
# Vizualization
density_plot <- ggplot(data=independence_sample, aes(x=competence, fill=convergence)) +
geom_density(alpha=.4) 
density_plot

density_plot_smoothed <- ggplot(data=independence_sample, aes(x=competence, fill=convergence)) + 
  geom_density(adjust=3, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  theme_pubr()
density_plot_smoothed
```

Here's a visualization of the difference in `accuracy` ratings between _convergence_ and _divergence_ conditions. 
```{r}
# Vizualization
density_plot <- ggplot(data=conflict_sample, aes(x=accuracy, fill=convergence)) +
geom_density(alpha=.4) 
density_plot

density_plot_smoothed <- ggplot(data=conflict_sample, aes(x=accuracy, fill=convergence)) + 
  geom_density(adjust=3, alpha=.4)+
  scale_x_continuous(name = "Accuracy ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  theme_pubr()
density_plot_smoothed
```

Here's a visualization of the difference in `competence` ratings between _convergence_ and _divergence_ conditions. 
```{r}
# Vizualization
density_plot <- ggplot(data=conflict_sample, aes(x=competence, fill=convergence)) +
geom_density(alpha=.4) 
density_plot

density_plot_smoothed <- ggplot(data=conflict_sample, aes(x=competence, fill=convergence)) + 
  geom_density(adjust=3, alpha=.4)+
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  theme_pubr()
density_plot_smoothed
```

### 2023-04-20 Experiment_3/cleaning.Rmd
initialize repository

---
title: "Data cleaning experiment 3"
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-09"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
```

## Import data
```{r import}
d <- read_csv("./data/qualtrics.csv")
head(d)

# delete first two rows
d <- d[3:201,]
```

## Attention check
```{r attention}
# attention check
# to see different answers given (i.e.levels), transform into factor
d$attention <- as.factor(d$attention)
# check levels to see different answer types
levels(d$attention) 
```

All levels are some variation of "I pay attention", so we do not exclude any participants.

## Re-shape data
```{r}
d <- d %>% 
  # build an easy-to-read ID variable for subjects
  mutate(ID = as.factor(1:nrow(.))) %>% 
  # order data by subjects
  arrange(ID) %>% 
  # from wide to long format
  gather(item, score, indep_conv_acc_a_I:confl_div_comp_b_II,
               factor_key=TRUE) %>% 
  # seperate variables 
  separate(item, into = c("independence", "convergence", "dependent_variable", 
                          "stimulus_variant", "sample"), 
           convert = T, sep = "\\_")  %>% 
  # remove NA's
  filter(!is.na(score)) %>% 
  # make sure data is still ordered by participants
  arrange(ID) %>%
  # now for each subject, each condition and each variant the three DVs
  # can be assigned
  pivot_wider(names_from = dependent_variable, values_from = score) %>% 
  # so far, all variables are coded characters (see e.g. str(d))
  # we want our dependent variables to be numeric
  mutate(competence = as.numeric(comp), 
         accuracy = as.numeric(acc))
```

## Re-code some values

```{r}
# competence
ggplot(d, aes(x = competence)) +
  geom_histogram()# seems like 7s have been coded as 8s

# accuracy
ggplot(d, aes(x = accuracy)) +
  geom_histogram()# same, 7s have been coded as 8s
```
For the two Likert scale measures (`confidence` and `accuracy`) the maximum of the scale has been coded as `8` instead of `7`.

```{r}
d <- d %>% 
  mutate(competence = ifelse(competence == 8, 7, competence), 
         accuracy = ifelse(accuracy ==8, 7, accuracy))

# check again
ggplot(d, aes(x = competence)) +
  geom_histogram()# looks good

ggplot(d, aes(x = accuracy)) +
  geom_histogram()# looks good
```

## Recode factors

Lastly, we recode the `convergence` variable. For now, the baseline category is _convergence_, which will yield a negative estimate for the effect of _no convergence_. We want to have the positive effect of _convergence_ instead. Likewise, we recode the levels of the `independence` variable to make sure the effect goes into the right direction (i.e. the effec of _independence_ compared to _conflict of interest_).

```{r}
d <- d %>%
  mutate(convergence = as.factor(convergence),
         convergence = fct_relevel(convergence, "div", "conv"),
         independence = as.factor(independence),
         independence = fct_relevel(independence, "confl", "indep")
         )
```

## Export data

```{r}
write_csv(d, "data/cleaned.csv")
```

### 2023-04-20 Experiment_4/analysis.Rmd
initialize repository

---
title: 
- "Analysis Experiment 4: Do inferences from convergence hold in a categorical choice setting?"
author: 
- "Hugo Mercier, Jan Pfänder" 
date: "2023-02-02"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = TRUE) 
```

```{r, message=FALSE}
library(tidyverse) # for everything
library(tidyr) # some additional, neat tidy reshaping functions
library(lme4) # for linear mixed models
library(lmerTest) # p-values for mixed models
library(broom) # for tidy outputs of regressions
library(broom.mixed) # for tidy outputs of linear mixed models
library(ggridges) # for plots
library(gghalves) # for plots
library(ggbeeswarm) # Special distribution-shaped point jittering
library(scales)  # for nice axis labels
library(modelsummary) # regression tables
library(patchwork) # combine plots
library(ggrepel) # fit labels in plots 
```

### Some themes and functions

```{r}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)

#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

### Read data
```{r, message=FALSE}
# read data
d <- read_csv("./data/cleaned.csv")
```

### H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.

We use a linear mixed effect model with random intercept and random slope per participant. Should this model yield convergence issues, we will use a model with random intercept only.

In all our models we treat `convergence` as a continuous variable. We will, however, include robustness checks where we treat convergence as a categorical variable, allowing to inspect difference between different levels.

```{r}
# model for accuracy
# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = d)

summary(model_accuracy)
```

### H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.

We will proceed in the same way for `competence` as we did for `accuracy` above.

```{r}
# model for competence
# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + 
                           (1 + convergence | id), data = d)

summary(model_competence)
```

# Robustness checks

In the models above, we treated convergence as a continuous variable. Based on the different levels, we will build a categorical variable, `convergence_categorical`.

```{r}
# make a categorical variable from `convergence`
d <- d %>% 
  mutate(convergence_categorical = recode_factor(convergence, 
                                                 `0` = "opposing majority", 
                                                 `1` = "divergence", 
                                                 `2` = "majority", 
                                                 `3` = "consensus",
                                                 .default = NA_character_)
         )

levels(d$convergence_categorical)

```

We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical`. This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e. "opposing majority").

### Plots

```{r}
# plot for accuracy
plot_accuracy <- ggplot(d,
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r}
# plot for competence
plot_competence <- ggplot(d,
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r}
plot_accuracy + plot_competence
```

### 2023-04-20 Experiment_4/cleaning.Rmd
initialize repository

---
title: "Data cleaning experiment 4"
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-09"
output: 
  html_document: 
    keep_md: yes
---

```{r packages, message=FALSE}
library(tidyverse)     # create plots with ggplot, manipulate data, etc.
library(broom.mixed)   # convert regression models into nice tables
library(modelsummary)  # combine multiple regression models into a single table
library(lme4)          # model specification / estimation 
library(lmerTest)      # provides p-values in the output
library(ggpubr)        # stile feature of ggplot
library(gghalves)      # do special plots in ggplot
library(kableExtra)    # for tables
```

## Import data
```{r import}
d <- read_csv("./data/qualtrics.csv")
names(d)
```

```{r}
# delete first two rows
d <- d %>% 
  slice(3: nrow(.)) 
```

## Attention checks 
```{r attention}
# attention check
# to see different answers given (i.e.levels), transform into factor
d$attention <- as.factor(d$attention)
# check levels to see different answer types
levels(d$attention) 
```

No one failed the attention check.

## Re-shape data
```{r}
d <- d %>% 
  # add an easy to read participant identifier
  mutate(id = 1:nrow(.)) %>% 
  # bring to long format
  pivot_longer(cols = c("0_acc_a_1" : "3_com_b"), 
               names_to = "condition", values_to = "score") %>% 
  # For some weird reason, qualtrics added `_1` at the end of each accuracy
  # stimulus name. We remove that
  mutate(across(condition, ~str_remove(., "_1"))) %>% 
  # separate conditions into CONVERGENCE_OUTCOME_STIMULUS
  separate_wider_delim(condition, "_", names = c("convergence", "outcome", 
                                                 "stimulus_variant")
                       ) %>%
  pivot_wider(names_from = outcome, values_from = score) %>% 
  # create better variable names
  rename(competence = com, 
         accuracy = acc) %>% 
  # all variables are coded as `character` - make key variables numeric
  mutate(across(c(convergence, competence, accuracy), as.numeric)
         )
```

## Export data

```{r}
write_csv(d, "data/cleaned.csv")
```

### 2023-04-20 Experiment_4/model.Rmd
initialize repository

---
title: "Model"
output: 
  html_document: 
    keep_md: yes
date: "2023-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages
```{r, Message=FALSE}
library(tidyverse)
library(gghalves) # for plots
library(patchwork)
```

## Set plot theme
```{r}
#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

## Explaining the model

### Define competence and its distribution

We define competence simply as the probability of making an accurate choice (among various options). We assume that individuals (i) vary in competence,(ii) that competence is continuous and ranges from pure chance (`1/options`) to certainly making the correct choice (`1`), and that it is distributed uniformly in the population.

```{r}
# Imagine a situation with 3 choice options
options <- 3
# generate a uniform competence distribution for a population of n = 10'000
population <- 10000
competence = runif(population, min = 1/options, max = 1)

# data generating function
ggplot() +
  stat_function(fun = dunif, args = list(min = 0, max = 1)) +
  labs(title = "Uniform distribution of competence from 1/3 to 1", x = "Competence", y = "P(Competence)") + 
  xlim(1/3, 1)

# generated population
ggplot() +
  geom_histogram(aes(x = competence)) + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
       y = "Count")
```
### Generate choices from competence

### Sample choice

### Iterate this process

## Function for single population

```{r}
simulate_single_population <- function(population, sample, options, 
                                       distribution = "uniform") {
  # Check if population is not divisible by sample without remainder
  if (population %% sample != 0) {
    warning("Population is not divisible by sample without remainder.")
  }

  # if necessary, alter population size so that it is divisible by sample without rest
  n_possible_samples <- floor(population/sample)

  # Biggest possible population to pick
  possible_population = n_possible_samples * sample

  # Issue warning if different population size used
  if (population %% sample != 0) {
    warning(paste0("Chosen population size is ", possible_population))
  }
  # change population value (if divisible without remainder, it'll be the same)
  population = possible_population

  # 1: Distribution of competence

  if (distribution == "uniform") {
  # randomly draw competence levels
  data <- tibble(id = 1:population,
                 competence = runif(population, min = 1/options, max = 1))
  }

  if (distribution == "normal") {
  # randomly draw competence levels
  data <- tibble(id = 1:population,
                 competence = rnorm(population, mean = 0.6, sd = 0.01))
  }

  # Create a relative measure of competence that allows to compare across
  # different numbers of choice options. 
  # This measure varies from 0 = chance to 1 = definitely right. 
  # We use min-max scaling, with min = 1/options and max = 1.
  data <- data %>% 
    mutate(relative_competence = (competence - 1/options) / (1 - 1/options)
           )

  # 2: randomly draw individual answers based on competence levels 
  data$answer <- data$competence %>% 
    purrr::map_chr(function(x){

      answer_options <- c("correct", paste0("false", 1:(options-1)))
      probabilities <- c(x, rep((1-x)/(options-1), options-1))

      answer = sample(answer_options, 
                      size = 1, 
                      prob = probabilities
      )
      return(answer)

    })

  # 3: randomly assign samples in population
  data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))

  # 4: identify constellations 
  data <- data %>% 
    # identify how often a one type of answer occurs in one group
    group_by(sample_id, answer) %>% 
  mutate(n_answer_in_sample = n()) %>% 
    # assign constellations
  group_by(sample_id) %>% 
    mutate(unique_answers = n_distinct(answer),
           # Build a first version of constellation variable
           constellation = case_when(unique_answers == sample ~ "dissensus", 
                                     unique_answers == 1 ~ "consensus",
                                     unique_answers <  sample ~ "majority"
           ),
           # identify minority answers in majority constellations
           minority = ifelse(constellation == "majority", 
                             # report whether occurences of answer within a group
                             # are the minority within that group
                             n_answer_in_sample == min(n_answer_in_sample),
                             # for all other constellations, simply code NA,
                             NA
           ), 
           # identify majority answers in majority constellations
           majority = ifelse(constellation == "majority", 
                             # report whether occurrences of answer within a group
                             # are the majority within that group
                             n_answer_in_sample == max(n_answer_in_sample),
                             # for all other constellations, simply code NA,
                             NA
           ), 
           # modify the constellation variable to distinguish between minority, 
           # intermediate majority, majority and dissensus that is composed of 
           # multiple answers (e.g. 3 options, sample of 6, two per option)
           constellation = case_when(
             is.na(minority) | is.na(majority) ~ constellation, 
             minority == TRUE & majority == TRUE ~ "dissensus",
             minority == TRUE ~ "minority", 
             majority == TRUE ~ "majority", 
             .default = "intermediate majority")
           ) %>% ungroup()

  # 5: calculate accuracy and competence levels per constellation

  # identify accurate responses
  data <- data %>%
    mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))

  # compute the summary statistics by constellation
  results <- data %>% 
    group_by(constellation) %>% 
    summarize(average_competence = mean(competence), 
              average_relative_competence = mean(relative_competence),
              average_accuracy = mean(accurate),
              count = n_distinct(sample_id)) %>% 
    # store simulation information
    mutate(population = population, 
           sample = sample, 
           options = options)

  return(results)
}

```

```{r}
# # test output (set return to data above)
# test <- simulate_single_population(population = 600, sample = 6, options = 3) %>% arrange(sample_id)
# levels(as.factor(test$constellation))
# 
# test %>% filter(constellation == "intermediate majority")
```

## Function for various populations
```{r}
simulate_various_populations <- function(iterations,...) {

  # create data frame with model results for generated samples
  various_populations <- 1:iterations %>% 
    purrr::map_df(function(x){
      # this is essentially a for loop - do the following for each 
      # element in 1:iterations

      results <- simulate_single_population(...)

      # identify iteration
      results$iteration <- x

      # To keep track of progress
      if (x %% 50 == 0) {print(paste("iteration number ", x))}

      return(results)

    }) %>% 
    # store simulation information
    mutate(total_iterations = iterations)

  return(various_populations)
}
```

## Results by varying choice options and sample size function

This function allows us to investigate how accuracy and competence change with varying the number of choice options and the sample size.

The simulations that this function executes will take quite some time. Therefore, we do not want to run it every time we render this document. Instead we want to store the output of the power simulation in a `.csv` file, and have an integrated "stop" mechanism to prevent execution when that file already exists. To achieve this, we make `file_name` a mandatory argument. If a file with that name already exists, the function will not be executed.

```{r}
# set name
file_name <- "model.csv" # change for new analyses / or delete file to re-use same name

vary_sample_options <- function(vary = "options", file_name, n, ...) {

  # only run analysis if a file with that name does not yet exists
  if (!file.exists(paste0("data/", file_name))) {

    # vary choice options for given sample size
    if (vary == "options") {

      # do the `calculate_power()` function for each sample size and store the results
      # in a new variable called `power`
      data <- n %>% purrr::map_df(function(n_option){
                                     # this is essentially a for loop - 
                                     # do the following for each 
                                     # element data$n_subj

                                     # To keep track of progress
                                     print(paste("tested option number = ", n_option))

                                     # run power calculation
                                     result <- simulate_various_populations(
                                       options = n_option, ...)

                                     return(result)

                                   })

      write_csv(data, paste0("data/", file_name))
    }

    # vary sample size for given choice option
     if (vary == "sample") {

      # do the `calculate_power()` function for each sample size and store the results
      # in a new variable called `power`
      data <- n %>% purrr::map_df(function(n_sample){
                                     # this is essentially a for loop - 
                                     # do the following for each 
                                     # element data$n_subj

                                     # To keep track of progress
                                     print(paste("tested sample number = ", n_sample))

                                     # run power calculation
                                     result <- simulate_various_populations(
                                       sample = n_sample, ...)

                                     return(result)

                                   })

      write_csv(data, paste0("data/", file_name))
    }
  }
}

# # You can test the function using the commented code below
# # (un-comment by highlighting, then command+shit+c)
# test <- do.call(power_by_sample_size,
#                 c(parameters,
#                   list(file_name = file_name, sample_sizes = c(5, 10),
#                        iterations = 100)
#                   )
#                 )
```

## Plot functions

### a

First, we want a function that plots results obtained by the `simulate_various_populations` function. 
```{r}
plot_results <- function(data, outcome = "everything") {

  d <- data

  # make constellation a factor with the right levels
  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")

  # extract simulation info
  simulation_info <- d %>% summarize(across(c(population, sample, options, total_iterations), mean)) %>% 
    pivot_longer(everything(), names_to = "parameter", values_to = "value") %>% 
    gridExtra::tableGrob(cols = NULL, rows = NULL)

  # make descriptive data
  descriptive <- d %>% 
    group_by(constellation) %>% 
    summarise(count = sum(count)) %>% 
    mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
    gridExtra::tableGrob(rows = NULL)

  # plot for accuracy
  plot_accuracy <- ggplot(d,
                          aes(x = constellation, y = average_accuracy, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Accuracy") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    guides(fill = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1))

  # plot for competence
  plot_competence <- ggplot(d,
                            aes(x = constellation, y = average_competence, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Competence") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    guides(fill = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1))

  # unite tables
  tables <- gridExtra::grid.arrange(descriptive, simulation_info, ncol = 2)

  if (outcome == "everything") {

    return((plot_accuracy | plot_competence) / tables)
  }

  if (outcome == "accuracy") {

    return(plot_accuracy)
  }

  if (outcome == "competence") {

    return(plot_competence)
  }

}
```

### b

Second, we want a function that plots results obtained by the `vary_sample_options` function.

```{r}
plot_results <- function(data, variable = options) {

  d <- data

  # Extract variable name as a string
  variable_name <- deparse(substitute(variable))

  # retrieve info from choice of variable
  if (variable_name == "options") {
    x_label = "Number of Choice Options"
    title = paste0("Sample: ", d$sample," \n iterations: ", d$total_iterations)
  }
    if (variable_name == "sample") {
    x_label = "Size of informant groups"
    title = paste0("Options: ", d$options," \n iterations: ", d$total_iterations)
  }

  # make constellation a factor with the right levels
  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")

  # plot for accuracy
  plot_accuracy <- ggplot(d,
                          aes(x = as.factor({{variable}}), y = average_accuracy, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = x_label, y = "Accuracy") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1, 
                         limits = rev(levels(d$constellation)),
                         direction = -1
    ) +
    plot_theme

  # plot for competence
  plot_competence <- ggplot(d,
                            aes(x = as.factor({{variable}}), y = average_relative_competence, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=2, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = x_label, y = "Relative competence", fill = NULL) + 
    scale_fill_viridis_d(option = "plasma", begin = 0.1, 
                         limits = rev(levels(d$constellation)),
                         direction = -1
    ) +
    plot_theme

  ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE) +
    plot_annotation(title = title)
}
```

```{r, message=FALSE}
n <- c(3, 10, 50, 100, 200)

vary_sample_options(n = n, vary = "options", iterations = 100, population = 999, sample = 3, file_name = "model_3_options.csv")
vary_sample_options(n = n, vary = "sample", iterations = 100, population = 999, options = 3, file_name = "model_3_samples.csv")

data_3_options <- read_csv("data/model_3_options.csv")
data_3_sample <- read_csv("data/model_3_samples.csv")

plot_results(data_3_options)
plot_results(data_3_sample, variable = sample)

```

```{r}
data_3_sample %>% group_by(sample, constellation) %>% summarise(across(c(average_accuracy, average_relative_competence), mean))

# make descriptive data
data_3_sample %>% 
  group_by(sample, constellation) %>% 
  summarise(count = sum(count)) %>% 
  mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
  print(n = 100)
```

## Simulate

Note that in the current state, we cannot vary `sample`. The function needs modification first. The problem is that it gets complicated with more than three informants, in the case of a majority constellation. There might be equal majorities (e.g. 5 informants, three options: 2 agree on 'A', 2 on 'B', one picks 'A'), equal minorities (e.g. 3 agree on 'A', one picks 'B', another picks 'C'), and unequal majorities (e.g. 3 agree on 'A', 2 agree on 'B').

I have to think about it again.

```{r}
data_3 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 3)

data_10 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 10)

data_100 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 100)

data_2 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 2)
```

```{r}
plot_results(data_3)
plot_results(data_10)

data_3 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
```

```{r}
data_10 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
```

```{r}
data_2 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
```

```{r}
plot_results(data_3)
model_accuracy <- plot_results(data_3, outcome = "accuracy")
model_competence <- plot_results(data_3, outcome = "competence")
```

```{r}
# # example to check essential parts of the function
# my_strings <- c("apple", "banana", "apple", "banana", "apple")
# 
# bla <- tibble(id = rep(1:5, each = 5), 
#        string = sample(c("apple", "banana"), 25, replace = TRUE, 
#                        prob = c(0.8, 0.2)))
# 
# bla %>% group_by(id, string) %>% 
#   mutate(n_answer_in_group = n()) %>% 
#   group_by(id) %>% 
#   mutate(
#     unique_answers = n_distinct(string),
#     minority = ifelse(id == 4, 
#                       n_answer_in_group == min(n_answer_in_group), 
#                       NA
#                       )
#     ) %>% arrange(id) 
```

## Combine with plots from paper

### Make plots from data
```{r, message = FALSE}
# read data
d <- read_csv("./data/cleaned.csv")

# make a categorical variable from `convergence`
d <- d %>% 
  mutate(convergence_categorical = recode_factor(convergence, 
                                                 `0` = "opposing majority", 
                                                 `1` = "divergence", 
                                                 `2` = "majority", 
                                                 `3` = "consensus",
                                                 .default = NA_character_)
         )
```

```{r}
# plot for accuracy
plot_accuracy <- ggplot(d,
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r}
# plot for competence
plot_competence <- ggplot(d,
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

### combine plots from data and from model

```{r}
# accuracy
plot_accuracy <- plot_accuracy + ggtitle("Participants")
model_accuracy<- model_accuracy + ggtitle("Model")

plot_accuracy + model_accuracy
```

```{r}
# competence
plot_competence <- plot_competence + ggtitle("Participants")
model_competence<- model_competence + ggtitle("Model")

plot_competence + model_competence
```

## Have only one population.

```{r}
simulate_single_population <- function(population, sample, options) {

  # randomly draw competence levels
  data <- tibble(id = 1:population,
                 competence = runif(population, min = 1/options, max = 1))

  # randomly draw individual answers based on competence levels 
  data$answer <- data$competence %>% 
    purrr::map_chr(function(x){

      answer_options <- c("correct", paste0("false", 1:(options-1)))
      probabilities <- c(x, rep((1-x)/(options-1), options-1))

      answer = sample(answer_options, 
                      size = 1, 
                      prob = probabilities
      )
      return(answer)

    })

  # randomly assign samples in population
  data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))

  # identify constellations 
  data <- data %>% 
    # identify how often a one type of answer occurs in one group
    group_by(sample_id, answer) %>% 
  mutate(n_answer_in_sample = n()) %>% 
    # assign constellations
  group_by(sample_id) %>% 
    mutate(unique_answers = n_distinct(answer),
           # Build a first version of constellation variable
           constellation = case_when(unique_answers == sample ~ "dissensus", 
                                     unique_answers == 1 ~ "consensus",
                                     unique_answers <  sample ~ "majority"
           ),
           # identify minority answers in majority constellations
           minority = ifelse(constellation == "majority", 
                             # report whether occurences of answer within a group
                             # are the minority within that group
                             n_answer_in_sample == min(n_answer_in_sample),
                             # for all other constellations, simply code NA,
                             NA
           ), 
           # identify majority answers in majority constellations
           majority = ifelse(constellation == "majority", 
                             # report whether occurrences of answer within a group
                             # are the minority within that group
                             n_answer_in_sample == max(n_answer_in_sample),
                             # for all other constellations, simply code NA,
                             NA
           ), 
           # modify the constellation variable to distinguish between minority, 
           # intermediate majority, majority and dissensus that is composed of 
           # multiple answers (e.g. 3 options, sample of 6, two per option)
           constellation = case_when(
             is.na(minority) | is.na(majority) ~ constellation, 
             minority == TRUE & majority == TRUE ~ "dissensus",
             minority == TRUE ~ "minority", 
             majority == TRUE ~ "majority", 
             .default = "intermediate majority")
           ) %>% ungroup()

  # identify accurate responses
  data <- data %>%
    mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))

  # compute the summary statistics by constellation
  results <- data %>% 
    group_by(sample_id, constellation) %>% 
    summarize(average_competence = mean(competence), 
              average_accuracy = mean(accurate)) %>% 
    # store simulation information
    mutate(population = population, 
           sample = sample, 
           options = options) %>%

  return(results)
}

```

```{r}
plot_results <- function(data) {

  d <- data
  # make constellation a factor with the right levels
  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")

  # make count data
  counts <- d %>% group_by(constellation) %>% count()

  # plot for accuracy
  plot_accuracy <- ggplot(d,
                          aes(x = constellation, y = average_accuracy, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=0.1, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Accuracy") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    scale_color_viridis_d(option = "plasma", begin = 0.7) +
    guides(fill = FALSE, color = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
    geom_text(data = counts, aes(x = constellation, y = 1.05, 
                                 label =  paste0("n =", n)
                                 )
              )

  # plot for competence
  plot_competence <- ggplot(d,
                            aes(x = constellation, y = average_competence, fill = constellation)) +
    geom_half_violin(position = position_nudge(x = -.2),
                     adjust=1, alpha = .8,
                     side = "l") +
    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
    # Add nice labels
    labs(x = "Convergence", y = "Competence") +
    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
    scale_color_viridis_d(option = "plasma", begin = 0.7) +
    guides(fill = FALSE) +
    plot_theme + 
    theme(axis.text.x = element_text(angle = 20, hjust = 1))

  (plot_accuracy | plot_competence)
}
```

```{r}
test <- simulate_single_population(population = 999, sample = 3, options = 3)
plot_results(test)
```

### 2023-04-20 Experiment_4/old_model.Rmd
initialize repository

--- 2023-04-20 Experiment_4/model.Rmd
@@
 
-## Explaining the model
+## Function for single population
+```{r}
+simulate_single_population <- function(population, sample, options) {
 
-### Define competence and its distribution
-
-We define competence simply as the probability of making an accurate choice (among various options).
-We assume that individuals (i) vary in competence,(ii) that competence is continuous and ranges from pure chance (`1/options`) to certainly making the correct choice (`1`), and that it is distributed uniformly in the population.
-
-```{r}
-# Imagine a situation with 3 choice options
-options <- 3
-# generate a uniform competence distribution for a population of n = 10'000
-population <- 10000
-competence = runif(population, min = 1/options, max = 1)
-
-# data generating function
-ggplot() +
-  stat_function(fun = dunif, args = list(min = 0, max = 1)) +
-  labs(title = "Uniform distribution of competence from 1/3 to 1", x = "Competence", y = "P(Competence)") + 
-  xlim(1/3, 1)
-
-# generated population
-ggplot() +
-  geom_histogram(aes(x = competence)) + 
-  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
-  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
-  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
-       y = "Count")
-```
-### Generate choices from competence
-
-### Sample choice
-
-### Iterate this process
-
-## Function for single population
-
-```{r}
-simulate_single_population <- function(population, sample, options, 
-                                       distribution = "uniform") {
-  # Check if population is not divisible by sample without remainder
-  if (population %% sample != 0) {
-    warning("Population is not divisible by sample without remainder.")
-  }
-
-  # if necessary, alter population size so that it is divisible by sample without rest
-  n_possible_samples <- floor(population/sample)
-
-  # Biggest possible population to pick
-  possible_population = n_possible_samples * sample
-
-  # Issue warning if different population size used
-  if (population %% sample != 0) {
-    warning(paste0("Chosen population size is ", possible_population))
-  }
-  # change population value (if divisible without remainder, it'll be the same)
-  population = possible_population
-
-  # 1: Distribution of competence
-
-  if (distribution == "uniform") {
   # randomly draw competence levels
@@
                  competence = runif(population, min = 1/options, max = 1))
-  }
 
-  if (distribution == "normal") {
-  # randomly draw competence levels
-  data <- tibble(id = 1:population,
-                 competence = rnorm(population, mean = 0.6, sd = 0.01))
-  }
-
-  # Create a relative measure of competence that allows to compare across
-  # different numbers of choice options. 
-  # This measure varies from 0 = chance to 1 = definitely right. 
-  # We use min-max scaling, with min = 1/options and max = 1.
-  data <- data %>% 
-    mutate(relative_competence = (competence - 1/options) / (1 - 1/options)
-           )
-
-  # 2: randomly draw individual answers based on competence levels 
+  # randomly draw individual answers based on competence levels 
   data$answer <- data$competence %>% 
@@
 
-  # 3: randomly assign samples in population
-  data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))
+  # randomly assign samples in population
+  data <- data %>% mutate(sample = rep(1:(population/sample), sample))
 
-  # 4: identify constellations 
-  data <- data %>% 
-    # identify how often a one type of answer occurs in one group
-    group_by(sample_id, answer) %>% 
-  mutate(n_answer_in_sample = n()) %>% 
-    # assign constellations
-  group_by(sample_id) %>% 
+  # assign constellations build a variable identifying accurate responses
+  data <- data %>% group_by(sample) %>% 
     mutate(unique_answers = n_distinct(answer),
            # Build a first version of constellation variable
-           constellation = case_when(unique_answers == sample ~ "dissensus", 
+           constellation = case_when(unique_answers == 3 ~ "dissensus", 
                                      unique_answers == 1 ~ "consensus",
-                                     unique_answers <  sample ~ "majority"
+                                     unique_answers == 2  ~ "majority"
            ), 
@@
            minority = ifelse(constellation == "majority", 
-                             # report whether occurences of answer within a group
-                             # are the minority within that group
-                             n_answer_in_sample == min(n_answer_in_sample),
+                             # negate duplicated answers
+                             !(duplicated(answer) | duplicated(answer, 
+                                                               # makes sure that the original value AND
+                                                               # the duplicate get marked as duplicates
+                                                               fromLast = TRUE)),
                              # for all other constellations, simply code NA,
@@
            ), 
-           # identify majority answers in majority constellations
-           majority = ifelse(constellation == "majority", 
-                             # report whether occurrences of answer within a group
-                             # are the majority within that group
-                             n_answer_in_sample == max(n_answer_in_sample),
-                             # for all other constellations, simply code NA,
-                             NA
-           ), 
-           # modify the constellation variable to distinguish between minority, 
-           # intermediate majority, majority and dissensus that is composed of 
-           # multiple answers (e.g. 3 options, sample of 6, two per option)
-           constellation = case_when(
-             is.na(minority) | is.na(majority) ~ constellation, 
-             minority == TRUE & majority == TRUE ~ "dissensus",
-             minority == TRUE ~ "minority", 
-             majority == TRUE ~ "majority", 
-             .default = "intermediate majority")
+           # modify the constellation variable to distinguish between minority 
+           # and majority
+           constellation = ifelse(is.na(minority) | minority == FALSE, 
+                                  constellation, 
+                                  "minority")
     ) %>% ungroup()
-
-  # 5: calculate accuracy and competence levels per constellation
 
@@
     summarize(average_competence = mean(competence), 
-              average_relative_competence = mean(relative_competence),
-              average_accuracy = mean(accurate),
-              count = n_distinct(sample_id)) %>% 
+              average_accuracy = mean(accurate)) %>% 
     # store simulation information
@@
 
-```
-
-```{r}
-# # test output (set return to data above)
-# test <- simulate_single_population(population = 600, sample = 6, options = 3) %>% arrange(sample_id)
-# levels(as.factor(test$constellation))
-# 
-# test %>% filter(constellation == "intermediate majority")
 ```
@@
 
-## Results by varying choice options and sample size function
-
-This function allows us to investigate how accuracy and competence change with varying the number of choice options and the sample size.
-
-The simulations that this function executes will take quite some time.
-Therefore, we do not want to run it every time we render this document.
-Instead we want to store the output of the power simulation in a `.csv` file, and have an integrated "stop" mechanism to prevent execution when that file already exists.
-To achieve this, we make `file_name` a mandatory argument.
-If a file with that name already exists, the function will not be executed.
-
+## Plot function
 ```{r}
-# set name
-file_name <- "model.csv" # change for new analyses / or delete file to re-use same name
-
-vary_sample_options <- function(vary = "options", file_name, n, ...) {
-
-  # only run analysis if a file with that name does not yet exists
-  if (!file.exists(paste0("data/", file_name))) {
-
-    # vary choice options for given sample size
-    if (vary == "options") {
-
-      # do the `calculate_power()` function for each sample size and store the results
-      # in a new variable called `power`
-      data <- n %>% purrr::map_df(function(n_option){
-                                     # this is essentially a for loop - 
-                                     # do the following for each 
-                                     # element data$n_subj
-
-                                     # To keep track of progress
-                                     print(paste("tested option number = ", n_option))
-
-                                     # run power calculation
-                                     result <- simulate_various_populations(
-                                       options = n_option, ...)
-
-                                     return(result)
-
-                                   })
-
-      write_csv(data, paste0("data/", file_name))
-    }
-
-    # vary sample size for given choice option
-     if (vary == "sample") {
-
-      # do the `calculate_power()` function for each sample size and store the results
-      # in a new variable called `power`
-      data <- n %>% purrr::map_df(function(n_sample){
-                                     # this is essentially a for loop - 
-                                     # do the following for each 
-                                     # element data$n_subj
-
-                                     # To keep track of progress
-                                     print(paste("tested sample number = ", n_sample))
-
-                                     # run power calculation
-                                     result <- simulate_various_populations(
-                                       sample = n_sample, ...)
-
-                                     return(result)
-
-                                   })
-
-      write_csv(data, paste0("data/", file_name))
-    }
-  }
-}
-
-# # You can test the function using the commented code below
-# # (un-comment by highlighting, then command+shit+c)
-# test <- do.call(power_by_sample_size,
-#                 c(parameters,
-#                   list(file_name = file_name, sample_sizes = c(5, 10),
-#                        iterations = 100)
-#                   )
-#                 )
-```
-
-## Plot functions
-
-### a
-
-First, we want a function that plots results obtained by the `simulate_various_populations` function. 
-```{r}
-plot_results <- function(data, outcome = "everything") {
+plot_results <- function(data) {
 
@@
   # make constellation a factor with the right levels
-  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")
+  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "majority", "consensus")
 
-  # extract simulation info
   simulation_info <- d %>% summarize(across(c(population, sample, options, total_iterations), mean)) %>% 
@@
     gridExtra::tableGrob(cols = NULL, rows = NULL)
-
-  # make descriptive data
-  descriptive <- d %>% 
-    group_by(constellation) %>% 
-    summarise(count = sum(count)) %>% 
-    mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
-    gridExtra::tableGrob(rows = NULL)
 
@@
 
-  # unite tables
-  tables <- gridExtra::grid.arrange(descriptive, simulation_info, ncol = 2)
-
-  if (outcome == "everything") {
-
-    return((plot_accuracy | plot_competence) / tables)
+  (plot_accuracy | plot_competence) / simulation_info
 }
-
-  if (outcome == "accuracy") {
-
-    return(plot_accuracy)
-  }
-
-  if (outcome == "competence") {
-
-    return(plot_competence)
-  }
-
-}
-```
-
-### b
-
-Second, we want a function that plots results obtained by the `vary_sample_options` function.
-
-```{r}
-plot_results <- function(data, variable = options) {
-
-  d <- data
-
-  # Extract variable name as a string
-  variable_name <- deparse(substitute(variable))
-
-  # retrieve info from choice of variable
-  if (variable_name == "options") {
-    x_label = "Number of Choice Options"
-    title = paste0("Sample: ", d$sample," \n iterations: ", d$total_iterations)
-  }
-    if (variable_name == "sample") {
-    x_label = "Size of informant groups"
-    title = paste0("Options: ", d$options," \n iterations: ", d$total_iterations)
-  }
-
-  # make constellation a factor with the right levels
-  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")
-
-  # plot for accuracy
-  plot_accuracy <- ggplot(d,
-                          aes(x = as.factor({{variable}}), y = average_accuracy, fill = constellation)) +
-    geom_half_violin(position = position_nudge(x = -.2),
-                     adjust=2, alpha = .8,
-                     side = "l") +
-    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-    # Add nice labels
-    labs(x = x_label, y = "Accuracy") +
-    scale_fill_viridis_d(option = "plasma", begin = 0.1, 
-                         limits = rev(levels(d$constellation)),
-                         direction = -1
-    ) +
-    plot_theme
-
-  # plot for competence
-  plot_competence <- ggplot(d,
-                            aes(x = as.factor({{variable}}), y = average_relative_competence, fill = constellation)) +
-    geom_half_violin(position = position_nudge(x = -.2),
-                     adjust=2, alpha = .8,
-                     side = "l") +
-    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-    # Add nice labels
-    labs(x = x_label, y = "Relative competence", fill = NULL) + 
-    scale_fill_viridis_d(option = "plasma", begin = 0.1, 
-                         limits = rev(levels(d$constellation)),
-                         direction = -1
-    ) +
-    plot_theme
-
-  ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE) +
-    plot_annotation(title = title)
-}
-```
-
-```{r, message=FALSE}
-n <- c(3, 10, 50, 100, 200)
-
-vary_sample_options(n = n, vary = "options", iterations = 100, population = 999, sample = 3, file_name = "model_3_options.csv")
-vary_sample_options(n = n, vary = "sample", iterations = 100, population = 999, options = 3, file_name = "model_3_samples.csv")
-
-data_3_options <- read_csv("data/model_3_options.csv")
-data_3_sample <- read_csv("data/model_3_samples.csv")
-
-plot_results(data_3_options)
-plot_results(data_3_sample, variable = sample)
-
-```
-
-```{r}
-data_3_sample %>% group_by(sample, constellation) %>% summarise(across(c(average_accuracy, average_relative_competence), mean))
-
-# make descriptive data
-data_3_sample %>% 
-  group_by(sample, constellation) %>% 
-  summarise(count = sum(count)) %>% 
-  mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
-  print(n = 100)
 ```
@@
 data_3 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 3)
-
 data_10 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 10)
-
-data_100 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 100)
-
-data_2 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 2)
 ```
@@
 plot_results(data_10)
-
-data_3 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
 ```
-
-```{r}
-data_10 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
-```
-
-```{r}
-data_2 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
-```
-
-```{r}
-plot_results(data_3)
-model_accuracy <- plot_results(data_3, outcome = "accuracy")
-model_competence <- plot_results(data_3, outcome = "competence")
-```
-
-```{r}
-# # example to check essential parts of the function
-# my_strings <- c("apple", "banana", "apple", "banana", "apple")
-# 
-# bla <- tibble(id = rep(1:5, each = 5), 
-#        string = sample(c("apple", "banana"), 25, replace = TRUE, 
-#                        prob = c(0.8, 0.2)))
-# 
-# bla %>% group_by(id, string) %>% 
-#   mutate(n_answer_in_group = n()) %>% 
-#   group_by(id) %>% 
-#   mutate(
-#     unique_answers = n_distinct(string),
-#     minority = ifelse(id == 4, 
-#                       n_answer_in_group == min(n_answer_in_group), 
-#                       NA
-#                       )
-#     ) %>% arrange(id) 
-```
-
-## Combine with plots from paper
-
-### Make plots from data
-```{r, message = FALSE}
-# read data
-d <- read_csv("./data/cleaned.csv")
-
-# make a categorical variable from `convergence`
-d <- d %>% 
-  mutate(convergence_categorical = recode_factor(convergence, 
-                                                 `0` = "opposing majority", 
-                                                 `1` = "divergence", 
-                                                 `2` = "majority", 
-                                                 `3` = "consensus",
-                                                 .default = NA_character_)
-         )
-```
-
-```{r}
-# plot for accuracy
-plot_accuracy <- ggplot(d,
-       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
-  geom_half_violin(position = position_nudge(x = -.2),
-                   adjust=2, alpha = .8,
-                   side = "l") +
-  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-  # Add nice labels
-  labs(x = "Convergence", y = "Accuracy") +
-  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
-  guides(fill = FALSE) +
-  plot_theme + 
-  theme(axis.text.x = element_text(angle = 20, hjust = 1))
-```
-
-```{r}
-# plot for competence
-plot_competence <- ggplot(d,
-       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
-  geom_half_violin(position = position_nudge(x = -.2),
-                   adjust=2, alpha = .8,
-                   side = "l") +
-  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-  # Add nice labels
-  labs(x = "Convergence", y = "Competence") +
-  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
-  guides(fill = FALSE) +
-  plot_theme + 
-  theme(axis.text.x = element_text(angle = 20, hjust = 1))
-```
-
-### combine plots from data and from model
-
-```{r}
-# accuracy
-plot_accuracy <- plot_accuracy + ggtitle("Participants")
-model_accuracy<- model_accuracy + ggtitle("Model")
-
-plot_accuracy + model_accuracy
-```
-
-```{r}
-# competence
-plot_competence <- plot_competence + ggtitle("Participants")
-model_competence<- model_competence + ggtitle("Model")
-
-plot_competence + model_competence
-```
-
-## Have only one population.
-
-```{r}
-simulate_single_population <- function(population, sample, options) {
-
-  # randomly draw competence levels
-  data <- tibble(id = 1:population,
-                 competence = runif(population, min = 1/options, max = 1))
-
-  # randomly draw individual answers based on competence levels 
-  data$answer <- data$competence %>% 
-    purrr::map_chr(function(x){
-
-      answer_options <- c("correct", paste0("false", 1:(options-1)))
-      probabilities <- c(x, rep((1-x)/(options-1), options-1))
-
-      answer = sample(answer_options, 
-                      size = 1, 
-                      prob = probabilities
-      )
-      return(answer)
-
-    })
-
-  # randomly assign samples in population
-  data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))
-
-  # identify constellations 
-  data <- data %>% 
-    # identify how often a one type of answer occurs in one group
-    group_by(sample_id, answer) %>% 
-  mutate(n_answer_in_sample = n()) %>% 
-    # assign constellations
-  group_by(sample_id) %>% 
-    mutate(unique_answers = n_distinct(answer),
-           # Build a first version of constellation variable
-           constellation = case_when(unique_answers == sample ~ "dissensus", 
-                                     unique_answers == 1 ~ "consensus",
-                                     unique_answers <  sample ~ "majority"
-           ),
-           # identify minority answers in majority constellations
-           minority = ifelse(constellation == "majority", 
-                             # report whether occurences of answer within a group
-                             # are the minority within that group
-                             n_answer_in_sample == min(n_answer_in_sample),
-                             # for all other constellations, simply code NA,
-                             NA
-           ), 
-           # identify majority answers in majority constellations
-           majority = ifelse(constellation == "majority", 
-                             # report whether occurrences of answer within a group
-                             # are the minority within that group
-                             n_answer_in_sample == max(n_answer_in_sample),
-                             # for all other constellations, simply code NA,
-                             NA
-           ), 
-           # modify the constellation variable to distinguish between minority, 
-           # intermediate majority, majority and dissensus that is composed of 
-           # multiple answers (e.g. 3 options, sample of 6, two per option)
-           constellation = case_when(
-             is.na(minority) | is.na(majority) ~ constellation, 
-             minority == TRUE & majority == TRUE ~ "dissensus",
-             minority == TRUE ~ "minority", 
-             majority == TRUE ~ "majority", 
-             .default = "intermediate majority")
-           ) %>% ungroup()
-
-  # identify accurate responses
-  data <- data %>%
-    mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))
-
-  # compute the summary statistics by constellation
-  results <- data %>% 
-    group_by(sample_id, constellation) %>% 
-    summarize(average_competence = mean(competence), 
-              average_accuracy = mean(accurate)) %>% 
-    # store simulation information
-    mutate(population = population, 
-           sample = sample, 
-           options = options) %>%
-
-  return(results)
-}
-
-```
-
-```{r}
-plot_results <- function(data) {
-
-  d <- data
-  # make constellation a factor with the right levels
-  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")
-
-  # make count data
-  counts <- d %>% group_by(constellation) %>% count()
-
-  # plot for accuracy
-  plot_accuracy <- ggplot(d,
-                          aes(x = constellation, y = average_accuracy, fill = constellation)) +
-    geom_half_violin(position = position_nudge(x = -.2),
-                     adjust=0.1, alpha = .8,
-                     side = "l") +
-    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-    # Add nice labels
-    labs(x = "Convergence", y = "Accuracy") +
-    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
-    scale_color_viridis_d(option = "plasma", begin = 0.7) +
-    guides(fill = FALSE, color = FALSE) +
-    plot_theme + 
-    theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
-    geom_text(data = counts, aes(x = constellation, y = 1.05, 
-                                 label =  paste0("n =", n)
-                                 )
-              )
-
-  # plot for competence
-  plot_competence <- ggplot(d,
-                            aes(x = constellation, y = average_competence, fill = constellation)) +
-    geom_half_violin(position = position_nudge(x = -.2),
-                     adjust=1, alpha = .8,
-                     side = "l") +
-    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-    # Add nice labels
-    labs(x = "Convergence", y = "Competence") +
-    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
-    scale_color_viridis_d(option = "plasma", begin = 0.7) +
-    guides(fill = FALSE) +
-    plot_theme + 
-    theme(axis.text.x = element_text(angle = 20, hjust = 1))
-
-  (plot_accuracy | plot_competence)
-}
-```
-
-```{r}
-test <- simulate_single_population(population = 999, sample = 3, options = 3)
-plot_results(test)
-```

### 2023-04-20 Experiment_5/analysis.Rmd
initialize repository

---
title: 
- "Analysis Experiment 5: Is convergence less trustworthy when informants are biased?"
author: 
- "Hugo Mercier, Jan Pfänder" 
date: "2023-02-02"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, error = TRUE) 
```

```{r, message=FALSE}
library(tidyverse) # for everything
library(tidyr) # some additional, neat tidy reshaping functions
library(lme4) # for linear mixed models
library(lmerTest) # p-values for mixed models
library(broom) # for tidy outputs of regressions
library(broom.mixed) # for tidy outputs of linear mixed models
library(ggridges) # for plots
library(gghalves) # for plots
library(ggbeeswarm) # Special distribution-shaped point jittering
library(scales)  # for nice axis labels
library(modelsummary) # regression tables
library(patchwork) # combine plots
library(ggrepel) # fit labels in plots 
```

### Some themes and functions

```{r, echo=FALSE}
# round numbers from models
rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)

#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

### Read data
```{r, message=FALSE}
# read data
d <- read_csv("./data/cleaned.csv")
```

### H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.

To test this hypothesis, we only consider participants assigned to the `independent` condition.

We use a linear mixed effect model with random intercept and random slope per participant.

In all our models we treat `convergence` as a continuous variable. We will, however, include robustness checks where we treat convergence as a categorical variable.

```{r, warning=FALSE, message=FALSE}
# models for accuracy

# random intercept and slope by participants
model_accuracy_independent <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                       data = d %>% filter(independence == "independent"))
```

### H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.

We will proceed in the same way for `competence` as we did for `accuracy` above.

```{r, warning=FALSE, message=FALSE}
# models for competence

# random intercept and slope by participants
model_competence_independent <- lmer(competence ~ convergence + 
                           (1 + convergence | id), 
                         data = d %>% filter(independence == "independent"))
```

### H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).

To test this hypothesis, we only consider the full data.

The resulting estimate of the interaction term will provide the test for our hypothesis.

```{r, warning=FALSE, message=FALSE}
# models for accuracy

# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = d)
```

### H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).

To test this hypothesis, we only consider the full data.

The resulting estimate of the interaction term will provide the test for our hypothesis.

```{r}
# models for competence

# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = d)

```

Show all results
```{r, echo=FALSE}
models <- list("Accuracy (independent only)" = model_accuracy_independent, 
               "Competence (independent only)" = model_competence_independent, 
               "Accuracy" = model_accuracy, 
               "Competence" = model_competence)
modelsummary::modelsummary(models, stars = TRUE)
```

# Robustness checks

In the models above, we treated convergence as a continuous variable. Based on the different levels, we will build a categorical variable, `convergence_categorical`.

```{r, echo=FALSE}
# make a categorical variable from `convergence`
d <- d %>% 
  mutate(convergence_categorical = recode_factor(convergence, 
                                                 `0` = "opposing majority", 
                                                 `1` = "divergence", 
                                                 `2` = "majority", 
                                                 `3` = "consensus",
                                                 .default = NA_character_)
         )

levels(d$convergence_categorical)

```

We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical`. This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e. "opposing majority").

### Plots

#### Independence Condition only

```{r, echo=FALSE}
# plot for accuracy
plot_accuracy <- ggplot(d %>% filter(independence == "independent"),
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r, echo=FALSE}
# plot for competence
plot_competence <- ggplot(d %>% filter(independence == "independent"),
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma") +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r, echo=FALSE}
plot_accuracy + plot_competence
```

#### Interaction

```{r, echo=FALSE}
# plot for accuracy
plot_accuracy <- ggplot(d,
       aes(x = convergence_categorical, y = accuracy, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r, echo=FALSE}
# plot for competence
plot_competence <- ggplot(d,
       aes(x = convergence_categorical, y = competence, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

```{r main-plot, echo=FALSE}
ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE)
```

### 2023-04-20 Experiment_5/cleaning.Rmd
initialize repository

--- 2023-04-20 Experiment_4/cleaning.Rmd
@@
 ---
-title: "Data cleaning experiment 4"
+title: "Data cleaning experiment 5"
 author: "Jan Pfänder, Hugo Mercier"
@@
 ## Import data
-```{r import}
+```{r import, message=FALSE, echo=FALSE}
 d <- read_csv("./data/qualtrics.csv")
@@
 
-```{r}
+```{r, echo=FALSE}
 # delete first two rows
@@
 
-## Attention checks 
-```{r attention}
+## Attention check
+```{r attention, echo=FALSE}
 # attention check
@@
 
-No one failed the attention check.
+There is one clearly failed attention check that we remove.
+
+```{r, echo=FALSE}
+# filter to only valid attention check responses
+d <- d %>% filter(str_detect(attention, "attention"))
+```
 
 ## Re-shape data
-```{r}
+```{r, echo=FALSE}
+# clean and re-shape
 d <- d %>% 
@@
   mutate(id = 1:nrow(.)) %>% 
+  # rename condition to `independence`
+  rename(independence = condition) %>% 
   # bring to long format

### 2023-04-20 Experiment_1/Sensitivity analysis.R

@@ -1,0 +1,146 @@
+# Sensitivity analysis (re-write the script so that number is in there as a control)
+
+library(tidyverse); library(purrr); library(broom); 
+set.seed(1000)
+
+# draw a sample
+create_data <- function(n_subj, n_low, n_high, beta_0, beta_1, 
+                        tau_0, tau_1, rho, sigma) { 
+  
+  items <- data.frame(
+    item_id = seq_len(n_low + n_high),
+    convergence = rep(c("low", "high"), c(n_low, n_high)),
+    X_i = rep(c(-0.5, 0.5), c(n_low, n_high)))
+  
+  # variance-covariance matrix for random effects by participant
+  cov_mx  <- matrix(
+    c(tau_0^2,             rho * tau_0 * tau_1,
+      rho * tau_0 * tau_1, tau_1^2            ),
+    nrow = 2, byrow = TRUE)
+  
+  subjects <- data.frame(subj_id = seq_len(n_subj),
+                         MASS::mvrnorm(n = n_subj,
+                                       mu = c(T_0s = 0, T_1s = 0),
+                                       Sigma = cov_mx))
+  
+  crossing(subjects, items)  %>%
+    mutate(e_si = rnorm(nrow(.), mean = 0, sd = sigma),
+           competence = beta_0 + T_0s + (beta_1 + T_1s) * X_i + e_si) %>%
+    select(subj_id, item_id, convergence, X_i, competence)
+}
+
+# estimate model based on sample
+est_model <- function(n_subj, n_low, n_high, beta_0, beta_1, tau_0, tau_1, rho, sigma)
+{ # residual (standard deviation)
+  
+  data_sim <- create_data(n_subj, n_low, n_high, beta_0, beta_1, tau_0, tau_1, rho, sigma)
+  model_sim <- lmer(competence ~ X_i + (1 + X_i | subj_id),
+                    data_sim)
+  
+  tidied <- broom.mixed::tidy(model_sim)
+  sig <- tidied$p.value[2] < .05
+  
+  return(sig)
+}
+
+
+# iterate sampling and estimation process 
+iterate <- function(n_subj, n_low, n_high, beta_0, beta_1, tau_0, tau_1, rho, sigma, iterations) 
+{ # number of iterations
+  
+  results <-  1:iterations %>%
+    map_dbl(function(x) {
+      # To keep track of progress
+      if (x %% 100 == 0) {print(paste("iteration number ", x))}
+      
+      # Run our model and return the result
[…]

### 2023-04-20 Experiment_3/power_random_effects.R

@@ -1,0 +1,293 @@
+# Power simulation for interaction model (a)
+## including random effects for participants
+
+#############
+# Attention: with the effect coding: competence scores are generated using
+# effect (-0.5, 05) and not dummy (0,1) coding for the factor variables. 
+# This is to specify meaningful random effects for the within factor 
+# 'convergence'.
+#
+# Among the two 'get_coefficients' functions, use the effect coding one to
+# generate coefficients from assumptions about condition means. 
+#############
+
+library(tidyverse); library(purrr); library(broom); library(lmerTest)
+set.seed(1000)
+
+# Generate data
+create_data <- function(n_subj, n_div, n_conv, beta_0, beta_1, beta_2, beta_3,
+                        tau_0, tau_1, rho, sigma) { 
+  
+  items <- data.frame(
+    item_id = seq_len(n_div + n_conv),
+    convergence = rep(c("div", "conv"), c(n_div, n_conv)),
+    # make the same as a numeric variable to generate competence
+    convergence_num = rep(c(-0.5, 0.5), c(n_div, n_conv)))
+  
+  # variance-covariance matrix
+  cov_mx  <- matrix(
+    c(tau_0^2,             rho * tau_0 * tau_1,
+      rho * tau_0 * tau_1, tau_1^2            ),
+    nrow = 2, byrow = TRUE)
+  
+  subjects <- data.frame(subj_id = seq_len(n_subj),
+                         MASS::mvrnorm(n = n_subj,
+                                       mu = c(T_0s = 0, T_1s = 0),
+                                       Sigma = cov_mx))
+  
+  crossing(subjects, items)  %>%
+    mutate(independence = rep(c("confl", "indep"), c(nrow(.)/2, nrow(.)/2)),
+           # make the same as a numeric variable to generate competence
+           independence_num = rep(c(-0.5, 0.5), c(nrow(.)/2, nrow(.)/2)),
+           # turn independence and convergence into factor variables
+           # for later analysis. Make sure the levels correspond to the 
+           # numeric variables we generate competence with
+           convergence = as.factor(convergence),
+           convergence = fct_relevel(convergence, "div", "conv"),
+           independence = as.factor(independence),
+           independence = fct_relevel(independence, "confl", "indep"),
+           # residual error
+           e_si = rnorm(nrow(.), mean = 0, sd = sigma),
+           # generate competence
[…]

### 2023-05-09 Experiment_4/model.Rmd
- update Y-axis scale in model plots

@@
   # different numbers of choice options. 
-  # This measure varies from 0 = chance to 1 = definitely right. 
+  # This measure varies from 0 = random chance to 1 = definitely right. 
   # We use min-max scaling, with min = 1/options and max = 1.
@@
 
-## Results by varying choice options and sample size function
+## Function for varying choice options and sample size function
 
@@
 
-### a
+### a) Plot various populations
 
@@
     plot_theme + 
-    theme(axis.text.x = element_text(angle = 20, hjust = 1))
+    theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
+    ylim(c(0, 1))
 
@@
 
-### b
+### b) Plot varying sample options
 
@@
 ```{r}
-plot_results <- function(data, variable = options) {
+plot_results_vary <- function(data, variable = options) {
 
@@
 
+## Simulate
+
 ```{r, message=FALSE}
@@
 
-plot_results(data_3_options)
-plot_results(data_3_sample, variable = sample)
+plot_results_vary(data_3_options)
+plot_results_vary(data_3_sample, variable = sample)
 
@@
 data_3 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 3)
-
-data_10 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 10)
-
-data_100 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 100)
-
-data_2 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 2)
 ```
@@
 ```{r}
-plot_results(data_3)
-plot_results(data_10)
-
 data_3 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
-```
-
-```{r}
-data_10 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
-```
-
-```{r}
-data_2 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
 ```
@@
 ```
-
-```{r}
-test <- simulate_single_population(population = 999, sample = 3, options = 3)
-plot_results(test)
-```

### 2023-05-22 Experiment_4/analysis.Rmd
add experiment 6 and slight model updates

@@
                                  data = d)
-
-summary(model_accuracy)
 ```
@@
                            (1 + convergence | id), data = d)
+```
 
-summary(model_competence)
+```{r}
+# combined results table
+modelsummary(list("Accuracy" = model_accuracy , "Competence" = model_competence), 
+             title = 'Models estimating effect of Convergence', 
+             stars = TRUE)
 ```
@@
 
-In the models above, we treated convergence as a continuous variable. Based on the different levels, we will build a categorical variable, `convergence_categorical`.
+In the models above, we treated convergence as a continuous variable.
+We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical` (where all levels are categories).
+This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e.
+"opposing majority").
 
 ```{r}
-# make a categorical variable from `convergence`
-d <- d %>% 
-  mutate(convergence_categorical = recode_factor(convergence, 
-                                                 `0` = "opposing majority", 
-                                                 `1` = "divergence", 
-                                                 `2` = "majority", 
-                                                 `3` = "consensus",
-                                                 .default = NA_character_)
-         )
-
-levels(d$convergence_categorical)
-
+# show levels of `convergence`
+levels(as.factor(d$convergence_categorical))
 ```
 
-We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical`.
-This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e.
-"opposing majority").
+```{r}
+# model for accuracy
+# random intercept and slope by participants
+model_accuracy <- lmer(accuracy ~ convergence_categorical + (1 + convergence_categorical | id), 
+                                 data = d)
+# model for competence
+# random intercept and slope by participants
+model_competence <- lmer(competence ~ convergence_categorical + 
+                           (1 + convergence_categorical | id), data = d)
+```
+
+```{r}
+# combined results table
+modelsummary(list("Accuracy" = model_accuracy , "Competence" = model_competence), 
+             title = 'Models estimating effect of Convergence', 
+             stars = TRUE)
+```
 

### 2023-05-22 Experiment_4/cleaning.Rmd
add experiment 6 and slight model updates

@@
   # all variables are coded as `character` - make key variables numeric
-  mutate(across(c(convergence, competence, accuracy), as.numeric)
+  mutate(across(c(convergence, competence, accuracy), as.numeric), 
+         # make a categorical variable from `convergence`
+         convergence_categorical = recode_factor(convergence, 
+                                                 `0` = "opposing majority", 
+                                                 `1` = "divergence", 
+                                                 `2` = "majority", 
+                                                 `3` = "consensus",
+                                                 .default = NA_character_)
          )

### 2023-05-22 Experiment_4/model.Rmd
add experiment 6 and slight model updates

@@
 
-We define competence simply as the probability of making an accurate choice (among various options).
-We assume that individuals (i) vary in competence,(ii) that competence is continuous and ranges from pure chance (`1/options`) to certainly making the correct choice (`1`), and that it is distributed uniformly in the population.
+We define competence as the probability of making an accurate choice (among various options).
+Competence is continuous and ranges from `0` to `1`, where `0` corresponds to pure chance (`1/options`) and `1` corresponds to certainly making the correct choice.
+
+In our code this measure is called `relative_competence`, while `competence` designates the actual probability of making an accurate choice, ranging from `1/options` to `1`).
+For example, in a 3-choice-options scenario, an individual with a probability of 1/3 (`competence == 0.333`) to pick the right answer has `relative_competence == 0`.
+
+We assume that individuals (i) vary in competence and (ii) that competence that it is distributed uniformly in the population.
 
@@
 population <- 10000
-competence = runif(population, min = 1/options, max = 1)
+data <- data.frame(competence = runif(population, min = 1/options, max = 1))
+
+# Create a relative measure of competence that allows to compare across
+# different numbers of choice options. 
+# This measure varies from 0 = random chance to 1 = definitely right. 
+# We use min-max scaling, with min = 1/options and max = 1.
+data <- data %>% 
+  mutate(relative_competence = (competence - 1/options) / (1 - 1/options) )
 
@@
   stat_function(fun = dunif, args = list(min = 0, max = 1)) +
-  labs(title = "Uniform distribution of competence from 1/3 to 1", x = "Competence", y = "P(Competence)") + 
-  xlim(1/3, 1)
+  labs(title = "Uniform distribution of competence from 0 to 1", x = "Competence", y = "P(Competence)") + 
+  xlim(0, 1)
 
 # generated population
-ggplot() +
-  geom_histogram(aes(x = competence)) + 
+ggplot(data, aes(x = relative_competence)) +
+  geom_histogram() + 
   scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
@@
 
-  # 2: randomly draw individual answers based on competence levels 
+  # 2: Draw individual answers based on competence levels 
   data$answer <- data$competence %>% 
@@
 ```{r}
-data_3_sample %>% group_by(sample, constellation) %>% summarise(across(c(average_accuracy, average_relative_competence), mean))
+regression_data <- data_3_options %>% 
+  filter(options %in% c(3, 10)) %>% 
+  group_by(iteration, constellation, options) %>% 
+  summarise(across(c(average_accuracy, average_relative_competence), mean)) %>% 
+  mutate(
+    convergence = case_when( 
+      constellation == "minority" ~ 0, 
+      constellation == "dissensus" ~ 1,
+      constellation == "majority" ~ 2,
+      constellation == "consensus" ~ 3), 
+    options = as.factor(options)
+  )
+
+accuracy <- lm(average_accuracy ~ convergence + options, 
+               data = regression_data) #%>% broom.mixed::tidy()
+competence <- lm(average_relative_competence ~ convergence + options, data = regression_data) %>% broom.mixed::tidy()
+
+summary(accuracy)
+```
+
+```{r}
+data_3_options %>% group_by(sample, constellation) %>% summarise(across(c(average_accuracy, average_relative_competence), mean))
 

### 2023-05-22 Experiment_6/analysis.Rmd
add experiment 6 and slight model updates

--- 2023-04-20 Experiment_5/analysis.Rmd
@@
 title: 
-- "Analysis Experiment 5: Is convergence less trustworthy when informants are biased?"
+- "Analysis Experiment 6: Is convergence less trustworthy when informants are biased?"
 author: 

### 2023-05-22 Experiment_6/power_Exp6.Rmd
add experiment 6 and slight model updates

---
title: "Power Analysis Experiment 6"
output: 
  html_document: 
    keep_md: yes
author: "Jan Pfänder, Hugo Mercier"
date: "2023-15-05"
bibliography: references.bib
---

```{r, message=FALSE}
# load required packages
library("lme4")        # model specification / estimation
library("lmerTest")    # get p-values for mixed models
library("broom.mixed") # extracting data from model fits 
library("tidyverse")   # data wrangling and visualisation
library("patchwork")    # combine plots

# ensure this script returns the same results on each run
set.seed(912783)
```

### Credit

The power simulation presented in this document draws heavily on @debruine2021. It is also inspired by @huntington-klein.

# Introduction

Statistical power is the probability by which we are able to detect an effect, assuming that it actually exists. The higher you set the benchmark for statistical power, the less likely you are to miss an effect in your sample.

The aim of a power analysis is find out how big a sample you need to achieve the statistical power you want. If there is an effect, more participants will make it more likely to detect it.

To give an overview of this power analysis, we will

1.  Make assumptions about the data generating process (i.e. a model and its parameters).
2.  "Draw" (i.e. simulate) a sample with a given sample size.
3.  Run the statistical analysis we plan to perform on that sample and store the estimates of the model.
4.  Calculate the power: Repeat steps two and three many times. See for how many samples the model recovers a statistically significant estimate for the parameter we set. Statistical power is just the ratio of significant samples among all samples we look at.
5.  Calculate power for different sample sizes: Repeat step four for various sample sizes. See at which sample size we achieve sufficiently high power.

# I. Building a power simulation function

## 1. Data generating process

### Variables

-   **Convergence**. We treat `convergence` as a continuous variable (0 = opposing majority; 1 = dissensus; 2 = majority; 3 = consensus).

-   **Number of choice options**. Binary variable indicating whether informants were choosing from `ten` or from `three` choice options.

-   **Accuracy**. We ask participants "What do you think is the probability of player [X] being correct?". Participants answer with a slider from 0 to 100.

-   **Competence**. We ask participants "How competent do you think player [X] is in games like these?" Participants answer on a 7-point Likert scale (from "not competent at all" to "extremely competent").

### Model

In our experiment, we manipulate two factors:

-   `convergence`, i.e. the extent to which several informants agree. We do so within subjects, meaning that each subject rates stimuli for all levels of convergence. Data points of the same subject are thus not independent. To account for that, in our simulation, we model subject differences with random effects.

-   `number_options`, i.e. whether informants were choosing from `ten` or from `three` choice options. We do so between subjects, meaning that a participant saw scenarios of only of one of the two levels.

The model we use to generate accuracy (and competence) is:

$$
\text{Accuracy} = (\beta_0 + b_{0, Subject}) + (\beta_c + b_{c, Subject}) \text{Convergence} + \beta_i \text{Number_options} + \beta_{ci} \text{Number_options*Convergence} + \epsilon
$$

with:

-   $b_{0, Subject} \sim N(0, \tau_0)$ (add footnote that this is not quite it yet, see below)
-   $b_{c, Subject} \sim N(0, \tau_c)$
-   $\epsilon \sim N(0, \sigma)$

where:

-   $\beta_0$ is the average intercept, i.e. average accuracy rating given that there is an opposing majority (`convergence = 0`)

-   $b_{0, Subject}$ is the subject-specific deviation from the average intercept

-   $\beta_c$ is the average effect of `convergence`, i.e. the average increase in `accuracy` caused by a one unit increase in `convergence`.

-   $b_{c, Subject}$ is the subject-specific deviation from the average convergence effect

-   $\beta_n$ is the average effect of `number_options`, i.e. the average difference in `accuracy` between the level `ten` and the level `three`.

-   $\beta_cn$ is the interaction effect between `number_options` and `convergence`, i.e. the difference in the effect of `convergence` on `accuracy` between the level `ten` and the level `three`.

-   $\epsilon$ is the error term that represents noise not accounted for by the random effects

-   $\sigma$ is the standard deviation of the normal distribution that of the error term.

-   $\tau_0)$ and $\tau_1)$ are the respective standard deviations of the normal distributions that model by-subject deviation from the average effects $\beta_0$ and $\beta_c$, respectively. We assume that the deviations of subjects are normally distributed around the average effects (hence a mean of `0`). Some subjects will have a positive deviation (i.e. larger values than the average); some will have a negative offset (i.e. smaller values than the average).

The modelling of $b_{0, Subject}$ and $b_{c, Subject}$ as described above was simplified. It assumed that both were gnerated from independent normal distributions. But in fact, we do not expect those distributions to be independent of each other. We expect a subject's intercept and effect convergence to be correlated. For example, a subject with a relatively *small intercept* (i.e. assigning very low accuracy when `convergence = 0`) might assign all the more accuracy with increasing convergence (i.e has a relatively *larger effect of convergence*). a larger effect of accuracy - case in which the distributions of both random effects are *negatively* correlated.

Instead of drawing to independent univariate distributions, we model two correlated distributions. We achieve this by modeling a bivariate normal distribution.

$$
\begin{bmatrix} b_{0, \text{Subject}} \\ b_{c, \text{Subject}} \end{bmatrix} \sim \text{N}\Bigg(\begin{bmatrix} 0 \\ 0 \end{bmatrix}, \begin{bmatrix} \tau_0^2 & \rho \tau_0 \tau_c \\ \rho \tau_0 \tau_c & \tau_c^2 \end{bmatrix}\Bigg)
$$

where

-   $\begin{bmatrix} \tau_0^2 & \rho \tau_0 \tau_c \\ \rho \tau_0 \tau_c & \tau_c^2 \end{bmatrix}$ is the variance-covariance matrix of the two distributions

-   $\rho$ is the correlation coefficient

### Parameters

We use the following prefixes to designate model parameters and sampled values:

-   `beta_*`: fixed effect parameters
-   `subj_*`: by-subject random effect parameters
-   `SU_*`: sampled values for by-subject random effects
-   `B_*`: sums of added beta's
-   `e_*`: residual sd

We use the following suffices:

-   `*_0`: intercept
-   `*_c`: convergence
-   `*_n`: number_options
-   `*_cn`: convergence\*number_options

Other terms:

-   `*_rho`: correlation for subject's random effects
-   `n_*`: sample size
-   `sigma`: residual (error) sd

The parameters we set differ with regard to our outcome variables. Later when we do the actual simulation, we will set two different sets of parameters, one for `accuracy` and one for `competence`, respectively. The steps before we introduce functions serve demonstration purposes only, so we'll just focus on `accuracy`.

We set the parameters for convergence and the by-participant random effects based on experiment 4.

We further use the modeling results to produce the expected difference between the categories.

```{r}
# data from model
model_data <- read_csv("../experiment_4/data/model_3_options.csv") %>% 
  filter(options %in% c(3, 10)) %>% 
  mutate(
    # make numeric version of convergence
    convergence = case_when(constellation == "minority" ~ 0, 
                                 constellation == "dissensus" ~ 1, 
                                 constellation == "majority" ~ 2, 
                                 constellation == "consensus" ~ 3),
    # scale accuracy from 0 to 100 (currently from 0 to 1)
    average_accuracy = 100*average_accuracy,
    # scale competence reaching from 1 to 7 (currently from 0 to 1)
    average_relative_competence = (average_relative_competence * (7 - 1)) + 1,
    # make choice options a factor
    options = as.factor(options)
         )

# we run the model on averages from different iterations
model_accuracy <- lm(average_accuracy ~ convergence + options + convergence*options, data = model_data) %>% 
  tidy()

model_competence <- lm(average_relative_competence ~ convergence + options + convergence*options, data = model_data) %>% 
  tidy()
```

```{r}
# parameters for modelling accuracy 
beta_0   <- 20 # intercept
beta_c   <- 10 # effect of convergence when conflict of interest
beta_i   <- 5 # effect of independence (compared to conflict of interest)
beta_ci  <- 5 # difference in effect of convergence when number of options = 10 (compared to conflict of interest)
subj_0   <- 10 # by-subject random intercept sd
subj_c   <- 5 # by-subject random slope sd
subj_rho <- .2 # correlation between intercept and slope 
sigma    <- 5 # residual (error) sd (i.e. all varition that the model cannot account for)
```

We store a complete list of parameters in a list so that we can call them for functions later.

```{r}
# complete list of parameters (including some that will be introduced later)
parameters <- list(
  # outcome variable
  outcome = "accuracy", # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  n_subj = 10,              # n participants
  levels_convergence = 0:3,  # levels of convergence
  n_stimuli_per_level = 2, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0 = 40,   # intercept
  beta_c = 10,   # effect of convergence, 
  beta_i = 5,    # # effect of independence (compared to conflict of interest)
  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0 = 10,   # by-subject random intercept sd
  subj_c = 5,   # by-subject random slope sd
  subj_rho = 0.5, # correlation between intercept and slope 
  sigma = 10     # residual (error) sd (i.e. all varition that the model cannot account for)
)
```

## 2. Simulate a single sample

### Stimuli

Each participant sees two stimuli per convergence level, i.e. `8` items in total.

```{r}
# set stimuli
levels_convergence <- c(0:3) 
n_stimuli_per_level  <-  2 
```

Simulate all stimuli a single participant sees.

```{r}
# create stimuli per participant data
stimuli <- data.frame(
  convergence = rep(levels_convergence, each = n_stimuli_per_level)
  )
```

### Subjects

```{r}
# set number of subjects and items
n_subj <- 100 # number of subjects we arbitrarily set for this example
```

When generating the subject data, we need to draw a pair of random effects (intercept + slope of convergence) for each participant. Earlier, we had defined the parameters for the bivariate normal distribution we draw from. We use the function `MASS::mvrnorm` to generate these draws. To be able to use this function, we define the variance-covariance matrix between the two by-subject random effects beforehand.

```{r}
# calculate random intercept / random slope covariance
covar <- subj_rho * subj_0 * subj_c

# put values into variance-covariance matrix
cov_mx  <- matrix(
  c(subj_0^2, covar,
    covar,   subj_c^2),
  nrow = 2, byrow = TRUE)

# generate the by-subject random effects
subject_rfx <- MASS::mvrnorm(n = n_subj,
                             mu = c(SU_0 = 0, SU_c = 0),
                             Sigma = cov_mx)

# make data 
subjects <- data.frame(
  # add a subject id
  id = seq_len(n_subj),
  # add by participant pair of random effects
  subject_rfx, 
  # add `independence` condition
  independence = rep(c("conflict", "independent"), length.out = n_subj)
  ) %>% 
  # make a numeric variable to be able to generate competence
  mutate(independence_num = ifelse(independence == "conflict", 0, 1))
```

Check how simulated values compare to the parameters we set earlier.

```{r}
data.frame(
  parameter = c("subj_0", "subj_c", "subj_rho"),
  value = c(subj_0, subj_c, subj_rho),
  simulated = c(sd(subjects$SU_0), sd(subjects$SU_c), 
                cor(subjects$SU_0, subjects$SU_c)
                )
)
```

### Trials (Subjects x Stimuli)

We combine the two data frames `subjects` and `stimuli` we generated. We also draw a residual error for each trial/observation (`e_ss` for error term for each combination of subject and stimulus).

```{r}
# cross subject and item IDs
trials <- expand_grid(subjects, stimuli)  %>%
  # add an error term
  # nrow(.) is the number of trials/observations, i.e. rows in the data frame
  mutate(e_ss = rnorm(nrow(.), mean = 0, sd = sigma))
```

#### Calculate accuracy values

Our data frame now contains all the information to calculate the accuracy values. Note that our`beta_` parameters are not part of the data frame, but they exist in the environment since we defined them earlier. Once we have added `accuracy` and we removed the helper variables we don't need anymore, we have the data frame that we can run our statistical analysis on.

We truncate the simulated `accuracy` values so that they remain on the scale from 0 to 100. This is to mimic our actual responses. It is conservative for the power calculation, since more extreme values are increasing effect size (thereby increasing power).

```{r}
dat_sim <- trials %>%
  mutate(accuracy = beta_0 + SU_0 + (beta_c + SU_c)*convergence + beta_i*independence_num + beta_ci*convergence*independence_num + e_ss,
         # truncate accuracy values so that they lie between 0 and 100 only
         accuracy = case_when(accuracy < 0 ~ 0, 
                              accuracy > 100 ~ 100,
                              TRUE ~ accuracy)) %>% 
  # retain only variables of interest
  select(id, convergence, independence, accuracy)
```

Plot the data to check values.

```{r}
# independent condition
independent_plot <- ggplot(dat_sim %>% filter(independence == "independent"), 
       aes(x = as.factor(convergence), y = accuracy, 
           color = as.factor(convergence)
       )
) +
  # predicted means
  geom_hline(yintercept = beta_0) +
  geom_hline(yintercept = (beta_0 + beta_c + beta_ci)) +
  geom_hline(yintercept = (beta_0 + 2*(beta_c + beta_ci))) +
  geom_hline(yintercept = (beta_0 + 3*(beta_c + beta_ci))) +
  # actual data
  geom_violin(alpha = 0, show.legend = FALSE) +
  stat_summary(fun = mean, geom="crossbar", show.legend = FALSE) +
  scale_color_viridis_d() +
  ggtitle("Predicted versus simulated values") + 
  facet_wrap(~independence) +
  ylim(0, 100)

# conflict condition
conflict_plot <-ggplot(dat_sim %>% filter(independence == "conflict"), 
       aes(x = as.factor(convergence), y = accuracy, 
           color = as.factor(convergence)
           )
       ) +
  # predicted means
  geom_hline(yintercept = beta_0) +
  geom_hline(yintercept = (beta_0 + beta_c)) +
  geom_hline(yintercept = (beta_0 + 2*beta_c)) +
  geom_hline(yintercept = (beta_0 + 3*beta_c)) +
  # actual data
  geom_violin(alpha = 0, show.legend = FALSE) +
  stat_summary(fun = mean, geom="crossbar", show.legend = FALSE) +
  scale_color_viridis_d() +
  ggtitle("Predicted versus simulated values")  + 
  facet_wrap(~independence) +
  ylim(0, 100)

independent_plot + conflict_plot
```

### Data generating function

We put all the steps to generate a single sample into a function we call `draw_single_sample`. We add an argument `outcome` to the function that allows to specify the outcome variable we want to simulate data for.

```{r}
# set up data generating function
draw_single_sample <- function(
  # outcome variable
  outcome, # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  n_subj,              # n participants
  levels_convergence,  # levels of convergence
  n_stimuli_per_level, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0,   # intercept
  beta_c,   # effect of convergence 
  beta_i = 5,    # # effect of independence (compared to conflict of interest)
  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0,   # by-subject random intercept sd
  subj_c,   # by-subject random slope sd
  subj_rho, # correlation between intercept and slope 
  sigma     # residual (error) sd (i.e. all variation that the model cannot account for)
) {

  # 1. create stimuli per subject data
  stimuli <- data.frame(
    convergence = rep(levels_convergence, each = n_stimuli_per_level)
  )

  # 2. create subject data

  # calculate random intercept / random slope covariance
  covar <- subj_rho * subj_0 * subj_c

  # put values into variance-covariance matrix
  cov_mx  <- matrix(
    c(subj_0^2, covar,
      covar,   subj_c^2),
    nrow = 2, byrow = TRUE)

  # generate the by-subject random effects
  subject_rfx <- MASS::mvrnorm(n = n_subj,
                               mu = c(SU_0 = 0, SU_c = 0),
                               Sigma = cov_mx)

  # make data 
  subjects <- data.frame(
    # add a subject id
    id = seq_len(n_subj),
    # add by participant pair of random effects
    subject_rfx,
    # add `independence` condition
    independence = rep(c("conflict", "independent"), length.out = n_subj)
    ) %>% 
    # make a numeric variable to be able to generate competence
    mutate(independence_num = ifelse(independence == "conflict", 0, 1))

  # 3. create trials data (subjects x stimuli) & simulate outcome variables

  ## cross subject and item IDs
  sample <- expand_grid(subjects, stimuli)  %>%
    # add an error term
    # nrow(.) is the number of trials/observations, i.e. rows in the data frame
    mutate(e_ss = rnorm(nrow(.), mean = 0, sd = sigma)) %>%
    # simulate outcome variable
    mutate(outcome = beta_0 + SU_0 + (beta_c + SU_c)*convergence + 
             beta_i*independence_num + beta_ci*convergence*independence_num + e_ss, 
           outcome_variable = {{outcome}}) %>% 
    # retain only variables of interest
    select(id, convergence, independence, outcome, outcome_variable)

  # Truncate results so that we only get values on the correct scale
  if (outcome == "accuracy") {

    sample <- sample %>% mutate(
      # truncate accuracy values so that they lie between 0 and 100 only
      outcome = case_when(outcome < 0 ~ 0,
                           outcome > 100 ~ 100,
                           TRUE ~ outcome)
      )
  }

  if (outcome == "competence") {

    sample <- sample %>% mutate(
      # truncate accuracy values so that they lie between 0 and 100 only
      outcome = case_when(outcome < 1 ~ 1,
                           outcome > 7 ~ 7,
                           TRUE ~ outcome)
      )
  }

  return(sample)
}

# # You can test the function using the commented code below 
# # (un-comment by highlighting, then command+shit+c)
# test <- do.call(draw_single_sample, parameters)
```

## 3. Analyze the simulated data

We run a linear mixed model using the `lme4` package on our simulated sample.

```{r}
mod_sim <- lmer(accuracy ~ convergence + independence + 
                  independence*convergence + (1 + convergence | id), 
                dat_sim)
summary(mod_sim)
```

We use the `broom.mixed` package to get a tidy table of the results. Below, we added column "parameter" and "value" to compare the estimate from the model to the parameters used to simulate the data.

```{r}
# get a tidy table of results
model_estimates <- broom.mixed::tidy(mod_sim) %>% 
  mutate_if(is.numeric, round, 3) %>%
  mutate(
    parameter = c("beta_0", "beta_c", "beta_i", "beta_ci", "subj_0", "subj_rho", "subj_c", "sigma"),
    value = c(beta_0, beta_c, beta_i, beta_ci, subj_0, subj_rho, subj_c, sigma),
  ) %>%
  select(effect:term, parameter, value, estimate:p.value)

model_estimates %>% knitr::kable()
```

This table provides us with several estimates. What we are ultimately interested in is whether the the `p.value` of the `convergence` effect is smaller than the significance threshold `alpha` we set.

The `p.value` is the probability of obtaining a test statistic at least as extreme (i.e. far away from `0`) as the one observed, given the null hypothesis is true. Even in a world where there is no true effect in the population altogether, there is still a chance that looking at a random sample of that population can (misleadingly) yield an effect, i.e. an estimate other than `0`. That chance is described by the p-value. Everything else equal, the larger the estimate of the effect of convergence, the smaller the p-value.

Our `alpha` is the threshold we set, at which we consider the p-value sufficiently small to "safely" reject the null hypothesis. The smaller we set alpha, the less risk we run in mistakenly detecting an effect. The alpha is commonly set at `0.05`.

We are interested in the p.value of three effects: `convergence`, `independent` and `convergence*independent`

```{r}
# set alpha
alpha <- 0.05

# estimates to extract
estimates = c("convergence", "independenceindependent", 
              "convergence:independenceindependent")

# check significance and store result
significance <- model_estimates %>% 
  filter(term %in% estimates) %>% 
  select(term, p.value) %>% 
  mutate(significant = ifelse(p.value < alpha, TRUE, FALSE))

significance
```

### Model estimation function

We put the model estimation into a function.

```{r}
# set up model estimation function
estimate_model <- function(alpha = 0.05, ...) {

  # Alpha is the probability we accept of rejecting the null hypothesis
  # when in fact it is true. The smaller alpha, the less likely the
  # results are to have occurred by "accident" (i.e. sampling variation).

  # ... is a shortcut that forwards any additional arguments to draw_single_sample()

  # draw a sample
  sample <- draw_single_sample(...)

  # estimate model for the drawn sample
  model_estimates <- lmer(outcome ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id),
                          sample) %>% 
    # put into tidy output
    broom.mixed::tidy()

  # estimates to extract
  estimates = c("convergence", "independenceindependent", 
                "convergence:independenceindependent")

  # check significance and store the result
  significance <- model_estimates %>% 
  filter(term %in% estimates)  %>% 
  mutate(significant = ifelse(p.value < alpha, TRUE, FALSE))

  # return the p_value (as tibble)
  return(significance)
}

# # You can test the function using the commented code below 
# # (un-comment by highlighting, then command+shit+c)
# test <- do.call(estimate_model, parameters)

```

## 4. Calculate Power

The idea of a power simulation is to draw many samples, run our statistical model on each, and check how often we get a significant estimate for `convergence`. The share of samples for which we get a significant result among all samples we generate is our estimate of the statistical power - our ability to detect an effect in a sample, *given that there is one in the population*.

To calculate the statistical power, we set the number of samples we want to draw (`iterations`) and then use the functions we created earlier.

```{r}
# define how many samples to generate
iterations <- 5

# we run the estimate_model() function as often as specified by `iterations` and store the data
power_data <- purrr::map_df(1:iterations, 
                              # We use do.call here since we stored the
                              # parameters in an external list. We could
                              # also define them directly in 
                              # estimate_model().
                              ~do.call(estimate_model, parameters))

# we group by estimate (or `term`) 
# the calculate the share of significant estimates
power <- power_data %>% 
  group_by(term) %>% 
  summarize(power = mean(significant))
```

### Power calculation function

We put the power calculation above into a function. We slightly modify this function compared to above - for example, since that function will take some time to run, we add a print message.

```{r}
calculate_power <- function(iterations, ...) {

  # create data frame with model results for generated samples
  power_data <- 1:iterations %>% 
    purrr::map_df(function(x){
      # this is essentially a for loop - do the following for each 
      # element in 1:iterations

      results <- estimate_model(...)

      # To keep track of progress
      if (x %% 50 == 0) {print(paste("iteration number ", x))}

      return(results)

    })

  # we group by estimate (or `term`) 
  # the calculate the share of significant estimates
  power <- power_data %>% 
    group_by(term) %>% 
    summarize(power = mean(significant))

  return(power)
}
# # You can test the function using the commented code below 
# # (un-comment by highlighting, then command+shit+c)
# test <- do.call(calculate_power, c(parameters, list(iterations = 500)))
```

## 5. Calculate Power for different sample sizes

The aim of a power analysis is to inform the number of participants to recruit, *given a desired level of statistical power*.

We set this level at 90%.

```{r}
power_threshold <- 0.9
```

We are looking for a sample size that let's us detect a statistically significant effect in (at least) 90% of samples.

To do so, we repeat the previous power calculation for various sample sizes. Here, we pick only two and a low number of iterations since it is only for illustration.

```{r warning=FALSE, message=FALSE}
# calcluate power for different sample sizes

# make a data frame with sample sizes
power_by_sample_size <- tibble(n_subj = seq(from = 10, to = 20, by = 10))

# Remove "n_subj" parameter from our initial parameter list since we want to replace it 
# with the respective sample sizes every time.
parameters[["n_subj"]] <- NULL

# do the `calculate_power()` function for each sample size and store the results
# in a new variable called `power`
power_by_sample_size <- map_df(power_by_sample_size$n_subj, 
                                      ~do.call(calculate_power, 
                                               c(parameters,
                                                 # sets the n_subj value according
                                                 # to the respective value in
                                                 # `power_by_sample_size`
                                                 list(n_subj = .x, 
                                                      iterations = 100))))

```

We can check the resulting data to find the first sample size where power is at least 90%.

```{r}
power_by_sample_size
```

### Power by sample size function

We can put the above power calculation for different sample sizes into a function - which, finally is the function we were looking for.

The simulations that this function executes will take quite some time. Therefore, we do not want to run it every time we render this document. Instead we want to store the output of the power simulation in a `.csv` file, and have an integrated "stop" mechanism to prevent execution when that file already exists. To achieve this, we make `file_name` a mandatory argument. If a file with that name already exists, the function will not be executed.

```{r}
# file_name <- "power_simulation.csv" # change for new analyses / or delete file to re-use same name

power_by_sample_size <- function(file_name, sample_sizes, iterations, outcome, ...) {

  # only run analysis if a file with that name does not yet exists
  if (!file.exists(paste0("data/", file_name))) {

    # do the `calculate_power()` function for each sample size and store the results
    # in a new variable called `power`
    power <- purrr::map_df(sample_sizes, 
                           function(n){
                             # this is essentially a for loop - 
                             # do the following for each 
                             # element data$n_subj

                             # To keep track of progress
                             print(paste("tested sample size = ", n))

                             # run power calculation
                             result <- calculate_power(n_subj = n, 
                                                       iterations = iterations, 
                                                       outcome = outcome, ...)
                             # identify respective sample size
                             result$n <- n

                             return(result)

                           })

    # add some variables with information about simulation
    data <- power %>% 
      mutate(
        # add identifier variable for number of iterations
        iterations = iterations, 
        # add identifier variable for outcome
        outcome = outcome)

    write_csv(data, paste0("data/", file_name))
  }
}

# # You can test the function using the commented code below
# # (un-comment by highlighting, then command+shit+c)
# test <- do.call(power_by_sample_size,
#                 c(parameters,
#                   list(file_name = file_name, sample_sizes = c(5, 10),
#                        iterations = 100)
#                   )
#                 )
```

# II. Running the power analyses

We will run two power analyses: one for `accuracy` and one for `competence`. We try to make reasonable parameter assumptions for both. Our choice of sample size will be based on the analysis that requires the larger sample.

### Set parameters

For accuracy.

```{r}
# list of parameters for accuracy
parameters_accuracy <- list(
  # outcome variable
  outcome = "accuracy", # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  sample_sizes  =  seq(from = 20, to = 200, by = 20), # different n's to simulate power for 
  levels_convergence = 0:3,  # levels of convergence
  n_stimuli_per_level = 2, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0 = 40,   # intercept
  beta_c = 5,   # effect of convergence 
  beta_i = 5,    # effect of independence (compared to conflict of interest)
  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0 = 10,   # by-subject random intercept sd
  subj_c = 5,   # by-subject random slope sd
  subj_rho = -0.5, # correlation between intercept and slope 
  sigma = 10,     # residual (error) sd (i.e. all varition that the model cannot account for)
  # number of iterations (per sample size)
  iterations = 1000
)
```

For competence.

```{r}
# list of parameters for competence
parameters_competence <- list(
  # outcome variable
  outcome = "competence", # takes names as input, e.g. "accuracy"
  # parameters for size of data frame
  sample_sizes  =  seq(from = 20, to = 300, by = 20), # different n's to simulate power for 
  levels_convergence = 0:3,  # levels of convergence
  n_stimuli_per_level = 2, # n stimuli per participant
  # parameters for modelling outcome variable
  beta_0 = 3.5,   # intercept
  beta_c = 0.25,   # effect of convergence 
  beta_i = 0.25,    # # effect of independence (compared to conflict of interest)
  beta_ci = 0.5, # difference in effect of convergence when independence (compared to conflict of interest)
  subj_0 = 1,   # by-subject random intercept sd
  subj_c = 0.25,   # by-subject random slope sd
  subj_rho = -0.5, # correlation between intercept and slope 
  sigma = 1,     # residual (error) sd (i.e. all varition that the model cannot account for)
  # number of iterations (per sample size)
  iterations = 1000
)
```

### Execute functions

For accuracy

```{r, message=FALSE}
do.call(power_by_sample_size, c(parameters_accuracy, 
                                        list(file_name = "power_accuracy.csv")
                                        )
                )

power_accuracy <- read_csv("data/power_accuracy.csv")
```

For competence

```{r, message=FALSE}
do.call(power_by_sample_size, c(parameters_competence, 
                                        list(file_name = "power_competence.csv")
                                        )
                )

power_competence <- read_csv("data/power_competence.csv")
```

### Plot results

We can plot the results of this power calculation.

```{r}
# plot results
plot_power <- function(data) {

  # make nicer labels for the term variable
  data <- data %>% 
    mutate(term = case_when(
      term == "independenceindependent" ~ "independence", 
      term == "convergence:independenceindependent" ~ "interaction", 
      TRUE ~ term)
    )

  ggplot(data, 
       aes(x = n, y = power, color = term)) +
  geom_line(size = 1.5) + 
  # add a horizontal line at 90%, our power_threshold
  geom_hline(aes(yintercept = .9), linetype = 'dashed') + 
  # display iterations
  annotate("text", x = 150, y = 0.25, label = paste("iterations per \nsample size =", max(data$iterations)))+
  # Prettify!
  theme_minimal() + 
    scale_colour_viridis_d(option = "plasma") + 
  scale_y_continuous(labels = scales::percent) + 
  labs(x = 'Sample Size', y = 'Power', 
       title = paste0("Power analysis for ", unique(data$outcome)),
       subtitle = "(accounting for participant random effects)")
}
```

```{r plot-power-competence}
competence_plot <- plot_power(power_competence)
competence_plot
```

```{r plot-power-accuracy}
accuracy_plot <- plot_power(power_accuracy)
accuracy_plot
```

## References

### 2023-05-22 Experiment_6/preregistration.Rmd
add experiment 6 and slight model updates

---
title: 'Preregistration Experiment 6: Is convergence more trustworthy when there are more choice options?'
author: "Jan Pfänder, Hugo Mercier"
date: "2023-03-08"
output:
  html_document:
    keep_md: yes
  pdf_document: 
      keep_md: yes
---

```{r, include=FALSE}
# load packages
library(tidyverse)
library(kableExtra)
library(lme4)
```

```{r, include=FALSE}
# generate some random data with respective variables to ensure 
# using code snippets that work

# simulate data
data <- tibble(
  convergence = sample(0:3, replace = TRUE, size=100),
  accuracy = sample(0:100, replace = TRUE, size=100),
  competence = sample(1:7, replace = TRUE, size=100),
  options = rep(c("3", "10"), each = 50),
  id = rep(1:5, length.out = 100)
  ) 
```

# I. Introduction

In the three previous experiments, we have shown that people infer trustworthiness from convergence - the extent to which different informants agree. Greater convergence makes (i) an information to be perceived as more accurate and (ii) the informants as more competent - that is, given certain context.

Some aspects of that context we held constant; others we varied to evaluate how they alter the effect of convergence. We held constant that participants did not have any priors whatsoever - not on the information they received (they knew nothing about the games that were played) nor about the informants (the players who played the games). We manipulated informants’ independence. We tested two cases of independence, by varying the baseline of what dependent meant: In experiment two, dependent informants had discussed “at great length” before making their estimates. In experiment three, dependent informants shared the same conflict of interest of making a certain estimate.

In this first series of experiments (experiments one, two and three) we tested inferences from convergence in a numerical choice setting: Participants saw (fictive) players’ estimates that would lie with 1000 and 2000. The degree of convergence varied by the distance between estimates.

In a second series of experiments (experiments four, five and six), we test inferences from convergence in a categorical choice setting. In the categorical scenario, the fictive players make choices on a set of response options (i.e. categories).  Convergence varies by the ratio of people agreeing on an option. Experiment four and five can be considered robustness checks as to whether the results of the first series hold in a categorical choice setting. Experiment six tests a new context factor: the number of choice options. In experiments four and five, scenarios always involved three choice options. Here, we will vary between three and ten options. The design we use to manipulate convergence is otherwise identical to experiment four (players playing games).

# II. Data collection

No data has been collected yet.

# III. Design

Participants see eight situtations in which three advisors give investment recommendations. They are then asked to judge one advisor's accuracy and competence. We will manipulate two factors: convergence (four levels, within participants) and number of options (two levels, between participants).

> **Introduction for participants:** *"To be able to understand the task, please read the following instructions carefully: Some people are playing games in which they have to select the correct answer among three answers. You will see the results of several of these games. Each game is different, with different solutions and involving different players. All players answer independently of each other. At first, you have no idea how competent each individual player is: they might be completely at chance, or be very good at the task. It's also possible that some players are really good while others are really bad. Some games might be difficult while others are easy. Your task will be to evaluate the performance of one of the players based on what everyone's answers are."*

```{r -stimulus, echo=FALSE, fig.cap="An example of a stimulus for the majority condition"}
knitr::include_graphics("figures/example_stimulus.png")
```

**Convergence**. Convergence varies by the ratio of players choosing the same response as the focal player (i.e. the one that participants evaluate). The levels of convergence are: (i) consensus, where all three players pick the same option [`coded value = 3`]; (ii) majority, where either the third or second player picks the same option as the first player [`coded value = 2`]; (iii) dissensus, where all three players pick different options [`coded value = 1`]; (iv) majority against the focal player's estimate, where the second and third player pick the same option, but one that is different from the first player's choice [`coded value = 0`]. In our analysis, we treat convergence as a continuous variable, assigning the values in squared parenthesis.

We manipulate convergence within participants. All participants see all four conditions of convergence, with two stimuli (i.e. game results) per condition. Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli) .

```{r stimuli-3, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(condition = c("opposing majority (0)", 
                                        "dissensus (1)", "majority (2)", 
                                        "consensus (3)"),
                          imgage_a = c("![](figures/stimuli/minority_3_a.png){ width=60% }", 
                        "![](figures/stimuli/divergence_3_a.png){ width=60% }",
                        "![](figures/stimuli/majority_3_a.png){ width=60% }", 
                        "![](figures/stimuli/consensus_3_a.png){ width=60% }"),
                        imgage_b = c("![](figures/stimuli/minority_3_b.png){ width=60% }", 
                        "![](figures/stimuli/divergence_3_b.png){ width=60% }",
                        "![](figures/stimuli/majority_3_b.png){ width=60% }", 
                        "![](figures/stimuli/consensus_3_b.png){ width=60% }"))

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "Stimuli for 3 options condition by levels of convergence")

```

**Options**. Number of options has two levels: (i) '3' and (ii) '10'.

We manipulate independence between participants, i.e. each participant gets assigned to either the '3' or the '10' condition. All stimuli that participants see in their respective condition will involve the same number of choice options.

```{r stimuli-options, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(convergence = c("consensus (3)"),
                          imgage_a = c("![](figures/stimuli/consensus_3_a.png){ width=60% }"),
                        imgage_b = c("![](figures/stimuli/consensus_10_a.png){ width=60% }"))

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Convergence", "Number of options: 3", "Number of options: 10"), 
                  align = "c",
                  caption = "Example of a consensus stimulus for the two 'Number of option' conditions")
```

As outcome variables, we will measure people's perceived accuracy and competence of player one.

**Accuracy**. We ask participants "What do you think is the probability of player 1 being correct?". Participants answer with a slider from 0 to 100.

**Competence**. We ask participants "How competent do you think player 1 is in games like these?" Participants answer on a 7-point Likert scale (from "not competent at all" to "extremely competent").

# IV. Hypotheses

Considering only `3` options condition of the experiment at hand, we expect the findings of experiment four to replicate. In experiment four, players would choose among `3` options.

### H1a: In the three options condition, participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.

To test this hypothesis, we only consider participants assigned to the `3` options condition.

We use a linear mixed effect model with random intercept and random slope per participant. Should this model yield convergence issues, we will use a model with random intercept only.

In all our models we treat `convergence` as a continuous variable. We will, however, include robustness checks where we treat convergence as a categorical variable, allowing to inspect difference between different levels.

```{r, warning=FALSE, message=FALSE}
# models for accuracy

# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + (1 + convergence | id), 
                       data = data %>% filter(options == "3"))

# in case of non-convergence: random intercept by participants only
alt_model_accuracy <- lmer(accuracy ~ convergence + (1 | id), 
                           data = data %>% filter(options == "3"))
```

### H1b:  In the three options condition, participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants, given three choice options.

To test this hypothesis, we only consider participants assigned to the `3` options condition.

We will proceed in the same way for `competence` as we did for `accuracy` above.

```{r, warning=FALSE, message=FALSE}
# models for competence

# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + 
                           (1 + convergence | id), 
                         data = data %>% filter(options == "3"))

# in case of non-convergence: random intercept by participants only
alt_model_competence <- lmer(competence ~ convergence + (1 | id), 
                             data = data %>% filter(options == "3"))
```

How about a context in which informants (i.e. the fictive players) could choose not among 3, but among 10 options? In line with our findings from experiment three, we predict that effects of convergence are more positive when there are more choice options:

### H2a: The effect of convergence on accuracy (H1a) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.

To test this hypothesis, we consider the full data.

The resulting estimate of the interaction term will provide the test for our hypothesis.

```{r, warning=FALSE, message=FALSE}
# models for accuracy

# random intercept and slope by participants
model_accuracy <- lmer(accuracy ~ convergence + options + 
                            options*convergence + (1 + convergence | id), 
                       data = data)

# in case of non-convergence: random intercept by participants only
alt_model_accuracy <- lmer(accuracy ~ convergence + options + 
                            options*convergence + (1 | id), 
                           data = data)
```

### H2b: The effect of convergence on competence (H1b) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.

To test this hypothesis, we consider the full data.

The resulting estimate of the interaction term will provide the test for our hypothesis.

```{r, warning=FALSE, message=FALSE}
# models for competence

# random intercept and slope by participants
model_competence <- lmer(competence ~ convergence + options + 
                            options*convergence + (1 + convergence | id), 
                       data = data)

# in case of non-convergence: random intercept by participants only
alt_model_competence <- lmer(competence ~ convergence + options + 
                            options*convergence + (1 | id), 
                           data = data)
```

# Robustness checks

## Convergence as categorical variable

In the models above, we treated convergence as a continuous variable. Based on the different levels, we will build a categorical variable, `convergence_categorical`.

```{r}
# make a categorical variable from `convergence`
data <- data %>% 
  mutate(convergence_categorical = recode_factor(convergence, 
                                                 `0` = "opposing majority", 
                                                 `1` = "divergence", 
                                                 `2` = "majority", 
                                                 `3` = "consensus",
                                                 .default = NA_character_)
         )

levels(data$convergence_categorical)

```

We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical`. This also allows us to inspect heterogeneity in differences between levels (with respect to the baseline, i.e. "opposing majority").

## Alternative stimuli with greater distance

We kept the same distance between 10 and 3. That's conservative - people might think that other options are not relevant and hence don't make a difference between 3 and 10 options. We therefore run the hypotheses set two comparing 3 to a version of 10 with larger distance.

Pick an example stimulus as earlier for illustrating choice options. See all stimuli in appendix.

# Exclusions

We will exclude participants failing (i.e. participants not answering the question or writing anything that does not at least resemble "I pay attention") the following attention check:

> *Imagine you are playing video games with a friend and at some point your friend says: "I don't want to play this game anymore! To make sure that you read the instructions, please write the three following words"I pay attention" in the box below. I really dislike this game, it's the most overrated game ever. Do you agree with your friend?*

# Power analysis

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found in the `power_Exp5.Rmd` document. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%.

The power simulation for `accuracy` suggested that for 80 participants, we would have a power of at least 90% for the interaction effect. The simulation for `competence` suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we also detect an effect of convergence.

However, due to uncertainty about our assumptions and because we are anticipating failed attention checks, *we will recruit a sample of `200` participants*.

```{r accuracy-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
#knitr::include_graphics("power_Exp5_files/figure-html/plot-power-accuracy-1.png")
```

```{r competence-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
#knitr::include_graphics("power_Exp5_files/figure-html/plot-power-competence-1.png")
```

## Appendix

```{r stimuli-10, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(condition = c("opposing majority (0)", 
                                        "dissensus (1)", "majority (2)", 
                                        "consensus (10)"),
                          imgage_a = c("![](figures/stimuli/minority_10_a.png){ width=60% }", 
                        "![](figures/stimuli/divergence_10_a.png){ width=60% }",
                        "![](figures/stimuli/majority_10_a.png){ width=60% }", 
                        "![](figures/stimuli/consensus_10_a.png){ width=60% }"),
                        imgage_b = c("![](figures/stimuli/minority_10_b.png){ width=60% }", 
                        "![](figures/stimuli/divergence_10_b.png){ width=60% }",
                        "![](figures/stimuli/majority_10_b.png){ width=60% }", 
                        "![](figures/stimuli/consensus_10_b.png){ width=60% }"))

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "Stimuli for 10 options condition by levels of convergence")

```

```{r stimuli-10-alt, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(condition = c("opposing majority (0)", 
                                        "dissensus (1)", "majority (2)", 
                                        "consensus (10)"),
                          imgage_a = c("![](figures/stimuli/minority_10_a_alt.png){ width=60% }", 
                        "![](figures/stimuli/divergence_10_a_alt.png){ width=60% }",
                        "![](figures/stimuli/majority_10_a_alt.png){ width=60% }", 
                        "![](figures/stimuli/consensus_10_a_alt.png){ width=60% }"),
                        imgage_b = c("![](figures/stimuli/minority_10_b_alt.png){ width=60% }", 
                        "![](figures/stimuli/divergence_10_b_alt.png){ width=60% }",
                        "![](figures/stimuli/majority_10_b_alt.png){ width=60% }", 
                        "![](figures/stimuli/consensus_10_b_alt.png){ width=60% }"))

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "Alternative stimuli for 10 options condition by levels of convergence")

```

### 2023-05-22 Experiment_6/preregistration.tex
add experiment 6 and slight model updates

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Preregistration Experiment 6: Is convergence more trustworthy when there are more choice options?},
  pdfauthor={Jan Pfänder, Hugo Mercier},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Preregistration Experiment 6: Is convergence more trustworthy
when there are more choice options?}
\author{Jan Pfänder, Hugo Mercier}
\date{2023-03-08}

\begin{document}
\maketitle

\hypertarget{i.-introduction}{%
\section{I. Introduction}\label{i.-introduction}}

In the three previous experiments, we have shown that people infer
trustworthiness from convergence - the extent to which different
informants agree. Greater convergence makes (i) an information to be
perceived as more accurate and (ii) the informants as more competent -
that is, given certain context.

Some aspects of that context we held constant; others we varied to
evaluate how they alter the effect of convergence. We held constant that
participants did not have any priors whatsoever - not on the information
they received (they knew nothing about the games that were played) nor
about the informants (the players who played the games). We manipulated
informants' independence. We tested two cases of independence, by
varying the baseline of what dependent meant: In experiment two,
dependent informants had discussed ``at great length'' before making
their estimates. In experiment three, dependent informants shared the
same conflict of interest of making a certain estimate.

In this first series of experiments (experiments one, two and three) we
tested inferences from convergence in a numerical choice setting:
Participants saw (fictive) players' estimates that would lie with 1000
and 2000. The degree of convergence varied by the distance between
estimates.

In a second series of experiments (experiments four, five and six), we
test inferences from convergence in a categorical choice setting. In the
categorical scenario, the fictive players make choices on a set of
response options (i.e.~categories). Convergence varies by the ratio of
people agreeing on an option. Experiment four and five can be considered
robustness checks as to whether the results of the first series hold in
a categorical choice setting. Experiment six tests a new context factor:
the number of choice options. In experiments four and five, scenarios
always involved three choice options. Here, we will vary between three
and ten options. The design we use to manipulate convergence is
otherwise identical to experiment four (players playing games).

\hypertarget{ii.-data-collection}{%
\section{II. Data collection}\label{ii.-data-collection}}

No data has been collected yet.

\hypertarget{iii.-design}{%
\section{III. Design}\label{iii.-design}}

Participants see eight situtations in which three advisors give
investment recommendations. They are then asked to judge one advisor's
accuracy and competence. We will manipulate two factors: convergence
(four levels, within participants) and options (two levels, between
participants).

\begin{quote}
\textbf{Introduction for participants:} \emph{``To be able to understand
the task, please read the following instructions carefully: Some people
are playing games in which they have to select the correct answer among
three answers. You will see the results of several of these games. Each
game is different, with different solutions and involving different
players. All players answer independently of each other. At first, you
have no idea how competent each individual player is: they might be
completely at chance, or be very good at the task. It's also possible
that some players are really good while others are really bad. Some
games might be difficult while others are easy. Your task will be to
evaluate the performance of one of the players based on what everyone's
answers are.''}
\end{quote}

\begin{figure}
\centering
\includegraphics{figures/example_stimulus.png}
\caption{An example of a stimulus for the majority condition}
\end{figure}

\textbf{Convergence}. Convergence varies by the ratio of players
choosing the same response as the focal player (i.e.~the one that
participants evaluate). The levels of convergence are: (i) consensus,
where all three players pick the same option
{[}\texttt{coded\ value\ =\ 3}{]}; (ii) majority, where either the third
or second player picks the same option as the first player
{[}\texttt{coded\ value\ =\ 2}{]}; (iii) dissensus, where all three
players pick different options {[}\texttt{coded\ value\ =\ 1}{]}; (iv)
majority against the focal player's estimate, where the second and third
player pick the same option, but one that is different from the first
player's choice {[}\texttt{coded\ value\ =\ 0}{]}. In our analysis, we
treat convergence as a continuous variable, assigning the values in
squared parenthesis.

We manipulate convergence within participants. All participants see all
four conditions of convergence, with two stimuli (i.e.~game results) per
condition. Each participant therefore sees eight stimuli in total (4
convergence levels x 2 stimuli) .

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1704}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4148}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4148}}@{}}
\caption{Stimuli for 3 options condition by levels of
convergence}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endhead
opposing majority (0) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_3_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_3_b.png} \\
dissensus (1) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_3_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_3_b.png} \\
majority (2) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_3_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_3_b.png} \\
consensus (3) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_3_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_3_b.png} \\
\bottomrule()
\end{longtable}

\textbf{Options}. Independence has two levels: (i) `3' and (ii) `10'.

We manipulate independence between participants, i.e.~each participant
gets assigned to either the `3' or the `10' condition. All stimuli that
participants see in their respective condition will involve the same
number of choice options.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1230}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4344}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4426}}@{}}
\caption{Example of a consensus stimulus for the two `Option'
conditions}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Convergence
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Options: 3
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Options: 10
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Convergence
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Options: 3
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Options: 10
\end{minipage} \\
\midrule()
\endhead
consensus (3) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_3_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_10_a.png} \\
\bottomrule()
\end{longtable}

As outcome variables, we will measure people's perceived accuracy and
competence of player one.

\textbf{Accuracy}. We ask participants ``What do you think is the
probability of player 1 being correct?''. Participants answer with a
slider from 0 to 100.

\textbf{Competence}. We ask participants ``How competent do you think
player 1 is in games like these?'' Participants answer on a 7-point
Likert scale (from ``not competent at all'' to ``extremely competent'').

\hypertarget{iv.-hypotheses}{%
\section{IV. Hypotheses}\label{iv.-hypotheses}}

Considering only \texttt{3} options condition of the experiment at hand,
we expect the findings of experiment four to replicate. In experiment
four, players would choose among \texttt{3} options.

\hypertarget{h1a-participants-perceive-an-estimate-of-an-independent-informant-as-more-accurate-the-more-it-converges-with-the-estimates-of-other-informants-given-three-choice-options.}{%
\subsubsection{H1a: Participants perceive an estimate of an independent
informant as more accurate the more it converges with the estimates of
other informants, given three choice
options.}\label{h1a-participants-perceive-an-estimate-of-an-independent-informant-as-more-accurate-the-more-it-converges-with-the-estimates-of-other-informants-given-three-choice-options.}}

To test this hypothesis, we only consider participants assigned to the
\texttt{3} options condition.

We use a linear mixed effect model with random intercept and random
slope per participant. Should this model yield convergence issues, we
will use a model with random intercept only.

In all our models we treat \texttt{convergence} as a continuous
variable. We will, however, include robustness checks where we treat
convergence as a categorical variable, allowing to inspect difference
between different levels.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# models for accuracy}

\CommentTok{\# random intercept and slope by participants}
\NormalTok{model\_accuracy }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(accuracy }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ convergence }\SpecialCharTok{|}\NormalTok{ id), }
                       \AttributeTok{data =}\NormalTok{ data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(options }\SpecialCharTok{==} \StringTok{"3"}\NormalTok{))}

\CommentTok{\# in case of non{-}convergence: random intercept by participants only}
\NormalTok{alt\_model\_accuracy }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(accuracy }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ id), }
                           \AttributeTok{data =}\NormalTok{ data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(options }\SpecialCharTok{==} \StringTok{"3"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\hypertarget{h1b-participants-perceive-an-independent-informant-as-more-competent-the-more-their-estimate-converges-with-the-estimates-of-other-informants-given-three-choice-options.}{%
\subsubsection{H1b: Participants perceive an independent informant as
more competent the more their estimate converges with the estimates of
other informants, given three choice
options.}\label{h1b-participants-perceive-an-independent-informant-as-more-competent-the-more-their-estimate-converges-with-the-estimates-of-other-informants-given-three-choice-options.}}

To test this hypothesis, we only consider participants assigned to the
\texttt{3} options condition.

We will proceed in the same way for \texttt{competence} as we did for
\texttt{accuracy} above.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# models for competence}

\CommentTok{\# random intercept and slope by participants}
\NormalTok{model\_competence }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(competence }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+} 
\NormalTok{                           (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ convergence }\SpecialCharTok{|}\NormalTok{ id), }
                         \AttributeTok{data =}\NormalTok{ data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(options }\SpecialCharTok{==} \StringTok{"3"}\NormalTok{))}

\CommentTok{\# in case of non{-}convergence: random intercept by participants only}
\NormalTok{alt\_model\_competence }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(competence }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ id), }
                             \AttributeTok{data =}\NormalTok{ data }\SpecialCharTok{\%\textgreater{}\%} \FunctionTok{filter}\NormalTok{(options }\SpecialCharTok{==} \StringTok{"3"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

How about a context in which informants (i.e.~the fictive players) could
choose not among 3, but among 10 options? In line with our findings from
experiment three, we predict that effects of convergence are more
positive when there are more choice options:

\hypertarget{h2a-the-effect-of-convergence-on-accuracy-h1a-is-more-positive-in-a-context-where-informants-can-choose-among-ten-response-options-compared-to-when-they-can-choose-among-only-three.}{%
\subsubsection{H2a: The effect of convergence on accuracy (H1a) is more
positive in a context where informants can choose among ten response
options compared to when they can choose among only
three.}\label{h2a-the-effect-of-convergence-on-accuracy-h1a-is-more-positive-in-a-context-where-informants-can-choose-among-ten-response-options-compared-to-when-they-can-choose-among-only-three.}}

To test this hypothesis, we consider the full data.

The resulting estimate of the interaction term will provide the test for
our hypothesis.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# models for accuracy}

\CommentTok{\# random intercept and slope by participants}
\NormalTok{model\_accuracy }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(accuracy }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+}\NormalTok{ options }\SpecialCharTok{+} 
\NormalTok{                            options}\SpecialCharTok{*}\NormalTok{convergence }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ convergence }\SpecialCharTok{|}\NormalTok{ id), }
                       \AttributeTok{data =}\NormalTok{ data)}

\CommentTok{\# in case of non{-}convergence: random intercept by participants only}
\NormalTok{alt\_model\_accuracy }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(accuracy }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+}\NormalTok{ options }\SpecialCharTok{+} 
\NormalTok{                            options}\SpecialCharTok{*}\NormalTok{convergence }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ id), }
                           \AttributeTok{data =}\NormalTok{ data)}
\end{Highlighting}
\end{Shaded}

\hypertarget{h2b-the-effect-of-convergence-on-competence-h1b-is-more-positive-in-a-context-where-informants-can-choose-among-ten-response-options-compared-to-when-they-can-choose-among-only-three.}{%
\subsubsection{H2b: The effect of convergence on competence (H1b) is
more positive in a context where informants can choose among ten
response options compared to when they can choose among only
three.}\label{h2b-the-effect-of-convergence-on-competence-h1b-is-more-positive-in-a-context-where-informants-can-choose-among-ten-response-options-compared-to-when-they-can-choose-among-only-three.}}

To test this hypothesis, we consider the full data.

The resulting estimate of the interaction term will provide the test for
our hypothesis.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# models for competence}

\CommentTok{\# random intercept and slope by participants}
\NormalTok{model\_competence }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(competence }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+}\NormalTok{ options }\SpecialCharTok{+} 
\NormalTok{                            options}\SpecialCharTok{*}\NormalTok{convergence }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{+}\NormalTok{ convergence }\SpecialCharTok{|}\NormalTok{ id), }
                       \AttributeTok{data =}\NormalTok{ data)}

\CommentTok{\# in case of non{-}convergence: random intercept by participants only}
\NormalTok{alt\_model\_competence }\OtherTok{\textless{}{-}} \FunctionTok{lmer}\NormalTok{(competence }\SpecialCharTok{\textasciitilde{}}\NormalTok{ convergence }\SpecialCharTok{+}\NormalTok{ options }\SpecialCharTok{+} 
\NormalTok{                            options}\SpecialCharTok{*}\NormalTok{convergence }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{1} \SpecialCharTok{|}\NormalTok{ id), }
                           \AttributeTok{data =}\NormalTok{ data)}
\end{Highlighting}
\end{Shaded}

\hypertarget{robustness-checks}{%
\section{Robustness checks}\label{robustness-checks}}

\hypertarget{convergence-as-categorical-variable}{%
\subsection{Convergence as categorical
variable}\label{convergence-as-categorical-variable}}

In the models above, we treated convergence as a continuous variable.
Based on the different levels, we will build a categorical variable,
\texttt{convergence\_categorical}.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# make a categorical variable from \textasciigrave{}convergence\textasciigrave{}}
\NormalTok{data }\OtherTok{\textless{}{-}}\NormalTok{ data }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{convergence\_categorical =} \FunctionTok{recode\_factor}\NormalTok{(convergence, }
                                                 \StringTok{\textasciigrave{}}\AttributeTok{0}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"opposing majority"}\NormalTok{, }
                                                 \StringTok{\textasciigrave{}}\AttributeTok{1}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"divergence"}\NormalTok{, }
                                                 \StringTok{\textasciigrave{}}\AttributeTok{2}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"majority"}\NormalTok{, }
                                                 \StringTok{\textasciigrave{}}\AttributeTok{3}\StringTok{\textasciigrave{}} \OtherTok{=} \StringTok{"consensus"}\NormalTok{,}
                                                 \AttributeTok{.default =} \ConstantTok{NA\_character\_}\NormalTok{)}
\NormalTok{         )}

\FunctionTok{levels}\NormalTok{(data}\SpecialCharTok{$}\NormalTok{convergence\_categorical)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## [1] "opposing majority" "divergence"        "majority"         
## [4] "consensus"
\end{verbatim}

We run the same models outlined in the hypotheses section, but replacing
\texttt{convergence} with \texttt{convergence\_categorical}. This also
allows us to inspect heterogeneity in differences between levels (with
respect to the baseline, i.e.~``opposing majority'').

\hypertarget{alternative-stimuli-with-greater-distance}{%
\subsection{Alternative stimuli with greater
distance}\label{alternative-stimuli-with-greater-distance}}

Pick an example stimulus as earlier for illustrating choice options. See
all stimuli in appendix.

\hypertarget{exclusions}{%
\section{Exclusions}\label{exclusions}}

We will exclude participants failing (i.e.~participants not answering
the question or writing anything that does not at least resemble ``I pay
attention'') the following attention check:

\begin{quote}
\emph{Imagine you are playing video games with a friend and at some
point your friend says: ``I don't want to play this game anymore! To
make sure that you read the instructions, please write the three
following words''I pay attention'' in the box below. I really dislike
this game, it's the most overrated game ever. Do you agree with your
friend?}
\end{quote}

\hypertarget{power-analysis}{%
\section{Power analysis}\label{power-analysis}}

We ran a power simulation to inform our choice of sample size. All
assumptions and details on the procedure can be found in the
\texttt{power\_Exp5.Rmd} document. We ran two different power analyses,
one for each outcome variable. We set the power threshold for our
experiment to 90\%.

The power simulation for \texttt{accuracy} suggested that for 80
participants, we would have a power of at least 90\% for the interaction
effect. The simulation for \texttt{competence} suggested that with
already 40 participants, we would detect an interaction, but only with
60 participants we also detect an effect of convergence.

However, due to uncertainty about our assumptions and because we are
anticipating failed attention checks, \emph{we will recruit a sample of
\texttt{200} participants}.

\hypertarget{appendix}{%
\subsection{Appendix}\label{appendix}}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1679}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4161}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4161}}@{}}
\caption{Stimuli for 10 options condition by levels of
convergence}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endhead
opposing majority (0) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_10_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_10_b.png} \\
dissensus (1) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_10_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_10_b.png} \\
majority (2) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_10_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_10_b.png} \\
consensus (10) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_10_a.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_10_b.png} \\
\bottomrule()
\end{longtable}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1608}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4196}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4196}}@{}}
\caption{Alternative stimuli for 3 options condition by levels of
convergence}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endhead
opposing majority (0) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_3_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_3_b_alt.png} \\
dissensus (1) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_3_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_3_b_alt.png} \\
majority (2) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_3_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_3_b_alt.png} \\
consensus (3) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_3_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_3_b_alt.png} \\
\bottomrule()
\end{longtable}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1586}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4207}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4207}}@{}}
\caption{Alternative stimuli for 10 options condition by levels of
convergence}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule()
\endhead
opposing majority (0) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_10_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/opp_majority_10_b_alt.png} \\
dissensus (1) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_10_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/divergence_10_b_alt.png} \\
majority (2) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_10_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/majority_10_b_alt.png} \\
consensus (10) &
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_10_a_alt.png}
&
\includegraphics[width=0.6\textwidth,height=\textheight]{figures/stimuli/consensus_10_b_alt.png} \\
\bottomrule()
\end{longtable}

\end{document}

### 2023-05-31 Experiment_1/model.Rmd
adding model for exp 1 to 3 and cleaning model for exp 4 to 6

---
title: "Model for numerical choice scenarios (experiments 1 to 3)"
output: 
  html_document: 
    keep_md: yes
date: "2023-04-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, echo=FALSE}
# Load packages
library(tidyverse)
library(gghalves) # for plots
library(patchwork)
library(lme4) # for linear mixed models
library(lmerTest) # p-values for mixed models
library(broom) # for tidy outputs of regressions
library(broom.mixed) # for tidy outputs of linear mixed models
```

```{r, echo = FALSE}
#set general theme for plots
plot_theme <- theme_minimal(base_size = 12) +
  theme(# Bold, bigger title
        plot.title = element_text(face = "bold", size = rel(1.7)),
        # Plain, slightly bigger subtitle that is grey
        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
        # Bold legend titles
        legend.title = element_text(face = "bold"),
        # Bold axis titles
        axis.title = element_text(face = "bold"))
```

```{r}
# ensure this script returns the same results on each run
set.seed(8726)
```

## Explaining the model

```{r}
# Imagine a situation in which answers range from 1000 to 2000
range <- 1000
# and a population of n = 999
population <- 999
# and we observe samples of 3
sample <- 3
```

### Define competence and its distribution

We define competence as the value of the standard deviation from which one's answer is drawn. We suppose that answers for all individuals are drawn from normal distributions. Each individual has their own normal distribution. All normal distributions are centered around the true answer, but they differ in their standard deviations. The higher the competence, the lower the standard deviation, i.e. the more certain a guess drawn from the normal distribution will be close to the true answer.

We (arbitrarily) set the *lowest* competence to the range of possible values, in our case 2000 - 1000 = 1000. We set the *highest* competence to 1/1000 of the range of possible values, in our case 1.

```{r}
# Define the x-axis values
x <- seq(1000, 2000, length.out = 1000)

# Define the PDFs for the two distributions
high_competence_pdf <- dnorm(x, mean = 1500, sd = 1)
low_competence_pdf <- dnorm(x, mean = 1500, sd = 1000)

# Create the plot
ggplot() +
  geom_line(aes(x, high_competence_pdf, color = "Highest Competence Individual \n (SD = 1, Mean = 1500)"), size = 1) +
  geom_line(aes(x, low_competence_pdf, color = "Lowest Competence Individual \n (SD = 1000, Mean = 1500)"), size = 1) +
  labs(x = "Competence Level", y = "Density", color = "Data generating function for") +
  ggtitle("Competence Level Distributions") +
  scale_color_manual(values = c("Highest Competence Individual \n (SD = 1, Mean = 1500)" = "blue", 
                                "Lowest Competence Individual \n (SD = 1000, Mean = 1500)" = "red")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

We suppose that these competence levels are drawn from a random uniform distribution. That is, in a population, each level of competence is equally likely to occur.

In order to be able to compare competence measures across different ranges, we want to scale it to reach from 0 (minimal competence) to 1 (maximal competence).

```{r}
# 1: Distribution of competence
data <- data.frame(competence = runif(population, 
                                      min = range/1000, 
                                      max = range)) %>% 
  # Scale competence to reach from 0 to 1 (instead of 1 to 1000).
  # Reverse the scale such that 1 becomes maximal competence (instead of 0).
  mutate(competence_scaled = 1 - ((competence - 1) / (1000-1)))

# generated population
ggplot(data, aes(x = competence_scaled)) +
  geom_histogram() + 
  scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.1)) + 
  annotate("text", x = 0.9, y = 100, label = paste0("n = ", population), color = "red") +
  labs(title = "(sampled) Population of competence drawn from \n uniform population distribution", 
       y = "Count")
```
### Generate a choice for each individual based on their competence.

The numbers that we generate from the normal distributions are tuncated such that they all lie within the pre-defined range (1000 to 2000)

```{r}
# 2: Draw individual answers based on competence levels 
mean <- 1500

data$answer <- data$competence %>% 
  purrr::map_dbl(function(x){

    sd <- x

    answer = truncnorm::rtruncnorm(1, mean = mean, sd = sd, a = 1000, b = 2000)
  }
  )

ggplot(data, aes(x = answer)) +
  geom_histogram()
```

### Measure accuracy

We measure accuracy as the (absolute value of) distance between the chosen answer and the true answer for each individual.

We let this distance be negative, such that higher values represent higher accuracy.

```{r}
# 3: measure accuracy
data <- data %>% 
  mutate(accuracy = -1 * abs(mean - answer))

ggplot(data, aes(x = accuracy)) +
  geom_histogram() +

ggplot(data, aes(x = competence_scaled, y = accuracy)) +
  geom_point()
```

### Draw samples, calculate convergence & compare average outcomes

### Randomly draw samples from the population

```{r}
# 4: randomly assign samples in population
data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))
```

### Calculate convergence

We define convergence as the standard deviation of the answers within a sample (multiplied by -1 so that higher values correspond to greater convegence)

```{r}
# 5: calculate convergence 
data <- data %>% 
  # identify how often a one type of answer occurs in one group
  group_by(sample_id) %>% 
  mutate(convergence = -1 * sd(answer)) %>% 
  ungroup()

ggplot(data, aes(x = convergence)) +
  geom_histogram()
```

### Compute average accuracy/competence per sample.

```{r}
# 6: calculate accuracy and competence per sample

# compute the summary statistics for each sample
results <- data %>% 
  group_by(sample_id, convergence) %>% 
  summarize(competence_mean = mean(competence_scaled),
            accuracy_mean = mean(accuracy)) %>% 
  # store simulation information
  mutate(population = population, 
         sample = sample)
```

```{r}
accuracy_plot <- ggplot(results, aes(x = convergence, y = accuracy_mean)) +
  geom_point() +
  # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Accuracy \n(average distance to true mean)") +
  plot_theme

competence_plot <- ggplot(results, aes(x = convergence, y = competence_mean)) +
  geom_point() +
  # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Competence \n(SD for data generating function, transposed)") +
  plot_theme

accuracy_plot + competence_plot
```

## Functions

### Data generating functions

#### Single population

```{r, echo=FALSE}
simulate_single_population <- function(population, sample, min_range, max_range, 
                                       competence_ratio = 1000) {

  # Check if population is not divisible by sample without remainder

  if (population %% sample != 0) {
    warning("Population is not divisible by sample without remainder.")
  }

  # if necessary, alter population size so that it is divisible by sample without rest
  n_possible_samples <- floor(population/sample)

  # Biggest possible population to pick
  possible_population = n_possible_samples * sample

  # Issue warning if different population size used
  if (population %% sample != 0) {
    warning(paste0("Chosen population size is ", possible_population))
  }
  # change population value (if divisible without remainder, it'll be the same)
  population = possible_population

  # Set variables
  range = max_range - min_range
  mean = min_range + (max_range - min_range)/2

  # 1: Distribution of competence
  data <- data.frame(competence = runif(population, 
                                        min = range/competence_ratio, 
                                        max = range)) %>% 
    # Scale competence to reach from 0 to 1.
    # Reverse the scale such that 1 becomes maximal competence (instead of 0).
    mutate(competence_scaled = 1 - ((competence - range/competence_ratio) / 
                                      (range - range/competence_ratio)))

  # 2: Draw individual answers based on competence levels

  data$answer <- data$competence %>% 
    purrr::map_dbl(function(x){

      sd <- x

      answer = truncnorm::rtruncnorm(1, mean = mean, sd = sd, a = min_range, 
                                     b = max_range)
    }
    )

  # 3: measure accuracy
  data <- data %>% 
    mutate(accuracy = -1 * abs(mean - answer))

  # 4: randomly assign samples in population
  data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))

  # 5: calculate convergence 
  data <- data %>% 
    group_by(sample_id) %>% 
    mutate(convergence = -1 * sd(answer)) %>% 
    ungroup()

  # 6: calculate accuracy and competence levels per sample
  results <- data %>% 
    group_by(sample_id, convergence) %>% 
    summarize(competence_mean = mean(competence_scaled),
              accuracy_mean = mean(accuracy)) %>% 
    # store simulation information
    mutate(population = population, 
           sample = sample) %>% 
    ungroup()

  return(results)
}

```

```{r, echo=FALSE}
# # # test output (set return to data above)
# test <- simulate_single_population(population = 500, sample = 3,
#                                    min_range = 1, max_range = 6,
#                                    competence_ratio = 1000)
```

#### Various populations

```{r, echo=FALSE}
simulate_various_populations <- function(iterations,...) {

  # create data frame with model results for generated samples
  various_populations <- 1:iterations %>% 
    purrr::map_df(function(x){
      # this is essentially a for loop - do the following for each 
      # element in 1:iterations

      results <- simulate_single_population(...)

      # identify iteration
      results$iteration <- x

      # To keep track of progress
      if (x %% 50 == 0) {print(paste("iteration number ", x))}

      return(results)

    }) %>% 
    # store simulation information
    mutate(total_iterations = iterations)

  return(various_populations)
}
```

```{r, echo=FALSE}
# # try out code
# test <- simulate_various_populations(population = 500, sample = 3,
#                                    min_range = 1, max_range = 6,
#                                    competence_ratio = 1000,
#                                    iterations = 100)
```

#### Vary sample size

This function allows us to investigate how accuracy and competence change with varying the sample size.

The simulations that this function executes will take quite some time. Therefore, we do not want to run it every time we render this document. Instead we want to store the output of the power simulation in a `.csv` file, and have an integrated "stop" mechanism to prevent execution when that file already exists. To achieve this, we make `file_name` a mandatory argument. If a file with that name already exists, the function will not be executed.

```{r, echo=FALSE}
# function for different sample sizes
vary_sample <- function(file_name, n, ...) {

  # only run analysis if a file with that name does not yet exists
  if (!file.exists(paste0("data/", file_name))) {

    # do the `simulate_various_populations()` function for each sample size 
    # and store the results in a new data_frame
    data <- n %>% purrr::map_df(function(n_sample){
      # this is essentially a for loop - 
      # do the following for each 
      # element data$n_subj

      # To keep track of progress
      print(paste("tested sample number = ", n_sample))

      # run power calculation
      result <- simulate_various_populations(
        sample = n_sample, ...)

      return(result)

    })

    write_csv(data, paste0("data/", file_name))
  }
}
```

```{r, echo=FALSE}
# # try out code
# n <- c(3, 10)
# test <- vary_sample(population = 999,
#                     min_range = 1000, max_range = 2000,
#                     competence_ratio = 1000,
#                     iterations = 100,
#                     file_name = "test.csv",
#                     n = n)
```

### Plot functions

#### a) Plot various populations

First, we want a function that plots results obtained by the `simulate_various_populations` function. 
```{r, echo=FALSE}
plot_results <- function(data, outcome = "everything") {

  d <- data

  # extract simulation info
  simulation_info <- d %>% 
    summarize(across(c(population, sample, total_iterations), max)) %>% 
    pivot_longer(everything(), names_to = "parameter", values_to = "value") %>% 
    gridExtra::tableGrob(cols = NULL, rows = NULL)

  plot_accuracy <- ggplot(d, aes(x = convergence, y = accuracy_mean)) +
    geom_hex() +
    # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Accuracy \n(average distance to true mean)") +
    plot_theme +
  scale_fill_viridis_c(option = "plasma", begin = 0.1)

  plot_competence <- ggplot(d, aes(x = convergence, y = competence_mean)) +
    geom_hex() +
    # Add nice labels
    labs(x = "Convergence \n(SDs within samples of three)", 
         y = "Competence \n(SD for data generating function, transposed)") +
    plot_theme +
  scale_fill_viridis_c(option = "plasma", begin = 0.1)

  if (outcome == "everything") {

    return((plot_accuracy | plot_competence) / simulation_info)
  }

  if (outcome == "accuracy") {

    return(plot_accuracy)
  }

  if (outcome == "competence") {

    return(plot_competence)
  }

}
```

#### b) Plot varying sample & options

Second, we want a function that plots results obtained by the `vary_sample()` function.

```{r, echo=FALSE}
plot_results_vary <- function(data, outcome = "everything", sample) {

  # get different samples as vector
  samples <- data %>% 
    reframe(unique(as.character(sample))) %>% pull()

  # empty list
  plot_list <- list()

  # make plots fro each moderator
  for (i in samples) {

    plot_data <- data %>%
      filter(sample == i)

    sample_plot <- plot_results(data = plot_data)

    plot_list[[i]] <- sample_plot
  }

  return(plot_list)
}
```

## Simulation

### Generate data

```{r, message=FALSE}
n <- c(3, 10, 20, 50)

# run simulation and store results in .csv files
vary_sample(population = 999,
            min_range = 1000, max_range = 2000,
            competence_ratio = 1000,
            iterations = 100,
            file_name = "model_vary_3_10_20_50.csv",
            n = n)

# read simulated data from .csv files
data <- read_csv("data/model_vary_3_10_20_50.csv")
```

### Plot

```{r}
# plot results
plot_results_vary(data)
```

### Analyze & Compare to participant data

It's not clear to me yet what we would want to compare.

In the experiment, we generated guesses from two different distribution widths: one with SD = 20 and another with SD = 150. In the model, this corresponds to the competence. However, we did not measure the observed SD of the guesses (which is the measure of convergence I rely on in the models).

Below, what I did was predict accuracy and competence for convergence levels of 20 and 150 based on a regression run on the model data. We could compare that to the participant data - but it's not quite the same thing...

```{r, echo=FALSE}
# possible analysis
regression_data <- data %>% 
  filter(sample %in% c(3, 10)) %>% 
  mutate(
    # makes sample a factor
    sample = as.factor(sample),
    # Scale accuracy to reach from 1 to 7 
    # a) scale the variable to range between 0 and 1
    scaled_accuracy_mean = (accuracy_mean - min(accuracy_mean)) / 
      (max(accuracy_mean) - min(accuracy_mean)),
    # b) scale to reach from 1 to 7
    accuracy_mean = 6*scaled_accuracy_mean + 1,
    # scale competence to reach from 1 to 7 (instead of 0 to 1)
    competence_mean = 6*competence_mean + 1
  )

accuracy <- lm(accuracy_mean ~ convergence + sample + convergence*sample, 
               data = regression_data) 
competence <- lm(competence_mean ~ convergence + sample + convergence*sample, 
               data = regression_data)

summary(competence)

# Create a new data frame with values for prediction
values_to_predict <- data.frame(convergence = c(-20, -150), 
                                sample = rep(c("3", "10"), each = 2))

# Predict values using the fitted model
predictions <- bind_cols(values_to_predict,
                         predict(accuracy, newdata = values_to_predict) %>% 
                           tidy() %>% 
                           pivot_longer(x, 
                                       names_to = "bla", 
                                       values_to = "predicted_accuracy") %>% 
                           select(predicted_accuracy),
                         predict(competence, newdata = values_to_predict) %>% 
                           tidy() %>% 
                           pivot_longer(x, 
                                       names_to = "bla", 
                                       values_to = "predicted_competence") %>% 
                           select(predicted_competence)
                         )

predictions

```

### 2023-05-31 Experiment_4/model.Rmd
adding model for exp 1 to 3 and cleaning model for exp 4 to 6

@@
 ---
-title: "Model"
+title: "Model for categorical choice scenarios (experiments 4 to 6)"
 output: 
@@
 
-## Load packages
-```{r, Message=FALSE}
+```{r, message=FALSE, echo=FALSE}
+# Load packages
 library(tidyverse)
@@
 library(patchwork)
+library(lme4) # for linear mixed models
+library(lmerTest) # p-values for mixed models
+library(broom) # for tidy outputs of regressions
+library(broom.mixed) # for tidy outputs of linear mixed models
 ```
 
-## Set plot theme
-```{r}
+```{r, echo=FALSE}
 #set general theme for plots
@@
 
+```{r}
+# ensure this script returns the same results on each run
+set.seed(7643)
+```
+
 ## Explaining the model
+
+```{r}
+# Imagine a situation with 3 choice options
+options <- 3
+# and a population of n = 999
+population <- 999
+# and we observe samples of 3
+sample <- 3
+```
 
@@
 ```{r}
-# Imagine a situation with 3 choice options
-options <- 3
-# generate a uniform competence distribution for a population of n = 10'000
-population <- 10000
+# 1: Distribution of competence
 data <- data.frame(competence = runif(population, min = 1/options, max = 1))
@@
 ```
-### Generate choices from competence
 
-### Sample choice
-
-### Iterate this process
-
-## Function for single population
+### Generate a choice for each individual based on their competence.
 
 ```{r}
+# 2: Draw individual answers based on competence levels 
+data$answer <- data$competence %>% 
+  purrr::map_chr(function(x){
+
+    answer_options <- c("correct", paste0("false", 1:(options-1)))
+    probabilities <- c(x, rep((1-x)/(options-1), options-1))
+
+    answer = sample(answer_options, 
+                    size = 1, 
+                    prob = probabilities
+                    )
+  }
+  )
+
+data <- data %>% 
+  mutate(correct = ifelse(str_detect(answer, "correct"), 
+                                        TRUE, FALSE))
+
+answers <- ggplot(data, aes(x = answer, fill = correct)) +
+  geom_bar() +
+  guides(fill = "none") +
+  plot_theme
+
+competence_by_answers <- ggplot(data, 
+       aes(x = answer, y = relative_competence, fill = correct)) +
+geom_half_violin(position = position_nudge(x = -.2),
+                     adjust=2, alpha = .8,
+                     side = "r") +
+  coord_flip() +
+  guides(fill = "none") +
+  plot_theme
+
+answers + competence_by_answers
+```
+
+### Draw samples, assign categories & compare average outcomes
+
+### Randomly draw samples from the population
+
+```{r}
+# 3: randomly assign samples in population
+data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))
+```
+
+### Categorize constellations.
+
+```{r}
+# 4: identify constellations 
+data <- data %>% 
+  # identify how often a one type of answer occurs in one group
+  group_by(sample_id, answer) %>% 
+  mutate(n_answer_in_sample = n()) %>% 
+  # assign constellations
+  group_by(sample_id) %>% 
+  mutate(unique_answers = n_distinct(answer),
+         # Build a first version of constellation variable
+         constellation = case_when(unique_answers == sample ~ "dissensus", 
+                                   unique_answers == 1 ~ "consensus",
+                                   unique_answers <  sample ~ "majority"
+         ),
+         # identify minority answers in majority constellations
+         minority = ifelse(constellation == "majority", 
+                           # report whether occurences of answer within a group
+                           # are the minority within that group
+                           n_answer_in_sample == min(n_answer_in_sample),
+                           # for all other constellations, simply code NA,
+                           NA
+         ), 
+         # identify majority answers in majority constellations
+         majority = ifelse(constellation == "majority", 
+                           # report whether occurrences of answer within a group
+                           # are the majority within that group
+                           n_answer_in_sample == max(n_answer_in_sample),
+                           # for all other constellations, simply code NA,
+                           NA
+         ), 
+         # modify the constellation variable to distinguish between minority, 
+         # intermediate majority, majority and dissensus that is composed of 
+         # multiple answers (e.g. 3 options, sample of 6, two per option)
+         constellation = case_when(
+           is.na(minority) | is.na(majority) ~ constellation, 
+           minority == TRUE & majority == TRUE ~ "dissensus",
+           minority == TRUE ~ "minority", 
+           majority == TRUE ~ "majority", 
+           .default = "intermediate majority")
+  ) %>% ungroup()
+```
+
+### Compute average accuracy/competence by constellation.
+
+```{r}
+# 5: calculate accuracy and competence levels per constellation
+
+# identify accurate responses
+data <- data %>%
+  mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))
+
+# compute the summary statistics by constellation
+results <- data %>% 
+  group_by(constellation) %>% 
+  summarize(average_competence = mean(competence), 
+            average_relative_competence = mean(relative_competence),
+            average_accuracy = mean(accurate)) %>% 
+  # store simulation information
+  mutate(population = population, 
+         sample = sample, 
+         options = options)
+```
+
+```{r}
+ggplot(results,
+       aes(x = constellation, y = average_accuracy, fill = constellation)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .8,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Accuracy") +
+  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
+  guides(fill = "none") +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
+  ylim(c(0, 1))
+```
+
+## Functions
+
+### Data generating functions
+
+#### Single population
+
+```{r, echo=FALSE}
 simulate_single_population <- function(population, sample, options, 
-                                       distribution = "uniform") {
+                                       distribution = "uniform", 
+                                       return = "summarized") {
   # Check if population is not divisible by sample without remainder
@@
 
+  if(return == "summarized") {
     return(results)
+  } else {
+    return(data)
+  }
 }
@@
 
-```{r}
+```{r, echo=FALSE}
 # # test output (set return to data above)
-# test <- simulate_single_population(population = 600, sample = 6, options = 3) %>% arrange(sample_id)
+# test <- simulate_single_population(population = 600, sample = 6, options = 3, 
+#                                    return = "raw data") %>% arrange(sample_id)
 # levels(as.factor(test$constellation))
@@
 
-## Function for various populations
-```{r}
+#### Various populations
+
+```{r, echo=FALSE}
 simulate_various_populations <- function(iterations,...) {
@@
 
-## Function for varying choice options and sample size function
+#### Vary choice options and sample size
 
@@
 
-```{r}
+```{r, echo=FALSE}
 # set name
@@
 }
-
-# # You can test the function using the commented code below
-# # (un-comment by highlighting, then command+shit+c)
-# test <- do.call(power_by_sample_size,
-#                 c(parameters,
-#                   list(file_name = file_name, sample_sizes = c(5, 10),
-#                        iterations = 100)
-#                   )
-#                 )
 ```
 
-## Plot functions
+### Plot functions
 
-### a) Plot various populations
+#### a) Plot various populations
 
 First, we want a function that plots results obtained by the `simulate_various_populations` function. 
-```{r}
+```{r, echo=FALSE}
 plot_results <- function(data, outcome = "everything") {
@@
   plot_competence <- ggplot(d,
-                            aes(x = constellation, y = average_competence, fill = constellation)) +
+                            aes(x = constellation, y =  average_relative_competence, fill = constellation)) +
     geom_half_violin(position = position_nudge(x = -.2),
@@
 
-### b) Plot varying sample options
+#### b) Plot varying sample & options
 
@@
 
-```{r}
+```{r, echo=FALSE}
 plot_results_vary <- function(data, variable = options) {
@@
     geom_half_violin(position = position_nudge(x = -.2),
-                     adjust=2, alpha = .8,
+                     adjust=2, alpha = .5,
                      side = "l") +
@@
 
-## Simulate
+## Simulation
 
-```{r, message=FALSE}
-n <- c(3, 10, 50, 100, 200)
+### Fixed sample and choice options
 
-vary_sample_options(n = n, vary = "options", iterations = 100, population = 999, sample = 3, file_name = "model_3_options.csv")
-vary_sample_options(n = n, vary = "sample", iterations = 100, population = 999, options = 3, file_name = "model_3_samples.csv")
+#### Generate data
 
-data_3_options <- read_csv("data/model_3_options.csv")
-data_3_sample <- read_csv("data/model_3_samples.csv")
+```{r}
+file_name <- "model_3_sample_3_options.csv" # change for new analyses / or delete file to re-use same name
 
-plot_results_vary(data_3_options)
-plot_results_vary(data_3_sample, variable = sample)
+if (!file.exists(paste0("data/", file_name))) {
+
+  # generate data
+  data <- simulate_various_populations(iterations = 1000, population = 999, sample = 3, options = 3)
+
+  write_csv(data, paste0("data/", file_name))
+}
+
+# read simulated data from .csv files
+data_3 <- read_csv(paste0("data/", file_name))
+
+# descriptive
+data_3 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
+```
+
+#### Plot
+
+```{r}
+plot_results(data_3)
+plot_accuracy_model <- plot_results(data_3, outcome = "accuracy")
+plot_competence_model <- plot_results(data_3, outcome = "competence")
+```
+
+#### Analyze
+
+```{r, echo=FALSE}
+# possible analysis
+regression_data <- data_3 %>%
+  mutate(
+    # make numeric version of convergence
+    convergence = case_when( 
+      constellation == "minority" ~ 0, 
+      constellation == "dissensus" ~ 1,
+      constellation == "majority" ~ 2,
+      constellation == "consensus" ~ 3), 
+    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
+    average_accuracy = average_accuracy*100,
+    # scale competence to reach from 1 to 7 (instead of 0 to 1)
+    average_relative_competence = 6*average_relative_competence + 1 
+  )
+
+accuracy_model <- lm(average_accuracy ~ convergence, 
+                     data = regression_data)
+competence_model <- lm(average_relative_competence ~ convergence, 
+                       data = regression_data)
 
@@
 
-```{r}
+### Varying sample and choice options
+
+### Generate data
+
+```{r, message=FALSE}
+n <- c(3, 5, 10, 20, 50, 100)
+
+# run simulation and store results in .csv files
+vary_sample_options(n = n, vary = "options", iterations = 1000, population = 999, sample = 3, file_name = "model_vary_options_3_sample.csv")
+vary_sample_options(n = n, vary = "sample", iterations = 1000, population = 999, options = 3, file_name = "model_vary_sample_3_options.csv")
+
+# read simulated data from .csv files
+data_3_options <- read_csv("data/model_vary_options_3_sample.csv")
+data_3_sample <- read_csv("data/model_vary_sample_3_options.csv")
+
+```
+
+### Plot
+
+```{r, echo=FALSE}
+# plot results
+plot_results_vary(data_3_options)
+plot_results_vary(data_3_sample, variable = sample)
+```
+
+### Analyze
+
+```{r, echo=FALSE}
+# possible analysis
 regression_data <- data_3_options %>% 
@@
       constellation == "consensus" ~ 3), 
-    options = as.factor(options)
+    options = as.factor(options), 
+    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
+    average_accuracy = average_accuracy*100,
+    # scale competence to reach from 1 to 7 (instead of 0 to 1)
+    average_relative_competence = 6*average_relative_competence + 1
   )
 
-accuracy <- lm(average_accuracy ~ convergence + options, 
-               data = regression_data) #%>% broom.mixed::tidy()
-competence <- lm(average_relative_competence ~ convergence + options, data = regression_data) %>% broom.mixed::tidy()
+accuracy <- lm(average_accuracy ~ convergence + options + convergence*options, 
+               data = regression_data) 
+competence <- lm(average_relative_competence ~ convergence + options, data = regression_data)
 
-summary(accuracy)
+summary(competence)
 ```
 
-```{r}
-data_3_options %>% group_by(sample, constellation) %>% summarise(across(c(average_accuracy, average_relative_competence), mean))
+## Compare model vs. participants
 
-# make descriptive data
-data_3_sample %>% 
-  group_by(sample, constellation) %>% 
-  summarise(count = sum(count)) %>% 
-  mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
-  print(n = 100)
-```
-
-## Simulate
-
-Note that in the current state, we cannot vary `sample`.
-The function needs modification first.
-The problem is that it gets complicated with more than three informants, in the case of a majority constellation.
-There might be equal majorities (e.g.
-5 informants, three options: 2 agree on 'A', 2 on 'B', one picks 'A'), equal minorities (e.g.
-3 agree on 'A', one picks 'B', another picks 'C'), and unequal majorities (e.g.
-3 agree on 'A', 2 agree on 'B').
-
-I have to think about it again.
-
-```{r}
-data_3 <- simulate_various_populations(iterations = 100, population = 999, sample = 3, options = 3)
-```
-
-```{r}
-data_3 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
-```
-
-```{r}
-plot_results(data_3)
-model_accuracy <- plot_results(data_3, outcome = "accuracy")
-model_competence <- plot_results(data_3, outcome = "competence")
-```
-
-```{r}
-# # example to check essential parts of the function
-# my_strings <- c("apple", "banana", "apple", "banana", "apple")
-# 
-# bla <- tibble(id = rep(1:5, each = 5), 
-#        string = sample(c("apple", "banana"), 25, replace = TRUE, 
-#                        prob = c(0.8, 0.2)))
-# 
-# bla %>% group_by(id, string) %>% 
-#   mutate(n_answer_in_group = n()) %>% 
-#   group_by(id) %>% 
-#   mutate(
-#     unique_answers = n_distinct(string),
-#     minority = ifelse(id == 4, 
-#                       n_answer_in_group == min(n_answer_in_group), 
-#                       NA
-#                       )
-#     ) %>% arrange(id) 
-```
-
-## Combine with plots from paper
-
-### Make plots from data
-```{r, message = FALSE}
-# read data
+```{r, message=FALSE}
+# read data from experiments
 d <- read_csv("./data/cleaned.csv")
@@
 
-```{r}
+### Table
+
+```{r, echo=FALSE}
+# Calculate models for participants
+
+# model for accuracy
+# random intercept and slope by participants
+accuracy_participants <- lmer(accuracy ~ convergence + (1 + convergence | id), 
+                                 data = d)
+# model for competence
+# random intercept and slope by participants
+competence_participants <- lmer(competence ~ convergence + 
+                           (1 + convergence | id), data = d)
+```
+
+```{r, echo=FALSE}
+# main result table
+modelsummary::modelsummary(list("Accuracy" = accuracy_participants, 
+                                "Competence" = competence_participants,
+                                "Accuracy" = accuracy_model, 
+                                "Competence" = competence_model
+                                ),
+                           title = 'Participants vs. Model', 
+                           stars = TRUE) %>%
+  kableExtra::add_header_above(c(" " = 1, "Participants" = 2, "Model" = 2))
+```
+
+### Plots
+
+#### Participant data
+
+```{r, echo=FALSE}
 # plot for accuracy
@@
 
-```{r}
+```{r, echo=FALSE}
 # plot for competence
@@
 
-### combine plots from data and from model
+#### Combine participant and model
 
-```{r}
+```{r, echo=FALSE}
 # accuracy
 plot_accuracy <- plot_accuracy + ggtitle("Participants")
-model_accuracy<- model_accuracy + ggtitle("Model")
+plot_accuracy_model <- plot_accuracy_model + ggtitle("Model")
 
-plot_accuracy + model_accuracy
+plot_accuracy + plot_accuracy_model
 ```
 
-```{r}
+```{r, echo=FALSE}
 # competence
 plot_competence <- plot_competence + ggtitle("Participants")
-model_competence<- model_competence + ggtitle("Model")
+plot_competence_model <- plot_competence_model + ggtitle("Model")
 
-plot_competence + model_competence
+plot_competence + plot_competence_model
 ```
-
-## Have only one population.
-
-```{r}
-simulate_single_population <- function(population, sample, options) {
-
-  # randomly draw competence levels
-  data <- tibble(id = 1:population,
-                 competence = runif(population, min = 1/options, max = 1))
-
-  # randomly draw individual answers based on competence levels 
-  data$answer <- data$competence %>% 
-    purrr::map_chr(function(x){
-
-      answer_options <- c("correct", paste0("false", 1:(options-1)))
-      probabilities <- c(x, rep((1-x)/(options-1), options-1))
-
-      answer = sample(answer_options, 
-                      size = 1, 
-                      prob = probabilities
-      )
-      return(answer)
-
-    })
-
-  # randomly assign samples in population
-  data <- data %>% mutate(sample_id = rep(1:(population/sample), sample))
-
-  # identify constellations 
-  data <- data %>% 
-    # identify how often a one type of answer occurs in one group
-    group_by(sample_id, answer) %>% 
-  mutate(n_answer_in_sample = n()) %>% 
-    # assign constellations
-  group_by(sample_id) %>% 
-    mutate(unique_answers = n_distinct(answer),
-           # Build a first version of constellation variable
-           constellation = case_when(unique_answers == sample ~ "dissensus", 
-                                     unique_answers == 1 ~ "consensus",
-                                     unique_answers <  sample ~ "majority"
-           ),
-           # identify minority answers in majority constellations
-           minority = ifelse(constellation == "majority", 
-                             # report whether occurences of answer within a group
-                             # are the minority within that group
-                             n_answer_in_sample == min(n_answer_in_sample),
-                             # for all other constellations, simply code NA,
-                             NA
-           ), 
-           # identify majority answers in majority constellations
-           majority = ifelse(constellation == "majority", 
-                             # report whether occurrences of answer within a group
-                             # are the minority within that group
-                             n_answer_in_sample == max(n_answer_in_sample),
-                             # for all other constellations, simply code NA,
-                             NA
-           ), 
-           # modify the constellation variable to distinguish between minority, 
-           # intermediate majority, majority and dissensus that is composed of 
-           # multiple answers (e.g. 3 options, sample of 6, two per option)
-           constellation = case_when(
-             is.na(minority) | is.na(majority) ~ constellation, 
-             minority == TRUE & majority == TRUE ~ "dissensus",
-             minority == TRUE ~ "minority", 
-             majority == TRUE ~ "majority", 
-             .default = "intermediate majority")
-           ) %>% ungroup()
-
-  # identify accurate responses
-  data <- data %>%
-    mutate(accurate = ifelse(answer == "correct", TRUE, FALSE))
-
-  # compute the summary statistics by constellation
-  results <- data %>% 
-    group_by(sample_id, constellation) %>% 
-    summarize(average_competence = mean(competence), 
-              average_accuracy = mean(accurate)) %>% 
-    # store simulation information
-    mutate(population = population, 
-           sample = sample, 
-           options = options) %>%
-
-  return(results)
-}
-
-```
-
-```{r}
-plot_results <- function(data) {
-
-  d <- data
-  # make constellation a factor with the right levels
-  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")
-
-  # make count data
-  counts <- d %>% group_by(constellation) %>% count()
-
-  # plot for accuracy
-  plot_accuracy <- ggplot(d,
-                          aes(x = constellation, y = average_accuracy, fill = constellation)) +
-    geom_half_violin(position = position_nudge(x = -.2),
-                     adjust=0.1, alpha = .8,
-                     side = "l") +
-    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-    # Add nice labels
-    labs(x = "Convergence", y = "Accuracy") +
-    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
-    scale_color_viridis_d(option = "plasma", begin = 0.7) +
-    guides(fill = FALSE, color = FALSE) +
-    plot_theme + 
-    theme(axis.text.x = element_text(angle = 20, hjust = 1)) +
-    geom_text(data = counts, aes(x = constellation, y = 1.05, 
-                                 label =  paste0("n =", n)
-                                 )
-              )
-
-  # plot for competence
-  plot_competence <- ggplot(d,
-                            aes(x = constellation, y = average_competence, fill = constellation)) +
-    geom_half_violin(position = position_nudge(x = -.2),
-                     adjust=1, alpha = .8,
-                     side = "l") +
-    stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
-    stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-    stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-    # Add nice labels
-    labs(x = "Convergence", y = "Competence") +
-    scale_fill_viridis_d(option = "plasma", begin = 0.1) +
-    scale_color_viridis_d(option = "plasma", begin = 0.7) +
-    guides(fill = FALSE) +
-    plot_theme + 
-    theme(axis.text.x = element_text(angle = 20, hjust = 1))
-
-  (plot_accuracy | plot_competence)
-}
-```

### 2023-06-05 Experiment_6/power_Exp6.Rmd
update experiment 6 before data collection

@@
 
-We set the parameters for convergence and the by-participant random effects based on experiment 4.
+We set the parameters for the by-participant random effects and the error term based on the estimates from experiment 4.
 
-We further use the modeling results to produce the expected difference between the categories.
+```{r, message=FALSE}
+# read data
+d <- read_csv("../experiment_4/data/cleaned.csv")
 
-```{r}
+# get random effects for accuracy
+
+# calculate model
+model_accuracy <- lmer(accuracy ~ convergence + (1 + convergence | id), 
+                                 data = d)
+
+# store estimates with according parameters in list
+random_effects_accuracy <- tidy(model_accuracy, effects = "ran_pars") %>% 
+  mutate(parameter = case_when(term == "sd__(Intercept)" ~ "subj_0", 
+                               term == "cor__(Intercept).convergence" ~ "subj_rho",
+                               term == "sd__convergence" ~ "subj_c", 
+                               term == "sd__Observation" ~ "sigma")
+         ) %>% 
+  select(parameter, estimate)
+
+# get random effects for competence
+
+# calculate model
+model_competence <- lmer(competence ~ convergence + 
+                           (1 + convergence | id), data = d)
+
+# store estimates with according parameters in list
+random_effects_competence <- tidy(model_competence, effects = "ran_pars") %>% 
+  mutate(parameter = case_when(term == "sd__(Intercept)" ~ "subj_0", 
+                               term == "cor__(Intercept).convergence" ~ "subj_rho",
+                               term == "sd__convergence" ~ "subj_c", 
+                               term == "sd__Observation" ~ "sigma")
+         ) %>% 
+  select(parameter, estimate)
+
+```
+
+We further use the modeling results to set the parameters for all fixed effects.
+
+```{r, message=FALSE}
 # data from model
-model_data <- read_csv("../experiment_4/data/model_3_options.csv") %>% 
+model_data <- read_csv("../experiment_4/data/model_vary_options_3_sample.csv") %>% 
   filter(options %in% c(3, 10)) %>% 
@@
 # we run the model on averages from different iterations
-model_accuracy <- lm(average_accuracy ~ convergence + options + convergence*options, data = model_data) %>% 
-  tidy()
 
-model_competence <- lm(average_relative_competence ~ convergence + options + convergence*options, data = model_data) %>% 
-  tidy()
+# for accuracy
+model_accuracy <- lm(average_accuracy ~ convergence + options + convergence*options, data = model_data)
+
+# store estimates with according parameters in list
+fixed_effects_accuracy <- tidy(model_accuracy) %>% 
+  mutate(parameter = case_when(term == "(Intercept)" ~ "beta_0", 
+                               term == "convergence" ~ "beta_c",
+                               term == "options10" ~ "beta_n",
+                               term == "convergence:options10" ~ "beta_cn")
+         ) %>% 
+  select(parameter, estimate)
+
+# for competence
+model_competence <- lm(average_relative_competence ~ convergence + options + convergence*options, data = model_data)
+
+# store estimates with according parameters in list
+fixed_effects_competence <- tidy(model_competence) %>% 
+  mutate(parameter = case_when(term == "(Intercept)" ~ "beta_0", 
+                               term == "convergence" ~ "beta_c",
+                               term == "options10" ~ "beta_n",
+                               term == "convergence:options10" ~ "beta_cn")
+         ) %>% 
+  select(parameter, estimate)
 ```
@@
 # parameters for modelling accuracy 
-beta_0   <- 20 # intercept
-beta_c   <- 10 # effect of convergence when conflict of interest
-beta_i   <- 5 # effect of independence (compared to conflict of interest)
-beta_ci  <- 5 # difference in effect of convergence when number of options = 10 (compared to conflict of interest)
-subj_0   <- 10 # by-subject random intercept sd
-subj_c   <- 5 # by-subject random slope sd
-subj_rho <- .2 # correlation between intercept and slope 
-sigma    <- 5 # residual (error) sd (i.e. all varition that the model cannot account for)
+parameters_accuracy <- bind_rows(random_effects_accuracy, fixed_effects_accuracy)
+# split() creates a list using the values in a dataframe column
+parameters_accuracy <- split(parameters_accuracy$estimate, parameters_accuracy$parameter)
+
+# parameters for modelling competence 
+parameters_competence <- bind_rows(random_effects_competence, fixed_effects_competence)
+parameters_competence <- split(parameters_competence$estimate, parameters_competence$parameter)
+
+parameters_accuracy
+parameters_competence
 ```
 
+To these effect parameters, we have to add a set of additional parameters.
 We store a complete list of parameters in a list so that we can call them for functions later.
+Note that here we'll only set parameters for accuracy, for demonstration purposes.
+We'll do the same for competence later when running the actual simulations.
 
 ```{r}
+# store parameters in separate objects to be able to call them for demonstration purposes
+# (for accuracy only)
+beta_0   <- parameters_accuracy$beta_0 # intercept
+beta_c   <- parameters_accuracy$beta_c # effect of convergence when 3 choice options
+beta_n   <- parameters_accuracy$beta_n # effect of 10 options (vs. 3) when convergence = 0
+beta_cn  <- parameters_accuracy$beta_cn # difference in effect of convergence when 10 options (vs. 3)
+subj_0   <- parameters_accuracy$subj_0 # by-subject random intercept sd
+subj_c   <- parameters_accuracy$subj_c # by-subject random slope sd
+subj_rho <- parameters_accuracy$subj_rho # correlation between intercept and slope 
+sigma    <- parameters_accuracy$sigma # residual (error) sd (i.e. all varition that the model cannot account for)
+
 # complete list of parameters (including some that will be introduced later)
-parameters <- list(
+parameters <- c(
+  parameters_accuracy,
+  list(
   # outcome variable
@@
   # parameters for size of data frame
-  n_subj = 10,              # n participants
+  n_subj = 100,              # n participants
   levels_convergence = 0:3,  # levels of convergence
-  n_stimuli_per_level = 2, # n stimuli per participant
-  # parameters for modelling outcome variable
-  beta_0 = 40,   # intercept
-  beta_c = 10,   # effect of convergence, 
-  beta_i = 5,    # # effect of independence (compared to conflict of interest)
-  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
-  subj_0 = 10,   # by-subject random intercept sd
-  subj_c = 5,   # by-subject random slope sd
-  subj_rho = 0.5, # correlation between intercept and slope 
-  sigma = 10     # residual (error) sd (i.e. all varition that the model cannot account for)
+  n_stimuli_per_level = 2 # n stimuli per participant
+  )
   )
@@
   # add `independence` condition
-  independence = rep(c("conflict", "independent"), length.out = n_subj)
+  number_options = rep(c("3", "10"), length.out = n_subj)
 ) %>% 
   # make a numeric variable to be able to generate competence
-  mutate(independence_num = ifelse(independence == "conflict", 0, 1))
+  mutate(number_options_num = ifelse(number_options == "3", 0, 1), 
+         # relevel the number_options variable
+         number_options = relevel(as.factor(number_options), ref = "3"))
 ```
@@
 
-#### Calculate accuracy values
+### Calculate accuracy values
 
@@
 dat_sim <- trials %>%
-  mutate(accuracy = beta_0 + SU_0 + (beta_c + SU_c)*convergence + beta_i*independence_num + beta_ci*convergence*independence_num + e_ss,
+  mutate(accuracy = beta_0 + SU_0 + (beta_c + SU_c)*convergence + beta_n*number_options_num + beta_cn*convergence*number_options_num + e_ss,
          # truncate accuracy values so that they lie between 0 and 100 only
@@
   # retain only variables of interest
-  select(id, convergence, independence, accuracy)
+  select(id, convergence, number_options, accuracy)
 ```
@@
 # independent condition
-independent_plot <- ggplot(dat_sim %>% filter(independence == "independent"), 
+options_10_plot <- ggplot(dat_sim %>% filter(number_options == "10"), 
        aes(x = as.factor(convergence), y = accuracy, 
@@
   geom_hline(yintercept = beta_0) +
-  geom_hline(yintercept = (beta_0 + beta_c + beta_ci)) +
-  geom_hline(yintercept = (beta_0 + 2*(beta_c + beta_ci))) +
-  geom_hline(yintercept = (beta_0 + 3*(beta_c + beta_ci))) +
+  geom_hline(yintercept = (beta_0 + beta_c + beta_cn)) +
+  geom_hline(yintercept = (beta_0 + 2*(beta_c + beta_cn))) +
+  geom_hline(yintercept = (beta_0 + 3*(beta_c + beta_cn))) +
   # actual data
@@
   ggtitle("Predicted versus simulated values") + 
-  facet_wrap(~independence) +
-  ylim(0, 100)
+  facet_wrap(~number_options) +
+  ylim(0, 120)
 
 # conflict condition
-conflict_plot <-ggplot(dat_sim %>% filter(independence == "conflict"), 
+options_3_plot <-ggplot(dat_sim %>% filter(number_options == "3"), 
        aes(x = as.factor(convergence), y = accuracy, 
@@
   ggtitle("Predicted versus simulated values")  + 
-  facet_wrap(~independence) +
-  ylim(0, 100)
+  facet_wrap(~number_options) +
+  ylim(0, 120)
 
-independent_plot + conflict_plot
+options_10_plot + options_3_plot
 ```
@@
   beta_c,   # effect of convergence 
-  beta_i = 5,    # # effect of independence (compared to conflict of interest)
-  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
+  beta_n,  # effect of independence (compared to conflict of interest)
+  beta_cn,  # difference in effect of convergence when independence (compared to conflict of interest)
   subj_0,   # by-subject random intercept sd
@@
     # add `independence` condition
-    independence = rep(c("conflict", "independent"), length.out = n_subj)
+    number_options = rep(c("3", "10"), length.out = n_subj)
     ) %>% 
     # make a numeric variable to be able to generate competence
-    mutate(independence_num = ifelse(independence == "conflict", 0, 1))
+    mutate(number_options_num = ifelse(number_options == "3", 0, 1), 
+           # relevel the number_options variable
+           number_options = relevel(as.factor(number_options), ref = "3")
+           )
 
@@
     mutate(outcome = beta_0 + SU_0 + (beta_c + SU_c)*convergence + 
-             beta_i*independence_num + beta_ci*convergence*independence_num + e_ss, 
+             beta_n*number_options_num + beta_cn*convergence*number_options_num + e_ss, 
            outcome_variable = {{outcome}}) %>% 
     # retain only variables of interest
-    select(id, convergence, independence, outcome, outcome_variable)
+    select(id, convergence, number_options, outcome, outcome_variable)
 
@@
 ```{r}
-mod_sim <- lmer(accuracy ~ convergence + independence + 
-                  independence*convergence + (1 + convergence | id), 
+mod_sim <- lmer(accuracy ~ convergence + number_options + 
+                  number_options*convergence + (1 + convergence | id), 
                 dat_sim)
@@
   mutate(
-    parameter = c("beta_0", "beta_c", "beta_i", "beta_ci", "subj_0", "subj_rho", "subj_c", "sigma"),
-    value = c(beta_0, beta_c, beta_i, beta_ci, subj_0, subj_rho, subj_c, sigma),
+    parameter = c("beta_0", "beta_c", "beta_n", "beta_cn", "subj_0", "subj_rho", "subj_c", "sigma"),
+    value = c(beta_0, beta_c, beta_n, beta_cn, subj_0, subj_rho, subj_c, sigma),
   ) %>%
@@
 # estimates to extract
-estimates = c("convergence", "independenceindependent", 
-              "convergence:independenceindependent")
+estimates = c("convergence", "number_options10", 
+              "convergence:number_options10")
 
@@
   # estimate model for the drawn sample
-  model_estimates <- lmer(outcome ~ convergence + independence + 
-                            independence*convergence + (1 + convergence | id),
+  model_estimates <- lmer(outcome ~ convergence + number_options + 
+                            number_options*convergence + (1 + convergence | id),
                           sample) %>% 
@@
   # estimates to extract
-  estimates = c("convergence", "independenceindependent", 
-                "convergence:independenceindependent")
+  estimates = c("convergence", "number_options10", 
+              "convergence:number_options10")
 
@@
 # list of parameters for accuracy
-parameters_accuracy <- list(
+parameters_accuracy <- c(parameters_accuracy, 
+                         list(
                            # outcome variable
@@
                            n_stimuli_per_level = 2, # n stimuli per participant
-  # parameters for modelling outcome variable
-  beta_0 = 40,   # intercept
-  beta_c = 5,   # effect of convergence 
-  beta_i = 5,    # effect of independence (compared to conflict of interest)
-  beta_ci = 5, # difference in effect of convergence when independence (compared to conflict of interest)
-  subj_0 = 10,   # by-subject random intercept sd
-  subj_c = 5,   # by-subject random slope sd
-  subj_rho = -0.5, # correlation between intercept and slope 
-  sigma = 10,     # residual (error) sd (i.e. all varition that the model cannot account for)
                            # number of iterations (per sample size)
                            iterations = 1000
+                           )
                          )
@@
 # list of parameters for competence
-parameters_competence <- list(
+parameters_competence <- c(parameters_competence, 
+                         list(
                            # outcome variable
-  outcome = "competence", # takes names as input, e.g. "accuracy"
+                           outcome = "competence", # takes names as input, e.g. "competence"
                            # parameters for size of data frame
@@
                            n_stimuli_per_level = 2, # n stimuli per participant
-  # parameters for modelling outcome variable
-  beta_0 = 3.5,   # intercept
-  beta_c = 0.25,   # effect of convergence 
-  beta_i = 0.25,    # # effect of independence (compared to conflict of interest)
-  beta_ci = 0.5, # difference in effect of convergence when independence (compared to conflict of interest)
-  subj_0 = 1,   # by-subject random intercept sd
-  subj_c = 0.25,   # by-subject random slope sd
-  subj_rho = -0.5, # correlation between intercept and slope 
-  sigma = 1,     # residual (error) sd (i.e. all varition that the model cannot account for)
                            # number of iterations (per sample size)
                            iterations = 1000
+                           )
                          )
@@
 
+### Calculate precise sample size threshold
+
+```{r}
+# accuracy
+power_accuracy %>% filter(term == "convergence:number_options10" & 
+                            power >= 0.9) %>%
+  arrange(power) %>% 
+  slice(1)
+```
+
+```{r}
+# competence
+power_competence %>% filter(term == "convergence:number_options10") %>%
+  arrange(desc(power)) 
+```
+
 ## References

### 2023-06-05 Experiment_6/preregistration.Rmd
update experiment 6 before data collection

@@
 output:
+  pdf_document: 
+      keep_md: yes
   html_document:
-    keep_md: yes
-  pdf_document: 
     keep_md: yes
@@
   competence = sample(1:7, replace = TRUE, size=100),
-  options = rep(c("3", "10"), each = 50),
+  condition = sample(c("3", "10", "10_distant"), replace = TRUE, 
+                   prob = c(0.5, 0.25, 0.25), size = 100),
   id = rep(1:5, length.out = 100)
+  ) %>% 
+  mutate(stimuli_10_version = ifelse(str_detect(condition, "3"), NA, 
+                                     condition), 
+         options = ifelse(str_detect(condition, "10"), "10", 
+                          condition)
                                      )
@@
 
-**Options**. Number of options has two levels: (i) '3' and (ii) '10'.
+**Number of choice options**. Number of options has two levels: '3' '10'.
 
-We manipulate independence between participants, i.e.
+We manipulate number of choice options between participants, i.e.
 each participant gets assigned to either the '3' or the '10' condition.
 All stimuli that participants see in their respective condition will involve the same number of choice options.
+
+As for the '10' options condition, participants will see one of two distinct set of stimuli (`stimuli_10_version`): one in which the range of the answers corresponds to the range of the '3' options condition (`stimuli_10_version == 10`), and another with increased range (`stimuli_10_version == 10_distant`; see Appendix).
+We add the increased range condition because we think that there is the possibility that participant might not consider all options as relevant when they only see scenarios in which all answers cluster.
+For the main analyses, we will simply merge these two stimuli sets and treat them as part of the same '10' options condition.
+However, we will investigate potential differences as a research question.
 
@@
 
-Considering only `3` options condition of the experiment at hand, we expect the findings of experiment four to replicate. In experiment four, players would choose among `3` options.
+Considering only `3` options condition of the experiment at hand, we expect the findings of experiment four to replicate, for it is the exact same experiment as experiment four in this case.
 
@@
 
-### H1b:  In the three options condition, participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants, given three choice options.
+### H1b:  In the three options condition, participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.
 
@@
 the fictive players) could choose not among 3, but among 10 options?
-In line with our findings from experiment three, we predict that effects of convergence are more positive when there are more choice options:
+Following the results from our model, we predict that effects of convergence are more positive when there are more choice options:
 
@@
 
+## Research question
+
+### RQ1: Within the 10 choice options condition, is the effect of convergence more positive for the set of stimuli with greater distance?
+
+```{r, warning=FALSE, message=FALSE}
+# models for accuracy
+
+# random intercept and slope by participants
+model_accuracy <- lmer(accuracy ~ convergence + stimuli_10_version + 
+                            stimuli_10_version*convergence + (1 + convergence | id), 
+                       data = data)
+
+# in case of non-convergence: random intercept by participants only
+alt_model_accuracy <- lmer(accuracy ~ convergence + stimuli_10_version + 
+                            stimuli_10_version*convergence + (1 | id), 
+                           data = data)
+
+# models for competence
+
+# random intercept and slope by participants
+model_competence <- lmer(competence ~ convergence + stimuli_10_version + 
+                            stimuli_10_version*convergence + (1 + convergence | id), 
+                       data = data)
+
+# in case of non-convergence: random intercept by participants only
+alt_model_competence <- lmer(competence ~ convergence + stimuli_10_version + 
+                            stimuli_10_version*convergence + (1 | id), 
+                           data = data)
+```
+
+Should we find the difference, we will re-run the analyses of the second set of hypotheses restraining the data to the subset of the more distant set of stimuli for the 10 options condition.
+
+Pick an example stimulus as earlier for illustrating choice options. See all stimuli in appendix.
+
 # Robustness checks
@@
 
-## Alternative stimuli with greater distance
-
-We kept the same distance between 10 and 3.
-That's conservative - people might think that other options are not relevant and hence don't make a difference between 3 and 10 options.
-We therefore run the hypotheses set two comparing 3 to a version of 10 with larger distance.
-
-Pick an example stimulus as earlier for illustrating choice options. See all stimuli in appendix.
-
 # Exclusions
@@
 We ran a power simulation to inform our choice of sample size.
-All assumptions and details on the procedure can be found in the `power_Exp5.Rmd` document.
-We ran two different power analyses, one for each outcome variable.
-We set the power threshold for our experiment to 90%.
+All assumptions and details on the procedure can be found in the `power_Exp6.Rmd` document.
+We used previous experiments and estimates of our models to inform our choice of parameter values.
 
-The power simulation for `accuracy` suggested that for 80 participants, we would have a power of at least 90% for the interaction effect.
-The simulation for `competence` suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we also detect an effect of convergence.
+We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%.
 
-However, due to uncertainty about our assumptions and because we are anticipating failed attention checks, *we will recruit a sample of `200` participants*.
+The power simulation for `accuracy` suggested that for 140 participants we would cross the power threshold of 90% for the interaction effect (power = 0.928).
+The simulation for `competence` suggested that with 300 participants, we would detect an interaction with a power of 87% (power = 0.872).
+
+Due to budget constraints, we will consider a sample of 300 participants as good enough in terms of power.
 
 ```{r accuracy-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
-#knitr::include_graphics("power_Exp5_files/figure-html/plot-power-accuracy-1.png")
+knitr::include_graphics("power_Exp6_files/figure-html/plot-power-accuracy-1.png")
 ```
@@
 ```{r competence-image, echo=FALSE, fig.cap="Results of power simulation for competence"}
-#knitr::include_graphics("power_Exp5_files/figure-html/plot-power-competence-1.png")
+knitr::include_graphics("power_Exp6_files/figure-html/plot-power-competence-1.png")
 ```

### 2023-06-06 Experiment_6/analysis.Rmd
analysis experiment 6

@@
 # read data
-d <- read_csv("./data/cleaned.csv")
+d <- read_csv("./data/cleaned.csv") %>% 
+  # make number_options a factor
+  mutate(number_options = relevel(as.factor(number_options), ref = "three"))
 ```
 
-### H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.
+### H1a: In the three options condition, participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.
 
-To test this hypothesis, we only consider participants assigned to the `independent` condition.
+To test this hypothesis, we only consider participants assigned to the `3` options condition.
 
-We use a linear mixed effect model with random intercept and random slope per participant.
+We use a linear mixed effect model with random intercept and random slope per participant. Should this model yield convergence issues, we will use a model with random intercept only.
 
-In all our models we treat `convergence` as a continuous variable. We will, however, include robustness checks where we treat convergence as a categorical variable.
+In all our models we treat `convergence` as a continuous variable.
+We will, however, include robustness checks where we treat convergence as a categorical variable, allowing to inspect difference between different levels.
 
@@
 # random intercept and slope by participants
-model_accuracy_independent <- lmer(accuracy ~ convergence + (1 + convergence | id), 
-                       data = d %>% filter(independence == "independent"))
+model_accuracy_3options <- lmer(accuracy ~ convergence + (1 + convergence | id), 
+                       data = d %>% filter(number_options == "three"))
+
 ```
 
-### H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.
+### H1b:  In the three options condition, participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.
 
@@
 # random intercept and slope by participants
-model_competence_independent <- lmer(competence ~ convergence + 
+model_competence_3options <- lmer(competence ~ convergence + 
                            (1 + convergence | id), 
-                         data = d %>% filter(independence == "independent"))
+                         data = d %>% filter(number_options == "three"))
 ```
 
-### H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are biased (i.e.
-share a conflict of interest to pick a given answer).
+### H2a: The effect of convergence on accuracy (H1a) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.
 
-To test this hypothesis, we only consider the full data.
+To test this hypothesis, we consider the full data.
 
@@
 # random intercept and slope by participants
-model_accuracy <- lmer(accuracy ~ convergence + independence + 
-                            independence*convergence + (1 + convergence | id), 
+model_accuracy <- lmer(accuracy ~ convergence + number_options + 
+                            number_options*convergence + (1 + convergence | id), 
                        data = d)
@@
 
-### H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are biased (i.e.
-share a conflict of interest to pick a given answer).
+### H2b: The effect of convergence on competence (H1b) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.
 
-To test this hypothesis, we only consider the full data.
+To test this hypothesis, we consider the full data.
 
@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 # models for competence
@@
 # random intercept and slope by participants
-model_competence <- lmer(competence ~ convergence + independence + 
-                            independence*convergence + (1 + convergence | id), 
+model_competence <- lmer(competence ~ convergence + number_options + 
+                            number_options*convergence + (1 + convergence | id), 
                        data = d)
-
 ```
@@
 ```{r, echo=FALSE}
-models <- list("Accuracy (independent only)" = model_accuracy_independent, 
-               "Competence (independent only)" = model_competence_independent, 
+models <- list("Accuracy (3 options condition)" = model_accuracy_3options, 
+               "Competence (3 options condition)" = model_competence_3options, 
                "Accuracy" = model_accuracy, 
@@
 
-# Robustness checks
+## Research question
 
-In the models above, we treated convergence as a continuous variable. Based on the different levels, we will build a categorical variable, `convergence_categorical`.
+### RQ1: Within the 10 choice options condition, is the effect of convergence more positive for the set of stimuli with greater distance?
+
+```{r, warning=FALSE, message=FALSE}
+# models for accuracy
+
+# random intercept and slope by participants
+model_accuracy_10options <- lmer(accuracy ~ convergence + stimuli_10_version + 
+                            stimuli_10_version*convergence + (1 + convergence | id), 
+                       data = d)
+
+# models for competence
+
+# random intercept and slope by participants
+model_competence_10options <- lmer(competence ~ convergence + stimuli_10_version + 
+                            stimuli_10_version*convergence + (1 + convergence | id), 
+                       data = d)
+
+```
+
+```{r}
+models <- list("Accuracy (10 options conditions)" = model_accuracy_10options, 
+               "Competence (10 options conditions)" = model_competence_10options)
+modelsummary::modelsummary(models, stars = TRUE)
+```
+
+### Plots
 
@@
 levels(d$convergence_categorical)
-
 ```
 
-We run the same models outlined in the hypotheses section, but replacing `convergence` with `convergence_categorical`.
-This also allows us to inspect heterogeniety in differences between levels (with respect to the baseline, i.e.
-"opposing majority").
-
-### Plots
-
-#### Independence Condition only
+#### Three options only
 
@@
 # plot for accuracy
-plot_accuracy <- ggplot(d %>% filter(independence == "independent"),
+plot_accuracy <- ggplot(d %>% filter(number_options == "three"),
        aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
@@
 # plot for competence
-plot_competence <- ggplot(d %>% filter(independence == "independent"),
+plot_competence <- ggplot(d %>% filter(number_options == "three"),
        aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
@@
 plot_accuracy <- ggplot(d,
-       aes(x = convergence_categorical, y = accuracy, fill = independence)) +
+       aes(x = convergence_categorical, y = accuracy, fill = number_options)) +
   geom_half_violin(position = position_nudge(x = -.2),
@@
 plot_competence <- ggplot(d,
-       aes(x = convergence_categorical, y = competence, fill = independence)) +
+       aes(x = convergence_categorical, y = competence, fill = number_options)) +
   geom_half_violin(position = position_nudge(x = -.2),

### 2023-11-08 Experiment_1/cleaning.Rmd
Remove cached *_files directories

@@
 d <- d %>%
-  mutate(convergence = as.factor(convergence),
-         convergence = fct_relevel(convergence, "noconv", "conv"),
+  mutate(
+    convergence = ifelse(convergence == "conv", "convergent", "divergent"),
+    convergence = as.factor(convergence),
+    convergence = fct_relevel(convergence, "divergent", "convergent"),
     number = as.factor(number),

### 2023-11-08 Experiment_1/model.Rmd
Remove cached *_files directories

@@
 library(broom.mixed) # for tidy outputs of linear mixed models
+library(ggpubr) # for plots and ggarrange()
 ```
@@
 ```{r, echo = FALSE}
-#set general theme for plots
+# set general theme for plots
 plot_theme <- theme_minimal(base_size = 12) +
-  theme(# Bold, bigger title
-        plot.title = element_text(face = "bold", size = rel(1.7)),
-        # Plain, slightly bigger subtitle that is grey
-        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
+  theme(# Bold title
+    plot.title = element_text(face = "bold", size = rel(0.7)),
+    # Plain subtitle that is grey
+    plot.subtitle = element_text(face = "plain", size = rel(1), color = "grey70"),
     # Bold legend titles
@@
     # Bold axis titles
-        axis.title = element_text(face = "bold"))
+    axis.title = element_text(face = "bold"), 
+    # Bold legend labels
+    legend.text = element_text(face = "bold")
+    )
 ```
@@
 sample <- 3
+# alpha and beta values for beta distributions modelling competence
+alpha <- 1
+beta <- 1
+# ramge of values for estimates
+min_range = 1000
+max_range = 2000
+
+# set arbitrary limit for competence distribution
+# (factor determining distance between lowest and highest possible competence, 
+# i.e. standard deviation)
+competence_ratio = 1000 
 ```
@@
 
-We suppose that these competence levels are drawn from a random uniform distribution. That is, in a population, each level of competence is equally likely to occur.
+We assume that individuals (i) vary in competence and (ii) that the distribution of competence in the population varies, as can be modeled by a beta distribution.
+In the case below, we assume a uniform competence distribution (equal to a beta distribution of alpha = 1 and beta = 1).
 
@@
 # 1: Distribution of competence
-data <- data.frame(competence = runif(population, 
-                                      min = range/1000, 
-                                      max = range)) %>% 
-  # Scale competence to reach from 0 to 1 (instead of 1 to 1000).
-  # Reverse the scale such that 1 becomes maximal competence (instead of 0).
-  mutate(competence_scaled = 1 - ((competence - 1) / (1000-1)))
+
+# Theoretical population distribution of competence
+ggplot() +
+  geom_function(fun = ~dbeta(.x, shape1 = 1, shape2 = 1))
+
+# randomly draw competence levels
+data <- tibble(id = 1:population,
+               competence = rbeta(population, shape1 = alpha, 
+                                          shape2 = beta)) %>% 
+    # For now competence tending towards one is more competent.
+    # For the data generating, we need the reverse and scale this competence measure.  
+    mutate(
+      # Since competence = standard deviation of estimate generating normal distribution, 
+      # reverse the scale such that 0 corresponds to maximal competence (instead of 1).
+      competence_reversed = 1 - competence, 
+      # Put values on a scale: 
+      # We (arbitrarily) set the lowest competence to the range of possible values, 
+      # in case of our experiments 2000 - 1000 = 1000. 
+      # We set the highest competence to 1/1000 of the range of possible values, 
+      # in our case 1.
+      competence_reversed_scaled = (competence_reversed * range) + range/competence_ratio)
 
 # generated population
-ggplot(data, aes(x = competence_scaled)) +
+ggplot(data, aes(x = competence)) +
   geom_histogram() + 
@@
 
-The numbers that we generate from the normal distributions are tuncated such that they all lie within the pre-defined range (1000 to 2000)
+The numbers that we generate from the normal distributions are truncated such that they all lie within the pre-defined range (1000 to 2000)
 
@@
 
-data$answer <- data$competence %>% 
+# use the reversed competence variable to generate the estimates
+data$answer <- data$competence_reversed_scaled %>% 
   purrr::map_dbl(function(x){
@@
 
-    answer = truncnorm::rtruncnorm(1, mean = mean, sd = sd, a = 1000, b = 2000)
+    answer = truncnorm::rtruncnorm(1, mean = mean, sd = sd, a = min_range, b = max_range)
   }
@@
 
-We measure accuracy as the (absolute value of) distance between the chosen answer and the true answer for each individual.
+We measure accuracy as the (squared) distance between the chosen answer and the true answer for each individual.
 
@@
 data <- data %>% 
-  mutate(accuracy = -1 * abs(mean - answer))
+  mutate(accuracy = -1 * (mean - answer)^2)
 
@@
 
-ggplot(data, aes(x = competence_scaled, y = accuracy)) +
+ggplot(data, aes(x = competence, y = accuracy)) +
   geom_point()
@@
 
-We define convergence as the standard deviation of the answers within a sample (multiplied by -1 so that higher values correspond to greater convegence)
+We define convergence as the standard deviation of the answers within a sample (multiplied by -1 so that higher values correspond to greater convergence)
 
@@
   group_by(sample_id, convergence) %>% 
-  summarize(competence_mean = mean(competence_scaled),
+  summarize(competence_mean = mean(competence),
             accuracy_mean = mean(accuracy)) %>% 
@@
     labs(x = "Convergence \n(SDs within samples of three)", 
-         y = "Accuracy \n(average distance to true mean)") +
+         y = "Accuracy \n(squared distance to true mean)") +
   plot_theme
@@
     labs(x = "Convergence \n(SDs within samples of three)", 
-         y = "Competence \n(SD for data generating function, transposed)") +
+         y = "Competence \n(SD of data generating function)") +
   plot_theme
@@
 simulate_single_population <- function(population, sample, min_range, max_range, 
-                                       competence_ratio = 1000) {
+                                       competence_ratio = 1000, alpha, beta) {
 
@@
   # 1: Distribution of competence
-  data <- data.frame(competence = runif(population, 
-                                        min = range/competence_ratio, 
-                                        max = range)) %>% 
-    # Scale competence to reach from 0 to 1.
-    # Reverse the scale such that 1 becomes maximal competence (instead of 0).
-    mutate(competence_scaled = 1 - ((competence - range/competence_ratio) / 
-                                      (range - range/competence_ratio)))
+  data <- tibble(id = 1:population,
+                 competence = rbeta(population, shape1 = alpha, 
+                                    shape2 = beta)) %>% 
+    # For now competence tending towards one is more competent.
+    # For the data generating, we need the reverse and scale this competence measure.  
+    mutate(
+      # Since competence = standard deviation of estimate generating normal distribution, 
+      # reverse the scale such that 0 corresponds to maximal competence (instead of 1).
+      competence_reversed = 1 - competence, 
+      # Put values on a scale: 
+      # We (arbitrarily) set the lowest competence to the range of possible values, 
+      # in case of our experiments 2000 - 1000 = 1000. 
+      # We set the highest competence to 1/1000 of the range of possible values, 
+      # in our case 1.
+      competence_reversed_scaled = (competence_reversed * range) + range/competence_ratio)
 
@@
 
-  data$answer <- data$competence %>% 
+  data$answer <- data$competence_reversed_scaled %>% 
     purrr::map_dbl(function(x){
@@
   data <- data %>% 
-    mutate(accuracy = -1 * abs(mean - answer))
+    mutate(accuracy = -1 * (mean - answer)^2)
 
@@
     group_by(sample_id, convergence) %>% 
-    summarize(competence_mean = mean(competence_scaled),
+    summarize(competence_mean = mean(competence),
               accuracy_mean = mean(accuracy)) %>% 
@@
 }
-
 ```
@@
 ```{r, echo=FALSE}
-# # # test output (set return to data above)
 # test <- simulate_single_population(population = 500, sample = 3,
 #                                    min_range = 1, max_range = 6,
-#                                    competence_ratio = 1000)
+#                                    competence_ratio = 1000,
+#                                    alpha = 1,
+#                                    beta = 1)
 ```
@@
 # function for different sample sizes
-vary_sample <- function(file_name, n, ...) {
+vary_sample <- function(file_name, n, format = "result", ...) {
 
@@
 
+    if(format == "result") {
+      return(data)
+    } else{
       write_csv(data, paste0("data/", file_name))
+    }
   }
@@
 
+#### Vary competence
+
+We now want to do the above (simulate a single population, repeat this process, do so for various sample sizes and choice options) for different competence distributions.
+We vary these distributions with the `alpha` and `beta` parameters for the beta distributions.
+
+```{r, echo=FALSE}
+# set name
+file_name <- "model.csv" # change for new analyses / or delete file to re-use same name
+
+vary_competence <- function(file_name, alpha, beta, names, ...) {
+
+  # only run analysis if a file with that name does not yet exists
+  if (!file.exists(paste0("data/", file_name))) {
+
+    competence_data <- tibble(alpha, beta, names) %>% 
+      mutate(competence = paste0(names, "\n(alpha = ", alpha, ", beta = ", beta, ")"))
+
+    # do all the above functions for different competence distributions
+    data <- competence_data$competence %>% purrr::map_df(function(competence_level){
+      # this is essentially a for loop
+
+      # To keep track of progress
+      print(paste("competence_level = ", competence_level))
+
+      # set alpha and beta
+      alpha = competence_data %>% filter(competence == competence_level) %>% pull(alpha)
+      beta = competence_data %>% filter(competence == competence_level) %>% pull(beta)
+
+      # run simulation
+      result <- vary_sample(
+        alpha = alpha, beta = beta, file_name = file_name, ...)
+
+      # add competence level
+      result <- result %>% mutate(competence = competence_level, 
+                                  alpha = alpha,
+                                  beta = beta)
+
+      return(result)
+
+    })
+
+    write_csv(data, paste0("data/", file_name))
+  }
+}
+```
+
 ### Plot functions
@@
   # extract simulation info
-  simulation_info <- d %>% 
-    summarize(across(c(population, sample, total_iterations), max)) %>% 
-    pivot_longer(everything(), names_to = "parameter", values_to = "value") %>% 
-    gridExtra::tableGrob(cols = NULL, rows = NULL)
+  caption = paste0("Sample size of informants: ", d$sample[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
 
@@
     # Add nice labels
-    labs(x = "Convergence \n(SDs within samples of three)", 
-         y = "Accuracy \n(average distance to true mean)") +
+    labs(x = "Convergence \n(SDs of samples)", 
+         y = "Accuracy \n(squared distance to true mean)", 
+         caption = paste0(caption, "; assuming uniform competence distribution")) +
     plot_theme +
@@
     # Add nice labels
-    labs(x = "Convergence \n(SDs within samples of three)", 
-         y = "Competence \n(SD for data generating function, transposed)") +
+    labs(x = "Convergence \n(SDs of samples)", 
+         y = "Competence \n(SD for data generating function)",
+         caption = paste0(caption, "; assuming uniform competence distribution")) +
     plot_theme +
@@
 
-    return((plot_accuracy | plot_competence) / simulation_info)
+    return(plot_accuracy | plot_competence)
   }
@@
 ```{r, echo=FALSE}
-plot_results_vary <- function(data, outcome = "everything", sample) {
+plot_results_vary <- function(data, ...) {
 
@@
 
-    sample_plot <- plot_results(data = plot_data)
+    sample_plot <- plot_results(data = plot_data, ...)
 
@@
 
+#### c) Plot competence distributions
+
+We want to plot the underlying competence distributions.
+
+```{r}
+# Plot competence distributions
+plot_competence_distributions <- function(data) {
+  # Your data frame
+  d <- data %>% 
+    group_by(competence) %>% 
+    reframe(alpha = unique(alpha), beta = unique(beta))
+
+  # Function to plot the different competence distributions 
+  plot_competence_distributions <- function(alpha, beta) {
+    ggplot() +
+      geom_function(
+        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
+      ) +
+      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
+           y = "Density", x = "Competence") +
+      plot_theme
+  }
+
+  # Specify the values of alpha and beta from the data
+  alpha_values <- d$alpha
+  beta_values <- d$beta
+
+  # Create a list of ggplot objects
+  plot_list <- map2(alpha_values, beta_values, plot_competence_distributions)
+
+  # Arrange and display the plots
+  competence_plot <- ggarrange(plotlist = plot_list) %>%
+    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
+                                                    face = "bold",
+                                                    color = "darkgrey",
+                                                    size = 14))
+
+  return(competence_plot) 
+}
+```
+
+#### d) Plot varying competence distributions by sample
+
+Fourth, we want another function that plots results by various competence levels by sample sizes.
+
+```{r}
+plot_competence <- function(data, outcome = "accuracy") {
+
+  d <- data
+
+  # extract simulation info
+  caption = paste0("Sample size of informants: ", d$sample[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
+
+  plot_accuracy <- ggplot(d, aes(x = convergence, y = accuracy_mean)) +
+    geom_hex(alpha = 0.9) +
+    #geom_smooth(method = "lm", se = FALSE) +
+    stat_smooth(geom='line', method = "lm", alpha=0.7, se=FALSE, color = "darkgrey") +
+    # Add nice labels
+    labs(x = "Convergence \n(SDs of samples)", 
+         y = "Accuracy \n(squared distance to true mean)",
+         caption = caption) +
+  scale_fill_viridis_c(option = "plasma", begin = 0.1) +
+    facet_wrap(~competence) +
+    plot_theme + 
+    theme(legend.position = "top", 
+          legend.key.width = unit(1.5, "cm"), 
+          legend.key.height = unit(0.3, "cm"))
+
+  plot_competence <- ggplot(d, aes(x = convergence, y = competence_mean)) +
+    geom_hex(alpha = 0.9) +
+    #geom_smooth(method = "lm", se = FALSE) +
+    stat_smooth(geom='line', method = "lm", alpha=0.7, se=FALSE, color = "darkgrey") +
+    # Add nice labels
+    labs(x = "Convergence \n(SDs of samples)", 
+         y = "Competence \n(SD for data generating function)",
+         caption = caption) +
+  scale_fill_viridis_c(option = "plasma", begin = 0.1) +
+    facet_wrap(~competence) +
+    plot_theme + 
+    theme(legend.position = "top", 
+          legend.key.width = unit(1.5, "cm"), 
+          legend.key.height = unit(0.3, "cm"))
+
+  if (outcome == "accuracy") {
+
+    return(plot_accuracy)
+  }
+
+  if (outcome == "competence") {
+
+    return(plot_competence)
+  }
+}
+
+plot_competence_vary <- function(data, ...) {
+
+  # get different samples as vector
+  samples <- data %>% 
+    reframe(unique(as.character(sample))) %>% pull()
+
+  # empty list
+  plot_list <- list()
+
+  # make plots fro each moderator
+  for (i in samples) {
+
+    plot_data <- data %>%
+      filter(sample == i)
+
+    sample_plot <- plot_competence(data = plot_data, ...)
+
+    plot_list[[i]] <- sample_plot
+  }
+
+  return(plot_list)
+}
+
+plot_competence_vary_relative_majority <- function(data, variable = options, outcome = "Accuracy") {
+
+  d <- data
+
+  # Extract variable name as a string
+  variable_name <- deparse(substitute(variable))
+
+  # retrieve info from choice of variable
+  if (variable_name == "options") {
+    caption = paste0("Sample size of informants: ", d$sample[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
+  }
+  if (variable_name == "sample") {
+    caption = paste0("Number of choice options: ", d$options[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
+  }
+
+  d <- d %>% 
+    group_by(relative_majority, {{variable}}, competence) %>%
+    summarise(competence_mean = mean(average_relative_competence),
+              accuracy_mean = mean(average_accuracy),
+              accuracy_sd = sd(average_accuracy),
+              competence_sd = sd(average_relative_competence)
+    )
+
+  # Assign outcome variable
+  if(outcome == "Accuracy") {
+    d$outcome <- d$accuracy_mean
+    d$outcome_sd <- d$accuracy_sd
+  }
+
+  if(outcome == "Competence") {
+    d$outcome <- d$competence_mean
+    d$outcome_sd <- d$competence_sd
+  }
+
+  plot_outcome <- ggplot(d, 
+                         aes(x = relative_majority, y = outcome,
+                             color = as.factor({{variable}}))) +
+    geom_point() +
+    geom_line() +
+    # Add nice labels
+    labs(x = "Relative majority \n (share of informants agreeing on an option)", 
+         y = outcome, caption = caption, 
+         color = variable_name) +
+    scale_color_viridis_d(option = "plasma", begin = 0.1, 
+                          #limits = rev(levels(d$constellation)),
+                          direction = -1
+    ) +
+    facet_wrap(~competence) +
+    plot_theme + 
+    theme(legend.position = "top")
+
+   return(plot_outcome)
+
+} 
+```
+
 ## Simulation
 
-### Generate data
+### Experiment 4
+
+#### Generate Data
 
@@
             iterations = 100,
-            file_name = "model_vary_3_10_20_50.csv",
-            n = n)
+            file_name = "sim_experiment_1.csv",
+            n = n, 
+            alpha = 1, 
+            beta = 1, 
+            format = ".csv")
 
 # read simulated data from .csv files
-data <- read_csv("data/model_vary_3_10_20_50.csv")
+data <- read_csv("data/sim_experiment_1.csv")
 ```
@@
 # plot results
-plot_results_vary(data)
+plot_results_vary(data, outcome = "accuracy")
+plot_results_vary(data, outcome = "competence")
 ```
@@
 Below, what I did was predict accuracy and competence for convergence levels of 20 and 150 based on a regression run on the model data.
-We could compare that to the participant data - but it's not quite the same thing...
+We could compare that to the participant data - but it's not quite the same thing..
 
@@
 ```
+
+## Vary competence
+
+```{r, message=FALSE}
+n <- c(3, 5, 10, 20, 50, 100)
+alpha = c(0.1, 2, 1, 1, 6, 0.9)
+beta = c(1, 6, 1, 0.1, 2, 0.9)
+names = c("very incompetent", "incompetent", "uniform", 
+          "very competent", "competent", "bimodal")
+
+# run simulation and store results in .csv files
+vary_competence(population = 999,
+            min_range = 1000, max_range = 2000,
+            competence_ratio = 1000,
+            iterations = 100,
+            file_name = "sim.csv",
+            n = n, 
+            alpha = alpha, 
+            beta = beta, 
+            names = names)
+
+# read simulated data from .csv files
+test <- read_csv("data/sim.csv")
+```
+
+```{r}
+plot_competence_distributions(test)
+```
+```{r}
+# for accuracy
+plot_competence_vary(data = test, outcome = "accuracy")
+```
+```{r}
+# for competence
+plot_competence_vary(data = test, outcome = "competence")
+```

### 2023-11-08 Experiment_3/cleaning.Rmd
Remove cached *_files directories

@@
 Likewise, we recode the levels of the `independence` variable to make sure the effect goes into the right direction (i.e.
-the effec of _independence_ compared to _conflict of interest_).
+the effect of _independence_ compared to _conflict of interest_).
 
@@
 d <- d %>%
-  mutate(convergence = as.factor(convergence),
-         convergence = fct_relevel(convergence, "div", "conv"),
+  mutate(
+    convergence = ifelse(convergence == "conv", "convergent", "divergent"),
+    convergence = as.factor(convergence),
+    convergence = fct_relevel(convergence, "divergent", "convergent"),
+    independence = ifelse(independence == "indep", "independent", "conflict"),
     independence = as.factor(independence),
-         independence = fct_relevel(independence, "confl", "indep")
+    independence = fct_relevel(independence, "conflict", "independent")
          )

### 2023-11-08 Experiment_4/model.Rmd
Remove cached *_files directories

@@
 library(broom.mixed) # for tidy outputs of linear mixed models
+library(ggpubr) # for plots
 ```
@@
 ```{r, echo=FALSE}
-#set general theme for plots
+# set general theme for plots
 plot_theme <- theme_minimal(base_size = 12) +
-  theme(# Bold, bigger title
-        plot.title = element_text(face = "bold", size = rel(1.7)),
-        # Plain, slightly bigger subtitle that is grey
-        plot.subtitle = element_text(face = "plain", size = rel(1.3), color = "grey70"),
+  theme(# Bold title
+    plot.title = element_text(face = "bold", size = rel(0.7)),
+    # Plain subtitle that is grey
+    plot.subtitle = element_text(face = "plain", size = rel(1), color = "grey70"),
     # Bold legend titles
@@
 
-We assume that individuals (i) vary in competence and (ii) that competence that it is distributed uniformly in the population.
+We assume that individuals (i) vary in competence and (ii) that the distribution of competence in the population varies, as can be modeled by a beta distribution.
+In the case below, we assume a uniform competence distribution (equal to a beta distribution of alpha = 1 and beta = 1).
 
@@
          # Build a first version of constellation variable
-         constellation = case_when(unique_answers == sample ~ "dissensus", 
-                                   unique_answers == 1 ~ "consensus",
-                                   unique_answers <  sample ~ "majority"
-         ),
-         # identify minority answers in majority constellations
-         minority = ifelse(constellation == "majority", 
-                           # report whether occurences of answer within a group
-                           # are the minority within that group
-                           n_answer_in_sample == min(n_answer_in_sample),
-                           # for all other constellations, simply code NA,
-                           NA
+         constellation = case_when(unique_answers == 1 ~ "consensus",
+                                   unique_answers <=  sample ~ "majority-minority-dissensus"
          ),
          # identify majority answers in majority constellations
-         majority = ifelse(constellation == "majority", 
+         majority = ifelse(constellation == "majority-minority-dissensus", 
                            # report whether occurrences of answer within a group
@@
          ), 
-         # modify the constellation variable to distinguish between minority, 
-         # intermediate majority, majority and dissensus that is composed of 
-         # multiple answers (e.g. 3 options, sample of 6, two per option)
+         # identify dissensus answers in majority constellations
+         dissensus = ifelse(constellation == "majority-minority-dissensus", 
+                            # report whether occurrences of answer within a group
+                            # are the minority within that group
+                            min(n_answer_in_sample) == max(n_answer_in_sample),
+                            # for all other constellations, simply code NA,
+                            NA
+         ), 
+         # build final constellation variable using helper variables 
          constellation = case_when(
-           is.na(minority) | is.na(majority) ~ constellation, 
-           minority == TRUE & majority == TRUE ~ "dissensus",
-           minority == TRUE ~ "minority", 
+           dissensus == TRUE ~ "dissensus",
            majority == TRUE ~ "majority",
-           .default = "intermediate majority")
+           majority == FALSE ~ "minority(ies)",
+           .default = constellation)
   ) %>% ungroup()
@@
 
-# compute the summary statistics by constellation
+# compute the summary statistics by sample id
 results <- data %>% 
-  group_by(constellation) %>% 
+  group_by(sample_id, constellation) %>% 
   summarize(average_competence = mean(competence), 
@@
                                        distribution = "uniform", 
-                                       return = "summarized") {
+                                       alpha, beta,
+                                       return = "by constellation") {
   # Check if population is not divisible by sample without remainder
@@
     data <- tibble(id = 1:population,
-                 competence = runif(population, min = 1/options, max = 1))
-  }
-
-  if (distribution == "normal") {
-  # randomly draw competence levels
-  data <- tibble(id = 1:population,
-                 competence = rnorm(population, mean = 0.6, sd = 0.01))
-  }
-
+                   competence = runif(population, min = 1/options, max = 1)) %>% 
       # Create a relative measure of competence that allows to compare across
@@
       # We use min-max scaling, with min = 1/options and max = 1.
-  data <- data %>% 
-    mutate(relative_competence = (competence - 1/options) / (1 - 1/options)
+      mutate(relative_competence = (competence - 1/options) / (1 - 1/options))
+  }
+
+  if (distribution == "beta") {
+  # randomly draw competence levels
+  data <- tibble(id = 1:population,
+                 relative_competence = rbeta(population, shape1 = alpha, 
+                                             shape2 = beta)) %>% 
+    mutate(
+      competence = relative_competence * (1 - 1/options) + 1/options
       )
+  }
 
@@
     mutate(unique_answers = n_distinct(answer),
+           # relative degree of majority
+           n_sample = n_distinct(id),
+           relative_majority = n_answer_in_sample / n_sample,
            # Build a first version of constellation variable
-           constellation = case_when(unique_answers == sample ~ "dissensus", 
-                                     unique_answers == 1 ~ "consensus",
-                                     unique_answers <  sample ~ "majority"
-           ),
-           # identify minority answers in majority constellations
-           minority = ifelse(constellation == "majority", 
-                             # report whether occurences of answer within a group
-                             # are the minority within that group
-                             n_answer_in_sample == min(n_answer_in_sample),
-                             # for all other constellations, simply code NA,
-                             NA
+           constellation = case_when(unique_answers == 1 ~ "consensus",
+                                     unique_answers <=  sample ~ "majority-minority-dissensus"
            ),
            # identify majority answers in majority constellations
-           majority = ifelse(constellation == "majority", 
+           majority = ifelse(constellation == "majority-minority-dissensus", 
                              # report whether occurrences of answer within a group
@@
            ), 
-           # modify the constellation variable to distinguish between minority, 
-           # intermediate majority, majority and dissensus that is composed of 
-           # multiple answers (e.g. 3 options, sample of 6, two per option)
+           # identify dissensus answers in majority constellations
+           dissensus = ifelse(constellation == "majority-minority-dissensus", 
+                             # report whether occurrences of answer within a group
+                             # are the minority within that group
+                             min(n_answer_in_sample) == max(n_answer_in_sample),
+                             # for all other constellations, simply code NA,
+                             NA
+           ), 
+           # build final constellation variable using helper variables 
            constellation = case_when(
-             is.na(minority) | is.na(majority) ~ constellation, 
-             minority == TRUE & majority == TRUE ~ "dissensus",
-             minority == TRUE ~ "minority", 
+             dissensus == TRUE ~ "dissensus",
              majority == TRUE ~ "majority",
-             .default = "intermediate majority")
+             majority == FALSE ~ "minority(ies)",
+             .default = constellation)
            ) %>% ungroup()
@@
 
-  # compute the summary statistics by constellation
+  if(return == "by participant") {
+    return(data)
+  } else {
+
+    if(return == "by degree of majority") {
       results <- data %>% 
-    group_by(constellation) %>% 
-    summarize(average_competence = mean(competence), 
+        group_by(sample_id, relative_majority)
+    }
+
+    if(return == "by constellation") {
+      results <- data %>% 
+        group_by(sample_id, constellation)
+    }
+
+    # compute the summary statistics by sample id (and constellation)
+    results <- results %>% summarize(average_competence = mean(competence), 
                                      average_relative_competence = mean(relative_competence),
@@
 
-  if(return == "summarized") {
     return(results)
-  } else {
-    return(data)
   }
 }
-
 ```
@@
 ```{r, echo=FALSE}
-# # test output (set return to data above)
-# test <- simulate_single_population(population = 600, sample = 6, options = 3, 
-#                                    return = "raw data") %>% arrange(sample_id)
-# levels(as.factor(test$constellation))
+# test output (set return to data above)
+# test <- simulate_single_population(population = 10000, sample = 10, options = 10,
+#                                    return = "by degree of majority", distribution = "beta",
+#                                    alpha = 2, beta = 6) %>% arrange(sample_id) %>%
+#   arrange(sample_id)
 # 
-# test %>% filter(constellation == "intermediate majority")
+# test %>% group_by(relative_majority) %>% summarise(across(c(average_relative_competence,
+#                                                         average_accuracy),
+#                                                       mean)
+#                                                )
 ```
@@
 
-vary_sample_options <- function(vary = "options", file_name, n, ...) {
+vary_sample_options <- function(vary = "options", file_name, n, 
+                                format = "result", ...) {
 
@@
 
-      # do the `calculate_power()` function for each sample size and store the results
-      # in a new variable called `power`
+      # run the populations generating function for different values of n
       data <- n %>% purrr::map_df(function(n_option){
@@
         # do the following for each 
-                                     # element data$n_subj
+        # element of n
 
@@
 
-                                     # run power calculation
+        # run simulation
         result <- simulate_various_populations(
@@
 
+      if(format == "result") {
+        return(data)
+      } else{
         write_csv(data, paste0("data/", file_name))
+      }
     }
@@
 
+      if(format == "result") {
+        return(data)
+      } else{
         write_csv(data, paste0("data/", file_name))
       }
+    }
+  }
+}
+```
+
+#### Vary competence distribution
+
+We now want to do the above (simulate a single population, repeat this process, do so for various sample sizes and choice options) for different competence distributions.
+We vary these distributions with the `alpha` and `beta` parameters for the beta distributions.
+
+```{r, echo=FALSE}
+# set name
+file_name <- "model.csv" # change for new analyses / or delete file to re-use same name
+
+vary_competence <- function(file_name, alpha, beta, names, ...) {
+
+  # only run analysis if a file with that name does not yet exists
+  if (!file.exists(paste0("data/", file_name))) {
+
+    competence_data <- tibble(alpha, beta, names) %>% 
+      mutate(competence = paste0(names, "\n(alpha = ", alpha, ", beta = ", beta, ")"))
+
+    # do all the above functions for different competence distributions
+    data <- competence_data$competence %>% purrr::map_df(function(competence_level){
+      # this is essentially a for loop
+
+      # To keep track of progress
+      print(paste("competence_level = ", competence_level))
+
+      # set alpha and beta
+      alpha = competence_data %>% filter(competence == competence_level) %>% pull(alpha)
+      beta = competence_data %>% filter(competence == competence_level) %>% pull(beta)
+
+      # run simulation
+      result <- vary_sample_options(
+        alpha = alpha, beta = beta, file_name = file_name, ...)
+
+      # add competence level
+      result <- result %>% mutate(competence = competence_level, 
+                                  alpha = alpha,
+                                  beta = beta)
+
+      return(result)
+
+    })
+
+    write_csv(data, paste0("data/", file_name))
   }
@@
   # make constellation a factor with the right levels
-  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")
+  d$constellation <- fct_relevel(d$constellation, "minority(ies)", "dissensus", "majority", "consensus")
 
   # extract simulation info
-  simulation_info <- d %>% summarize(across(c(population, sample, options, total_iterations), mean)) %>% 
-    pivot_longer(everything(), names_to = "parameter", values_to = "value") %>% 
-    gridExtra::tableGrob(cols = NULL, rows = NULL)
+  simulation_info <- d %>% summarize(across(c(population, sample, options, total_iterations), mean))
+
+  caption = paste0("Number of choice options: ", simulation_info$options,
+                   "; Sample size: ", simulation_info$sample,
+                   "; Iterations : ", simulation_info$total_iterations, 
+                   "; Population size per iteration: ", simulation_info$population)
 
   # make descriptive data
-  descriptive <- d %>% 
-    group_by(constellation) %>% 
-    summarise(count = sum(count)) %>% 
-    mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
-    gridExtra::tableGrob(rows = NULL)
+  # descriptive <- d %>% 
+  #   group_by(constellation) %>% 
+  #   summarise(count = sum(count)) %>% 
+  #   mutate(rel_freq = round(count / sum(count), digits = 2)) %>% 
+  #   gridExtra::tableGrob(rows = NULL)
 
@@
     # Add nice labels
-    labs(x = "Convergence", y = "Competence") +
+    labs(x = "Convergence", y = "Competence", caption = caption) +
     scale_fill_viridis_d(option = "plasma", begin = 0.1) +
@@
   # unite tables
-  tables <- gridExtra::grid.arrange(descriptive, simulation_info, ncol = 2)
+  # tables <- gridExtra::grid.arrange(descriptive, simulation_info, ncol = 2)
 
@@
 
-    return((plot_accuracy | plot_competence) / tables)
+    return(plot_accuracy | plot_competence )
   }
@@
 
-    return(plot_accuracy)
+    return(plot_accuracy + labs(caption = caption))
   }
@@
   # make constellation a factor with the right levels
-  d$constellation <- fct_relevel(d$constellation, "minority", "dissensus", "intermediate majority", "majority", "consensus")
+  d$constellation <- fct_relevel(d$constellation, "minority(ies)", "dissensus", "majority", "consensus")
 
@@
 
+#### c) Plot varying competence distributions
+
+Third, we want a function that plots results by various competence levels.
+
+```{r}
+plot_competence_vary <- function(data, variable = options, outcome = "Accuracy", 
+                                 return = "simulations") {
+
+  d <- data
+
+  # Extract variable name as a string
+  variable_name <- deparse(substitute(variable))
+
+  # retrieve info from choice of variable
+  if (variable_name == "options") {
+    x_label = "Number of Choice Options"
+    caption = paste0("Sample size of informants: ", d$sample[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
+  }
+  if (variable_name == "sample") {
+    x_label = "Size of informant groups"
+    caption = paste0("Number of choice options: ", d$options[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
+  }
+
+  # make constellation a factor with the right levels
+  d$constellation <- fct_relevel(d$constellation, "minority(ies)", "dissensus", "majority", "consensus")
+
+  d <- d %>% 
+    group_by(constellation, {{variable}}, competence) %>% 
+    summarise(competence_mean = mean(average_relative_competence),
+              accuracy_mean = mean(average_accuracy),
+              accuracy_sd = sd(average_accuracy),
+              competence_sd = sd(average_relative_competence)
+    )
+
+  # Assign outcome variable
+  if(outcome == "Accuracy") {
+    d$outcome <- d$accuracy_mean
+    d$outcome_sd <- d$accuracy_sd
+  }
+
+  if(outcome == "Competence") {
+    d$outcome <- d$competence_mean
+    d$outcome_sd <- d$competence_sd
+  }
+
+  # plot for accuracy
+  plot_outcome <- ggplot(d,
+                         aes(x = {{variable}}, y = outcome,
+                             color = constellation)) +
+    # geom_half_violin(position = position_nudge(x = -.2),
+    #                  adjust=2, alpha = .5,
+    #                  side = "l") +
+    geom_pointrange(aes(ymin = outcome - outcome_sd, 
+                        ymax = outcome + outcome_sd), 
+                    size = 0.1) +
+    geom_line() +
+    # Add nice labels
+    labs(x = x_label, y = outcome, caption = caption) +
+    scale_color_viridis_d(option = "plasma", begin = 0.1, 
+                          limits = rev(levels(d$constellation)),
+                          direction = -1
+    ) +
+    facet_wrap(~competence) +
+    plot_theme + 
+    theme(legend.position = "top")
+
+  # Plot competence distributions
+  # Your data frame
+  d <- data %>% 
+    group_by(competence) %>% 
+    reframe(alpha = unique(alpha), beta = unique(beta))
+
+  # Function to plot the different competence distributions 
+  plot_competence_distributions <- function(alpha, beta) {
+    ggplot() +
+      geom_function(
+        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
+      ) +
+      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
+           y = "Density", x = "Competence") +
+      plot_theme
+  }
+
+  # Specify the values of alpha and beta from the data
+  alpha_values <- d$alpha
+  beta_values <- d$beta
+
+  # Create a list of ggplot objects
+  plot_list <- map2(alpha_values, beta_values, plot_competence_distributions)
+
+  # Arrange and display the plots
+  competence_plot <- ggarrange(plotlist = plot_list) %>%
+    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
+                                                    face = "bold",
+                                                    color = "darkgrey",
+                                                    size = 14))
+
+  if(return == "competence distributions") {
+    return(competence_plot)
+  } else {
+   return(plot_outcome) 
+  }
+} 
+```
+
+#### d) Plot varying competence distributions by relative majority
+
+Fourth, we want another function that plots results by various competence levels - but this time not for  constellations but the numeric value of the relative majority.
+
+```{r}
+plot_competence_vary_relative_majority <- function(data, variable = options, outcome = "Accuracy") {
+
+  d <- data
+
+  # Extract variable name as a string
+  variable_name <- deparse(substitute(variable))
+
+  # retrieve info from choice of variable
+  if (variable_name == "options") {
+    caption = paste0("Sample size of informants: ", d$sample[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
+  }
+  if (variable_name == "sample") {
+    caption = paste0("Number of choice options: ", d$options[1],
+                     "; Iterations per sample size: ", d$total_iterations[1], 
+                     "; Population size per iteration: ", d$population[1])
+  }
+
+  d <- d %>% 
+    group_by(relative_majority, {{variable}}, competence) %>%
+    summarise(competence_mean = mean(average_relative_competence),
+              accuracy_mean = mean(average_accuracy),
+              accuracy_sd = sd(average_accuracy),
+              competence_sd = sd(average_relative_competence)
+    )
+
+  # Assign outcome variable
+  if(outcome == "Accuracy") {
+    d$outcome <- d$accuracy_mean
+    d$outcome_sd <- d$accuracy_sd
+  }
+
+  if(outcome == "Competence") {
+    d$outcome <- d$competence_mean
+    d$outcome_sd <- d$competence_sd
+  }
+
+  plot_outcome <- ggplot(d, 
+                         aes(x = relative_majority, y = outcome,
+                             color = as.factor({{variable}}))) +
+    geom_point() +
+    geom_line() +
+    # Add nice labels
+    labs(x = "Relative majority \n (share of informants agreeing on an option)", 
+         y = outcome, caption = caption, 
+         color = variable_name) +
+    scale_color_viridis_d(option = "plasma", begin = 0.1, 
+                          #limits = rev(levels(d$constellation)),
+                          direction = -1
+    ) +
+    facet_wrap(~competence) +
+    plot_theme + 
+    theme(legend.position = "top")
+
+   return(plot_outcome)
+
+} 
+```
+
+#### e) Plot competence
+
+Finally, we want to plot the underlying competence distributions
+
+```{r}
+# Plot competence distributions
+plot_competence_distributions <- function(data) {
+  # Your data frame
+  d <- data %>% 
+    group_by(competence) %>% 
+    reframe(alpha = unique(alpha), beta = unique(beta))
+
+  # Function to plot the different competence distributions 
+  plot_competence_distributions <- function(alpha, beta) {
+    ggplot() +
+      geom_function(
+        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
+      ) +
+      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
+           y = "Density", x = "Competence") +
+      plot_theme
+  }
+
+  # Specify the values of alpha and beta from the data
+  alpha_values <- d$alpha
+  beta_values <- d$beta
+
+  # Create a list of ggplot objects
+  plot_list <- map2(alpha_values, beta_values, plot_competence_distributions)
+
+  # Arrange and display the plots
+  competence_plot <- ggarrange(plotlist = plot_list) %>%
+    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
+                                                    face = "bold",
+                                                    color = "darkgrey",
+                                                    size = 14))
+
+  return(competence_plot) 
+}
+```
+
 ## Simulation
 
-### Fixed sample and choice options
+```{r}
+# Tryout beta functions
+# ggplot() +
+#   geom_function(fun = ~dbeta(.x, shape1 = 0.9, shape2 = 0.9))
+```
 
-#### Generate data
+### Experiment 4
 
-```{r}
-file_name <- "model_3_sample_3_options.csv" # change for new analyses / or delete file to re-use same name
+#### Generate Data
+
+First, we want to simulate data under assumptions that correspond to our experimental design.
+
+For the first, simulation, we fix all parameters. We assume:
+- a uniform competence distribution (beta = 1 and alpha = 1)
+- 3 choice options,
+- samples of 3 (i.e. 3 informants)
+
+We generate a sample of 999x1000 = 999000 individuals.
+
+```{r, message=FALSE}
+file_name <- "sim_experiment_4.csv" # change for new analyses / or delete file to re-use same name
 
@@
 # read simulated data from .csv files
-data_3 <- read_csv(paste0("data/", file_name))
-
-# descriptive
-data_3 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
+sim_experiment_4 <- read_csv(paste0("data/", file_name))
 ```
 
-#### Plot
+#### Plot result
+
+We make then make a plot corresponding to the figure from the experiments.
 
 ```{r}
-plot_results(data_3)
-plot_accuracy_model <- plot_results(data_3, outcome = "accuracy")
-plot_competence_model <- plot_results(data_3, outcome = "competence")
+# descriptive
+sim_experiment_4 %>% group_by(constellation) %>% summarise(across(c(average_accuracy, average_competence), mean))
+
+plot_results(sim_experiment_4)
 ```
 
-#### Analyze
+#### Analyze data
 
@@
 # possible analysis
-regression_data <- data_3 %>%
+regression_data <- sim_experiment_4 %>%
   mutate(
@@
                        data = regression_data)
-
 ```
 
-### Varying sample and choice options
-
-### Generate data
-
-```{r, message=FALSE}
-n <- c(3, 5, 10, 20, 50, 100)
-
-# run simulation and store results in .csv files
-vary_sample_options(n = n, vary = "options", iterations = 1000, population = 999, sample = 3, file_name = "model_vary_options_3_sample.csv")
-vary_sample_options(n = n, vary = "sample", iterations = 1000, population = 999, options = 3, file_name = "model_vary_sample_3_options.csv")
-
-# read simulated data from .csv files
-data_3_options <- read_csv("data/model_vary_options_3_sample.csv")
-data_3_sample <- read_csv("data/model_vary_sample_3_options.csv")
-
-```
-
-### Plot
-
-```{r, echo=FALSE}
-# plot results
-plot_results_vary(data_3_options)
-plot_results_vary(data_3_sample, variable = sample)
-```
-
-### Analyze
-
-```{r, echo=FALSE}
-# possible analysis
-regression_data <- data_3_options %>% 
-  filter(options %in% c(3, 10)) %>% 
-  group_by(iteration, constellation, options) %>% 
-  summarise(across(c(average_accuracy, average_relative_competence), mean)) %>% 
-  mutate(
-    convergence = case_when( 
-      constellation == "minority" ~ 0, 
-      constellation == "dissensus" ~ 1,
-      constellation == "majority" ~ 2,
-      constellation == "consensus" ~ 3), 
-    options = as.factor(options), 
-    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
-    average_accuracy = average_accuracy*100,
-    # scale competence to reach from 1 to 7 (instead of 0 to 1)
-    average_relative_competence = 6*average_relative_competence + 1
-  )
-
-accuracy <- lm(average_accuracy ~ convergence + options + convergence*options, 
-               data = regression_data) 
-competence <- lm(average_relative_competence ~ convergence + options, data = regression_data)
-
-summary(competence)
-```
-
-## Compare model vs. participants
+#### Compare participants and simulations
 
@@
 
-### Table
+##### Table
 
@@
                            (1 + convergence | id), data = d)
-```
 
-```{r, echo=FALSE}
 # main result table
@@
 
-### Plots
-
-#### Participant data
+##### Plots
 
 ```{r, echo=FALSE}
+# Participant data
+
 # plot for accuracy
@@
   theme(axis.text.x = element_text(angle = 20, hjust = 1))
-```
 
-```{r, echo=FALSE}
 # plot for competence
@@
 
-#### Combine participant and model
+```{r}
+# Model data
+plot_accuracy_model <- plot_results(sim_experiment_4, outcome = "accuracy")
+plot_competence_model <- plot_results(sim_experiment_4, outcome = "competence")
+```
 
 ```{r, echo=FALSE}
+# Combine participant and model
 # accuracy
@@
 ```
+
+### Experiment 6
+
+#### Generate data
+
+In experiment 6, we varied choice options (from 3 to 10).
+
+In this simulation, we assume:
+- a sample size of 3 
+- a uniform competence distribution (beta = 1 and alpha = 1)
+
+```{r, message=FALSE}
+n <- c(3, 5, 10, 20, 50, 100)
+
+# run simulation and store results in .csv files
+vary_sample_options(n = n, format = "csv file", vary = "options", iterations = 5, population = 10000, 
+                    sample = 3, file_name = "sim_vary_options_3_sample.csv", 
+                    distribution = "beta", alpha = 1, beta = 1)
+
+# read simulated data from .csv files
+data_3_options <- read_csv("data/sim_vary_options_3_sample.csv")
+```
+
+#### Plot result
+
+```{r}
+# example for plot
+plot_results_vary(data_3_options, 
+                  variable = options)
+```
+
+#### Analyze data
+
+```{r, echo=FALSE}
+# possible analysis for varying choice options (3 vs. 10)
+regression_data <- data_3_options %>%
+  filter(options %in% c(3, 10)) %>%
+  # group_by(iteration, constellation, options) %>%
+  # summarise(across(c(average_accuracy, average_relative_competence), mean)) %>%
+  mutate(
+    convergence = case_when(
+      constellation == "minority" ~ 0,
+      constellation == "dissensus" ~ 1,
+      constellation == "majority" ~ 2,
+      constellation == "consensus" ~ 3),
+    options = as.factor(options),
+    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
+    average_accuracy = average_accuracy*100,
+    # scale competence to reach from 1 to 7 (instead of 0 to 1)
+    average_relative_competence = 6*average_relative_competence + 1
+  )
+
+accuracy <- lm(average_accuracy ~ convergence + options + convergence*options,
+               data = regression_data)
+competence <- lm(average_relative_competence ~ convergence + options + convergence*options, 
+                 data = regression_data)
+
+summary(competence)
+```
+
+### Vary competence
+
+Finally, we want to vary the underlying competence distribution. Along with it, we vary either sample size or number of choice options (holding the respective other constant).
+
+#### By constellation
+
+We classify samples into categories: dissensus, minority (any of potentially several minorites), majority (the answer with the most counts), and consensus.
+
+```{r, message=FALSE}
+n <- c(3, 5, 10, 20, 50, 100)
+alpha = c(0.1, 2, 1, 1, 6, 0.9)
+beta = c(1, 6, 1, 0.1, 2, 0.9)
+names = c("very incompetent", "incompetent", "uniform", 
+          "very competent", "competent", "bimodal")
+
+vary_competence(file_name = "sim_vary_competence_sample_3_options.csv", 
+                distribution = "beta", 
+                alpha = alpha, beta = beta, names = names, vary = "sample", n = n, 
+                iterations = 5, 
+                population = 1000, options = 3)
+
+varied_competence <- read_csv("data/sim_vary_competence_sample_3_options.csv")
+```
+
+```{r}
+plot_competence_vary(varied_competence,  return = "competence distributions")
+plot_competence_vary(varied_competence , outcome = "Competence", variable = sample)
+plot_competence_vary(varied_competence , outcome = "Accuracy", variable = sample)
+```
+#### By relative majority
+
+Here, we do not group data into categories, but simply sort it by the share of participants (within a sample of informants) that agreed with an options. We call this relative majority.
+
+##### Vary sample size
+
+```{r}
+n <- c(3, 5, 10, 20, 50, 100)
+alpha = c(0.1, 2, 1, 1, 6, 0.9)
+beta = c(1, 6, 1, 0.1, 2, 0.9)
+names = c("very incompetent", "incompetent", "uniform", 
+          "very competent", "competent", "bimodal")
+
+# Vary the sample size, while holding constant on 3 choice options
+vary_competence(file_name = "sim_rel_majority_vary_competence_sample_3_options.csv", 
+                distribution = "beta", 
+                alpha = alpha, beta = beta, names = names, vary = "sample", n = n, 
+                iterations = 100, 
+                population = 1000, options = 3, return = "by degree of majority")
+
+vary_competence_sample <- read_csv("data/sim_rel_majority_vary_competence_sample_3_options.csv")
+```
+
+```{r}
+# Competence distributions
+plot_competence_distributions(vary_competence_sample)
+```
+
+```{r}
+plot_competence_vary_relative_majority(vary_competence_sample,
+                                       outcome = "Accuracy", variable = sample)
+```
+```{r}
+plot_competence_vary_relative_majority(vary_competence_sample,
+                                       outcome = "Competence", variable = sample)
+```
+
+##### Vary choice options
+
+```{r}
+n <- c(3, 5, 10, 20, 50, 100)
+alpha = c(0.1, 2, 1, 1, 6, 0.9)
+beta = c(1, 6, 1, 0.1, 2, 0.9)
+names = c("very incompetent", "incompetent", "uniform", 
+          "very competent", "competent", "bimodal")
+
+# Vary the sample size, while holding constant on 3 choice options
+vary_competence(file_name = "sim_rel_majority_vary_competence_options_3_sample.csv", 
+                distribution = "beta", 
+                alpha = alpha, beta = beta, names = names, vary = "options", n = n, 
+                iterations = 100, 
+                population = 1000, sample = 3, return = "by degree of majority")
+
+vary_competence_options <- read_csv("data/sim_rel_majority_vary_competence_options_3_sample.csv")
+```
+
+```{r}
+# Competence distributions
+plot_competence_distributions(vary_competence_options)
+```
+
+```{r}
+plot_competence_vary_relative_majority(vary_competence_options,
+                                       outcome = "Accuracy", variable = options)
+```
+
+```{r}
+plot_competence_vary_relative_majority(vary_competence_options,
+                                       outcome = "Competence", variable = options)
+```

### 2023-11-08 Pfander_DeCourson_Mercier_2023.Rmd
Remove cached *_files directories

---
title             : "How wise is the crowd? \n Can we infer people are accurate and competent merely because they agree with each other?"
shorttitle        : "How wise is the crowd?"

header-includes:  | # to prevent floats from moving past certain points (for the appendix)
  \usepackage{placeins}

author: 
  - name          : "Jan Pfänder"
    affiliation   : "1"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Data collection"
      - "Statistical Analyses"
      - "Writing - Original Draft"
      - "Writing - Review & Editing"
  - name          : "Benoît De Courson"
    affiliation   : "2"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Models and Simulations"
      - "Writing - Review & Editing"
  - name          : "Hugo Mercier"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "hugo.mercier@gmail.com"
    address: " "
    role:
      - "Conceptualization"
      - "Writing - Original Draft"
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Institut Jean Nicod, Département d’études cognitives, ENS, EHESS, PSL University, CNRS, France"
  - id            : "2"
    institution   : "Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany"

authornote: |
  JP received funding from the SCALUP ANR grant ANR-21-CE28-0016-01

  HM received funding from XX

abstract: |
  Are people who agree on something more likely to be right? Evidence suggests that people tend to make this inference. However, standard wisdom of crowds approaches do not provide normative grounds for this behavior. Using simulations and analytical arguments, we argue that in stylized scenarios where individuals make independent and unbiased estimates, more convergent groups individuals indeed tend to be more competent and accurate. Mirroring the stylized setting of the simulations, we then show that people make these inferences in a series of experiments. Importantly, in our experiments people do not have any other cues available than the social information (i.e. no priors). These inferences from convergence might help explain why people respect scientists' competence, even if they do not understand much about how scientific results are reached.

keywords          :

floatsintext      : yes
linenumbers       : no 
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "doc" # "doc" for nice look, "man" for manuscripty
output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors. For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g. set always allow html as true in the yaml heading

always_allow_html: true

appendix:
bibliography: references.bib
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  # fig.retina = 3, dpi = conditional_dpi,
  # fig.width = 7, fig.asp = 0.618,
  # fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  },
  options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)
```

```{r packages, include=FALSE}
# load required packages
library("papaja")      # For APA style manuscript   
library("lme4")        # model specification / estimation
library("lmerTest")    # provides p-values in the output
library("tidyverse")   # data wrangling and visualisation
library("afex")        # anova and deriving p-values from lmer
library("broom")       # extracting data from model fits 
library("broom.mixed") # extracting data from mixed models
library("metafor")     # doing mata analysis
library("metaviz")     # vizualization of meta analysis
library("stringr")     # for dmetar p-curve function to work
library("poibin")      # for dmetar p-curve function to work
library("patchwork")   # put several plots together
library("ggridges")    # for plots
library("gghalves")    # for plots
library("ggbeeswarm")  # Special distribution-shaped point jittering
library("knitr")       # for tables
library("kableExtra")  # also for tables
library("sf")          # for maps
library("ggpubr")      # for combining plots with ggarrange() 
library("viridis")     # for generating color palette for map 
```

```{r functions}
# load plot theme
source("functions/plot_theme.R")

# load other functions
source("functions/own_functions.R")
```

# Introduction

Imagine that you live in ancient Greece, and a fellow called Eratostenes claims the circumference of the earth is 252000 stades (approximately 40000 kilometers). You know nothing about this man, and you don't understand trigonometry. Moreover, you know little about the circumference of the Earth, and even less of how one could measure such a thing. As a result, you discard his opinion and take him for a pretentious loon. But what if other scholars had arrived at very similar measurements, independently of Eratosthenes? Or even if they had carefully checked his measurement, with a critical eye? Wouldn't that give you enough ground to believe not only that the estimates might be correct, but also that Eratosthenes and his fellow scholars must be quite bright, to be able to achieve such a feat?

In this article, we explore how, under some circumstances, we infer that a group of people who provided the same, or very similar answers to a question are likely to be correct, and to be competent in the relevant area, even if we had no prior about either what the correct answer was, or about their competence.

We first review experiments suggesting that people, including young children, appear to trust groups with more convergent opinions to be more accurate, although the evidence is more conclusive in numerical than in categorical choice contexts. We continue by explaining that two well-studied forms of the wisdom of crowds--averaging and the Condorcet Jury Theorem--do not provide normative grounds for this behavior. We then present the results of simulations, suggesting that under certain conditions, inferences from convergence to accuracy and to underlying competence are sound. Finally, we show that participants draw these inferences in a series of six experiments.

## Numerical choices

Informants can converge in numerical (e.g. how many calories are there in an apple?) and categorical (e.g. which option is correct - A or B?) choice contexts. In numerical contexts, convergence varies by the proximity of estimates; in categorical contexts, convergence varies by the ratio of people agreeing on a choice. We start by reviewing numerical choices.

For numerical choice contexts, evidence that people take convergence into account can be found in advice-taking experiments. In these experiments, participants are called judges, and they need to make numerical estimates--sometimes on factual knowledge, e.g. 'What year was the Suez Canal opened first?' [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g. 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020]. To help answer these questions, participants are given estimates from others, the advisors--sometimes presented as fellow participants, sometimes as experts. The overarching question in those studies is how people integrate social information from the advisors when making estimates.

A substantial share of studies using this judge-advisor paradigm is irrelevant for studying inferences from convergence, because they present participants only with single estimates. In some studies, that single estimate is coming from one advisor [e.g. @bednarikEffectivenessImperfectWeighting2015a; @sollStrategiesRevisingJudgment2009; @yanivReceivingOtherPeople2004; @yanivAdviceTakingDecision2000; @harveyTakingAdviceAccepting1997]. In other studies, that estimate is presented as an average coming from a group of advisors [e.g. @jaylesHowSocialInformation2017a; @mannesAreWeWise2009], but no information about the distribution of initial estimates is given. By contrast, other advice-taking studies provide participants with a set of estimates. One subset of these studies manipulate the degree of convergence between groups of advisors, e.g. via variance of estimates [@mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009], or their range [@budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007]. These studies find that participants are more confident about, or rely more on, estimates from groups of advisors who converge more. More indirect evidence comes from another subset of studies varied the degree of convergence within groups of advisors. They present participants with sets of estimates where some estimates are close to each other (or overlapping, in cases where estimates were intervals), while others were clearly outliers [@yanivWeightingTrimmingHeuristics1997, study 3 & 4; @harriesCombiningAdviceWeight2004]. These studies find that participants heavily discount outliers when aggregating estimates. These results are thus consistent with the idea that people rely more on convergent opinions, although testing extreme cases of singular dissident opinions. Another study varied convergence between estimates from single advisors [@yanivWeightingTrimmingHeuristics1997, study 1 & 2]. In this study, participants were shown ranges of estimates from two advisors - one with a small range, the other with a big range. Participants relied more on the precise, small range advice. These results can be interpreted as a case of inferring accuracy from (internal) convergence of single informants, perceived by participants perhaps as consistency or for example confidence. However, in this study, the small-range advisor (e.g. 3 to 5) was always within the range of the big-range adivsor (e.g. 1 to 10). If people relied more on the small-range advisor, then it was likely not only because of greater convergence, but also because of the overlap between the two advisors.

## Categorical choices

For categorical choice contexts, experimental evidence on people inferring accuracy from convergence is scarce. There is an extensive body of literature suggesting that people are susceptible to adopting majority behavior and opinions. But it is is difficult to tear apart whether they do so in a strive for normative conformity or for accuracy [@mercierMajorityRulesHow2019]. However, some aspects of the experimental paradigm should favor motivations for accuracy, for example private answers and individual rewards for accuracy. As for human adults, we are aware of only one article that assessing majority effects using such accuracy motivations: In a series of experiments using such accuracy motivations, @morganEvolutionaryBasisHuman2012 systematically varied the proportion of informants agreeing on a given answer, from 50 to 100%. They found that the larger the (relative) size of the majority, the more likely participants were to adopt the majority answer. As for children, relevant evidence comes from studies on social learning in the field of developmental psychology. Several such studies rely on a similar paradigm revolving around a naming tasks [@fusaroChildrenAssessInformant2008; @corriveauGoingFlowPreschoolers2009; @bernardChildrenWeighNumber2015a; @bernardFourSixYearOldChildren2015; @chenChildrenTrustConsensus2013]. In @corriveauGoingFlowPreschoolers2009, for example, children had to find out which of three unfamiliar objects corresponds to the made up name "modi". Three adult informants pointed to one object, but a fourth adult pointed to a different one. The children were then asked for their opinion. All these studies consistently find that children adopt the majority opinion. It is, however, not clear how the naming paradigm taps into intuitions for inferring accuracy - after all, assigning made-up names to objects has no objectively correct solution and, accordingly, these studies are often interpreted to reveal conformity in children [see e.g. @vanleeuwenConformityItsLookalikes2015]. Other experimental paradigms yield similar results. @morganDevelopmentAdaptiveConformity2015 had children (aged between 3 and 7) count which of two pictures displayed more dots. Among 10 adult informants, they varied the proportion of those agreeing on one picture. They find that children relied more on advice from larger (relative) majorities, although younger children (aged 3 to 4) were only affected in cases where the majority would stand against a single dissident. In @haunMajorityBiasedTransmissionChimpanzees2012, 2-year-olds tried to use an unfamiliar box to deliver a reward. The children were more likely to copy an action demonstrated by three informants rather than a different action demonstrated three times by the same informant. In @herrmannStickScriptEffect2013, children (aged 3 to 6) were required to copy a necklace-making demonstration from video. Children showed more fidelity to the demonstrations after watching two identical demonstrations by two different adults than after watching the same individual twice. In @vanleeuwenDevelopmentHumanSocial2018, children could use a choice box with three different pipes to insert a boll, to then receive a toy. Before using the box, they were shown a video of four other children using the box. Three of these children used one particular pipe, each throwing in one ball and receiving one toy per insertion, sequentially. One additional demonstrator used one of the two other pipes three times in a row, also receiving one toy per insertion. Children were then given one ball and had to chose which pipe to use. Children generally followed the majority choice.

Although the evidence is not conclusive, especially for categorical choice scenarios, the current literature suggests that people believe more convergent opinions to be more accurate. But is this inference sound?

## Standard wisdom of crowds approaches

Literature on the wisdom of crowds provides normative ground for inferring accuracy from others' aggregated opinions [@surowieckiWisdomCrowds2005]. What role does this literature assign to convergence? Here we look at two of the main phenomena that can generate wisdom of crowds effects: averaging, and the CJT.

### Averaging

In the beginning of the 20th century, Francis Galton famously demonstrated accuracy gains from averaging: in his “vox populi” experiment, he asked 787 participants to estimate the weight of an ox, and found the average estimate of an ox weight to be just one pound from reality [@galtonVoxPopuli1907]. Since then, it has been shown both theoretically and empirically that the mean or median answer of a group is typically closer to the truth than the mean individual is [see e.g. @yanivBenefitAdditionalOpinions2004; @mannesWisdomSelectCrowds2014]. For instance, when considering a range of numerical estimates that deviate more or less from a correct answer, the error of the mean answer will always be either lower than the mean error (if the correct answer is within the range of all the answers provided), or the same as the mean error (otherwise) [@larrickIntuitionsCombiningOpinions2006]. In fact, the error of the mean is often uncannily small compared to the mean error, a phenomenon which has allowed averaging to considerably improve performance on a variety of problems ranging from political predictions to medical diagnoses [@surowieckiWisdomCrowds2005].

Averaging works independently of the degree of convergence of a set of estimates. It is not clear how it leads to more accurate results, without making assumptions about the data generating process of these estimates.

[One might ask why talk about averaging at all - good question. I guess just because it happens in a numeric context and is used to justify aggregating opinions. But the link is not straightforward.]

### Condorcet Jury Theorem

In the late 18th Century, Condorcet established a formal argument in favor of large scale majority voting, which would later become known as the Condorcet Jury Theorem (CJT) [@decondorcetEssaiApplicationAnalyse2014]. He demonstrated that, for a binary decision (e.g. correct vs. false), the probability that the majority vote is correct is larger than the probability of each individual being correct. Moreover, as the number of votes increases, the probability that the majority is correct converges towards one. Crucially, the original CJT assumes that (i) each individual has the same probability p to select the correct answer, and (ii) that this probability is at least slightly better than chance (p \> 0.5) (there are also assumptions about the independence and lack of strategic motives, for reviews, see, e.g. @ladhaCondorcetJuryTheorem1992, @austen-smithInformationAggregationRationality1996, @dietrichEpistemicDemocracyDefensible2013).

At least in its basic version, the CJT glosses over variation in convergence, i.e. the relative size of the majority. It allows to make statements such as 'given an individual-level competence p, the probability that a (any) majority vote is correct is Y'. This latter probability comprises votes with 99% majorities as well 51% majorities. To justify inferences of accuracy and competence from convergence, CJT would need to allow making statements such as 'given a majority of X the probability that the majority vote is correct is Y, and the individual-level competence is Z.' In an extension of the CJT, @romeijnLearningJurorCompetence2011 provide normative grounds for making such statements. They show that with increasing relative majority, group members are more likely to be competent and the majority decision to be accurate (??). Their framework is limited, however, to binary decisions.

In sum, the wisdom of crowd literature provides at best insufficient normative grounds for inferring accuracy and competence from convergence. Its main purpose is to justify that aggregating across many opinions is a better strategy than picking any individual's opinion. The proposed aggregation strategies - whether averaging or following the majority - are about extracting wisdom from any crowd. But not all crowds are equally wise. Competence should play a crucial role in the confidence we have in any wisdom of crowd inferences. In the absence of more informative cues, our literature review shows that people place more trust in more convergent crowds. The wisdom of crowd literature does not provide generalizable justification for this behavior.

## Inferring competence from the convergence of opinions

As Eratosthenes’ example above illustrates, intuitively it seems there are cases in which we can infer that the members of a group are more likely to be correct, or, equivalently, more competent, when their answers converge. When are we justified to infer that people are right (and thus competent) when they converge?

To provide a normative answer, we built analytical and agent-based models (see \@ref(simulations)). We consider two scenarios in which opinions can converge: a numerical and categorical one. In a numerical setting, we measure convergences by the empirical variance. The closer the estimates, i.e. the smaller the empirical variance, the greater convergence. In the categorical setting, we measure convergence by the share of votes for an option. The larger the share of votes for one option, the greater convergence.

### Numeric

### Categorical

Agents answer a categorical question, with m+1 answers. We define the competence p of an individual as the probability to choose the right answer.  Each of the wrong answers has the same probability (1-p)/m to be chosen. We observe the answers of a population of n+1 agents, with diverse levels of competence drawn from a beta distribution – a flexible probability distribution that can be uniform, unimodal or bimodal, depending on the parameters. We assume that the shape parameters of the beta distribution ($\alpha$ and $\beta$) are known, but the right answer is not. We aim to infer the competence of an individual, based on the population answers.
In our analytical model, we use Bayes formula to compute the posterior competence distribution of competence for an individual, knowing that a number K of other participants agree with him. The posterior distribution is a mixture of two beta distributions, Beta($\alpha$+1, $\beta$) and Beta($\alpha$, $\alpha$+1), weighted respectively by the probabilities that K individuals would choose the right or the wrong answer. The model suggests that the more consensual an individual's vote is, the more that individual is estimated to be competent (Fig. \@ref(fig:figure-Benoit)).

(ref:figure-Benoit) Average competence of an informant as a function of K, the number of individuals who agree. In this example, we assume that the number of informants - not counting the focal informant - in a sample is n=10, m=5, and that the population competence ($\alpha$ = $\beta$ = 1).

```{r figure-Benoit, echo=FALSE, fig.cap="(ref:figure-Benoit)", out.width= "50%", fig.align="left"}
knitr::include_graphics("figures/figure_Benoit.png")
```

That said, the average competence is bounded between  and $\frac{\alpha}{\alpha+\beta+1}$ and $\frac{\alpha+1}{\alpha+\beta+1}$. In other words, even if all other individuals agree with an informant, a case in which the informant very likely chose the correct answer (add an accuracy figure, maybe?), the competence we should assign to this informant, according to our model, is not higher than $\frac{\alpha+1}{\alpha+\beta+1}$ (or $2/3$ in our example in Fig. \@ref(fig:figure-Benoit)). This counter-intuitive result is due to our framework: we assume that a certain competence distribution as a given, and only observe one decision per individual. Intuitively, in a world where even some highly incompetent people sometimes pick the correct answer by chance, being right (with the majority) once is not a sufficient signal of highest competence.

## Overview experiments

We test whether people infer that the members of a group are correct, and thus likely competent, when their answers converge in two series of experiments. In all experiments, participants see the estimates or choices of different individuals, which we will call informants. Participants were were not given any information about the estimation tasks and the informants - they had no priors.

In the first series (experiment 1 to 3), the scenario is about numerical estimates. In the second series (experiment 4 to 6), the scenario is about categorical choices.

In experiment 1 (numerical) and 4 (categorical), participants evaluated several groups of informants, with each informant providing an answer independently. We manipulated convergence, i.e how close estimates were (Exp. 1, numerical setting), or how consensual an answer was (Exp.4, categorical setting). We found that the more convergence, the more participants inferred accuracy and competence.

In experiments 3 (numerical) and 5 (categorical), we tested whether cues of informational dependency inferences hamper inferences from convergence. Besides convergence, we manipulated whether informants were independent (mimicking experiments 1 and 4) or shared a conflict of interest. We found that (i) participants inferred accuracy and competence from convergence even when informants were in a conflict of interest situation, but (ii) that these  inference are more enhanced when informants are independent. Using a slightly different experimental setting, we replicate our results from experiments 1 and 4.

In experiment 2, we tested a different form of informational dependency: informants were either independent informants or "had discussed at great length with each other" before making their estimate. We did not find a difference between the two conditions. This experiment was the only one in which we did not manipulate convergence - all informant groups were convergent.

In experiment 5, we tested whether more choice options (ten vs. three) enhance inferences from convergence. We reasoned that with more options, consensus is less likely to occur by chance, and thus people might find it more impressive. We did not find that the number of choice options altered effects of convergence.

All experiments were pre-registered. Pre-registration documents, data and code can be found on Open Science Framework (OSF, <https://osf.io/6abqy/>). All analyses were conducted in R (version 4.2.2) using R Studio.

# Numerical estimates: Experiments 1 to 3

## Experiment 1

```{r exp1}
# Analyze data of experiments and store results

# Experiment 1

# read data
exp1 <- read_csv("Experiment_1/data/cleaned.csv") %>% 
  # import experimental variables as factors and set levels
  mutate(
    across(c(convergence, number), ~as.factor(.x)),
    convergence = fct_relevel(convergence, "divergent", "convergent"), 
    number = fct_relevel(number, "small", "large"))

# run mixed model (random slope and intercept) with participants as random factor
exp1_model_accuracy = lmer(confidence ~ convergence + number + (1 + convergence | ID), data = exp1)
exp1_model_competence = lmer(confidence ~ convergence + number + (1 + convergence | ID), data = exp1)

# extract descriptives for inline reporting
exp1_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp1$ID),
  # Model results
  model_accuracy = text_ready(exp1_model_accuracy), 
  model_competence = text_ready(exp1_model_competence)
)
```

The first experiment was designed to test the effect of convergence on the (average) perceived accuracy of estimates (H1) and the perceived competence of a group of informants (H2). We decided to measure accuracy as confidence in one's estimate based on a group of informants estimates. Our two hypotheses therefore read:

***H1: When making a guess based on the estimates of (independent) informants, participants will be more confident about their guess when these estimates converge compared to when they diverge.***

***H2: Participants perceive (independent) informants whose estimates converge more as more competent than informants whose estimates diverge.***

### Participants

We recruited `r exp1_descriptives$n_subj` participants from the UK via Prolific. Not a single participant failed our attention check. The sample size was determined on the basis of a power analysis for a t-test to detect the difference between two dependent means ("matched pairs") run on G\*Power3. The analysis suggested that a combined sample of 199 would provide us with 80% power to detect a true effect size of Cohen's d $\geq$ 0.2 (alpha = .05, two-tailed).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "Some people are playing games in which they have to estimate various quantities. Each game is different. You have no idea how competent the people are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be hard while others are easy. Across various games, we will give you the answers of several players, and ask you questions about how good they are. As it so happens, for all the games, the estimates have to be between 1000 and 2000, but all the games are completely different otherwise, and might require different skills, be of different difficulties, etc. Each player in the game makes their own estimate, completely independent of the others". They were then presented to the results of eight such games and had to answer questions (Fig.1).

(ref:stimulus-exp1) Results of one of eight games that participants have to rate. The stimulus corresponds to a divergent, 10 players condition

```{r stimulus-exp1, echo=FALSE, fig.cap="(ref:stimulus-exp1)", out.width= "50%", fig.align="center"}
knitr::include_graphics("figures/example_stimulus_exp1.png")
```

#### Design

We manipulated two experimental factors: First, the convergence of the estimates (how close they were); second, the number of estimates (how many players there were). We did not make explicit hypotheses on the latter, but included some research questions (cf. pre-registration document). Each factor had two levels: convergent vs. divergent and low (three) vs. high (ten). We used a 2(convergence: divergent/convergent) x 2(number: low/high) within-participant design, with each participant seeing all of the conditions. Per condition, participants saw two sets of estimates (game results). Thus, each participant saw eight sets of estimates in total.

#### Stimuli

We generated sets of estimates with random draws from normal distributions. First, we varied the standard deviation of these distributions to simulate the degree of convergence (150 for divergence, 20 for convergence; estimate scale ranged from 1000 to 2000). Second, we varied the number of draws from these distributions per set (three for low number, 10 for high number). For each of the four possible resulting condition, we generated two random draws from the respective normal distribution.

The means of the normal distributions that we draw our estimates from were distinct between sets of estimates. Considering our within-participant design, we wanted to ensure that participants understood each set of estimates as being the result of a different, unrelated game. In order to assure that random draws from the distributions will (most likely) appear on the response scale (1000 to 2000), we constrained the means of all normal distributions to lie between the first and third quartile of the response scale (i.e. smallest possible mean was 1225 and largest 1775). We define a set of eight means--one for each set of estimates--that cover the range from first to third quartile of the predefined scale with an equal interval (1250, 1325, 1400, 1475, 1550, 1625, 1700, 1775). We randomly paired means with conditions when generating the stimuli. We then drew the set of estimates from the respective normal distributions given the assigned means and the condition constraints. We repeated this three times, resulting in three different series of eight sets of estimates. We randomly assign participants to one of these series. Additionally, for each participant, we randomize the order of appearance of the sets of estimates within the respective series. Images of all sets of estimates can be found on the OSF.

#### Dependent variables

For each set of estimates participants responded to several questions. We first asked participants to make a guess about the correct answer based on the estimates they see ("What would you guess is the correct answer, if there is one?"). Participants indicated their numeric guess using a slider on a line identical with the one they saw the estimates on. We did not analyze those guesses. They merely served as a basis for the next question, intended to measure perceived accuracy: "How confident are you that your answer is at least approximately correct?" on a 7-point Likert scale ("not confident at all" to "extremely confident")[^1]. Finally, participants were asked about the competence of the group of players whose estimates they saw: "On average, how good do you think these players are at the game?"), also on a 7-point Likert scale (from "not good at all" to "extremely good").

[^1]: We changed this measure in study 3. Participants did not make a guess and were asked "How accurate do you think...". The results are very similar.

### Results

To account from dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slope for participants, using the `lme4` package and its `lmer()` function in R. In the models, we control for the number of estimates (our second experimental factor). Figure \@ref(fig:exp1-plot) visualizes the results and table \@ref(tab:exp1-table) contains descriptive results. We find a positive effect of convergence on accuracy: Participants were more confident about their estimate in convergent scenarios ($\Delta$ Accuracy = `r exp1_descriptives$model_accuracy$convergenceconvergent$estimate` `r exp1_descriptives$model_accuracy$convergenceconvergent$ci`, p = `r exp1_descriptives$model_accuracy$convergenceconvergent$p.value`) than in divergent ones (Accuracy = `r exp1_descriptives$model_accuracy$intercept$estimate` `r exp1_descriptives$model_accuracy$intercept$ci`, p = `r exp1_descriptives$model_accuracy$intercept$p.value`). We also find a positive effect of convergence on competence: participants rated players as more competent in convergent scenarios ($\Delta$ Competence = `r exp1_descriptives$model_competence$convergenceconvergent$estimate` `r exp1_descriptives$model_competence$convergenceconvergent$ci`, p = `r exp1_descriptives$model_competence$convergenceconvergent$p.value`) than in divergent ones (Competence = `r exp1_descriptives$model_competence$intercept$estimate` `r exp1_descriptives$model_competence$intercept$ci`, p = `r exp1_descriptives$model_competence$intercept$p.value`).

```{r exp1-table}
exp1 %>% 
  group_by(convergence) %>% 
  summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(confidence_mean, " (sd = ", confidence_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")")
         ) %>% 
  rename(Convergence = convergence) %>% 
  select(Convergence, Accuracy, Competence) %>% 
  apa_table()
```

(ref:exp1-plot) Distributions of accuracy and competence by level of convergence.

```{r exp1-plot, fig.cap="(ref:exp1-plot)"}
# make mean data for plot
x_nudge <- 0.8

means <- exp1 %>%
  group_by(convergence) %>%
  summarize(confidence = mean(confidence), 
            competence = mean(competence)) %>% 
  pivot_longer(c(confidence, competence), 
               names_to = "outcome", 
               values_to = "value") %>% 
  mutate(x_position = ifelse(convergence == "convergent", value + x_nudge, 
                                value - x_nudge))

confidence_plot <- ggplot(data=exp1, aes(x=confidence, fill=convergence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
  # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "confidence"), 
             aes(xintercept = value, 
                 color = convergence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "confidence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "plasma") + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=exp1, aes(x=competence, fill=convergence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
  # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "competence"), 
             aes(xintercept = value, 
                 color = convergence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "competence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "plasma") + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, 
                    competence_plot + 
                      theme(axis.text.y = element_blank(),
                            axis.ticks.y = element_blank(),
                            axis.title.y = element_blank()), 
                    common.legend = T) +
  theme(legend.position="top") 
figure
```

## Experiment 2

```{r exp2}
# Analyze data of experiments and store results

# Experiment 2

# read data
exp2 <- read_csv("Experiment_2/data/cleaned.csv") %>% 
  # import experimental variable as factors and set levels
  mutate(
    independence = as.factor(independence),
    independence = fct_relevel(independence, "dependent", "independent")
  )

# run mixed model (random intercept) with participants as random factor
exp2_model_accuracy = lmer(confidence ~ independence + (1 |ID), data = exp2)
exp2_model_competence = lmer(competence ~ independence + (1 |ID), data = exp2)

# extract descriptives for inline reporting
exp2_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp2$ID),
  # Model results
  model_accuracy = text_ready(exp2_model_accuracy), 
  model_competence = text_ready(exp2_model_competence)
)
```

In experiment 1, informants were always independent. In experiment 2, we set out to test whether informational dependency alters inferences from convergence. We used the same setting as in experiment one, in which participants saw the fictional results of an estimation game. However, this time participants only saw convergent sets of estimates. We manipulated informational dependency: players were independent either independent or could "discuss at lenght" before making an estimate. Our hypotheses were:

***H1: When making a guess based on convergent estimates of informants, participants will be more confident about their guess when informants were independent compared to when they weren't (i.e. they could discuss before).***

***H2: Participants perceive informants whose estimates converge as more competent when they are independent, compared to when they weren't (i.e. they could discuss before).***

### Participants

We recruited `r exp2_descriptives$n_subj` participants from the UK via Prolific. Not a single participant failed our attention check. As for experiment 1, the sample size was determined on the basis of a power analysis for a t-test to detect the difference between two dependent means ("matched pairs") run on G\*Power3. The analysis suggested that a combined sample of 199 would provide us with 80% power to detect a true effect size of Cohen's d $\geq$ 0.2 (alpha = .05, two-tailed).

### Design

We manipulated informational dependency. According to the condition, players read a different introduction before seeing a set of estimates (see Table \@ref(tab:exp2-conditions)).

```{r exp2-conditions}
# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
data.frame(
  Condition = c("Independence", "Discussion"),
  Description = c(
    "Players are asked to make completely independent decisions – they cannot see each other’s estimates, or talk with each other before giving their estimates.",
    "Players are asked to talk with each other about the game at length before giving their estimates."
  )
) #%>% 
  # apa_table(align = c("m{3cm}", "m{12cm}")) 
```

#### Stimuli

We a subset of the stimuli we had generated for experiment one, namely convergent ones with three estimates. By contrast to experiment one, participants saw only two stimuli in total (one set of estimates per condition). Otherwise, we proceeded just as in experiment one: we randomly assigned individual participants to one of the three series of stimuli, and for each participant, we randomized the order of appearance of conditions.

#### Dependent variables

We relied on the same set of questions as in experiment one.

### Results

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept for participants, using the `lme4` package and its `lmer()` function in R. Figure \@ref(fig:exp2-plot) visualizes the results and table \@ref(tab:exp2-table) contains descriptive results. The data does not support our hypotheses. Participants were slightly less confident about their estimates when the converging informants were independent $\Delta$ Accuracy = `r exp2_descriptives$model_accuracy$independenceindependent$estimate` `r exp2_descriptives$model_accuracy$independenceindependent$ci`, p = `r exp2_descriptives$model_accuracy$independenceindependent$p.value`), compared to when they discussed (Accuracy = `r exp2_descriptives$model_accuracy$intercept$estimate` `r exp2_descriptives$model_accuracy$intercept$ci`, p = `r exp2_descriptives$model_accuracy$intercept$p.value`). The effect is small, but in the opposite direction of what we had predicted. We do not find an effect regarding competence.

```{r exp2-table}
exp2 %>% 
  group_by(independence) %>% 
  summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(confidence_mean, " (sd = ", confidence_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")")
         ) %>% 
  rename(Independence = independence) %>% 
  select(Independence, Accuracy, Competence) %>% 
  apa_table()
```

(ref:exp2-plot) Distributions of accuracy and competence by level of informational dependency.

```{r exp2-plot, fig.cap="(ref:exp2-plot)"}
# make mean data for plot
x_nudge <- 0.8

means <- exp2 %>%
  group_by(independence) %>%
  summarize(confidence = mean(confidence), 
            competence = mean(competence)) %>% 
  pivot_longer(c(confidence, competence), 
               names_to = "outcome", 
               values_to = "value") %>% 
  mutate(x_position = ifelse(independence == "dependent", value + x_nudge, 
                                value - x_nudge))

confidence_plot <- ggplot(data=exp2, aes(x=confidence, fill=independence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
 # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "confidence"), 
             aes(xintercept = value, 
                 color = independence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "confidence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1) +
  scale_color_brewer(palette = "Accent",
                    direction = 1) +
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=exp2, aes(x=competence, fill=independence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
 # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "competence"), 
             aes(xintercept = value, 
                 color = independence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "competence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1) +
  scale_color_brewer(palette = "Accent",
                    direction = 1) +
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, 
                    competence_plot + 
                      theme(axis.text.y = element_blank(),
                            axis.ticks.y = element_blank(),
                            axis.title.y = element_blank()), 
                    common.legend = T) +
  theme(legend.position="top") 
figure
```

## Experiment 3

```{r exp3}
# Analyze data of experiments and store results

# Experiment 3

# read data
exp3 <- read_csv("Experiment_3/data/cleaned.csv") %>% 
  # import experimental variable as factors and set levels
  mutate(
    across(c(convergence, independence), ~as.factor(.x)),
    convergence = fct_relevel(convergence, "divergent", "convergent"),
    independence = fct_relevel(independence, "conflict", "independent")
  )

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the independence condition
exp3_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence |ID), 
                           data = exp3 %>% filter(independence == "independent"))
exp3_model_h1_competence = lmer(competence ~ convergence + (1 + convergence |ID), 
                             data = exp3 %>% filter(independence == "independent"))

# H2
# run mixed model (random intercept and slope for convergence) with participants as random factor
exp3_model_accuracy = lmer(accuracy ~ convergence + independence + 
                            convergence*independence + (1 + convergence | ID), 
                           data = exp3)

exp3_model_competence = lmer(competence~ convergence + independence + 
                            convergence*independence + (1 + convergence | ID), 
                           data = exp3)

# extract descriptives for inline reporting
exp3_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp3$ID),
  # Model results
  model_h1_accuracy = text_ready(exp3_model_h1_accuracy), 
  model_h1_competence = text_ready(exp3_model_h1_competence),
  model_accuracy = text_ready(exp3_model_accuracy), 
  model_competence = text_ready(exp3_model_competence)
)
```

The aim of experiment 3 was twofold: First, we wanted to replicate the results of experiment 1 in a slightly more ecological scenario. In experiment 3, participants were told they saw (fictional) predictions of experts for stock values. Since we expected the findings of experiment one to replicate, our first set of hypotheses was:

***H1a: Participants perceive predictions of independent informants as more accurate when they converge compared to when they diverge.***

***H1b: Participants perceive independent informants as more competent when their predictions converge compared to when they diverge.***

Second, as in experiment two, we wanted to test the independence assumption -- but this time, against a different case of informational dependency, namely when informants shared a conflict of interest. In this condition, experts were described to gain from estimating the value of a stock at a certain price -- and that price always happened to be the mean of the the expert's estimates. Our second set of hypotheses was:

***H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are in a conflict of interest.***

***H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are in a conflict of interest.***

### Participants

The interaction design of our third experiment made the power analysis more complex and less standard than for experiments one and two. Because we could build upon data from the first experiment, we ran a power analysis by simulation. The simulation code is available on OSF, and the procedure is described in the pre-registration document. The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% (alpha for significance = 0.05) of the time. Due to uncertainty about the assumptions we made and because we had resources for a larger sample, we recruited `r exp3_descriptives$n_subj` participants for this study -- again, from the UK and via Prolific. Again, not a single participant failed our attention check.

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "You will see four scenarios in which several experts predict the future value of a stock. You have no idea how competent the experts are. It's also possible that some are really good while others are really bad. As it so happens, in all scenarios, the predictions for the value of the stock have to lie between 1000 and 2000. Other than that, the scenarios are completely unrelated: it is different experts predicting the values of different stocks every time." Participants then saw the four scenarios, each introduced by a text according to which condition the participant was assigned to (see Table \@ref(tab:conditions-exp3); for more detailed materials see pre-registration document on OSF).

```{r conditions-exp3}
# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
data.frame(
  Condition = c("Independence", "Conflict of interest"),
  Description = c(
    "Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.",
    "All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future."
  )
) #%>% 
  # apa_table(align = c("m{3cm}", "m{12cm}"))  
```

#### Design

We manipulated two factors: informational dependency (two levels, independence and conflict of interest, see Table \@ref(tab:conditions); between participants) and convergence (two levels, convergence and divergence; within participants).

#### Stimuli

Our design required that each participant sees four different sets of predictions--two convergent and two divergent ones. We generated new stimuli: As in experiment 1, estimates would appear on a scale from 1000 to 2000. By contrast with experiment 1, we generated the sets of estimates with random draws from uniform distributions (instead of normal distributions). We varied the range of these distributions according to convergence (60 for convergence, 600 for divergence). Switching from normal distributions to uniform distributions made 'unlucky' draws in which conditions are visually not distinguishable less likely.

Similar to the previous experiments, we also vary the value on the prediction scale (from 1000 to 2000) across which the range is centered. Considering our within-participant design, this, again, made it seem more likely that participants understand each set of predictions as being the result of a different stock, with a different true value. In order to assure all random draws from the distributions would appear on the response scale, we constrained the center of the uniform distributions to lie between 1300 and 1700. We define four center values -- one per set of predictions -- that divide this interval in (rounded) quartiles (1300, 1430, 1570, 1700). Given a center and a range, we then draw the predictions from uniform distributions. For example, in a draw for a divergent set of estimates with a center of 1700, each value within a range of 1400 and 2000 is equally likely to get selected. To avoid that single draws overlap too much within the same set, we defined a minimum space of 5 between the three predictions of a set.

To minimize confounding of convergence with a certain placement of the center value, we paired center values with conditions such that each condition appears once in each half of the scale and each condition gets one of the extreme values. For example, in one set the convergence condition would get assigned center values of 1300 and 1570, the divergent condition center values of 1430 and 1700). We generated four such series (all possible combinations) and randomly assigned participants to one of them. Additionally, for each participant, we randomize the order of appearance of the sets of predictions within the respective series.

For each set of predictions, we calculated the empirical mean based on the randomly drawn estimates. In the conflict of interest condition, this mean was inserted as the value that participants were be told experts would gain from. Consequently, the convergent predictions converge around what is said to be the incentivized value for the experts to choose.

#### Dependent variables

By contrast to the previous experiments, we operationalized accuracy as an actual question on accuracy this time. For each set of estimates participants responded to the following questions: First, "On average, how accurate do you think these three predictions are?" on a 7-point Likert scale ("not accurate at all" to "extremely accurate"); second, "On average, how good do you think these three experts are at predicting the value of stocks?", also assessed on a 7-point Likert scale (from "not good at all" to "extremely good").

### Results

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants, using the `lme4` package and its `lmer()` function in R. Figure \@ref(fig:exp3-plot) visualizes the results and Table \@ref(tab:exp3-table) contains descriptive results.

We find evidence all four hypotheses. As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample of experiment three to half of the participants, namely those who were assigned to the independence condition. On this reduced sample, we ran the exact same analyses as in experiment 1 and replicated the results (see left side of Fig. \@ref(fig:exp3-plot)). We find a positive effect of convergence on accuracy ($\Delta$ Accuracy = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$estimate` `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$ci`, p = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$p.value`; baseline Accuracy divergent: `r exp3_descriptives$model_h1_accuracy$intercept$estimate` `r exp3_descriptives$model_h1_accuracy$intercept$ci`, p = `r exp3_descriptives$model_h1_accuracy$intercept$p.value`). We find very similar results for competence ($\Delta$ Competence = `r exp3_descriptives$model_h1_competence$convergenceconvergent$estimate` `r exp3_descriptives$model_h1_competence$convergenceconvergent$ci`, p = `r exp3_descriptives$model_h1_competence$convergenceconvergent$p.value`; baseline Competence divergent: `r exp3_descriptives$model_h1_competence$intercept$estimate` `r exp3_descriptives$model_h1_competence$intercept$ci`, p = `r exp3_descriptives$model_h1_competence$intercept$p.value`).

The second set of hypotheses targeted the interaction of informational dependency and convergence (for a visual representation of these interactions, see Fig. \@ref(fig:exp3-plot)). In the independence condition, the effect of convergence on accuracy was more positive ($\Delta$ Convergence = `r exp3_descriptives$model_accuracy$interaction$estimate` `r exp3_descriptives$model_accuracy$interaction$ci`, p = `r exp3_descriptives$model_accuracy$interaction$p.value`) than in the conflict of interest condition (Convergence = `r exp3_descriptives$model_accuracy$convergenceconvergent$estimate` `r exp3_descriptives$model_accuracy$convergenceconvergent$ci`, p = `r exp3_descriptives$model_accuracy$convergenceconvergent$p.value`). Likewise the effect of convergence on competence is more positive ($\Delta$ Convergence = `r exp3_descriptives$model_competence$interaction$estimate` `r exp3_descriptives$model_competence$interaction$ci`, p = `r exp3_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition (Convergence = `r exp3_descriptives$model_competence$convergenceconvergent$estimate` `r exp3_descriptives$model_competence$convergenceconvergent$ci`, p = `r exp3_descriptives$model_competence$convergenceconvergent$p.value`).

```{r exp3-table}
exp3_accuracy <- exp3 %>% 
  group_by(independence, convergence) %>% 
  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
  select(independence, convergence, Accuracy) %>% 
    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
                "Convergent")) %>%  
  pivot_wider(names_from = convergence,
              values_from = Accuracy) %>% 
  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
                "Independent") 
         ) %>% 
  rename(" " = independence) %>% 
  as.matrix()

exp3_competence <- exp3 %>% 
  group_by(independence, convergence) %>% 
  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
  select(independence, convergence, Competence) %>% 
    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
                "Convergent")) %>%  
  pivot_wider(names_from = convergence,
              values_from = Competence) %>% 
  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
                "Independent") 
         ) %>% 
  ungroup() %>% 
  select(-independence) %>% 
  as.matrix()

table <- cbind(exp3_accuracy, exp3_competence)

table %>%
  kable(booktabs = TRUE, 
        # need to add a (empty) caption for cross referencing to work
        caption = "") %>%
  kable_styling() %>% 
  add_header_above(c("", "Accuracy" = 2, "Competence" = 2))

```

(ref:exp3-plot) Distributions of accuracy and competence by convergence and informational dependency.

```{r exp3-plot, fig.cap="(ref:exp3-plot)"}
interaction_accuracy <- ggplot(exp3, aes(x=independence, y=accuracy, fill = convergence, 
                     shape = convergence,
                     group = convergence,
                     color = convergence)) +
  scale_x_discrete(limits = c("conflict", "independent"), 
                    labels = c("Conflict of interest", "Independent")) +
  geom_half_violin(data = exp3 %>% filter(independence=="conflict"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = exp3 %>% filter(independence=="independent"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Accuracy") +
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7)) +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none",
         fill = guide_legend(title = NULL)) +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

interaction_competence <- ggplot(exp3, aes(x=independence, y=competence, 
                                        fill = convergence, 
                                        shape = convergence, 
                                        group = convergence, 
                                        color = convergence)) +
  scale_x_discrete(limits = c("conflict", "independent"), 
                    labels = c("Conflict of interest", "Independent")) +
  geom_half_violin(data = exp3 %>% filter(independence=="conflict"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = exp3 %>% filter(independence=="independent"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Competence")+
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7)) +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none") +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

figure <- ggarrange(interaction_accuracy, interaction_competence, 
                    common.legend = T) +
  theme(legend.position="top") 
figure
```

# Categorical estimates: Experiments 4 to 6

In the first series of experiments (experiments 1-3) we tested inferences from convergence in a numerical choice setting: Participants saw (fictive) players' numeric estimates on a scale from 1000 to 2000. The degree of convergence varied by the distance between estimates.

In the second series of experiments (experiments 4-6), we test inferences from convergence in a categorical choice setting. In the categorical scenario, the fictive players make choices on a set of response options (i.e. categories). Convergence varies by the ratio of people agreeing on an option. Experiment four and five can be considered robustness checks as to whether the results of the first series hold in a categorical choice setting. Experiment six tests a new context factor: the number of choice options.

## Experiment 4

```{r exp4}
# Analyze data of experiments and store results

# Experiment 4

# read data
exp4 <- read_csv("Experiment_4/data/cleaned.csv") %>% 
  # import experimental variable as factor and set levels
  mutate(
    # set levels for `convergence`
    convergence_categorical = recode_factor(convergence, 
                                            `0` = "opposing majority", 
                                            `1` = "divergence", 
                                            `2` = "majority", 
                                            `3` = "consensus",
                                            .default = NA_character_))

# run mixed model (random slope and intercept) with participants as random factor
exp4_model_accuracy = lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = exp4)
exp4_model_competence = lmer(competence ~ convergence + (1 + convergence | id), 
                                 data = exp4)

# extract descriptives for inline reporting
exp4_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp1$ID),
  # Model results
  model_accuracy = text_ready(exp4_model_accuracy), 
  model_competence = text_ready(exp4_model_competence)
)
```

In line with our findings in numerical choice settings, we predicted that:

***H1: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.***

***H2: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.***

### Participants

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found in the `power_Exp4.Rmd` document. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%. The power simulation for `accuracy` suggested that even for as few as 10 participants (the minimum sample size we simulated data for), we would have a power of close to 100%. The simulation for `competence` suggested that we achieve statistical power of at least 90% with a sample size of `30`. Due to uncertainty about our assumptions, we recruited `r exp4_descriptives$n_subj` participants.

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "To be able to understand the task, please read the following instructions carefully: Some people are playing games in which they have to select the correct answer among three answers. You will see the results of several of these games. Each game is different, with different solutions and involving different players. All players answer independently of each other. At first, you have no idea how competent each individual player is: they might be completely at chance, or be very good at the task. It's also possible that some players are really good while others are really bad. Some games might be difficult while others are easy. Your task will be to evaluate the performance of one of the players based on what everyone's answers are." They were then presented to the results of eight such games and had to answer questions (see Fig. \@ref(fig:stimulus-exp4)).

(ref:stimulus-exp4) An example of a stimulus for the majority condition

```{r stimulus-exp4, echo=FALSE, fig.cap="(ref:stimulus-exp4)"}
knitr::include_graphics("figures/example_stimulus_exp4.png")
```

#### Design

We manipulated convergence by varying the ratio of players choosing the same response as a focal player (i.e. the one that participants evaluate). The levels of convergence are: (i) consensus, where all three players pick the same option [`coded value = 3`]; (ii) majority, where either the third or second player picks the same option as the first player [`coded value = 2`]; (iii) dissensus, where all three players pick different options [`coded value = 1`]; (iv) majority against the focal player's estimate, where the second and third player pick the same option, but one that is different from the first player's choice [`coded value = 0`]. In our analysis, we treat convergence as a continuous variable, assigning the values in squared parenthesis.

#### Stimuli

We manipulate convergence within participants. All participants see all four conditions, with two stimuli per condition. Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli).

#### Dependent variables

To assess accuracy, we asked: "What do you think is the probability of player 1 being correct?". Participants answered with a slider on a scale from 0 to 100. As for competence, we asked participants: "How competent do you think player 1 is in games like these?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

### Results

To account from dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slope for participants, using the `lme4` package and its `lmer()` function in R. Figure \@ref(fig:exp4-plot) visualizes the results and table \@ref(tab:exp4-table) contains descriptive results.

As in the numerical setting, we found a positive effect of convergence on both accuracy (Accuracy = `r exp4_descriptives$model_accuracy$convergence$estimate` `r exp4_descriptives$model_accuracy$convergence$ci`, p = `r exp4_descriptives$model_accuracy$convergence$p.value`) and competence (Competence = `r exp4_descriptives$model_competencey$convergence$estimate` `r exp4_descriptives$model_competence$convergence$ci`, p = `r exp4_descriptives$model_competence$convergence$p.value`).

```{r exp4-table}
exp4 %>% 
  group_by(convergence_categorical, convergence) %>% 
  summarise(across(c(competence, accuracy), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")"),
         Convergence = paste0(convergence_categorical, " (", convergence, ")")
         ) %>% 
  ungroup() %>% 
  select(Convergence, Accuracy, Competence) %>% 
  apa_table()
```

(ref:exp4-plot) Distributions of accuracy and competence by level of convergence.

```{r exp4-plot, fig.cap="(ref:exp4-plot)"}
# plot for accuracy
plot_accuracy <- ggplot(exp4,
       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_competence <- ggplot(exp4,
       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .8,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  scale_y_continuous(limits = c(1,7), breaks = seq(1,7)) +
  labs(x = "Convergence", y = "Competence") +
  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
  guides(fill = FALSE) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

plot_accuracy + plot_competence
```

## Experiment 5

```{r exp5}
# Analyze data of experiments and store results

# Experiment 5

# read data
exp5 <- read_csv("Experiment_5/data/cleaned.csv") %>% 
  # import experimental variable as factor and set levels
  mutate(
    # set levels for `convergence`
    convergence_categorical = recode_factor(convergence, 
                                            `0` = "opposing majority", 
                                            `1` = "divergence", 
                                            `2` = "majority", 
                                            `3` = "consensus",
                                            .default = NA_character_))

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the independence condition
exp5_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence | id), 
                              data = exp5 %>% filter(independence == "independent"))
exp5_model_h1_competence = lmer(competence ~ convergence + (1 + convergence | id), 
                              data = exp5 %>% filter(independence == "independent"))

# H2 
# run mixed model (random slope and intercept) with participants as random factor
exp5_model_accuracy = lmer(accuracy ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = exp5)
exp5_model_competence = lmer(competence ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = exp5)

# extract descriptives for inline reporting
exp5_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp1$ID),
  # Model results
  model_h1_accuracy = text_ready(exp5_model_h1_accuracy), 
  model_h1_competence = text_ready(exp5_model_h1_competence),
  model_accuracy = text_ready(exp5_model_accuracy), 
  model_competence = text_ready(exp5_model_competence)
)
```

In line with our findings in numerical choice settings, we predicted that:

***H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.***

***H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.***

***H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).***

***H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).***

### Participants

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found in the `power_Exp5.Rmd` document. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%.

The power simulation for `accuracy` suggested that for 80 participants, we would have a power of at least 90% for the interaction effect. The simulation for `competence` suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we also detect an effect of convergence. Due to uncertainty about our assumptions and because resources were available for a larger sample, we recruited `r exp5_descriptives$n_subj` participants.

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "We will show you three financial advisors who are giving recommendations on investment decisions. They can choose between three investment options. Their task is to recommend one. You will see several such situations. They are completely unrelated: it is different advisors evaluating different investments every time. At first you have no idea how competent the advisors are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be difficult while others are easy. Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are." They were then presented to the results of eight such games and had to answer questions (see Fig. \@ref(fig:stimulus-exp4)).

#### Design

We manipulated convergence within participants in the same way we did in experiment 4. In addition, between participants, we manipulated informational dependence, akin to experiment 3. In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is. In the independent condition, there was no such conflict of interest and experts were described as independent.

```{r}
# Participants read only one of these introductions before seeing the sets of estimates, depending on which condition they had been assigned to
data.frame(
  Condition = c("Independence condition", "Conflict of interest condition"),
  Description = c(
    "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.",
    "The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations."
  )
) #%>% 
  # apa_table(align = c("m{3cm}", "m{12cm}")) 
```

#### Stimuli

Identical to experiment 4, all participants see all four conditions, with two stimuli (i.e. game results) per condition. Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli).

#### Dependent variables

To assess accuracy, we asked participants: "What do you think is the probability of advisor 1 making the best investment recommendation?". Participants answered with a slider on a scale from 0 to 100. As for competence, we asked participants: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

### Results

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants, using the `lme4` package and its `lmer()` function in R. Figure \@ref(fig:exp5-plot) visualizes the results and table \@ref(tab:exp5-table) contains descriptive results.

Akin to the results of experiment 3 in the numerical setting, we find evidence for all four hypotheses. As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample to half of the participants, namely those who were assigned to the independence condition. On this reduced sample, we ran the exact same analyses as in experiment 4 and replicated the results (see blue colored distributions in Fig. \@ref(fig:exp5-plot)). We find a positive effect of convergence on both accuracy (Convergence = `r exp5_descriptives$model_h1_accuracy$convergence$estimate` `r exp5_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp5_descriptives$model_h1_accuracy$convergence$p.value`) and competence (Convergence = `r exp5_descriptives$model_h1_competence$convergence$estimate` `r exp5_descriptives$model_h1_competence$convergence$ci`, p = `r exp5_descriptives$model_h1_competence$convergence$p.value`). The second set of hypotheses targeted the interaction of informational dependency and convergence (for a visual representation of these interactions, see Fig. \@ref(fig:exp5-plot)). In the independence condition, the effect of convergence on accuracy was more positive ($\Delta$ Convergence = `r exp5_descriptives$model_accuracy$interaction$estimate` `r exp5_descriptives$model_accuracy$interaction$ci`, p = `r exp5_descriptives$model_accuracy$interaction$p.value`) than in the conflict of interest condition (Convergence = `r exp5_descriptives$model_accuracy$convergence$estimate` `r exp5_descriptives$model_accuracy$convergence$ci`, p = `r exp5_descriptives$model_accuracy$convergence$p.value`). Likewise, the effect of convergence on competence was more positive ($\Delta$ Convergence = `r exp5_descriptives$model_competence$interaction$estimate` `r exp5_descriptives$model_competence$interaction$ci`, p = `r exp5_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition (Convergence = `r exp5_descriptives$model_competence$convergence$estimate` `r exp5_descriptives$model_competence$convergence$ci`, p = `r exp5_descriptives$model_competence$convergence$p.value`).

```{r exp5-table}
exp5_accuracy <- exp5 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(independence, Convergence, convergence) %>% 
  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
  select(independence, Accuracy, Convergence, convergence) %>% 
mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
                "Independent") 
         ) %>% 
  pivot_wider(names_from = independence,
              values_from = Accuracy) %>% 
  arrange(convergence) %>% 
  ungroup() %>% 
  select(-convergence) %>% 
  rename(" " = Convergence) %>% 
  as.matrix()

exp5_competence <- exp5 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(independence, Convergence, convergence) %>% 
  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
  select(independence, Competence, Convergence, convergence) %>% 
mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
                "Independent") 
         ) %>% 
  pivot_wider(names_from = independence,
              values_from = Competence) %>% 
  ungroup() %>% 
  arrange(convergence) %>% 
  select(-c(convergence, Convergence)) %>% 
  as.matrix()

table <- cbind(exp5_accuracy, exp5_competence)

table %>%
  kable(booktabs = TRUE, 
        # need to add a (empty) caption for cross referencing to work
        caption = "") %>%
  kable_styling() %>% 
  add_header_above(c("", "Accuracy" = 2, "Competence" = 2)) %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down") 
```

(ref:exp5-plot) Interaction of convergence and informational dependency.

```{r exp5-plot, fig.cap="(ref:exp5-plot)"}
# plot for accuracy
plot_accuracy <- ggplot(exp5,
       aes(x = convergence_categorical, y = accuracy, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_competence <- ggplot(exp5,
       aes(x = convergence_categorical, y = competence, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE)
```

## Experiment 6

```{r exp6}
# Analyze data of experiments and store results

# Experiment 6

# read data
exp6 <- read_csv("Experiment_6/data/cleaned.csv") %>% 
  # import experimental variable as factor and set levels
  mutate(
    # set levels for `convergence`
    convergence_categorical = recode_factor(convergence, 
                                            `0` = "opposing majority", 
                                            `1` = "divergence", 
                                            `2` = "majority", 
                                            `3` = "consensus",
                                            .default = NA_character_))

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the three options condition
exp6_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence | id), 
                              data = exp6 %>% filter(number_options == "three"))
exp6_model_h1_competence = lmer(competence ~ convergence + (1 + convergence | id), 
                              data = exp6 %>% filter(number_options == "three"))

# H2 
# run mixed model (random slope and intercept) with participants as random factor
exp6_model_accuracy = lmer(accuracy ~ convergence + number_options + 
                            number_options*convergence + (1 + convergence | id), 
                       data = exp6)
exp6_model_competence = lmer(competence ~ convergence + number_options + 
                            number_options*convergence + (1 + convergence | id), 
                       data = exp6)

# extract descriptives for inline reporting
exp6_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp1$ID),
  # Model results
  model_h1_accuracy = text_ready(exp6_model_h1_accuracy), 
  model_h1_competence = text_ready(exp6_model_h1_competence),
  model_accuracy = text_ready(exp6_model_accuracy), 
  model_competence = text_ready(exp6_model_competence)
)
```

Experiment four and five tested if the results from the first series of experiments hold in a categorical choice setting. In Experiment six we tested a new context factor: the number of choice options. In experiments four and five, scenarios always involved three choice options. Here, we varied between three and ten options. The design we use to manipulate convergence is otherwise identical to experiment 4 (players playing games).

First, considering only the three options condition, we ran a direct replication of experiment 4. Second, following the results from our model, we predict that

***H1: The effect of convergence on accuracy (H1a) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.***

***H2: The effect of convergence on competence (H1b) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.***

### Participants

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found in the `power_Exp6.Rmd` document. We used previous experiments and estimates of our models to inform our choice of parameter values. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%. The power simulation for `accuracy` suggested that for 140 participants we would cross the power threshold of 90% for the interaction effect (power = 0.928). The simulation for `competence` suggested that with 300 participants, we would detect an interaction with a power of 87% (power = 0.872). Due to budget constraints, we considered a sample of 300 participants as good enough, although slightly below our threshold.

### Procedure

We used the same experimental setting as in experiment 4.

#### Design

We manipulated convergence within participants in the same way we did in experiment 4. In addition, between participants, we manipulated the number of choice options. Participants were randomly assigned to either to see stimuli with '3' options, or with '10' options. Participants assigned to the '10' options condition, participants were assigned to one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the '3' options condition, and another with increased range (see Appendix XX). We added the increased range condition because we anticipated the possibility that participants might not consider all options as relevant when they only see scenarios in which all answers cluster. We found no differences between the two sub-conditions and collapsed them into a single '10' options condition.

```{r stimuli-options, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(
  imgage_a = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_3_a.png){ width=60% }"),
  imgage_b = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_10_a.png){ width=60% }")
  )

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Number of options: 3", "Number of options: 10"), 
                  align = "l",
                  caption = "Example of a consensus stimulus for the two 'Number of option' conditions") %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down") 
```

#### Stimuli

Identical to experiment 4, all participants saw all four conditions, with two stimuli (i.e. game results) per condition. Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).

#### Dependent variables

We used the same items as in experiment 4. To assess accuracy, we asked: "What do you think is the probability of player 1 being correct?". Participants answered with a slider on a scale from 0 to 100. As for competence, we asked participants: "How competent do you think player 1 is in games like these?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

### Results

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants, using the `lme4` package and its `lmer()` function in R. Figure \@ref(fig:exp5-plot) visualizes the results and table \@ref(tab:exp5-table) contains descriptive results.

We replicate the results of experiment 4, but do not find evidence for an interaction between convergence and the number of choice options. To match the setting of experiment one, we reduced the sample to half of the participants, namely those who were assigned to the three options condition. On this reduced sample, we ran the exact same analyses as in experiment 4 and replicated the results (see orange colored distributions in Fig. \@ref(fig:exp6-plot)). We find a positive effect of convergence on both accuracy (Convergence = `r exp6_descriptives$model_h1_accuracy$convergence$estimate` `r exp6_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp6_descriptives$model_h1_accuracy$convergence$p.value`) and competence (Convergence = `r exp6_descriptives$model_h1_competence$convergence$estimate` `r exp6_descriptives$model_h1_competence$convergence$ci`, p = `r exp6_descriptives$model_h1_competence$convergence$p.value`). We do not find evidence that the number of choice options alters the effect of convergence on either accuracy or competence.

```{r exp6-table}
exp6_accuracy <- exp6 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(number_options, Convergence, convergence) %>% 
  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
  select(number_options, Accuracy, Convergence, convergence) %>% 
mutate(number_options = ifelse(number_options == "three", "3 options", 
                "10 options") 
         ) %>% 
  pivot_wider(names_from = number_options,
              values_from = Accuracy) %>% 
  arrange(convergence) %>% 
  ungroup() %>% 
  select(-convergence) %>% 
  rename(" " = Convergence) %>% 
  as.matrix()

exp6_competence <- exp6 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(number_options, Convergence, convergence) %>% 
  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
  select(number_options, Competence, Convergence, convergence) %>% 
mutate(number_options = ifelse(number_options == "three", "3 options", 
                "10 options") 
         ) %>% 
  pivot_wider(names_from = number_options,
              values_from = Competence) %>% 
  ungroup() %>% 
  arrange(convergence) %>% 
  select(-c(convergence, Convergence)) %>% 
  as.matrix()

table <- cbind(exp6_accuracy, exp6_competence)

table %>%
  kable(booktabs = TRUE, 
        # need to add a (empty) caption for cross referencing to work
        caption = "") %>%
  kable_styling() %>% 
  add_header_above(c("", "Accuracy" = 2, "Competence" = 2)) %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down") 
```

(ref:exp6-plot) Interaction of convergence and informational dependency.

```{r exp6-plot, fig.cap="(ref:exp6-plot)"}
# colors were identified with the wesanderson package (library(wesanderson)) as:
# wes_palette("Zissou1", 100, type = "continuous")[A number]

# plot for accuracy
plot_accuracy <- ggplot(exp6,
       aes(x = convergence_categorical, y = accuracy, fill = number_options)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy", fill = NULL) +
  scale_fill_manual(values = c("#EE3800", "#3D9BB2"), 
                    name = "Number of choice options: ") +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_competence <- ggplot(exp6,
       aes(x = convergence_categorical, y = competence, fill = number_options)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence", fill = NULL) +
  scale_fill_manual(values = c("#EE3800", "#3D9BB2")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE)
```

# Discussion

This literature review shows that there is evidence for both adults and children to be susceptible to convergence. It is not always clear, however, whether people infer accuracy from. It also shows (although not really outlined above) that this is only true in boundary conditions - when no bettter information (e.g. strong priors) or information about the competence o informants is available.

When that is the case, peopl tend to favour their own estimates, a phenomenon known as egocentric discounting, or prefer in

Inferences of convergence is one mechanism that work inside a system of epistemic vigiliance.

In many cases, when better knowledge is available, this cognitive mechanism is outperformed by others.

-   egocentric discounting in adults
-   perceptual cues + performance hints in children
-   older children are more critical of consensus

there are, however, many real-world scenarios in which people have little priors on an estimation task, and little knowledge about the competence of the advisor. In these cases, people have been shown to be especially sensitive to consensus.

## Limitations

However, as the original CJT, they still require a minimum of individual-level competence for a majority decision to be accurate (also, non-strategizing and independence). Therefore, without any information about individual voters' competence, the CJT does not justify inferring accuracy from a majority vote.

### Data availability

The extracted data used to produce our results are available on the OSF project page (<https://osf.io/96zbp/?view_only=d2f3147f652e44e2a0414d7d6d9a6c29>).

### Code availability

The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (<https://osf.io/96zbp/?view_only=d2f3147f652e44e2a0414d7d6d9a6c29>).

### Competing interest

The authors declare having no competing interests.

\FloatBarrier

# References

<div id="refs"></div>

\newpage

# Appendix A {simulations}

Are we justified in inferring competence and accuracy from convergence? To provide a normative answer, we ran simulations, intended to mirror our experimental setup. We find that - under certain conditions - more convergent groups indeed tend to be more competent and accurate. In this appendix, we describe these simulations in detail.

## Numerical choice context

```{r read-simulation-data-numeric, message=FALSE}
# read data from simulations
simulation_numeric <- read_csv("Experiment_1/data/sim.csv")
```

When several people estimate a quantity (numeric scenario), their convergence can be measured for example by the empirical variance. The closer the estimates, i.e. the smaller the empirical variance, the greater convergence. This measure is at the group level.

To provide a normative answer, we ran simulations for a scenario in which individuals provide an estimate on a scale from 1000 to 2000. In our simulations, we suppose that an individual’s answer is drawn from a normal distribution. Each individual has their own normal distribution. All normal distributions are centered around the true answer - but they differ in their standard deviations. The value of that standard deviation is what we define as an individual’s competence. The lower the standard deviation, the higher the competence, i.e. the more likely a guess drawn from the normal distribution will be close to the true answer. We (arbitrarily) define a range of competence: we set the lowest competence equal to the range of possible values, i.e.the largest standard deviation (2000 - 1000) = 1000. We set the highest competence to 0.1% of the range of possible values, i.e.the smallest standard deviation (0.001 x 1000) = 1 (see Fig. \@ref(fig:example-competence-numeric)).

(ref:example-competence-numeric) Range of possible data generating functions for individuals.

```{r example-competence-numeric, fig.cap="(ref:example-competence-numeric)", out.width= "75%"}
# Define the x-axis values
x <- seq(1000, 2000, length.out = 1000)

# Define the PDFs for the two distributions
high_competence_pdf <- dnorm(x, mean = 1500, sd = 1)
low_competence_pdf <- dnorm(x, mean = 1500, sd = 1000)

# Create the plot
ggplot() +
  geom_line(aes(x, high_competence_pdf, color = "Highest Competence Individual \n (SD = 1, Mean = 1500)"), size = 1) +
  geom_line(aes(x, low_competence_pdf, color = "Lowest Competence Individual \n (SD = 1000, Mean = 1500)"), size = 1) +
  labs(x = "Competence Level", y = "Density", color = "Data generating function for") +
  ggtitle("Competence Level Distributions") +
  scale_color_manual(values = c("Highest Competence Individual \n (SD = 1, Mean = 1500)" = "blue", 
                                "Lowest Competence Individual \n (SD = 1000, Mean = 1500)" = "red")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

We suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution. This competence distribution can take vary different shapes, depending on the alpha and beta parameters that describe the distribution (see fig. \@ref(fig:competence-distributions)).

(ref:competence-distributions) The different population competence distributions we considered in our simulations.

```{r competence-distributions, fig.cap="(ref:competence-distributions)", out.width= "75%"}
plot_competence_distributions(simulation_numeric)
```

We draw an estimate for each individual based on their respective competence distribution. For each individual, we then measure accuracy as the (squared) distance between their estimate and the true answer. Having a competence and an accuracy value for each individual, we randomly assign individuals to groups of three. For each group, we calculate the average of the three individuals' competence and accuracy. We measure the convergence of a group by calculating the standard deviation of the estimates. We run this simulation on a sample size of roughly 100000 (varying slightly as a function of sample size). We repeat this simulation process for various sample sizes and competence distributions. The results are displayed in Fig. \@ref(fig:simulation-accuracy-numeric-1) to \@ref(fig:simulation-accuracy-numeric-6), for accuracy, and Fig. \@ref(fig:simulation-competence-numeric-1) to \@ref(fig:simulation-competence-numeric-6), for competence. Across all underlying competence distributions, we find a positive correlation between convergence and accuracy, which tends towards 1 as sample size increases (see Fig. \@ref(fig:simulation-numeric-correlations)). As for accuracy, we find a positive correlation between convergence and competence across all underlying competence distributions. However, this correlations are weaker than for accuracy, and do not increase with sample size (see Fig. \@ref(fig:simulation-numeric-correlations)).

(ref:simulation-numeric-correlations) Correlation between accuracy (left) or competence (right) and convergence, as a function of sample size, grouped by the data underlying competence distributions.

```{r simulation-numeric-correlations, fig.cap="(ref:simulation-numeric-correlations)", out.width="75%"}
# calculate correlations
simulation_numeric <- simulation_numeric %>% 
  group_by(competence, sample) %>% 
  mutate(correlation_accuracy = cor(accuracy_mean, convergence), 
         correlation_competence = cor(competence_mean, convergence)) %>% 
  ungroup()

# accuracy plot
accuracy_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_accuracy, color = competence)) +
  geom_point() + 
  geom_line(alpha = 0.5) + 
  labs(y = "Correlation: Accuracy/Convergence", 
       x = "Sample size", 
       color = "Competence distribution") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_brewer(palette = "Set1") +
  plot_theme + 
  theme(
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8))

# competence plot
competence_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_competence, color = competence)) +
  geom_point() + 
  geom_line(alpha = 0.5) + 
  labs(y = "Correlation: Competence/Convergence", 
       x = "Sample size", 
       color = "Competence distribution") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_brewer(palette = "Set1") +
  plot_theme + 
  theme(
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8))

ggarrange(accuracy_plot, competence_plot, common.legend = TRUE)
```

## Categorical choice context

```{r read-simulation-data-categorical, message=FALSE}
# read data from simulations
simulation_sample_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_sample_3_options.csv")

simulation_options_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_options_3_sample.csv")
```

When people make a choice based on several categories, their answers cannot be ranked by their nature (i.e. they are nominal, not ordinal), and that there are fewer of them (e.g. one of three possible products to choose, instead of an estimate between one and two thousand). In this case, convergence can be measured by the share of people agreeing on an answer. The larger the share of informants agreeing on an answer, the greater convergence. This measure is at the response level, nested within the group level.

As for the numeric scenario, we ran simulations to provide a normative answer as to whether it is justified to infer accuracy and competence from greater convergence. We, again, suppose that an individual's answer is drawn from an internal distribution - in this case, a multinomial distribution, that describes how likely the individual is to choose each available option. If there are m choice options, an individual has the probability p of picking the right one, and the probability of (1-p)/(m-1) to pick any other option. Each individual has their own multinomial distribution. We define competence as the probability of making the correct choice. The higher the competence, the greater the probability that an individual will choose the correct option. Competence values range from being at chance for selecting the correct option (p = 1/m) to certainly selecting the correct option (p = 1). As before in the numeric case, suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution (see fig. \@ref(fig:competence-distributions)). Based on their competence level, we draw an estimate for each individual. We measure an individual's accuracy as a binary outcome, namely whether they picked the correct option, or not. We then randomly assign individuals to groups of informants (we vary the sample size from one simulation to another). Within these groups, we calculate the share of individuals voting for an answer option. For example, in a scenario in which three individuals pick among three options (A, B and C), two individuals might vote C and one B. In this case we obtain an average accuracy and an average competence value for a share of 2/3 (option C) and for a share of 1/3 (option B). We simulate this on a population of 999000 individuals. We repeat this procedure varying the underlying population competence distributions, and additionally varying either (a) the sample size of informants, or (b) the number of choice options. If we vary the sample size, we hold the number of choice options constant at n = 3, and vice versa when varying the number of choice options. Fig. \@ref(fig:simulation-accuracy-sample-categorical) shows the average accuracy, and Fig. \@ref(fig:simulation-competence-sample-categorical) the average competence value for each share of votes, for different competence levels and varying the sample size. Fig. \@ref(fig:simulation-accuracy-options-categorical) and Fig. \@ref(fig:simulation-competence-options-categorical) display the same relationship, but varying the number of choice options instead. The figures dispaly that, across all sample sizes and competence levels, the larger the share of votes for an option, the more accurate the option is on average. That relationship appears to follow some sigmoid curve which switches from an average accuracy of 0 to an average accuracy of 1 before a share of 0.5 is attained, and which is steeper for larger sample sizes. For competence, we observe a similar sigmoid-like relationship, but of lesser amplitude and varying considerably as a function of the underlying population competence distributions.

(ref:simulation-accuracy-sample-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and sample sizes. Points represent averages across all simulations within the respective segment. The number of choice options is three.

```{r simulation-accuracy-sample-categorical, fig.cap="(ref:simulation-accuracy-sample-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Accuracy", variable = sample)
```

(ref:simulation-competence-sample-categorical) Competence as a function of vote share for an option, for different population competence distributions and sample sizes. Points represent averages across all simulations within the respective segment. The number of choice options is three.

```{r simulation-competence-sample-categorical, fig.cap="(ref:simulation-competence-sample-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Competence", variable = sample)
```

(ref:simulation-accuracy-options-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and number of choice options. Points represent averages across all simulations within the respective segment. The sample size is three.

```{r simulation-accuracy-options-categorical, fig.cap="(ref:simulation-accuracy-options-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_options_categorical,
                                       outcome = "Accuracy", variable = options)
```

(ref:simulation-competence-options-categorical) Competence as a function of vote share for an option, for different population competence distributions and number of choice options. Points represent averages across all simulations within the respective segment. The sample size is three.

```{r simulation-competence-options-categorical, fig.cap="(ref:simulation-competence-options-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_options_categorical,
                                       outcome = "Competence", variable = options)
```

In sum, given the set of specific assumptions we made, our simulations suggest that people are indeed justified in inferring accuracy and competence from convergence in both numeric and categorical choice settings.

(ref:simulation-accuracy-numeric) Simulation results showing the relationship between convergence and accuracy for different population competence distributions.

```{r simulation-accuracy-numeric-1, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)

plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[1]]
```

```{r simulation-accuracy-numeric-2, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[2]]
```

```{r simulation-accuracy-numeric-3, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[3]]
```

```{r simulation-accuracy-numeric-4, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[4]]
```

```{r simulation-accuracy-numeric-5, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[5]]
```

```{r simulation-accuracy-numeric-6, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[6]]
```

(ref:simulation-competence-numeric) Simulation results showing the relationship between convergence and competence for different population competence distributions.

```{r simulation-competence-numeric-1, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[1]]
```

```{r simulation-competence-numeric-2, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[2]]
```

```{r simulation-competence-numeric-3, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[3]]
```

```{r simulation-competence-numeric-4, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[4]]
```

```{r simulation-competence-numeric-5, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[5]]
```

```{r simulation-competence-numeric-6, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[6]]
```

### 2023-11-08 functions/own_functions.R

@@ -1,0 +1,317 @@
+# round numbers from models
+rounded_numbers <- function(x) mutate_if(x, is.numeric, round, 3)
+
+# get model outputs ready for inline text reporting
+text_ready <- function(model_output) {
+  
+  result <- tidy(model_output, conf.int = TRUE) %>% 
+    filter(effect == "fixed") %>% 
+    # report p.value according to apa standards
+    mutate(p.value = case_when(p.value < 0.001 ~ "< .001",
+                               TRUE ~ sprintf("p = %.3f", p.value)
+    )
+    ) %>% 
+    # all other terms
+    rounded_numbers() %>% 
+    mutate(
+      term = ifelse(term == "(Intercept)", "intercept", term),
+      ci = glue::glue("[{conf.low}, {conf.high}]"), 
+      # if there is an interaction (recognized by ":"), name it just interaction
+      term = str_replace(term, ".*:.*", "interaction")
+    ) %>% 
+    select(term, estimate, std.error, ci, p.value) %>% 
+    split(.$term)
+  
+  return(result)
+}
+
+# Plot results from simulation
+
+# Plot competence distributions
+plot_competence_distributions <- function(data) {
+  # Your data frame
+  d <- data %>% 
+    group_by(competence) %>% 
+    reframe(alpha = unique(alpha), beta = unique(beta))
+  
+  # Function to plot the different competence distributions 
+  plot_competence_distribution <- function(alpha, beta) {
+    ggplot() +
+      geom_function(
+        fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
+      ) +
+      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
+           y = "Density", x = "Competence") +
+      plot_theme
+  }
+  
+  # Specify the values of alpha and beta from the data
+  alpha_values <- d$alpha
+  beta_values <- d$beta 
+  
+  # Create a list of ggplot objects
+  plot_list <- map2(alpha_values, beta_values, plot_competence_distribution)
+  
+  # Arrange and display the plots
+  competence_plot <- ggarrange(plotlist = plot_list) %>%
+    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
+                                                    face = "bold",
+                                                    color = "darkgrey",
+                                                    size = 14))
+  
+  
+  return(competence_plot) 
+}
+
+# Categorical scenario
+
+# Vary competence
+# group by constellations
[…]

### 2024-01-18 Experiment_4/model.Rmd
version before more detailed text changes

@@
 # ggplot() +
-#   geom_function(fun = ~dbeta(.x, shape1 = 0.9, shape2 = 0.9))
+#   geom_function(fun = ~dbeta(.x, shape1 = 0.1, shape2 = 10))
 ```
@@
 ```
+
+### Explanation Hugo
+
+We take a very incompetent population
+
+```{r}
+# Tryout beta functions
+ggplot() +
+  geom_function(fun = ~dbeta(.x, shape1 = 0.1, shape2 = 5))
+```
+
+We assume many many choice options (approximating a numeric scenario). We look at a very large sample to (hopefully), sample a couple of really competent individuals.
+
+```{r}
+# test output (set return to data above)
+# test <- simulate_single_population(population = 10000, sample = 10000, options = 10000,
+#                                    return = "by degree of majority", distribution = "beta",
+#                                    alpha = 0.1, beta = 3) %>% arrange(sample_id) %>%
+#   arrange(sample_id)
+# 
+# test %>% group_by(relative_majority) %>% summarise(across(c(average_relative_competence,
+#                                                         average_accuracy),
+#                                                       mean)
+#                                                )
+```
+
+## next
+
+```{r}
+# read data for various sample sizes and 3 choice options
+test <- read_csv("data/sim_rel_majority_vary_competence_sample_3_options.csv")
+```
+
+```{r}
+z <- test %>% 
+  # filter to scenario with samples of 3
+  filter(sample == 3) %>% 
+  group_by(competence, relative_majority) %>% 
+  summarize(
+    mean_competence = mean(average_relative_competence),
+    mean_accuracy = mean(average_accuracy),
+    absolute_count = n()
+    ) %>% 
+  group_by(competence) %>% 
+  mutate(relative_share = absolute_count / sum(absolute_count),
+         bla = sum(absolute_count), 
+         relative_majority = round(relative_majority, digits = 2)
+         )
+
+ggplot(z, aes(x = as.factor(relative_majority), y = relative_share, fill = competence)) + 
+  geom_col(position = "dodge", width = 0.7, color = "white") +
+  labs(title = "Grouped Bar Graph",
+       x = "Relative Majority",
+       y = "Relative Share") +
+  scale_fill_viridis_d(option = "rocket") +
+  theme_minimal()
+
+z_step2 <- z %>% 
+  filter(relative_majority == 0.67)
+
+ggplot(z_step2, aes(x = as.factor(relative_majority), y = relative_share, fill = competence)) + 
+  geom_col(position = "dodge", width = 0.7, color = "white") +
+  labs(title = "Grouped Bar Graph",
+       x = "Relative Majority",
+       y = "Relative Share") +
+  scale_fill_viridis_d(option = "rocket") +
+  theme_minimal()
+
+z_step3 <- z_step2 %>% 
+         filter(competence == "very incompetent\n(alpha = 0.1, beta = 1)")
+
+ggplot(z_step3, 
+       aes(x = as.factor(relative_majority), y = mean_competence, fill = competence)) + 
+  geom_col(position = "dodge", width = 0.7, color = "white") +
+  geom_text(aes(label = round(mean_competence, 3)),
+            position = position_dodge(width = 0.7),
+            vjust = -0.5,    # Adjust vertical position of labels
+            size = 3.5,      # Adjust text size
+            color = "black") +
+  labs(title = "Competence for most likely distribution",
+       x = "Relative Majority",
+       y = "Average competence") +
+  scale_fill_viridis_d(option = "rocket") +
+  theme_minimal()
+
+View(z)
+```

### 2024-01-18 Pfander_DeCourson_Mercier_2023.Rmd
version before more detailed text changes

@@
 no priors).
-These inferences from convergence might help explain why people respect scientists' competence, even if they do not understand much about how scientific results are reached.
+These inferences from convergence might help explain why people respect scientists’ competence, even if they do not understand much about how scientific results are reached.
 
@@
 classoption       : "doc" # "doc" for nice look, "man" for manuscripty
-output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
+output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
 For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g.
@@
 appendix:
+  - "appendix_exp1.Rmd"
+  - "appendix_exp2.Rmd"
+  - "appendix_exp3.Rmd"
+
 bibliography: references.bib
@@
 library("viridis")     # for generating color palette for map 
+library("grid")        # for image plots   
+library("gridExtra")   # for image plots
+library("png")         # for image plots
 ```
@@
 Imagine that you live in ancient Greece, and a fellow called Eratostenes claims the circumference of the earth is 252000 stades (approximately 40000 kilometers).
-You know nothing about this man, and you don't understand trigonometry.
-Moreover, you know little about the circumference of the Earth, and even less of how one could measure such a thing.
-As a result, you discard his opinion and take him for a pretentious loon.
+You know nothing about this man, the circumference of the Earth, or how one could measure such a thing.
+As a result, you discard Eratostenes' opinion and (mis)take him for a pretentious loon.
 But what if other scholars had arrived at very similar measurements, independently of Eratosthenes?
 Or even if they had carefully checked his measurement, with a critical eye?
-Wouldn't that give you enough ground to believe not only that the estimates might be correct, but also that Eratosthenes and his fellow scholars must be quite bright, to be able to achieve such a feat?
-
-In this article, we explore how, under some circumstances, we infer that a group of people who provided the same, or very similar answers to a question are likely to be correct, and to be competent in the relevant area, even if we had no prior about either what the correct answer was, or about their competence.
-
-We first review experiments suggesting that people, including young children, appear to trust groups with more convergent opinions to be more accurate, although the evidence is more conclusive in numerical than in categorical choice contexts.
-We continue by explaining that two well-studied forms of the wisdom of crowds--averaging and the Condorcet Jury Theorem--do not provide normative grounds for this behavior.
-We then present the results of simulations, suggesting that under certain conditions, inferences from convergence to accuracy and to underlying competence are sound.
-Finally, we show that participants draw these inferences in a series of six experiments.
-
-## Numerical choices
-
-Informants can converge in numerical (e.g.
-how many calories are there in an apple?) and categorical (e.g.
-which option is correct - A or B?) choice contexts.
-In numerical contexts, convergence varies by the proximity of estimates; in categorical contexts, convergence varies by the ratio of people agreeing on a choice.
-We start by reviewing numerical choices.
-
-For numerical choice contexts, evidence that people take convergence into account can be found in advice-taking experiments.
-In these experiments, participants are called judges, and they need to make numerical estimates--sometimes on factual knowledge, e.g.
-'What year was the Suez Canal opened first?' [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g.
+Wouldn't that give you enough ground to believe not only that the estimates might be correct, but also that Eratosthenes and his fellow scholars must be quite bright, to be able to achieve such a feat as measuring the Earth?
+
+In this article, we explore how, under some circumstances, we should, and we do infer that a group of individuals whose answers converge are likely to be correct, and to be competent in the relevant area, even if we had no prior belief about either what the correct answer was, or about these individuals' competence.
+
+We first review experiments testing whether people infer that individuals whose answers converge are more likely to be right, and to be competent.
+We then argue that little attention has been paid to the normative models that might explain when and why such an inference from convergence to competence is warranted.
+We develop such models, both analytically and with simulations.
+Finally, we present a series of experiments in which participants evaluate the competence of individuals in groups as a function of how much they converge on a given answer, in the absence of any priors about these individuals' competence, or what the correct answer is.
+
+## Do people infer that individuals whose answers converge tend to be right, and to be competent?
+
+There are two ways for answers to converge: when the answers are continuous, they can have a smaller range or a lower variance; when the answers are categorical (i.e.
+there are only a few, often two, options), more can agree on the same answer.
+These two cases have been studied with different paradigms.
+
+In the continuous case, the most relevant evidence comes from the literature on \'advice-taking.\' In these experiments, participants are called \'judges,\' and they need to make numerical estimates--sometimes on factual knowledge, e.g. \'What year was the Suez Canal opened first?
+[@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g.
 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020].
-To help answer these questions, participants are given estimates from others, the advisors--sometimes presented as fellow participants, sometimes as experts.
-The overarching question in those studies is how people integrate social information from the advisors when making estimates.
+To help answer these questions, participants are given estimates from others, the advisors.
 
@@
 @jaylesHowSocialInformation2017a; @mannesAreWeWise2009], but no information about the distribution of initial estimates is given.
-By contrast, other advice-taking studies provide participants with a set of estimates.
+By contrast, advice-taking studies relevant to our argument provide participants with a set of estimates.
 One subset of these studies manipulate the degree of convergence between groups of advisors, e.g.
-via variance of estimates [@mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009], or their range [@budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007].
+via the variance of estimates [@mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009], or their range [@budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007].
 These studies find that participants are more confident about, or rely more on, estimates from groups of advisors who converge more.
-More indirect evidence comes from another subset of studies varied the degree of convergence within groups of advisors.
-They present participants with sets of estimates where some estimates are close to each other (or overlapping, in cases where estimates were intervals), while others were clearly outliers [@yanivWeightingTrimmingHeuristics1997, study 3 & 4; @harriesCombiningAdviceWeight2004].
+More indirect evidence comes from another subset of studies on outlier opinions.
+In these studies, participants were presented with sets of estimates where all but one estimate--the outlier--were close to each other [@harriesCombiningAdviceWeight2004], or overlapping, when estimates were intervals [@yanivWeightingTrimmingHeuristics1997, study 3 & 4].
+The outlier was defined, for example, as having "[...] a z-score greater than 2.0 and the remaining opinions had z-scores less than 2.0" [@harriesCombiningAdviceWeight2004].
 These studies find that participants heavily discount outliers when aggregating estimates.
 These results are thus consistent with the idea that people rely more on convergent opinions, although testing extreme cases of singular dissident opinions.
-Another study varied convergence between estimates from single advisors [@yanivWeightingTrimmingHeuristics1997, study 1 & 2].
-In this study, participants were shown ranges of estimates from two advisors - one with a small range, the other with a big range.
+Yet more indirect evidence comes from experiments that varied convergence of estimates from single advisors [@yanivWeightingTrimmingHeuristics1997, study 1 & 2].
+In this study, participants were shown ranges of estimates from two advisors - one with a small range (e.g.
+3 to 5), the other with a big range (e.g.
+1 to 10).
 Participants relied more on the precise, small range advice.
 These results can be interpreted as a case of inferring accuracy from (internal) convergence of single informants, perceived by participants perhaps as consistency or for example confidence.
-However, in this study, the small-range advisor (e.g.
-3 to 5) was always within the range of the big-range adivsor (e.g.
-1 to 10).
-If people relied more on the small-range advisor, then it was likely not only because of greater convergence, but also because of the overlap between the two advisors.
+However, in this study, the small-range advisor was always within the range of the big-range adivsor.
+Consequently, if people relied more on the small-range advisor, then it was likely not only because of greater convergence, but also because of the overlap between the two advisors.
 
@@
 
-In the beginning of the 20th century, Francis Galton famously demonstrated accuracy gains from averaging: in his “vox populi” experiment, he asked 787 participants to estimate the weight of an ox, and found the average estimate of an ox weight to be just one pound from reality [@galtonVoxPopuli1907].
+In the beginning of the 20th century, Francis Galton famously demonstrated accuracy gains from averaging: in his "vox populi" experiment, he asked 787 participants to estimate the weight of an ox, and found the average estimate of an ox weight to be just one pound from reality [@galtonVoxPopuli1907].
 Since then, it has been shown both theoretically and empirically that the mean or median answer of a group is typically closer to the truth than the mean individual is [see e.g.
@@
 
-As Eratosthenes’ example above illustrates, intuitively it seems there are cases in which we can infer that the members of a group are more likely to be correct, or, equivalently, more competent, when their answers converge.
+As Eratosthenes' example above illustrates, intuitively it seems there are cases in which we can infer that the members of a group are more likely to be correct, or, equivalently, more competent, when their answers converge.
 When are we justified to infer that people are right (and thus competent) when they converge?
@@
 Each of the wrong answers has the same probability (1-p)/m to be chosen.
-We observe the answers of a population of n+1 agents, with diverse levels of competence drawn from a beta distribution – a flexible probability distribution that can be uniform, unimodal or bimodal, depending on the parameters.
+We observe the answers of a population of n+1 agents, with diverse levels of competence drawn from a beta distribution -- a flexible probability distribution that can be uniform, unimodal or bimodal, depending on the parameters.
 We assume that the shape parameters of the beta distribution ($\alpha$ and $\beta$) are known, but the right answer is not.
@@
 All analyses were conducted in R (version 4.2.2) using R Studio.
-
-# Numerical estimates: Experiments 1 to 3
 
@@
 
+# Hypotheses
+
 # run mixed model (random slope and intercept) with participants as random factor
 exp1_model_accuracy = lmer(confidence ~ convergence + number + (1 + convergence | ID), data = exp1)
-exp1_model_competence = lmer(confidence ~ convergence + number + (1 + convergence | ID), data = exp1)
+exp1_model_competence = lmer(competence ~ convergence + number + (1 + convergence | ID), data = exp1)
 
@@
 )
-```
+
+# Research questions
+
+# RQ2 (as separate t-test):
+accuracy <- t.test(confidence ~ number, data = exp1, paired = TRUE)
+competence <- t.test(competence ~ number, data = exp1, paired = TRUE)
+
+# RQ2 and RQ3 (as single mixed model):
+
+# effect code variables
+exp1 <- exp1 %>% 
+  mutate(convergence_effect_code = recode(convergence, "divergent" = -0.5, "convergent" = +0.5),
+         number_effect_code = recode(number, "small" = -0.5, "large" = +0.5))
+
+model_interaction_competence = lmer(competence ~ convergence_effect_code + number_effect_code +
+                                      convergence_effect_code*number_effect_code + 
+                                      (1 + convergence_effect_code | ID), data = exp1)
+
+model_interaction_accuracy = lmer(confidence ~ convergence_effect_code + number_effect_code +
+                                    convergence_effect_code*number_effect_code + 
+                                    (1 + convergence_effect_code | ID), data = exp1)
+
+# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? 
+exp1_explorative <- exp1 %>% 
+  pivot_longer(
+    c(confidence, competence),
+    names_to = "outcome", 
+    values_to = "score") %>% 
+  # effect code variables
+  mutate(outcome_effect_code = recode(outcome, "confidence" = -0.5, "competence" = +0.5)
+         )
+
+# run mixed model (random slope and intercept) with participants as random factor
+exploratory_analysis <- lmer(score ~ convergence_effect_code + outcome + 
+                               convergence_effect_code*outcome + 
+                               (1 + convergence_effect_code + outcome | ID), 
+                             data = exp1_explorative)
+
+exp1_research_questions <- list(
+  # speparate t-tests for RQ2
+  means = exp1 %>% 
+    group_by(number) %>% 
+    summarise(across(c(confidence, competence), 
+                     mean)) %>% 
+    rounded_numbers() %>% 
+    split(.$number), 
+  accuracy = clean_t_test(accuracy), 
+  competence = clean_t_test(competence),
+  # RQ2 and RQ3 in a single mixed model
+  model_interaction_competence = text_ready(model_interaction_competence), 
+  model_interaction_accuracy = text_ready(model_interaction_accuracy),
+  # Exploratory analysis
+  exploratory_analysis = text_ready(exploratory_analysis)
+)
+```
+
+## RQ3 - Interaction between number and convergence on competence
 
@@
 ***H2: Participants perceive (independent) informants whose estimates converge more as more competent than informants whose estimates diverge.***
+
+We had three additional research questions about the number of informants:
+
+***RQ1: Do H1 and H2 hold for both a small [3] and a large [10] number of estimates?***
+
+***RQ2: When making a guess based on the opinions of (independent) informants, will participants be more confident about their guess when there is a larger number of estimates compared to when this number is smaller?***
+
+***RQ3: Is there an interaction effect between the number of estimates and convergence on perceived competence of informants?***
 
@@
 
-(ref:stimulus-exp1) Results of one of eight games that participants have to rate. The stimulus corresponds to a divergent, 10 players condition
+(ref:stimulus-exp1) Results of one of eight games that participants have to rate. The stimulus corresponds to a convergent, 10 players condition
 
@@
 
-#### Stimuli
+#### Materials
 
@@
 First, we varied the standard deviation of these distributions to simulate the degree of convergence (150 for divergence, 20 for convergence; estimate scale ranged from 1000 to 2000).
-Second, we varied the number of draws from these distributions per set (three for low number, 10 for high number).
-For each of the four possible resulting condition, we generated two random draws from the respective normal distribution.
-
-The means of the normal distributions that we draw our estimates from were distinct between sets of estimates.
-Considering our within-participant design, we wanted to ensure that participants understood each set of estimates as being the result of a different, unrelated game.
-In order to assure that random draws from the distributions will (most likely) appear on the response scale (1000 to 2000), we constrained the means of all normal distributions to lie between the first and third quartile of the response scale (i.e.
-smallest possible mean was 1225 and largest 1775).
-We define a set of eight means--one for each set of estimates--that cover the range from first to third quartile of the predefined scale with an equal interval (1250, 1325, 1400, 1475, 1550, 1625, 1700, 1775).
-We randomly paired means with conditions when generating the stimuli.
-We then drew the set of estimates from the respective normal distributions given the assigned means and the condition constraints.
-We repeated this three times, resulting in three different series of eight sets of estimates.
-We randomly assign participants to one of these series.
-Additionally, for each participant, we randomize the order of appearance of the sets of estimates within the respective series.
-Images of all sets of estimates can be found on the OSF.
+Second, we varied the number of draws (either three or 10) from these distributions.
+For each of the four possible resulting conditions, we generated two random draws from the respective normal distribution.
+More information on how the stimuli were created can be found in Appendix \@ref(exp1).
 
@@
 
-### Results
-
-To account from dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slope for participants, using the `lme4` package and its `lmer()` function in R.
-In the models, we control for the number of estimates (our second experimental factor).
-Figure \@ref(fig:exp1-plot) visualizes the results and table \@ref(tab:exp1-table) contains descriptive results.
+### Results and discussion
+
+To account from dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slopes of convergence for participants, using the `lme4` package and its `lmer()` function in R.
+In the models for our hypotheses, we control for the number of estimates (our second experimental factor).
+Visualizations and descriptive statistics can be found in Appendix \@ref(exp1).
 We find a positive effect of convergence on accuracy: Participants were more confident about their estimate in convergent scenarios ($\Delta$ Accuracy = `r exp1_descriptives$model_accuracy$convergenceconvergent$estimate` `r exp1_descriptives$model_accuracy$convergenceconvergent$ci`, p = `r exp1_descriptives$model_accuracy$convergenceconvergent$p.value`) than in divergent ones (Accuracy = `r exp1_descriptives$model_accuracy$intercept$estimate` `r exp1_descriptives$model_accuracy$intercept$ci`, p = `r exp1_descriptives$model_accuracy$intercept$p.value`).
@@
 
-```{r exp1-table}
-exp1 %>% 
-  group_by(convergence) %>% 
-  summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Accuracy = paste0(confidence_mean, " (sd = ", confidence_sd, ")"),
-         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")")
-         ) %>% 
-  rename(Convergence = convergence) %>% 
-  select(Convergence, Accuracy, Competence) %>% 
-  apa_table()
-```
-
-(ref:exp1-plot) Distributions of accuracy and competence by level of convergence.
-
-```{r exp1-plot, fig.cap="(ref:exp1-plot)"}
-# make mean data for plot
-x_nudge <- 0.8
-
-means <- exp1 %>%
-  group_by(convergence) %>%
-  summarize(confidence = mean(confidence), 
-            competence = mean(competence)) %>% 
-  pivot_longer(c(confidence, competence), 
-               names_to = "outcome", 
-               values_to = "value") %>% 
-  mutate(x_position = ifelse(convergence == "convergent", value + x_nudge, 
-                                value - x_nudge))
-
-confidence_plot <- ggplot(data=exp1, aes(x=confidence, fill=convergence)) +
-  # density plot
-  geom_density(adjust=3, alpha=.6) +
-  # mean lines
-  geom_vline(data=means %>% 
-               filter(outcome == "confidence"), 
-             aes(xintercept = value, 
-                 color = convergence),
-             show.legend = FALSE,
-             linetype="dashed") +
-  geom_label(data = means %>% 
-               filter(outcome == "confidence"), 
-             aes(x = x_position, y = 0.3, 
-                 label =  paste("mean", round(value, digits = 2),
-                                sep = " = ")), 
-             alpha = 0.6, show.legend = FALSE, size = 3
-            ) +
-  # scales
-  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
-                     limits=c(1, 7)) +
-  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
-  scale_fill_viridis_d(option = "plasma") +
-  scale_color_viridis_d(option = "plasma") + 
-  guides(fill = guide_legend(title = NULL)) +
-  plot_theme
-
-competence_plot <- ggplot(data=exp1, aes(x=competence, fill=convergence)) +
-  # density plot
-  geom_density(adjust=3, alpha=.6) +
-  # mean lines
-  geom_vline(data=means %>% 
-               filter(outcome == "competence"), 
-             aes(xintercept = value, 
-                 color = convergence),
-             show.legend = FALSE,
-             linetype="dashed") +
-  geom_label(data = means %>% 
-               filter(outcome == "competence"), 
-             aes(x = x_position, y = 0.3, 
-                 label =  paste("mean", round(value, digits = 2),
-                                sep = " = ")), 
-             alpha = 0.6, show.legend = FALSE, size = 3
-            ) +
-  # scales
-  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
-                     limits=c(1, 7)) +
-  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
-  scale_fill_viridis_d(option = "plasma") +
-  scale_color_viridis_d(option = "plasma") + 
-  guides(fill = guide_legend(title = NULL)) +
-  plot_theme
-
-figure <- ggarrange(confidence_plot, 
-                    competence_plot + 
-                      theme(axis.text.y = element_blank(),
-                            axis.ticks.y = element_blank(),
-                            axis.title.y = element_blank()), 
-                    common.legend = T) +
-  theme(legend.position="top") 
-figure
-```
+To test our research questions, we ran the same mixed-models but this time including an interaction between number of informants and convergence.
+We use effect-coded versions of our variables, allowing us to detect main effects for each factor.
+As for RQ2, pooling across convergent and divergent conditions, we find a main effect of number, such that participants had more confidence in their estimate when they relied on ten informants compared to three informants (beta = `r exp1_research_questions$model_interaction_accuracy$number_effect_code$estimate` `r exp1_research_questions$model_interaction_accuracy$number_effect_code$ci`, p = `r exp1_research_questions$model_interaction_accuracy$number_effect_code$p.value`).
+As for RQ3, we find an interaction between number and convergence regarding competence: participants perceived ten informants as slightly more competent than three informants in convergent scenarios, but not in divergent ones (beta = `r exp1_research_questions$model_interaction_competence$interaction$estimate` `r exp1_research_questions$model_interaction_competence$interaction$ci`, p = `r exp1_research_questions$model_interaction_competence$interaction$p.value`).
+There was no interaction regarding accuracy (beta = `r exp1_research_questions$model_interaction_accuracy$interaction$estimate` `r exp1_research_questions$model_interaction_accuracy$interaction$ci`, p = `r exp1_research_questions$model_interaction_accuracy$interaction$p.value`)
+
+In an exploratory, not pre-registered analysis, we tested whether the effect of convergence is bigger on accuracy than on competence.
+To do so, we regressed the outcome score on convergence and it's interaction with a binary variable indicating which outcome was asked for (accuracy or competence).
+We find that pooled across divergent and convergent conditions, participants rated participants reported lower perceived accuracy than competence (beta = `r exp1_research_questions$exploratory_analysis$outcomeconfidence$estimate` `r exp1_research_questions$exploratory_analysis$outcomeconfidence$ci`, `r exp1_research_questions$exploratory_analysis$outcomeconfidence$p.value`).
+However, we do not find an interaction effect indicating a difference of this effect between convergent and divergent conditions (beta = `r exp1_research_questions$exploratory_analysis$interaction$estimate` `r exp1_research_questions$exploratory_analysis$interaction$ci`, `r exp1_research_questions$exploratory_analysis$interaction$p.value`).
+
+In summary, as predicted, participants were more confident when the estimates were more convergent, which indicates they believed the estimates to have been more accurate, and they thought the individuals who had made more convergent estimates were more competent.
 
@@
 
-In experiment 1, informants were always independent.
-In experiment 2, we set out to test whether informational dependency alters inferences from convergence.
-We used the same setting as in experiment one, in which participants saw the fictional results of an estimation game.
-However, this time participants only saw convergent sets of estimates.
-We manipulated informational dependency: players were independent either independent or could "discuss at lenght" before making an estimate.
-Our hypotheses were:
+It is rational to infer that convergent estimates are more likely to be accurate, and to have been made by competent individuals, only if the convergence is best explained by the accuracy of the estimates.
+However, convergence could also be the outcome of other factors.
+If the individuals do not make their estimates independently of each other, a single individual might exert a strong influence on the others, making their convergence a poor cue to their accuracy.
+Alternatively, all individuals might have an incentive to provide a similar, but not accurate answer.
+In Experiment 2, we investigate the first possibility, and the second in Experiment 3.
+In particular, for Experiment 2 we rely on past results showing that participants, under some circumstances, put less weight on opinions that have been formed through discussion, by contrast with more independent opinions [@harkinsInformationUtilityMultiple1987; see also @lopesSocialValidationEveryday2007; @hessPsychologicalAdaptationsAssessing2006; @einavThinkingThemselvesEffect2018].
+We sought to replicate this finding in the context of convergent estimates, formulating the following hypotheses:
 
@@
 
-We manipulated informational dependency. According to the condition, players read a different introduction before seeing a set of estimates (see Table \@ref(tab:exp2-conditions)).
-
-```{r exp2-conditions}
-# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
-data.frame(
-  Condition = c("Independence", "Discussion"),
-  Description = c(
-    "Players are asked to make completely independent decisions – they cannot see each other’s estimates, or talk with each other before giving their estimates.",
-    "Players are asked to talk with each other about the game at length before giving their estimates."
-  )
-) #%>% 
-  # apa_table(align = c("m{3cm}", "m{12cm}")) 
-```
+In a within-participants design, participants saw both an independence condition, in which they were told "Players are asked to make completely independent decisions -- they cannot see each other's estimates, or talk with each other before giving their estimates," and a dependence condition, in which they were told "Players are asked to talk with each other about the game at length before giving their estimates."
 
@@
 
-We a subset of the stimuli we had generated for experiment one, namely convergent ones with three estimates.
+We used the materials generated for the convergent condition of Experiment 1.
 By contrast to experiment one, participants saw only two stimuli in total (one set of estimates per condition).
@@
 
-### Results
+### Results and discussion
 
 To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept for participants, using the `lme4` package and its `lmer()` function in R.
-Figure \@ref(fig:exp2-plot) visualizes the results and table \@ref(tab:exp2-table) contains descriptive results.
+Visualizations and descriptive statistics can be found in Appendix \@ref(exp2).
 The data does not support our hypotheses.
@@
 
-```{r exp2-table}
-exp2 %>% 
-  group_by(independence) %>% 
-  summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Accuracy = paste0(confidence_mean, " (sd = ", confidence_sd, ")"),
-         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")")
-         ) %>% 
-  rename(Independence = independence) %>% 
-  select(Independence, Accuracy, Competence) %>% 
-  apa_table()
-```
-
-(ref:exp2-plot) Distributions of accuracy and competence by level of informational dependency.
-
-```{r exp2-plot, fig.cap="(ref:exp2-plot)"}
-# make mean data for plot
-x_nudge <- 0.8
-
-means <- exp2 %>%
-  group_by(independence) %>%
-  summarize(confidence = mean(confidence), 
-            competence = mean(competence)) %>% 
-  pivot_longer(c(confidence, competence), 
-               names_to = "outcome", 
-               values_to = "value") %>% 
-  mutate(x_position = ifelse(independence == "dependent", value + x_nudge, 
-                                value - x_nudge))
-
-confidence_plot <- ggplot(data=exp2, aes(x=confidence, fill=independence)) +
-  # density plot
-  geom_density(adjust=3, alpha=.6) +
- # mean lines
-  geom_vline(data=means %>% 
-               filter(outcome == "confidence"), 
-             aes(xintercept = value, 
-                 color = independence),
-             show.legend = FALSE,
-             linetype="dashed") +
-  geom_label(data = means %>% 
-               filter(outcome == "confidence"), 
-             aes(x = x_position, y = 0.3, 
-                 label =  paste("mean", round(value, digits = 2),
-                                sep = " = ")), 
-             alpha = 0.6, show.legend = FALSE, size = 3
-            ) +
-  # scales
-  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
-                     limits=c(1, 7)) +
-  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
-  scale_fill_brewer(palette = "Accent",
-                    direction = 1) +
-  scale_color_brewer(palette = "Accent",
-                    direction = 1) +
-  guides(fill = guide_legend(title = NULL)) +
-  plot_theme
-
-competence_plot <- ggplot(data=exp2, aes(x=competence, fill=independence)) +
-  # density plot
-  geom_density(adjust=3, alpha=.6) +
- # mean lines
-  geom_vline(data=means %>% 
-               filter(outcome == "competence"), 
-             aes(xintercept = value, 
-                 color = independence),
-             show.legend = FALSE,
-             linetype="dashed") +
-  geom_label(data = means %>% 
-               filter(outcome == "competence"), 
-             aes(x = x_position, y = 0.3, 
-                 label =  paste("mean", round(value, digits = 2),
-                                sep = " = ")), 
-             alpha = 0.6, show.legend = FALSE, size = 3
-            ) +
-  # scales
-  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
-                     limits=c(1, 7)) +
-  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
-  scale_fill_brewer(palette = "Accent",
-                    direction = 1) +
-  scale_color_brewer(palette = "Accent",
-                    direction = 1) +
-  guides(fill = guide_legend(title = NULL)) +
-  plot_theme
-
-figure <- ggarrange(confidence_plot, 
-                    competence_plot + 
-                      theme(axis.text.y = element_blank(),
-                            axis.ticks.y = element_blank(),
-                            axis.title.y = element_blank()), 
-                    common.legend = T) +
-  theme(legend.position="top") 
-figure
-```
+Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently, to be less accurate, or made by less competent individuals.
+This suggests that participants might not be sensitive to at least some forms of informational dependency between the individuals making the estimates.
+Although previous studies have found that participants sometimes discount the opinions of groups compared to those of independent individuals, the superior performance of groups over individuals for a range of tasks [for review, see, e.g., @mercierArgumentativeTheoryPredictions2016a], including numerical estimates [e.g.
+@mercierDoesDiscussionMake2022] suggests that this discounting might have been misguided.
+As a result, the participants in the current experiment might have been behaving rationally when they did not discount the estimates made after discussion.
 
@@
 
-The aim of experiment 3 was twofold: First, we wanted to replicate the results of experiment 1 in a slightly more ecological scenario.
-In experiment 3, participants were told they saw (fictional) predictions of experts for stock values.
-Since we expected the findings of experiment one to replicate, our first set of hypotheses was:
+Experiment 3 tests whether participants are sensitive to another potential source of dependency between convergent estimates: when the individuals making the estimate share an incentive to bias their estimates in a given direction, independently of its accuracy.
+Even though it is formally similar to Experiment 1, the setting is different, as participants were told that they would be looking at (fictional) predictions of experts for stock values, instead of the answers of individuals in abstract games.
+In the conflict of interest condition, the experts had an incentive to value the stock in a given way, while they had no such conflict of interest in the independence condition.
+Given the nature of this manipulation, participants might have discounted the opinions of the conflicted experts, irrespective of the degree of convergence of their estimates.
+As a result, we could not directly compare the participants' answers across conditions.
+Instead, we must test whether the effect of greater convergence is reduced when the individuals making the estimates are systematically biased, compared to when they are not.
+On this basis, we formulate four hypotheses, two which are identical to those of Experiment 1, and only apply in the independent condition, and two that bear on the comparison between the conditions.
 
@@
 ***H1b: Participants perceive independent informants as more competent when their predictions converge compared to when they diverge.***
-
-Second, as in experiment two, we wanted to test the independence assumption -- but this time, against a different case of informational dependency, namely when informants shared a conflict of interest.
-In this condition, experts were described to gain from estimating the value of a stock at a certain price -- and that price always happened to be the mean of the the expert's estimates.
-Our second set of hypotheses was:
 
@@
 The simulation code is available on OSF, and the procedure is described in the pre-registration document.
-The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% (alpha for significance = 0.05) of the time.
-Due to uncertainty about the assumptions we made and because we had resources for a larger sample, we recruited `r exp3_descriptives$n_subj` participants for this study -- again, from the UK and via Prolific.
+The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% of the time, given an alpha threshold for significance of 0.05.
+Due to uncertainty about our effect size assumptions and because we had resources for a larger sample, we recruited `r exp3_descriptives$n_subj` participants for this study -- again, from the UK and via Prolific.
 Again, not a single participant failed our attention check.
@@
 As it so happens, in all scenarios, the predictions for the value of the stock have to lie between 1000 and 2000.
-Other than that, the scenarios are completely unrelated: it is different experts predicting the values of different stocks every time." Participants then saw the four scenarios, each introduced by a text according to which condition the participant was assigned to (see Table \@ref(tab:conditions-exp3); for more detailed materials see pre-registration document on OSF).
-
-```{r conditions-exp3}
-# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
-data.frame(
-  Condition = c("Independence", "Conflict of interest"),
-  Description = c(
-    "Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.",
-    "All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future."
-  )
-) #%>% 
-  # apa_table(align = c("m{3cm}", "m{12cm}"))  
-```
+Other than that, the scenarios are completely unrelated: it is different experts predicting the values of different stocks every time." Participants then saw the four scenarios, each introduced by a text according to which condition the participant was assigned to.
+To remove any potential ambiguity about participants' inferences on the accuracy of the estimates, we replaced the question about confidence to one bearing directly on accuracy: "On average, how accurate do you think these three predictions are?" on a 7-point Likert scale ("not accurate at all" to "extremely accurate").
+The question about competence read: "On average, how good do you think these three experts are at predicting the value of stocks?", also assessed on a 7-point Likert scale (from "not good at all" to "extremely good").
 
@@
 
-We manipulated two factors: informational dependency (two levels, independence and conflict of interest, see Table \@ref(tab:conditions); between participants) and convergence (two levels, convergence and divergence; within participants).
-
-#### Stimuli
-
-Our design required that each participant sees four different sets of predictions--two convergent and two divergent ones.
-We generated new stimuli: As in experiment 1, estimates would appear on a scale from 1000 to 2000.
-By contrast with experiment 1, we generated the sets of estimates with random draws from uniform distributions (instead of normal distributions).
-We varied the range of these distributions according to convergence (60 for convergence, 600 for divergence).
-Switching from normal distributions to uniform distributions made 'unlucky' draws in which conditions are visually not distinguishable less likely.
-
-Similar to the previous experiments, we also vary the value on the prediction scale (from 1000 to 2000) across which the range is centered.
-Considering our within-participant design, this, again, made it seem more likely that participants understand each set of predictions as being the result of a different stock, with a different true value.
-In order to assure all random draws from the distributions would appear on the response scale, we constrained the center of the uniform distributions to lie between 1300 and 1700.
-We define four center values -- one per set of predictions -- that divide this interval in (rounded) quartiles (1300, 1430, 1570, 1700).
-Given a center and a range, we then draw the predictions from uniform distributions.
-For example, in a draw for a divergent set of estimates with a center of 1700, each value within a range of 1400 and 2000 is equally likely to get selected.
-To avoid that single draws overlap too much within the same set, we defined a minimum space of 5 between the three predictions of a set.
-
-To minimize confounding of convergence with a certain placement of the center value, we paired center values with conditions such that each condition appears once in each half of the scale and each condition gets one of the extreme values.
-For example, in one set the convergence condition would get assigned center values of 1300 and 1570, the divergent condition center values of 1430 and 1700).
-We generated four such series (all possible combinations) and randomly assigned participants to one of them.
-Additionally, for each participant, we randomize the order of appearance of the sets of predictions within the respective series.
-
-For each set of predictions, we calculated the empirical mean based on the randomly drawn estimates.
-In the conflict of interest condition, this mean was inserted as the value that participants were be told experts would gain from.
-Consequently, the convergent predictions converge around what is said to be the incentivized value for the experts to choose.
-
-#### Dependent variables
-
-By contrast to the previous experiments, we operationalized accuracy as an actual question on accuracy this time.
-For each set of estimates participants responded to the following questions: First, "On average, how accurate do you think these three predictions are?" on a 7-point Likert scale ("not accurate at all" to "extremely accurate"); second, "On average, how good do you think these three experts are at predicting the value of stocks?", also assessed on a 7-point Likert scale (from "not good at all" to "extremely good").
-
-### Results
+We manipulated two factors: informational dependency (two levels, independence and conflict of interest; between participants) and convergence (two levels, convergence and divergence; within participants).
+In the independence condition, the participants read "Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock." In the conflict of interest condition, the participants read "All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future." The distributions presented were similar to those of Experiment 1, although generated in a slightly different manner (see Appendix \@ref(exp3)).
+
+### Results and discussion
 
 To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants, using the `lme4` package and its `lmer()` function in R.
-Figure \@ref(fig:exp3-plot) visualizes the results and Table \@ref(tab:exp3-table) contains descriptive results.
+Figure \@ref(fig:exp3-plot) visualizes the results.
 
@@
 Likewise the effect of convergence on competence is more positive ($\Delta$ Convergence = `r exp3_descriptives$model_competence$interaction$estimate` `r exp3_descriptives$model_competence$interaction$ci`, p = `r exp3_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition (Convergence = `r exp3_descriptives$model_competence$convergenceconvergent$estimate` `r exp3_descriptives$model_competence$convergenceconvergent$ci`, p = `r exp3_descriptives$model_competence$convergenceconvergent$p.value`).
-
-```{r exp3-table}
-exp3_accuracy <- exp3 %>% 
-  group_by(independence, convergence) %>% 
-  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
-  select(independence, convergence, Accuracy) %>% 
-    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
-                "Convergent")) %>%  
-  pivot_wider(names_from = convergence,
-              values_from = Accuracy) %>% 
-  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
-                "Independent") 
-         ) %>% 
-  rename(" " = independence) %>% 
-  as.matrix()
-
-exp3_competence <- exp3 %>% 
-  group_by(independence, convergence) %>% 
-  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
-  select(independence, convergence, Competence) %>% 
-    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
-                "Convergent")) %>%  
-  pivot_wider(names_from = convergence,
-              values_from = Competence) %>% 
-  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
-                "Independent") 
-         ) %>% 
-  ungroup() %>% 
-  select(-independence) %>% 
-  as.matrix()
-
-table <- cbind(exp3_accuracy, exp3_competence)
-
-table %>%
-  kable(booktabs = TRUE, 
-        # need to add a (empty) caption for cross referencing to work
-        caption = "") %>%
-  kable_styling() %>% 
-  add_header_above(c("", "Accuracy" = 2, "Competence" = 2))
-
-```
 
@@
 
-# Categorical estimates: Experiments 4 to 6
+Experiment 3 shows that, when the individuals making the estimates are systematically biased, then participants put less weight on the convergence of their estimates to infer that the estimates are accurate, and that the individuals making them are competent.
+
+*Cut this transition out probably*: \_\_\_\_\_\_\_ \# Categorical estimates: Experiments 4 to 6
 
@@
 Experiment six tests a new context factor: the number of choice options.
+\_\_\_\_\_\_
 
@@
 
-In line with our findings in numerical choice settings, we predicted that:
+In a second series of experiments, we test similar predictions to those of the previous experiments, but in a categorical choice context.
+The set-up is similar to that of Experiment 1, except that the outcomes seen by the participants are not numerical estimates, but choices made between a few options.
+In Experiment 4 tests hypotheses that are analogous to those of Experiment 1:
 
@@
 We ran a power simulation to inform our choice of sample size.
-All assumptions and details on the procedure can be found in the `power_Exp4.Rmd` document.
+All assumptions and details on the procedure can be found on the OSF.
 We ran two different power analyses, one for each outcome variable.
@@
 The simulation for `competence` suggested that we achieve statistical power of at least 90% with a sample size of `30`.
-Due to uncertainty about our assumptions, we recruited `r exp4_descriptives$n_subj` participants.
+Due to uncertainty about our assumptions and because it was within our budget, we recruited `r exp4_descriptives$n_subj` participants.
 
@@
 \@ref(fig:stimulus-exp4)).
-
-(ref:stimulus-exp4) An example of a stimulus for the majority condition
+To assess perceived accuracy, we asked: "What do you think is the probability of player 1 being correct?".
+Participants answered with a slider on a scale from 0 to 100.
+To assess perceived competence, we asked participants: "How competent do you think player 1 is in games like these?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
+
+(ref:stimulus-exp4) One set of stimuli by level of convergence.
+In the study, we used as second set of stimuli where each constellation was mirrored.
+A full set of stimuli can be found in Appendix \@ref(exp4).
 
 ```{r stimulus-exp4, echo=FALSE, fig.cap="(ref:stimulus-exp4)"}
-knitr::include_graphics("figures/example_stimulus_exp4.png")
+# Function to create a plot with image and title
+create_image_plot <- function(image_path, title) {
+  img <- rasterGrob(readPNG(image_path), interpolate = TRUE)
+  title_grob <- textGrob(title, gp = gpar(fontsize = 8, fontface = "bold"))
+
+  # Combine image and title in a single plot
+  plot <- ggplot() +
+    annotation_custom(img, xmin = 0, xmax = 1, ymin = 0, ymax = 1) +
+    annotation_custom(title_grob, xmin = 0, xmax = 1, ymin = 1, ymax = 1) +
+    theme_void()  # Set the panel background to be transparent
+
+  return(plot)
+}
+
+# File paths to four images
+image_paths <- c(
+  "Experiment_4/figures/stimuli/opp_majority_a.png",
+  "Experiment_4/figures/stimuli/divergence_a.png",
+  "Experiment_4/figures/stimuli/majority_a.png",
+  "Experiment_4/figures/stimuli/consensus_a.png"
+)
+
+# Titles for the images
+image_titles <- c(
+  "minority (0)", 
+  "dissensus (1)", 
+  "majority (2)", 
+  "consensus (3)"
+)
+
+# Create a list of plots
+plots <- lapply(seq_along(image_paths), function(i) {
+  create_image_plot(image_paths[i], image_titles[i])
+})
+
+# Arrange the plots in a grid
+grid.arrange(grobs = plots, ncol = 2)
 ```
@@
 
-#### Stimuli
-
-We manipulate convergence within participants.
-All participants see all four conditions, with two stimuli per condition.
-Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli).
-
-#### Dependent variables
-
-To assess accuracy, we asked: "What do you think is the probability of player 1 being correct?".
-Participants answered with a slider on a scale from 0 to 100.
-As for competence, we asked participants: "How competent do you think player 1 is in games like these?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
-
-### Results
+Convergence was manipulated within participants.
+All participants saw all four conditions, with two stimuli per condition.
+Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).
+
+### Results and discussion
 
@@
 
-As in the numerical setting, we found a positive effect of convergence on both accuracy (Accuracy = `r exp4_descriptives$model_accuracy$convergence$estimate` `r exp4_descriptives$model_accuracy$convergence$ci`, p = `r exp4_descriptives$model_accuracy$convergence$p.value`) and competence (Competence = `r exp4_descriptives$model_competencey$convergence$estimate` `r exp4_descriptives$model_competence$convergence$ci`, p = `r exp4_descriptives$model_competence$convergence$p.value`).
-
-```{r exp4-table}
-exp4 %>% 
-  group_by(convergence_categorical, convergence) %>% 
-  summarise(across(c(competence, accuracy), list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ")"),
-         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")"),
-         Convergence = paste0(convergence_categorical, " (", convergence, ")")
-         ) %>% 
-  ungroup() %>% 
-  select(Convergence, Accuracy, Competence) %>% 
-  apa_table()
-```
+As in the numerical setting, we found a positive effect of convergence on both accuracy (Accuracy = `r exp4_descriptives$model_accuracy$convergence$estimate` `r exp4_descriptives$model_accuracy$convergence$ci`, p = `r exp4_descriptives$model_accuracy$convergence$p.value`; on a scale from 0 to 100) and competence (Competence = `r exp4_descriptives$model_competence$convergence$estimate` `r exp4_descriptives$model_competence$convergence$ci`, p = `r exp4_descriptives$model_competence$convergence$p.value`; on a scale from 1 to 7).
 
@@
 
-In line with our findings in numerical choice settings, we predicted that:
+Experiment 5 is a conceptual replication of Experiment 3 in the categorical instead of numerical case: are participants less likely to infer that more convergent estimates are more accurate, and the individuals who made them more competent, when the estimates are made by individuals with a conflict of interest pushing them to all provide a given answer, compared to when they are made by independent participants?
+The independence condition of Experiment 5 also serves as a replication of Experiment 4, leading to the following hypotheses:
 
@@
 We ran a power simulation to inform our choice of sample size.
-All assumptions and details on the procedure can be found in the `power_Exp5.Rmd` document.
+All assumptions and details on the procedure can be found on the OSF.
 We ran two different power analyses, one for each outcome variable.
-We set the power threshold for our experiment to 90%.
+We set the power threshold for both to 90%.
 
@@
 Some tasks might be difficult while others are easy.
-Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are." They were then presented to the results of eight such games and had to answer questions (see Fig.
-\@ref(fig:stimulus-exp4)).
+Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are."
+
+To assess perceptions of accuracy, we asked: "What do you think is the probability of advisor 1 making the best investment recommendation?".
+Participants answered with a slider on a scale from 0 to 100.
+To assess perceptions of accuracy, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
 
@@
 
-We manipulated convergence within participants in the same way we did in experiment 4.
-In addition, between participants, we manipulated informational dependence, akin to experiment 3.
-In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is.
-In the independent condition, there was no such conflict of interest and experts were described as independent.
-
-```{r}
-# Participants read only one of these introductions before seeing the sets of estimates, depending on which condition they had been assigned to
-data.frame(
-  Condition = c("Independence condition", "Conflict of interest condition"),
-  Description = c(
-    "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.",
-    "The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations."
-  )
-) #%>% 
-  # apa_table(align = c("m{3cm}", "m{12cm}")) 
-```
-
-#### Stimuli
-
-Identical to experiment 4, all participants see all four conditions, with two stimuli (i.e.
+We manipulated convergence within participants, and conflict of interest between participants.
+In the conflict of interest condition, experts were introduced this way: "The three advisors have already invested in one of the three options, the same option for all three.
+As a result, they have an incentive to push that option in their recommendations." For the independence condition: "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations."
+
+Participants saw all four convergence conditions (identical to those of Experiment 4), with two stimuli (i.e.
 game results) per condition.
-Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli).
-
-#### Dependent variables
-
-To assess accuracy, we asked participants: "What do you think is the probability of advisor 1 making the best investment recommendation?".
-Participants answered with a slider on a scale from 0 to 100.
-As for competence, we asked participants: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
-
-### Results
+Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).
+
+### Results and discussion
 
 To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants, using the `lme4` package and its `lmer()` function in R.
-Figure \@ref(fig:exp5-plot) visualizes the results and table \@ref(tab:exp5-table) contains descriptive results.
-
-Akin to the results of experiment 3 in the numerical setting, we find evidence for all four hypotheses.
-As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample to half of the participants, namely those who were assigned to the independence condition.
-On this reduced sample, we ran the exact same analyses as in experiment 4 and replicated the results (see blue colored distributions in Fig.
+
+We find evidence for all four hypotheses (see Fig.
 \@ref(fig:exp5-plot)).
+To test H1a and H1b, we use the same analyses as in Experiment 4, replicating the results.
 We find a positive effect of convergence on both accuracy (Convergence = `r exp5_descriptives$model_h1_accuracy$convergence$estimate` `r exp5_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp5_descriptives$model_h1_accuracy$convergence$p.value`) and competence (Convergence = `r exp5_descriptives$model_h1_competence$convergence$estimate` `r exp5_descriptives$model_h1_competence$convergence$ci`, p = `r exp5_descriptives$model_h1_competence$convergence$p.value`).
@@
 Likewise, the effect of convergence on competence was more positive ($\Delta$ Convergence = `r exp5_descriptives$model_competence$interaction$estimate` `r exp5_descriptives$model_competence$interaction$ci`, p = `r exp5_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition (Convergence = `r exp5_descriptives$model_competence$convergence$estimate` `r exp5_descriptives$model_competence$convergence$ci`, p = `r exp5_descriptives$model_competence$convergence$p.value`).
-
-```{r exp5-table}
-exp5_accuracy <- exp5 %>% 
-  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
-  group_by(independence, Convergence, convergence) %>% 
-  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
-  select(independence, Accuracy, Convergence, convergence) %>% 
-mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
-                "Independent") 
-         ) %>% 
-  pivot_wider(names_from = independence,
-              values_from = Accuracy) %>% 
-  arrange(convergence) %>% 
-  ungroup() %>% 
-  select(-convergence) %>% 
-  rename(" " = Convergence) %>% 
-  as.matrix()
-
-exp5_competence <- exp5 %>% 
-  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
-  group_by(independence, Convergence, convergence) %>% 
-  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
-  select(independence, Competence, Convergence, convergence) %>% 
-mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
-                "Independent") 
-         ) %>% 
-  pivot_wider(names_from = independence,
-              values_from = Competence) %>% 
-  ungroup() %>% 
-  arrange(convergence) %>% 
-  select(-c(convergence, Convergence)) %>% 
-  as.matrix()
-
-table <- cbind(exp5_accuracy, exp5_competence)
-
-table %>%
-  kable(booktabs = TRUE, 
-        # need to add a (empty) caption for cross referencing to work
-        caption = "") %>%
-  kable_styling() %>% 
-  add_header_above(c("", "Accuracy" = 2, "Competence" = 2)) %>%
-  # make smaller to fit
-  kable_styling(latex_options = "scale_down") 
-```
 
@@
 # run mixed model (random slope and intercept) with participants as random factor
-exp6_model_accuracy = lmer(accuracy ~ convergence + number_options + 
-                            number_options*convergence + (1 + convergence | id), 
+
+# For this model, we effect code `number_options`, so that we get a coefficient 
+# for the main effect of convergence as well as for the interaction
+exp6 <- exp6 %>% 
+  mutate(number_options_effect_code = recode(number_options, "three" = -0.5, "ten" = +0.5))
+
+exp6_model_accuracy = lmer(accuracy ~ convergence + number_options_effect_code + 
+                            number_options_effect_code*convergence + (1 + convergence | id), 
                        data = exp6)
-exp6_model_competence = lmer(competence ~ convergence + number_options + 
-                            number_options*convergence + (1 + convergence | id), 
+exp6_model_competence = lmer(competence ~ convergence + number_options_effect_code + 
+                            number_options_effect_code*convergence + (1 + convergence | id), 
                        data = exp6)
@@
 )
+
+summary(exp6_model_accuracy)
 ```
@@
 We ran a power simulation to inform our choice of sample size.
-All assumptions and details on the procedure can be found in the `power_Exp6.Rmd` document.
+All assumptions and details on the procedure can be found on the OSF.
 We used previous experiments and estimates of our models to inform our choice of parameter values.
@@
 The power simulation for `accuracy` suggested that for 140 participants we would cross the power threshold of 90% for the interaction effect (power = 0.928).
-The simulation for `competence` suggested that with 300 participants, we would detect an interaction with a power of 87% (power = 0.872).
+The simulation for `competence` suggested that with 300 participants, we would detect an interaction with a power of 87%.
 Due to budget constraints, we considered a sample of 300 participants as good enough, although slightly below our threshold.
@@
 
-We used the same experimental setting as in experiment 4.
+We used the same procedure as in Experiment 4, with the addition of one condition described below.
 
@@
 
-We manipulated convergence within participants in the same way we did in experiment 4.
-In addition, between participants, we manipulated the number of choice options.
-Participants were randomly assigned to either to see stimuli with '3' options, or with '10' options.
-Participants assigned to the '10' options condition, participants were assigned to one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the '3' options condition, and another with increased range (see Appendix XX).
-We added the increased range condition because we anticipated the possibility that participants might not consider all options as relevant when they only see scenarios in which all answers cluster.
-We found no differences between the two sub-conditions and collapsed them into a single '10' options condition.
-
-```{r stimuli-options, echo=FALSE}
-# Create a matrix of image file paths as Markdown-formatted strings
-image_paths <- data.frame(
-  imgage_a = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_3_a.png){ width=60% }"),
-  imgage_b = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_10_a.png){ width=60% }")
-  )
-
-# Use kable() to create the table and print it as Markdown
-kableExtra::kable(image_paths, format = "markdown",
-                  col.names = c("Number of options: 3", "Number of options: 10"), 
-                  align = "l",
-                  caption = "Example of a consensus stimulus for the two 'Number of option' conditions") %>%
-  # make smaller to fit
-  kable_styling(latex_options = "scale_down") 
-```
-
-#### Stimuli
-
-Identical to experiment 4, all participants saw all four conditions, with two stimuli (i.e.
-game results) per condition.
-Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).
-
-#### Dependent variables
-
-We used the same items as in experiment 4.
-To assess accuracy, we asked: "What do you think is the probability of player 1 being correct?".
-Participants answered with a slider on a scale from 0 to 100.
-As for competence, we asked participants: "How competent do you think player 1 is in games like these?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
-
-### Results
+The number of choice options was manipulated between participants.
+Participants were randomly assigned to either to see stimuli with three options (as in Experiment 4), or with ten options.
+Participants assigned to the ten options condition were divided into one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the three options condition, and another with increased range (see Appendix \@ref(exp6)).
+We found no differences between the two sub-conditions and collapsed them into a single ten options condition.
+
+### Results and discussion
 
 To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants, using the `lme4` package and its `lmer()` function in R.
-Figure \@ref(fig:exp5-plot) visualizes the results and table \@ref(tab:exp5-table) contains descriptive results.
 
@@
 We find a positive effect of convergence on both accuracy (Convergence = `r exp6_descriptives$model_h1_accuracy$convergence$estimate` `r exp6_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp6_descriptives$model_h1_accuracy$convergence$p.value`) and competence (Convergence = `r exp6_descriptives$model_h1_competence$convergence$estimate` `r exp6_descriptives$model_h1_competence$convergence$ci`, p = `r exp6_descriptives$model_h1_competence$convergence$p.value`).
-We do not find evidence that the number of choice options alters the effect of convergence on either accuracy or competence.
-
-```{r exp6-table}
-exp6_accuracy <- exp6 %>% 
-  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
-  group_by(number_options, Convergence, convergence) %>% 
-  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
-  select(number_options, Accuracy, Convergence, convergence) %>% 
-mutate(number_options = ifelse(number_options == "three", "3 options", 
-                "10 options") 
-         ) %>% 
-  pivot_wider(names_from = number_options,
-              values_from = Accuracy) %>% 
-  arrange(convergence) %>% 
-  ungroup() %>% 
-  select(-convergence) %>% 
-  rename(" " = Convergence) %>% 
-  as.matrix()
-
-exp6_competence <- exp6 %>% 
-  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
-  group_by(number_options, Convergence, convergence) %>% 
-  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
-  rounded_numbers() %>% 
-  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
-  select(number_options, Competence, Convergence, convergence) %>% 
-mutate(number_options = ifelse(number_options == "three", "3 options", 
-                "10 options") 
-         ) %>% 
-  pivot_wider(names_from = number_options,
-              values_from = Competence) %>% 
-  ungroup() %>% 
-  arrange(convergence) %>% 
-  select(-c(convergence, Convergence)) %>% 
-  as.matrix()
-
-table <- cbind(exp6_accuracy, exp6_competence)
-
-table %>%
-  kable(booktabs = TRUE, 
-        # need to add a (empty) caption for cross referencing to work
-        caption = "") %>%
-  kable_styling() %>% 
-  add_header_above(c("", "Accuracy" = 2, "Competence" = 2)) %>%
-  # make smaller to fit
-  kable_styling(latex_options = "scale_down") 
-```
-
-(ref:exp6-plot) Interaction of convergence and informational dependency.
-
-```{r exp6-plot, fig.cap="(ref:exp6-plot)"}
-# colors were identified with the wesanderson package (library(wesanderson)) as:
-# wes_palette("Zissou1", 100, type = "continuous")[A number]
-
-# plot for accuracy
-plot_accuracy <- ggplot(exp6,
-       aes(x = convergence_categorical, y = accuracy, fill = number_options)) +
-  geom_half_violin(position = position_nudge(x = -.2),
-                   adjust=2, alpha = .4,
-                   side = "l") +
-  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
-  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-  # Add nice labels
-  labs(x = "Convergence", y = "Accuracy", fill = NULL) +
-  scale_fill_manual(values = c("#EE3800", "#3D9BB2"), 
-                    name = "Number of choice options: ") +
-  plot_theme + 
-  theme(axis.text.x = element_text(angle = 20, hjust = 1))
-
-# plot for competence
-plot_competence <- ggplot(exp6,
-       aes(x = convergence_categorical, y = competence, fill = number_options)) +
-  geom_half_violin(position = position_nudge(x = -.2),
-                   adjust=2, alpha = .4,
-                   side = "l") +
-  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
-  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
-  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
-  # Add nice labels
-  labs(x = "Convergence", y = "Competence", fill = NULL) +
-  scale_fill_manual(values = c("#EE3800", "#3D9BB2")) +
-  plot_theme + 
-  theme(axis.text.x = element_text(angle = 20, hjust = 1))
-
-ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE)
-```
-
-# Discussion
+
+This finding also holds when looking at main effects of convergence across the entire sample ($beta_\text{Accuracy}$ = `r exp6_descriptives$model_accuracy$convergence$estimate` `r exp6_descriptives$model_accuracy$convergence$ci`, `r exp6_descriptives$model_accuracy$convergence$p.value`; $beta_\text{Competence}$ = `r exp6_descriptives$model_competence$convergence$estimate` `r exp6_descriptives$model_competence$convergence$ci`, `r exp6_descriptives$model_competence$convergence$p.value`).
+We do not find evidence of an interaction, i.e.
+evidence that the number of choice options alters the effect of convergence ($beta_\text{Accuracy}$ = `r exp6_descriptives$model_accuracy$interaction$estimate` `r exp6_descriptives$model_accuracy$interaction$ci`, `r exp6_descriptives$model_accuracy$interaction$p.value`; $beta_\text{Competence}$ = `r exp6_descriptives$model_competence$interaction$estimate` `r exp6_descriptives$model_competence$interaction$ci`, `r exp6_descriptives$model_competence$interaction$p.value`).
+
+## General discussion
+
+In two experiments (Experiment 1, and independence condition of Experiment 3), we find that participants presented with a set of more (rather than less) convergent numerical estimates find the estimates more accurate, and the individuals making the estimates more competent.
+Participants thus appear to draw normatively justified inferences.
+Experiment 2 suggests that participants do not think that a discussion between the individuals making the estimates could explain away the convergence of their estimates.
+By contrast, Experiment 3 reveals that, when the individuals making the estimates are systematically biased by a conflict of interest, then participants put less weight on the convergence of their estimates to infer that the estimates are accurate, or that the individuals making them are competent.
+
+Similar results are obtained in a categorical choice context, in which participants see the answers of individuals made within a limited set of options.
+Experiments 4, 5, and 6 show that, the more the answers converge, the more they are thought to be accurate, and the more the individuals who made them are thought to be competent.
+Experiment 5 shows that these inferences are weakened when the convergence can be explained by a conflict of interest (as in Experiment 4).
+Experiment 6 fails to find an effect of the number of options.
+
+## Conclusion
+
+When people see that others agree with each other, they tend to believe that they are right.
+This inference has been evidenced in several experiments, both for numerical estimates (refs), and for categorical choices (refs).
+However, in these experiments, the participants arguably assumed a degree of competence among the individuals whose answers they saw.
+For instance, when children are confronted with several individuals who agree on how to name a novel object (e.g.
+ref), they can assume that these (adult) individuals tend to know what the name of objects is.
+If the competence of the individuals is assumed, then well-known results from the literature on judgment aggregation--the wisdom of crowds--show that the average opinion of a set of individuals is, in a wide range of circumstances, more likely to be accurate than that of a single individual (ref).
+
+Here, we do not assume that the individuals answering are competent, asking the question: if we see a set of individuals, whose competence is unknown, converge on the same answer, is it rational to infer that this answer is more likely to be correct, and that the individuals are likely to be competent?
+We show that the answer is yes on both counts--assuming there is no systematic bias among the individuals answering.
+A series of XXX [math, simulations], reveal that, for both the numerical choice context and the categorical choice context, the more individuals agree on an answer, the more likely the answers are to be correct, the more likely the individuals are to be competent, with the former effect being stronger than the latter.
+Moreover, this is true for a wide range of distributions of competence.
+This means that, unless there are reasons to believe that the convergence of the answers is due to some external cause, such as a common bias among the individuals, people can safely infer that the more answers tend to converge, the more they are likely to be correct, and the more likely they are to have been made by competent individuals.
+
+In a set of experiments, we show that participants (US) draw these inferences: when presented with more convergent answers, they tend to believe the answers are more likely to be correct, and that the individuals who made them are more likely to be competent.
+This is true for numerical estimates and for categorical answers.
+These beliefs are weakened when the individuals making the estimates are systematically biased by a common conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).
+Finally, these beliefs are not strengthened when the number of individuals whose answers are converging increases, which might not be very surprising given that the effects of this number tend to plateau quite early in the simulations.
+
+The results--both simulations and experiments--are a novel contribution to the wisdom of crowd literature.
+In this literature--in particular that relying on the Condorcet Jury Theorem--a degree of competence is assumed in the individuals providing some answers.
+From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate.
+Here we show that the reverse inference--from agreement to competence--is also warranted.
+We also show that participants, by and large, are able to draw rational inferences, inferring accuracy and competence from an observation of convergence, and doing so more when there are no alternative explanations for the observed convergence.
+
+People might draw this inference in a variety of contexts, but the most prominent one might be science.
+Science is, arguably, the institution in which individuals end up converging the most in their opinions.
+For instance, scientists within the relevant disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA.
+This represents an incredible degree of convergence.
+When people hear that scientists have measured the distance between the solar system and the center of the galaxy, if they assume that there is a broad agreement within the relevant experts, this should lead them to infer that this measure is accurate, and that the scientists who made it are competent.
+Experiments have already shown that increasing the degree of perceived consensus among scientists tends to increase acceptance of the consensual belief [@vanstekelenburgScientificConsensusCommunicationContested2022; @deryuginaEffectInformationProvision2016; @dixonApplyingGatewayBelief2016; @kerrChangesPerceivedScientific2018; @lewandowskyPivotalRolePerceived2013; @vanderlindenHighlightingConsensusMedical2015a; @vanderlindenScientificConsensusClimate2015; @vanderlindenHowCommunicateScientific2014; but see @dixonImprovingClimateChange2017; @landrumExaminingImpactExpert2019], but it hasn't been shown that the degree of consensus also affects the perceived competence of scientists.
+
+In the case of science, the relationship between convergence and accuracy is broadly justified.
+However, at some points of history, there has been broad agreement on misbeliefs, such as when Christian theologians had calculated that the Earth was approximately six thousand years old.
+To the extent that people were aware of this broad agreement, and believed the theologians to have reached it independently of each other, this might have not only fostered acceptance of this estimate of the age of the Earth, but also a perception of the theologians as competent.
+
+The current study has a number of limitations.
+If the very abstract materials allow us to remove most of the priors the participants might have, they might also reduce the ecological validity of the experiments.
+Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized.
+Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, Americans) [@coppockGeneralizingSurveyExperiments2019].
+However, we do not know whether they would generalize to other cultures.
+
+Future studies could overcome these limitations by replicating the present results in different cultures, using more ecologically valid stimuli.
+For instance, it would be interesting to test whether the inference described here, from convergence to competence, might be partly responsible for the fact that people tend to believe scientists to be competent [REF MANY LABS PRE-PRINT].
+
+# Stuff
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::
 
@@
 
-# Appendix A {simulations}
-
-Are we justified in inferring competence and accuracy from convergence?
-To provide a normative answer, we ran simulations, intended to mirror our experimental setup.
-We find that - under certain conditions - more convergent groups indeed tend to be more competent and accurate.
-In this appendix, we describe these simulations in detail.
-
-## Numerical choice context
-
-```{r read-simulation-data-numeric, message=FALSE}
-# read data from simulations
-simulation_numeric <- read_csv("Experiment_1/data/sim.csv")
-```
-
-When several people estimate a quantity (numeric scenario), their convergence can be measured for example by the empirical variance.
-The closer the estimates, i.e.
-the smaller the empirical variance, the greater convergence.
-This measure is at the group level.
-
-To provide a normative answer, we ran simulations for a scenario in which individuals provide an estimate on a scale from 1000 to 2000.
-In our simulations, we suppose that an individual’s answer is drawn from a normal distribution.
-Each individual has their own normal distribution.
-All normal distributions are centered around the true answer - but they differ in their standard deviations.
-The value of that standard deviation is what we define as an individual’s competence.
-The lower the standard deviation, the higher the competence, i.e.
-the more likely a guess drawn from the normal distribution will be close to the true answer.
-We (arbitrarily) define a range of competence: we set the lowest competence equal to the range of possible values, i.e.the largest standard deviation (2000 - 1000) = 1000.
-We set the highest competence to 0.1% of the range of possible values, i.e.the smallest standard deviation (0.001 x 1000) = 1 (see Fig.
-\@ref(fig:example-competence-numeric)).
-
-(ref:example-competence-numeric) Range of possible data generating functions for individuals.
-
-```{r example-competence-numeric, fig.cap="(ref:example-competence-numeric)", out.width= "75%"}
-# Define the x-axis values
-x <- seq(1000, 2000, length.out = 1000)
-
-# Define the PDFs for the two distributions
-high_competence_pdf <- dnorm(x, mean = 1500, sd = 1)
-low_competence_pdf <- dnorm(x, mean = 1500, sd = 1000)
-
-# Create the plot
-ggplot() +
-  geom_line(aes(x, high_competence_pdf, color = "Highest Competence Individual \n (SD = 1, Mean = 1500)"), size = 1) +
-  geom_line(aes(x, low_competence_pdf, color = "Lowest Competence Individual \n (SD = 1000, Mean = 1500)"), size = 1) +
-  labs(x = "Competence Level", y = "Density", color = "Data generating function for") +
-  ggtitle("Competence Level Distributions") +
-  scale_color_manual(values = c("Highest Competence Individual \n (SD = 1, Mean = 1500)" = "blue", 
-                                "Lowest Competence Individual \n (SD = 1000, Mean = 1500)" = "red")) +
-  theme_minimal() +
-  theme(plot.title = element_text(hjust = 0.5))
-```
-
-We suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution.
-This competence distribution can take vary different shapes, depending on the alpha and beta parameters that describe the distribution (see fig.
-\@ref(fig:competence-distributions)).
-
-(ref:competence-distributions) The different population competence distributions we considered in our simulations.
-
-```{r competence-distributions, fig.cap="(ref:competence-distributions)", out.width= "75%"}
-plot_competence_distributions(simulation_numeric)
-```
-
-We draw an estimate for each individual based on their respective competence distribution.
-For each individual, we then measure accuracy as the (squared) distance between their estimate and the true answer.
-Having a competence and an accuracy value for each individual, we randomly assign individuals to groups of three.
-For each group, we calculate the average of the three individuals' competence and accuracy.
-We measure the convergence of a group by calculating the standard deviation of the estimates.
-We run this simulation on a sample size of roughly 100000 (varying slightly as a function of sample size).
-We repeat this simulation process for various sample sizes and competence distributions.
-The results are displayed in Fig.
-\@ref(fig:simulation-accuracy-numeric-1) to \@ref(fig:simulation-accuracy-numeric-6), for accuracy, and Fig.
-\@ref(fig:simulation-competence-numeric-1) to \@ref(fig:simulation-competence-numeric-6), for competence.
-Across all underlying competence distributions, we find a positive correlation between convergence and accuracy, which tends towards 1 as sample size increases (see Fig.
-\@ref(fig:simulation-numeric-correlations)).
-As for accuracy, we find a positive correlation between convergence and competence across all underlying competence distributions.
-However, this correlations are weaker than for accuracy, and do not increase with sample size (see Fig.
-\@ref(fig:simulation-numeric-correlations)).
-
-(ref:simulation-numeric-correlations) Correlation between accuracy (left) or competence (right) and convergence, as a function of sample size, grouped by the data underlying competence distributions.
-
-```{r simulation-numeric-correlations, fig.cap="(ref:simulation-numeric-correlations)", out.width="75%"}
-# calculate correlations
-simulation_numeric <- simulation_numeric %>% 
-  group_by(competence, sample) %>% 
-  mutate(correlation_accuracy = cor(accuracy_mean, convergence), 
-         correlation_competence = cor(competence_mean, convergence)) %>% 
-  ungroup()
-
-# accuracy plot
-accuracy_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_accuracy, color = competence)) +
-  geom_point() + 
-  geom_line(alpha = 0.5) + 
-  labs(y = "Correlation: Accuracy/Convergence", 
-       x = "Sample size", 
-       color = "Competence distribution") +
-  scale_y_continuous(limits = c(0, 1)) +
-  scale_color_brewer(palette = "Set1") +
-  plot_theme + 
-  theme(
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 8))
-
-# competence plot
-competence_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_competence, color = competence)) +
-  geom_point() + 
-  geom_line(alpha = 0.5) + 
-  labs(y = "Correlation: Competence/Convergence", 
-       x = "Sample size", 
-       color = "Competence distribution") +
-  scale_y_continuous(limits = c(0, 1)) +
-  scale_color_brewer(palette = "Set1") +
-  plot_theme + 
-  theme(
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 8))
-
-ggarrange(accuracy_plot, competence_plot, common.legend = TRUE)
-```
-
-## Categorical choice context
-
-```{r read-simulation-data-categorical, message=FALSE}
-# read data from simulations
-simulation_sample_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_sample_3_options.csv")
-
-simulation_options_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_options_3_sample.csv")
-```
-
-When people make a choice based on several categories, their answers cannot be ranked by their nature (i.e.
-they are nominal, not ordinal), and that there are fewer of them (e.g.
-one of three possible products to choose, instead of an estimate between one and two thousand).
-In this case, convergence can be measured by the share of people agreeing on an answer.
-The larger the share of informants agreeing on an answer, the greater convergence.
-This measure is at the response level, nested within the group level.
-
-As for the numeric scenario, we ran simulations to provide a normative answer as to whether it is justified to infer accuracy and competence from greater convergence.
-We, again, suppose that an individual's answer is drawn from an internal distribution - in this case, a multinomial distribution, that describes how likely the individual is to choose each available option.
-If there are m choice options, an individual has the probability p of picking the right one, and the probability of (1-p)/(m-1) to pick any other option.
-Each individual has their own multinomial distribution.
-We define competence as the probability of making the correct choice.
-The higher the competence, the greater the probability that an individual will choose the correct option.
-Competence values range from being at chance for selecting the correct option (p = 1/m) to certainly selecting the correct option (p = 1).
-As before in the numeric case, suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution (see fig.
-\@ref(fig:competence-distributions)).
-Based on their competence level, we draw an estimate for each individual.
-We measure an individual's accuracy as a binary outcome, namely whether they picked the correct option, or not.
-We then randomly assign individuals to groups of informants (we vary the sample size from one simulation to another).
-Within these groups, we calculate the share of individuals voting for an answer option.
-For example, in a scenario in which three individuals pick among three options (A, B and C), two individuals might vote C and one B.
-In this case we obtain an average accuracy and an average competence value for a share of 2/3 (option C) and for a share of 1/3 (option B).
-We simulate this on a population of 999000 individuals.
-We repeat this procedure varying the underlying population competence distributions, and additionally varying either (a) the sample size of informants, or (b) the number of choice options.
-If we vary the sample size, we hold the number of choice options constant at n = 3, and vice versa when varying the number of choice options.
-Fig.
-\@ref(fig:simulation-accuracy-sample-categorical) shows the average accuracy, and Fig.
-\@ref(fig:simulation-competence-sample-categorical) the average competence value for each share of votes, for different competence levels and varying the sample size.
-Fig.
-\@ref(fig:simulation-accuracy-options-categorical) and Fig.
-\@ref(fig:simulation-competence-options-categorical) display the same relationship, but varying the number of choice options instead.
-The figures dispaly that, across all sample sizes and competence levels, the larger the share of votes for an option, the more accurate the option is on average.
-That relationship appears to follow some sigmoid curve which switches from an average accuracy of 0 to an average accuracy of 1 before a share of 0.5 is attained, and which is steeper for larger sample sizes.
-For competence, we observe a similar sigmoid-like relationship, but of lesser amplitude and varying considerably as a function of the underlying population competence distributions.
-
-(ref:simulation-accuracy-sample-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and sample sizes.
-Points represent averages across all simulations within the respective segment.
-The number of choice options is three.
-
-```{r simulation-accuracy-sample-categorical, fig.cap="(ref:simulation-accuracy-sample-categorical)", out.width="75%"}
-plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
-                                       outcome = "Accuracy", variable = sample)
-```
-
-(ref:simulation-competence-sample-categorical) Competence as a function of vote share for an option, for different population competence distributions and sample sizes.
-Points represent averages across all simulations within the respective segment.
-The number of choice options is three.
-
-```{r simulation-competence-sample-categorical, fig.cap="(ref:simulation-competence-sample-categorical)", out.width="75%"}
-plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
-                                       outcome = "Competence", variable = sample)
-```
-
-(ref:simulation-accuracy-options-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and number of choice options.
-Points represent averages across all simulations within the respective segment.
-The sample size is three.
-
-```{r simulation-accuracy-options-categorical, fig.cap="(ref:simulation-accuracy-options-categorical)", out.width="75%"}
-plot_competence_vary_relative_majority_categorical(simulation_options_categorical,
-                                       outcome = "Accuracy", variable = options)
-```
-
-(ref:simulation-competence-options-categorical) Competence as a function of vote share for an option, for different population competence distributions and number of choice options.
-Points represent averages across all simulations within the respective segment.
-The sample size is three.
-
-```{r simulation-competence-options-categorical, fig.cap="(ref:simulation-competence-options-categorical)", out.width="75%"}
-plot_competence_vary_relative_majority_categorical(simulation_options_categorical,
-                                       outcome = "Competence", variable = options)
-```
-
-In sum, given the set of specific assumptions we made, our simulations suggest that people are indeed justified in inferring accuracy and competence from convergence in both numeric and categorical choice settings.
-
-(ref:simulation-accuracy-numeric) Simulation results showing the relationship between convergence and accuracy for different population competence distributions.
-
-```{r simulation-accuracy-numeric-1, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
-# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)
-
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[1]]
-```
-
-```{r simulation-accuracy-numeric-2, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[2]]
-```
-
-```{r simulation-accuracy-numeric-3, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[3]]
-```
-
-```{r simulation-accuracy-numeric-4, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[4]]
-```
-
-```{r simulation-accuracy-numeric-5, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[5]]
-```
-
-```{r simulation-accuracy-numeric-6, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[6]]
-```
-
-(ref:simulation-competence-numeric) Simulation results showing the relationship between convergence and competence for different population competence distributions.
-
-```{r simulation-competence-numeric-1, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
-# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[1]]
-```
-
-```{r simulation-competence-numeric-2, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[2]]
-```
-
-```{r simulation-competence-numeric-3, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[3]]
-```
-
-```{r simulation-competence-numeric-4, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[4]]
-```
-
-```{r simulation-competence-numeric-5, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[5]]
-```
-
-```{r simulation-competence-numeric-6, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[6]]
-```
+# (APPENDIX) Appendix {.unnumbered}
+
+```{r child = "appendix_exp1.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_exp2.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_exp3.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_exp4.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_exp5.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_exp6.Rmd"}
+```
+
+\clearpage
+
+```{r child = "appendix_simulations.Rmd"}
+```

### 2024-01-18 appendix_exp1.Rmd
version before more detailed text changes

# Experiment 1 {#exp1}

## Stimuli

\FloatBarrier

The means of the normal distributions that we draw our estimates from were distinct between sets of estimates. Considering our within-participant design, we wanted to ensure that participants understood each set of estimates as being the result of a different, unrelated game. In order to assure that random draws from the distributions will (most likely) appear on the response scale (1000 to 2000), we constrained the means of all normal distributions to lie between the first and third quartile of the response scale (i.e. smallest possible mean was 1225 and largest 1775). We define a set of eight means--one for each set of estimates--that cover the range from first to third quartile of the predefined scale with an equal interval (1250, 1325, 1400, 1475, 1550, 1625, 1700, 1775). We randomly paired means with conditions when generating the stimuli. We then drew the set of estimates from the respective normal distributions given the assigned means and the condition constraints. We repeated this three times, resulting in three different series of eight sets of estimates. We randomly assign participants to one of these series. Additionally, for each participant, we randomize the order of appearance of the sets of estimates within the respective series. Images of all sets of estimates can be found on the OSF.

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Results

Figure \@ref(fig:exp1-plot) visualizes the results and table \@ref(tab:exp1-table) contains descriptive results.

```{r exp1-table}
exp1 %>% 
  group_by(convergence, number) %>% 
  summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(confidence_mean, " (sd = ", confidence_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")")
         ) %>% 
  rename(Convergence = convergence,
         Number = number) %>% 
  select(Convergence, Number, Accuracy, Competence) %>% 
  apa_table()
```

(ref:exp1-plot) Distributions of accuracy and competence by level of convergence.

```{r exp1-plot, fig.cap="(ref:exp1-plot)"}
# make mean data for plot
x_nudge <- 0.8

means <- exp1 %>%
  group_by(convergence) %>%
  summarize(confidence = mean(confidence), 
            competence = mean(competence)) %>% 
  pivot_longer(c(confidence, competence), 
               names_to = "outcome", 
               values_to = "value") %>% 
  mutate(x_position = ifelse(convergence == "convergent", value + x_nudge, 
                                value - x_nudge))

confidence_plot <- ggplot(data=exp1, aes(x=confidence, fill=convergence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
  # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "confidence"), 
             aes(xintercept = value, 
                 color = convergence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "confidence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "plasma") + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=exp1, aes(x=competence, fill=convergence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
  # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "competence"), 
             aes(xintercept = value, 
                 color = convergence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "competence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.3)) +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "plasma") + 
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, 
                    competence_plot + 
                      theme(axis.text.y = element_blank(),
                            axis.ticks.y = element_blank(),
                            axis.title.y = element_blank()), 
                    common.legend = T) +
  theme(legend.position="top") 
figure
```

### 2024-01-18 appendix_exp2.Rmd
version before more detailed text changes

# Experiment 2 {#exp2}

## Design

\FloatBarrier

We manipulated informational dependency. According to the condition, players read a different introduction before seeing a set of estimates (see Table \@ref(tab:exp2-conditions)).

```{r exp2-conditions}
# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
data.frame(
  Condition = c("Independence", "Discussion"),
  Description = c(
    "Players are asked to make completely independent decisions – they cannot see each other’s estimates, or talk with each other before giving their estimates.",
    "Players are asked to talk with each other about the game at length before giving their estimates."
  )
) #%>% 
  # apa_table(align = c("m{3cm}", "m{12cm}")) 
```

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Results

Figure \@ref(fig:exp2-plot) visualizes the results and table \@ref(tab:exp2-table) contains descriptive results.

```{r exp2-table}
exp2 %>% 
  group_by(independence) %>% 
  summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(confidence_mean, " (sd = ", confidence_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")")
         ) %>% 
  rename(Independence = independence) %>% 
  select(Independence, Accuracy, Competence) %>% 
  apa_table()
```

(ref:exp2-plot) Distributions of accuracy and competence by level of informational dependency.

```{r exp2-plot, fig.cap="(ref:exp2-plot)"}
# make mean data for plot
x_nudge <- 0.8

means <- exp2 %>%
  group_by(independence) %>%
  summarize(confidence = mean(confidence), 
            competence = mean(competence)) %>% 
  pivot_longer(c(confidence, competence), 
               names_to = "outcome", 
               values_to = "value") %>% 
  mutate(x_position = ifelse(independence == "dependent", value + x_nudge, 
                                value - x_nudge))

confidence_plot <- ggplot(data=exp2, aes(x=confidence, fill=independence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
 # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "confidence"), 
             aes(xintercept = value, 
                 color = independence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "confidence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Accuracy (confidence) ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1) +
  scale_color_brewer(palette = "Accent",
                    direction = 1) +
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

competence_plot <- ggplot(data=exp2, aes(x=competence, fill=independence)) +
  # density plot
  geom_density(adjust=3, alpha=.6) +
 # mean lines
  geom_vline(data=means %>% 
               filter(outcome == "competence"), 
             aes(xintercept = value, 
                 color = independence),
             show.legend = FALSE,
             linetype="dashed") +
  geom_label(data = means %>% 
               filter(outcome == "competence"), 
             aes(x = x_position, y = 0.3, 
                 label =  paste("mean", round(value, digits = 2),
                                sep = " = ")), 
             alpha = 0.6, show.legend = FALSE, size = 3
            ) +
  # scales
  scale_x_continuous(name = "Competence ratings", breaks = seq(1, 7), 
                     limits=c(1, 7)) +
  scale_y_continuous(name = "Density", limits=c(0, 0.4)) +
  scale_fill_brewer(palette = "Accent",
                    direction = 1) +
  scale_color_brewer(palette = "Accent",
                    direction = 1) +
  guides(fill = guide_legend(title = NULL)) +
  plot_theme

figure <- ggarrange(confidence_plot, 
                    competence_plot + 
                      theme(axis.text.y = element_blank(),
                            axis.ticks.y = element_blank(),
                            axis.title.y = element_blank()), 
                    common.legend = T) +
  theme(legend.position="top") 
figure
```

### 2024-01-18 appendix_exp3.Rmd
version before more detailed text changes

# Experiment 3 {#exp3}

## Design

\FloatBarrier

We manipulated two factors: informational dependency (two levels, independence and conflict of interest, see Table \@ref(tab:conditions-exp3); between participants) and convergence (two levels, convergence and divergence; within participants). Participants saw four scenarios, one for each combination of the two factors convergence and informational dependency.

```{r conditions-exp3}
# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
data.frame(
  Condition = c("Independence", "Conflict of interest"),
  Description = c(
    "Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.",
    "All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future."
  )
) #%>% 
  # apa_table(align = c("m{3cm}", "m{12cm}"))  
```

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Stimuli

Our design required that each participant sees four different sets of predictions--two convergent and two divergent ones. We generated new stimuli: As in experiment 1, estimates would appear on a scale from 1000 to 2000. By contrast with experiment 1, we generated the sets of estimates with random draws from uniform distributions (instead of normal distributions). We varied the range of these distributions according to convergence (60 for convergence, 600 for divergence). Switching from normal distributions to uniform distributions made 'unlucky' draws in which conditions are visually not distinguishable less likely.

Similar to the previous experiments, we also vary the value on the prediction scale (from 1000 to 2000) across which the range is centered. Considering our within-participant design, this, again, made it seem more likely that participants understand each set of predictions as being the result of a different stock, with a different true value. In order to assure all random draws from the distributions would appear on the response scale, we constrained the center of the uniform distributions to lie between 1300 and 1700. We define four center values -- one per set of predictions -- that divide this interval in (rounded) quartiles (1300, 1430, 1570, 1700). Given a center and a range, we then draw the predictions from uniform distributions. For example, in a draw for a divergent set of estimates with a center of 1700, each value within a range of 1400 and 2000 is equally likely to get selected. To avoid that single draws overlap too much within the same set, we defined a minimum space of 5 between the three predictions of a set.

To minimize confounding of convergence with a certain placement of the center value, we paired center values with conditions such that each condition appears once in each half of the scale and each condition gets one of the extreme values. For example, in one set the convergence condition would get assigned center values of 1300 and 1570, the divergent condition center values of 1430 and 1700). We generated four such series (all possible combinations) and randomly assigned participants to one of them. Additionally, for each participant, we randomize the order of appearance of the sets of predictions within the respective series.

For each set of predictions, we calculated the empirical mean based on the randomly drawn estimates. In the conflict of interest condition, this mean was inserted as the value that participants were be told experts would gain from. Consequently, the convergent predictions converge around what is said to be the incentivized value for the experts to choose.

## Results

Table \@ref(tab:exp3-table) contains descriptive results.

```{r exp3-table}
exp3_accuracy <- exp3 %>% 
  group_by(independence, convergence) %>% 
  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
  select(independence, convergence, Accuracy) %>% 
    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
                "Convergent")) %>%  
  pivot_wider(names_from = convergence,
              values_from = Accuracy) %>% 
  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
                "Independent") 
         ) %>% 
  rename(" " = independence) %>% 
  as.matrix()

exp3_competence <- exp3 %>% 
  group_by(independence, convergence) %>% 
  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
  select(independence, convergence, Competence) %>% 
    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
                "Convergent")) %>%  
  pivot_wider(names_from = convergence,
              values_from = Competence) %>% 
  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
                "Independent") 
         ) %>% 
  ungroup() %>% 
  select(-independence) %>% 
  as.matrix()

table <- cbind(exp3_accuracy, exp3_competence)

table %>%
  kable(booktabs = TRUE, 
        # need to add a (empty) caption for cross referencing to work
        caption = "") %>%
  kable_styling() %>% 
  add_header_above(c("", "Accuracy" = 2, "Competence" = 2))

```

### 2024-01-18 appendix_exp4.Rmd
version before more detailed text changes

# Experiment 4 {#exp4}

\FloatBarrier

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Stimuli

We manipulate convergence within participants. All participants see all four conditions, with two stimuli (i.e. game results) per condition (see Table \@ref(tab:exp4-stimuli)). Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli)

```{r exp4-stimuli, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(condition = c("opposing majority (0)", 
                                        "dissensus (1)", "majority (2)", 
                                        "consensus (3)"),
                          imgage_a = c("![](Experiment_4/figures/stimuli/opp_majority_a.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/divergence_a.png){ width=60% }",
                        "![](Experiment_4/figures/stimuli/majority_a.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/consensus_a.png){ width=60% }"),
                        imgage_b = c("![](Experiment_4/figures/stimuli/opp_majority_b.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/divergence_b.png){ width=60% }",
                        "![](Experiment_4/figures/stimuli/majority_b.png){ width=60% }", 
                        "![](Experiment_4/figures/stimuli/consensus_b.png){ width=60% }"))

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "All stimuli by levels of convergence")

```

## Results

Table \@ref(tab:exp4-table) contains descriptive results.

```{r exp4-table}
exp4 %>% 
  group_by(convergence_categorical, convergence) %>% 
  summarise(across(c(competence, accuracy), list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ")"),
         Competence = paste0(competence_mean, " (sd = ", competence_sd, ")"),
         Convergence = paste0(convergence_categorical, " (", convergence, ")")
         ) %>% 
  ungroup() %>% 
  select(Convergence, Accuracy, Competence) %>% 
  apa_table()
```

### 2024-01-18 appendix_exp5.Rmd
version before more detailed text changes

--- 2024-01-18 appendix_exp3.Rmd
@@
-# Experiment 3 {#exp3}
+# Experiment 5 {#exp5}
 
@@
 
-We manipulated two factors: informational dependency (two levels, independence and conflict of interest, see Table \@ref(tab:conditions-exp3); between participants) and convergence (two levels, convergence and divergence; within participants).
-Participants saw four scenarios, one for each combination of the two factors convergence and informational dependency.
+We manipulated convergence within participants in the same way we did in experiment 4.
+In addition, between participants, we manipulated informational dependence, akin to experiment 3 (Table \@ref(exp5-condition)).
+In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is.
+In the independent condition, there was no such conflict of interest and experts were described as independent.
 
-```{r conditions-exp3}
-# The different introductions that participants read before seeing the sets of estimates, according to the conditions 
+```{r exp5-condition}
+# Participants read only one of these introductions before seeing the sets of estimates, depending on which condition they had been assigned to
 data.frame(
-  Condition = c("Independence", "Conflict of interest"),
+  Condition = c("Independence condition", "Conflict of interest condition"),
   Description = c(
-    "Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.",
-    "All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future."
+    "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.",
+    "The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations."
   )
@@
 
-Table \@ref(tab:exp3-table) contains descriptive results.
+Table \@ref(tab:exp5-table) contains descriptive results.
 
-```{r exp3-table}
-exp3_accuracy <- exp3 %>% 
-  group_by(independence, convergence) %>% 
+```{r exp5-table}
+exp5_accuracy <- exp5 %>% 
+  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
+  group_by(independence, Convergence, convergence) %>% 
   summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
@@
   mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
-  select(independence, convergence, Accuracy) %>% 
-    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
-                "Convergent")) %>%  
-  pivot_wider(names_from = convergence,
-              values_from = Accuracy) %>% 
-  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
+  select(independence, Accuracy, Convergence, convergence) %>% 
+mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
                 "Independent") 
          ) %>% 
-  rename(" " = independence) %>% 
+  pivot_wider(names_from = independence,
+              values_from = Accuracy) %>% 
+  arrange(convergence) %>% 
+  ungroup() %>% 
+  select(-convergence) %>% 
+  rename(" " = Convergence) %>% 
   as.matrix()
 
-exp3_competence <- exp3 %>% 
-  group_by(independence, convergence) %>% 
+exp5_competence <- exp5 %>% 
+  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
+  group_by(independence, Convergence, convergence) %>% 
   summarise(across(competence, list(mean = mean, sd = sd))) %>% 
@@
   mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
-  select(independence, convergence, Competence) %>% 
-    mutate(convergence = ifelse(convergence == "divergent", "Divergent", 
-                "Convergent")) %>%  
-  pivot_wider(names_from = convergence,
-              values_from = Competence) %>% 
-  mutate(independence = ifelse(independence == "conflict", "Conflict of interest", 
+  select(independence, Competence, Convergence, convergence) %>% 
+mutate(independence = ifelse(independence == "conflict of interest", "Conflict of interest", 
                 "Independent") 
          ) %>% 
+  pivot_wider(names_from = independence,
+              values_from = Competence) %>% 
   ungroup() %>% 
-  select(-independence) %>% 
+  arrange(convergence) %>% 
+  select(-c(convergence, Convergence)) %>% 
   as.matrix()
 
-table <- cbind(exp3_accuracy, exp3_competence)
+table <- cbind(exp5_accuracy, exp5_competence)
 
@@
   kable_styling() %>% 
-  add_header_above(c("", "Accuracy" = 2, "Competence" = 2))
-
+  add_header_above(c("", "Accuracy" = 2, "Competence" = 2)) %>%
+  # make smaller to fit
+  kable_styling(latex_options = "scale_down") 
 ```

### 2024-01-18 appendix_exp6.Rmd
version before more detailed text changes

# Experiment 6 {#exp6}

## Design

\FloatBarrier

We manipulated convergence within participants in the same way we did in experiment 4. In addition, between participants, we manipulated the number of choice options. Participants were randomly assigned to either to see stimuli with '3' options, or with '10' options. Participants assigned to the '10' options condition, participants were assigned to one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the '3' options condition, and another with increased range (see Appendix XX). We added the increased range condition because we anticipated the possibility that participants might not consider all options as relevant when they only see scenarios in which all answers cluster. We found no differences between the two sub-conditions and collapsed them into a single '10' options condition.

```{r stimuli-options, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(
  imgage_a = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_3_a.png){ width=60% }"),
  imgage_b = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_10_a.png){ width=60% }")
  )

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Number of options: 3", "Number of options: 10"), 
                  align = "l",
                  caption = "Example of a consensus stimulus for the two 'Number of option' conditions") %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down") 
```

## Attention check

Imagine you are playing video games with a friend and at some point your friend says: “I don’t want to play this game anymore! To make sure that you read the instructions, please write the three following words "I pay attention" in the box below. I really dislike this game, it's the most overrated game ever.”

Do you agree with your friend? (Yes/No)

## Stimuli

```{r stimuli-10, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(
  condition = c("opposing majority (0)", "dissensus (1)", "majority (2)", "consensus (10)"),
  imgage_a = c(
    "![](Experiment_6/figures/stimuli/minority_10_a.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/divergence_10_a.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/majority_10_a.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/consensus_10_a.png){ width=60% }"
  ),
  imgage_b = c(
    "![](Experiment_6/figures/stimuli/minority_10_b.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/divergence_10_b.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/majority_10_b.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/consensus_10_b.png){ width=60% }"
  )
)

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "Stimuli for 10 options condition by levels of convergence")

```

```{r stimuli-10-alt, echo=FALSE}
# Create a matrix of image file paths as Markdown-formatted strings
image_paths <- data.frame(
  condition = c("opposing majority (0)", "dissensus (1)", "majority (2)", "consensus (10)"),
  imgage_a = c(
    "![](Experiment_6/figures/stimuli/minority_10_a_alt.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/divergence_10_a_alt.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/majority_10_a_alt.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/consensus_10_a_alt.png){ width=60% }"
  ),
  imgage_b = c(
    "![](Experiment_6/figures/stimuli/minority_10_b_alt.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/divergence_10_b_alt.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/majority_10_b_alt.png){ width=60% }",
    "![](Experiment_6/figures/stimuli/consensus_10_b_alt.png){ width=60% }"
  )
)

# Use kable() to create the table and print it as Markdown
kableExtra::kable(image_paths, format = "markdown",
                  col.names = c("Level", "Version a)", "Version b)"), 
                  align = "c",
                  caption = "Alternative stimuli for 10 options condition by levels of convergence")

```

## Results

Figure \@ref(fig:exp5-plot) visualizes the results and table \@ref(tab:exp5-table) contains descriptive results.

```{r exp6-table}
exp6_accuracy <- exp6 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(number_options, Convergence, convergence) %>% 
  summarise(across(accuracy, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Accuracy = paste0(accuracy_mean, " (sd = ", accuracy_sd, ") ")) %>% 
  select(number_options, Accuracy, Convergence, convergence) %>% 
mutate(number_options = ifelse(number_options == "three", "3 options", 
                "10 options") 
         ) %>% 
  pivot_wider(names_from = number_options,
              values_from = Accuracy) %>% 
  arrange(convergence) %>% 
  ungroup() %>% 
  select(-convergence) %>% 
  rename(" " = Convergence) %>% 
  as.matrix()

exp6_competence <- exp6 %>% 
  mutate(Convergence = paste0(convergence_categorical, " (", convergence, ")")) %>% 
  group_by(number_options, Convergence, convergence) %>% 
  summarise(across(competence, list(mean = mean, sd = sd))) %>% 
  rounded_numbers() %>% 
  mutate(Competence = paste0(competence_mean, " (sd = ", competence_sd, ") ")) %>% 
  select(number_options, Competence, Convergence, convergence) %>% 
mutate(number_options = ifelse(number_options == "three", "3 options", 
                "10 options") 
         ) %>% 
  pivot_wider(names_from = number_options,
              values_from = Competence) %>% 
  ungroup() %>% 
  arrange(convergence) %>% 
  select(-c(convergence, Convergence)) %>% 
  as.matrix()

table <- cbind(exp6_accuracy, exp6_competence)

table %>%
  kable(booktabs = TRUE, 
        # need to add a (empty) caption for cross referencing to work
        caption = "") %>%
  kable_styling() %>% 
  add_header_above(c("", "Accuracy" = 2, "Competence" = 2)) %>%
  # make smaller to fit
  kable_styling(latex_options = "scale_down") 
```

(ref:exp6-plot) Interaction of convergence and informational dependency.

```{r exp6-plot, fig.cap="(ref:exp6-plot)"}
# colors were identified with the wesanderson package (library(wesanderson)) as:
# wes_palette("Zissou1", 100, type = "continuous")[A number]

# plot for accuracy
plot_accuracy <- ggplot(exp6,
       aes(x = convergence_categorical, y = accuracy, fill = number_options)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy", fill = NULL) +
  scale_fill_manual(values = c("#EE3800", "#3D9BB2"), 
                    name = "Number of choice options: ") +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_competence <- ggplot(exp6,
       aes(x = convergence_categorical, y = competence, fill = number_options)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence", fill = NULL) +
  scale_fill_manual(values = c("#EE3800", "#3D9BB2")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE)
```

### 2024-01-18 appendix_simulations.Rmd
version before more detailed text changes

# Simulations {simulations}

Are we justified in inferring competence and accuracy from convergence? To provide a normative answer, we ran simulations, intended to mirror our experimental setup. We find that - under certain conditions - more convergent groups indeed tend to be more competent and accurate. In this appendix, we describe these simulations in detail.

## Numerical choice context

```{r read-simulation-data-numeric, message=FALSE}
# read data from simulations
simulation_numeric <- read_csv("Experiment_1/data/sim.csv") %>% 
  # remove alpha and beta values from competence variable strings
  mutate(competence = sub("\n.*", "", competence))
```

When several people estimate a quantity (numeric scenario), their convergence can be measured for example by the empirical variance. The closer the estimates, i.e. the smaller the empirical variance, the greater convergence. This measure is at the group level.

To provide a normative answer, we ran simulations for a scenario in which individuals provide an estimate on a scale from 1000 to 2000. In our simulations, we suppose that an individual's answer is drawn from a normal distribution. Each individual has their own normal distribution. All normal distributions are centered around the true answer - but they differ in their standard deviations. The value of that standard deviation is what we define as an individual's competence. The lower the standard deviation, the higher the competence, i.e. the more likely a guess drawn from the normal distribution will be close to the true answer. We (arbitrarily) define a range of competence: we set the lowest competence equal to the range of possible values, i.e.the largest standard deviation (2000 - 1000) = 1000. We set the highest competence to 0.1% of the range of possible values, i.e.the smallest standard deviation (0.001 x 1000) = 1 (see Fig. \@ref(fig:example-competence-numeric)).

(ref:example-competence-numeric) Range of possible data generating functions for individuals.

```{r example-competence-numeric, fig.cap="(ref:example-competence-numeric)", out.width= "75%"}
# Define the x-axis values
x <- seq(1000, 2000, length.out = 1000)

# Define the PDFs for the two distributions
high_competence_pdf <- dnorm(x, mean = 1500, sd = 1)
low_competence_pdf <- dnorm(x, mean = 1500, sd = 1000)

# Create the plot
ggplot() +
  geom_line(aes(x, high_competence_pdf, color = "Highest Competence Individual \n (SD = 1, Mean = 1500)"), size = 1) +
  geom_line(aes(x, low_competence_pdf, color = "Lowest Competence Individual \n (SD = 1000, Mean = 1500)"), size = 1) +
  labs(x = "Competence Level", y = "Density", color = "Data generating function for") +
  ggtitle("Competence Level Distributions") +
  scale_color_manual(values = c("Highest Competence Individual \n (SD = 1, Mean = 1500)" = "blue", 
                                "Lowest Competence Individual \n (SD = 1000, Mean = 1500)" = "red")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

We suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution. This competence distribution can take vary different shapes, depending on the alpha and beta parameters that describe the distribution (see fig. \@ref(fig:competence-distributions)).

(ref:competence-distributions) The different population competence distributions we considered in our simulations.

```{r competence-distributions, fig.cap="(ref:competence-distributions)", out.width= "75%"}
plot_competence_distributions(simulation_numeric)
```

We draw an estimate for each individual based on their respective competence distribution. For each individual, we then measure accuracy as the (squared) distance between their estimate and the true answer. Having a competence and an accuracy value for each individual, we randomly assign individuals to groups of three. For each group, we calculate the average of the three individuals' competence and accuracy. We measure the convergence of a group by calculating the standard deviation of the estimates. We run this simulation on a sample size of roughly 100000 (varying slightly as a function of sample size). We repeat this simulation process for various sample sizes and competence distributions. The results are displayed in Fig. \@ref(fig:simulation-accuracy-numeric-1) to \@ref(fig:simulation-accuracy-numeric-6), for accuracy, and Fig. \@ref(fig:simulation-competence-numeric-1) to \@ref(fig:simulation-competence-numeric-6), for competence. Across all underlying competence distributions, we find a positive correlation between convergence and accuracy, which tends towards 1 as sample size increases (see Fig. \@ref(fig:simulation-numeric-correlations)). As for accuracy, we find a positive correlation between convergence and competence across all underlying competence distributions. However, this correlations are weaker than for accuracy, and do not increase with sample size (see Fig. \@ref(fig:simulation-numeric-correlations)).

(ref:simulation-numeric-correlations) Correlation between accuracy (left) or competence (right) and convergence, as a function of sample size, grouped by the data underlying competence distributions.

```{r simulation-numeric-correlations, fig.cap="(ref:simulation-numeric-correlations)", out.width="75%"}
# calculate correlations
simulation_numeric <- simulation_numeric %>% 
  group_by(competence, sample) %>% 
  mutate(correlation_accuracy = cor(accuracy_mean, convergence), 
         correlation_competence = cor(competence_mean, convergence)) %>% 
  ungroup()

# accuracy plot
accuracy_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_accuracy, color = competence)) +
  geom_point() + 
  geom_line(alpha = 0.5) + 
  labs(y = "Correlation: Accuracy/Convergence", 
       x = "Sample size", 
       color = "Competence distribution") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_brewer(palette = "Set1") +
  plot_theme + 
  theme(
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8))

# competence plot
competence_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_competence, color = competence)) +
  geom_point() + 
  geom_line(alpha = 0.5) + 
  labs(y = "Correlation: Competence/Convergence", 
       x = "Sample size", 
       color = "Competence distribution") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_color_brewer(palette = "Set1") +
  plot_theme + 
  theme(
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8))

ggarrange(accuracy_plot, competence_plot, common.legend = TRUE)
```

## Categorical choice context

```{r read-simulation-data-categorical, message=FALSE}
# read data from simulations
simulation_sample_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_sample_3_options.csv")

simulation_options_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_options_3_sample.csv")
```

When people make a choice based on several categories, their answers cannot be ranked by their nature (i.e. they are nominal, not ordinal), and that there are fewer of them (e.g. one of three possible products to choose, instead of an estimate between one and two thousand). In this case, convergence can be measured by the share of people agreeing on an answer. The larger the share of informants agreeing on an answer, the greater convergence. This measure is at the response level, nested within the group level.

As for the numeric scenario, we ran simulations to provide a normative answer as to whether it is justified to infer accuracy and competence from greater convergence. We, again, suppose that an individual's answer is drawn from an internal distribution - in this case, a multinomial distribution, that describes how likely the individual is to choose each available option. If there are m choice options, an individual has the probability p of picking the right one, and the probability of (1-p)/(m-1) to pick any other option. Each individual has their own multinomial distribution. We define competence as the probability of making the correct choice. The higher the competence, the greater the probability that an individual will choose the correct option. Competence values range from being at chance for selecting the correct option (p = 1/m) to certainly selecting the correct option (p = 1). As before in the numeric case, suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution (see fig. \@ref(fig:competence-distributions)). Based on their competence level, we draw an estimate for each individual. We measure an individual's accuracy as a binary outcome, namely whether they picked the correct option, or not. We then randomly assign individuals to groups of informants (we vary the sample size from one simulation to another). Within these groups, we calculate the share of individuals voting for an answer option. For example, in a scenario in which three individuals pick among three options (A, B and C), two individuals might vote C and one B. In this case we obtain an average accuracy and an average competence value for a share of 2/3 (option C) and for a share of 1/3 (option B). We simulate this on a population of 999000 individuals. We repeat this procedure varying the underlying population competence distributions, and additionally varying either (a) the sample size of informants, or (b) the number of choice options. If we vary the sample size, we hold the number of choice options constant at n = 3, and vice versa when varying the number of choice options. Fig. \@ref(fig:simulation-accuracy-sample-categorical) shows the average accuracy, and Fig. \@ref(fig:simulation-competence-sample-categorical) the average competence value for each share of votes, for different competence levels and varying the sample size. Fig. \@ref(fig:simulation-accuracy-options-categorical) and Fig. \@ref(fig:simulation-competence-options-categorical) display the same relationship, but varying the number of choice options instead. The figures dispaly that, across all sample sizes and competence levels, the larger the share of votes for an option, the more accurate the option is on average. That relationship appears to follow some sigmoid curve which switches from an average accuracy of 0 to an average accuracy of 1 before a share of 0.5 is attained, and which is steeper for larger sample sizes. For competence, we observe a similar sigmoid-like relationship, but of lesser amplitude and varying considerably as a function of the underlying population competence distributions.

(ref:simulation-accuracy-sample-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and sample sizes. Points represent averages across all simulations within the respective segment. The number of choice options is three.

```{r simulation-accuracy-sample-categorical, fig.cap="(ref:simulation-accuracy-sample-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Accuracy", variable = sample)
```

(ref:simulation-competence-sample-categorical) Competence as a function of vote share for an option, for different population competence distributions and sample sizes. Points represent averages across all simulations within the respective segment. The number of choice options is three.

```{r simulation-competence-sample-categorical, fig.cap="(ref:simulation-competence-sample-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Competence", variable = sample)
```

(ref:simulation-accuracy-options-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and number of choice options. Points represent averages across all simulations within the respective segment. The sample size is three.

```{r simulation-accuracy-options-categorical, fig.cap="(ref:simulation-accuracy-options-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_options_categorical,
                                       outcome = "Accuracy", variable = options)
```

(ref:simulation-competence-options-categorical) Competence as a function of vote share for an option, for different population competence distributions and number of choice options. Points represent averages across all simulations within the respective segment. The sample size is three.

```{r simulation-competence-options-categorical, fig.cap="(ref:simulation-competence-options-categorical)", out.width="75%"}
plot_competence_vary_relative_majority_categorical(simulation_options_categorical,
                                       outcome = "Competence", variable = options)
```

In sum, given the set of specific assumptions we made, our simulations suggest that people are indeed justified in inferring accuracy and competence from convergence in both numeric and categorical choice settings.

(ref:simulation-accuracy-numeric) Simulation results showing the relationship between convergence and accuracy for different population competence distributions.

```{r simulation-accuracy-numeric-1, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)

plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[1]]
```

```{r simulation-accuracy-numeric-2, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[2]]
```

```{r simulation-accuracy-numeric-3, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[3]]
```

```{r simulation-accuracy-numeric-4, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[4]]
```

```{r simulation-accuracy-numeric-5, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[5]]
```

```{r simulation-accuracy-numeric-6, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[6]]
```

(ref:simulation-competence-numeric) Simulation results showing the relationship between convergence and competence for different population competence distributions.

```{r simulation-competence-numeric-1, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[1]]
```

```{r simulation-competence-numeric-2, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[2]]
```

```{r simulation-competence-numeric-3, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[3]]
```

```{r simulation-competence-numeric-4, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[4]]
```

```{r simulation-competence-numeric-5, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[5]]
```

```{r simulation-competence-numeric-6, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[6]]
```

### 2024-01-18 functions/own_functions.R

@@ -25,6 +25,23 @@ text_ready <- function(model_output) {
   return(result)
 }
 
+# clean data frame results for paired t-test
+clean_t_test <- function(t.test) {
+  
+  t.test %>% 
+    tidy() %>% 
+    select(-c(method, alternative)) %>% 
+    # report p.value according to apa standards
+    mutate(p.value = case_when(p.value < 0.001 ~ "< .001",
+                               TRUE ~ sprintf("p = %.3f", p.value)
+    )
+    ) %>% 
+    # all other terms
+    rounded_numbers() %>%
+    mutate_all(~ {attributes(.) <- NULL; .})
+  
+}
+
 # Plot results from simulation
 
 # Plot competence distributions
@@ -35,33 +52,37 @@ plot_competence_distributions <- function(data) {
     reframe(alpha = unique(alpha), beta = unique(beta))
   
   # Function to plot the different competence distributions 
-  plot_competence_distribution <- function(alpha, beta) {
+  plot_competence_distribution <- function(alpha, beta, competence) {
     ggplot() +
       geom_function(
         fun = ~dbeta(.x, shape1 = alpha, shape2 = beta)
       ) +
-      labs(title = paste0("alpha = ", alpha, ", beta = ", beta), 
+      labs(title = paste0(competence,"\n", "alpha = ", alpha, ", beta = ", beta), 
            y = "Density", x = "Competence") +
-      plot_theme
+      plot_theme +
+      rremove("ylab") +  # Remove y-axis label
+      rremove("xlab")   # Remove x-axis label
   }
   
   # Specify the values of alpha and beta from the data
   alpha_values <- d$alpha
   beta_values <- d$beta 
+  competence_values <- d$competence
   
   # Create a list of ggplot objects
-  plot_list <- map2(alpha_values, beta_values, plot_competence_distribution)
+  plot_list <- pmap(list(alpha = alpha_values, beta = beta_values, competence = competence_values),
+                    plot_competence_distribution)
   
   # Arrange and display the plots
-  competence_plot <- ggarrange(plotlist = plot_list) %>%
-    ggpubr::annotate_figure(top = ggpubr::text_grob("Competence distributions", 
-                                                    face = "bold",
-                                                    color = "darkgrey",
-                                                    size = 14))
+  competence_plot <- ggarrange(plotlist = plot_list) %>% 
+    annotate_figure(figure,
[…]

### 2024-01-18 Pfander_DeCourson_Mercier_2023.Rmd
before changes to format of inline reporting of results

@@
 classoption       : "doc" # "doc" for nice look, "man" for manuscripty
-output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
+output            : papaja::apa6_doc # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors.
 For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g.
@@
 
-In the continuous case, the most relevant evidence comes from the literature on \'advice-taking.\' In these experiments, participants are called \'judges,\' and they need to make numerical estimates--sometimes on factual knowledge, e.g. \'What year was the Suez Canal opened first?
+In the continuous case, the most relevant evidence comes from the literature on 'advice-taking.' In these experiments, participants are called 'judges,' and they need to make numerical estimates--sometimes on factual knowledge, e.g. 'What year was the Suez Canal opened first?
 [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g.
 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020].
-To help answer these questions, participants are given estimates from others, the advisors.
-
-A substantial share of studies using this judge-advisor paradigm is irrelevant for studying inferences from convergence, because they present participants only with single estimates.
-In some studies, that single estimate is coming from one advisor [e.g.
-@bednarikEffectivenessImperfectWeighting2015a; @sollStrategiesRevisingJudgment2009; @yanivReceivingOtherPeople2004; @yanivAdviceTakingDecision2000; @harveyTakingAdviceAccepting1997].
-In other studies, that estimate is presented as an average coming from a group of advisors [e.g.
-@jaylesHowSocialInformation2017a; @mannesAreWeWise2009], but no information about the distribution of initial estimates is given.
-By contrast, advice-taking studies relevant to our argument provide participants with a set of estimates.
-One subset of these studies manipulate the degree of convergence between groups of advisors, e.g.
-via the variance of estimates [@mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009], or their range [@budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007].
+To help answer these questions, participants are given estimates from others, the advisors [REVIEW ARTICLE].
+
+Most of this literature is irrelevant to the point at hand since participants are presented with single estimates, either from a single advisor [e.g.
+@bednarikEffectivenessImperfectWeighting2015a; @sollStrategiesRevisingJudgment2009; @yanivReceivingOtherPeople2004; @yanivAdviceTakingDecision2000; @harveyTakingAdviceAccepting1997], or as an average coming from a group of advisors [e.g.
+@jaylesHowSocialInformation2017a; @mannesAreWeWise2009], but without any information about the distribution of initial estimates.
+
+By contrast, other advice-taking studies provide participants with a set of individual estimates.
+One subset of these studies manipulate the degree of convergence between groups of advisors, through the variance of estimates [@mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009], or their range [@budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007].
 These studies find that participants are more confident about, or rely more on, estimates from groups of advisors who converge more.
-More indirect evidence comes from another subset of studies on outlier opinions.
-In these studies, participants were presented with sets of estimates where all but one estimate--the outlier--were close to each other [@harriesCombiningAdviceWeight2004], or overlapping, when estimates were intervals [@yanivWeightingTrimmingHeuristics1997, study 3 & 4].
-The outlier was defined, for example, as having "[...] a z-score greater than 2.0 and the remaining opinions had z-scores less than 2.0" [@harriesCombiningAdviceWeight2004].
+
+Other studies varied the degree of convergence within a group of advisors.
+They present participants with a set of estimates where some estimates are close to each other (or overlapping, in cases where estimates were intervals), while others are outliers [@harriesCombiningAdviceWeight2004; [@yanivWeightingTrimmingHeuristics1997, study 3 & 4].
 These studies find that participants heavily discount outliers when aggregating estimates.
-These results are thus consistent with the idea that people rely more on convergent opinions, although testing extreme cases of singular dissident opinions.
-Yet more indirect evidence comes from experiments that varied convergence of estimates from single advisors [@yanivWeightingTrimmingHeuristics1997, study 1 & 2].
-In this study, participants were shown ranges of estimates from two advisors - one with a small range (e.g.
-3 to 5), the other with a big range (e.g.
-1 to 10).
-Participants relied more on the precise, small range advice.
-These results can be interpreted as a case of inferring accuracy from (internal) convergence of single informants, perceived by participants perhaps as consistency or for example confidence.
-However, in this study, the small-range advisor was always within the range of the big-range adivsor.
-Consequently, if people relied more on the small-range advisor, then it was likely not only because of greater convergence, but also because of the overlap between the two advisors.
-
-## Categorical choices
-
-For categorical choice contexts, experimental evidence on people inferring accuracy from convergence is scarce.
-There is an extensive body of literature suggesting that people are susceptible to adopting majority behavior and opinions.
-But it is is difficult to tear apart whether they do so in a strive for normative conformity or for accuracy [@mercierMajorityRulesHow2019].
-However, some aspects of the experimental paradigm should favor motivations for accuracy, for example private answers and individual rewards for accuracy.
-As for human adults, we are aware of only one article that assessing majority effects using such accuracy motivations: In a series of experiments using such accuracy motivations, @morganEvolutionaryBasisHuman2012 systematically varied the proportion of informants agreeing on a given answer, from 50 to 100%.
-They found that the larger the (relative) size of the majority, the more likely participants were to adopt the majority answer.
-As for children, relevant evidence comes from studies on social learning in the field of developmental psychology.
-Several such studies rely on a similar paradigm revolving around a naming tasks [@fusaroChildrenAssessInformant2008; @corriveauGoingFlowPreschoolers2009; @bernardChildrenWeighNumber2015a; @bernardFourSixYearOldChildren2015; @chenChildrenTrustConsensus2013].
-In @corriveauGoingFlowPreschoolers2009, for example, children had to find out which of three unfamiliar objects corresponds to the made up name "modi".
-Three adult informants pointed to one object, but a fourth adult pointed to a different one.
-The children were then asked for their opinion.
-All these studies consistently find that children adopt the majority opinion.
-It is, however, not clear how the naming paradigm taps into intuitions for inferring accuracy - after all, assigning made-up names to objects has no objectively correct solution and, accordingly, these studies are often interpreted to reveal conformity in children [see e.g.
-@vanleeuwenConformityItsLookalikes2015].
-Other experimental paradigms yield similar results.
-@morganDevelopmentAdaptiveConformity2015 had children (aged between 3 and 7) count which of two pictures displayed more dots.
-Among 10 adult informants, they varied the proportion of those agreeing on one picture.
-They find that children relied more on advice from larger (relative) majorities, although younger children (aged 3 to 4) were only affected in cases where the majority would stand against a single dissident.
-In @haunMajorityBiasedTransmissionChimpanzees2012, 2-year-olds tried to use an unfamiliar box to deliver a reward.
-The children were more likely to copy an action demonstrated by three informants rather than a different action demonstrated three times by the same informant.
-In @herrmannStickScriptEffect2013, children (aged 3 to 6) were required to copy a necklace-making demonstration from video.
-Children showed more fidelity to the demonstrations after watching two identical demonstrations by two different adults than after watching the same individual twice.
-In @vanleeuwenDevelopmentHumanSocial2018, children could use a choice box with three different pipes to insert a boll, to then receive a toy.
-Before using the box, they were shown a video of four other children using the box.
-Three of these children used one particular pipe, each throwing in one ball and receiving one toy per insertion, sequentially.
-One additional demonstrator used one of the two other pipes three times in a row, also receiving one toy per insertion.
-Children were then given one ball and had to chose which pipe to use.
-Children generally followed the majority choice.
-
-Although the evidence is not conclusive, especially for categorical choice scenarios, the current literature suggests that people believe more convergent opinions to be more accurate.
-But is this inference sound?
-
-## Standard wisdom of crowds approaches
-
-Literature on the wisdom of crowds provides normative ground for inferring accuracy from others' aggregated opinions [@surowieckiWisdomCrowds2005].
-What role does this literature assign to convergence?
-Here we look at two of the main phenomena that can generate wisdom of crowds effects: averaging, and the CJT.
-
-### Averaging
-
-In the beginning of the 20th century, Francis Galton famously demonstrated accuracy gains from averaging: in his "vox populi" experiment, he asked 787 participants to estimate the weight of an ox, and found the average estimate of an ox weight to be just one pound from reality [@galtonVoxPopuli1907].
-Since then, it has been shown both theoretically and empirically that the mean or median answer of a group is typically closer to the truth than the mean individual is [see e.g.
-@yanivBenefitAdditionalOpinions2004; @mannesWisdomSelectCrowds2014].
-For instance, when considering a range of numerical estimates that deviate more or less from a correct answer, the error of the mean answer will always be either lower than the mean error (if the correct answer is within the range of all the answers provided), or the same as the mean error (otherwise) [@larrickIntuitionsCombiningOpinions2006].
-In fact, the error of the mean is often uncannily small compared to the mean error, a phenomenon which has allowed averaging to considerably improve performance on a variety of problems ranging from political predictions to medical diagnoses [@surowieckiWisdomCrowds2005].
-
-Averaging works independently of the degree of convergence of a set of estimates.
-It is not clear how it leads to more accurate results, without making assumptions about the data generating process of these estimates.
-
-[One might ask why talk about averaging at all - good question. I guess just because it happens in a numeric context and is used to justify aggregating opinions. But the link is not straightforward.]
-
-### Condorcet Jury Theorem
-
-In the late 18th Century, Condorcet established a formal argument in favor of large scale majority voting, which would later become known as the Condorcet Jury Theorem (CJT) [@decondorcetEssaiApplicationAnalyse2014].
-He demonstrated that, for a binary decision (e.g.
-correct vs.
-false), the probability that the majority vote is correct is larger than the probability of each individual being correct.
-Moreover, as the number of votes increases, the probability that the majority is correct converges towards one.
-Crucially, the original CJT assumes that (i) each individual has the same probability p to select the correct answer, and (ii) that this probability is at least slightly better than chance (p \> 0.5) (there are also assumptions about the independence and lack of strategic motives, for reviews, see, e.g.
-@ladhaCondorcetJuryTheorem1992, @austen-smithInformationAggregationRationality1996, @dietrichEpistemicDemocracyDefensible2013).
-
-At least in its basic version, the CJT glosses over variation in convergence, i.e.
-the relative size of the majority.
-It allows to make statements such as 'given an individual-level competence p, the probability that a (any) majority vote is correct is Y'.
-This latter probability comprises votes with 99% majorities as well 51% majorities.
-To justify inferences of accuracy and competence from convergence, CJT would need to allow making statements such as 'given a majority of X the probability that the majority vote is correct is Y, and the individual-level competence is Z.' In an extension of the CJT, @romeijnLearningJurorCompetence2011 provide normative grounds for making such statements.
-They show that with increasing relative majority, group members are more likely to be competent and the majority decision to be accurate (??).
-Their framework is limited, however, to binary decisions.
-
-In sum, the wisdom of crowd literature provides at best insufficient normative grounds for inferring accuracy and competence from convergence.
-Its main purpose is to justify that aggregating across many opinions is a better strategy than picking any individual's opinion.
-The proposed aggregation strategies - whether averaging or following the majority - are about extracting wisdom from any crowd.
-But not all crowds are equally wise.
-Competence should play a crucial role in the confidence we have in any wisdom of crowd inferences.
-In the absence of more informative cues, our literature review shows that people place more trust in more convergent crowds.
-The wisdom of crowd literature does not provide generalizable justification for this behavior.
-
-## Inferring competence from the convergence of opinions
-
-As Eratosthenes' example above illustrates, intuitively it seems there are cases in which we can infer that the members of a group are more likely to be correct, or, equivalently, more competent, when their answers converge.
-When are we justified to infer that people are right (and thus competent) when they converge?
-
-To provide a normative answer, we built analytical and agent-based models (see \@ref(simulations)).
-We consider two scenarios in which opinions can converge: a numerical and categorical one.
-In a numerical setting, we measure convergences by the empirical variance.
-The closer the estimates, i.e.
-the smaller the empirical variance, the greater convergence.
-In the categorical setting, we measure convergence by the share of votes for an option.
-The larger the share of votes for one option, the greater convergence.
-
-### Numeric
-
-### Categorical
-
-Agents answer a categorical question, with m+1 answers.
-We define the competence p of an individual as the probability to choose the right answer.
-Each of the wrong answers has the same probability (1-p)/m to be chosen.
-We observe the answers of a population of n+1 agents, with diverse levels of competence drawn from a beta distribution -- a flexible probability distribution that can be uniform, unimodal or bimodal, depending on the parameters.
-We assume that the shape parameters of the beta distribution ($\alpha$ and $\beta$) are known, but the right answer is not.
-We aim to infer the competence of an individual, based on the population answers.
-In our analytical model, we use Bayes formula to compute the posterior competence distribution of competence for an individual, knowing that a number K of other participants agree with him.
-The posterior distribution is a mixture of two beta distributions, Beta($\alpha$+1, $\beta$) and Beta($\alpha$, $\alpha$+1), weighted respectively by the probabilities that K individuals would choose the right or the wrong answer.
-The model suggests that the more consensual an individual's vote is, the more that individual is estimated to be competent (Fig.
-\@ref(fig:figure-Benoit)).
-
-(ref:figure-Benoit) Average competence of an informant as a function of K, the number of individuals who agree.
-In this example, we assume that the number of informants - not counting the focal informant - in a sample is n=10, m=5, and that the population competence ($\alpha$ = $\beta$ = 1).
-
-```{r figure-Benoit, echo=FALSE, fig.cap="(ref:figure-Benoit)", out.width= "50%", fig.align="left"}
-knitr::include_graphics("figures/figure_Benoit.png")
-```
-
-That said, the average competence is bounded between and $\frac{\alpha}{\alpha+\beta+1}$ and $\frac{\alpha+1}{\alpha+\beta+1}$.
-In other words, even if all other individuals agree with an informant, a case in which the informant very likely chose the correct answer (add an accuracy figure, maybe?), the competence we should assign to this informant, according to our model, is not higher than $\frac{\alpha+1}{\alpha+\beta+1}$ (or $2/3$ in our example in Fig.
-\@ref(fig:figure-Benoit)).
-This counter-intuitive result is due to our framework: we assume that a certain competence distribution as a given, and only observe one decision per individual.
-Intuitively, in a world where even some highly incompetent people sometimes pick the correct answer by chance, being right (with the majority) once is not a sufficient signal of highest competence.
+
+Studies on advice taking thus suggest participants believe that more convergent opinions are more likely to be correct.
+None of these studies investigated whether participants also believe that those whose opinions converge are also more likely to possess some underlying competence.
+
+In categorical choice contexts, there is ample and long-standing [e.g.
+@crutchfieldConformityCharacter1955] evidence from social psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger (both in absolute and in relative terms) [e.g., @morganEvolutionaryBasisHuman2012; for review, see @mercierMajorityRulesHow2019).
+This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role [e.g.
+because the answers are private, see @aschStudiesIndependenceConformity1956].
+Similar results have been obtained with young children [e.g.
+@fusaroChildrenAssessInformant2008; @corriveauGoingFlowPreschoolers2009; @bernardChildrenWeighNumber2015a; @bernardFourSixYearOldChildren2015; @chenChildrenTrustConsensus2013; @herrmannStickScriptEffect2013; @morganDevelopmentAdaptiveConformity2015].
+
+If many studies have demonstrated participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergent was indicative of the individuals' competence.
+One potential exception is a study with preschoolers [@corriveauGoingFlowPreschoolers2009] in which the children were more likely to believe the opinion of an informant who had previously been the member of a majority over that of an informant who had dissented from the majority.
+However, it is not clear whether the children thought the members of the majority were particularly competent, since their task--naming an object--was one in which children should already expect a high degree of competence from (adult) informants.
+This result might thus indicate simply that children infer that someone who disagrees with several others on how to call something is likely wrong, and thus likely less competent at least in that domain.
+
+## Is inferring competence from convergence justified?
+
+There is clear evidence that, both in the continuous and the categorical case, participants infer that convergent opinions are more likely to be correct.
+There is also tentative evidence that participants also infer that individuals who hold more convergent opinions are more likely to be competent.
+Would that second inference be justified?
+More precisely, if we consider individuals who have convergent opinions, by contrast with individuals whose opinions are less convergent, or whose opinion diverges from that of a majority, are we justified in thinking that the individuals with more convergent opinions are more likely to be right not only for this specific question, but that they would also be right for other (related) questions?
+
+This question can be broken down in two questions, at two different scales.
+First, can we infer that a population where there is a stronger degree of consensus is, on average, more competent?
+Second, can we infer that inside this population, individuals who are closer to the consensus are more competent?
+[add the accuracy/competence distinction?] In the numerical case [trivial analytical result?], it is easy to show that the answer is yes in both cases.
+We can assume that individuals' answers are a normally distributed random variable, centered on the truth -- that is, participants have no systematic bias.
+The variance of this variable represents the competence of individuals.
+Then, both at the population and the individual level, the "sample variance" in their answers -- the mean squared distance to the mean response -- is the best way to estimate competence.
+Thus, the degree of convergence in the population can be used as a cue of the population competence, and the closeness to consensus of an individual as a cue of individual competence.
+The categorical case is more intricate.
+Here, we assume that agents answer a question, with several categorical answers.
+We define the competence of an individual as the probability to choose the right answer.
+In this case, we showed that the size of the majority was a good approximation of the average competence of the population.
+To make an inference at the individual level, we need to specify a population distribution of competence.
+Here, we assume that the mean competence is equal to the majority size, and a particular value for variance.
+Using Bayes theorem, we can show that the more agents agree with a focal participant, the more the focal agent can be inferred to be competent (Fig.
+\@ref(fig:figure-Benoit-2)).
+Additionally, the more competence varies in the population, the more the degree of convergence is informative on an individual competence.
+It makes sense: if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence.
+More generally, the fact that the focal individual is right (his 'accuracy') can be inferred more strongly than his competence (Fig.
+\@ref(fig:figure-Benoit-2)).
+
+(ref:figure-Benoit-2) Average estimated competence and accuracy for an individual, depending on how many participants agree with him.
+Here, we assume that 20 individuals answer a 5-choice questions.
+For values higher than 4, we assume that the individual is part of the majority.
+Below 4, we assume that the focus individual is part of a minority, and that the majority size is 4.
+
+```{r figure-Benoit-2, echo=FALSE, fig.cap="(ref:figure-Benoit-2)", out.width= "50%", fig.align="left"}
+knitr::include_graphics("figures/figure_Benoit-2.png")
+```
+
+This makes sense: the only reason why a member of the majority can be inferred to be more competent is because he is more likely to be right (Fig. \@ref(fig:figure-Benoit-2)).
+
+(ref:diagram-Benoit) Diagram of the data generating process.
+Due to noise, there is information loss between each step.
+Because competence precedes accuracy in the causal chain, more certain inferences can be made about accuracy, compared to competence.
+
+```{r diagram-Benoit, echo=FALSE, fig.cap="(ref:diagram-Benoit)", out.width= "50%", fig.align="left"}
+knitr::include_graphics("figures/diagram_Benoit.png")
+```
+
+In order to better understand the influence of different parameters (e.g.
+degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence, we conducted simulations.
+
+### Continuous case
+
+```{r read-simulation-data-numeric, message=FALSE}
+# read data from simulations
+simulation_numeric <- read_csv("Experiment_1/data/sim.csv") %>% 
+  # remove alpha and beta values from competence variable strings
+  mutate(competence = sub("\n.*", "", competence))
+```
+
+In these simulations, a set of agents with varying degrees of competence provide numerical answers.
+We measure how accurate these answers are, and how much they converge (i.e.
+how low their variance is).
+We then correlate the degree of convergence with the accuracy and with the competence of the agents.
+
+More specifically, agents provide an estimate on a scale from 1000 to 2000 (chosen to facilitate the experiments below).
+Each agent is characterized by a normal distribution of possible answers.
+All of the agents' distributions are centered around the correct answer, but their standard deviation varies, denoting varying degrees of competence.
+The agents' standard deviation varied from 1 (highest competence) to 1000 (lowest competence).
+Agents' competence are drawn from a population competence distribution, expressed by a beta distribution, which can take different shapes.
+We conducted simulations with a variety of beta distributions which cover a large range of possible competence populations (see Fig.
+\@ref(fig:competence-distributions)).
+
+(ref:competence-distributions) The different population competence distributions we considered in our simulations. 
+In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1.
+In the categorical simulations, a competence value of 0 corresponds to chance, e.g.
+in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$, while competence of 1 corresponds to definitely picking the correct answer.
+
+```{r competence-distributions, fig.cap="(ref:competence-distributions)", out.width= "75%"}
+plot_competence_distributions(simulation_numeric)
+```
+
+In the simulation, a population of around 100000 (CHECK THAT AGAIN) agents with different competence is generated.
+An answer is drawn for each agent, based on their competence.
+The accuracy of this answer is defined as the squared distance to the true answer.
+Having a competence and an accuracy value for each agent, we randomly assign agents to groups of, e.g., three.
+For each group, we calculate the average of the agents' competence and accuracy.
+We measure the convergence of a group by calculating the standard deviation of the answers.
+We repeat this process for different sample sizes for the groups, and different competence distributions.
+
+The simulation thus provides, for groups of agents of different size, and drawn from different competence distributions, a measure of their average accuracy, their average competence, and the degree of convergence of their answers, allowing us to represent the relation between these variables.
+Of particular interest, we depict in Fig.
+XX the correlation between the convergence of the answers and their accuracy (left), and between the convergence of the answers and the competence of the agents who produced them (right) (for more detailed figures, see Electronic Supplementary Materials).
+We draw broad conclusions from these results after reporting the outcome of the simulations with categorical answers.
+
+(ref:simulation-numeric-correlations) Correlation between the convergence of the answers of a group of agents and (left) the average accuracy of their answers; (right) the average competence of the agents, as a function of how many agents are in the group, and of the competence distribution of the agents (see Fig.
+\@ref(fig:competence-distributions)).
+
+```{r simulation-numeric-correlations, fig.cap="(ref:simulation-numeric-correlations)", out.width="75%"}
+# calculate correlations
+simulation_numeric <- simulation_numeric %>% 
+  group_by(competence, sample) %>% 
+  mutate(correlation_accuracy = cor(accuracy_mean, convergence), 
+         correlation_competence = cor(competence_mean, convergence)) %>% 
+  ungroup()
+
+# accuracy plot
+accuracy_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_accuracy, color = competence)) +
+  geom_point() + 
+  geom_line(alpha = 0.5) + 
+  labs(y = "Correlation \nAccuracy/Convergence", 
+       x = "Group size", 
+       color = "Competence distribution") +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_color_brewer(palette = "Set1") +
+  plot_theme + 
+  theme(
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)) +
+  rremove("xlab")
+
+# competence plot
+competence_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_competence, color = competence)) +
+  geom_point() + 
+  geom_line(alpha = 0.5) + 
+  labs(y = "Correlation \nCompetence/Convergence", 
+       x = "Group size", 
+       color = "Competence distribution") +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_color_brewer(palette = "Set1") +
+  plot_theme + 
+  theme(
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)) +
+  rremove("xlab")
+
+ggarrange(accuracy_plot, competence_plot, common.legend = TRUE) %>% 
+    annotate_figure(bottom = textGrob("Group size", gp = gpar(cex = 1, fontface = "bold")))
+```
+
+### Categorical case
+
+```{r read-simulation-data-categorical, message=FALSE}
+# read data from simulations
+simulation_sample_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_sample_3_options.csv")  %>% 
+  # remove alpha and beta values from competence variable strings
+  mutate(competence = sub("\n.*", "", competence))
+
+simulation_options_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_options_3_sample.csv")  %>% 
+  # remove alpha and beta values from competence variable strings
+  mutate(competence = sub("\n.*", "", competence))
+```
+
+In the case of categorical answers, convergence can be measured as the relative size of informants picking the same option.
+The Condorcet jury theorem [@decondorcetEssaiApplicationAnalyse2014; for recent treatments, see, e.g., @ladhaCondorcetJuryTheorem1992, @austen-smithInformationAggregationRationality1996, @dietrichEpistemicDemocracyDefensible2013] and its extensions to situations with more than two options (REF) already shows that the answer defended by the majority/plurality is more likely to be correct.
+Regarding competence, @romeijnLearningJurorCompetence2011 have shown that when individuals are more competent, the share of majority vote tends to increase.
+However, they have only studied a case with a uniform distribution of competence.
+By contrast, here, we investigate a wide range of distributions of competence, and we do not assume that all individuals are equally competent, meaning that, from an observed set of answers, we can attribute an average competence to individuals whose answers form a majority/plurality vs.
+individuals whose answers form a minority.
+
+We simulate agents who have to decide between a number of options, one of which is correct, and whose competence varies.
+Competence is defined as a value p, which can differ for each agent, and which corresponds to the probability of selecting the right answer (the agents then have a probability (1-p)/(m-1), with m being the number of options, of selecting any other option).
+Competence values range from chance level (p = 1/m) to always selecting the correct option (p = 1).
+Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see Fig.
+\@ref(fig:competence-distributions)).
+Based on their competence level, we draw an answer for each agent.
+We measure an agent's accuracy as a binary outcome, namely whether they selected the correct option or not.
+In each simulation 999000 agents are generated, and then randomly assigned to groups (of varying size in different simulations).
+Within these groups, we (i) calculate the share of individuals voting for each answer, allowing us to measure convergence (ii) compute an average competence of agents for each level of convergence occurring within that group.
+Across all groups, we then compute the average of the average within group competence for each degree of convergence.
+
+We repeat this procedure varying population competence distributions, and either (a) the size of informant groups, or (b) the number of choice options.
+If we vary the group size, we hold the number of choice options constant at n = 3, and vice versa when varying the number of choice options.
+Fig.
+\@ref(fig:simulation-accuracy-sample-categorical) shows the average accuracy, and Fig.
+\@ref(fig:simulation-competence-sample-categorical) the average competence value for each share of votes, and for different competence levels and varying the group size (see Appendix \@ref(simulations) for results varying the number of choice options instead of group size).
+
+(ref:simulation-accuracy-sample-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and sample sizes.
+Points represent averages across all simulations within the respective segment.
+The number of choice options is three.
+
+```{r simulation-accuracy-sample-categorical, fig.cap="(ref:simulation-accuracy-sample-categorical)", out.width="75%"}
+plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
+                                       outcome = "Accuracy", variable = sample)
+```
+
+(ref:simulation-competence-sample-categorical) Competence as a function of vote share for an option, for different population competence distributions and sample sizes.
+Points represent averages across all simulations within the respective segment.
+The number of choice options is three.
+
+```{r simulation-competence-sample-categorical, fig.cap="(ref:simulation-competence-sample-categorical)", out.width="75%"}
+plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
+                                       outcome = "Competence", variable = sample)
+```
+
+The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows:
+(i) Irrespective of group size and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate. 
+(ii) For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents.
+The strength of this relation is not much affected by the number of agents whose answers are converging, but, although it is always positive, it ranges from very weak to very strong depending on the initial competence distribution.
+(iii) The relation between convergence and accuracy is always much stronger than the relation between convergence and average competence of the agents.
 
@@
 
-We test whether people infer that the members of a group are correct, and thus likely competent, when their answers converge in two series of experiments.
-In all experiments, participants see the estimates or choices of different individuals, which we will call informants.
-Participants were were not given any information about the estimation tasks and the informants - they had no priors.
-
-In the first series (experiment 1 to 3), the scenario is about numerical estimates. In the second series (experiment 4 to 6), the scenario is about categorical choices.
-
-In experiment 1 (numerical) and 4 (categorical), participants evaluated several groups of informants, with each informant providing an answer independently.
-We manipulated convergence, i.e how close estimates were (Exp.
-1, numerical setting), or how consensual an answer was (Exp.4, categorical setting).
-We found that the more convergence, the more participants inferred accuracy and competence.
-
-In experiments 3 (numerical) and 5 (categorical), we tested whether cues of informational dependency inferences hamper inferences from convergence.
-Besides convergence, we manipulated whether informants were independent (mimicking experiments 1 and 4) or shared a conflict of interest.
-We found that (i) participants inferred accuracy and competence from convergence even when informants were in a conflict of interest situation, but (ii) that these inference are more enhanced when informants are independent.
-Using a slightly different experimental setting, we replicate our results from experiments 1 and 4.
-
-In experiment 2, we tested a different form of informational dependency: informants were either independent informants or "had discussed at great length with each other" before making their estimate.
-We did not find a difference between the two conditions.
-This experiment was the only one in which we did not manipulate convergence - all informant groups were convergent.
-
-In experiment 5, we tested whether more choice options (ten vs.
-three) enhance inferences from convergence.
-We reasoned that with more options, consensus is less likely to occur by chance, and thus people might find it more impressive.
-We did not find that the number of choice options altered effects of convergence.
+We test whether people infer that individuals are more likely to give accurate answers, and to be competent, when their answers converge, both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6).
+By contrast with previous experiments, participants were not were not given any information about the tasks–how difficult they were–and the informants–how competent they might be.
+We also manipulated (Experiments 2, 3, and 5) whether the convergence of the answers could be explained by something else than accuracy, which should reduce participants’ reliance on the convergence of the answer as a cue to accuracy and to competence.
 
@@
 All analyses were conducted in R (version 4.2.2) using R Studio.
+For most statistical models, we relied on the `lme4` package and its `lmer()` function.
 

### 2024-01-18 appendix_simulations.Rmd
before changes to format of inline reporting of results

@@
 ## Numerical choice context
-
-```{r read-simulation-data-numeric, message=FALSE}
-# read data from simulations
-simulation_numeric <- read_csv("Experiment_1/data/sim.csv") %>% 
-  # remove alpha and beta values from competence variable strings
-  mutate(competence = sub("\n.*", "", competence))
-```
 
@@
 
-(ref:competence-distributions) The different population competence distributions we considered in our simulations.
-
-```{r competence-distributions, fig.cap="(ref:competence-distributions)", out.width= "75%"}
-plot_competence_distributions(simulation_numeric)
-```
-
 We draw an estimate for each individual based on their respective competence distribution.
@@
 
-(ref:simulation-numeric-correlations) Correlation between accuracy (left) or competence (right) and convergence, as a function of sample size, grouped by the data underlying competence distributions.
-
-```{r simulation-numeric-correlations, fig.cap="(ref:simulation-numeric-correlations)", out.width="75%"}
-# calculate correlations
-simulation_numeric <- simulation_numeric %>% 
-  group_by(competence, sample) %>% 
-  mutate(correlation_accuracy = cor(accuracy_mean, convergence), 
-         correlation_competence = cor(competence_mean, convergence)) %>% 
-  ungroup()
-
-# accuracy plot
-accuracy_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_accuracy, color = competence)) +
-  geom_point() + 
-  geom_line(alpha = 0.5) + 
-  labs(y = "Correlation: Accuracy/Convergence", 
-       x = "Sample size", 
-       color = "Competence distribution") +
-  scale_y_continuous(limits = c(0, 1)) +
-  scale_color_brewer(palette = "Set1") +
-  plot_theme + 
-  theme(
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 8))
-
-# competence plot
-competence_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_competence, color = competence)) +
-  geom_point() + 
-  geom_line(alpha = 0.5) + 
-  labs(y = "Correlation: Competence/Convergence", 
-       x = "Sample size", 
-       color = "Competence distribution") +
-  scale_y_continuous(limits = c(0, 1)) +
-  scale_color_brewer(palette = "Set1") +
-  plot_theme + 
-  theme(
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 8))
-
-ggarrange(accuracy_plot, competence_plot, common.legend = TRUE)
-```
-
 ## Categorical choice context
-
-```{r read-simulation-data-categorical, message=FALSE}
-# read data from simulations
-simulation_sample_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_sample_3_options.csv")
-
-simulation_options_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_options_3_sample.csv")
-```
 
@@
 For competence, we observe a similar sigmoid-like relationship, but of lesser amplitude and varying considerably as a function of the underlying population competence distributions.
-
-(ref:simulation-accuracy-sample-categorical) Accuracy as a function of vote share for an option, for different population competence distributions and sample sizes.
-Points represent averages across all simulations within the respective segment.
-The number of choice options is three.
-
-```{r simulation-accuracy-sample-categorical, fig.cap="(ref:simulation-accuracy-sample-categorical)", out.width="75%"}
-plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
-                                       outcome = "Accuracy", variable = sample)
-```
-
-(ref:simulation-competence-sample-categorical) Competence as a function of vote share for an option, for different population competence distributions and sample sizes.
-Points represent averages across all simulations within the respective segment.
-The number of choice options is three.
-
-```{r simulation-competence-sample-categorical, fig.cap="(ref:simulation-competence-sample-categorical)", out.width="75%"}
-plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
-                                       outcome = "Competence", variable = sample)
-```
 

### 2024-01-18 functions/own_functions.R

@@ -75,7 +75,7 @@ plot_competence_distributions <- function(data) {
   
   # Arrange and display the plots
   competence_plot <- ggarrange(plotlist = plot_list) %>% 
-    annotate_figure(figure,
+    annotate_figure(
                     left = textGrob("Density", rot = 90, vjust = 1, gp = gpar(cex = 1)),
                     bottom = textGrob("Competence", gp = gpar(cex = 1)))
   

### 2024-01-22 appendix_analytical.Rmd
add latest version of analytical argument Benoit

# Analytical argument (#analytical)

Agents answer a categorical question, with m+1 answers. We define the competence p of an individual as the probability to choose the right answer. Each of the wrong answers has the same probability (1-p)/m to be chosen. We observe the answers of a population of n+1 agents, with diverse levels of competence drawn from a beta distribution -- a flexible probability distribution that can be uniform, unimodal or bimodal, depending on the parameters. We assume that the shape parameters of the beta distribution ($\alpha$ and $\beta$) are known, but the right answer is not. We aim to infer the competence of an individual, based on the population answers. In our analytical model, we use Bayes formula to compute the posterior competence distribution of competence for an individual, knowing that a number K of other participants agree with him. The posterior distribution is a mixture of two beta distributions, Beta($\alpha$+1, $\beta$) and Beta($\alpha$, $\alpha$+1), weighted respectively by the probabilities that K individuals would choose the right or the wrong answer. The model suggests that the more consensual an individual's vote is, the more that individual is estimated to be competent (Fig. \@ref(fig:figure-Benoit)).

(ref:figure-Benoit) Average competence of an informant as a function of K, the number of individuals who agree. In this example, we assume that the number of informants - not counting the focal informant - in a sample is n=10, m=5, and that the population competence ($\alpha$ = $\beta$ = 1).

```{r figure-Benoit, echo=FALSE, fig.cap="(ref:figure-Benoit)", out.width= "50%", fig.align="left"}
knitr::include_graphics("figures/figure_Benoit.png")
```

That said, the average competence is bounded between and $\frac{\alpha}{\alpha+\beta+1}$ and $\frac{\alpha+1}{\alpha+\beta+1}$. In other words, even if all other individuals agree with an informant, a case in which the informant very likely chose the correct answer (add an accuracy figure, maybe?), the competence we should assign to this informant, according to our model, is not higher than $\frac{\alpha+1}{\alpha+\beta+1}$ (or $2/3$ in our example in Fig. \@ref(fig:figure-Benoit)). This counter-intuitive result is due to our framework: we assume that a certain competence distribution as a given, and only observe one decision per individual. Intuitively, in a world where even some highly incompetent people sometimes pick the correct answer by chance, being right (with the majority) once is not a sufficient signal of highest competence.

### 2024-03-26 Cogsci_version_raw.Rmd
major changes before adding analytical part

---
title             : "How wise is the crowd? \n Can we infer people are accurate and competent merely because they agree with each other?"
shorttitle        : "How wise is the crowd?"

header-includes:  | # to prevent floats from moving past certain points (for the appendix)
  \usepackage{placeins}

author: 
  - name          : "Anonymous"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : " "

authornote: |

abstract: |
  Are people who agree on something more likely to be right and competent? Evidence suggests that people tend to make this inference. However, standard wisdom of crowds approaches only provide limited normative grounds for this behavior. Using simulations, we argue that in stylized scenarios where individuals make independent and unbiased estimates, more convergent groups of individuals indeed tend to be more competent and accurate. Mirroring the stylized setting of the simulations, we then show that people make these inferences in two experiments. These inferences from convergence might help explain why people respect scientists’ competence, even if they do not understand much about how scientific results are reached.

keywords          : consensus; convergence; majority rules; wisdom of crowds; trust; trust in science

floatsintext      : yes
linenumbers       : no 
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "doc" # "doc" for nice look, "man" for manuscripty
output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors. For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g. set always allow html as true in the yaml heading

always_allow_html: true

appendix:

bibliography: references.bib
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  # fig.retina = 3, dpi = conditional_dpi,
  # fig.width = 7, fig.asp = 0.618,
  # fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  },
  options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)
```

```{r packages, include=FALSE}
# load required packages
library("papaja")      # For APA style manuscript   
library("lme4")        # model specification / estimation
library("lmerTest")    # provides p-values in the output
library("tidyverse")   # data wrangling and visualisation
library("afex")        # anova and deriving p-values from lmer
library("broom")       # extracting data from model fits 
library("broom.mixed") # extracting data from mixed models
library("metafor")     # doing mata analysis
library("metaviz")     # vizualization of meta analysis
library("stringr")     # for dmetar p-curve function to work
library("poibin")      # for dmetar p-curve function to work
library("patchwork")   # put several plots together
library("ggridges")    # for plots
library("gghalves")    # for plots
library("ggbeeswarm")  # Special distribution-shaped point jittering
library("knitr")       # for tables
library("kableExtra")  # also for tables
library("sf")          # for maps
library("ggpubr")      # for combining plots with ggarrange() 
library("viridis")     # for generating color palette for map 
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
```

```{r functions}
# load plot theme
source("functions/plot_theme.R")

# load other functions
source("functions/own_functions.R")
```

***This is a "raw" version. The final version was edited in overleaf to meet formatting criteria***

# Introduction

Imagine that you live in ancient Greece, and a fellow called Eratostenes claims the circumference of the earth is 252000 stades (approximately 40000 kilometers). You know nothing about this man, the circumference of the Earth, or how one could measure such a thing. As a result, you discard Eratostenes' opinion and (mis)take him for a pretentious loon. But what if other scholars had arrived at very similar measurements, independently of Eratosthenes? Or even if they had carefully checked his measurement, with a critical eye? Wouldn't that give you enough ground to believe not only that the estimates might be correct, but also that Eratosthenes and his fellow scholars must be quite bright, to be able to achieve such a feat as measuring the Earth?

In this article, we explore how, under some circumstances, we should, and we do infer that a group of individuals whose answers converge are likely to be correct, and to be competent in the relevant area, even if we had no prior belief about either what the correct answer was, or about these individuals' competence. There are two ways for answers to converge: when the answers are continuous, they can have a smaller range or a lower variance; when the answers are categorical (i.e. there are only a few, often two, options), more can agree on the same answer. These two cases have been studied with different paradigms.

In the continuous case, the most relevant evidence comes from the literature on 'advice-taking.' In these experiments, participants are called 'judges,' and they need to make numerical estimates--sometimes on factual knowledge, e.g. 'What year was the Suez Canal opened first? [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g. 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020]. To help answer these questions, participants are given estimates from others, the advisors.

One subset of these studies manipulate the degree of convergence between groups of advisors, through the variance of estimates [@mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009], or their range [@budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007]. These studies find that participants are more confident about, or rely more on, estimates from groups of advisors who converge more.

Other studies varied the degree of convergence within a group of advisors. They present participants with a set of estimates where some estimates are close to each other (or overlapping, in cases where estimates were intervals), while others are outliers [@harriesCombiningAdviceWeight2004; @yanivWeightingTrimmingHeuristics1997, study 3 & 4]. These studies find that participants heavily discount outliers when aggregating estimates.

In categorical choice contexts, there is ample and long-standing [e.g. @crutchfieldConformityCharacter1955] evidence from social psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger (both in absolute and in relative terms) [e.g., @morganEvolutionaryBasisHuman2012; for review, see @mercierMajorityRulesHow2019]. This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role [e.g. because the answers are private, see @aschStudiesIndependenceConformity1956]. Similar results have been obtained with young children [e.g. @fusaroChildrenAssessInformant2008; @corriveauGoingFlowPreschoolers2009; @bernardChildrenWeighNumber2015a; @bernardFourSixYearOldChildren2015; @chenChildrenTrustConsensus2013; @herrmannStickScriptEffect2013; @morganDevelopmentAdaptiveConformity2015].

If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the individuals' competence. One potential exception is a study with preschoolers [@corriveauGoingFlowPreschoolers2009] in which the children were more likely to believe the opinion of an informant who had previously been the member of a majority over that of an informant who had dissented from the majority. However, it is not clear whether the children thought the members of the majority were particularly competent, since their task--naming an object--was one in which children should already expect a high degree of competence from (adult) informants. This result might thus indicate simply that children infer that someone who disagrees with several others on how to call something is likely wrong, and thus likely less competent at least in that domain.

## Is inferring competence from convergence justified?

```{r read-simulation-data-numeric, message=FALSE}
# read data from simulations
simulation_numeric <- read_csv("Experiment_1/data/sim.csv") %>% 
  # remove alpha and beta values from competence variable strings
  mutate(competence = sub("\n.*", "", competence))
```

```{r read-simulation-data-categorical, message=FALSE}
# read data from simulations
simulation_sample_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_sample_3_options.csv")  %>% 
  # remove alpha and beta values from competence variable strings
  mutate(competence = sub("\n.*", "", competence))

simulation_options_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_options_3_sample.csv")  %>% 
  # remove alpha and beta values from competence variable strings
  mutate(competence = sub("\n.*", "", competence))
```

To provide a normative answer, we conducted simulations, both for a continuous and a categorical scenario.

### Continuous case

In these simulations, groups of agents, with each agent holding varying degrees of competence, provide numerical answers. We measure how accurate these answers are, and how much they converge (i.e. how low their variance is). We then correlate the degree of convergence with the accuracy and with the competence of the agents.

More specifically, agents provide an estimate on a scale from 1000 to 2000 (chosen to facilitate the experiments below). Each agent is characterized by a normal distribution of possible answers. All of the agents' distributions are centered around the correct answer, but their standard deviation varies, denoting varying degrees of competence. The agents' standard deviation varied from 1 (highest competence) to 1000 (lowest competence). Each agent's competence level is drawn from a population competence distribution, expressed by a beta distribution, which can take different shapes. We conducted simulations with a variety of beta distributions which cover a large range of possible competence populations (see Fig. \@ref(fig:simulations) C).

In the simulation, a population of around 990000 agents with different competence levels is generated. An answer is drawn for each agent, based on their respective competence distribution. The accuracy of this answer is defined as the squared distance to the true answer. Having a competence and an accuracy value for each agent, we randomly assign agents to groups of, e.g., three. For each group, we calculate the average of the agents' competence and accuracy. We measure the convergence of a group by calculating the standard deviation of the agents' answers. We repeat this process for different sample sizes for the groups, and different competence distributions. Fig. \@ref(fig:simulations) B displays the relation of convergence with accuracy (left), and competence (right) across different underlying competence distributions and group sizes.

### Categorical case

We simulate agents who have to decide between a number of options, one of which is correct, and whose competence varies. Competence is defined as a value p, which can differ for each agent, and which corresponds to the probability of selecting the right answer (the agents then have a probability (1-p)/(m-1), with m being the number of options, of selecting any other option). Competence values range from chance level (p = 1/m) to always selecting the correct option (p = 1). Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see see Fig. \@ref(fig:simulations) C). Based on their competence level, we draw an answer for each agent. We measure an agent's accuracy as a binary outcome, namely whether they selected the correct option or not. In each simulation 999000 agents are generated, and then randomly assigned to groups (of varying size in different simulations). Within these groups, we first calculate the share of individuals voting for each answer, allowing us to measure convergence. For each level of convergence occurring within a group, we then compute (i) the average accuracy and (ii) the average competence of agents. Across all groups, we then compute the averages of these within-group averages, for each level of convergence.

We repeat this procedure varying population competence distributions, and the size of informant groups, holding the number of choice options constant at n = 3. Fig. \@ref(fig:simulations) A shows the average accuracy (left), and the average competence (right) value for each share of votes, across different underlying competence distributions and group sizes.

The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows: (i) Irrespective of group size and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate. (ii) For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents. The strength of this relation is not much affected by the number of agents whose answers are converging, but, although it is always positive, it ranges from very weak to very strong depending on the initial competence distribution. (iii) The relation between convergence and accuracy is always much stronger than the relation between convergence and average competence of the agents.

(ref:simulations) **A** Shows the results of simulations in a categorical setting with 3 choice options. Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes. **B** Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left)/competence (right) and convergence (measured by squared distance to true mean and reversed such that greater convergence corresponds to being more right on the x-axis). **C** Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1. In the categorical simulations, a competence value of 0 corresponds to chance (e.g. in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer.

```{r simulations, fig.cap="(ref:simulations)", fig.height= 10, fig.width=10}
# Plot continuous choice scenario
plot_continuous_accuracy <- plot_continuous_cogsci(simulation_numeric, outcome = "accuracy") + 
  rremove("xlab") +
  guides(color = "none") +
  labs(caption = NULL) 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plot_continuous_competence <- plot_continuous_cogsci(simulation_numeric, outcome = "competence") + 
  rremove("xlab") +
  guides(color = "none") +
  labs(caption = NULL) 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plots_continuous <- ggarrange(plot_continuous_accuracy, plot_continuous_competence) %>% 
  annotate_figure(bottom = textGrob("Convergence (SDs of samples)", gp = gpar(cex = 1, fontface = "bold")))

# Plot categorical choice scenario
plot_categorical_accuracy <- plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Accuracy", variable = sample) + 
  rremove("xlab") +
  labs(caption = NULL) 
  # theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plot_categorical_competence <- plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Competence", variable = sample) + 
  rremove("xlab") +
  labs(caption = NULL) 
  # theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plots_categorical <- ggarrange(plot_categorical_accuracy, plot_categorical_competence, common.legend = TRUE) %>% 
  annotate_figure(bottom = textGrob("Convergence (share of votes for an option)", gp = gpar(cex = 1, fontface = "bold")))

# common plot
patchwork <-  plots_categorical / plots_continuous / (plot_competence_distributions(simulation_numeric) + plot_spacer())

patchwork + 
  plot_annotation(tag_levels = 'A')

```

## Overview experiments

We test whether people infer that individuals--we will call them informants henceforth--are more likely to give accurate answers, and to be competent, when their answers converge, both in a continuous tasks (Experiment 1), and in a categorical tasks (Experiment 2). By contrast with previous experiments, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be. We also manipulated whether the convergence of the answers could be explained by something else than accuracy, which should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.

All experiments and analyses were pre-registered, unless explicitly said so. Pre-registration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b) (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the `lme4` package and its `lmer()` function. Note that on the OSF, the enumeration of experiments differs, since we report only a subset of experiments here (pattern: article at hand ~ OSF; Exp.1 ~ Exp.3; Exp. 2 ~ Exp.5)

# Experiment 1

```{r exp3}
# Analyze data of experiments and store results

# Experiment 3

# read data
exp3 <- read_csv("Experiment_3/data/cleaned.csv") %>% 
  # import experimental variable as factors and set levels
  mutate(
    across(c(convergence, independence), ~as.factor(.x)),
    convergence = fct_relevel(convergence, "divergent", "convergent"),
    independence = fct_relevel(independence, "conflict", "independent")
  )

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the independence condition
exp3_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence |ID), 
                           data = exp3 %>% filter(independence == "independent"))
exp3_model_h1_competence = lmer(competence ~ convergence + (1 + convergence |ID), 
                             data = exp3 %>% filter(independence == "independent"))

# H2
# run mixed model (random intercept and slope for convergence) with participants as random factor
exp3_model_accuracy = lmer(accuracy ~ convergence + independence + 
                            convergence*independence + (1 + convergence | ID), 
                           data = exp3)

exp3_model_competence = lmer(competence~ convergence + independence + 
                            convergence*independence + (1 + convergence | ID), 
                           data = exp3)

# extract descriptives for inline reporting
exp3_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp3$ID),
  gender = exp3 %>% group_by(gender) %>% summarize(n = n_distinct(ID)) %>% split(.$gender),
  age = exp3 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_h1_accuracy = text_ready(exp3_model_h1_accuracy), 
  model_h1_competence = text_ready(exp3_model_h1_competence),
  model_accuracy = text_ready(exp3_model_accuracy), 
  model_competence = text_ready(exp3_model_competence),
  # Means
  means_H1 = exp3 %>% 
    filter(independence == "independent") %>% 
    group_by(convergence) %>% 
    summarise(across(c(accuracy, competence), list(mean = mean, sd = sd))) %>% 
    rounded_numbers() %>% 
    split(.$convergence), 
  means_H2 = exp3 %>% 
    group_by(convergence, independence) %>% 
    summarise(across(c(accuracy, competence), list(mean = mean, sd = sd))) %>% 
    rounded_numbers() %>% 
    super_split(convergence, independence)
)

# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? 
exp3_explorative <- exp3 %>% 
  pivot_longer(
    c(accuracy, competence),
    names_to = "outcome", 
    values_to = "score") %>% 
  # effect code variables
  mutate(convergence_effect_code = recode(convergence, "divergent" = -0.5, 
                                          "convergent" = 0.5))

# run mixed model (random slope and intercept) with participants as random factor
exp3_exploratory_analysis <- lmer(score ~ convergence_effect_code + outcome + 
                               convergence_effect_code*outcome + independence +
                               (1 + convergence_effect_code + outcome | ID), 
                             data = exp3_explorative) %>% 
  text_ready()
```
In Experiment 1, participants were told that they would be looking at (fictional) predictions of experts for stock values. They were provided with a set of numerical estimates which were more or less convergent, and asked whether they thought the estimates were accurate, and whether the experts making the estimates were competent. In a conflict of interest condition, the experts had an incentive to value the stock in a given way, while they had no such conflict of interest in the independence condition. We formulated four hypotheses:

***H1a: Participants perceive predictions of independent informants as more accurate when they converge compared to when they diverge.***

***H1b: Participants perceive independent informants as more competent when their predictions converge compared to when they diverge.***

***H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are in a conflict of interest.***

***H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are in a conflict of interest.***

## Methods

### Participants

We ran a power analysis by simulation. The simulation code is available on OSF, and the procedure is described in the pre-registration document. The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% of the time, given an alpha threshold for significance of 0.05. Due to uncertainty about our effect size assumptions and because we had resources for a larger sample, we recruited `r exp3_descriptives$n_subj` participants for this study, from the UK and via Prolific (`r exp3_descriptives$gender$female$n` female, `r exp3_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp3_descriptives$age$mean`, $age_\text{sd}$: `r exp3_descriptives$age$sd`, $age_\text{median}$: `r exp3_descriptives$age$median`).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "You will see four scenarios in which several experts predict the future value of a stock. You have no idea how competent the experts are. It's also possible that some are really good while others are really bad. As it so happens, in all scenarios, the predictions for the value of the stock have to lie between 1000 and 2000. Other than that, the scenarios are completely unrelated: it is different experts predicting the values of different stocks every time." Participants then saw four scenarios (Fig. \@ref(fig:stimuli) A), each introduced by a text according to the condition the participant was assigned to (see 'Design'). Participants were then asked to rate the experts' accuracy ("On average, how accurate do you think these three predictions are?" on a 7-point Likert scale - (1) "not accurate at all" to (7) "extremely accurate"), and their competence ("On average, how good do you think these three experts are at predicting the value of stocks?" - (1) "not good at all" to (7) "extremely good").

(ref:stimuli) **A** Example of stimuli as used in Experiment 1 (independence condition). **B** Example of stimuli as used in Experiment 2. In these experiments, we treated convergence as a continuous variable, assigned the values in brackets to the respective conditions.

```{r stimuli, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimuli)"}
# Function to create a plot with image and title
create_image_plot <- function(image_path, title) {
  img <- rasterGrob(readPNG(image_path), interpolate = TRUE)
  title_grob <- textGrob(title, gp = gpar(fontsize = 11, fontface = "bold"), just = "left")

  # Combine image and title in a single plot
  plot <- ggplot() +
    annotation_custom(img, xmin = 0, xmax = 1, ymin = 0, ymax = 1) +
    annotation_custom(title_grob, xmin = 0, xmax = 0.2, ymin = 1, ymax = 1) +
    theme_void()  # Set the panel background to be transparent +
    theme(plot.margin = margin(0, 0, 0, 0, "cm"))  # Set margins to zero

  return(plot)
}

# Experiment 1

# File paths to four images
image_paths <- c(
  "figures/example_stimulus_exp3_convergent.png",
  "figures/example_stimulus_exp3_divergent.png"
)

# Titles for the images
image_titles <- c(
  "convergent", 
  "divergent"
)

# Create a list of plots
stimuli_exp1 <- lapply(seq_along(image_paths), function(i) {
  create_image_plot(image_paths[i], image_titles[i])
})

# Experiment 4

# File paths to four images
image_paths <- c(
  "Experiment_4/figures/stimuli/large_opp_majority_a.png",
  "Experiment_4/figures/stimuli/large_divergence_a.png",
  "Experiment_4/figures/stimuli/large_majority_a.png",
  "Experiment_4/figures/stimuli/large_consensus_a.png"
)

# Titles for the images
image_titles <- c(
  "minority (0)", 
  "dissensus (1)", 
  "majority (2)", 
  "consensus (3)"
)

# Create a list of plots
stimuli_exp4 <- lapply(seq_along(image_paths), function(i) {
  create_image_plot(image_paths[i], image_titles[i])
})

exp1_plot <- ggarrange(plotlist = stimuli_exp1, ncol = 1)
exp4_plot <- ggarrange(plotlist = stimuli_exp4, ncol = 2, nrow = 2)

# Combine the plots with the new annotations
patch <- exp1_plot | exp4_plot

patch <- patch + plot_layout(tag_level = 'new')

patch + 
  plot_annotation(tag_levels = 'A')

```

(ref:stimuli-exp1) **A** Example of stimuli as used in Experiment 1.

```{r stimuli-exp1, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimuli-exp1)"}
# Function to create a plot with image and title
create_image_plot <- function(image_path, title) {
  img <- rasterGrob(readPNG(image_path), interpolate = TRUE)
  title_grob <- textGrob(title, gp = gpar(fontsize = 11, fontface = "bold"), just = "left")

  # Combine image and title in a single plot
  plot <- ggplot() +
    annotation_custom(img, xmin = 0, xmax = 1, ymin = 0, ymax = 1) +
    annotation_custom(title_grob, xmin = 0, xmax = 0.1, ymin = 1, ymax = 1.05) +
    theme_void()  # Set the panel background to be transparent +
    theme(plot.margin = margin(0, 0, 0, 0, "cm"))  # Set margins to zero

  return(plot)
}

# Experiment 1

# File paths to four images
image_paths <- c(
  "figures/example_stimulus_exp3_convergent.png",
  "figures/example_stimulus_exp3_divergent.png"
)

# Titles for the images
image_titles <- c(
  "convergent", 
  "divergent"
)

# Create a list of plots
stimuli_exp1 <- lapply(seq_along(image_paths), function(i) {
  create_image_plot(image_paths[i], image_titles[i])
})

exp1_plot <- ggarrange(plotlist = stimuli_exp1, ncol = 1)

exp1_plot
```

(ref:stimuli-exp2) **B** Example of stimuli as used in Experiment 2.

```{r stimuli-exp2, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimuli-exp2)"}
# Function to create a plot with image and title
create_image_plot <- function(image_path, title) {
  img <- rasterGrob(readPNG(image_path), interpolate = TRUE)
  title_grob <- textGrob(title, gp = gpar(fontsize = 11, fontface = "bold"), just = "left")

  # Combine image and title in a single plot
  plot <- ggplot() +
    annotation_custom(img, xmin = 0, xmax = 1, ymin = 0, ymax = 1) +
    annotation_custom(title_grob, xmin = 0, xmax = 0.2, ymin = 1, ymax = 1) +
    theme_void()  # Set the panel background to be transparent +
    theme(plot.margin = margin(0, 0, 0, 0, "cm"))  # Set margins to zero

  return(plot)
}

# Experiment 4

# File paths to four images
image_paths <- c(
  "Experiment_5/figures/stimuli/large_minority_a.png",
  "Experiment_5/figures/stimuli/large_divergence_a.png",
  "Experiment_5/figures/stimuli/large_majority_a.png",
  "Experiment_5/figures/stimuli/large_consensus_a.png"
)

# Titles for the images
image_titles <- c(
  "minority (0)", 
  "dissensus (1)", 
  "majority (2)", 
  "consensus (3)"
)

# Create a list of plots
stimuli_exp4 <- lapply(seq_along(image_paths), function(i) {
  create_image_plot(image_paths[i], image_titles[i])
})

exp4_plot <- ggarrange(plotlist = stimuli_exp4, ncol = 2, nrow = 2)

exp4_plot

```

### Design

We manipulated two factors: informational dependency (two levels, independence and conflict of interest; between participants) and convergence (two levels, convergence and divergence; within participants). In the independence condition, the participants read "Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock." In the conflict of interest condition, the participants read "All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future."

### Materials

We generated sets of estimates from uniform distributions with varying range (60 for convergence, 600 for divergence; estimate scale ranged from 1000 to 2000). The same sets of estimates were used in the conflict of interest and in the independence condition. Each participant rated four scenarios, two for each level of convergence. More information on how the stimuli were created can be found on the OSF.

## Results and discussion

To account for dependencies of observations due to our within-participant design regarding convergence, we ran mixed models, with participants as random factor for the intercept and the convergence slope. We find evidence for all four hypotheses. For the first set of hypotheses, we reduced the sample to half of the participants, namely those who were assigned to the independence condition. As for accuracy, participants rated informants in convergent scenarios (mean = `r exp3_descriptives$means_H1$convergent$accuracy_mean`, sd = `r exp3_descriptives$means_H1$convergent$accuracy_sd`) as more accurate than in divergent ones (mean = `r exp3_descriptives$means_H1$divergent$accuracy_mean`, sd = `r exp3_descriptives$means_H1$divergent$accuracy_sd`; $\hat{b}_{\text{Accuracy}}$ = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$estimate` `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$ci`, p = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$p.value`). As for competence, participants rated informants in convergent scenarios (mean = `r exp3_descriptives$means_H1$convergent$competence_mean`, sd = `r exp3_descriptives$means_H1$convergent$competence_sd`) as more competent than in divergent ones (mean = `r exp3_descriptives$means_H1$divergent$competence_mean`, sd = `r exp3_descriptives$means_H1$divergent$competence_sd`; $\hat{b}_{\text{Competence}}$ = `r exp3_descriptives$model_h1_competence$convergenceconvergent$estimate` `r exp3_descriptives$model_h1_competence$convergenceconvergent$ci`, p = `r exp3_descriptives$model_h1_competence$convergenceconvergent$p.value`).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \@ref(fig:experiments) A). In the independence condition, the effect of convergence on accuracy was more positive ($\hat{b}_{\text{interaction, Accuracy}}$ = `r exp3_descriptives$model_accuracy$interaction$estimate` `r exp3_descriptives$model_accuracy$interaction$ci`, p = `r exp3_descriptives$model_accuracy$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ = `r exp3_descriptives$model_accuracy$convergenceconvergent$estimate` `r exp3_descriptives$model_accuracy$convergenceconvergent$ci`, p = `r exp3_descriptives$model_accuracy$convergenceconvergent$p.value`). Likewise the effect of convergence on competence was more positive ($\hat{b}_{\text{interaction, Competence}}$ = `r exp3_descriptives$model_competence$interaction$estimate` `r exp3_descriptives$model_competence$interaction$ci`, p = `r exp3_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ = `r exp3_descriptives$model_competence$convergenceconvergent$estimate` `r exp3_descriptives$model_competence$convergenceconvergent$ci`, p = `r exp3_descriptives$model_competence$convergenceconvergent$p.value`).

In an exploratory, not pre-registered analysis, we tested whether the effect of convergence is bigger on accuracy than on competence. To do so, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction effect, indicating that pooled across independent and conflict of interest conditions, the effect of convergence had a smaller effect on competence than on accuracy ($\hat{b}$ = `r exp3_exploratory_analysis$interaction$estimate` `r exp3_exploratory_analysis$interaction$ci`, `r exp3_exploratory_analysis$interaction$p.value`).

Experiment 1 shows that, believed sets of estimates to have been more accurate, and the individuals who had made them to be more competent, participants were more confident when the estimates were more convergent, when the estimates were more convergent. These inferences were less pronounced when the informants are systematically biased.

# Experiment 2

```{r exp5}
# Analyze data of experiments and store results

# Experiment 5

# read data
exp5 <- read_csv("Experiment_5/data/cleaned.csv") %>% 
  # import experimental variable as factor and set levels
  mutate(
    # set levels for `convergence`
    convergence_categorical = recode_factor(convergence, 
                                            `0` = "minority", 
                                            `1` = "divergence", 
                                            `2` = "majority", 
                                            `3` = "consensus",
                                            .default = NA_character_))

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the independence condition
exp5_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence | id), 
                              data = exp5 %>% filter(independence == "independent"))
exp5_model_h1_competence = lmer(competence ~ convergence + (1 + convergence | id), 
                              data = exp5 %>% filter(independence == "independent"))

# H2 
# run mixed model (random slope and intercept) with participants as random factor
exp5_model_accuracy = lmer(accuracy ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = exp5)
exp5_model_competence = lmer(competence ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = exp5)

# extract descriptives for inline reporting
exp5_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp5$id),
  gender = exp5 %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp5 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_h1_accuracy = text_ready(exp5_model_h1_accuracy), 
  model_h1_competence = text_ready(exp5_model_h1_competence),
  model_accuracy = text_ready(exp5_model_accuracy), 
  model_competence = text_ready(exp5_model_competence)
)

# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? 
exp5_explorative <- exp5 %>% 
  mutate(accuracy_scaled = scale(accuracy), 
         competence_scaled = scale(competence)) %>% 
  pivot_longer(
    c(accuracy_scaled, competence_scaled),
    names_to = "outcome", 
    values_to = "score")

# run mixed model (random slope and intercept) with participants as random factor
exp5_exploratory_analysis <- lmer(score ~ convergence + outcome + 
                               convergence*outcome + independence +
                               (1 + convergence | id), 
                             data = exp5_explorative) %>% 
  text_ready()
```

Experiment 2 is a conceptual replication of Experiment 1 in a categorical instead of a numerical case:

***H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.***

***H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.***

***H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).***

***H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).***

## Methods

### Participants

We ran two different power analyses, one for each outcome variable. inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. The power simulation for accuracy suggested that for 80 participants, we would have a power of at least 90% for the interaction effect. The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we would also detect a main effect of convergence. Due to uncertainty about our assumptions and because resources were available for a larger sample, we recruited `r exp5_descriptives$n_subj` participants, again in the UK and via Prolific (`r exp5_descriptives$gender$female$n` female, `r exp5_descriptives$gender$male$n`, 1 non-identified; $age_\text{mean}$: `r exp5_descriptives$age$mean`, $age_\text{sd}$: `r exp5_descriptives$age$sd`, $age_\text{median}$: `r exp5_descriptives$age$median`).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "We will show you three financial advisors who are giving recommendations on investment decisions. They can choose between three investment options. Their task is to recommend one. You will see several such situations. They are completely unrelated: it is different advisors evaluating different investments every time. At first you have no idea how competent the advisors are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be difficult while others are easy. Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are." They were then presented with eight recommendation scenarios (Fig. \@ref(fig:stimuli)). To assess perceptions of accuracy, we asked: "What do you think is the probability of advisor 1 making the best investment recommendation?". Participants answered with a slider on a scale from 0 to 100. To assess perceptions of accuracy, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

### Design

We manipulated convergence within participants, by varying the ratio of players choosing the same response as a focal player (i.e. the one that participants evaluate). The levels of convergence are: (i) consensus, where all three players pick the same option [`coded value = 3`]; (ii) majority, where either the third or second player picks the same option as the first player [`coded value = 2`]; (iii) dissensus, where all three players pick different options [`coded value = 1`]; (iv) minority, where the second and third player pick the same option, but one that is different from the first player's choice [`coded value = 0`]. In our analysis, we treat convergence as a continuous variable, assigning the coded values in squared parenthesis here. We manipulated conflict of interest between participants. In the conflict of interest condition, experts were introduced this way: "The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations." For the independence condition: "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations."

### Materials

All participants saw all four conditions of convergence (Fig. \@ref(fig:stimuli) B), with two stimuli per condition, i.e. eight stimuli in total (4 convergence levels x 2 stimuli).

## Results and discussion

To account for dependencies of observations due to our within-participant design regarding convergence, we ran mixed models, with participants as random factor for the intercept and the convergence slope.

We find evidence for all four hypotheses. To test H1a and H1b, we restricted our data on the independent condition. We find a positive effect of convergence on both accuracy ($\hat{b}_{\text{Accuracy}}$ = `r exp5_descriptives$model_h1_accuracy$convergence$estimate` `r exp5_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp5_descriptives$model_h1_accuracy$convergence$p.value`) and competence ($\hat{b}_{\text{Competence}}$ = `r exp5_descriptives$model_h1_competence$convergence$estimate` `r exp5_descriptives$model_h1_competence$convergence$ci`, p = `r exp5_descriptives$model_h1_competence$convergence$p.value`).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \@ref(fig:experiments) B). In the independence condition, the effect of convergence on accuracy was more positive ($\hat{b}_{\text{interaction, Accuracy}}$ = `r exp5_descriptives$model_accuracy$interaction$estimate` `r exp5_descriptives$model_accuracy$interaction$ci`, p = `r exp5_descriptives$model_accuracy$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ = `r exp5_descriptives$model_accuracy$convergence$estimate` `r exp5_descriptives$model_accuracy$convergence$ci`, p = `r exp5_descriptives$model_accuracy$convergence$p.value`). Likewise, the effect of convergence on competence was more positive ($\hat{b}_{\text{interaction, Competence}}$ = `r exp5_descriptives$model_competence$interaction$estimate` `r exp5_descriptives$model_competence$interaction$ci`, p = `r exp5_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ = `r exp5_descriptives$model_competence$convergence$estimate` `r exp5_descriptives$model_competence$convergence$ci`, p = `r exp5_descriptives$model_competence$convergence$p.value`).

(ref:experiments) Interaction of convergence and informational dependency in **A** Experiment 1 (continuous) and **B** Experiment 2 (categorical).

```{r experiments, fig.cap="(ref:experiments)"}
# Experiment 3 figure
plot_exp3_accuracy <- ggplot(exp3, aes(x=convergence, y=accuracy, fill = independence, 
                     shape = independence,
                     group = independence,
                     color = independence)) +
  scale_x_discrete(limits = c("divergent", "convergent"), 
                    labels = c("Divergent", "Convergent")) +
  geom_half_violin(data = exp3 %>% filter(convergence=="divergent"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = exp3 %>% filter(convergence=="convergent"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Accuracy") +
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7)) +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("Conflict of interest", "Independent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("Conflict of interest", "Independent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none",
         fill = guide_legend(title = NULL)) +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

plot_exp3_competence <-ggplot(exp3, aes(x=convergence, y=competence, fill = independence, 
                     shape = independence,
                     group = independence,
                     color = independence)) +
  scale_x_discrete(limits = c("divergent", "convergent"), 
                    labels = c("Divergent", "Convergent")) +
  geom_half_violin(data = exp3 %>% filter(convergence=="divergent"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = exp3 %>% filter(convergence=="convergent"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Competence") +
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7)) +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("Conflict of interest", "Independent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("Conflict of interest", "Independent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none",
         fill = guide_legend(title = NULL)) +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

# Experiment 5 figure

# plot for accuracy
plot_exp5_accuracy <- ggplot(exp5,
       aes(x = convergence_categorical, y = accuracy, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_exp5_competence <-  ggplot(exp5,
       aes(x = convergence_categorical, y = competence, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# Combine plot
exp3_figure  <- ggarrange(
  plot_exp3_accuracy + 
    labs(y = " ")  + 
    rremove("xlab"), 
plot_exp3_competence + 
    labs(y = " ")  + 
    rremove("xlab"),
  nrow = 1,
common.legend = T) +
  theme(legend.position="top")

exp5_figure  <-  ggarrange(plot_exp5_accuracy + 
                             labs(y = " ") +
                             rremove("xlab") +
                             guides(color = "none", fill = "none") +
                             labs(caption = NULL),
                           plot_exp5_competence + 
                             labs(y = " ")  + 
                             rremove("xlab") +
                             guides(color = "none", fill = "none") +
                             labs(caption = NULL), 
                           common.legend = TRUE,
                           nrow = 1)

# common plot
patchwork <-  exp3_figure  / exp5_figure

# patchwork + 
#   plot_annotation(tag_levels = 'A') %>% 
#   annotate_figure(left= textGrob("Accuracy", gp = gpar(cex = 1, fontface = "bold"), 
#                                  rot = 90)) %>% 
#   annotate_figure(bottom = textGrob("Convergence", gp = gpar(cex = 1, fontface = "bold"), 
#                                  rot = 90))

patchwork +
  annotate("text", x = 0.02, y = 1, label = "Accuracy", angle = 90, size = rel(5), fontface = "bold" 
  ) +
  annotate("text", x = 0.52, y = 1, label = "Competence", angle = 90, size = rel(5), fontface = "bold" ) +
  #annotate("text", x = 0.5, y = 0.02, label = "Convergence", size = rel(5), fontface = "bold" ) +
  plot_layout(guides = 'collect') + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 12, hjust = -3))
```

In an exploratory, not pre-registered analysis, we tested whether the effect of convergence is bigger on accuracy than on competence. To do so, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy ($\hat{b}$ = `r exp5_exploratory_analysis$interaction$estimate` `r exp5_exploratory_analysis$interaction$ci`, `r exp5_exploratory_analysis$interaction$p.value`; units in standard deviations).

# General Discussion

In simulations, we have shown that the more a group of informants agrees with each other, the more they tend to be right and competent--assuming they are not systematically biased towards a false response. This is true for scenarios involving numerical estimates and for scenarios involving categorical answers, across a wide range of population competence distributions. In two experiments in which participants are deprived of any prior knowledge about the task of the informants' competence, we find that participants (UK) make these inferences. We also find that these inferences are weakened when the informants are systematically biased by a common conflict of interest.

People might draw this inference in a variety of contexts, but the most prominent one might be science. Science is, arguably, the institution in which individuals end up converging the most in their opinions. For instance, scientists within the relevant disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA. This represents an incredible degree of convergence. When people hear that scientists have measured the distance between the solar system and the center of the galaxy, if they assume that there is a broad agreement within the relevant experts, this should lead them to infer that this measure is accurate, and that the scientists who made it are competent. Experiments have already shown that increasing the degree of perceived consensus among scientists tends to increase acceptance of the consensual belief [@vanstekelenburgScientificConsensusCommunicationContested2022], but it hasn't been shown that the degree of consensus also affects the perceived competence of scientists.

In the case of science, the relationship between convergence and accuracy is broadly justified. However, at some points of history, there has been broad agreement on misbeliefs, such as when Christian theologians had calculated that the Earth was approximately six thousand years old. To the extent that people were aware of this broad agreement, and believed the theologians to have reached it independently of each other, this might have not only fostered acceptance of this estimate of the age of the Earth, but also a perception of the theologians as competent.

The current study has a number of limitations. If the very abstract materials allow us to remove most of the priors the participants might have, they might also reduce the ecological validity of the experiments. Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019]. However, we do not know whether they would generalize to other cultures.

\FloatBarrier

# References

::: {#refs}
:::

### 2024-03-26 README.qmd
major changes before adding analytical part

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  fig.retina = 3
)

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("paper.Rmd"))
title <- yaml_metadata$title
abstract <- yaml_metadata$abstract
```

# `r title`

---

## Abstract

`r abstract`

---

## How to download and replicate

To reproduce the findings and re-run the analysis, do the following:

1. Download and install this repository. You can use GitHub to clone or fork the repository (see the green "Clone or download" button at the top of the GitHub page).
2. Open `consensus_trust.Rproj` to open an [RStudio Project](https://r4ds.had.co.nz/workflow-projects.html).
3. Within the RStudio Project, open the `paper.Rmd` file and run it.

Please find the [session info below](#session-info).

*Estimated installation time* : If you have R and Rstudio installed, then adding additional packages required in our R scripts is quick (a couple of minutes)

*Estimated run time*: Rendering the `anonymized.Rmd` document takes about 2mins on our machine (see [session info](#session-info))

## Files glossary {#glossary}

- `appendix*` files contain the appendices that are automatically added `paper.Rmd` when rendering the file.
- `Cogsci*` files are documents used to generate a shorter version of our paper submitted for the CogSci 2024 conference. 
- `Experiment*` folders contain all files related to the respective experiment. The `/data` folders contain the raw data (`qualtrics.csv`) and cleaned data (`cleaned.csv`). The cleaning can be reproduced by running the `cleaning.Rmd`. `Experiment_1` and `Experiment_4` also contain the code for the simulations (`model.Rmd`) and several data files with simulated data. 
- `functions/` folder contains functions that we rely upon in our analysis scripts. To keep the `paper.Rmd` document neat and readable, we wrote all lengthy code bits as external functions.
- figures includes images used in the `paper.Rmd` file
- `howwise_0131.pdf` is the final submission to the CogSci 2024 conference, based on the raw files but edited in [overleaf](https://www.overleaf.com/). 
- `output` folder lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.
- `paper.pdf` is the rendered version of `paper.Rmd`
- `references.bib` contains the references
- `stimuli` folder contains all stimuli

## Session info {#session-info}

```{r, echo=FALSE}
sessionInfo()
```

## Licenses

**Text and figures:** All prose and images are licensed under Creative Commons ([CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)).

**Code:** All code is licensed under the [MIT License](LICENSE.md).

### 2024-03-26 appendix_exp1.Rmd
major changes before adding analytical part

@@
 
+```{r simuli, include=FALSE, out.width= "40%", fig.align="center", fig.show="hold"}
+# knitr::include_graphics("figures/example_stimulus_exp1_convergent.png")
+# knitr::include_graphics("figures/example_stimulus_exp1_divergent.png")
+```
+
 ## Attention check
@@
 Figure \@ref(fig:exp1-plot) visualizes the results and table \@ref(tab:exp1-table) contains descriptive results.
+Figure \@ref(fig:exp1-outcomes) visualizes score differences between the two outcomes, accuracy and competence.
 
@@
 ```
+
+## Research questions
+
+We had three additional research questions regarding the number of informants:
+
+***RQ1: Do H1 and H2 hold for both a small [3] and a large [10] number of estimates?***
+
+***RQ2: When making a guess based on the opinions of (independent) informants, will participants be more confident about their guess when there is a larger number of estimates compared to when this number is smaller?***
+
+***RQ3: Is there an interaction effect between the number of estimates and convergence on perceived competence of informants?***
+
+Regarding RQ1, we ran the same mixed models as for the two hypotheses, but without number as a covariate.
+Instead, we ran separate models for the three and the ten informants conditions.
+We find that for both sub-samples, there is a positive effect of convergence on both accuracy ($\hat{b}_{\text{three informants}}$ = `r exp1_research_questions$model_small_accuracy$convergenceconvergent$estimate` `r exp1_research_questions$model_small_accuracy$convergenceconvergent$ci`, p = `r exp1_research_questions$model_small_accuracy$convergenceconvergent$p.value`; $\hat{b}_{\text{ten informants}}$ = `r exp1_research_questions$model_large_accuracy$convergenceconvergent$estimate` `r exp1_research_questions$model_large_accuracy$convergenceconvergent$ci`, p = `r exp1_research_questions$model_large_accuracy$convergenceconvergent$p.value`) and competence ($\hat{b}_{\text{three informants}}$ = `r exp1_research_questions$model_small_competence$convergenceconvergent$estimate` `r exp1_research_questions$model_small_competence$convergenceconvergent$ci`, p = `r exp1_research_questions$model_small_competence$convergenceconvergent$p.value`; $\hat{b}_{\text{ten informants}}$ = `r exp1_research_questions$model_large_competence$convergenceconvergent$estimate` `r exp1_research_questions$model_large_competence$convergenceconvergent$ci`, p = `r exp1_research_questions$model_large_competence$convergenceconvergent$p.value`).
+
+To test the other two research questions, we ran the same mixed-models as for the hypotheses, but this time including an interaction between number of informants and convergence.
+We use deviation-coded versions of our convergent variable (divergent = -0.5, convergent = 0.5), allowing us to have a coefficient measuring the main effect of the number of informants in our model.
+Regarding RQ2, pooling across convergent and divergent conditions, we find a main effect of number, such that participants had more confidence in their estimate when they relied on ten informants compared to three informants ($\hat{b}_{\text{Accuracy}}$ = `r exp1_research_questions$model_interaction_accuracy$number_effect_code$estimate` `r exp1_research_questions$model_interaction_accuracy$number_effect_code$ci`, p = `r exp1_research_questions$model_interaction_accuracy$number_effect_code$p.value`).
+Regarding RQ3, we find an interaction between number of informants and convergence on competence: the positive effect of convergence was stronger in scenarios involving ten informants compared to three informants ($\hat{b}_{\text{Competence}}$ = `r exp1_research_questions$model_interaction_competence$interaction$estimate` `r exp1_research_questions$model_interaction_competence$interaction$ci`, p = `r exp1_research_questions$model_interaction_competence$interaction$p.value`).
+We do not find a statistically significant interaction on accuracy ($\hat{b}_{\text{Accuracy}}$ = `r exp1_research_questions$model_interaction_accuracy$interaction$estimate` `r exp1_research_questions$model_interaction_accuracy$interaction$ci`, p = `r exp1_research_questions$model_interaction_accuracy$interaction$p.value`).
+
+(ref:exp1-outcomes) Differences between accuracy and competence ratings, by level of convergence.
+
+```{r exp1-outcomes, fig.cap="(ref:exp1-outcomes)"}
+# Differences in outcome between accuracy and competence
+exp1_long <- exp1 %>%
+  pivot_longer(cols = c(confidence, competence), 
+               names_to = "outcome", 
+               values_to = "score")
+
+interaction_plot <- ggplot(exp1_long, aes(x = outcome, y = score, 
+                                         fill = convergence, 
+                                         shape = convergence, 
+                                         group = convergence, 
+                                         color = convergence)) +
+  scale_x_discrete(limits = c("confidence", "competence"), 
+                   labels = c("Accuracy", "Competence")) +
+  geom_half_violin(data = exp1_long %>% filter(outcome == "confidence"), 
+                   position = position_nudge(x = -.2), adjust = 2, alpha = .4,
+                   side = "l") +
+  geom_half_violin(data = exp1_long %>% filter(outcome == "competence"),
+                   position = position_nudge(x = .2), adjust = 2, alpha = .4,
+                   side = "r") +
+  xlab("Outcome") +
+  ylab("Score") +
+  scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
+  stat_summary(fun = "mean", geom = "point", size = 3) +
+  stat_summary(fun = "mean", geom = "line") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  scale_fill_manual(name = NULL,
+                    labels = c("divergent", "convergent"),
+                    values = c("#E69F00", "#56B4E9")) +
+  scale_color_manual(name = NULL,
+                     labels = c("divergent", "convergent"),
+                     values = c("#E69F00", "#56B4E9")) +
+  guides(shape = "none",
+         fill = guide_legend(title = NULL)) +
+  plot_theme +
+  # change font sizes
+  theme(axis.text = element_text(size = 10)) +
+  theme(axis.title = element_text(size = 15)) +
+  theme(legend.text = element_text(size = 10))
+
+print(interaction_plot)
+```
+
+```{r, include = FALSE}
+# manual version for points / confidence intervals for the above plot
+
+# exp1_long <- exp1 %>%
+#   pivot_longer(cols = c(confidence, competence), 
+#                names_to = "outcome", 
+#                values_to = "score")
+# 
+# # Calculate mean and standard error
+# mean_se <- exp1_long %>%
+#   group_by(outcome, convergence) %>%
+#   summarize(mean_score = mean(score), 
+#             se_score = sd(score) / sqrt(n())) %>% 
+#   mutate(ci_low = mean_score - se_score, 
+#          ci_high = mean_score + se_score)
+# 
+# interaction_plot <- ggplot(exp1_long, aes(x = outcome, y = score, 
+#                                          fill = convergence, 
+#                                          shape = convergence, 
+#                                          group = convergence, 
+#                                          color = convergence)) +
+#   scale_x_discrete(limits = c("confidence", "competence"), 
+#                    labels = c("Accuracy", "Competence")) +
+#   geom_half_violin(data = exp1_long %>% filter(outcome == "confidence"), 
+#                    position = position_nudge(x = -.2), adjust = 2, alpha = .4,
+#                    side = "l") +
+#   geom_half_violin(data = exp1_long %>% filter(outcome == "competence"),
+#                    position = position_nudge(x = .2), adjust = 2, alpha = .4,
+#                    side = "r") +
+#   xlab("Outcome") +
+#   ylab("Score") +
+#   scale_y_continuous(breaks = c(1, 2, 3, 4, 5, 6, 7)) +
+#   geom_point(data = mean_se, aes(x = outcome, y = mean_score), size = 3) +
+#   geom_line(data = mean_se, aes(x = outcome, y = mean_score)) +
+#   geom_errorbar(data = mean_se,
+#                 aes(x = outcome,
+#                     y = mean_score,
+#                     ymin = mean_score - se_score,
+#                     ymax = mean_score + se_score),
+#                 width = 0.2) +
+#   scale_fill_manual(name = NULL,
+#                     labels = c("divergent", "convergent"),
+#                     values = c("#E69F00", "#56B4E9")) +
+#   scale_color_manual(name = NULL,
+#                      labels = c("divergent", "convergent"),
+#                      values = c("#E69F00", "#56B4E9")) +
+#   guides(shape = "none",
+#          fill = guide_legend(title = NULL)) +
+#   plot_theme +
+#   # change font sizes
+#   theme(axis.text = element_text(size = 10)) +
+#   theme(axis.title = element_text(size = 15)) +
+#   theme(legend.text = element_text(size = 10))
+# 
+# print(interaction_plot)
+```

### 2024-03-26 appendix_exp2.Rmd
major changes before adding analytical part

@@
   )
-) #%>% 
-  # apa_table(align = c("m{3cm}", "m{12cm}")) 
+) %>% 
+  apa_table(align = c("m{3cm}", "m{12cm}")) 
 ```

### 2024-03-26 appendix_exp3.Rmd
major changes before adding analytical part

@@
   )
-) #%>% 
-  # apa_table(align = c("m{3cm}", "m{12cm}"))  
+) %>% 
+  apa_table(align = c("m{3cm}", "m{12cm}")) 
 ```
@@
 We generated four such series (all possible combinations) and randomly assigned participants to one of them.
-Additionally, for each participant, we randomize the order of appearance of the sets of predictions within the respective series.
+Additionally, for each participant, we randomized the order of appearance of the sets of predictions within the respective series.
 

### 2024-03-26 appendix_exp4.Rmd
major changes before adding analytical part

@@
 # Create a matrix of image file paths as Markdown-formatted strings
-image_paths <- data.frame(condition = c("opposing majority (0)", 
+image_paths <- data.frame(condition = c("minority (0)", 
                                         "dissensus (1)", "majority (2)", 
@@
                   align = "c",
-                  caption = "All stimuli by levels of convergence")
+                  caption = "All stimuli by levels of convergence") %>%
+  column_spec(1, width = "2em") %>%
+  column_spec(2, width = "15em") %>%
+  column_spec(3, width = "15em")
 
@@
 
-Table \@ref(tab:exp4-table) contains descriptive results.
+Table \@ref(tab:exp4-table) contains descriptive results and Figure \@ref(fig:exp4-plot) visualizes the results.
 
@@
 ```
+
+(ref:exp4-plot) Distributions of accuracy and competence by level of convergence.
+
+```{r exp4-plot, fig.cap="(ref:exp4-plot)"}
+# plot for accuracy
+plot_accuracy <- ggplot(exp4,
+       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .8,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Accuracy") +
+  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
+  guides(fill = FALSE) +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +  
+  rremove("xlab")   # Remove x-axis label
+
+# plot for competence
+plot_competence <- ggplot(exp4,
+       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .8,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  scale_y_continuous(limits = c(1,7), breaks = seq(1,7)) +
+  labs(x = "Convergence", y = "Competence") +
+  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
+  guides(fill = FALSE) +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1)) +  
+  rremove("xlab")   # Remove x-axis label
+
+ggarrange(plot_accuracy, plot_competence) %>% 
+  annotate_figure(bottom = textGrob("Convergence",gp = gpar(cex = 1, fontface = "bold")))
+```
+
+## Comparison with simulation
+
+We compared the accuracy and competence ratings of the participants to actual accuracy and competence values from simulated data.
+The data was generated using the model described in the main paper, and assuming a uniform competence distribution.
+Compared to this data generating model, participants underestimate the effect of convergence for both accuracy and competence, but more for accuracy.
+Table \@ref(tab:model-vs-participant-table) compares the results of regressions on participant data and regressions on simulated data.
+Fig.
+\@ref(fig:exp4-model) visualizes the results.
+
+```{r}
+# read simulated data from .csv files
+sim_experiment_4 <- read_csv("Experiment_4/data/sim_experiment_4.csv")
+```
+
+```{r}
+# Run model on simulated data
+regression_data <- sim_experiment_4 %>%
+  mutate(
+    # make numeric version of convergence
+    convergence = case_when( 
+      constellation == "minority" ~ 0, 
+      constellation == "dissensus" ~ 1,
+      constellation == "majority" ~ 2,
+      constellation == "consensus" ~ 3), 
+    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
+    average_accuracy = average_accuracy*100,
+    # scale competence to reach from 1 to 7 (instead of 0 to 1)
+    average_relative_competence = 6*average_relative_competence + 1 
+  )
+
+accuracy_model <- lm(average_accuracy ~ convergence, 
+                     data = regression_data)
+competence_model <- lm(average_relative_competence ~ convergence, 
+                       data = regression_data)
+```
+
+```{r}
+# read participant data
+d <- exp4
+```
+
+```{r model-vs-participant-table}
+# Calculate models for participants
+
+# model for accuracy
+# random intercept and slope by participants
+accuracy_participants <- lmer(accuracy ~ convergence + (1 + convergence | id), 
+                                 data = d)
+# model for competence
+# random intercept and slope by participants
+competence_participants <- lmer(competence ~ convergence + 
+                           (1 + convergence | id), data = d)
+
+# main result table
+modelsummary::modelsummary(list("Accuracy" = accuracy_participants, 
+                                "Competence" = competence_participants,
+                                "Accuracy" = accuracy_model, 
+                                "Competence" = competence_model
+                                ),
+                           title = 'Participants vs. Model (Exp. 4)', 
+                           stars = TRUE) %>%
+  kableExtra::add_header_above(c(" " = 1, "Participants" = 2, "Model" = 2))
+```
+
+```{r, echo=FALSE}
+# Participant data
+
+# plot for accuracy
+plot_accuracy <- ggplot(d,
+       aes(x = convergence_categorical, y = accuracy, fill = convergence_categorical)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .8,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Accuracy") +
+  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
+  guides(fill = FALSE) +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1))
+
+# plot for competence
+plot_competence <- ggplot(d,
+       aes(x = convergence_categorical, y = competence, fill = convergence_categorical)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .8,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 1, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Competence") +
+  scale_fill_viridis_d(option = "plasma", begin = 0.1) +
+  guides(fill = FALSE) +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1))
+```
+
+```{r}
+# Model data
+plot_accuracy_model <- plot_results(sim_experiment_4, outcome = "accuracy")
+plot_competence_model <- plot_results(sim_experiment_4, outcome = "competence")
+```
+
+(ref:exp4-model) Comparison of accuracy (top) and competence (bottom) ratings in participant data (left) to observed accuracy in simulated data (right), by level of convergence.
+
+```{r exp4-model, fig.cap="(ref:exp4-model)"}
+# Combine participant and model
+# accuracy
+plot_accuracy <- plot_accuracy + ggtitle("Participants") + labs(caption = NULL)
+plot_accuracy_model <- plot_accuracy_model + ggtitle("Model") + labs(caption = NULL)
+
+# competence
+plot_competence <- plot_competence 
+plot_competence_model <- plot_competence_model
+
+(plot_accuracy + plot_accuracy_model) / (plot_competence + plot_competence_model)
+```

### 2024-03-26 appendix_exp5.Rmd
major changes before adding analytical part

@@
 We manipulated convergence within participants in the same way we did in experiment 4.
-In addition, between participants, we manipulated informational dependence, akin to experiment 3 (Table \@ref(exp5-condition)).
+In addition, between participants, we manipulated informational dependence, akin to experiment 3 (Table \@ref(tab:exp5-condition)).
 In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is.
@@
 Do you agree with your friend? (Yes/No)
-
-## Stimuli
-
-Our design required that each participant sees four different sets of predictions--two convergent and two divergent ones.
-We generated new stimuli: As in experiment 1, estimates would appear on a scale from 1000 to 2000.
-By contrast with experiment 1, we generated the sets of estimates with random draws from uniform distributions (instead of normal distributions).
-We varied the range of these distributions according to convergence (60 for convergence, 600 for divergence).
-Switching from normal distributions to uniform distributions made 'unlucky' draws in which conditions are visually not distinguishable less likely.
-
-Similar to the previous experiments, we also vary the value on the prediction scale (from 1000 to 2000) across which the range is centered.
-Considering our within-participant design, this, again, made it seem more likely that participants understand each set of predictions as being the result of a different stock, with a different true value.
-In order to assure all random draws from the distributions would appear on the response scale, we constrained the center of the uniform distributions to lie between 1300 and 1700.
-We define four center values -- one per set of predictions -- that divide this interval in (rounded) quartiles (1300, 1430, 1570, 1700).
-Given a center and a range, we then draw the predictions from uniform distributions.
-For example, in a draw for a divergent set of estimates with a center of 1700, each value within a range of 1400 and 2000 is equally likely to get selected.
-To avoid that single draws overlap too much within the same set, we defined a minimum space of 5 between the three predictions of a set.
-
-To minimize confounding of convergence with a certain placement of the center value, we paired center values with conditions such that each condition appears once in each half of the scale and each condition gets one of the extreme values.
-For example, in one set the convergence condition would get assigned center values of 1300 and 1570, the divergent condition center values of 1430 and 1700).
-We generated four such series (all possible combinations) and randomly assigned participants to one of them.
-Additionally, for each participant, we randomize the order of appearance of the sets of predictions within the respective series.
-
-For each set of predictions, we calculated the empirical mean based on the randomly drawn estimates.
-In the conflict of interest condition, this mean was inserted as the value that participants were be told experts would gain from.
-Consequently, the convergent predictions converge around what is said to be the incentivized value for the experts to choose.
 

### 2024-03-26 appendix_exp6.Rmd
major changes before adding analytical part

@@
 
-We manipulated convergence within participants in the same way we did in experiment 4.
-In addition, between participants, we manipulated the number of choice options.
-Participants were randomly assigned to either to see stimuli with '3' options, or with '10' options.
-Participants assigned to the '10' options condition, participants were assigned to one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the '3' options condition, and another with increased range (see Appendix XX).
-We added the increased range condition because we anticipated the possibility that participants might not consider all options as relevant when they only see scenarios in which all answers cluster.
-We found no differences between the two sub-conditions and collapsed them into a single '10' options condition.
+The main focus of experiment 6 was on comparing 3-options conditions to 10-options conditions (see Table \@ref(tab:stimuli-options)).
+Additionally, we varied between two versions of the 10-options conditions: one in which the range of the answers corresponded to the range of the three options condition (see Table \@ref(tab:stimuli-10)), and another with increased range (see Table \@ref(tab:stimuli-10-alt)).
 
@@
 image_paths <- data.frame(
-  imgage_a = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_3_a.png){ width=60% }"),
-  imgage_b = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_10_a.png){ width=60% }")
+  imgage_a = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_3_a.png){ width=40% }"),
+  imgage_b = c("![](stimuli/Experiment_6/stimuli_Exp6/consensus_10_a.png){ width=40% }")
   )
@@
   imgage_a = c(
-    "![](Experiment_6/figures/stimuli/minority_10_a.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/divergence_10_a.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/majority_10_a.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/consensus_10_a.png){ width=60% }"
+    "![](Experiment_6/figures/stimuli/minority_10_a.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/divergence_10_a.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/majority_10_a.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/consensus_10_a.png){ width=40% }"
   ),
   imgage_b = c(
-    "![](Experiment_6/figures/stimuli/minority_10_b.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/divergence_10_b.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/majority_10_b.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/consensus_10_b.png){ width=60% }"
+    "![](Experiment_6/figures/stimuli/minority_10_b.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/divergence_10_b.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/majority_10_b.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/consensus_10_b.png){ width=40% }"
   )
@@
                   align = "c",
-                  caption = "Stimuli for 10 options condition by levels of convergence")
+                  caption = "Stimuli for 10 options condition by levels of convergence", 
+                  full_width = T) %>%
+  column_spec(1, width = "2em") %>%
+  column_spec(2, width = "20em") %>%
+  column_spec(3, width = "20em") %>%
+  # make smaller to fit
+  kable_styling(latex_options = "scale_down")
 
@@
   imgage_a = c(
-    "![](Experiment_6/figures/stimuli/minority_10_a_alt.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/divergence_10_a_alt.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/majority_10_a_alt.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/consensus_10_a_alt.png){ width=60% }"
+    "![](Experiment_6/figures/stimuli/minority_10_a_alt.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/divergence_10_a_alt.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/majority_10_a_alt.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/consensus_10_a_alt.png){ width=40% }"
   ),
   imgage_b = c(
-    "![](Experiment_6/figures/stimuli/minority_10_b_alt.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/divergence_10_b_alt.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/majority_10_b_alt.png){ width=60% }",
-    "![](Experiment_6/figures/stimuli/consensus_10_b_alt.png){ width=60% }"
+    "![](Experiment_6/figures/stimuli/minority_10_b_alt.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/divergence_10_b_alt.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/majority_10_b_alt.png){ width=40% }",
+    "![](Experiment_6/figures/stimuli/consensus_10_b_alt.png){ width=40% }"
   )
@@
 ```
+
+## Comparison with simulation
+
+```{r}
+# read simulated data from .csv files
+data_3_options <- read_csv("Experiment_4/data/sim_vary_options_3_sample.csv")
+```
+
+Since we didn't find a difference between three and ten choice option scenarios in the participant data, we wanted to have a normative point of reference of what should have been the expected difference, given our model.
+We therefor compared participant ratings to observed values from simulated data.
+The data was generated using the model described in the main paper, and assuming a uniform competence distribution.
+Compared to this data generating model, participants considerably underestimated how the number of choice options alters the effect of convergence for both accuracy and competence.
+Table \@ref(tab:model-vs-participant-table) compares the results of regressions on participant data and regressions on simulated data.
+Fig.
+\@ref(fig:exp6-model) visualizes the results.
+
+```{r}
+# possible analysis for varying choice options (3 vs. 10)
+regression_data <- data_3_options %>%
+  filter(options %in% c(3, 10)) %>%
+  # group_by(iteration, constellation, options) %>%
+  # summarise(across(c(average_accuracy, average_relative_competence), mean)) %>%
+  mutate(
+    convergence = case_when(
+      constellation == "minority(ies)" ~ 0,
+      constellation == "dissensus" ~ 1,
+      constellation == "majority" ~ 2,
+      constellation == "consensus" ~ 3),
+    # make a categorical version
+    convergence_categorical = recode_factor(convergence, 
+                                            `0` = "minority", 
+                                            `1` = "divergence", 
+                                            `2` = "majority", 
+                                            `3` = "consensus",
+                                            .default = NA_character_),
+    # make a version of number for plot
+    number_options = ifelse(options == 3, "three", "ten"),
+    # we'll use an effect coded version for the analysis
+    number_options_effect_code = ifelse(options == 3, -0.5, 0.5),
+    # scale accuracy to reach from 0 to 100 (instead from 0 to 1)
+    average_accuracy = average_accuracy*100,
+    # scale competence to reach from 1 to 7 (instead of 0 to 1)
+    average_relative_competence = 6*average_relative_competence + 1
+  )
+
+accuracy <- lm(average_accuracy ~ convergence + number_options_effect_code + convergence*number_options_effect_code,
+               data = regression_data)
+competence <- lm(average_relative_competence ~ convergence + number_options_effect_code + convergence*number_options_effect_code, 
+                 data = regression_data)
+```
+
+```{r}
+# read data
+exp6 <- read_csv("Experiment_6/data/cleaned.csv") %>% 
+  # import experimental variable as factor and set levels
+  mutate(
+    # set levels for `convergence`
+    convergence_categorical = recode_factor(convergence, 
+                                            `0` = "minority", 
+                                            `1` = "divergence", 
+                                            `2` = "majority", 
+                                            `3` = "consensus",
+                                            .default = NA_character_))  %>% 
+  # For this model, we effect code `number_options`, so that we get a coefficient 
+  # for the main effect of convergence as well as for the interaction
+  mutate(number_options_effect_code = recode(number_options, "three" = -0.5, "ten" = +0.5))
+
+# run mixed model (random slope and intercept) with participants as random factor
+exp6_model_accuracy = lmer(accuracy ~ convergence + number_options_effect_code + 
+                            number_options_effect_code*convergence + (1 + convergence | id), 
+                       data = exp6)
+exp6_model_competence = lmer(competence ~ convergence + number_options_effect_code + 
+                            number_options_effect_code*convergence + (1 + convergence | id), 
+                       data = exp6)
+```
+
+```{r}
+modelsummary::modelsummary(list("Accuracy" = exp6_model_accuracy, 
+                                "Competence" = exp6_model_competence,
+                                "Accuracy" = accuracy, 
+                                "Competence" = competence
+                                ),
+                           title = 'Participants vs. Model (Exp. 6)', 
+                           stars = TRUE) %>%
+  kableExtra::add_header_above(c(" " = 1, "Participants" = 2, "Model" = 2))
+```
+
+(ref:exp6-model) Comparison of accuracy (top) and competence (bottom) ratings in participant data (left) to observations in simulated data (right), by level of convergence.
+
+```{r exp6-model, fig.cap="(ref:exp6-model)"}
+# colors were identified with the wesanderson package (library(wesanderson)) as:
+# wes_palette("Zissou1", 100, type = "continuous")[A number]
+
+# plot for accuracy
+plot_accuracy_participants <- ggplot(exp6,
+       aes(x = convergence_categorical, y = accuracy, fill = number_options)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .4,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Accuracy", fill = NULL, title = "Participants") +
+  scale_fill_manual(values = c("#EE3800", "#3D9BB2"), 
+                    name = "Number of choice options: ") +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1))
+
+plot_accuracy_model <- ggplot(regression_data,
+       aes(x = convergence_categorical, y = average_accuracy, fill = number_options)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .4,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Accuracy", fill = NULL, title = "Model") +
+  scale_fill_manual(values = c("#EE3800", "#3D9BB2"), 
+                    name = "Number of choice options: ") +
+  guides(fill = "none") +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1))
+
+# plot for competence
+plot_competence_participants <- ggplot(exp6,
+       aes(x = convergence_categorical, y = competence, fill = number_options)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .4,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Competence", fill = NULL) +
+  scale_fill_manual(values = c("#EE3800", "#3D9BB2")) +
+  guides(fill = "none") +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1))
+
+plot_competence_model <- ggplot(regression_data,
+       aes(x = convergence_categorical, y = average_competence, fill = number_options)) +
+  geom_half_violin(position = position_nudge(x = -.2),
+                   adjust=2, alpha = .4,
+                   side = "l") +
+  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
+  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
+  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
+  # Add nice labels
+  labs(x = "Convergence", y = "Competence", fill = NULL) +
+  scale_fill_manual(values = c("#EE3800", "#3D9BB2")) +
+  guides(fill = "none") +
+  plot_theme + 
+  theme(axis.text.x = element_text(angle = 20, hjust = 1))
+
+participants <- ggpubr::ggarrange(plot_accuracy_participants,plot_accuracy_model, common.legend = TRUE) 
+model <- ggpubr::ggarrange(plot_competence_participants, plot_competence_model)
+
+participants / model + plot_layout(guides = "collect")
+```

### 2024-03-26 appendix_simulations.Rmd
major changes before adding analytical part

@@
-# Simulations {simulations}
+# Simulations {#simulations}
 
@@
 We suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution.
-This competence distribution can take vary different shapes, depending on the alpha and beta parameters that describe the distribution (see fig.
-\@ref(fig:competence-distributions)).
+This competence distribution can take vary different shapes, depending on the alpha and beta parameters that describe the distribution (see Fig.
+\@ref(fig:simulations)).
 
@@
 We measure the convergence of a group by calculating the standard deviation of the estimates.
-We run this simulation on a sample size of roughly 100000 (varying slightly as a function of sample size).
+We run this simulation on a sample size of roughly 99900 (varying slightly as a function of sample size).
 We repeat this simulation process for various sample sizes and competence distributions.
 The results are displayed in Fig.
-\@ref(fig:simulation-accuracy-numeric-1) to \@ref(fig:simulation-accuracy-numeric-6), for accuracy, and Fig.
-\@ref(fig:simulation-competence-numeric-1) to \@ref(fig:simulation-competence-numeric-6), for competence.
+\@ref(fig:simulation-accuracy-numeric) for accuracy, and Fig.
+\@ref(fig:simulation-competence-numeric) for competence.
 Across all underlying competence distributions, we find a positive correlation between convergence and accuracy, which tends towards 1 as sample size increases (see Fig.
 \@ref(fig:simulation-numeric-correlations)).
-As for accuracy, we find a positive correlation between convergence and competence across all underlying competence distributions.
-However, this correlations are weaker than for accuracy, and do not increase with sample size (see Fig.
+As for competence, we find a positive correlation between convergence and competence across all underlying competence distributions.
+However, these correlations are weaker than for accuracy, and do not increase with sample size (see Fig.
 \@ref(fig:simulation-numeric-correlations)).
 
+(ref:simulation-numeric-correlations) Correlation between the convergence of the answers of a group of agents and (left) the average accuracy of their answers; (right) the average competence of the agents, as a function of how many agents are in the group, and of the competence distribution of the agents (see Fig.
+\@ref(fig:simulations)).
+
+```{r simulation-numeric-correlations, fig.cap="(ref:simulation-numeric-correlations)", out.width="75%"}
+# calculate correlations
+simulation_numeric <- simulation_numeric %>% 
+  group_by(competence, sample) %>% 
+  mutate(correlation_accuracy = cor(accuracy_mean, convergence), 
+         correlation_competence = cor(competence_mean, convergence)) %>% 
+  ungroup()
+
+# accuracy plot
+accuracy_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_accuracy, color = competence)) +
+  geom_point() + 
+  geom_line(alpha = 0.5) + 
+  labs(y = "Correlation \nAccuracy/Convergence", 
+       x = "Group size", 
+       color = "Competence distribution") +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_color_brewer(palette = "Set1") +
+  plot_theme + 
+  theme(
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)) +
+  rremove("xlab")
+
+# competence plot
+competence_plot <- ggplot(simulation_numeric, aes(x = sample, y = correlation_competence, color = competence)) +
+  geom_point() + 
+  geom_line(alpha = 0.5) + 
+  labs(y = "Correlation \nCompetence/Convergence", 
+       x = "Group size", 
+       color = "Competence distribution") +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_color_brewer(palette = "Set1") +
+  plot_theme + 
+  theme(
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)) +
+  rremove("xlab")
+
+ggarrange(accuracy_plot, competence_plot, common.legend = TRUE) %>% 
+    annotate_figure(bottom = textGrob("Group size", gp = gpar(cex = 1, fontface = "bold")))
+```
+
 ## Categorical choice context
+
+(ref:simulation-accuracy-sample-categorical) Accuracy as a function of convergence, for different population competence distributions and sample sizes. The number of choice options is three.
+
+```{r simulation-accuracy-sample-categorical, fig.cap="(ref:simulation-accuracy-sample-categorical)", out.width="100%"}
+plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
+                                       outcome = "Accuracy", variable = sample) 
+```
+
+(ref:simulation-competence-sample-categorical) Competence as a function of convergence (i.e.
+vote share for an option), for different population competence distributions and sample sizes.
+The number of choice options is three.
+
+```{r simulation-competence-sample-categorical, fig.cap="(ref:simulation-competence-sample-categorical)", out.width="100%"}
+plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
+                                       outcome = "Competence", variable = sample)
+```
 
@@
 Competence values range from being at chance for selecting the correct option (p = 1/m) to certainly selecting the correct option (p = 1).
-As before in the numeric case, suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution (see fig.
-\@ref(fig:competence-distributions)).
+As before in the numeric case, suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution (see Fig.
+\@ref(fig:simulations)).
 Based on their competence level, we draw an estimate for each individual.
@@
 In this case we obtain an average accuracy and an average competence value for a share of 2/3 (option C) and for a share of 1/3 (option B).
-We simulate this on a population of 999000 individuals.
+We simulate this on a population of 99900 individuals.
 We repeat this procedure varying the underlying population competence distributions, and additionally varying either (a) the sample size of informants, or (b) the number of choice options.
@@
 \@ref(fig:simulation-competence-options-categorical) display the same relationship, but varying the number of choice options instead.
-The figures dispaly that, across all sample sizes and competence levels, the larger the share of votes for an option, the more accurate the option is on average.
+The figures display that, across all sample sizes and competence levels, the larger the share of votes for an option, the more accurate the option is on average.
 That relationship appears to follow some sigmoid curve which switches from an average accuracy of 0 to an average accuracy of 1 before a share of 0.5 is attained, and which is steeper for larger sample sizes.
@@
 
-```{r simulation-accuracy-numeric-1, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
-# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)
-
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[1]]
-```
-
-```{r simulation-accuracy-numeric-2, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[2]]
-```
-
-```{r simulation-accuracy-numeric-3, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[3]]
-```
-
-```{r simulation-accuracy-numeric-4, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[4]]
-```
-
-```{r simulation-accuracy-numeric-5, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[5]]
-```
-
-```{r simulation-accuracy-numeric-6, fig.cap="(ref:simulation-accuracy-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[6]]
+```{r, simulation-accuracy-numeric, fig.cap="(ref:simulation-accuracy-numeric)", fig.height= 10, fig.width=10}
+plot_continuous_combined(data = simulation_numeric, outcome = "accuracy")
 ```
@@
 
-```{r simulation-competence-numeric-1, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
+```{r, simulation-competence-numeric, fig.cap="(ref:simulation-competence-numeric)", fig.height= 10, fig.width=10}
+plot_continuous_combined(data = simulation_numeric, outcome = "competence")
+```
+
+```{r simulation-accuracy-single, fig.cap="(ref:simulation-accuracy-single)", out.width="75%"}
 # plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")
 # ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[1]]
+
+# Display single plots
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[1]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[2]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[3]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[4]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[5]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "accuracy")[[6]]
 ```
 
-```{r simulation-competence-numeric-2, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[2]]
+```{r simulation-competence-single, fig.cap="(ref:simulation-competence-single)", out.width="75%"}
+# plot_list <- plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")
+# ggarrange(plotlist = plot_list, common.legend = TRUE, ncol = 1)
+
+# Display single plots
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[1]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[2]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[3]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[4]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[5]]
+# plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[6]]
 ```
-
-```{r simulation-competence-numeric-3, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[3]]
-```
-
-```{r simulation-competence-numeric-4, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[4]]
-```
-
-```{r simulation-competence-numeric-5, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[5]]
-```
-
-```{r simulation-competence-numeric-6, fig.cap="(ref:simulation-competence-numeric)", out.width="75%"}
-plot_competence_vary_numeric(data = simulation_numeric, outcome = "competence")[[6]]
-```

### 2024-03-26 paper.Rmd
major changes before adding analytical part

---
title             : "How wise is the crowd? \n Can we infer people are accurate and competent merely because they agree with each other?"
shorttitle        : "How wise is the crowd?"

header-includes:  | # to prevent floats from moving past certain points (for the appendix)
  \usepackage{placeins}

affiliation:
  - id            : " "
    institution   : " "

abstract: |
  Are people who agree on something more likely to be right and competent? Evidence suggests that people tend to make this inference. However, standard wisdom of crowds approaches only provide limited normative grounds . Using simulations and analytical arguments, we argue that when individuals make independent and unbiased estimates, under a wide range of parameters, individuals whose answers converge with each other tend to have more accurate answers and to be more competent. In 6 experiments (UK participants, total N = 1197), we show that participants infer that informants who agree have more accurate answers and are more competent, even when they have no priors, and that these inferences are weakened when the informants were systematically biased. In conclusion, we speculate that  inferences from convergence to accuracy and competence might help explain why people deem scientists competent, even if they have little understanding of science.

keywords          :

floatsintext      : yes
linenumbers       : no 
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "doc" # "doc" for nice look, "man" for manuscripty
output            : papaja::apa6_pdf # "_doc" for word; however, note that some of the features of kableExtra are not available for word and will yield errors. For example to knit to word, you'll have to comment out all "add_header_above()" functions for model output tables, or e.g. set always allow html as true in the yaml heading

always_allow_html: true

appendix:
  - "appendix_analytical.Rmd"
  - "appendix_simulations.Rmd"
  - "appendix_exp1.Rmd"
  - "appendix_exp2.Rmd"
  - "appendix_exp3.Rmd"
  - "appendix_exp4.Rmd"
  - "appendix_exp5.Rmd"
  - "appendix_exp6.Rmd"

bibliography: references.bib
---

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  # tidy.opts = list(width.cutoff = 120),  # Code width
  # fig.retina = 3, dpi = conditional_dpi,
  # fig.width = 7, fig.asp = 0.618,
  # fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  },
  options(scipen = 99999999)  # Prevent scientific notation
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_factory_default = table_format,  # Set modelsummary backend
  modelsummary_format_numeric_latex = "plain"  # Don't use siunitx
)
```

```{r packages, include=FALSE}
# load required packages
library("papaja")      # For APA style manuscript   
library("lme4")        # model specification / estimation
library("lmerTest")    # provides p-values in the output
library("tidyverse")   # data wrangling and visualisation
library("afex")        # anova and deriving p-values from lmer
library("broom")       # extracting data from model fits 
library("broom.mixed") # extracting data from mixed models
library("metafor")     # doing mata analysis
library("patchwork")   # put several plots together
library("ggridges")    # for plots
library("gghalves")    # for plots
library("ggbeeswarm")  # Special distribution-shaped point jittering
library("knitr")       # for tables
library("kableExtra")  # also for tables
library("ggpubr")      # for combining plots with ggarrange() 
library("grid")        # for image plots   
library("gridExtra")   # for image plots
library("png")         # for image plots
```

```{r functions}
# load plot theme
source("functions/plot_theme.R")

# load other functions
source("functions/own_functions.R")
```

# Introduction

Imagine that you live in ancient Greece, and a fellow called Eratostenes claims the circumference of the earth is 252000 stades (approximately 40000 kilometers). You know nothing about this man, the circumference of the Earth, or how one could measure such a thing. As a result, you discard Eratostenes' opinion and (mis)take him for a pretentious loon. But what if other scholars had arrived at very similar measurements, independently of Eratosthenes? Or even if they had carefully checked his measurement, with a critical eye? Wouldn't that give you enough ground to believe not only that the estimates might be correct, but also that Eratosthenes and his fellow scholars must be quite bright, to be able to achieve such a feat as measuring the Earth?

In this article, we explore how, under some circumstances, we should, and we do infer that a group of individuals whose answers converge are likely to be correct, and to be competent in the relevant area, even if we had no prior belief about either what the correct answer was, or about these individuals' competence.

We first review experiments testing whether people infer that individuals whose answers converge are more likely to be right, and to be competent. We then argue that little attention has been paid to the normative models that might explain when and why an inference from convergence to competence is warranted. We develop such models, both analytically and with simulations. Finally, we present a series of experiments in which participants evaluate the accuracy and competence of individuals as a function of how much their answers converge on a given answer, in the absence of any priors about these individuals’ competence, or what the correct answer is.

## Do people infer that individuals whose answers converge tend to be right, and to be competent?

There are two ways for answers to converge: when the answers are continuous, they can have a smaller range, or a lower variance; when the answers are categorical (i.e. there are only a few options), more answers can be the same. These two cases have been studied with different paradigms.

In the continuous case, the most relevant evidence comes from the literature on 'advice-taking.' In these experiments, participants are called 'judges' who need to make numerical estimates--sometimes on factual knowledge, e.g. 'What year was the Suez Canal opened first? [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g. 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020]. To help answer these questions, participants are given estimates from others, the 'advisors'.

Most of this literature is irrelevant to the point at hand since participants are presented with single estimates, either from a single advisor [e.g. @bednarikEffectivenessImperfectWeighting2015a; @sollStrategiesRevisingJudgment2009; @yanivReceivingOtherPeople2004; @yanivAdviceTakingDecision2000; @harveyTakingAdviceAccepting1997], or as an average coming from a group of advisors [e.g. @jaylesHowSocialInformation2017a; @mannesAreWeWise2009], but without any information about the distribution of initial estimates, so that we cannot tell whether participants put more weight on more convergent answers.

Some advice-taking studies provide participants with a set of individual estimates. One subset of these studies manipulates the degree of convergence between groups of advisors, through the variance of estimates [@mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009], or their range [@budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007]. These studies find that participants are more confident about, or rely more on, estimates from groups of advisors that converge more.

Other studies manipulated the degree of convergence within a group of advisors. These studies present participants with a set of estimates, some of which are close to each other , while others are outliers [@harriesCombiningAdviceWeight2004; @yanivWeightingTrimmingHeuristics1997, study 3 & 4]. These studies find that participants discount outliers when aggregating estimates.

Studies on advice taking thus suggest participants believe that more convergent opinions are more likely to be correct. None of these studies investigated whether participants also believe that those whose opinions converge are also more likely to possess greater underlying competence.

In categorical choice contexts, there is ample and long-standing [e.g. @crutchfieldConformityCharacter1955] evidence from social psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms [e.g., @morganEvolutionaryBasisHuman2012; for review, see @mercierMajorityRulesHow2019]. This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role [e.g. because the answers are private, see @aschStudiesIndependenceConformity1956]. Similar results have been obtained with young children [e.g. @fusaroChildrenAssessInformant2008; @corriveauGoingFlowPreschoolers2009; @bernardChildrenWeighNumber2015a; @bernardFourSixYearOldChildren2015; @chenChildrenTrustConsensus2013; @herrmannStickScriptEffect2013; @morganDevelopmentAdaptiveConformity2015].

If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the individuals' competence. One potential exception is a study with preschoolers in which the children were more likely to believe the opinion of an informant who had previously been the member of a majority over that of an informant who had dissented from the majority [@corriveauGoingFlowPreschoolers2009]. However, it is not clear whether the children thought the members of the majority were particularly competent, since their task--naming an object--was one in which children should already expect a high degree of competence from (adult) informants. This result might thus indicate simply that children infer that someone who disagrees with several others on how to call something is likely wrong, and thus likely less competent at least in that domain.

## Is inferring competence from convergence justified?

There is clear evidence that, both in the continuous and the categorical case, participants infer that convergent opinions are more likely to be correct. There is tentative evidence that participants also infer that individuals who hold more convergent opinions are more likely to be competent. Would that second inference be justified? More precisely, if we consider individuals who have convergent opinions, by contrast with individuals whose opinions are less convergent, or whose opinion diverges from that of a majority, are we justified in thinking that the individuals with more convergent opinions are more likely to be right not only for this specific question, but that they would also be right for other (related) questions? We first provide a general analytic answer, before presenting simulations for the continuous, and then the categorical case.

### Analytical argument

Regarding the analytical answer, the question can be broken down into two questions. First, can we infer that a population of informants whose answers converge more is, on average, more competent? Second, can we infer that, within this population, individuals who are closer to the consensus or to the average answer are more competent?

In the continuous case, we imagine a population of informants. The answers of each informant are drawn from a normally distributed random variable, centered on the correct answer (i.e., informants have no systematic bias). The variance of this distribution represents the individuals’ competence: the larger the variance, the lower the competence. If we observe a sample of this population, the sample mean (i.e. the mean of the answers of all the informants) is then the best estimate of the correct answer, and the sample variance (i.e. mean squared distance between the answers and the sample mean) is the best estimate of the population’s average competence (best is understood here as the least volatile estimator). This means that a population of informants whose answers converge more (lower sample variance) is, on average, more competent (informants tend to answer with a lower variance).

We can extend this argument from populations to individuals: Consider that competence varies within a population, such that each informant's answer is drawn from their own individual distribution--always centered on the correct answer, but with different variances. In this case, the distance between each informant's answer and the sample mean provides the best estimate of that informant's competence (i.e. individuals whose answers are further away from the sample mean tend to be less competent).

In the categorical case, we assume that informants answer a question with several categorical answers. We define the competence of an informant as the probability of choosing the correct answer. Due to the law of large numbers, the relative size of the majority--the share of informants who choose the most chosen answer--is, when the population is large enough, a good approximation of the average competence of the population (e.g. if the average competence is .66, and there are enough informants, then approximately 66% of informants will select the right answer). This relationship still exists, but is more noisy for small populations. This means that, at the population level, the more the answers converge, the more the population is, on average, competent (and the larger the population, the stronger the inference).

To make inferences on individual informants' competence on the basis of whether or not their answers converge with those of the majority in the sample, we must assume some distribution of competence in the population (assuming, as above, that the true average population competence is equal to the observed majority size in the sample). Using Bayes theorem, we can then show (see Electronic Supplementary Materials, ESM, section \@ref(analytical)) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig. \@ref(fig:figure-Benoit-2)). Additionally, the more competence varies in the population, the more the degree of convergence is informative of an informant's competence. It makes sense: if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence. More generally, the fact that the focal individual is right (their 'accuracy') can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa). Figure \@ref(fig:figure-Benoit-2) provides an example of a specific categorical choice, for two different distributions of competence.

(ref:figure-Benoit-2) Average estimated competence and accuracy for an individual, depending on how many participants agree with him, in a categorical choice scenario. Here, we assume that 20 individuals answer a 5-choice question. For values higher than four, we assume that the individual is part of the majority. Below four, we assume that the focus individual is part of a minority, and that the majority size is four. Two situations are represented: one in which there is low variance in the distribution of competence in the population (orange line in the insert), and one in which there is high variance in this distribution (blue line in the insert).

```{r figure-Benoit-2, echo=FALSE, fig.cap="(ref:figure-Benoit-2)", out.width= "50%", fig.align="left"}
knitr::include_graphics("figures/figure_Benoit-2.png")
```

### Simulations

In order to better understand the influence of different parameters (e.g. degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence of informants, we conducted simulations.

#### Continuous case

```{r read-simulation-data-numeric, message=FALSE}
# read data from simulations
simulation_numeric <- read_csv("Experiment_1/data/sim.csv") %>% 
  # remove alpha and beta values from competence variable strings
  mutate(competence = sub("\n.*", "", competence))
```

Groups of agents, with each agent’s competence varying, provide numerical answers. We measure how accurate these answers are, and how much they converge (i.e. how low their variance is). We then look at the relationship between convergence and both the accuracy of the answers and the competence of the agents.

More specifically, agents provide an estimate on a scale from 1000 to 2000 (chosen to match the experiments below). Each agent is characterized by a normal distribution of possible answers. All of the agents’ distributions are centered around the correct answer, but their standard deviation varies, representing varying degrees of competence. The agents’ standard deviation varies from 1 (highest competence) to 1000 (lowest competence). Each agent’s competence is drawn from a population competence distribution, expressed by a beta distribution, which can take different shapes. We conducted simulations with a variety of beta distributions which cover a wide range of possible competence populations (see Fig. \@ref(fig:simulations) A).

A population of around 990000 agents (varying slightly with as a function of group sizes) with different competence levels is generated. An answer is drawn for each agent, based on their respective competence distribution. The accuracy of this answer is defined as the squared distance to the true answer. Having a competence and an accuracy value for each agent, we randomly assign agents to groups of, e.g., three. For each group, we calculate the average of the agents' competence and accuracy. We measure the convergence of a group’s answers by calculating the standard deviation of the agents' answers. We repeat this process for different sample sizes for the groups, and different competence distributions. Fig. \@ref(fig:simulations) C displays the resulting relation of convergence with accuracy (left), and competence (right) for different underlying competence distributions and group sizes. We draw broad conclusions from these results after reporting the outcome of the simulations with categorical answers.

#### Categorical case

```{r read-simulation-data-categorical, message=FALSE}
# read data from simulations
simulation_sample_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_sample_3_options.csv")  %>% 
  # remove alpha and beta values from competence variable strings
  mutate(competence = sub("\n.*", "", competence))

simulation_options_categorical <- read_csv("Experiment_4/data/sim_rel_majority_vary_competence_options_3_sample.csv")  %>% 
  # remove alpha and beta values from competence variable strings
  mutate(competence = sub("\n.*", "", competence))
```

In the case of categorical answers, convergence can be measured as the relative size of informants picking the same option. The Condorcet jury theorem [@decondorcetEssaiApplicationAnalyse2014; for a recent treatment, see @dietrichEpistemicDemocracyDefensible2013] and its extensions to situations with more than two options [e.g., @hastieRobustBeautyMajority2005] already shows that the answer defended by the majority/plurality is more likely to be correct. Regarding competence, @romeijnLearningJurorCompetence2011 have shown that when individuals are more competent, the share of the majority vote tends to increase. However, they have only studied a case with a uniform distribution of competence. By contrast, here, we investigate a wide range of distributions of competence, and we do not assume that all individuals are equally competent, meaning that, from an observed set of answers, we can attribute an average competence to individuals whose answers form a majority/plurality vs. individuals whose answers form a minority.

We simulate agents whose competence varies and who have to decide between a number of options, one of which is correct. Competence is defined as a value p which corresponds to the probability of selecting the right answer (the agents then have a probability (1-p)/(m-1), with m being the number of options, of selecting any other option). Competence values range from chance level (p = 1/m) to always selecting the correct option (p = 1). Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see see Fig. \@ref(fig:simulations) A). Based on their competence level, we draw an answer for each agent. We measure an agent's accuracy as a binary outcome, namely whether they selected the correct option or not. In each simulation around 99900 agents (varying slightly as a function of the group size) are generated, and then randomly assigned to groups (of varying size in different simulations). Within these groups, we first calculate the share of individuals voting for each answer, allowing us to measure convergence. For each level of convergence occurring within a group, we then compute (i) the average accuracy, (ii) the average competence of agents. Across all groups, we then compute the averages of these within-group averages, for each level of convergence.

We repeat this procedure varying population competence distributions, and the size of informant groups, holding the number of choice options constant at n = 3 (for simulations with varying choice options, see ESM section \@ref(simulations)). Fig. \@ref(fig:simulations) B shows the average accuracy (left), and the average competence (right) value for each share of votes, across different underlying competence distributions and group sizes.

The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows: (i) Irrespective of group size (and number of choice options) and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate. (ii) For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents. The strength of this relation is not much affected by the number of agents whose answers are converging, but, although it is always positive, it ranges from very weak to very strong depending on the population's competence distribution. (iii) The relation between convergence and accuracy is always much stronger than the relation between convergence and average competence of the agents.

(ref:simulations) **A** Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1. In the categorical simulations, a competence value of 0 corresponds to chance (e.g. in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer. **B** Shows the results of simulations in a categorical setting with three choice options. Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes. **C** Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the scale) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).

```{r simulations, fig.cap="(ref:simulations)", fig.height= 10, fig.width=10}
# Plot continuous choice scenario
plot_continuous_accuracy <- plot_continuous_cogsci(simulation_numeric, outcome = "accuracy") + 
  rremove("xlab") +
  guides(color = "none") +
  labs(caption = NULL) 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plot_continuous_competence <- plot_continuous_cogsci(simulation_numeric, outcome = "competence") + 
  rremove("xlab") +
  guides(color = "none") +
  labs(caption = NULL) 
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plots_continuous <- ggarrange(plot_continuous_accuracy, plot_continuous_competence) %>% 
  annotate_figure(bottom = textGrob("Convergence (SDs of samples)", gp = gpar(cex = 1, fontface = "bold")))

# Plot categorical choice scenario
plot_categorical_accuracy <- plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Accuracy", variable = sample) + 
  rremove("xlab") +
  labs(caption = NULL) 
  # theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plot_categorical_competence <- plot_competence_vary_relative_majority_categorical(simulation_sample_categorical,
                                       outcome = "Competence", variable = sample) + 
  rremove("xlab") +
  labs(caption = NULL) 
  # theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

plots_categorical <- ggarrange(plot_categorical_accuracy, plot_categorical_competence, common.legend = TRUE) %>% 
  annotate_figure(bottom = textGrob("Convergence (share of votes for an option)", gp = gpar(cex = 1, fontface = "bold")))

# common plot
patchwork <- (plot_competence_distributions(simulation_numeric) + plot_spacer()) / 
  plots_categorical /
  plots_continuous

patchwork + 
  plot_annotation(tag_levels = 'A')

```

## Overview of the experiments

We test whether people infer that informants are more likely to give accurate answers, and to be competent, when their answers converge, both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6). By contrast with previous studies, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be. We also manipulated (Experiments 2, 3, and 5) whether the convergence of the answers could be explained by something else than accuracy, which should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.

All experiments were preregistered. Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/) (<https://osf.io/6abqy/>). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the `lme4` package and its `lmer()` function.

# Experiment 1

```{r exp1}
# Analyze data of experiments and store results

# Experiment 1

# read data
exp1 <- read_csv("Experiment_1/data/cleaned.csv") %>% 
  # import experimental variables as factors and set levels
  mutate(
    across(c(convergence, number), ~as.factor(.x)),
    convergence = fct_relevel(convergence, "divergent", "convergent"), 
    number = fct_relevel(number, "small", "large"))

# Hypotheses

# run mixed model (random slope and intercept) with participants as random factor
exp1_model_accuracy = lmer(confidence ~ convergence + number + (1 + convergence | ID), data = exp1)
exp1_model_competence = lmer(competence ~ convergence + number + (1 + convergence | ID), data = exp1)

# extract descriptives for inline reporting
exp1_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp1$ID),
  gender = exp1 %>% group_by(gender) %>% summarize(n = n_distinct(ID)) %>% split(.$gender),
  age = exp1 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_accuracy = text_ready(exp1_model_accuracy), 
  model_competence = text_ready(exp1_model_competence),
  # Means
  means = exp1 %>% 
    group_by(convergence) %>% 
    summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
    rounded_numbers() %>% 
    split(.$convergence)
)

# Research questions

# RQ1
# run mixed model from main hypotheses, but without number as covariate and separately for 10 or 3 informant conditions
model_small_accuracy = lmer(confidence ~ convergence + (1 + convergence | ID), 
                                  data = exp1 %>% filter(number =="small"))
model_large_accuracy = lmer(confidence ~ convergence + (1 + convergence | ID), 
                                  data = exp1 %>% filter(number =="large"))
model_small_competence = lmer(competence ~ convergence + (1 + convergence | ID), 
                                  data = exp1 %>% filter(number =="small"))
model_large_competence = lmer(competence ~ convergence + (1 + convergence | ID), 
                                  data = exp1 %>% filter(number =="large"))

# RQ2 (as separate t-test):
accuracy <- t.test(confidence ~ number, data = exp1, paired = TRUE)
competence <- t.test(competence ~ number, data = exp1, paired = TRUE)

# RQ2 and RQ3 (as single mixed model):

# effect code variables
exp1 <- exp1 %>% 
  mutate(convergence_effect_code = recode(convergence, "divergent" = -0.5, "convergent" = +0.5),
         number_effect_code = recode(number, "small" = -0.5, "large" = +0.5))

model_interaction_competence = lmer(competence ~ convergence_effect_code + number_effect_code +
                                      convergence_effect_code*number_effect_code + 
                                      (1 + convergence_effect_code | ID), data = exp1)

model_interaction_accuracy = lmer(confidence ~ convergence_effect_code + number_effect_code +
                                    convergence_effect_code*number_effect_code + 
                                    (1 + convergence_effect_code | ID), data = exp1)

# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? 
exp1_explorative <- exp1 %>% 
  pivot_longer(
    c(confidence, competence),
    names_to = "outcome", 
    values_to = "score")

# run mixed model (random slope and intercept) with participants as random factor
exploratory_analysis <- lmer(score ~ convergence_effect_code + outcome + 
                               convergence_effect_code*outcome + number +
                               (1 + convergence_effect_code + outcome | ID), 
                             data = exp1_explorative)

exp1_research_questions <- list(
  # RQ1
  model_small_accuracy = text_ready(model_small_accuracy),
  model_large_accuracy = text_ready(model_large_accuracy),
  model_small_competence = text_ready(model_small_competence),
  model_large_competence = text_ready(model_large_competence),
  # speparate t-tests for RQ2
  means = exp1 %>% 
    group_by(number) %>% 
    summarise(across(c(confidence, competence), 
                     mean)) %>% 
    rounded_numbers() %>% 
    split(.$number), 
  accuracy = clean_t_test(accuracy), 
  competence = clean_t_test(competence),
  # RQ2 and RQ3 in a single mixed model
  model_interaction_competence = text_ready(model_interaction_competence), 
  model_interaction_accuracy = text_ready(model_interaction_accuracy),
  # Exploratory analysis
  exploratory_analysis = text_ready(exploratory_analysis)
)
```

In Experiment 1, participants were provided with a set of numerical estimates which were more or less convergent, and asked whether they thought the estimates were accurate, and whether the informants making the estimates were competent. Perceptions of accuracy were measured as the confidence in what the participants thought the correct answer was, on the basis of the numerical estimates provided: the more participants think they can confidently infer the correct answer, the more they must think the estimates accurate, on average. Our hypotheses were:

***H1: When making a guess based on the estimates of (independent) informants, participants will be more confident about their guess when these estimates converge compared to when they diverge.***

***H2: Participants perceive (independent) informants whose estimates converge more as more competent than informants whose estimates diverge.***

We had three research questions regarding the number of informants which report in the ESM (section \@ref(exp1)).

## Methods

### Participants

We recruited `r exp1_descriptives$n_subj` participants from the UK via Prolific (`r exp1_descriptives$gender$female$n` female, `r exp1_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp1_descriptives$age$mean`, $age_\text{sd}$: `r exp1_descriptives$age$sd`, $age_\text{median}$: `r exp1_descriptives$age$median`). Not a single participant failed our attention check. The sample size was determined on the basis of a power analysis for a t-test to detect the difference between two dependent means ("matched pairs") run on G\*Power3. The analysis suggested that a combined sample of 199 would provide us with 80% power to detect a true effect size of Cohen's d $\geq$ 0.2 (alpha = .05, two-tailed).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "Some people are playing games in which they have to estimate various quantities. Each game is different. You have no idea how competent the people are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be hard while others are easy. Across various games, we will give you the answers of several players, and ask you questions about how good they are. As it so happens, for all the games, the estimates have to be between 1000 and 2000, but all the games are completely different otherwise, and might require different skills, be of different difficulties, etc. Each player in the game makes their own estimate, completely independent of the others". After being presented with the results of a game (Fig. \@ref(fig:stimulus-exp1)), participants had to (i) make a guess about the correct answer based on the estimates they see, "What would you guess is the correct answer, if there is one?", (ii) estimate their confidence in this guess, "How confident are you that your answer is at least approximately correct?" on a 7-point Likert scale ("not confident at all" to "extremely confident"), (iii) estimate the competence of the group of players whose estimates they saw, "On average, how good do you think these players are at the game?", also on a 7-point Likert scale (from "not good at all" to "extremely good").

(ref:stimulus-exp1) Example of two stimuli from Experiment 1, both in the 10 players condition, **A** corresponding to the convergent, **B** to the divergent condition. Similar stimuli are used in Experiments 2 and 3.

```{r stimulus-exp1, echo=FALSE, out.width= "100%", fig.align="left", fig.show="hold", fig.cap="(ref:stimulus-exp1)"}
# Function to create a plot with image and title
create_image_plot <- function(image_path, title) {
  img <- rasterGrob(readPNG(image_path), interpolate = TRUE)
  title_grob <- textGrob(title, gp = gpar(fontsize = 8, fontface = "bold"), just = "left")

  # Combine image and title in a single plot
  plot <- ggplot() +
    annotation_custom(img, xmin = 0, xmax = 1, ymin = 0, ymax = 1) +
    annotation_custom(title_grob, xmin = 0, xmax = 0.2, ymin = 1, ymax = 1) +
    theme_void()  # Set the panel background to be transparent +
    theme(plot.margin = margin(0, 0, 0, 0, "cm"))  # Set margins to zero

  return(plot)
}

# File paths to four images
image_paths <- c(
  "figures/example_stimulus_exp1_convergent.png",
  "figures/example_stimulus_exp1_divergent.png"
)

# Titles for the images
image_titles <- c(
  "A", 
  "B"
)

# Create a list of plots
plots <- lapply(seq_along(image_paths), function(i) {
  create_image_plot(image_paths[i], image_titles[i])
})

# Arrange the plots in a grid
grid.arrange(grobs = plots, ncol = 1)
```

### Design

We manipulated two experimental factors, with two levels each: the convergence of the estimates (how close they were to each other; levels: divergent/convergent), and the number of estimates (how many players there were; levels: low/high). This latter factor was chiefly included to make our results more robust, and was not attached to specific hypotheses. We used a 2 (convergence: divergent/convergent) x 2 (number: low/high) within-participant design, with each participant seeing all the conditions. Participants saw two different sets of estimates per condition, for a total of eight sets of estimates per participant.

### Materials

We generated sets of estimates with random draws from normal distributions. First, we varied the standard deviation of these distributions to simulate the degree of convergence (150 for divergence, 20 for convergence; estimate scale ranged from 1000 to 2000). Second, we varied the number of draws (either three or ten) from these distributions. For each of the four possible resulting conditions, we generated two random draws. We repeated this process for three different sets of estimates, and participants were randomly assigned to one of these sets. More information on how the stimuli were created can be found in the ESM, section \@ref(exp1).

## Results and discussion

To account for dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slopes of convergence for participants. In the models for our hypotheses, we control for the number of estimates provided to the participants (three or ten). Visualizations and descriptive statistics can be found in ESM, section \@ref(exp1). We find a positive effect of convergence on accuracy: Participants were more confident about their estimate in convergent scenarios (mean = `r exp1_descriptives$means$convergent$confidence_mean`, sd = `r exp1_descriptives$means$convergent$confidence_sd`) than in divergent ones (mean = `r exp1_descriptives$means$divergent$confidence_mean`, sd = `r exp1_descriptives$means$divergent$confidence_sd`; $\hat{b}_{\text{Accuracy}}$ = `r exp1_descriptives$model_accuracy$convergenceconvergent$estimate` `r exp1_descriptives$model_accuracy$convergenceconvergent$ci`, p = `r exp1_descriptives$model_accuracy$convergenceconvergent$p.value`). We also find a positive effect of convergence on competence: participants rated players as more competent in convergent scenarios (mean = `r exp1_descriptives$means$convergent$competence_mean`, sd = `r exp1_descriptives$means$convergent$competence_sd`) than in divergent ones (mean = `r exp1_descriptives$means$divergent$competence_mean`, sd = `r exp1_descriptives$means$divergent$competence_sd`; $\hat{b}_{\text{Competence}}$ = `r exp1_descriptives$model_competence$convergenceconvergent$estimate` `r exp1_descriptives$model_competence$convergenceconvergent$ci`, p = `r exp1_descriptives$model_competence$convergenceconvergent$p.value`).

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To this end, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for the number of informants. We do not find a statistically significant interaction that would indicate a difference of the effect of convergence ($\hat{b}$ = `r exp1_research_questions$exploratory_analysis$interaction$estimate` `r exp1_research_questions$exploratory_analysis$interaction$ci`, `r exp1_research_questions$exploratory_analysis$interaction$p.value`). We find that pooled across divergent and convergent conditions, participants reported lower perceived accuracy than competence ($\hat{b}$ = `r exp1_research_questions$exploratory_analysis$outcomeconfidence$estimate` `r exp1_research_questions$exploratory_analysis$outcomeconfidence$ci`, `r exp1_research_questions$exploratory_analysis$outcomeconfidence$p.value`)

In summary, as predicted, when the informants’ answers were more convergent, participants were more confident that their answers were correct, and they believed the informants to be more competent. This was true whether there were three or ten estimates.

# Experiment 2

```{r exp2}
# Analyze data of experiments and store results

# Experiment 2

# read data
exp2 <- read_csv("Experiment_2/data/cleaned.csv") %>% 
  # import experimental variable as factors and set levels
  mutate(
    independence = as.factor(independence),
    independence = fct_relevel(independence, "dependent", "independent")
  )

# run mixed model (random intercept) with participants as random factor
exp2_model_accuracy = lmer(confidence ~ independence + (1 |ID), data = exp2)
exp2_model_competence = lmer(competence ~ independence + (1 |ID), data = exp2)

# extract descriptives for inline reporting
exp2_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp2$ID),
  gender = exp2 %>% group_by(gender) %>% summarize(n = n_distinct(ID)) %>% split(.$gender),
  age = exp2 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_accuracy = text_ready(exp2_model_accuracy), 
  model_competence = text_ready(exp2_model_competence),
  # Means
  means = exp2 %>% 
    group_by(independence) %>% 
    summarise(across(c(competence, confidence), list(mean = mean, sd = sd))) %>% 
    rounded_numbers() %>% 
    split(.$independence)
)
```

It is rational to infer that convergent estimates are more likely to be accurate, and to have been made by competent individuals, only if the convergence is best explained by the accuracy of the estimates. However, convergence could also be the outcome of other factors. If the individuals do not make their estimates independently of each other, a single individual might exert a strong influence on the others, making their convergence a poor cue to their accuracy. Alternatively, all individuals might have an incentive to provide a similar, but not accurate answer. In Experiment 2, we investigate the first possibility, and the second in Experiment 3. In particular, for Experiment 2 we rely on past results showing that participants, under some circumstances, put less weight on opinions that have been formed through discussion, by contrast with more independent opinions [@harkinsInformationUtilityMultiple1987; see also @lopesSocialValidationEveryday2007; @hessPsychologicalAdaptationsAssessing2006; @einavThinkingThemselvesEffect2018]. We sought to replicate this finding in the context of convergent estimates, formulating the following hypotheses:

***H1: When making a guess based on convergent estimates of informants, participants will be more confident about their guess when informants were independent compared to when they weren't (i.e. they could discuss before).***

***H2: Participants perceive informants whose estimates converge as more competent when they are independent, compared to when they weren't (i.e. they could discuss before).***

## Methods

### Participants

We recruited `r exp2_descriptives$n_subj` participants from the UK via Prolific (`r exp2_descriptives$gender$female$n` female, `r exp2_descriptives$gender$male$n` male, 1 not-identified; $age_\text{mean}$: `r exp2_descriptives$age$mean`, $age_\text{sd}$: `r exp2_descriptives$age$sd`, $age_\text{median}$: `r exp2_descriptives$age$median`). Not a single participant failed our attention check. As for experiment 1, the sample size was determined on the basis of a power analysis for a t-test to detect the difference between two dependent means ("matched pairs") run on G\*Power3. The analysis suggested that a combined sample of 199 would provide us with 80% power to detect a true effect size of Cohen's d $\geq$ 0.2 (alpha = .05, two-tailed).

### Design

In a within-participants design, participants saw both an independence condition, in which they were told "Players are asked to make completely independent decisions -- they cannot see each other's estimates, or talk with each other before giving their estimates," and a dependence condition, in which they were told "Players are asked to talk with each other about the game at length before giving their estimates."

### Materials

We used the materials generated for the convergent condition of Experiment 1. By contrast to Experiment 1, participants saw only two stimuli in total (one set of estimates per condition). Otherwise, we proceeded just as in experiment one: we randomly assigned individual participants to one of the three series of stimuli, and for each participant, we randomized the order of appearance of conditions.

## Results and discussion

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept for participants. Visualizations and descriptive statistics can be found in ESM, section \@ref(exp2). The data does not support our hypotheses. Participants were slightly less confident about their estimates when the converging informants were independent (mean = `r exp2_descriptives$means$independent$confidence_mean`, sd = `r exp2_descriptives$means$independent$confidence_sd`), compared to when they discussed (mean = `r exp2_descriptives$means$dependent$confidence_mean`, sd = `r exp2_descriptives$means$dependent$confidence_sd`; $\hat{b}_{\text{Accuracy}}$ = `r exp2_descriptives$model_accuracy$independenceindependent$estimate` `r exp2_descriptives$model_accuracy$independenceindependent$ci`, p = `r exp2_descriptives$model_accuracy$independenceindependent$p.value`). The effect is small, but in the opposite direction of what we had predicted. We do not find an effect regarding competence ($\hat{b}_{\text{Competence}}$ = `r exp2_descriptives$model_competence$independenceindependent$estimate` `r exp2_descriptives$model_competence$independenceindependent$ci`, p = `r exp2_descriptives$model_competence$independenceindependent$p.value`).

Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently, to be less accurate, or made by less competent individuals. This suggests that participants might not be sensitive to at least some forms of informational dependency between the individuals making the estimates. Although previous studies have found that participants sometimes discount the opinions of groups compared to those of independent individuals, the superior performance of groups over individuals for a range of tasks [for review, see, e.g., @mercierArgumentativeTheoryPredictions2016a], including numerical estimates [e.g. @mercierDoesDiscussionMake2022] suggests that this discounting might have been misguided. As a result, the participants in the current experiment might have been behaving rationally when they did not discount the estimates made after discussion.

# Experiment 3

```{r exp3}
# Analyze data of experiments and store results

# Experiment 3

# read data
exp3 <- read_csv("Experiment_3/data/cleaned.csv") %>% 
  # import experimental variable as factors and set levels
  mutate(
    across(c(convergence, independence), ~as.factor(.x)),
    convergence = fct_relevel(convergence, "divergent", "convergent"),
    independence = fct_relevel(independence, "conflict", "independent")
  )

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the independence condition
exp3_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence |ID), 
                           data = exp3 %>% filter(independence == "independent"))
exp3_model_h1_competence = lmer(competence ~ convergence + (1 + convergence |ID), 
                             data = exp3 %>% filter(independence == "independent"))

# H2
# run mixed model (random intercept and slope for convergence) with participants as random factor
exp3_model_accuracy = lmer(accuracy ~ convergence + independence + 
                            convergence*independence + (1 + convergence | ID), 
                           data = exp3)

exp3_model_competence = lmer(competence~ convergence + independence + 
                            convergence*independence + (1 + convergence | ID), 
                           data = exp3)

# extract descriptives for inline reporting
exp3_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp3$ID),
  gender = exp3 %>% group_by(gender) %>% summarize(n = n_distinct(ID)) %>% split(.$gender),
  age = exp3 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_h1_accuracy = text_ready(exp3_model_h1_accuracy), 
  model_h1_competence = text_ready(exp3_model_h1_competence),
  model_accuracy = text_ready(exp3_model_accuracy), 
  model_competence = text_ready(exp3_model_competence),
  # Means
  means_H1 = exp3 %>% 
    filter(independence == "independent") %>% 
    group_by(convergence) %>% 
    summarise(across(c(accuracy, competence), list(mean = mean, sd = sd))) %>% 
    rounded_numbers() %>% 
    split(.$convergence), 
  means_H2 = exp3 %>% 
    group_by(convergence, independence) %>% 
    summarise(across(c(accuracy, competence), list(mean = mean, sd = sd))) %>% 
    rounded_numbers() %>% 
    super_split(convergence, independence)
)

# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? (in the independence condition)
exp3_explorative <- exp3 %>% 
  pivot_longer(
    c(accuracy, competence),
    names_to = "outcome", 
    values_to = "score") %>% 
  # effect code variables
  mutate(convergence_effect_code = recode(convergence, "divergent" = -0.5, 
                                          "convergent" = 0.5))

# run mixed model (random slope and intercept) with participants as random factor
exp3_exploratory_analysis <- lmer(score ~ convergence_effect_code + outcome + 
                               convergence_effect_code*outcome + independence +
                               (1 + convergence_effect_code + outcome | ID), 
                             data = exp3_explorative) %>% 
  text_ready()
```

Experiment 3 tests whether participants are sensitive to another potential source of dependency between convergent estimates: when the individuals making the estimate share an incentive to bias their estimates and disregarde accuracy. Even though Experiment 3 is formally similar to Experiment 1, the setting is different, as participants were told that they would be looking at (fictional) predictions of experts for stock values, instead of the answers of individuals in abstract games. In the conflict of interest condition, the experts had an incentive to value the stock in a given way, while they had no such conflict of interest in the independence condition. Given the nature of this manipulation, participants might have discounted the opinions of the conflicted experts, irrespective of the degree of convergence of their estimates. As a result, we could not just condition on convergent answers, as we did in Experiment 2. Instead, we tested for an interaction, namely whether the positive effect of convergence is reduced when informants are systematically biased, compared to when they are not. On this basis, we formulate four hypotheses, two of which are identical to those of Experiment 1, and only apply in the independent condition, and two that bear on the interaction:

***H1a: Participants perceive predictions of independent informants as more accurate when they converge compared to when they diverge.***

***H1b: Participants perceive independent informants as more competent when their predictions converge compared to when they diverge.***

***H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are in a conflict of interest.***

***H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are in a conflict of interest.***

## Methods

### Participants

The interaction design of our third experiment made the power analysis more complex and less standard than for experiments one and two. Because we could build upon data from the first experiment, we ran a power analysis by simulation. The simulation code is available on OSF, and the procedure is described in the preregistration document. The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% of the time, given an alpha threshold for significance of 0.05. Due to uncertainty about our effect size assumptions and because we had resources for a larger sample, we recruited `r exp3_descriptives$n_subj` participants for this study -- again, from the UK and via Prolific (`r exp3_descriptives$gender$female$n` female, `r exp3_descriptives$gender$male$n` male; $age_\text{mean}$: `r exp3_descriptives$age$mean`, $age_\text{sd}$: `r exp3_descriptives$age$sd`, $age_\text{median}$: `r exp3_descriptives$age$median`).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "You will see four scenarios in which several experts predict the future value of a stock. You have no idea how competent the experts are. It's also possible that some are really good while others are really bad. As it so happens, in all scenarios, the predictions for the value of the stock have to lie between 1000 and 2000. Other than that, the scenarios are completely unrelated: it is different experts predicting the values of different stocks every time." Participants then saw the four scenarios, each introduced by a text according to the condition the participant was assigned to. To remove any potential ambiguity about participants' inferences on the accuracy of the estimates, we replaced the question about confidence to one bearing directly on accuracy: "On average, how accurate do you think these three predictions are?" on a 7-point Likert scale ("not accurate at all" to "extremely accurate"). The question about competence read: "On average, how good do you think these three experts are at predicting the value of stocks?", also assessed on a 7-point Likert scale (from "not good at all" to "extremely good").

### Design

We manipulated two factors: informational dependency (two levels, independence and conflict of interest; between participants) and convergence (two levels, convergence and divergence; within participants). In the independence condition, the participants read "Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock." In the conflict of interest condition, the participants read "All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future."

### Materials

The distributions presented were similar to those of Experiment 1, although generated in a slightly different manner (see ESM, section \@ref(exp3)). Each participant rated four scenarios, two for each level of convergence.

## Results and discussion

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants. We find evidence for all four hypotheses. As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample of experiment three to half of the participants, namely those who were assigned to the independence condition. On this reduced sample, we ran the exact same analyses as in experiment 1 and replicated the results. As for accuracy, participants rated informants in convergent scenarios (mean = `r exp3_descriptives$means_H1$convergent$accuracy_mean`, sd = `r exp3_descriptives$means_H1$convergent$accuracy_sd`) as more accurate than in divergent ones (mean = `r exp3_descriptives$means_H1$divergent$accuracy_mean`, sd = `r exp3_descriptives$means_H1$divergent$accuracy_sd`; $\hat{b}_{\text{Accuracy}}$ = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$estimate` `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$ci`, p = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$p.value`). As for competence, participants rated informants in convergent scenarios (mean = `r exp3_descriptives$means_H1$convergent$competence_mean`, sd = `r exp3_descriptives$means_H1$convergent$competence_sd`) as more competent than in divergent ones (mean = `r exp3_descriptives$means_H1$divergent$competence_mean`, sd = `r exp3_descriptives$means_H1$divergent$competence_sd`; $\hat{b}_{\text{Competence}}$ = `r exp3_descriptives$model_h1_competence$convergenceconvergent$estimate` `r exp3_descriptives$model_h1_competence$convergenceconvergent$ci`, p = `r exp3_descriptives$model_h1_competence$convergenceconvergent$p.value`).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \@ref(fig:exp3-plot)). In the independence condition, the effect of convergence on accuracy was more positive ($\hat{b}_{\text{interaction, Accuracy}}$ = `r exp3_descriptives$model_accuracy$interaction$estimate` `r exp3_descriptives$model_accuracy$interaction$ci`, p = `r exp3_descriptives$model_accuracy$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ = `r exp3_descriptives$model_accuracy$convergenceconvergent$estimate` `r exp3_descriptives$model_accuracy$convergenceconvergent$ci`, p = `r exp3_descriptives$model_accuracy$convergenceconvergent$p.value`). Likewise the effect of convergence on competence is more positive ($\hat{b}_{\text{interaction, Competence}}$ = `r exp3_descriptives$model_competence$interaction$estimate` `r exp3_descriptives$model_competence$interaction$ci`, p = `r exp3_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ = `r exp3_descriptives$model_competence$convergenceconvergent$estimate` `r exp3_descriptives$model_competence$convergenceconvergent$ci`, p = `r exp3_descriptives$model_competence$convergenceconvergent$p.value`).

(ref:exp3-plot) Distributions of accuracy and competence by convergence and informational dependency.

```{r exp3-plot, fig.cap="(ref:exp3-plot)"}
interaction_accuracy <- ggplot(exp3, aes(x=independence, y=accuracy, fill = convergence, 
                     shape = convergence,
                     group = convergence,
                     color = convergence)) +
  scale_x_discrete(limits = c("conflict", "independent"), 
                    labels = c("Conflict of interest", "Independent")) +
  geom_half_violin(data = exp3 %>% filter(independence=="conflict"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = exp3 %>% filter(independence=="independent"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Accuracy") +
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7)) +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none",
         fill = guide_legend(title = NULL)) +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

interaction_competence <- ggplot(exp3, aes(x=independence, y=competence, 
                                        fill = convergence, 
                                        shape = convergence, 
                                        group = convergence, 
                                        color = convergence)) +
  scale_x_discrete(limits = c("conflict", "independent"), 
                    labels = c("Conflict of interest", "Independent")) +
  geom_half_violin(data = exp3 %>% filter(independence=="conflict"), 
                   position = position_nudge(x = -.2), adjust=2, alpha = .4,
                   side = "l") +
  geom_half_violin(data = exp3 %>% filter(independence=="independent"), 
                   position = position_nudge(x = .2), adjust=2, alpha = .4,
                   side = "r") + 
  xlab("Condition") +
  ylab("Competence")+
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7)) +
  stat_summary(fun = "mean", geom = "point", size = 3) +
  stat_summary(fun = "mean", geom = "line") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  scale_fill_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  scale_color_manual(name = NULL,
    labels = c("divergent", "convergent"),
    values = c("#E69F00", "#56B4E9")) +
  guides(shape = "none") +
  plot_theme +
  # change font sizes
  theme(axis.text = element_text(size = 10)) +
  theme(axis.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))

figure <- ggarrange(interaction_accuracy, interaction_competence, 
                    common.legend = T) +
  theme(legend.position="top") 
figure
```

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To this end, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction effect, indicating that pooled across independent and conflict of interest conditions, the effect of convergence had as smaller effect on competence than on accuracy ($\hat{b}$ = `r exp3_exploratory_analysis$interaction$estimate` `r exp3_exploratory_analysis$interaction$ci`, `r exp3_exploratory_analysis$interaction$p.value`). Pooled across all conditions, participants reported higher perceived competence than accuracy ($\hat{b}$ = `r exp3_exploratory_analysis$outcomecompetence$estimate` `r exp3_exploratory_analysis$outcomecompetence$ci`, `r exp3_exploratory_analysis$outcomecompetence$p.value`).

Experiment 3 shows that, when the individuals making the estimates are systematically biased, participants put less weight on the convergence of their estimates to infer that the estimates are accurate, and that the individuals making them are competent.

# Experiment 4

```{r exp4}
# Analyze data of experiments and store results

# Experiment 4

# read data
exp4 <- read_csv("Experiment_4/data/cleaned.csv") %>% 
  # import experimental variable as factor and set levels
  mutate(
    # set levels for `convergence`
    convergence_categorical = recode_factor(convergence, 
                                            `0` = "minority", 
                                            `1` = "divergence", 
                                            `2` = "majority", 
                                            `3` = "consensus",
                                            .default = NA_character_))

# run mixed model (random slope and intercept) with participants as random factor
exp4_model_accuracy = lmer(accuracy ~ convergence + (1 + convergence | id), 
                                 data = exp4)
exp4_model_competence = lmer(competence ~ convergence + (1 + convergence | id), 
                                 data = exp4)

# extract descriptives for inline reporting
exp4_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp4$id),
  gender = exp4 %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp4 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_accuracy = text_ready(exp4_model_accuracy), 
  model_competence = text_ready(exp4_model_competence)
)

# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? 
exp4_explorative <- exp4 %>% 
  mutate(accuracy_scaled = scale(accuracy), 
         competence_scaled = scale(competence)) %>% 
  pivot_longer(
    c(accuracy_scaled, competence_scaled),
    names_to = "outcome", 
    values_to = "score")

# run mixed model (random slope and intercept) with participants as random factor
exp4_exploratory_analysis <- lmer(score ~ convergence + outcome + 
                               convergence*outcome + 
                               (1 + convergence | id), 
                             data = exp4_explorative) %>% 
  text_ready()
```

In a second series of experiments, we test similar predictions to those of the previous experiments, but in a categorical choice context. The set-up is similar to that of Experiment 1, except that the outcomes seen by the participants are not numerical estimates, but choices made between a few options. Experiment 4 tests hypotheses that are analogous to those of Experiment 1:

***H1: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.***

***H2: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.***

## Methods

### Participants

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%. The power simulation for accuracy suggested that even for as few as 10 participants (the minimum sample size we simulated data for), we would have a power of close to 100%. The simulation for competence suggested that we achieve statistical power of at least 90% with a sample size of 30. Due to uncertainty about our assumptions and because it was within our budget, we recruited `r exp4_descriptives$n_subj` participants, from the UK and via Prolific (`r exp4_descriptives$gender$female$n` female, `r exp4_descriptives$gender$male$n`; $age_\text{mean}$: `r exp4_descriptives$age$mean`, $age_\text{sd}$: `r exp4_descriptives$age$sd`, $age_\text{median}$: `r exp4_descriptives$age$median`).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "To be able to understand the task, please read the following instructions carefully: Some people are playing games in which they have to select the correct answer among three answers. You will see the results of several of these games. Each game is different, with different solutions and involving different players. All players answer independently of each other. At first, you have no idea how competent each individual player is: they might be completely at chance, or be very good at the task. It's also possible that some players are really good while others are really bad. Some games might be difficult while others are easy. Your task will be to evaluate the performance of one of the players based on what everyone's answers are." They were then presented to the results of eight such games (Fig. \@ref(fig:stimulus-exp4)). To assess perceived accuracy, we asked: "What do you think is the probability of player 1 being correct?". Participants answered with a slider on a scale from 0 to 100. To assess perceived competence, we asked participants: "How competent do you think player 1 is in games like these?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

(ref:stimulus-exp4) One set of stimuli by level of convergence, in Experiment 4 (similar stimuli are used in Experiments 5 and 6). A full set of stimuli can be found in the ESM, section \@ref(exp4).

```{r stimulus-exp4, echo=FALSE, fig.cap="(ref:stimulus-exp4)"}
# Function to create a plot with image and title
create_image_plot <- function(image_path, title) {
  img <- rasterGrob(readPNG(image_path), interpolate = TRUE)
  title_grob <- textGrob(title, gp = gpar(fontsize = 8, fontface = "bold"))

  # Combine image and title in a single plot
  plot <- ggplot() +
    annotation_custom(img, xmin = 0, xmax = 1, ymin = 0, ymax = 1) +
    annotation_custom(title_grob, xmin = 0, xmax = 1, ymin = 1, ymax = 1) +
    theme_void()  # Set the panel background to be transparent

  return(plot)
}

# File paths to four images
image_paths <- c(
  "Experiment_4/figures/stimuli/large_opp_majority_a.png",
  "Experiment_4/figures/stimuli/large_divergence_a.png",
  "Experiment_4/figures/stimuli/large_majority_a.png",
  "Experiment_4/figures/stimuli/large_consensus_a.png"
)

# Titles for the images
image_titles <- c(
  "minority (0)", 
  "dissensus (1)", 
  "majority (2)", 
  "consensus (3)"
)

# Create a list of plots
plots <- lapply(seq_along(image_paths), function(i) {
  create_image_plot(image_paths[i], image_titles[i])
})

# Arrange the plots in a grid
grid.arrange(grobs = plots, ncol = 2)
```

### Design

We manipulated convergence within participants, by varying the ratio of players choosing the same response as a focal player (i.e. the one that participants evaluate). The levels of convergence are: (i) consensus, where all three players pick the same option [coded value = 3]; (ii) majority, where either the third or second player picks the same option as the first player [coded value = 2]; (iii) dissensus, where all three players pick different options [coded value = 1]; (iv) minority, where the second and third player pick the same option, but one that is different from the first player's choice [coded value = 0]. In our analysis, we treat convergence as a continuous variable, assigning the coded values in squared parenthesis here.

### Materials

All participants saw all four conditions, with two stimuli per condition. Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).

## Results and discussion

To account for dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slope for participants.

As in the numerical setting, we found a positive effect of convergence on both accuracy ($\hat{b}_{\text{Accuracy}}$ = `r exp4_descriptives$model_accuracy$convergence$estimate` `r exp4_descriptives$model_accuracy$convergence$ci`, p = `r exp4_descriptives$model_accuracy$convergence$p.value`; on a scale from 0 to 100) and competence ($\hat{b}_{\text{Competence}}$ = `r exp4_descriptives$model_competence$convergence$estimate` `r exp4_descriptives$model_competence$convergence$ci`, p = `r exp4_descriptives$model_competence$convergence$p.value`; on a scale from 1 to 7).

In the ESM (section \@ref(exp4)), we show participants underestimate the effect of convergence on both accuracy and competence, but especially on accuracy.

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To do so, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence). We find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy ($\hat{b}$ = `r exp4_exploratory_analysis$interaction$estimate` `r exp4_exploratory_analysis$interaction$ci`, `r exp4_exploratory_analysis$interaction$p.value`; units in standard deviations).

# Experiment 5

```{r exp5}
# Analyze data of experiments and store results

# Experiment 5

# read data
exp5 <- read_csv("Experiment_5/data/cleaned.csv") %>% 
  # import experimental variable as factor and set levels
  mutate(
    # set levels for `convergence`
    convergence_categorical = recode_factor(convergence, 
                                            `0` = "opposing majority", 
                                            `1` = "divergence", 
                                            `2` = "majority", 
                                            `3` = "consensus",
                                            .default = NA_character_))

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the independence condition
exp5_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence | id), 
                              data = exp5 %>% filter(independence == "independent"))
exp5_model_h1_competence = lmer(competence ~ convergence + (1 + convergence | id), 
                              data = exp5 %>% filter(independence == "independent"))

# H2 
# run mixed model (random slope and intercept) with participants as random factor
exp5_model_accuracy = lmer(accuracy ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = exp5)
exp5_model_competence = lmer(competence ~ convergence + independence + 
                            independence*convergence + (1 + convergence | id), 
                       data = exp5)

# extract descriptives for inline reporting
exp5_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp5$id),
  gender = exp5 %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp5 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_h1_accuracy = text_ready(exp5_model_h1_accuracy), 
  model_h1_competence = text_ready(exp5_model_h1_competence),
  model_accuracy = text_ready(exp5_model_accuracy), 
  model_competence = text_ready(exp5_model_competence)
)

# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? 
exp5_explorative <- exp5 %>% 
  mutate(accuracy_scaled = scale(accuracy), 
         competence_scaled = scale(competence)) %>% 
  pivot_longer(
    c(accuracy_scaled, competence_scaled),
    names_to = "outcome", 
    values_to = "score")

# run mixed model (random slope and intercept) with participants as random factor
exp5_exploratory_analysis <- lmer(score ~ convergence + outcome + 
                               convergence*outcome + independence +
                               (1 + convergence | id), 
                             data = exp5_explorative) %>% 
  text_ready()
```

Experiment 5 is a conceptual replication of Experiment 3 in a categorical instead of a numerical case: are participants less likely to infer that more convergent estimates are more accurate, and the individuals who made them more competent, when the estimates are made by individuals with a conflict of interest pushing them to all provide a given answer, compared to when they are made by independent participants? The independence condition of Experiment 5 also serves as a replication of Experiment 4, leading to the following hypotheses:

***H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.***

***H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.***

***H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).***

***H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are biased (i.e. share a conflict of interest to pick a given answer).***

## Methods

### Participants

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. We ran two different power analyses, one for each outcome variable. We set the power threshold for both to 90%.

The power simulation for accuracy suggested that for 80 participants, we would have a power of at least 90% for the interaction effect. The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we would also detect an effect of convergence. Due to uncertainty about our assumptions and because resources were available for a larger sample, we recruited `r exp5_descriptives$n_subj` participants, in the UK and via Prolific (`r exp5_descriptives$gender$female$n` female, `r exp5_descriptives$gender$male$n`, 1 non-identified; $age_\text{mean}$: `r exp5_descriptives$age$mean`, $age_\text{sd}$: `r exp5_descriptives$age$sd`, $age_\text{median}$: `r exp5_descriptives$age$median`).

### Procedure

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: "We will show you three financial advisors who are giving recommendations on investment decisions. They can choose between three investment options. Their task is to recommend one. You will see several such situations. They are completely unrelated: it is different advisors evaluating different investments every time. At first you have no idea how competent the advisors are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be difficult while others are easy. Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are." To assess perceptions of accuracy, we asked: "What do you think is the probability of advisor 1 making the best investment recommendation?". Participants answered with a slider on a scale from 0 to 100. To assess perceptions of accuracy, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").

### Design

We manipulated convergence within participants, and conflict of interest between participants. In the conflict of interest condition, experts were introduced this way: "The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations." For the independence condition: "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations."

### Materials

We used the same stimuli as in Experiment 4. Identical to Experiment 4, participants saw all four convergence conditions, with two stimuli (i.e. expert predictions) per condition. Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).

## Results and discussion

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants.

We find evidence for all four hypotheses (see Fig. \@ref(fig:exp5-plot)). To test H1a and H1b, we use the same analyses as in Experiment 4, restricted on the independent condition, and replicate the results. We find a positive effect of convergence on both accuracy ($\hat{b}_{\text{Accuracy}}$ = `r exp5_descriptives$model_h1_accuracy$convergence$estimate` `r exp5_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp5_descriptives$model_h1_accuracy$convergence$p.value`) and competence ($\hat{b}_{\text{Competence}}$ = `r exp5_descriptives$model_h1_competence$convergence$estimate` `r exp5_descriptives$model_h1_competence$convergence$ci`, p = `r exp5_descriptives$model_h1_competence$convergence$p.value`).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \@ref(fig:exp5-plot)). In the independence condition, the effect of convergence on accuracy was more positive ($\hat{b}_{\text{interaction, Accuracy}}$ = `r exp5_descriptives$model_accuracy$interaction$estimate` `r exp5_descriptives$model_accuracy$interaction$ci`, p = `r exp5_descriptives$model_accuracy$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ `r exp5_descriptives$model_accuracy$convergence$estimate` `r exp5_descriptives$model_accuracy$convergence$ci`, p = `r exp5_descriptives$model_accuracy$convergence$p.value`). Likewise, the effect of convergence on competence was more positive ($\hat{b}_{\text{interaction, Competence}}$ = `r exp5_descriptives$model_competence$interaction$estimate` `r exp5_descriptives$model_competence$interaction$ci`, p = `r exp5_descriptives$model_competence$interaction$p.value`) than in the conflict of interest condition ($\hat{b}_{\text{baseline}}$ = `r exp5_descriptives$model_competence$convergence$estimate` `r exp5_descriptives$model_competence$convergence$ci`, p = `r exp5_descriptives$model_competence$convergence$p.value`).

(ref:exp5-plot) Interaction of convergence and informational dependency.

```{r exp5-plot, fig.cap="(ref:exp5-plot)"}
# plot for accuracy
plot_accuracy <- ggplot(exp5,
       aes(x = convergence_categorical, y = accuracy, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Accuracy", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# plot for competence
plot_competence <- ggplot(exp5,
       aes(x = convergence_categorical, y = competence, fill = independence)) +
  geom_half_violin(position = position_nudge(x = -.2),
                   adjust=2, alpha = .4,
                   side = "l") +
  stat_summary(fun = "mean", geom = "point", size = 2, shape = 21) +
  stat_summary(fun = "mean", geom = "line", size = 1, linetype = "dashed") +
  stat_summary(fun.data = "mean_se", geom = "errorbar", width = .2) +
  # Add nice labels
  labs(x = "Convergence", y = "Competence", fill = NULL) +
  scale_fill_manual(values = c("#E69F00", "#56B4E9")) +
  plot_theme + 
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

ggpubr::ggarrange(plot_accuracy, plot_competence, common.legend = TRUE)
```

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To do so, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy ($\hat{b}$ = `r exp5_exploratory_analysis$interaction$estimate` `r exp5_exploratory_analysis$interaction$ci`, `r exp5_exploratory_analysis$interaction$p.value`; units in standard deviations).

## Experiment 6

```{r exp6}
# Analyze data of experiments and store results

# Experiment 6

# read data
exp6 <- read_csv("Experiment_6/data/cleaned.csv") %>% 
  # import experimental variable as factor and set levels
  mutate(
    # set levels for `convergence`
    convergence_categorical = recode_factor(convergence, 
                                            `0` = "opposing majority", 
                                            `1` = "divergence", 
                                            `2` = "majority", 
                                            `3` = "consensus",
                                            .default = NA_character_))

# H1
# run mixed model (random intercept and slope for convergence) with participants as random factor
# use only those participants assigned to the three options condition
exp6_model_h1_accuracy = lmer(accuracy ~ convergence + (1 + convergence | id), 
                              data = exp6 %>% filter(number_options == "three"))
exp6_model_h1_competence = lmer(competence ~ convergence + (1 + convergence | id), 
                              data = exp6 %>% filter(number_options == "three"))

# H2 
# run mixed model (random slope and intercept) with participants as random factor

# For this model, we effect code `number_options`, so that we get a coefficient 
# for the main effect of convergence as well as for the interaction
exp6 <- exp6 %>% 
  mutate(number_options_effect_code = recode(number_options, "three" = -0.5, "ten" = +0.5))

exp6_model_accuracy = lmer(accuracy ~ convergence + number_options_effect_code + 
                            number_options_effect_code*convergence + (1 + convergence | id), 
                       data = exp6)
exp6_model_competence = lmer(competence ~ convergence + number_options_effect_code + 
                            number_options_effect_code*convergence + (1 + convergence | id), 
                       data = exp6)

# extract descriptives for inline reporting
exp6_descriptives <- list(
  # N subjects
  n_subj = n_distinct(exp6$id),
  gender = exp6 %>% group_by(gender) %>% summarize(n = n_distinct(id)) %>% split(.$gender),
  age = exp6 %>% summarize(across(age,list(mean = mean, median = median, sd = sd), 
                                  .names = "{.fn}")) %>% rounded_numbers(),
  # Model results
  model_h1_accuracy = text_ready(exp6_model_h1_accuracy), 
  model_h1_competence = text_ready(exp6_model_h1_competence),
  model_accuracy = text_ready(exp6_model_accuracy), 
  model_competence = text_ready(exp6_model_competence)
)

# Exploratory analysis: Is effect of convergence bigger on accuracy then on competence? 
exp6_explorative <- exp6 %>% 
  mutate(accuracy_scaled = scale(accuracy), 
         competence_scaled = scale(competence)) %>% 
  pivot_longer(
    c(accuracy_scaled, competence_scaled),
    names_to = "outcome", 
    values_to = "score")

# run mixed model (random slope and intercept) with participants as random factor
exp6_exploratory_analysis <- lmer(score ~ convergence + outcome + 
                               convergence*outcome + number_options +
                               (1 + convergence | id), 
                             data = exp6_explorative) %>% 
  text_ready()
```

Experiment 6 is a replication and extension of Experiment 4 in which we test the effect of the number of choice options (three and ten, instead of only three). Our simulations suggested that, at least for some underlying population competence distributions, consensus should be more indicative of competence when there are more choice options, compared to fewer (see ESM, section \@ref(simulations) and \@ref(exp6)).

First, considering only the three options condition, we ran a direct replication of experiment 4. Second, following the results from our model, we predict that

***H1: The effect of convergence on accuracy (H1a) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.***

***H2: The effect of convergence on competence (H1b) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.***

## Methods

### Participants

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. We used previous experiments and estimates of our models to inform our choice of parameter values. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90%. The power simulation for accuracy suggested that for 140 participants we would cross the power threshold of 90% for the interaction effect (power = 0.928). The simulation for competence suggested that with 300 participants, we would detect an interaction with a power of 87%. Due to budget constraints, we considered aiming for a sample of 300 participants as good enough, although slightly below our threshold. Due to two failed attention checks, our final sample consisted of `r exp6_descriptives$n_subj` subjects, recruited, as in all experiments, in the UK and via Prolific (`r exp6_descriptives$gender$female$n` female, `r exp6_descriptives$gender$male$n`, 1 non-identified; $age_\text{mean}$: `r exp6_descriptives$age$mean`, $age_\text{sd}$: `r exp6_descriptives$age$sd`, $age_\text{median}$: `r exp6_descriptives$age$median`).

### Procedure

We used the same procedure as in Experiment 4, with the addition of one condition described below.

### Design

The number of choice options was manipulated between participants. Participants were randomly assigned to either to see stimuli with three options (as in Experiment 4), or with ten options. Participants assigned to the ten options condition were divided into one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the three options condition, and another with increased range (see ESM, section \@ref(exp6)). We found no differences between the two sub-conditions and collapsed them into a single ten options condition.

### Materials

For the three options condition, we used the same stimuli as in Experiments 4 and 5. For the ten options condition, we created new sets of stimuli (see ESM, section \@ref(exp6)). Identical to Experiments 4 and 5, participants saw all four convergence conditions, with two stimuli per condition. Each participant therefore saw eight stimuli in total.

## Results and discussion

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants.

We replicate the results of experiment 4, but do not find evidence for an interaction between convergence and the number of choice options. To match the setting of experiment one, we reduced the sample to half of the participants, namely those who were assigned to the three options condition. On this reduced sample, we ran the exact same analyses as in experiment 4 and replicated the results. We find a positive effect of convergence on both accuracy ($\hat{b}_{\text{Accuracy}}$ = `r exp6_descriptives$model_h1_accuracy$convergence$estimate` `r exp6_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp6_descriptives$model_h1_accuracy$convergence$p.value`) and competence ($\hat{b}_{\text{Competence}}$ = `r exp6_descriptives$model_h1_competence$convergence$estimate` `r exp6_descriptives$model_h1_competence$convergence$ci`, p = `r exp6_descriptives$model_h1_competence$convergence$p.value`). This finding also holds across the entire sample, pooling three and ten choice option observations ($\hat{b}_{\text{Accuracy}}$ = `r exp6_descriptives$model_accuracy$convergence$estimate` `r exp6_descriptives$model_accuracy$convergence$ci`, `r exp6_descriptives$model_accuracy$convergence$p.value`; $\hat{b}_{\text{Competence}}$ = `r exp6_descriptives$model_competence$convergence$estimate` `r exp6_descriptives$model_competence$convergence$ci`, `r exp6_descriptives$model_competence$convergence$p.value`).

We do not find evidence of an interaction, i.e. evidence that the number of choice options changes the effect of convergence ($\hat{b}_{\text{interaction, Accuracy}}$ = `r exp6_descriptives$model_accuracy$interaction$estimate` `r exp6_descriptives$model_accuracy$interaction$ci`, `r exp6_descriptives$model_accuracy$interaction$p.value`; $\hat{b}_{\text{interaction, Competence}}$ = `r exp6_descriptives$model_competence$interaction$estimate` `r exp6_descriptives$model_competence$interaction$ci`, `r exp6_descriptives$model_competence$interaction$p.value`).

We tested whether the effect of convergence is larger on accuracy than on competence, in an exploratory, non-preregistered analysis. To this end, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for the number of choice options. In line with all previous experiments, find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy ($\hat{b}$ = `r exp6_exploratory_analysis$interaction$estimate` `r exp6_exploratory_analysis$interaction$ci`, `r exp6_exploratory_analysis$interaction$p.value`).

Experiment 6 replicates Experiments 3 and 4 (independence condition), but suggests that the number of options the informants choose from does not powerfully affect participants’ estimates of the informants’ accuracy or competence. In the ESM, section \@ref(exp6), we show that, compared to our model, participants underestimate the effect of the number of choice options.

# General discussion

```{r}
# Calculate total number of participants

# Define a list of datasets
datasets <- list(exp1, exp2, exp3, exp4, exp5, exp6)

# Initialize a variable to store the total sum
total_sum <- 0

# Loop through each dataset
for (i in seq_along(datasets)) {
  # Check if the dataset contains "ID" column, if not, check for "id"
  if ("ID" %in% colnames(datasets[[i]])) {
    id_col <- "ID"
  } else if ("id" %in% colnames(datasets[[i]])) {
    id_col <- "id"
  } else {
    stop("Dataset doesn't contain 'ID' or 'id' column.")
  }

  # Calculate the sum of IDs for the current dataset
  sum_of_ids <- n_distinct(datasets[[i]][[id_col]])

  # Add the sum to the total sum
  total_sum <- total_sum + sum_of_ids
}

# Print the total sum
#total_sum
```

Using both analytical arguments and simulations, we showed that, under a wide range of parameters, more convergent answers tend to be more accurate, and to have been made by more competent informants.

In two experiments (Experiment 1, and independence condition of Experiment 3), we find that participants presented with a set of more (rather than less) convergent numerical estimates find the estimates more accurate, and the individuals making the estimates more competent, thus drawing normatively justified inferences. Experiment 2 suggests that participants do not think that a discussion between the individuals makes their convergence less indicative of accuracy or their competence. By contrast, Experiment 3 reveals that, when the individuals making the estimates are systematically biased by a conflict of interest, participants put less weight on the convergence of their estimates.

Similar results are obtained in a categorical choice context, in which participants see the answers of individuals made within a limited set of options. Experiments 4, 5 (independence condition), and 6 show that, the more the answers converge, the more participants believe them to be accurate and the individuals who made to be competent, again drawing normatively justified inferences. Experiment 5 shows that these inferences are weakened when the convergence can be explained by a conflict of interest (as in Experiment 3). Experiment 6 fails to find an effect of the number of options.

We also observe that, in line with the simulations, participants draw stronger inferences from convergence to accuracy than to competence in five of the six experiments.

On the whole, participants thus appear to draw normatively justified inferences: (i) they infer that more convergent answers are more accurate; (ii) they infer that more convergent answers are more competent informants; (iii) inference (ii) tends to be stronger than inference (i); (iv) both inference (i) and (ii) are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other. The only exception appears to be that participants do not take into account the number of options in categorical choices scenarios.

# Conclusion

When people see that others agree with each other, they tend to believe that they are right. This inference has been evidenced in several experiments, both for numerical estimates [e.g. @mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009; @budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007], and for categorical choices [e.g., @morganEvolutionaryBasisHuman2012; for review, see @mercierMajorityRulesHow2019]. However, in these experiments, the participants arguably assumed a degree of competence among the individuals whose answers they saw. For instance, when children are confronted with several individuals who agree on how to name a novel object [e.g. @corriveauGoingFlowPreschoolers2009], they can assume that these (adult) individuals tend to know what the names of objects are. If a certain competence of the informants is assumed, then well-known results from the literature on judgment aggregation--the wisdom of crowds--show that the average opinion of a set of individuals is, in a wide range of circumstances, more likely to be accurate than that of a single individual [e.g. @larrickIntuitionsCombiningOpinions2006].

Here, we assume no prior knowledge of individual competence, asking the question: if we see a set of individuals, whose competence is unknown, converge on an answer, is it rational to infer that this answer is more likely to be correct, and that the individuals are likely to be competent? We show that the answer is yes on both counts--assuming there is no systematic bias among the individuals answering. An analytical argument and a series of simulations reveal that, for both the numerical choice context and the categorical choice context, the more individuals agree on an answer, the more likely the answer is to be correct, and the more likely the individuals are competent, with the former effect being stronger than the latter. Moreover, this is true for a wide range of assumed population distributions of competence.

In a series of experiments, we show that participants (UK) draw these inferences: when presented with more convergent answers, they tend to believe the answers are more likely to be correct, and that the individuals who made them are more likely to be competent. This is true for numerical estimates and for categorical answers. These beliefs are weakened when the individuals making the estimates are systematically biased by a common conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).

The results--both simulations and experiments--are a novel contribution to the wisdom of crowds literature. In this literature--in particular that relying on the Condorcet Jury Theorem--a degree of competence is assumed in the individuals providing some answers. From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate. Here we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances. We also show that participants, by and large, are able to draw rational inferences, inferring accuracy and competence from an observation of convergence, and doing so more when there are no alternative explanations for the observed convergence.

People might draw this inference in a variety of contexts, but the most prominent one might be science. Science is, arguably, the institution in which individuals end up converging the most in their opinions. For instance, scientists within many disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA. This represents an incredible degree of convergence. When people hear that scientists have measured the distance between the solar system and the center of the galaxy, if they assume that there is a broad agreement within the relevant experts, this should lead them to infer that this measure is accurate, and that the scientists who made it are competent. Experiments have already shown that increasing the degree of perceived consensus among scientists tends to increase acceptance of the consensual belief [@vanstekelenburgScientificConsensusCommunicationContested2022], but it hasn't been shown yet that the degree of consensus also affects the perceived competence of scientists.

```{r}
# additional refs
# [@vanstekelenburgScientificConsensusCommunicationContested2022; @deryuginaEffectInformationProvision2016; @dixonApplyingGatewayBelief2016; @kerrChangesPerceivedScientific2018; @lewandowskyPivotalRolePerceived2013; @vanderlindenHighlightingConsensusMedical2015a; @vanderlindenScientificConsensusClimate2015; @vanderlindenHowCommunicateScientific2014; but see @dixonImprovingClimateChange2017; @landrumExaminingImpactExpert2019]
```

In the case of science, the relationship between convergence and accuracy is broadly justified. However, at some points of history, there has been broad agreement on misbeliefs, such as when Christian theologians had calculated that the Earth was approximately six thousand years old. To the extent that people were aware of this broad agreement, and believed the theologians to have reached it independently of each other, this might have not only fostered acceptance of this estimate of the age of the Earth, but also a perception of the theologians as competent.

The current study has a number of limitations. If the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the experiments. Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019]. However, we do not know whether they would generalize to other cultures.

Future studies could overcome these limitations by replicating the present results in different cultures, using more ecologically valid stimuli. For instance, it would be interesting to test whether the inference described here, from convergence to competence, might be partly responsible for the fact that people tend to believe scientists to be competent [@colognaTrustScientistsTheir2024].

### Data availability

The extracted data used to produce our results are available on the OSF project page (<https://osf.io/6abqy/>).

### Code availability

The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (<https://osf.io/6abqy/>).

### Competing interest

The authors declare having no competing interests.

\FloatBarrier

# References

::: {#refs}
:::

\newpage

# (APPENDIX) ESM, section {.unnumbered}

```{r child = "appendix_analytical.Rmd"}
```

\clearpage

```{r child = "appendix_simulations.Rmd"}
```

\clearpage

```{r child = "appendix_exp1.Rmd"}
```

\clearpage

```{r child = "appendix_exp2.Rmd"}
```

\clearpage

```{r child = "appendix_exp3.Rmd"}
```

\clearpage

```{r child = "appendix_exp4.Rmd"}
```

\clearpage

```{r child = "appendix_exp5.Rmd"}
```

\clearpage

```{r child = "appendix_exp6.Rmd"}
```

### 2024-03-26 functions/own_functions.R

@@ -8,7 +8,7 @@ text_ready <- function(model_output) {
     filter(effect == "fixed") %>% 
     # report p.value according to apa standards
     mutate(p.value = case_when(p.value < 0.001 ~ "< .001",
-                               TRUE ~ sprintf("p = %.3f", p.value)
+                               TRUE ~ sprintf("%.3f", p.value)
     )
     ) %>% 
     # all other terms
@@ -59,6 +59,7 @@ plot_competence_distributions <- function(data) {
       ) +
       labs(title = paste0(competence,"\n", "alpha = ", alpha, ", beta = ", beta), 
            y = "Density", x = "Competence") +
+      scale_x_continuous(breaks = scales::pretty_breaks()) +
       plot_theme +
       rremove("ylab") +  # Remove y-axis label
       rremove("xlab")   # Remove x-axis label
@@ -81,6 +82,21 @@ plot_competence_distributions <- function(data) {
   
   
   return(competence_plot) 
+}
+
+# Function for splitting data along several variables (useful for inline reporting)
+# taken from here: https://www.tjmahr.com/lists-knitr-secret-weapon/
+super_split <- function(.data, ...) {
+  dots <- rlang::enquos(...)
+  for (var in seq_along(dots)) {
+    var_name <- rlang::as_name(dots[[var]])
+    .data <- purrr::map_depth(
+      .x = .data,
+      .depth = var - 1,
+      .f = function(xs) split(xs, xs[var_name])
+    )
+  }
+  .data
 }
 
 
@@ -191,6 +207,84 @@ plot_competence_vary_categorical <- function(data, variable = options, outcome =
   }
 } 
 
+# function used in participant vs. model comparisons
+plot_results <- function(data, outcome = "everything") {
+  
+  d <- data
+  
+  # make constellation a factor with the right levels
+  d$constellation <- fct_relevel(d$constellation, "minority(ies)", "dissensus", "majority", "consensus")
+  
+  # extract simulation info
+  simulation_info <- d %>% summarize(across(c(population, sample, options, total_iterations), mean)) 
+  
+  caption = paste0("Number of choice options: ", simulation_info$options,
+                   "; Sample size: ", simulation_info$sample,
+                   "; Iterations : ", simulation_info$total_iterations, 
+                   "; Population size per iteration: ", simulation_info$population)
+  
+  # make descriptive data
[…]

### 2024-04-16 Cogsci_version_raw.Rmd
revised submission Cogsci, main article not yet including polished analytical argument

@@
 
-(ref:simulations) **A** Shows the results of simulations in a categorical setting with 3 choice options.
-Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes.
-**B** Shows the results in a continuous setting.
-Regression lines represent the correlation between accuracy (left)/competence (right) and convergence (measured by squared distance to true mean and reversed such that greater convergence corresponds to being more right on the x-axis).
-**C** Shows the different population competence distributions we considered in our simulations.
+(ref:simulations) **A** Shows the different population competence distributions we considered in our simulations.
 In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1.
@@
 in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer.
+**B** Shows the results of simulations in a categorical setting with three choice options.
+Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes.
+**C** Shows the results in a continuous setting.
+Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the figure) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).
 
@@
 # common plot
-patchwork <-  plots_categorical / plots_continuous / (plot_competence_distributions(simulation_numeric) + plot_spacer())
+patchwork <- (plot_competence_distributions(simulation_numeric) + plot_spacer()) / 
+  plots_categorical /
+  plots_continuous
 

### 2024-04-16 README.qmd
revised submission Cogsci, main article not yet including polished analytical argument

@@
 - `appendix*` files contain the appendices that are automatically added `paper.Rmd` when rendering the file.
-- `Cogsci*` files are documents used to generate a shorter version of our paper submitted for the CogSci 2024 conference. 
+- `Cogsci*` files are documents used to generate the initial submission of a shorter version of our paper for the CogSci 2024 conference.
 - `Experiment*` folders contain all files related to the respective experiment.
@@
 - figures includes images used in the `paper.Rmd` file
-- `howwise_0131.pdf` is the final submission to the CogSci 2024 conference, based on the raw files but edited in [overleaf](https://www.overleaf.com/). 
+- `howwise_0131_initial_submission.pdf` is the initial, and `howwise_0131.pdf` the final submission (after review) to the CogSci 2024 conference, based on the raw files but edited in [overleaf](https://www.overleaf.com/).
 - `output` folder lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.

### 2024-04-16 paper.Rmd
revised submission Cogsci, main article not yet including polished analytical argument

@@
 Evidence suggests that people tend to make this inference.
-However, standard wisdom of crowds approaches only provide limited normative grounds .
+However, standard wisdom of crowds approaches only provide limited normative grounds.
 Using simulations and analytical arguments, we argue that when individuals make independent and unbiased estimates, under a wide range of parameters, individuals whose answers converge with each other tend to have more accurate answers and to be more competent.
@@
 
-There are two ways for answers to converge: when the answers are continuous, they can have a smaller range, or a lower variance; when the answers are categorical (i.e.
-there are only a few options), more answers can be the same.
-These two cases have been studied with different paradigms.
+The literature on the wisdom of crowds has treated separately situations with continuous answers, such as the weight of an ox in Galton’s famous observation [@galtonVoxPopuli1907], and with categorical answers, as when voters have to choose between two options, in the standard Condorcet Jury Theorem [@decondorcetEssaiApplicationAnalyse2014].
+The continuous and the categorical case are typically modeled with different tools, and they have usually been studied in different empirical literatures (see below).
+Given that they both represent common ways for answers to converge more or less (e.g.
+when people give numerical estimates vs.
+vote on one of a limited number of options), we treat them both here, with different simulations and experiments.
 
@@
 There is clear evidence that, both in the continuous and the categorical case, participants infer that convergent opinions are more likely to be correct.
-There is tentative evidence that participants also infer that individuals who hold more convergent opinions are more likely to be competent.
-Would that second inference be justified?
-More precisely, if we consider individuals who have convergent opinions, by contrast with individuals whose opinions are less convergent, or whose opinion diverges from that of a majority, are we justified in thinking that the individuals with more convergent opinions are more likely to be right not only for this specific question, but that they would also be right for other (related) questions?
-We first provide a general analytic answer, before presenting simulations for the continuous, and then the categorical case.
+There is also tentative evidence that participants infer that individuals holding more convergent opinions are more likely to be competent.
+Is that second inference justified?
+More precisely, is it normatively warranted to claim that individuals holding convergent opinions are more likely to be right not only for this specific question, but also be right for other (related) questions?
+We first provide a general analytic answer, before presenting simulations for the continuous and categorical cases.
 
@@
 
-In the continuous case, we imagine a population of informants.
-The answers of each informant are drawn from a normally distributed random variable, centered on the correct answer (i.e., informants have no systematic bias).
-The variance of this distribution represents the individuals’ competence: the larger the variance, the lower the competence.
-If we observe a sample of this population, the sample mean (i.e.
-the mean of the answers of all the informants) is then the best estimate of the correct answer, and the sample variance (i.e.
-mean squared distance between the answers and the sample mean) is the best estimate of the population’s average competence (best is understood here as the least volatile estimator).
+In the continuous case, let us imagine a population of informants.
+Individual opinions are drawn from a normal distribution, centered on the correct answer [i.e., informants have no systematic bias, as for instance in Galton’s classic demonstration, @galtonVoxPopuli1907].
+The variance of this distribution represents the individuals’ competence: the larger the variance, the lower the competence.  We observe the individual answers.
+In this setting, it is well known in statistics (see also Electronic Supplementary Materials, ESM) that the sample mean (i.e. the mean of the answers of all the informants) is the best estimator of the correct answer, and the sample variance (i.e. the mean squared distance between the answers and the sample mean) is the best estimator of the population’s average competence (best understood here as the least volatile estimator).
 This means that a population of informants whose answers converge more (lower sample variance) is, on average, more competent (informants tend to answer with a lower variance).
 
-We can extend this argument from populations to individuals: Consider that competence varies within a population, such that each informant's answer is drawn from their own individual distribution--always centered on the correct answer, but with different variances.
+We can extend this argument from populations to individuals: Consider that competence varies within a population, such that each informant's answer is drawn from their own distribution -- always centered on the correct answer, but with different variances.
 In this case, the distance between each informant's answer and the sample mean provides the best estimate of that informant's competence (i.e.
@@
 
-In the categorical case, we assume that informants answer a question with several categorical answers.
-We define the competence of an informant as the probability of choosing the correct answer.
-Due to the law of large numbers, the relative size of the majority--the share of informants who choose the most chosen answer--is, when the population is large enough, a good approximation of the average competence of the population (e.g.
-if the average competence is .66, and there are enough informants, then approximately 66% of informants will select the right answer).
-This relationship still exists, but is more noisy for small populations.
-This means that, at the population level, the more the answers converge, the more the population is, on average, competent (and the larger the population, the stronger the inference).
-
-To make inferences on individual informants' competence on the basis of whether or not their answers converge with those of the majority in the sample, we must assume some distribution of competence in the population (assuming, as above, that the true average population competence is equal to the observed majority size in the sample).
-Using Bayes theorem, we can then show (see Electronic Supplementary Materials, ESM, section \@ref(analytical)) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig.
+In the categorical case, we define the competence of an informant as the probability of choosing the correct answer.
+The law of large numbers implies that the relative size of the majority – the share of informants who choose the most chosen answer – is, when the population is large enough, a good approximation of the average competence of the population (e.g. if the average competence is .66, and there are enough informants, then approximately 66% of informants will select the right answer).
+For smaller populations, the relationship holds, with some degree of noise.
+In other words, the more the answers converge, the more the population can be inferred to be competent (and the larger the population, the more reliable the inference).
+
+Now, can we infer that informants belonging to the majority are more competent?
+To do so, we must assume some distribution of competence in the population.
+Using Bayes theorem, we can then show (see ESM, section \@ref(analytical)) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig.
 \@ref(fig:figure-Benoit-2)).
-Additionally, the more competence varies in the population, the more the degree of convergence is informative of an informant's competence.
-It makes sense: if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence.
-More generally, the fact that the focal individual is right (their 'accuracy') can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa).
-Figure \@ref(fig:figure-Benoit-2) provides an example of a specific categorical choice, for two different distributions of competence.
+We also find that , the more competence varies in the population, the more the degree of convergence is indicative of an informant’s competence.
+This reflects the fact that, if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence.
+More generally, the fact that the focal individual is right (their ‘accuracy’) can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa) (see ESM).
+Figure \@ref(fig:figure-Benoit-2) provides an example with a specific categorical choice scenario, under two different population distributions of competence.
 
@@
 degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence of informants, we conducted simulations.
+In all simulations, we assume agents to be unbiased and independent in their answers, but varying in their competence.
 
@@
 
-Groups of agents, with each agent’s competence varying, provide numerical answers.
+Groups of agents, with each agents' competence varying, provide numerical answers.
 We measure how accurate these answers are, and how much they converge (i.e.
@@
 (ref:simulations) **A** Shows the different population competence distributions we considered in our simulations.
-In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1.
+In the continuous simulations, competence values of 0 correspond to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1.
 In the categorical simulations, a competence value of 0 corresponds to chance (e.g.
@@
 **C** Shows the results in a continuous setting.
-Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the scale) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).
+Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the figure) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).
 
@@
 For most statistical models, we relied on the `lme4` package and its `lmer()` function.
+Unless mentioned otherwise, we report unstandardized model coefficients that can be interpreted in units of the scales we use for our dependent variables.
 
@@
 Participants answered with a slider on a scale from 0 to 100.
-To assess perceptions of accuracy, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
+To assess perceptions of competence, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
 
@@
 
-On the whole, participants thus appear to draw normatively justified inferences: (i) they infer that more convergent answers are more accurate; (ii) they infer that more convergent answers are more competent informants; (iii) inference (ii) tends to be stronger than inference (i); (iv) both inference (i) and (ii) are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other.
+On the whole, participants thus appear to draw normatively justified inferences: (i) they infer that more convergent answers are more accurate; (ii) they infer that more convergent answers are coming from more competent informants; (iii) inference (i) tends to be stronger than inference (ii); (iv) both inference (i) and (ii) are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other.
 The only exception appears to be that participants do not take into account the number of options in categorical choices scenarios.
@@
 From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate.
-Here we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances.
-We also show that participants, by and large, are able to draw rational inferences, inferring accuracy and competence from an observation of convergence, and doing so more when there are no alternative explanations for the observed convergence.
-
-People might draw this inference in a variety of contexts, but the most prominent one might be science.
+Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be sufficient in determining who tends to be an expert.
+This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert [@nguyenCognitiveIslandsRunaway2020; @hahnHowGoodYour2018].
+
+People might draw these inferences from convergence in a variety of contexts, but the most prominent one might be science.
 Science is, arguably, the institution in which individuals end up converging the most in their opinions.
@@
 The current study has a number of limitations.
+In our simulations, we assume agents to be independent.
+Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency [@ladhaCondorcetJuryTheorem1992], more robust simulations would show that–while still assuming no systematic bias–our results hold even when agents influence each others' answers.
 If the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the experiments.

### 2024-04-19 paper.Rmd
First submission Cognition

@@
 ---
-title             : "How wise is the crowd? \n Can we infer people are accurate and competent merely because they agree with each other?"
+title             : "How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?"
 shorttitle        : "How wise is the crowd?"
@@
 
-## Do people infer that individuals whose answers converge tend to be right, and to be competent?
+# Do people infer that individuals whose answers converge tend to be right, and to be competent?
 
@@
 In the continuous case, the most relevant evidence comes from the literature on 'advice-taking.' In these experiments, participants are called 'judges' who need to make numerical estimates--sometimes on factual knowledge, e.g.
-'What year was the Suez Canal opened first?
-[@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g.
+'What year was the Suez Canal opened first?' [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g.
 'How many animals were on the screen you saw briefly?' [@mollemanStrategiesIntegratingDisparate2020].
@@
 Other studies manipulated the degree of convergence within a group of advisors.
-These studies present participants with a set of estimates, some of which are close to each other , while others are outliers [@harriesCombiningAdviceWeight2004; @yanivWeightingTrimmingHeuristics1997, study 3 & 4].
+These studies present participants with a set of estimates, some of which are close to each other, while others are outliers [@harriesCombiningAdviceWeight2004; @yanivWeightingTrimmingHeuristics1997, study 3 & 4].
 These studies find that participants discount outliers when aggregating estimates.
@@
 
-## Is inferring competence from convergence justified?
+# Is inferring competence from convergence justified?
 
@@
 
-### Analytical argument
+## Analytical argument
 
@@
 \@ref(fig:figure-Benoit-2)).
-We also find that , the more competence varies in the population, the more the degree of convergence is indicative of an informant’s competence.
-This reflects the fact that, if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence.
-More generally, the fact that the focal individual is right (their ‘accuracy’) can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa) (see ESM).
+We also find that, the more competence varies in the population, the more the degree of convergence is indicative of an informant’s competence.
+For example, if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence.
+More generally, the fact that the focal individual is right (their ‘accuracy’) can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa).
 Figure \@ref(fig:figure-Benoit-2) provides an example with a specific categorical choice scenario, under two different population distributions of competence.
 
-(ref:figure-Benoit-2) Average estimated competence and accuracy for an individual, depending on how many participants agree with him, in a categorical choice scenario.
+(ref:figure-Benoit-2) Average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario.
 Here, we assume that 20 individuals answer a 5-choice question.
-For values higher than four, we assume that the individual is part of the majority.
-Below four, we assume that the focus individual is part of a minority, and that the majority size is four.
+For values higher than four, we assume that the focal individual is part of the majority.
+Below four, we assume that the focal individual is part of a minority, and that the majority size is four.
 Two situations are represented: one in which there is low variance in the distribution of competence in the population (orange line in the insert), and one in which there is high variance in this distribution (blue line in the insert).
 
-```{r figure-Benoit-2, echo=FALSE, fig.cap="(ref:figure-Benoit-2)", out.width= "50%", fig.align="left"}
+```{r figure-Benoit-2, echo=FALSE, fig.cap="(ref:figure-Benoit-2)", out.width= "75%", fig.align="left"}
 knitr::include_graphics("figures/figure_Benoit-2.png")
@@
 
-### Simulations
+## Simulations
 
@@
 In all simulations, we assume agents to be unbiased and independent in their answers, but varying in their competence.
-
-#### Continuous case
+All code and data regarding the simulations can be found on Open Science Framework [project page](https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b) (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).
+
+### Continuous case
 
@@
 
-#### Categorical case
+### Categorical case
 
@@
 
-## Overview of the experiments
+# Overview of the experiments
 
@@
 All experiments were preregistered.
-Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/) (<https://osf.io/6abqy/>).
+Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b) (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).
 All analyses were conducted in R (version 4.2.2) using R Studio.
@@
 Because we could build upon data from the first experiment, we ran a power analysis by simulation.
-The simulation code is available on OSF, and the procedure is described in the preregistration document.
+The simulation code is available on the OSF, and the procedure is described in the preregistration document.
 The simulation suggested that 100 participants provide a significant interaction term between 95% and 97% of the time, given an alpha threshold for significance of 0.05.
@@
 This is true for numerical estimates and for categorical answers.
-These beliefs are weakened when the individuals making the estimates are systematically biased by a common conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).
+These beliefs are weakened when the individuals making the estimates are systematically biased by a shared conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).
 
@@
 From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate.
-Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be sufficient in determining who tends to be an expert.
+Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be a valid cue when determining who tends to be an expert.
 This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert [@nguyenCognitiveIslandsRunaway2020; @hahnHowGoodYour2018].
@@
 Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency [@ladhaCondorcetJuryTheorem1992], more robust simulations would show that–while still assuming no systematic bias–our results hold even when agents influence each others' answers.
-If the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the experiments.
-Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized.
+Regarding our experiments, if the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the results.
+Although the main results replicate well across our experiments, and we can thus be reasonably certain of their robustness, it's not clear how much they can be generalized.
 Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019].
@@
 
-The extracted data used to produce our results are available on the OSF project page (<https://osf.io/6abqy/>).
+Data for all experiments and the simulations is available on the OSF project page (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).
 
@@
 
-The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (<https://osf.io/6abqy/>).
+The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).
 
@@
 \clearpage
+\newpage
 

### 2024-09-11 appendix_exp1.Rmd
before final edits of revisions round 1

@@
 ```
+
+## Minor deviation from preregistration
+
+```{r}
+# T-test
+t_test_accuracy <- t.test(confidence ~ convergence, data = exp1, paired = TRUE) %>% clean_t_test()
+t_test_competence <- t.test(competence ~ convergence, data = exp1, paired = TRUE) %>% clean_t_test()
+```
+
+In our preregistration, we stated we would run paired t-tests and only conduct mixed models as a robustness check.
+However, it became clear to us later that mixed models are more appropriate in light of our data structure.
+In the main article, we therefore only report the results of the mixed models.
+Note, however, that using paired t-tests, we confirm our results (Accuracy: `r t_test_accuracy$estimate`, t-statistic = `r t_test_accuracy$statistic`, p `r t_test_accuracy$p.value`; Competence: `r t_test_competence$estimate`, t-statistic = `r t_test_competence$statistic`, p `r t_test_competence$p.value`)
+
+## Sensitivity simulation
+
+We had based our power calculations on a paired t-test (using the widely used tool G*Power), which does not correspond to the mixed-models we eventually ran.
+To address this point, here, we seek to identify the minimum effect size that our analysis with mixed models could have detected, given a 90% power, an N of 200 participants, and other parameters derived from the estimates observed in the sample of Experiment 1 (see \@ref(tab:exp1-regression-table)).
+By contrast to a standard power simulation (which varies sample size), a sensitivity simulation varies the effect size, while holding the sample size and desired power constant.
+
+```{r exp1-regression-table}
+modelsummary::modelsummary(list("Accuracy" = exp1_model_accuracy, 
+                                "Competence" = exp1_model_competence), 
+                           title = "Model Results Experiment 1")
+```
+
+```{r}
+# Sensitivity analysis Experiment 1
+
+library(tidyverse)
+library(purrr)
+library(broom)
+library(lmerTest)
+
+set.seed(1000)
+
+# Generate data
+create_data <- function(n_subj, n_div, n_conv, beta_0, beta_1, 
+                        tau_0, tau_1, rho, sigma) {
+
+  items <- data.frame(
+    item_id = seq_len(n_div + n_conv),
+    convergence = rep(c("div", "conv"), c(n_div, n_conv)),
+    # make the same as a numeric variable to generate competence
+    convergence_num = rep(c(-0.5, 0.5), c(n_div, n_conv)))
+
+  # variance-covariance matrix
+  cov_mx  <- matrix(
+    c(tau_0^2,             rho * tau_0 * tau_1,
+      rho * tau_0 * tau_1, tau_1^2            ),
+    nrow = 2, byrow = TRUE)
+
+  subjects <- data.frame(subj_id = seq_len(n_subj),
+                         MASS::mvrnorm(n = n_subj,
+                                       mu = c(T_0s = 0, T_1s = 0),
+                                       Sigma = cov_mx))
+
+  crossing(subjects, items)  %>%
+    mutate(
+           # turn convergence into factor variable
+           # for later analysis. Make sure the levels correspond to the 
+           # numeric variables we generate competence with
+           convergence = as.factor(convergence),
+           convergence = fct_relevel(convergence, "div", "conv"),
+           # residual error
+           e_si = rnorm(nrow(.), mean = 0, sd = sigma),
+           # generate competence
+           competence = beta_0 + T_0s + (beta_1 + T_1s) * convergence_num + e_si) %>%
+    select(subj_id, item_id, convergence, competence, convergence_num)
+}
+
+# test
+# data <- create_data(
+#   n_subj = 100,
+#   n_div  =  2,   # number of low convergence stimuli
+#   n_conv =  2,   # number of high convergence stimuli
+#   beta_0 = 4,
+#   beta_1 = 0.5,  # effect of convergence
+#   tau_0      = 0.7,   # by-subject random intercept sd
+#   tau_1      =  0.9,   # by-subject random slope sd
+#   rho        = 0.1,   # correlation between intercept and slope by-subject
+#   sigma      = 0.99) # residual (standard deviation)
+
+# Our estimation function
+est_model <- function(n_subj, n_div, n_conv, beta_0, beta_1,
+                      tau_0, tau_1, rho, sigma) {
+
+  data_sim <- create_data(n_subj, n_div, n_conv, beta_0, beta_1,
+                          tau_0, tau_1, rho, sigma)
+  model_sim <- lmer(competence ~ convergence + (1 + convergence | subj_id),
+                    data_sim)
+
+  tidied <- broom.mixed::tidy(model_sim)
+  sig <- tidied$p.value[2] < .05
+
+  return(sig)
+}
+
+# test
+# model_result <- est_model(
+#   n_subj = 100,
+#   n_div  =  2,   # number of low convergence stimuli
+#   n_conv =  2,   # number of high convergence stimuli
+#   beta_0 = 4,    # intercept (average of divergent condition)
+#   beta_1 = 0.5,  # effect of convergence
+#   tau_0      = 0.7,   # by-subject random intercept sd
+#   tau_1      =  0.9,   # by-subject random slope sd
+#   rho        = 0.1,   # correlation between intercept and slope by-subject
+#   sigma      = 0.99) # residual (standard deviation)
+
+# Iteration function
+iterate <- function(n_subj, n_div, n_conv, beta_0, beta_1,
+                    tau_0, tau_1, rho, sigma, iterations) 
+{ # number of iterations
+
+  results <-  1:iterations %>%
+    map_dbl(function(x) {
+      # To keep track of progress
+      if (x %% 100 == 0) {print(paste("iteration number ", x))}
+
+      # Run our model and return the result
+      return(est_model(n_subj, n_div, n_conv, beta_0, beta_1,
+                       tau_0, tau_1, rho, sigma))
+    })
+
+  # We want to know statistical power, 
+  # i.e., the proportion of significant results
+  return(mean(results))
+}
+
+# test
+# power <- iterate(
+#   n_subj = 100,
+#   n_div  =  2,   # number of low convergence stimuli
+#   n_conv =  2,   # number of high convergence stimuli
+#   beta_0 = 4, 
+#   beta_1 = 0.5,  
+#   tau_0      = 0.7,   # by-subject random intercept sd
+#   tau_1      =  1,   # by-subject random slope sd
+#   rho        = 0.1,   # correlation between intercept and slope by-subject
+#   sigma      = 1,  # residual (standard deviation)
+#   iterations = 500)
+
+sensitivity_analysis <- function(file_name, effect_sizes, ...) {
+
+  # only run analysis if a file with that name does not yet exists
+  if (!file.exists(paste0("Experiment_1/data/", file_name))) {
+
+    # do the `iterate()` function for each effect size and store the results
+    # in a common data frame
+    power <- purrr::map_df(effect_sizes, 
+              function(beta_1){
+                # this is essentially a for loop - 
+                # do the following for each 
+                # element of effect_sizes
+
+                # To keep track of progress
+                print(paste("tested effect size = ", beta_1))
+
+                # run power calculation
+                power <- iterate(beta_1 = beta_1, ...)
+
+                # make data frame
+                result <- tibble(effect_size = beta_1, 
+                                 power = power)
+
+                return(result)
+
+              })
+
+    # write out file
+    write_csv(power, paste0("Experiment_1/data/", file_name))
+  }
+}
+
+# test (note that this requires building the parameter lists below before)
+# effect_sizes <- c(1, 1.2)
+# test <- purrr::map_df(effect_sizes, 
+#               function(beta_1){
+#                 # this is essentially a for loop - 
+#                 # do the following for each 
+#                 # element of effect_sizes
+#                 
+#                 # To keep track of progress
+#                 print(paste("tested effect size = ", beta_1))
+#                 
+#                 # run power calculation
+#                 power <- do.call(iterate, c(parameters_accuracy, 
+#                                              beta_1 = beta_1))
+#                 
+#                 # make data frame
+#                 result <- tibble(effect_size = beta_1, 
+#                                  power = power)
+#                 
+#                 return(result)
+#                 
+#               })
+```
+
+```{r}
+# Define parameters
+
+# Extract parameters from model estimates of Experiment 1
+estimates_accuracy <- exp1_model_accuracy %>% 
+  tidy() %>% 
+  split(.$term)
+
+estimates_competence <- exp1_model_competence %>% 
+  tidy() %>% 
+  split(.$term)
+
+# Set parameter lists
+parameters_accuracy <- list(
+  n_subj = 200,
+  n_div  =  2,   # number of low convergence stimuli
+  n_conv =  2,   # number of high convergence stimuli
+  beta_0 = estimates_accuracy$'(Intercept)'$estimate, # intercept (mean of divergent)
+  #beta_1 = estimates_accuracy$convergenceconvergent$estimate, # mean difference convergence vs divergence
+  tau_0      = estimates_accuracy$'sd__(Intercept)'$estimate,   # by-subject random intercept sd
+  tau_1      =  estimates_accuracy$sd__convergenceconvergent$estimate,   # by-subject random slope sd
+  rho        = estimates_accuracy$'cor__(Intercept).convergenceconvergent'$estimate,   # correlation between intercept and slope by-subject
+  sigma      = estimates_accuracy$sd__Observation$estimate,  # residual (standard deviation)
+  iterations = 1000)
+
+# Set parameter lists
+parameters_competence <- list(
+  n_subj = 200,
+  n_div  =  2,   # number of low convergence stimuli
+  n_conv =  2,   # number of high convergence stimuli
+  beta_0 = estimates_competence$'(Intercept)'$estimate, # intercept (mean of divergent)
+  # beta_1 = estimates_competence$convergenceconvergent$estimate, # mean difference convergence vs divergence
+  tau_0      = estimates_competence$'sd__(Intercept)'$estimate,   # by-subject random intercept sd
+  tau_1      =  estimates_competence$sd__convergenceconvergent$estimate,   # by-subject random slope sd
+  rho        = estimates_competence$'cor__(Intercept).convergenceconvergent'$estimate,   # correlation between intercept and slope by-subject
+  sigma      = estimates_competence$sd__Observation$estimate,  # residual (standard deviation)
+  iterations = 1000)
+```
+
+```{r}
+# Run analysis
+
+# Accuracy
+do.call(sensitivity_analysis, c(parameters_accuracy, 
+                                        list(file_name = "sensitivity_accuracy.csv", 
+                                             effect_sizes = c(0.1, 0.2, 0.4, 0.6, 0.8, 1, estimates_accuracy$convergenceconvergent$estimate)
+                                             )
+                                        )
+                )
+
+sensitivity_accuracy <- read_csv("Experiment_1/data/sensitivity_accuracy.csv")
+
+# Competence
+do.call(sensitivity_analysis, c(parameters_competence, 
+                                        list(file_name = "sensitivity_competence.csv", 
+                                             effect_sizes = c(0.1, 0.2, 0.4, 0.6, 0.8, 1, estimates_competence$convergenceconvergent$estimate)
+                                             )
+                                        )
+                )
+
+sensitivity_competence <- read_csv("Experiment_1/data/sensitivity_competence.csv")
+
+```
+(ref:sensitivity) Results of the sensitivity analysis.
+
+```{r sensitivity, fig.cap="(ref:sensitivity)"}
+# Plot
+
+# combine data
+plot_data <- bind_rows(sensitivity_accuracy %>% 
+                         mutate(outcome = "Accuracy"), 
+                       sensitivity_competence %>% 
+                         mutate(outcome = "Competence"))
+
+# plot results
+plot_power <- function(data) {
+
+  ggplot(data, 
+       aes(x = effect_size, y = power)) +
+  geom_point(color = "red", size = 1) +
+  geom_line(size = 1, alpha = 0.5) + 
+  # add a horizontal line at 90%, our power_threshold
+  geom_hline(aes(yintercept = .9), linetype = 'dashed') + 
+  # display iterations
+  annotate("text", x = 1, y = 0.25, label = "iterations per \n effect size = 1'000")+
+  # Prettify!
+  theme_minimal() + 
+    scale_colour_viridis_d(option = "plasma") + 
+  scale_y_continuous(labels = scales::percent, breaks = ) + 
+  labs(x = 'Effect size', y = 'Power', 
+       title = "Sensitivity Analysis",
+       subtitle = "(Finding the minimum detectable effect)") +
+    facet_wrap(~outcome)
+}
+
+plot_power(plot_data)
+```
+
+The results of this can be seen in Fig.
+\@ref(fig:sensitivity).
+In our sample, we found a statistically significant effect size of `r estimates_accuracy$convergenceconvergent$estimate` for accuracy and an effect size of `r estimates_competence$convergenceconvergent$estimate` for competence (both on their original scales from 1 to 7).
+The simulations show that, given our parameter assumptions (based on the observations in our sample) and a sample size of N = 200, we would have detected an effect as small as 0.4 with a power of at least 90% (in fact nearly 100%) for both accuracy and competence.

### 2024-09-11 paper.Rmd
before final edits of revisions round 1

@@
 
-We first review experiments testing whether people infer that individuals whose answers converge are more likely to be right, and to be competent.
-We then argue that little attention has been paid to the normative models that might explain when and why an inference from convergence to competence is warranted.
-We develop such models, both analytically and with simulations.
-Finally, we present a series of experiments in which participants evaluate the accuracy and competence of individuals as a function of how much their answers converge on a given answer, in the absence of any priors about these individuals’ competence, or what the correct answer is.
+We begin by reviewing existing studies showing that people infer that competent informants who converge in their opinions are likely to be accurate.
+The wisdom of crowds literature provides normative grounds for this inference.
+We then argue that both the experimental and theoretical literature have paid little attention to extending this inference to cases in which there is no information about the informants’ competence, and to inferences about the competence of the informants.
+We first develop normative models, both analytically and with simulations, to show that inferences from convergence to accuracy and to competence are warranted under a wide range of parameters.
+Second, we present a series of experiments in which participants evaluate both the accuracy and competence of informants as a function of how much their answers converge on a given problem, in the absence of any priors about these individuals' competence, or what the correct answer is.
 
@@
 
-In the continuous case, the most relevant evidence comes from the literature on 'advice-taking.' In these experiments, participants are called 'judges' who need to make numerical estimates--sometimes on factual knowledge, e.g.
+In the continuous case, the most relevant evidence comes from the literature on 'advice-taking' [for review, see, @kammerSystematicReviewEmpirical2023].
+In these experiments, participants are called 'judges' who need to make numerical estimates--sometimes on factual knowledge, e.g.
 'What year was the Suez Canal opened first?' [@yanivReceivingOtherPeople2004], sometimes on knowledge controlled by the experimenters, e.g.
@@
 In categorical choice contexts, there is ample and long-standing [e.g.
-@crutchfieldConformityCharacter1955] evidence from social psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms [e.g., @morganEvolutionaryBasisHuman2012; for review, see @mercierMajorityRulesHow2019].
+@crutchfieldConformityCharacter1955] evidence from experimental psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms [e.g., @morganEvolutionaryBasisHuman2012; for review, see @mercierMajorityRulesHow2019].
 This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role [e.g.
@@
 
-If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the individuals' competence.
+If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the informants' competence.
 One potential exception is a study with preschoolers in which the children were more likely to believe the opinion of an informant who had previously been the member of a majority over that of an informant who had dissented from the majority [@corriveauGoingFlowPreschoolers2009].
@@
 
-# Is inferring competence from convergence justified?
-
-There is clear evidence that, both in the continuous and the categorical case, participants infer that convergent opinions are more likely to be correct.
-There is also tentative evidence that participants infer that individuals holding more convergent opinions are more likely to be competent.
-Is that second inference justified?
-More precisely, is it normatively warranted to claim that individuals holding convergent opinions are more likely to be right not only for this specific question, but also be right for other (related) questions?
-We first provide a general analytic answer, before presenting simulations for the continuous and categorical cases.
+# What inferences from convergence should we expect people to draw?
+
+Should we expect that people be able to infer that more convergent answers likely indicate not only more accurate answers, but also that those who gave the answers were competent?
+In order to make the best of communicated information, humans have to be able to evaluate it, so as to discard inaccurate or harmful information, while accepting accurate and beneficial information [@smithAnimalSignals2003].
+It has been argued that a suite of cognitive mechanisms–mechanisms of epistemic vigilance–evolved to serve this function [@sperberEpistemicVigilance2010; @mercierNotBornYesterday2020].
+Since the opinion of more than one individual is often available to us, there should be mechanisms of epistemic vigilance dedicated to processing such situations.
+It would be these mechanisms that lead us to put more weight on an opinion that is shared by a larger majority (in relative or absolute terms), and, in some cases at least, to discount majority opinion when the opinions haven’t been formed independently of each other [@mercierUtilizingSimpleCues2019].
+Evidence suggests that these mechanisms rely on heuristics which become more refined with age [@morganDevelopmentAdaptiveConformity2015], and which are far from perfect [in particular, they ignore many cases of informational dependencies, see, e.g.
+@yousifIllusionConsensusFailure2019; @mercierUtilizingSimpleCues2019].
+As mentioned above, in the experiments evaluating how people process convergent information, the participants had grounds to believe that the information came from competent informants.
+However, the same mechanisms could lead people to perform the same inference when they do not have such information (especially since, as is shown presently, such an inference is warranted).
+
+Regarding the inference from convergence to competence, other cognitive mechanisms allow us to infer how competent people are, based on a variety of cues, from visual access [@pillowEarlyUnderstandingPerception1989], to the time it takes to answer a question [@richardsonThinkingTakesTime2022].
+One of the most basic of these mechanisms infers from the fact that someone was right, that they possess some underlying competence [e.g.
+@koenigTrustTestimonyChildren2004].
+As a result, if participants infer that convergent opinions are more likely to be accurate, they should also infer that the informants who provided the opinions are competent.
+Before testing whether participants infer accuracy and competence from convergence, we show that these inferences are normatively warranted, both in the categorical and in the continuous case.
 
@@
 
-(ref:figure-Benoit-2) Average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario.
+(ref:figure-Benoit-2) Results of the analytic argument.
+The figure shows the average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario.
 Here, we assume that 20 individuals answer a 5-choice question.
@@
 
-A population of around 990000 agents (varying slightly with as a function of group sizes) with different competence levels is generated.
+A population of around 990000 agents (varying slightly as a function of group sizes) with different competence levels is generated.
 An answer is drawn for each agent, based on their respective competence distribution.
@@
 Competence values range from chance level (p = 1/m) to always selecting the correct option (p = 1).
-Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see see Fig.
+Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see Fig.
 \@ref(fig:simulations) A).
@@
 Within these groups, we first calculate the share of individuals voting for each answer, allowing us to measure convergence.
-For each level of convergence occurring within a group, we then compute (i) the average accuracy, (ii) the average competence of agents.
-Across all groups, we then compute the averages of these within-group averages, for each level of convergence.
-
-We repeat this procedure varying population competence distributions, and the size of informant groups, holding the number of choice options constant at n = 3 (for simulations with varying choice options, see ESM section \@ref(simulations)).
+For example, in a scenario with three choice options and three individuals, two might vote for option A, and one for option C, resulting in two levels of convergence, $2/3$ for A and $1/3$ for C.
+For each level of convergence occurring within a group, we then compute (i) the accuracy (either 1 if the correct option or 0 else),, (ii) the average competence of agents.
+Across all groups, we then compute the averages of these values, for each level of convergence.
+
+We repeat this procedure varying population competence distributions and the size of informant groups, holding the number of choice options constant at n = 3 (for simulations with varying choice options, see ESM section \@ref(simulations)).
 Fig.
-\@ref(fig:simulations) B shows the average accuracy (left), and the average competence (right) value for each share of votes, across different underlying competence distributions and group sizes.
-
-The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows: (i) Irrespective of group size (and number of choice options) and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate.
-(ii) For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents.
+\@ref(fig:simulations) B shows the average accuracy (left), and the average competence (right) value as a function of convergence, across different underlying competence distributions and group sizes.
+
+The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows:
+
+1.
+Irrespective of group size (and number of choice options) and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate.
+2.
+For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents.
 The strength of this relation is not much affected by the number of agents whose answers are converging, but, although it is always positive, it ranges from very weak to very strong depending on the population's competence distribution.
-(iii) The relation between convergence and accuracy is always much stronger than the relation between convergence and average competence of the agents.
-
-(ref:simulations) **A** Shows the different population competence distributions we considered in our simulations.
-In the continuous simulations, competence values of 0 correspond to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1.
+3.  The relation between convergence and accuracy is always much stronger than the relation between convergence and competence of the agents.
+
+(ref:simulations) Results of the simulations.
+**A** Shows the different population competence distributions we considered in our simulations.
+In the continuous simulations, competence values of 0 correspond to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a very small standard deviation (1, on the same scale).
 In the categorical simulations, a competence value of 0 corresponds to chance (e.g.
-in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while competence of 1 corresponds to definitely picking the correct answer.
+in a 3-choice-options scenario, an individual picking the correct answer with a probability of $1/3$), while a competence value of 1 corresponds to definitely picking the correct answer.
 **B** Shows the results of simulations in a categorical setting with three choice options.
-Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes.
+Points represent average accuracy (left)/competence (right) values by degree of convergence (measured by the share of votes for an option), for different population competence distributions (panels) and sample sizes (colors).
 **C** Shows the results in a continuous setting.
-Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the figure) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).
+Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to being further up on the y-axis) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).
 
@@
 
-We test whether people infer that informants are more likely to give accurate answers, and to be competent, when their answers converge, both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6).
+Our models indicate that groups of informants are more likely to have given accurate answers, and to be competent, when their answers converge.
+In a series of experiments, we test whether people draw these inferences both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6).
 By contrast with previous studies, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be.
-We also manipulated (Experiments 2, 3, and 5) whether the convergence of the answers could be explained by something else than accuracy, which should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.
+There has recently been much interest in investigating whether participants are able to take informational dependencies into account when evaluating convergent information [@yousifIllusionConsensusFailure2019; @mercierUtilizingSimpleCues2019; @desaiGettingSourceIllusion2022; @xieSensitivityEvidentialDependencies2022; @yinSmartlyFollowingOthers2024].
+To test whether participants took into account dependencies between the informants in the present context, this factor was manipulated in two different ways (in Experiment 2, some informants had discussed with each other; in Experiments 3 and 5, some informants had an incentive to provide the same answer).
+We predicted that dependencies between the informants’ answers should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.
 
 All experiments were preregistered.
-Preregistration documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b) (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).
+All documents, data and code can be found on Open Science Framework [project page](https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b) (<https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b>).
 All analyses were conducted in R (version 4.2.2) using R Studio.
@@
 In Experiment 1, participants were provided with a set of numerical estimates which were more or less convergent, and asked whether they thought the estimates were accurate, and whether the informants making the estimates were competent.
-Perceptions of accuracy were measured as the confidence in what the participants thought the correct answer was, on the basis of the numerical estimates provided: the more participants think they can confidently infer the correct answer, the more they must think the estimates accurate, on average.
+Perceptions of accuracy were measured as the confidence in what the participants thought the correct answer was, on the basis of the numerical estimates provided: the more participants think they can confidently infer the correct answer, the more they must think the estimates accurate, on average (the results replicate with a more direct measure of accuracy, see Experiment 3).
 Our hypotheses were:
@@
 
-We manipulated two experimental factors, with two levels each: the convergence of the estimates (how close they were to each other; levels: divergent/convergent), and the number of estimates (how many players there were; levels: low/high).
+We manipulated two experimental factors, with two levels each: the convergence of the estimates (how close they were to each other; levels: divergent/convergent), and the number of estimates (how many players there were; levels: three/ten).
 This latter factor was chiefly included to make our results more robust, and was not attached to specific hypotheses.
-We used a 2 (convergence: divergent/convergent) x 2 (number: low/high) within-participant design, with each participant seeing all the conditions.
+We used a 2 (convergence: divergent/convergent) x 2 (number: three/ten) within-participant design, with each participant seeing all the conditions.
 Participants saw two different sets of estimates per condition, for a total of eight sets of estimates per participant.
@@
 
-To account for dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slopes of convergence for participants.
+To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slopes of convergence for participants.
 In the models for our hypotheses, we control for the number of estimates provided to the participants (three or ten).
@@
 We do not find a statistically significant interaction that would indicate a difference of the effect of convergence ($\hat{b}$ = `r exp1_research_questions$exploratory_analysis$interaction$estimate` `r exp1_research_questions$exploratory_analysis$interaction$ci`, `r exp1_research_questions$exploratory_analysis$interaction$p.value`).
-We find that pooled across divergent and convergent conditions, participants reported lower perceived accuracy than competence ($\hat{b}$ = `r exp1_research_questions$exploratory_analysis$outcomeconfidence$estimate` `r exp1_research_questions$exploratory_analysis$outcomeconfidence$ci`, `r exp1_research_questions$exploratory_analysis$outcomeconfidence$p.value`)
-
-In summary, as predicted, when the informants’ answers were more convergent, participants were more confident that their answers were correct, and they believed the informants to be more competent.
-This was true whether there were three or ten estimates.
+Pooled across divergent and convergent conditions, we find that participants reported lower perceived accuracy than competence ($\hat{b}$ = `r exp1_research_questions$exploratory_analysis$outcomeconfidence$estimate` `r exp1_research_questions$exploratory_analysis$outcomeconfidence$ci`, `r exp1_research_questions$exploratory_analysis$outcomeconfidence$p.value`)
+
+In summary, as predicted, when the informants' answers were more convergent, participants were more confident that their answers were correct, and they believed the informants to be more competent.
+This was true both when there were three informants and when there were ten informants.
 
@@
 
-It is rational to infer that convergent estimates are more likely to be accurate, and to have been made by competent individuals, only if the convergence is best explained by the accuracy of the estimates.
-However, convergence could also be the outcome of other factors.
+We have shown that it is rational to infer that convergent estimates are more likely to be accurate, and to have been made by competent individuals, only if these individuals were independent and unbiased.
+However, convergence could come about differently.
 If the individuals do not make their estimates independently of each other, a single individual might exert a strong influence on the others, making their convergence a poor cue to their accuracy.
@@
 
+Note that these predictions only stem from past empirical results, and are not necessarily normatively justified [modeling the effects of discussion would be beyond the scope of this paper, but see @dietrichDeliberationWisdomCrowds2024].
+
 ## Methods
@@
 We used the materials generated for the convergent condition of Experiment 1.
-By contrast to Experiment 1, participants saw only two stimuli in total (one set of estimates per condition).
-Otherwise, we proceeded just as in experiment one: we randomly assigned individual participants to one of the three series of stimuli, and for each participant, we randomized the order of appearance of conditions.
+By contrast to Experiment 1, participants saw only two stimuli in total (one set of estimates per condition), and we only used stimuli involving groups of three informants.
+Otherwise, we proceeded just as in Experiment 1: we randomly assigned individual participants to one of the three series of stimuli, and for each participant, we randomized the order of appearance of conditions.
 
@@
 
-Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently, to be less accurate, or made by less competent individuals.
-This suggests that participants might not be sensitive to at least some forms of informational dependency between the individuals making the estimates.
-Although previous studies have found that participants sometimes discount the opinions of groups compared to those of independent individuals, the superior performance of groups over individuals for a range of tasks [for review, see, e.g., @mercierArgumentativeTheoryPredictions2016a], including numerical estimates [e.g.
-@mercierDoesDiscussionMake2022] suggests that this discounting might have been misguided.
+Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently made estimates, to be less accurate, or produced by less competent individuals.
+This might stem from the fact that participants, in various situations, neglect informational dependencies [@yousifIllusionConsensusFailure2019], or from the fact that discussing groups actually perform better than non-discussing groups in a range of tasks [for review, see, e.g., @mercierArgumentativeTheoryPredictions2016a], including numerical estimates [e.g.
+@mercierDoesDiscussionMake2022].
 As a result, the participants in the current experiment might have been behaving rationally when they did not discount the estimates made after discussion.
@@
 
-Experiment 3 tests whether participants are sensitive to another potential source of dependency between convergent estimates: when the individuals making the estimate share an incentive to bias their estimates and disregarde accuracy.
+Experiment 3 tests whether participants are sensitive to another potential source of dependency between convergent estimates: when the individuals making the estimate share an incentive to bias their estimates and disregard accuracy.
 Even though Experiment 3 is formally similar to Experiment 1, the setting is different, as participants were told that they would be looking at (fictional) predictions of experts for stock values, instead of the answers of individuals in abstract games.
 In the conflict of interest condition, the experts had an incentive to value the stock in a given way, while they had no such conflict of interest in the independence condition.
-Given the nature of this manipulation, participants might have discounted the opinions of the conflicted experts, irrespective of the degree of convergence of their estimates.
-As a result, we could not just condition on convergent answers, as we did in Experiment 2.
-Instead, we tested for an interaction, namely whether the positive effect of convergence is reduced when informants are systematically biased, compared to when they are not.
+We tested for an interaction, namely whether the positive effect of convergence is reduced when informants are systematically biased, compared to when they are not.
 On this basis, we formulate four hypotheses, two of which are identical to those of Experiment 1, and only apply in the independent condition, and two that bear on the interaction:
@@
 ***H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are in a conflict of interest.***
+
+We have not conducted simulations to validate these predictions.
+However, given the operationalization chosen, in which convergence should provide little evidence of either accuracy or competence, we believe the predictions regarding the superiority of the independent informants stem naturally from the model of independent informants presented above.
 
@@
 Each participant rated four scenarios, two for each level of convergence.
+By contrast to experiment one, all scenarios only involved groups of three informants.
 
@@
 We find evidence for all four hypotheses.
-As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample of experiment three to half of the participants, namely those who were assigned to the independence condition.
-On this reduced sample, we ran the exact same analyses as in experiment 1 and replicated the results.
+As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample of Experiment 3 to half of the participants, namely those who were assigned to the independence condition.
+On this reduced sample, we ran the exact same analyses as in Experiment 1 and replicated the results.
 As for accuracy, participants rated informants in convergent scenarios (mean = `r exp3_descriptives$means_H1$convergent$accuracy_mean`, sd = `r exp3_descriptives$means_H1$convergent$accuracy_sd`) as more accurate than in divergent ones (mean = `r exp3_descriptives$means_H1$divergent$accuracy_mean`, sd = `r exp3_descriptives$means_H1$divergent$accuracy_sd`; $\hat{b}_{\text{Accuracy}}$ = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$estimate` `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$ci`, p = `r exp3_descriptives$model_h1_accuracy$convergenceconvergent$p.value`).
@@
 
-(ref:exp3-plot) Distributions of accuracy and competence by convergence and informational dependency.
+(ref:exp3-plot) Results of Experiment 3, showing the distributions of accuracy and competence ratings by convergence and informational dependency.
 
@@
 The set-up is similar to that of Experiment 1, except that the outcomes seen by the participants are not numerical estimates, but choices made between a few options.
+An additional difference is that participants rate a focal informant, and not a group of informants.
+There were two reasons for this choice: First, accuracy is not on a continuum as in the first three experiments (an option was either correct or not), so forming an average across informants who chose different options was less sensible.
+Second, rating a focal individual allowed us to have a minority condition, which would not have been possible when providing an average rating for a group.
 Experiment 4 tests hypotheses that are analogous to those of Experiment 1:
@@
 
-To account for dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slope for participants.
+To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for participants.
 
@@
 
-In the ESM (section \@ref(exp4)), we show participants underestimate the effect of convergence on both accuracy and competence, but especially on accuracy.
+In the ESM (section \@ref(exp4)), we show that compared to what the normative models would predict, participants underestimate the effect of convergence on both accuracy and competence, but especially on accuracy.
 
@@
     convergence_categorical = recode_factor(convergence, 
-                                            `0` = "opposing majority", 
+                                            `0` = "minority", 
                                             `1` = "divergence", 
@@
 
-Experiment 5 is a conceptual replication of Experiment 3 in a categorical instead of a numerical case: are participants less likely to infer that more convergent estimates are more accurate, and the individuals who made them more competent, when the estimates are made by individuals with a conflict of interest pushing them to all provide a given answer, compared to when they are made by independent participants?
+Experiment 5 is a conceptual replication of Experiment 3 in a categorical instead of a numerical case: are participants less likely to infer that more convergent estimates are more accurate, and the individuals who made them more competent, when the estimates are made by individuals with a conflict of interest pushing them to all provide a given answer, compared to when they are made by independent individuals?
 The independence condition of Experiment 5 also serves as a replication of Experiment 4, leading to the following hypotheses:
@@
 The power simulation for accuracy suggested that for 80 participants, we would have a power of at least 90% for the interaction effect.
-The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we would also detect an effect of convergence.
+The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants would we also detect an effect of convergence.
 Due to uncertainty about our assumptions and because resources were available for a larger sample, we recruited `r exp5_descriptives$n_subj` participants, in the UK and via Prolific (`r exp5_descriptives$gender$female$n` female, `r exp5_descriptives$gender$male$n`, 1 non-identified; $age_\text{mean}$: `r exp5_descriptives$age$mean`, $age_\text{sd}$: `r exp5_descriptives$age$sd`, $age_\text{median}$: `r exp5_descriptives$age$median`).
@@
 Participants answered with a slider on a scale from 0 to 100.
-To assess perceptions of competence, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (2)"extremely competent").
+To assess perceptions of competence, we asked: "How competent do you think advisor 1 is regarding such investment recommendations?" Participants answered on a 7-point Likert scale (from (1)"not competent at all" to (7)"extremely competent").
 
@@
 In the conflict of interest condition, experts were introduced this way: "The three advisors have already invested in one of the three options, the same option for all three.
-As a result, they have an incentive to push that option in their recommendations." For the independence condition: "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations."
+As a result, they have an incentive to push that option in their recommendations." Participants assigned to the independence condition read: "The three advisors are independent of each other, and have no conflict of interest in making investment recommendations."
 
@@
 \@ref(fig:exp5-plot)).
-To test H1a and H1b, we use the same analyses as in Experiment 4, restricted on the independent condition, and replicate the results.
+To test H1a and H1b, we use the same analyses as in Experiment 4, restricted on the independence condition, and replicate the results.
 We find a positive effect of convergence on both accuracy ($\hat{b}_{\text{Accuracy}}$ = `r exp5_descriptives$model_h1_accuracy$convergence$estimate` `r exp5_descriptives$model_h1_accuracy$convergence$ci`, p = `r exp5_descriptives$model_h1_accuracy$convergence$p.value`) and competence ($\hat{b}_{\text{Competence}}$ = `r exp5_descriptives$model_h1_competence$convergence$estimate` `r exp5_descriptives$model_h1_competence$convergence$ci`, p = `r exp5_descriptives$model_h1_competence$convergence$p.value`).
@@
 
-## Experiment 6
+# Experiment 6
 
@@
 The number of choice options was manipulated between participants.
-Participants were randomly assigned to either to see stimuli with three options (as in Experiment 4), or with ten options.
+Participants were randomly assigned to either see stimuli with three options (as in Experiment 4), or stimuli with ten options.
 Participants assigned to the ten options condition were divided into one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the three options condition, and another with increased range (see ESM, section \@ref(exp6)).
@@
 
-We tested whether the effect of convergence is larger on accuracy than on competence, in an exploratory, non-preregistered analysis.
+We tested whether the effect of convergence is larger on accuracy than on competence in an exploratory, non-preregistered analysis.
 To this end, we first standardized both outcome scores to account for the different scales.
 We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for the number of choice options.
-In line with all previous experiments, find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy ($\hat{b}$ = `r exp6_exploratory_analysis$interaction$estimate` `r exp6_exploratory_analysis$interaction$ci`, `r exp6_exploratory_analysis$interaction$p.value`).
+In line with all previous experiments, we find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy ($\hat{b}$ = `r exp6_exploratory_analysis$interaction$estimate` `r exp6_exploratory_analysis$interaction$ci`, `r exp6_exploratory_analysis$interaction$p.value`).
 
@@
 
-Using both analytical arguments and simulations, we showed that, under a wide range of parameters, more convergent answers tend to be more accurate, and to have been made by more competent informants.
+Using both analytical arguments and simulations, we have shown that, under a wide range of parameters, more convergent answers tend to be more accurate, and to have been made by more competent informants.
 
@@
 
-We also observe that, in line with the simulations, participants draw stronger inferences from convergence to accuracy than to competence in five of the six experiments.
-
-On the whole, participants thus appear to draw normatively justified inferences: (i) they infer that more convergent answers are more accurate; (ii) they infer that more convergent answers are coming from more competent informants; (iii) inference (i) tends to be stronger than inference (ii); (iv) both inference (i) and (ii) are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other.
+We also observe that, in line with our simulations, participants draw stronger inferences from convergence to accuracy than to competence in five of the six experiments.
+
+On the whole, participants thus appear to draw normatively justified inferences: 1.
+They infer that more convergent answers are more accurate; 2.
+They infer that more convergent answers are coming from more competent informants; 3.
+The inference on accuracy tends to be stronger than the inference on competence ; 4.
+Both the inference on accuracy and on the inference on competence are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other.
 The only exception appears to be that participants do not take into account the number of options in categorical choices scenarios.
@@
 @mollemanStrategiesIntegratingDisparate2020; @yanivSpuriousConsensusOpinion2009; @budescuConfidenceAggregationExpert2000; @budescuEffectsAsymmetryAdvisors2003; @budescuAggregationOpinionsBased2007], and for categorical choices [e.g., @morganEvolutionaryBasisHuman2012; for review, see @mercierMajorityRulesHow2019].
-However, in these experiments, the participants arguably assumed a degree of competence among the individuals whose answers they saw.
+However, these experiments do not test whether this inference of accuracy of the information extends to an inference of competence of the informants.
+Moreover, by their design, participants arguably assumed a degree of competence among the informants.
 For instance, when children are confronted with several individuals who agree on how to name a novel object [e.g.
@@
 
-Here, we assume no prior knowledge of individual competence, asking the question: if we see a set of individuals, whose competence is unknown, converge on an answer, is it rational to infer that this answer is more likely to be correct, and that the individuals are likely to be competent?
-We show that the answer is yes on both counts--assuming there is no systematic bias among the individuals answering.
-An analytical argument and a series of simulations reveal that, for both the numerical choice context and the categorical choice context, the more individuals agree on an answer, the more likely the answer is to be correct, and the more likely the individuals are competent, with the former effect being stronger than the latter.
+Here, we assumed no prior knowledge of individual competence, asking the question: if we see informants, whose competence is unknown, converge on an answer, is it rational to infer that this answer is more likely to be correct, and that the informants are likely to be competent?
+We have shown that the answer is yes on both counts--assuming there is no systematic bias among the informants.
+An analytical argument and a series of simulations revealed that, for both the numerical choice context and the categorical choice context, the more individuals agree on an answer, the more likely the answer is correct, and the more likely the individuals are competent, with the former effect being stronger than the latter.
 Moreover, this is true for a wide range of assumed population distributions of competence.
-
-In a series of experiments, we show that participants (UK) draw these inferences: when presented with more convergent answers, they tend to believe the answers are more likely to be correct, and that the individuals who made them are more likely to be competent.
-This is true for numerical estimates and for categorical answers.
-These beliefs are weakened when the individuals making the estimates are systematically biased by a shared conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).
+In a series of experiments, we have shown that participants (UK) draw these inferences, but that they do so less when there is reason to assume a bias among informants.
 
@@
 From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate.
-Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be a valid cue when determining who tends to be an expert.
+Here, we have shown that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be a valid cue when determining who tends to be an expert.
 This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert [@nguyenCognitiveIslandsRunaway2020; @hahnHowGoodYour2018].
 
-People might draw these inferences from convergence in a variety of contexts, but the most prominent one might be science.
-Science is, arguably, the institution in which individuals end up converging the most in their opinions.
+The present experiments show that people draw inferences from convergence to accuracy and to competence, but then do not precisely show how they do it.
+As suggested in the introduction, it is plausible that participants use the combination of two heuristics: one leading them from convergence to accuracy, and one from accuracy to competence.
+This two step process is coherent with the observation that the inference from convergence to accuracy (one step) is stronger than that from convergence to competence (two steps).
+However, our results are also compatible with participants performing two inferences from convergence, one to accuracy and one to competence, the latter being weaker because the inferences roughly follow the normative model.
+Our results are coherent with work on other mechanisms of epistemic vigilance that process the combination of several informants’ opinions, showing that these mechanisms rely on sensible cues, but also systematically fail to take some subtle cues into account [@mercierMajorityRulesHow2019].
+
+A most prominent context in which the inferences uncovered here might play a role is that of science.
+Much of science is counterintuitive, and most people do not have the background knowledge to evaluate most scientific evidence.
+However, science is, arguably, the institution in which individuals end up converging the most in their opinions [on consensus being the defining trait of science by contrast with other intellectual enterprises, see @collinsSociologyPhilosophiesGlobal2002].
 For instance, scientists within many disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA.
@@
 The current study has a number of limitations.
-In our simulations, we assume agents to be independent.
+In our simulations, we assume agents to be independent and unbiased.
 Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency [@ladhaCondorcetJuryTheorem1992], more robust simulations would show that–while still assuming no systematic bias–our results hold even when agents influence each others' answers.
@@
 Although the main results replicate well across our experiments, and we can thus be reasonably certain of their robustness, it's not clear how much they can be generalized.
-Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) [@coppockGeneralizingSurveyExperiments2019].
+Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from–here, UK citizens [@coppockGeneralizingSurveyExperiments2019].
 However, we do not know whether they would generalize to other cultures.
 
-Future studies could overcome these limitations by replicating the present results in different cultures, using more ecologically valid stimuli.
+These limitations could be overcome by replicating the present results in different cultures, using more ecologically valid stimuli.
 For instance, it would be interesting to test whether the inference described here, from convergence to competence, might be partly responsible for the fact that people tend to believe scientists to be competent [@colognaTrustScientistsTheir2024].
+Finally, future studies could also attempt to systematically model cases of informational dependencies more subtle than those used in Experiments 3 and 5, to derive and test normative predictions.
 
@@
 
+```{=tex}
 \clearpage
 \newpage
-
+```
 ```{r child = "appendix_simulations.Rmd"}

### 2024-09-11 Experiment_3/power_random_effects.R

@@ -248,44 +248,7 @@ ggplot(mss,
 
 
 
-# Let's find the minimum detectable effect
-power_levels <- c()
-effects_to_try <- c(0.1, 0.2, 0.4, 0.6, 0.8, 1, 1.2)
 
-for (i in 1:length(effects_to_try)) {
-  
-  # keep track of progress
-  print(paste("effect to try: ", effects_to_try[i]))
-  
-  # iterate
-  power_levels[i] <- iterate(
-    n_subj = 200,
-    n_low  =  4,   # number of low convergence stimuli
-    n_high =  4,   # number of high convergence stimuli
-    beta_0     = 4.13,   # grand mean
-    beta_1     =  effects_to_try[i],   # effect of convergence
-    tau_0      = 0.64,   # by-subject random intercept sd
-    tau_1      =  0.99,   # by-subject random slope sd
-    rho        = 0.11,   # correlation between intercept and slope by-subject
-    sigma      = 0.955,  # residual (standard deviation)
-    iterations = 1000)
-}
-
-# Where do we cross 90%?
-power_results <- tibble(effect = effects_to_try,
-                        power = power_levels)
-power_results
-
-# plot results
-ggplot(power_results, 
-       aes(x = effect, y = power)) +
-  geom_line(color = 'red', size = 1.5) + 
-  # add a horizontal line at 90%
-  geom_hline(aes(yintercept = .9), linetype = 'dashed') + 
-  # Prettify!
-  theme_minimal() + 
-  scale_y_continuous(labels = scales::percent) + 
-  labs(x = 'Mixed model effect size of convergence', y = 'Power')
 
 
 

### 2024-09-11 Cogsci_version_raw.tex
update paper revisions round 1

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  doc,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{consensus; convergence; majority rules; wisdom of crowds; trust; trust in science}
\usepackage{csquotes}
\usepackage{placeins}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={How wise is the crowd? Can we infer people are accurate and competent merely because they agree with each other?},
  pdfauthor={Anonymous1},
  pdflang={en-EN},
  pdfkeywords={consensus; convergence; majority rules; wisdom of crowds; trust; trust in science},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{How wise is the crowd?
Can we infer people are accurate and competent merely because they agree with each other?}
\author{Anonymous\textsuperscript{1}}
\date{}

\shorttitle{How wise is the crowd?}

\affiliation{\vspace{0.5cm}\textsuperscript{1} }

\abstract{%
Are people who agree on something more likely to be right and competent? Evidence suggests that people tend to make this inference. However, standard wisdom of crowds approaches only provide limited normative grounds for this behavior. Using simulations, we argue that in stylized scenarios where individuals make independent and unbiased estimates, more convergent groups of individuals indeed tend to be more competent and accurate. Mirroring the stylized setting of the simulations, we then show that people make these inferences in two experiments. These inferences from convergence might help explain why people respect scientists' competence, even if they do not understand much about how scientific results are reached.
}

\begin{document}
\maketitle

\textbf{\emph{This is a ``raw'' version. The final version was edited in overleaf to meet formatting criteria}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Imagine that you live in ancient Greece, and a fellow called Eratostenes claims the circumference of the earth is 252000 stades (approximately 40000 kilometers). You know nothing about this man, the circumference of the Earth, or how one could measure such a thing. As a result, you discard Eratostenes' opinion and (mis)take him for a pretentious loon. But what if other scholars had arrived at very similar measurements, independently of Eratosthenes? Or even if they had carefully checked his measurement, with a critical eye? Wouldn't that give you enough ground to believe not only that the estimates might be correct, but also that Eratosthenes and his fellow scholars must be quite bright, to be able to achieve such a feat as measuring the Earth?

In this article, we explore how, under some circumstances, we should, and we do infer that a group of individuals whose answers converge are likely to be correct, and to be competent in the relevant area, even if we had no prior belief about either what the correct answer was, or about these individuals' competence. There are two ways for answers to converge: when the answers are continuous, they can have a smaller range or a lower variance; when the answers are categorical (i.e.~there are only a few, often two, options), more can agree on the same answer. These two cases have been studied with different paradigms.

In the continuous case, the most relevant evidence comes from the literature on `advice-taking.' In these experiments, participants are called `judges,' and they need to make numerical estimates--sometimes on factual knowledge, e.g.~`What year was the Suez Canal opened first? (Yaniv, 2004), sometimes on knowledge controlled by the experimenters, e.g.~'How many animals were on the screen you saw briefly?' (Molleman et al., 2020). To help answer these questions, participants are given estimates from others, the advisors.

One subset of these studies manipulate the degree of convergence between groups of advisors, through the variance of estimates (Molleman et al., 2020; Yaniv, Choshen-Hillel, \& Milyavsky, 2009), or their range (Budescu \& Rantilla, 2000; Budescu, Rantilla, Yu, \& Karelitz, 2003; Budescu \& Yu, 2007). These studies find that participants are more confident about, or rely more on, estimates from groups of advisors who converge more.

Other studies varied the degree of convergence within a group of advisors. They present participants with a set of estimates where some estimates are close to each other (or overlapping, in cases where estimates were intervals), while others are outliers (Harries, Yaniv, \& Harvey, 2004; Yaniv, 1997, study 3 \& 4). These studies find that participants heavily discount outliers when aggregating estimates.

In categorical choice contexts, there is ample and long-standing (e.g. Crutchfield, 1955) evidence from social psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger (both in absolute and in relative terms) (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see Mercier \& Morin, 2019). This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role (e.g.~because the answers are private, see Asch, 1956). Similar results have been obtained with young children (Bernard, Harris, Terrier, \& Clément, 2015; Bernard, Proust, \& Clément, 2015; Chen, Corriveau, \& Harris, 2013; Corriveau, Fusaro, \& Harris, 2009; e.g. Fusaro \& Harris, 2008; Herrmann, Legare, Harris, \& Whitehouse, 2013; Morgan, Laland, \& Harris, 2015).

If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the individuals' competence. One potential exception is a study with preschoolers (Corriveau et al., 2009) in which the children were more likely to believe the opinion of an informant who had previously been the member of a majority over that of an informant who had dissented from the majority. However, it is not clear whether the children thought the members of the majority were particularly competent, since their task--naming an object--was one in which children should already expect a high degree of competence from (adult) informants. This result might thus indicate simply that children infer that someone who disagrees with several others on how to call something is likely wrong, and thus likely less competent at least in that domain.

\hypertarget{is-inferring-competence-from-convergence-justified}{%
\subsection{Is inferring competence from convergence justified?}\label{is-inferring-competence-from-convergence-justified}}

To provide a normative answer, we conducted simulations, both for a continuous and a categorical scenario.

\hypertarget{continuous-case}{%
\subsubsection{Continuous case}\label{continuous-case}}

In these simulations, groups of agents, with each agent holding varying degrees of competence, provide numerical answers. We measure how accurate these answers are, and how much they converge (i.e.~how low their variance is). We then correlate the degree of convergence with the accuracy and with the competence of the agents.

More specifically, agents provide an estimate on a scale from 1000 to 2000 (chosen to facilitate the experiments below). Each agent is characterized by a normal distribution of possible answers. All of the agents' distributions are centered around the correct answer, but their standard deviation varies, denoting varying degrees of competence. The agents' standard deviation varied from 1 (highest competence) to 1000 (lowest competence). Each agent's competence level is drawn from a population competence distribution, expressed by a beta distribution, which can take different shapes. We conducted simulations with a variety of beta distributions which cover a large range of possible competence populations (see Fig. \ref{fig:simulations} C).

In the simulation, a population of around 990000 agents with different competence levels is generated. An answer is drawn for each agent, based on their respective competence distribution. The accuracy of this answer is defined as the squared distance to the true answer. Having a competence and an accuracy value for each agent, we randomly assign agents to groups of, e.g., three. For each group, we calculate the average of the agents' competence and accuracy. We measure the convergence of a group by calculating the standard deviation of the agents' answers. We repeat this process for different sample sizes for the groups, and different competence distributions. Fig. \ref{fig:simulations} B displays the relation of convergence with accuracy (left), and competence (right) across different underlying competence distributions and group sizes.

\hypertarget{categorical-case}{%
\subsubsection{Categorical case}\label{categorical-case}}

We simulate agents who have to decide between a number of options, one of which is correct, and whose competence varies. Competence is defined as a value p, which can differ for each agent, and which corresponds to the probability of selecting the right answer (the agents then have a probability (1-p)/(m-1), with m being the number of options, of selecting any other option). Competence values range from chance level (p = 1/m) to always selecting the correct option (p = 1). Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see see Fig. \ref{fig:simulations} C). Based on their competence level, we draw an answer for each agent. We measure an agent's accuracy as a binary outcome, namely whether they selected the correct option or not. In each simulation 999000 agents are generated, and then randomly assigned to groups (of varying size in different simulations). Within these groups, we first calculate the share of individuals voting for each answer, allowing us to measure convergence. For each level of convergence occurring within a group, we then compute (i) the average accuracy and (ii) the average competence of agents. Across all groups, we then compute the averages of these within-group averages, for each level of convergence.

We repeat this procedure varying population competence distributions, and the size of informant groups, holding the number of choice options constant at n = 3. Fig. \ref{fig:simulations} A shows the average accuracy (left), and the average competence (right) value for each share of votes, across different underlying competence distributions and group sizes.

The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows: (i) Irrespective of group size and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate. (ii) For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents. The strength of this relation is not much affected by the number of agents whose answers are converging, but, although it is always positive, it ranges from very weak to very strong depending on the initial competence distribution. (iii) The relation between convergence and accuracy is always much stronger than the relation between convergence and average competence of the agents.

\begin{figure}
\centering
\includegraphics{output/figures/simulations.pdf}
\caption{\label{fig:simulations}\textbf{A} Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 corresponds to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1. In the categorical simulations, a competence value of 0 corresponds to chance (e.g.~in a 3-choice-options scenario, an individual picking the correct answer with a probability of \(1/3\)), while competence of 1 corresponds to definitely picking the correct answer. \textbf{B} Shows the results of simulations in a categorical setting with three choice options. Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes. \textbf{C} Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the figure) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).}
\end{figure}

\hypertarget{overview-experiments}{%
\subsection{Overview experiments}\label{overview-experiments}}

We test whether people infer that individuals--we will call them informants henceforth--are more likely to give accurate answers, and to be competent, when their answers converge, both in a continuous tasks (Experiment 1), and in a categorical tasks (Experiment 2). By contrast with previous experiments, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be. We also manipulated whether the convergence of the answers could be explained by something else than accuracy, which should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.

All experiments and analyses were pre-registered, unless explicitly said so. Pre-registration documents, data and code can be found on Open Science Framework \href{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}{project page} (\url{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the \texttt{lme4} package and its \texttt{lmer()} function. Note that on the OSF, the enumeration of experiments differs, since we report only a subset of experiments here (pattern: article at hand \textasciitilde{} OSF; Exp.1 \textasciitilde{} Exp.3; Exp. 2 \textasciitilde{} Exp.5)

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

In Experiment 1, participants were told that they would be looking at (fictional) predictions of experts for stock values. They were provided with a set of numerical estimates which were more or less convergent, and asked whether they thought the estimates were accurate, and whether the experts making the estimates were competent. In a conflict of interest condition, the experts had an incentive to value the stock in a given way, while they had no such conflict of interest in the independence condition. We formulated four hypotheses:

\textbf{\emph{H1a: Participants perceive predictions of independent informants as more accurate when they converge compared to when they diverge.}}

\textbf{\emph{H1b: Participants perceive independent informants as more competent when their predictions converge compared to when they diverge.}}

\textbf{\emph{H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are in a conflict of interest.}}

\textbf{\emph{H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are in a conflict of interest.}}

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{participants}{%
\subsubsection{Participants}\label{participants}}

We ran a power analysis by simulation. The simulation code is available on OSF, and the procedure is described in the pre-registration document. The simulation suggested that 100 participants provide a significant interaction term between 95\% and 97\% of the time, given an alpha threshold for significance of 0.05. Due to uncertainty about our effect size assumptions and because we had resources for a larger sample, we recruited 199 participants for this study, from the UK and via Prolific (99 female, 100 male; \(age_\text{mean}\): 40.30, \(age_\text{sd}\): 12.72, \(age_\text{median}\): 38).

\hypertarget{procedure}{%
\subsubsection{Procedure}\label{procedure}}

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: ``You will see four scenarios in which several experts predict the future value of a stock. You have no idea how competent the experts are. It's also possible that some are really good while others are really bad. As it so happens, in all scenarios, the predictions for the value of the stock have to lie between 1000 and 2000. Other than that, the scenarios are completely unrelated: it is different experts predicting the values of different stocks every time.'' Participants then saw four scenarios (Fig. \ref{fig:stimuli} A), each introduced by a text according to the condition the participant was assigned to (see `Design'). Participants were then asked to rate the experts' accuracy (``On average, how accurate do you think these three predictions are?'' on a 7-point Likert scale - (1) ``not accurate at all'' to (7) ``extremely accurate''), and their competence (``On average, how good do you think these three experts are at predicting the value of stocks?'' - (1) ``not good at all'' to (7) ``extremely good'').

\begin{figure}

\includegraphics[width=1\linewidth]{output/figures/stimuli} \hfill{}

\caption{\textbf{A} Example of stimuli as used in Experiment 1 (independence condition). \textbf{B} Example of stimuli as used in Experiment 2. In these experiments, we treated convergence as a continuous variable, assigned the values in brackets to the respective conditions.}\label{fig:stimuli}
\end{figure}

\begin{figure}

\includegraphics[width=1\linewidth]{output/figures/stimuli-exp1} \hfill{}

\caption{\textbf{A} Example of stimuli as used in Experiment 1.}\label{fig:stimuli-exp1}
\end{figure}

\begin{figure}

\includegraphics[width=1\linewidth]{output/figures/stimuli-exp2} \hfill{}

\caption{\textbf{B} Example of stimuli as used in Experiment 2.}\label{fig:stimuli-exp2}
\end{figure}

\hypertarget{design}{%
\subsubsection{Design}\label{design}}

We manipulated two factors: informational dependency (two levels, independence and conflict of interest; between participants) and convergence (two levels, convergence and divergence; within participants). In the independence condition, the participants read ``Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.'' In the conflict of interest condition, the participants read ``All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at {[}mean of respective distribution{]} in the future.''

\hypertarget{materials}{%
\subsubsection{Materials}\label{materials}}

We generated sets of estimates from uniform distributions with varying range (60 for convergence, 600 for divergence; estimate scale ranged from 1000 to 2000). The same sets of estimates were used in the conflict of interest and in the independence condition. Each participant rated four scenarios, two for each level of convergence. More information on how the stimuli were created can be found on the OSF.

\hypertarget{results-and-discussion}{%
\subsection{Results and discussion}\label{results-and-discussion}}

To account for dependencies of observations due to our within-participant design regarding convergence, we ran mixed models, with participants as random factor for the intercept and the convergence slope. We find evidence for all four hypotheses. For the first set of hypotheses, we reduced the sample to half of the participants, namely those who were assigned to the independence condition. As for accuracy, participants rated informants in convergent scenarios (mean = 5.28, sd = 1.05) as more accurate than in divergent ones (mean = 3.40, sd = 1.08; \(\hat{b}_{\text{Accuracy}}\) = 1.88 {[}1.658, 2.102{]}, p = \textless{} .001). As for competence, participants rated informants in convergent scenarios (mean = 5.24, sd = 0.99) as more competent than in divergent ones (mean = 3.61, sd = 1.11; \(\hat{b}_{\text{Competence}}\) = 1.62 {[}1.411, 1.839{]}, p = \textless{} .001).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \ref{fig:experiments} A). In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 0.99 {[}0.634, 1.348{]}, p = \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.89 {[}0.636, 1.142{]}, p = \textless{} .001). Likewise the effect of convergence on competence was more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.80 {[}0.474, 1.13{]}, p = \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.82 {[}0.591, 1.056{]}, p = \textless{} .001).

In an exploratory, not pre-registered analysis, we tested whether the effect of convergence is bigger on accuracy than on competence. To do so, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction effect, indicating that pooled across independent and conflict of interest conditions, the effect of convergence had a smaller effect on competence than on accuracy (\(\hat{b}\) = -0.16 {[}-0.294, -0.028{]}, 0.018).

Experiment 1 shows that, believed sets of estimates to have been more accurate, and the individuals who had made them to be more competent, participants were more confident when the estimates were more convergent, when the estimates were more convergent. These inferences were less pronounced when the informants are systematically biased.

\hypertarget{experiment-2}{%
\section{Experiment 2}\label{experiment-2}}

Experiment 2 is a conceptual replication of Experiment 1 in a categorical instead of a numerical case:

\textbf{\emph{H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.}}

\textbf{\emph{H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.}}

\textbf{\emph{H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are biased (i.e.~share a conflict of interest to pick a given answer).}}

\textbf{\emph{H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are biased (i.e.~share a conflict of interest to pick a given answer).}}

\hypertarget{methods-1}{%
\subsection{Methods}\label{methods-1}}

\hypertarget{participants-1}{%
\subsubsection{Participants}\label{participants-1}}

We ran two different power analyses, one for each outcome variable. inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. The power simulation for accuracy suggested that for 80 participants, we would have a power of at least 90\% for the interaction effect. The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we would also detect a main effect of convergence. Due to uncertainty about our assumptions and because resources were available for a larger sample, we recruited 200 participants, again in the UK and via Prolific (99 female, 100, 1 non-identified; \(age_\text{mean}\): 41.88, \(age_\text{sd}\): 13.94, \(age_\text{median}\): 39).

\hypertarget{procedure-1}{%
\subsubsection{Procedure}\label{procedure-1}}

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: ``We will show you three financial advisors who are giving recommendations on investment decisions. They can choose between three investment options. Their task is to recommend one. You will see several such situations. They are completely unrelated: it is different advisors evaluating different investments every time. At first you have no idea how competent the advisors are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be difficult while others are easy. Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are.'' They were then presented with eight recommendation scenarios (Fig. \ref{fig:stimuli}). To assess perceptions of accuracy, we asked: ``What do you think is the probability of advisor 1 making the best investment recommendation?''. Participants answered with a slider on a scale from 0 to 100. To assess perceptions of accuracy, we asked: ``How competent do you think advisor 1 is regarding such investment recommendations?'' Participants answered on a 7-point Likert scale (from (1)``not competent at all'' to (2)``extremely competent'').

\hypertarget{design-1}{%
\subsubsection{Design}\label{design-1}}

We manipulated convergence within participants, by varying the ratio of players choosing the same response as a focal player (i.e.~the one that participants evaluate). The levels of convergence are: (i) consensus, where all three players pick the same option {[}\texttt{coded\ value\ =\ 3}{]}; (ii) majority, where either the third or second player picks the same option as the first player {[}\texttt{coded\ value\ =\ 2}{]}; (iii) dissensus, where all three players pick different options {[}\texttt{coded\ value\ =\ 1}{]}; (iv) minority, where the second and third player pick the same option, but one that is different from the first player's choice {[}\texttt{coded\ value\ =\ 0}{]}. In our analysis, we treat convergence as a continuous variable, assigning the coded values in squared parenthesis here. We manipulated conflict of interest between participants. In the conflict of interest condition, experts were introduced this way: ``The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations.'' For the independence condition: ``The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.''

\hypertarget{materials-1}{%
\subsubsection{Materials}\label{materials-1}}

All participants saw all four conditions of convergence (Fig. \ref{fig:stimuli} B), with two stimuli per condition, i.e.~eight stimuli in total (4 convergence levels x 2 stimuli).

\hypertarget{results-and-discussion-1}{%
\subsection{Results and discussion}\label{results-and-discussion-1}}

To account for dependencies of observations due to our within-participant design regarding convergence, we ran mixed models, with participants as random factor for the intercept and the convergence slope.

We find evidence for all four hypotheses. To test H1a and H1b, we restricted our data on the independent condition. We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 12.34 {[}10.362, 14.311{]}, p = \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.56 {[}0.459, 0.665{]}, p = \textless{} .001).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \ref{fig:experiments} B). In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 3.01 {[}0.027, 5.988{]}, p = 0.048) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 9.33 {[}7.232, 11.426{]}, p = \textless{} .001). Likewise, the effect of convergence on competence was more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.16 {[}0.014, 0.316{]}, p = 0.032) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.40 {[}0.291, 0.503{]}, p = \textless{} .001).

\begin{figure}
\centering
\includegraphics{output/figures/experiments.pdf}
\caption{\label{fig:experiments}Interaction of convergence and informational dependency in \textbf{A} Experiment 1 (continuous) and \textbf{B} Experiment 2 (categorical).}
\end{figure}

In an exploratory, not pre-registered analysis, we tested whether the effect of convergence is bigger on accuracy than on competence. To do so, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy (\(\hat{b}\) = -0.08 {[}-0.113, -0.044{]}, \textless{} .001; units in standard deviations).

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

In simulations, we have shown that the more a group of informants agrees with each other, the more they tend to be right and competent--assuming they are not systematically biased towards a false response. This is true for scenarios involving numerical estimates and for scenarios involving categorical answers, across a wide range of population competence distributions. In two experiments in which participants are deprived of any prior knowledge about the task of the informants' competence, we find that participants (UK) make these inferences. We also find that these inferences are weakened when the informants are systematically biased by a common conflict of interest.

People might draw this inference in a variety of contexts, but the most prominent one might be science. Science is, arguably, the institution in which individuals end up converging the most in their opinions. For instance, scientists within the relevant disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA. This represents an incredible degree of convergence. When people hear that scientists have measured the distance between the solar system and the center of the galaxy, if they assume that there is a broad agreement within the relevant experts, this should lead them to infer that this measure is accurate, and that the scientists who made it are competent. Experiments have already shown that increasing the degree of perceived consensus among scientists tends to increase acceptance of the consensual belief (Van Stekelenburg, Schaap, Veling, Van 'T Riet, \& Buijzen, 2022), but it hasn't been shown that the degree of consensus also affects the perceived competence of scientists.

In the case of science, the relationship between convergence and accuracy is broadly justified. However, at some points of history, there has been broad agreement on misbeliefs, such as when Christian theologians had calculated that the Earth was approximately six thousand years old. To the extent that people were aware of this broad agreement, and believed the theologians to have reached it independently of each other, this might have not only fostered acceptance of this estimate of the age of the Earth, but also a perception of the theologians as competent.

The current study has a number of limitations. If the very abstract materials allow us to remove most of the priors the participants might have, they might also reduce the ecological validity of the experiments. Although the main results replicate well, and we can thus be reasonably certain of their robustness with the present samples, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) (Coppock, 2019). However, we do not know whether they would generalize to other cultures.

\FloatBarrier

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-aschStudiesIndependenceConformity1956}{}}%
Asch, S. E. (1956). Studies of independence and conformity: I. A minority of one against a unanimous majority. \emph{Psychological Monographs: General and Applied}, \emph{70}(9), 1--70. \url{https://doi.org/10.1037/h0093718}

\leavevmode\vadjust pre{\hypertarget{ref-bernardChildrenWeighNumber2015a}{}}%
Bernard, S., Harris, P., Terrier, N., \& Clément, F. (2015). Children weigh the number of informants and perceptual uncertainty when identifying objects. \emph{Journal of Experimental Child Psychology}, \emph{136}, 70--81. \url{https://doi.org/10.1016/j.jecp.2015.03.009}

\leavevmode\vadjust pre{\hypertarget{ref-bernardFourSixYearOldChildren2015}{}}%
Bernard, S., Proust, J., \& Clément, F. (2015). Four- to Six-Year-Old Children's Sensitivity to Reliability Versus Consensus in the Endorsement of Object Labels. \emph{Child Development}, \emph{86}(4), 1112--1124. \url{https://doi.org/10.1111/cdev.12366}

\leavevmode\vadjust pre{\hypertarget{ref-budescuConfidenceAggregationExpert2000}{}}%
Budescu, D. V., \& Rantilla, A. K. (2000). Confidence in aggregation of expert opinions. \emph{Acta Psychologica}, \emph{104}(3), 371--398. \url{https://doi.org/10.1016/S0001-6918(00)00037-8}

\leavevmode\vadjust pre{\hypertarget{ref-budescuEffectsAsymmetryAdvisors2003}{}}%
Budescu, D. V., Rantilla, A. K., Yu, H.-T., \& Karelitz, T. M. (2003). The effects of asymmetry among advisors on the aggregation of their opinions. \emph{Organizational Behavior and Human Decision Processes}, \emph{90}(1), 178--194. \url{https://doi.org/10.1016/S0749-5978(02)00516-2}

\leavevmode\vadjust pre{\hypertarget{ref-budescuAggregationOpinionsBased2007}{}}%
Budescu, D. V., \& Yu, H.-T. (2007). Aggregation of opinions based on correlated cues and advisors. \emph{Journal of Behavioral Decision Making}, \emph{20}(2), 153--177. \url{https://doi.org/10.1002/bdm.547}

\leavevmode\vadjust pre{\hypertarget{ref-chenChildrenTrustConsensus2013}{}}%
Chen, E. E., Corriveau, K. H., \& Harris, P. L. (2013). Children Trust a Consensus Composed of Outgroup Members-But Do Not Retain That Trust. \emph{Child Development}, \emph{84}(1), 269--282. \url{https://doi.org/10.1111/j.1467-8624.2012.01850.x}

\leavevmode\vadjust pre{\hypertarget{ref-coppockGeneralizingSurveyExperiments2019}{}}%
Coppock, A. (2019). Generalizing from Survey Experiments Conducted on Mechanical Turk: A Replication Approach. \emph{Political Science Research and Methods}, \emph{7}(3), 613--628. \url{https://doi.org/10.1017/psrm.2018.10}

\leavevmode\vadjust pre{\hypertarget{ref-corriveauGoingFlowPreschoolers2009}{}}%
Corriveau, K. H., Fusaro, M., \& Harris, P. L. (2009). Going With the Flow: Preschoolers Prefer Nondissenters as Informants. \emph{Psychological Science}, \emph{20}(3), 372--377. \url{https://doi.org/10.1111/j.1467-9280.2009.02291.x}

\leavevmode\vadjust pre{\hypertarget{ref-crutchfieldConformityCharacter1955}{}}%
Crutchfield, R. S. (1955). Conformity and character. \emph{American Psychologist}, \emph{10}(5), 191--198. \url{https://doi.org/10.1037/h0040237}

\leavevmode\vadjust pre{\hypertarget{ref-fusaroChildrenAssessInformant2008}{}}%
Fusaro, M., \& Harris, P. L. (2008). Children assess informant reliability using bystanders{'} non-verbal cues. \emph{Developmental Science}, \emph{11}(5), 771--777. \url{https://doi.org/10.1111/j.1467-7687.2008.00728.x}

\leavevmode\vadjust pre{\hypertarget{ref-harriesCombiningAdviceWeight2004}{}}%
Harries, C., Yaniv, I., \& Harvey, N. (2004). Combining advice: the weight of a dissenting opinion in the consensus. \emph{Journal of Behavioral Decision Making}, \emph{17}(5), 333--348. \url{https://doi.org/10.1002/bdm.474}

\leavevmode\vadjust pre{\hypertarget{ref-herrmannStickScriptEffect2013}{}}%
Herrmann, P. A., Legare, C. H., Harris, P. L., \& Whitehouse, H. (2013). Stick to the script: The effect of witnessing multiple actors on children{'}s imitation. \emph{Cognition}, \emph{129}(3), 536--543. \url{https://doi.org/10.1016/j.cognition.2013.08.010}

\leavevmode\vadjust pre{\hypertarget{ref-mercierMajorityRulesHow2019}{}}%
Mercier, H., \& Morin, O. (2019). Majority rules: how good are we at aggregating convergent opinions? \emph{Evolutionary Human Sciences}, \emph{1}, e6. \url{https://doi.org/10.1017/ehs.2019.6}

\leavevmode\vadjust pre{\hypertarget{ref-mollemanStrategiesIntegratingDisparate2020}{}}%
Molleman, L., Tump, A. N., Gradassi, A., Herzog, S., Jayles, B., Kurvers, R. H. J. M., \& Bos, W. van den. (2020). Strategies for integrating disparate social information. \emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{287}(1939), 20202413. \url{https://doi.org/10.1098/rspb.2020.2413}

\leavevmode\vadjust pre{\hypertarget{ref-morganDevelopmentAdaptiveConformity2015}{}}%
Morgan, T. J. H., Laland, K. N., \& Harris, P. L. (2015). The development of adaptive conformity in young children: effects of uncertainty and consensus. \emph{Developmental Science}, \emph{18}(4), 511--524. \url{https://doi.org/10.1111/desc.12231}

\leavevmode\vadjust pre{\hypertarget{ref-morganEvolutionaryBasisHuman2012}{}}%
Morgan, T. J. H., Rendell, L. E., Ehn, M., Hoppitt, W., \& Laland, K. N. (2012). The evolutionary basis of human social learning. \emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{279}(1729), 653--662. \url{https://doi.org/10.1098/rspb.2011.1172}

\leavevmode\vadjust pre{\hypertarget{ref-vanstekelenburgScientificConsensusCommunicationContested2022}{}}%
Van Stekelenburg, A., Schaap, G., Veling, H., Van 'T Riet, J., \& Buijzen, M. (2022). Scientific-Consensus Communication About Contested Science: A Preregistered Meta-Analysis. \emph{Psychological Science}, \emph{33}(12), 1989--2008. \url{https://doi.org/10.1177/09567976221083219}

\leavevmode\vadjust pre{\hypertarget{ref-yanivWeightingTrimmingHeuristics1997}{}}%
Yaniv, I. (1997). \emph{Weighting and Trimming: Heuristics for Aggregating Judgments under Uncertainty}. 13.

\leavevmode\vadjust pre{\hypertarget{ref-yanivReceivingOtherPeople2004}{}}%
Yaniv, I. (2004). Receiving other people{'}s advice: Influence and benefit. \emph{Organizational Behavior and Human Decision Processes}, \emph{93}(1), 1--13. \url{https://doi.org/10.1016/j.obhdp.2003.08.002}

\leavevmode\vadjust pre{\hypertarget{ref-yanivSpuriousConsensusOpinion2009}{}}%
Yaniv, I., Choshen-Hillel, S., \& Milyavsky, M. (2009). Spurious consensus and opinion revision: Why might people be more confident in their less accurate judgments? \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{35}(2), 558--563. \url{https://doi.org/10.1037/a0014589}

\end{CSLReferences}

\end{document}

### 2024-09-11 appendix_exp6.Rmd
update paper revisions round 1

@@
 image_paths <- data.frame(
-  condition = c("opposing majority (0)", "dissensus (1)", "majority (2)", "consensus (10)"),
+  condition = c("minority (0)", "dissensus (1)", "majority (2)", "consensus (3)"),
   imgage_a = c(

### 2024-09-11 paper.tex
update paper revisions round 1

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  doc,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\makeatletter
\usepackage{etoolbox}
\patchcmd{\maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\usepackage{csquotes}
\usepackage{placeins}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?},
  pdflang={en-EN},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?}
\author{\phantom{0}}
\date{}

\shorttitle{How wise is the crowd?}

\affiliation{\vspace{0.5cm}\textsuperscript{} }

\abstract{%
Are people who agree on something more likely to be right and competent? Evidence suggests that people tend to make this inference. However, standard wisdom of crowds approaches only provide limited normative grounds. Using simulations and analytical arguments, we argue that when individuals make independent and unbiased estimates, under a wide range of parameters, individuals whose answers converge with each other tend to have more accurate answers and to be more competent. In 6 experiments (UK participants, total N = 1197), we show that participants infer that informants who agree have more accurate answers and are more competent, even when they have no priors, and that these inferences are weakened when the informants were systematically biased. In conclusion, we speculate that inferences from convergence to accuracy and competence might help explain why people deem scientists competent, even if they have little understanding of science.
}

\begin{document}
\maketitle

\section{Introduction}\label{introduction}

Imagine that you live in ancient Greece, and a fellow called Eratostenes claims the circumference of the earth is 252000 stades (approximately 40000 kilometers). You know nothing about this man, the circumference of the Earth, or how one could measure such a thing. As a result, you discard Eratostenes' opinion and (mis)take him for a pretentious loon. But what if other scholars had arrived at very similar measurements, independently of Eratosthenes? Or even if they had carefully checked his measurement, with a critical eye? Wouldn't that give you enough ground to believe not only that the estimates might be correct, but also that Eratosthenes and his fellow scholars must be quite bright, to be able to achieve such a feat as measuring the Earth?

In this article, we explore how, under some circumstances, we should, and we do infer that a group of individuals whose answers converge are likely to be correct, and to be competent in the relevant area, even if we had no prior belief about either what the correct answer was, or about these individuals' competence.

We begin by reviewing existing studies showing that people infer that competent informants who converge in their opinions are likely to be accurate. The wisdom of crowds literature provides normative grounds for this inference. We then argue that both the experimental and theoretical literature have paid little attention to extending this inference to cases in which there is no information about the informants' competence, and to inferences about the competence of the informants. We first develop normative models, both analytically and with simulations, to show that inferences from convergence to accuracy and to competence are warranted under a wide range of parameters. Second, we present a series of experiments in which participants evaluate both the accuracy and competence of informants as a function of how much their answers converge on a given problem, in the absence of any priors about these individuals' competence, or what the correct answer is.

\section{Do people infer that individuals whose answers converge tend to be right, and to be competent?}\label{do-people-infer-that-individuals-whose-answers-converge-tend-to-be-right-and-to-be-competent}

The literature on the wisdom of crowds has treated separately situations with continuous answers, such as the weight of an ox in Galton's famous observation (Galton, 1907), and with categorical answers, as when voters have to choose between two options, in the standard Condorcet Jury Theorem (De Condorcet, 2014). The continuous and the categorical case are typically modeled with different tools, and they have usually been studied in different empirical literatures (see below). Given that they both represent common ways for answers to converge more or less (e.g.~when people give numerical estimates vs.~vote on one of a limited number of options), we treat them both here, with different simulations and experiments.

In the continuous case, the most relevant evidence comes from the literature on `advice-taking' (for review, see, Kämmer, Choshen-Hillel, Müller-Trede, Black, \& Weibler, 2023). In these experiments, participants are called `judges' who need to make numerical estimates--sometimes on factual knowledge, e.g.~`What year was the Suez Canal opened first?' (Yaniv, 2004), sometimes on knowledge controlled by the experimenters, e.g.~`How many animals were on the screen you saw briefly?' (Molleman et al., 2020). To help answer these questions, participants are given estimates from others, the `advisors'.

Most of this literature is irrelevant to the point at hand since participants are presented with single estimates, either from a single advisor (e.g. Bednarik \& Schultze, 2015; Harvey \& Fischer, 1997; Soll \& Larrick, 2009; Yaniv, 2004; Yaniv \& Kleinberger, 2000), or as an average coming from a group of advisors (e.g. Jayles et al., 2017; Mannes, 2009), but without any information about the distribution of initial estimates, so that we cannot tell whether participants put more weight on more convergent answers.

Some advice-taking studies provide participants with a set of individual estimates. One subset of these studies manipulates the degree of convergence between groups of advisors, through the variance of estimates (Molleman et al., 2020; Yaniv, Choshen-Hillel, \& Milyavsky, 2009), or their range (Budescu \& Rantilla, 2000; Budescu, Rantilla, Yu, \& Karelitz, 2003; Budescu \& Yu, 2007). These studies find that participants are more confident about, or rely more on, estimates from groups of advisors that converge more.

Other studies manipulated the degree of convergence within a group of advisors. These studies present participants with a set of estimates, some of which are close to each other, while others are outliers (Harries, Yaniv, \& Harvey, 2004; Yaniv, 1997, study 3 \& 4). These studies find that participants discount outliers when aggregating estimates.

Studies on advice taking thus suggest participants believe that more convergent opinions are more likely to be correct. None of these studies investigated whether participants also believe that those whose opinions converge are also more likely to possess greater underlying competence.

In categorical choice contexts, there is ample and long-standing (e.g. Crutchfield, 1955) evidence from experimental psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see H. Mercier \& Morin, 2019). This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role (e.g.~because the answers are private, see Asch, 1956). Similar results have been obtained with young children (Bernard, Harris, Terrier, \& Clément, 2015; Bernard, Proust, \& Clément, 2015; Chen, Corriveau, \& Harris, 2013; Corriveau, Fusaro, \& Harris, 2009; e.g. Fusaro \& Harris, 2008; Herrmann, Legare, Harris, \& Whitehouse, 2013; Morgan, Laland, \& Harris, 2015).

If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the informants' competence. One potential exception is a study with preschoolers in which the children were more likely to believe the opinion of an informant who had previously been the member of a majority over that of an informant who had dissented from the majority (Corriveau et al., 2009). However, it is not clear whether the children thought the members of the majority were particularly competent, since their task--naming an object--was one in which children should already expect a high degree of competence from (adult) informants. This result might thus indicate simply that children infer that someone who disagrees with several others on how to call something is likely wrong, and thus likely less competent at least in that domain.

\section{What inferences from convergence should we expect people to draw?}\label{what-inferences-from-convergence-should-we-expect-people-to-draw}

Should we expect that people be able to infer that more convergent answers likely indicate not only more accurate answers, but also that those who gave the answers were competent? In order to make the best of communicated information, humans have to be able to evaluate it, so as to discard inaccurate or harmful information, while accepting accurate and beneficial information (Smith \& Harper, 2003). It has been argued that a suite of cognitive mechanisms--mechanisms of epistemic vigilance--evolved to serve this function (Hugo Mercier, 2020; Sperber et al., 2010). Since the opinion of more than one individual is often available to us, there should be mechanisms of epistemic vigilance dedicated to processing such situations. It would be these mechanisms that lead us to put more weight on an opinion that is shared by a larger majority (in relative or absolute terms), and, in some cases at least, to discount majority opinion when the opinions haven't been formed independently of each other (Hugo Mercier \& Miton, 2019). Evidence suggests that these mechanisms rely on heuristics which become more refined with age (Morgan et al., 2015), and which are far from perfect (Hugo Mercier \& Miton, 2019; in particular, they ignore many cases of informational dependencies, see, e.g. Yousif, Aboody, \& Keil, 2019). As mentioned above, in the experiments evaluating how people process convergent information, the participants had grounds to believe that the information came from competent informants. However, the same mechanisms could lead people to perform the same inference when they do not have such information (especially since, as is shown presently, such an inference is warranted).

Regarding the inference from convergence to competence, other cognitive mechanisms allow us to infer how competent people are, based on a variety of cues, from visual access (Pillow, 1989), to the time it takes to answer a question (Richardson \& Keil, 2022). One of the most basic of these mechanisms infers from the fact that someone was right, that they possess some underlying competence (e.g. Koenig, Clément, \& Harris, 2004). As a result, if participants infer that convergent opinions are more likely to be accurate, they should also infer that the informants who provided the opinions are competent. Before testing whether participants infer accuracy and competence from convergence, we show that these inferences are normatively warranted, both in the categorical and in the continuous case.

\subsection{Analytical argument}\label{analytical-argument}

Regarding the analytical answer, the question can be broken down into two questions. First, can we infer that a population of informants whose answers converge more is, on average, more competent? Second, can we infer that, within this population, individuals who are closer to the consensus or to the average answer are more competent?

In the continuous case, let us imagine a population of informants. Individual opinions are drawn from a normal distribution, centered on the correct answer (i.e., informants have no systematic bias, as for instance in Galton's classic demonstration, Galton, 1907). The variance of this distribution represents the individuals' competence: the larger the variance, the lower the competence.~ We observe the individual answers. In this setting, it is well known in statistics (see also Electronic Supplementary Materials, ESM) that the sample mean (i.e.~the mean of the answers of all the informants) is the best estimator of the correct answer, and the sample variance (i.e.~the mean squared distance between the answers and the sample mean) is the best estimator of the population's average competence (best understood here as the least volatile estimator). This means that a population of informants whose answers converge more (lower sample variance) is, on average, more competent (informants tend to answer with a lower variance).

We can extend this argument from populations to individuals: Consider that competence varies within a population, such that each informant's answer is drawn from their own distribution -- always centered on the correct answer, but with different variances. In this case, the distance between each informant's answer and the sample mean provides the best estimate of that informant's competence (i.e.~individuals whose answers are further away from the sample mean tend to be less competent).

In the categorical case, we define the competence of an informant as the probability of choosing the correct answer. The law of large numbers implies that the relative size of the majority -- the share of informants who choose the most chosen answer -- is, when the population is large enough, a good approximation of the average competence of the population (e.g.~if the average competence is .66, and there are enough informants, then approximately 66\% of informants will select the right answer). For smaller populations, the relationship holds, with some degree of noise. In other words, the more the answers converge, the more the population can be inferred to be competent (and the larger the population, the more reliable the inference).

Now, can we infer that informants belonging to the majority are more competent? To do so, we must assume some distribution of competence in the population. Using Bayes theorem, we can then show (see ESM, section \ref{analytical}) that the more informants agree with a focal informant, the more the focal informant can be inferred to be competent (Fig. \ref{fig:figure-Benoit-2}). We also find that, the more competence varies in the population, the more the degree of convergence is indicative of an informant's competence. For example, if competence is roughly uniform in the population, then being part of a minority likely reflects bad luck, rather than incompetence. More generally, the fact that the focal individual is right (their `accuracy') can be inferred more strongly than their competence, as there is noise in the answer choice (an incompetent individual can always pick the correct answer by chance and vice versa). Figure \ref{fig:figure-Benoit-2} provides an example with a specific categorical choice scenario, under two different population distributions of competence.

\begin{figure}

\includegraphics[width=0.75\linewidth]{figures/figure_Benoit-2} \hfill{}

\caption{Results of the analytic argument. The figure shows the average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario. Here, we assume that 20 individuals answer a 5-choice question. For values higher than four, we assume that the focal individual is part of the majority. Below four, we assume that the focal individual is part of a minority, and that the majority size is four. Two situations are represented: one in which there is low variance in the distribution of competence in the population (orange line in the insert), and one in which there is high variance in this distribution (blue line in the insert).}\label{fig:figure-Benoit-2}
\end{figure}

\subsection{Simulations}\label{simulations}

In order to better understand the influence of different parameters (e.g.~degree of convergence, number of individuals) on the relationship between convergence, accuracy, and competence of informants, we conducted simulations. In all simulations, we assume agents to be unbiased and independent in their answers, but varying in their competence. All code and data regarding the simulations can be found on Open Science Framework \href{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}{project page} (\url{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}).

\subsubsection{Continuous case}\label{continuous-case}

Groups of agents, with each agents' competence varying, provide numerical answers. We measure how accurate these answers are, and how much they converge (i.e.~how low their variance is). We then look at the relationship between convergence and both the accuracy of the answers and the competence of the agents.

More specifically, agents provide an estimate on a scale from 1000 to 2000 (chosen to match the experiments below). Each agent is characterized by a normal distribution of possible answers. All of the agents' distributions are centered around the correct answer, but their standard deviation varies, representing varying degrees of competence. The agents' standard deviation varies from 1 (highest competence) to 1000 (lowest competence). Each agent's competence is drawn from a population competence distribution, expressed by a beta distribution, which can take different shapes. We conducted simulations with a variety of beta distributions which cover a wide range of possible competence populations (see Fig. \ref{fig:simulations} A).

A population of around 990000 agents (varying slightly as a function of group sizes) with different competence levels is generated. An answer is drawn for each agent, based on their respective competence distribution. The accuracy of this answer is defined as the squared distance to the true answer. Having a competence and an accuracy value for each agent, we randomly assign agents to groups of, e.g., three. For each group, we calculate the average of the agents' competence and accuracy. We measure the convergence of a group's answers by calculating the standard deviation of the agents' answers. We repeat this process for different sample sizes for the groups, and different competence distributions. Fig. \ref{fig:simulations} C displays the resulting relation of convergence with accuracy (left), and competence (right) for different underlying competence distributions and group sizes. We draw broad conclusions from these results after reporting the outcome of the simulations with categorical answers.

\subsubsection{Categorical case}\label{categorical-case}

In the case of categorical answers, convergence can be measured as the relative size of informants picking the same option. The Condorcet jury theorem (De Condorcet, 2014; for a recent treatment, see Dietrich \& Spiekermann, 2013) and its extensions to situations with more than two options (e.g., Hastie \& Kameda, 2005) already shows that the answer defended by the majority/plurality is more likely to be correct. Regarding competence, Romeijn and Atkinson (2011) have shown that when individuals are more competent, the share of the majority vote tends to increase. However, they have only studied a case with a uniform distribution of competence. By contrast, here, we investigate a wide range of distributions of competence, and we do not assume that all individuals are equally competent, meaning that, from an observed set of answers, we can attribute an average competence to individuals whose answers form a majority/plurality vs.~individuals whose answers form a minority.

We simulate agents whose competence varies and who have to decide between a number of options, one of which is correct. Competence is defined as a value p which corresponds to the probability of selecting the right answer (the agents then have a probability (1-p)/(m-1), with m being the number of options, of selecting any other option). Competence values range from chance level (p = 1/m) to always selecting the correct option (p = 1). Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see Fig. \ref{fig:simulations} A). Based on their competence level, we draw an answer for each agent. We measure an agent's accuracy as a binary outcome, namely whether they selected the correct option or not. In each simulation around 99900 agents (varying slightly as a function of the group size) are generated, and then randomly assigned to groups (of varying size in different simulations). Within these groups, we first calculate the share of individuals voting for each answer, allowing us to measure convergence. For example, in a scenario with three choice options and three individuals, two might vote for option A, and one for option C, resulting in two levels of convergence, \(2/3\) for A and \(1/3\) for C. For each level of convergence occurring within a group, we then compute (i) the accuracy (either 1 if the correct option or 0 else),, (ii) the average competence of agents. Across all groups, we then compute the averages of these values, for each level of convergence.

We repeat this procedure varying population competence distributions and the size of informant groups, holding the number of choice options constant at n = 3 (for simulations with varying choice options, see ESM section \ref{simulations}). Fig. \ref{fig:simulations} B shows the average accuracy (left), and the average competence (right) value as a function of convergence, across different underlying competence distributions and group sizes.

The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Irrespective of group size (and number of choice options) and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate.
\item
  For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents. The strength of this relation is not much affected by the number of agents whose answers are converging, but, although it is always positive, it ranges from very weak to very strong depending on the population's competence distribution.
\item
  The relation between convergence and accuracy is always much stronger than the relation between convergence and competence of the agents.
\end{enumerate}

\begin{figure}
\centering
\includegraphics{output/figures/simulations.pdf}
\caption{\label{fig:simulations}Results of the simulations. \textbf{A} Shows the different population competence distributions we considered in our simulations. In the continuous simulations, competence values of 0 correspond to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a very small standard deviation (1, on the same scale). In the categorical simulations, a competence value of 0 corresponds to chance (e.g.~in a 3-choice-options scenario, an individual picking the correct answer with a probability of \(1/3\)), while a competence value of 1 corresponds to definitely picking the correct answer. \textbf{B} Shows the results of simulations in a categorical setting with three choice options. Points represent average accuracy (left)/competence (right) values by degree of convergence (measured by the share of votes for an option), for different population competence distributions (panels) and sample sizes (colors). \textbf{C} Shows the results in a continuous setting. Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to being further up on the y-axis) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).}
\end{figure}

\section{Overview of the experiments}\label{overview-of-the-experiments}

Our models indicate that groups of informants are more likely to have given accurate answers, and to be competent, when their answers converge. In a series of experiments, we test whether people draw these inferences both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6). By contrast with previous studies, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be. There has recently been much interest in investigating whether participants are able to take informational dependencies into account when evaluating convergent information (Desai, Xie, \& Hayes, 2022; Hugo Mercier \& Miton, 2019; Xie \& Hayes, 2022; Yin, Xu, Lin, Zhou, \& Guo, 2024; Yousif et al., 2019). To test whether participants took into account dependencies between the informants in the present context, this factor was manipulated in two different ways (in Experiment 2, some informants had discussed with each other; in Experiments 3 and 5, some informants had an incentive to provide the same answer). We predicted that dependencies between the informants' answers should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.

All experiments were preregistered. All documents, data and code can be found on Open Science Framework \href{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}{project page} (\url{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}). All analyses were conducted in R (version 4.2.2) using R Studio. For most statistical models, we relied on the \texttt{lme4} package and its \texttt{lmer()} function. Unless mentioned otherwise, we report unstandardized model coefficients that can be interpreted in units of the scales we use for our dependent variables.

\section{Experiment 1}\label{experiment-1}

In Experiment 1, participants were provided with a set of numerical estimates which were more or less convergent, and asked whether they thought the estimates were accurate, and whether the informants making the estimates were competent. Perceptions of accuracy were measured as the confidence in what the participants thought the correct answer was, on the basis of the numerical estimates provided: the more participants think they can confidently infer the correct answer, the more they must think the estimates accurate, on average (the results replicate with a more direct measure of accuracy, see Experiment 3). Our hypotheses were:

\textbf{\emph{H1: When making a guess based on the estimates of (independent) informants, participants will be more confident about their guess when these estimates converge compared to when they diverge.}}

\textbf{\emph{H2: Participants perceive (independent) informants whose estimates converge more as more competent than informants whose estimates diverge.}}

We had three research questions regarding the number of informants which report in the ESM (section \ref{exp1}).

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

We recruited 200 participants from the UK via Prolific (100 female, 100 male; \(age_\text{mean}\): 39.73, \(age_\text{sd}\): 15.39, \(age_\text{median}\): 35.50). Not a single participant failed our attention check. The sample size was determined on the basis of a power analysis for a t-test to detect the difference between two dependent means (``matched pairs'') run on G*Power3. The analysis suggested that a combined sample of 199 would provide us with 80\% power to detect a true effect size of Cohen's d \(\geq\) 0.2 (alpha = .05, two-tailed).

\subsubsection{Procedure}\label{procedure}

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: ``Some people are playing games in which they have to estimate various quantities. Each game is different. You have no idea how competent the people are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be hard while others are easy. Across various games, we will give you the answers of several players, and ask you questions about how good they are. As it so happens, for all the games, the estimates have to be between 1000 and 2000, but all the games are completely different otherwise, and might require different skills, be of different difficulties, etc. Each player in the game makes their own estimate, completely independent of the others''. After being presented with the results of a game (Fig. \ref{fig:stimulus-exp1}), participants had to (i) make a guess about the correct answer based on the estimates they see, ``What would you guess is the correct answer, if there is one?'', (ii) estimate their confidence in this guess, ``How confident are you that your answer is at least approximately correct?'' on a 7-point Likert scale (``not confident at all'' to ``extremely confident''), (iii) estimate the competence of the group of players whose estimates they saw, ``On average, how good do you think these players are at the game?'', also on a 7-point Likert scale (from ``not good at all'' to ``extremely good'').

\begin{figure}

\includegraphics[width=1\linewidth]{output/figures/stimulus-exp1} \hfill{}

\caption{Example of two stimuli from Experiment 1, both in the 10 players condition, \textbf{A} corresponding to the convergent, \textbf{B} to the divergent condition. Similar stimuli are used in Experiments 2 and 3.}\label{fig:stimulus-exp1}
\end{figure}

\subsubsection{Design}\label{design}

We manipulated two experimental factors, with two levels each: the convergence of the estimates (how close they were to each other; levels: divergent/convergent), and the number of estimates (how many players there were; levels: three/ten). This latter factor was chiefly included to make our results more robust, and was not attached to specific hypotheses. We used a 2 (convergence: divergent/convergent) x 2 (number: three/ten) within-participant design, with each participant seeing all the conditions. Participants saw two different sets of estimates per condition, for a total of eight sets of estimates per participant.

\subsubsection{Materials}\label{materials}

We generated sets of estimates with random draws from normal distributions. First, we varied the standard deviation of these distributions to simulate the degree of convergence (150 for divergence, 20 for convergence; estimate scale ranged from 1000 to 2000). Second, we varied the number of draws (either three or ten) from these distributions. For each of the four possible resulting conditions, we generated two random draws. We repeated this process for three different sets of estimates, and participants were randomly assigned to one of these sets. More information on how the stimuli were created can be found in the ESM, section \ref{exp1}.

\subsection{Results and discussion}\label{results-and-discussion}

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slopes of convergence for participants. In the models for our hypotheses, we control for the number of estimates provided to the participants (three or ten). Visualizations and descriptive statistics can be found in ESM, section \ref{exp1}. We find a positive effect of convergence on accuracy: Participants were more confident about their estimate in convergent scenarios (mean = 4.56, sd = 1.45) than in divergent ones (mean = 3.19, sd = 1.39; \(\hat{b}_{\text{Accuracy}}\) = 1.37 {[}1.225, 1.51{]}, p = \textless{} .001). We also find a positive effect of convergence on competence: participants rated players as more competent in convergent scenarios (mean = 4.75, sd = 1.24) than in divergent ones (mean = 3.52, sd = 1.27; \(\hat{b}_{\text{Competence}}\) = 1.23 {[}1.065, 1.4{]}, p = \textless{} .001).

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To this end, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for the number of informants. We do not find a statistically significant interaction that would indicate a difference of the effect of convergence (\(\hat{b}\) = 0.14 {[}-0.001, 0.271{]}, 0.052). Pooled across divergent and convergent conditions, we find that participants reported lower perceived accuracy than competence (\(\hat{b}\) = -0.26 {[}-0.359, -0.156{]}, \textless{} .001)

In summary, as predicted, when the informants' answers were more convergent, participants were more confident that their answers were correct, and they believed the informants to be more competent. This was true both when there were three informants and when there were ten informants.

\section{Experiment 2}\label{experiment-2}

We have shown that it is rational to infer that convergent estimates are more likely to be accurate, and to have been made by competent individuals, only if these individuals were independent and unbiased. However, convergence could come about differently. If the individuals do not make their estimates independently of each other, a single individual might exert a strong influence on the others, making their convergence a poor cue to their accuracy. Alternatively, all individuals might have an incentive to provide a similar, but not accurate answer. In Experiment 2, we investigate the first possibility, and the second in Experiment 3. In particular, for Experiment 2 we rely on past results showing that participants, under some circumstances, put less weight on opinions that have been formed through discussion, by contrast with more independent opinions (Harkins \& Petty, 1987; Einav, 2018; Hess \& Hagen, 2006; see also Lopes, Vala, \& Garcia-Marques, 2007). We sought to replicate this finding in the context of convergent estimates, formulating the following hypotheses:

\textbf{\emph{H1: When making a guess based on convergent estimates of informants, participants will be more confident about their guess when informants were independent compared to when they weren't (i.e.~they could discuss before).}}

\textbf{\emph{H2: Participants perceive informants whose estimates converge as more competent when they are independent, compared to when they weren't (i.e.~they could discuss before).}}

Note that these predictions only stem from past empirical results, and are not necessarily normatively justified (modeling the effects of discussion would be beyond the scope of this paper, but see Dietrich \& Spiekermann, 2024).

\subsection{Methods}\label{methods-1}

\subsubsection{Participants}\label{participants-1}

We recruited 200 participants from the UK via Prolific (100 female, 99 male, 1 not-identified; \(age_\text{mean}\): 40.54, \(age_\text{sd}\): 13.56, \(age_\text{median}\): 38.50). Not a single participant failed our attention check. As for experiment 1, the sample size was determined on the basis of a power analysis for a t-test to detect the difference between two dependent means (``matched pairs'') run on G*Power3. The analysis suggested that a combined sample of 199 would provide us with 80\% power to detect a true effect size of Cohen's d \(\geq\) 0.2 (alpha = .05, two-tailed).

\subsubsection{Design}\label{design-1}

In a within-participants design, participants saw both an independence condition, in which they were told ``Players are asked to make completely independent decisions -- they cannot see each other's estimates, or talk with each other before giving their estimates,'' and a dependence condition, in which they were told ``Players are asked to talk with each other about the game at length before giving their estimates.''

\subsubsection{Materials}\label{materials-1}

We used the materials generated for the convergent condition of Experiment 1. By contrast to Experiment 1, participants saw only two stimuli in total (one set of estimates per condition), and we only used stimuli involving groups of three informants. Otherwise, we proceeded just as in Experiment 1: we randomly assigned individual participants to one of the three series of stimuli, and for each participant, we randomized the order of appearance of conditions.

\subsection{Results and discussion}\label{results-and-discussion-1}

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept for participants. Visualizations and descriptive statistics can be found in ESM, section \ref{exp2}. The data does not support our hypotheses. Participants were slightly less confident about their estimates when the converging informants were independent (mean = 3.78, sd = 1.50), compared to when they discussed (mean = 4.03, sd = 1.39; \(\hat{b}_{\text{Accuracy}}\) = -0.26 {[}-0.462, -0.048{]}, p = 0.016). The effect is small, but in the opposite direction of what we had predicted. We do not find an effect regarding competence (\(\hat{b}_{\text{Competence}}\) = -0.12 {[}-0.272, 0.032{]}, p = 0.120).

Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently made estimates, to be less accurate, or produced by less competent individuals. This might stem from the fact that participants, in various situations, neglect informational dependencies (Yousif et al., 2019), or from the fact that discussing groups actually perform better than non-discussing groups in a range of tasks (for review, see, e.g., H. Mercier, 2016), including numerical estimates (e.g. H. Mercier \& Claidière, 2022). As a result, the participants in the current experiment might have been behaving rationally when they did not discount the estimates made after discussion.

\section{Experiment 3}\label{experiment-3}

Experiment 3 tests whether participants are sensitive to another potential source of dependency between convergent estimates: when the individuals making the estimate share an incentive to bias their estimates and disregard accuracy. Even though Experiment 3 is formally similar to Experiment 1, the setting is different, as participants were told that they would be looking at (fictional) predictions of experts for stock values, instead of the answers of individuals in abstract games. In the conflict of interest condition, the experts had an incentive to value the stock in a given way, while they had no such conflict of interest in the independence condition. We tested for an interaction, namely whether the positive effect of convergence is reduced when informants are systematically biased, compared to when they are not. On this basis, we formulate four hypotheses, two of which are identical to those of Experiment 1, and only apply in the independent condition, and two that bear on the interaction:

\textbf{\emph{H1a: Participants perceive predictions of independent informants as more accurate when they converge compared to when they diverge.}}

\textbf{\emph{H1b: Participants perceive independent informants as more competent when their predictions converge compared to when they diverge.}}

\textbf{\emph{H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are in a conflict of interest.}}

\textbf{\emph{H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are in a conflict of interest.}}

We have not conducted simulations to validate these predictions. However, given the operationalization chosen, in which convergence should provide little evidence of either accuracy or competence, we believe the predictions regarding the superiority of the independent informants stem naturally from the model of independent informants presented above.

\subsection{Methods}\label{methods-2}

\subsubsection{Participants}\label{participants-2}

The interaction design of our third experiment made the power analysis more complex and less standard than for experiments one and two. Because we could build upon data from the first experiment, we ran a power analysis by simulation. The simulation code is available on the OSF, and the procedure is described in the preregistration document. The simulation suggested that 100 participants provide a significant interaction term between 95\% and 97\% of the time, given an alpha threshold for significance of 0.05. Due to uncertainty about our effect size assumptions and because we had resources for a larger sample, we recruited 199 participants for this study -- again, from the UK and via Prolific (99 female, 100 male; \(age_\text{mean}\): 40.30, \(age_\text{sd}\): 12.72, \(age_\text{median}\): 38).

\subsubsection{Procedure}\label{procedure-1}

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: ``You will see four scenarios in which several experts predict the future value of a stock. You have no idea how competent the experts are. It's also possible that some are really good while others are really bad. As it so happens, in all scenarios, the predictions for the value of the stock have to lie between 1000 and 2000. Other than that, the scenarios are completely unrelated: it is different experts predicting the values of different stocks every time.'' Participants then saw the four scenarios, each introduced by a text according to the condition the participant was assigned to. To remove any potential ambiguity about participants' inferences on the accuracy of the estimates, we replaced the question about confidence to one bearing directly on accuracy: ``On average, how accurate do you think these three predictions are?'' on a 7-point Likert scale (``not accurate at all'' to ``extremely accurate''). The question about competence read: ``On average, how good do you think these three experts are at predicting the value of stocks?'', also assessed on a 7-point Likert scale (from ``not good at all'' to ``extremely good'').

\subsubsection{Design}\label{design-2}

We manipulated two factors: informational dependency (two levels, independence and conflict of interest; between participants) and convergence (two levels, convergence and divergence; within participants). In the independence condition, the participants read ``Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.'' In the conflict of interest condition, the participants read ``All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at {[}mean of respective distribution{]} in the future.''

\subsubsection{Materials}\label{materials-2}

The distributions presented were similar to those of Experiment 1, although generated in a slightly different manner (see ESM, section \ref{exp3}). Each participant rated four scenarios, two for each level of convergence. By contrast to Experiment 1, all scenarios only involved groups of three informants.

\subsection{Results and discussion}\label{results-and-discussion-2}

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants. We find evidence for all four hypotheses. As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample of Experiment 3 to half of the participants, namely those who were assigned to the independence condition. On this reduced sample, we ran the exact same analyses as in Experiment 1 and replicated the results. As for accuracy, participants rated informants in convergent scenarios (mean = 5.28, sd = 1.05) as more accurate than in divergent ones (mean = 3.40, sd = 1.08; \(\hat{b}_{\text{Accuracy}}\) = 1.88 {[}1.658, 2.102{]}, p = \textless{} .001). As for competence, participants rated informants in convergent scenarios (mean = 5.24, sd = 0.99) as more competent than in divergent ones (mean = 3.61, sd = 1.11; \(\hat{b}_{\text{Competence}}\) = 1.62 {[}1.411, 1.839{]}, p = \textless{} .001).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \ref{fig:exp3-plot}). In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 0.99 {[}0.634, 1.348{]}, p = \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.89 {[}0.636, 1.142{]}, p = \textless{} .001). Likewise the effect of convergence on competence is more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.80 {[}0.474, 1.13{]}, p = \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.82 {[}0.591, 1.056{]}, p = \textless{} .001).

\begin{figure}
\centering
\includegraphics{output/figures/exp3-plot.pdf}
\caption{\label{fig:exp3-plot}Results of Experiment 3, showing the distributions of accuracy and competence ratings by convergence and informational dependency.}
\end{figure}

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To this end, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction effect, indicating that pooled across independent and conflict of interest conditions, the effect of convergence had as smaller effect on competence than on accuracy (\(\hat{b}\) = -0.16 {[}-0.294, -0.028{]}, 0.018). Pooled across all conditions, participants reported higher perceived competence than accuracy (\(\hat{b}\) = 0.11 {[}0.025, 0.191{]}, 0.011).

Experiment 3 shows that, when the individuals making the estimates are systematically biased, participants put less weight on the convergence of their estimates to infer that the estimates are accurate, and that the individuals making them are competent.

\section{Experiment 4}\label{experiment-4}

In a second series of experiments, we test similar predictions to those of the previous experiments, but in a categorical choice context. The set-up is similar to that of Experiment 1, except that the outcomes seen by the participants are not numerical estimates, but choices made between a few options. An additional difference is that participants rate a focal informant, and not a group of informants. There were two reasons for this choice: First, accuracy is not on a continuum as in the first three experiments (an option was either correct or not), so forming an average across informants who chose different options was less sensible. Second, rating a focal individual allowed us to have a minority condition, which would not have been possible when providing an average rating for a group. Experiment 4 tests hypotheses that are analogous to those of Experiment 1:

\textbf{\emph{H1: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.}}

\textbf{\emph{H2: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.}}

\subsection{Methods}\label{methods-3}

\subsubsection{Participants}\label{participants-3}

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90\%. The power simulation for accuracy suggested that even for as few as 10 participants (the minimum sample size we simulated data for), we would have a power of close to 100\%. The simulation for competence suggested that we achieve statistical power of at least 90\% with a sample size of 30. Due to uncertainty about our assumptions and because it was within our budget, we recruited 100 participants, from the UK and via Prolific (50 female, 50; \(age_\text{mean}\): 37.32, \(age_\text{sd}\): 11.53, \(age_\text{median}\): 36).

\subsubsection{Procedure}\label{procedure-2}

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: ``To be able to understand the task, please read the following instructions carefully: Some people are playing games in which they have to select the correct answer among three answers. You will see the results of several of these games. Each game is different, with different solutions and involving different players. All players answer independently of each other. At first, you have no idea how competent each individual player is: they might be completely at chance, or be very good at the task. It's also possible that some players are really good while others are really bad. Some games might be difficult while others are easy. Your task will be to evaluate the performance of one of the players based on what everyone's answers are.'' They were then presented to the results of eight such games (Fig. \ref{fig:stimulus-exp4}). To assess perceived accuracy, we asked: ``What do you think is the probability of player 1 being correct?''. Participants answered with a slider on a scale from 0 to 100. To assess perceived competence, we asked participants: ``How competent do you think player 1 is in games like these?'' Participants answered on a 7-point Likert scale (from (1)``not competent at all'' to (2)``extremely competent'').

\begin{figure}
\centering
\includegraphics{output/figures/stimulus-exp4.pdf}
\caption{\label{fig:stimulus-exp4}One set of stimuli by level of convergence, in Experiment 4 (similar stimuli are used in Experiments 5 and 6). A full set of stimuli can be found in the ESM, section \ref{exp4}.}
\end{figure}

\subsubsection{Design}\label{design-3}

We manipulated convergence within participants, by varying the ratio of players choosing the same response as a focal player (i.e.~the one that participants evaluate). The levels of convergence are: (i) consensus, where all three players pick the same option {[}coded value = 3{]}; (ii) majority, where either the third or second player picks the same option as the first player {[}coded value = 2{]}; (iii) dissensus, where all three players pick different options {[}coded value = 1{]}; (iv) minority, where the second and third player pick the same option, but one that is different from the first player's choice {[}coded value = 0{]}. In our analysis, we treat convergence as a continuous variable, assigning the coded values in squared parenthesis here.

\subsubsection{Materials}\label{materials-3}

All participants saw all four conditions, with two stimuli per condition. Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).

\subsection{Results and discussion}\label{results-and-discussion-3}

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for participants.

As in the numerical setting, we found a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 16.84 {[}15.009, 18.668{]}, p = \textless{} .001; on a scale from 0 to 100) and competence (\(\hat{b}_{\text{Competence}}\) = 0.68 {[}0.578, 0.788{]}, p = \textless{} .001; on a scale from 1 to 7).

In the ESM (section \ref{exp4}), we show that compared to what the normative models would predict, participants underestimate the effect of convergence on both accuracy and competence, but especially on accuracy.

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To do so, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence). We find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy (\(\hat{b}\) = -0.10 {[}-0.136, -0.053{]}, \textless{} .001; units in standard deviations).

\section{Experiment 5}\label{experiment-5}

Experiment 5 is a conceptual replication of Experiment 3 in a categorical instead of a numerical case: are participants less likely to infer that more convergent estimates are more accurate, and the individuals who made them more competent, when the estimates are made by individuals with a conflict of interest pushing them to all provide a given answer, compared to when they are made by independent individuals? The independence condition of Experiment 5 also serves as a replication of Experiment 4, leading to the following hypotheses:

\textbf{\emph{H1a: Participants perceive an estimate of an independent informant as more accurate the more it converges with the estimates of other informants.}}

\textbf{\emph{H1b: Participants perceive an independent informant as more competent the more their estimate converges with the estimates of other informants.}}

\textbf{\emph{H2a: The effect of convergence on accuracy (H1a) is more positive in a context where informants are independent compared to when they are biased (i.e.~share a conflict of interest to pick a given answer).}}

\textbf{\emph{H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are biased (i.e.~share a conflict of interest to pick a given answer).}}

\subsection{Methods}\label{methods-4}

\subsubsection{Participants}\label{participants-4}

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. We ran two different power analyses, one for each outcome variable. We set the power threshold for both to 90\%.

The power simulation for accuracy suggested that for 80 participants, we would have a power of at least 90\% for the interaction effect. The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants would we also detect an effect of convergence. Due to uncertainty about our assumptions and because resources were available for a larger sample, we recruited 200 participants, in the UK and via Prolific (99 female, 100, 1 non-identified; \(age_\text{mean}\): 41.88, \(age_\text{sd}\): 13.94, \(age_\text{median}\): 39).

\subsubsection{Procedure}\label{procedure-3}

After providing their consent to participate in the study and passing an attention check, participants read the following introduction: ``We will show you three financial advisors who are giving recommendations on investment decisions. They can choose between three investment options. Their task is to recommend one. You will see several such situations. They are completely unrelated: it is different advisors evaluating different investments every time. At first you have no idea how competent the advisors are: they might be completely at chance, or be very good at the task. It's also possible that some are really good while others are really bad. Some tasks might be difficult while others are easy. Your task will be to evaluate the performance of one of the advisors based on what everyone's answers are.'' To assess perceptions of accuracy, we asked: ``What do you think is the probability of advisor 1 making the best investment recommendation?''. Participants answered with a slider on a scale from 0 to 100. To assess perceptions of competence, we asked: ``How competent do you think advisor 1 is regarding such investment recommendations?'' Participants answered on a 7-point Likert scale (from (1)``not competent at all'' to (7)``extremely competent'').

\subsubsection{Design}\label{design-4}

We manipulated convergence within participants, and conflict of interest between participants. In the conflict of interest condition, experts were introduced this way: ``The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations.'' Participants assigned to the independence condition read: ``The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.''

\subsubsection{Materials}\label{materials-4}

We used the same stimuli as in Experiment 4. Identical to Experiment 4, participants saw all four convergence conditions, with two stimuli (i.e.~expert predictions) per condition. Each participant therefore saw eight stimuli in total (4 convergence levels x 2 stimuli).

\subsection{Results and discussion}\label{results-and-discussion-4}

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants.

We find evidence for all four hypotheses (see Fig. \ref{fig:exp5-plot}). To test H1a and H1b, we use the same analyses as in Experiment 4, restricted on the independence condition, and replicate the results. We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 12.34 {[}10.362, 14.311{]}, p = \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.56 {[}0.459, 0.665{]}, p = \textless{} .001).

The second set of hypotheses targeted the interaction of informational dependency and convergence (Fig. \ref{fig:exp5-plot}). In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 3.01 {[}0.027, 5.988{]}, p = 0.048) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) 9.33 {[}7.232, 11.426{]}, p = \textless{} .001). Likewise, the effect of convergence on competence was more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.16 {[}0.014, 0.316{]}, p = 0.032) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.40 {[}0.291, 0.503{]}, p = \textless{} .001).

\begin{figure}
\centering
\includegraphics{output/figures/exp5-plot.pdf}
\caption{\label{fig:exp5-plot}Interaction of convergence and informational dependency.}
\end{figure}

In an exploratory, non-preregistered analysis, we tested whether the effect of convergence is larger on accuracy than on competence. To do so, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency. We find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy (\(\hat{b}\) = -0.08 {[}-0.113, -0.044{]}, \textless{} .001; units in standard deviations).

\section{Experiment 6}\label{experiment-6}

Experiment 6 is a replication and extension of Experiment 4 in which we test the effect of the number of choice options (three and ten, instead of only three). Our simulations suggested that, at least for some underlying population competence distributions, consensus should be more indicative of competence when there are more choice options, compared to fewer (see ESM, section \ref{simulations} and \ref{exp6}).

First, considering only the three options condition, we ran a direct replication of experiment 4. Second, following the results from our model, we predict that

\textbf{\emph{H1: The effect of convergence on accuracy (H1a) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.}}

\textbf{\emph{H2: The effect of convergence on competence (H1b) is more positive in a context when informants can choose among ten response options compared to when they can choose among only three.}}

\subsection{Methods}\label{methods-5}

\subsubsection{Participants}\label{participants-5}

We ran a power simulation to inform our choice of sample size. All assumptions and details on the procedure can be found on the OSF. We used previous experiments and estimates of our models to inform our choice of parameter values. We ran two different power analyses, one for each outcome variable. We set the power threshold for our experiment to 90\%. The power simulation for accuracy suggested that for 140 participants we would cross the power threshold of 90\% for the interaction effect (power = 0.928). The simulation for competence suggested that with 300 participants, we would detect an interaction with a power of 87\%. Due to budget constraints, we considered aiming for a sample of 300 participants as good enough, although slightly below our threshold. Due to two failed attention checks, our final sample consisted of 298 subjects, recruited, as in all experiments, in the UK and via Prolific (149 female, 149, 1 non-identified; \(age_\text{mean}\): 42.09, \(age_\text{sd}\): 13.06, \(age_\text{median}\): 40).

\subsubsection{Procedure}\label{procedure-4}

We used the same procedure as in Experiment 4, with the addition of one condition described below.

\subsubsection{Design}\label{design-5}

The number of choice options was manipulated between participants. Participants were randomly assigned to either see stimuli with three options (as in Experiment 4), or stimuli with ten options. Participants assigned to the ten options condition were divided into one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the three options condition, and another with increased range (see ESM, section \ref{exp6}). We found no differences between the two sub-conditions and collapsed them into a single ten options condition.

\subsubsection{Materials}\label{materials-5}

For the three options condition, we used the same stimuli as in Experiments 4 and 5. For the ten options condition, we created new sets of stimuli (see ESM, section \ref{exp6}). Identical to Experiments 4 and 5, participants saw all four convergence conditions, with two stimuli per condition. Each participant therefore saw eight stimuli in total.

\subsection{Results and discussion}\label{results-and-discussion-5}

To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for convergence for participants.

We replicate the results of experiment 4, but do not find evidence for an interaction between convergence and the number of choice options. To match the setting of experiment one, we reduced the sample to half of the participants, namely those who were assigned to the three options condition. On this reduced sample, we ran the exact same analyses as in experiment 4 and replicated the results. We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 15.68 {[}14.112, 17.246{]}, p = \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.65 {[}0.564, 0.736{]}, p = \textless{} .001). This finding also holds across the entire sample, pooling three and ten choice option observations (\(\hat{b}_{\text{Accuracy}}\) = 16.30 {[}15.124, 17.485{]}, \textless{} .001; \(\hat{b}_{\text{Competence}}\) = 0.68 {[}0.611, 0.739{]}, \textless{} .001).

We do not find evidence of an interaction, i.e.~evidence that the number of choice options changes the effect of convergence (\(\hat{b}_{\text{interaction, Accuracy}}\) = 1.25 {[}-1.11, 3.613{]}, 0.298; \(\hat{b}_{\text{interaction, Competence}}\) = 0.05 {[}-0.078, 0.178{]}, 0.442).

We tested whether the effect of convergence is larger on accuracy than on competence in an exploratory, non-preregistered analysis. To this end, we first standardized both outcome scores to account for the different scales. We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for the number of choice options. In line with all previous experiments, we find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy (\(\hat{b}\) = -0.05 {[}-0.075, -0.023{]}, \textless{} .001).

Experiment 6 replicates Experiments 3 and 4 (independence condition), but suggests that the number of options the informants choose from does not powerfully affect participants' estimates of the informants' accuracy or competence. In the ESM, section \ref{exp6}, we show that, compared to our model, participants underestimate the effect of the number of choice options.

\section{General discussion}\label{general-discussion}

Using both analytical arguments and simulations, we have shown that, under a wide range of parameters, more convergent answers tend to be more accurate, and to have been made by more competent informants.

In two experiments (Experiment 1, and independence condition of Experiment 3), we find that participants presented with a set of more (rather than less) convergent numerical estimates find the estimates more accurate, and the individuals making the estimates more competent, thus drawing normatively justified inferences. Experiment 2 suggests that participants do not think that a discussion between the individuals makes their convergence less indicative of accuracy or their competence. By contrast, Experiment 3 reveals that, when the individuals making the estimates are systematically biased by a conflict of interest, participants put less weight on the convergence of their estimates.

Similar results are obtained in a categorical choice context, in which participants see the answers of individuals made within a limited set of options. Experiments 4, 5 (independence condition), and 6 show that, the more the answers converge, the more participants believe them to be accurate and the individuals who made to be competent, again drawing normatively justified inferences. Experiment 5 shows that these inferences are weakened when the convergence can be explained by a conflict of interest (as in Experiment 3). Experiment 6 fails to find an effect of the number of options.

We also observe that, in line with our simulations, participants draw stronger inferences from convergence to accuracy than to competence in five of the six experiments.

On the whole, participants thus appear to draw normatively justified inferences: 1. They infer that more convergent answers are more accurate; 2. They infer that more convergent answers are coming from more competent informants; 3. The inference on accuracy tends to be stronger than the inference on competence ; 4. Both the inference on accuracy and on the inference on competence are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other. The only exception appears to be that participants do not take into account the number of options in categorical choices scenarios.

\section{Conclusion}\label{conclusion}

When people see that others agree with each other, they tend to believe that they are right. This inference has been evidenced in several experiments, both for numerical estimates (Budescu \& Rantilla, 2000; Budescu et al., 2003; Budescu \& Yu, 2007; e.g. Molleman et al., 2020; Yaniv et al., 2009), and for categorical choices (e.g., Morgan et al., 2012; for review, see H. Mercier \& Morin, 2019). However, these experiments do not test whether this inference of accuracy of the information extends to an inference of competence of the informants. Moreover, by their design, participants arguably assumed a degree of competence among the informants. For instance, when children are confronted with several individuals who agree on how to name a novel object (e.g. Corriveau et al., 2009), they can assume that these (adult) individuals tend to know what the names of objects are. If a certain competence of the informants is assumed, then well-known results from the literature on judgment aggregation--the wisdom of crowds--show that the average opinion of a set of individuals is, in a wide range of circumstances, more likely to be accurate than that of a single individual (e.g. Larrick \& Soll, 2006).

Here, we assumed no prior knowledge of individual competence, asking the question: if we see informants, whose competence is unknown, converge on an answer, is it rational to infer that this answer is more likely to be correct, and that the informants are likely to be competent? We have shown that the answer is yes on both counts--assuming there is no systematic bias among the informants. An analytical argument and a series of simulations revealed that, for both the numerical choice context and the categorical choice context, the more individuals agree on an answer, the more likely the answer is correct, and the more likely the individuals are competent, with the former effect being stronger than the latter. Moreover, this is true for a wide range of assumed population distributions of competence. In a series of experiments, we have shown that participants (UK) draw these inferences, but that they do so less when there is reason to assume a bias among informants.

The results--both simulations and experiments--are a novel contribution to the wisdom of crowds literature. In this literature--in particular that relying on the Condorcet Jury Theorem--a degree of competence is assumed in the individuals providing some answers. From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate. Here, we have shown that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be a valid cue when determining who tends to be an expert. This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert (Hahn, Merdes, \& Sydow, 2018; Nguyen, 2020).

The present experiments show that people draw inferences from convergence to accuracy and to competence, but then do not precisely show how they do it. As suggested in the introduction, it is plausible that participants use the combination of two heuristics: one leading them from convergence to accuracy, and one from accuracy to competence. This two step process is coherent with the observation that the inference from convergence to accuracy (one step) is stronger than that from convergence to competence (two steps). However, our results are also compatible with participants performing two inferences from convergence, one to accuracy and one to competence, the latter being weaker because the inferences roughly follow the normative model. Our results are coherent with work on other mechanisms of epistemic vigilance that process the combination of several informants' opinions, showing that these mechanisms rely on sensible cues, but also systematically fail to take some subtle cues into account (H. Mercier \& Morin, 2019).

A most prominent context in which the inferences uncovered here might play a role is that of science. Much of science is counterintuitive, and most people do not have the background knowledge to evaluate most scientific evidence. However, science is, arguably, the institution in which individuals end up converging the most in their opinions (on consensus being the defining trait of science by contrast with other intellectual enterprises, see Collins, 2002). For instance, scientists within many disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA. This represents an incredible degree of convergence. When people hear that scientists have measured the distance between the solar system and the center of the galaxy, if they assume that there is a broad agreement within the relevant experts, this should lead them to infer that this measure is accurate, and that the scientists who made it are competent. Experiments have already shown that increasing the degree of perceived consensus among scientists tends to increase acceptance of the consensual belief (Van Stekelenburg, Schaap, Veling, Van 'T Riet, \& Buijzen, 2022), but it hasn't been shown yet that the degree of consensus also affects the perceived competence of scientists.

In the case of science, the relationship between convergence and accuracy is broadly justified. However, at some points of history, there has been broad agreement on misbeliefs, such as when Christian theologians had calculated that the Earth was approximately six thousand years old. To the extent that people were aware of this broad agreement, and believed the theologians to have reached it independently of each other, this might have not only fostered acceptance of this estimate of the age of the Earth, but also a perception of the theologians as competent.

The current study has a number of limitations. In our simulations, we assume agents to be independent and unbiased. Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency (Ladha, 1992), more robust simulations would show that--while still assuming no systematic bias--our results hold even when agents influence each others' answers. Regarding our experiments, if the very abstract materials allow us to remove most of the priors that participants might have, they might also reduce the ecological validity of the results. Although the main results replicate well across our experiments, and we can thus be reasonably certain of their robustness, it's not clear how much they can be generalized. Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from--here, UK citizens (Coppock, 2019). However, we do not know whether they would generalize to other cultures.

These limitations could be overcome by replicating the present results in different cultures, using more ecologically valid stimuli. For instance, it would be interesting to test whether the inference described here, from convergence to competence, might be partly responsible for the fact that people tend to believe scientists to be competent (Cologna et al., 2024). Finally, future studies could also attempt to systematically model cases of informational dependencies more subtle than those used in Experiments 3 and 5, to derive and test normative predictions.

\subsubsection{Data availability}\label{data-availability}

Data for all experiments and the simulations is available on the OSF project page (\url{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}).

\subsubsection{Code availability}\label{code-availability}

The code used to create all results (including tables and figures) of this manuscript is also available on the OSF project page (\url{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}).

\subsubsection{Competing interest}\label{competing-interest}

The authors declare having no competing interests.

\FloatBarrier

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-aschStudiesIndependenceConformity1956}
Asch, S. E. (1956). Studies of independence and conformity: I. A minority of one against a unanimous majority. \emph{Psychological Monographs: General and Applied}, \emph{70}(9), 1--70. \url{https://doi.org/10.1037/h0093718}

\bibitem[\citeproctext]{ref-bednarikEffectivenessImperfectWeighting2015a}
Bednarik, P., \& Schultze, T. (2015). The effectiveness of imperfect weighting in advice taking. \emph{Judgment and Decision Making}, \emph{10}(3), 265--276. \url{https://doi.org/10.1017/S1930297500004666}

\bibitem[\citeproctext]{ref-bernardChildrenWeighNumber2015a}
Bernard, S., Harris, P., Terrier, N., \& Clément, F. (2015). Children weigh the number of informants and perceptual uncertainty when identifying objects. \emph{Journal of Experimental Child Psychology}, \emph{136}, 70--81. \url{https://doi.org/10.1016/j.jecp.2015.03.009}

\bibitem[\citeproctext]{ref-bernardFourSixYearOldChildren2015}
Bernard, S., Proust, J., \& Clément, F. (2015). Four- to Six-Year-Old Children's Sensitivity to Reliability Versus Consensus in the Endorsement of Object Labels. \emph{Child Development}, \emph{86}(4), 1112--1124. \url{https://doi.org/10.1111/cdev.12366}

\bibitem[\citeproctext]{ref-budescuConfidenceAggregationExpert2000}
Budescu, D. V., \& Rantilla, A. K. (2000). Confidence in aggregation of expert opinions. \emph{Acta Psychologica}, \emph{104}(3), 371--398. \url{https://doi.org/10.1016/S0001-6918(00)00037-8}

\bibitem[\citeproctext]{ref-budescuEffectsAsymmetryAdvisors2003}
Budescu, D. V., Rantilla, A. K., Yu, H.-T., \& Karelitz, T. M. (2003). The effects of asymmetry among advisors on the aggregation of their opinions. \emph{Organizational Behavior and Human Decision Processes}, \emph{90}(1), 178--194. \url{https://doi.org/10.1016/S0749-5978(02)00516-2}

\bibitem[\citeproctext]{ref-budescuAggregationOpinionsBased2007}
Budescu, D. V., \& Yu, H.-T. (2007). Aggregation of opinions based on correlated cues and advisors. \emph{Journal of Behavioral Decision Making}, \emph{20}(2), 153--177. \url{https://doi.org/10.1002/bdm.547}

\bibitem[\citeproctext]{ref-chenChildrenTrustConsensus2013}
Chen, E. E., Corriveau, K. H., \& Harris, P. L. (2013). Children Trust a Consensus Composed of Outgroup Members-But Do Not Retain That Trust. \emph{Child Development}, \emph{84}(1), 269--282. \url{https://doi.org/10.1111/j.1467-8624.2012.01850.x}

\bibitem[\citeproctext]{ref-collinsSociologyPhilosophiesGlobal2002}
Collins, R. (2002). \emph{The sociology of philosophies: a global theory of intellectual change} (4. print., 1. Harvard Univ. Pr. paperback ed., 2000). Cambridge, Mass. London: Belknap Press of Harvard Univ. Press.

\bibitem[\citeproctext]{ref-colognaTrustScientistsTheir2024}
Cologna, V., Mede, N. G., Berger, S., Besley, J., Brick, C., Joubert, M., \ldots{} Linden, D. S. van der. (2024). \emph{Trust in scientists and their role in society across 67 countries}. \url{https://doi.org/10.31219/osf.io/6ay7s}

\bibitem[\citeproctext]{ref-coppockGeneralizingSurveyExperiments2019}
Coppock, A. (2019). Generalizing from Survey Experiments Conducted on Mechanical Turk: A Replication Approach. \emph{Political Science Research and Methods}, \emph{7}(3), 613--628. \url{https://doi.org/10.1017/psrm.2018.10}

\bibitem[\citeproctext]{ref-corriveauGoingFlowPreschoolers2009}
Corriveau, K. H., Fusaro, M., \& Harris, P. L. (2009). Going With the Flow: Preschoolers Prefer Nondissenters as Informants. \emph{Psychological Science}, \emph{20}(3), 372--377. \url{https://doi.org/10.1111/j.1467-9280.2009.02291.x}

\bibitem[\citeproctext]{ref-crutchfieldConformityCharacter1955}
Crutchfield, R. S. (1955). Conformity and character. \emph{American Psychologist}, \emph{10}(5), 191--198. \url{https://doi.org/10.1037/h0040237}

\bibitem[\citeproctext]{ref-decondorcetEssaiApplicationAnalyse2014}
De Condorcet, N. (2014). \emph{Essai sur l'application de l'analyse à la probabilité des décisions rendues à la pluralité des voix}. Cambridge University Press.

\bibitem[\citeproctext]{ref-desaiGettingSourceIllusion2022}
Desai, S. C., Xie, B., \& Hayes, B. K. (2022). Getting to the source of the illusion of consensus. \emph{Cognition}, \emph{223}, 105023. \url{https://doi.org/10.1016/j.cognition.2022.105023}

\bibitem[\citeproctext]{ref-dietrichEpistemicDemocracyDefensible2013}
Dietrich, F., \& Spiekermann, K. (2013). Epistemic Democracy with Defensible Premises. \emph{Economics and Philosophy}, \emph{29}(1), 87--120. \url{https://doi.org/10.1017/S0266267113000096}

\bibitem[\citeproctext]{ref-dietrichDeliberationWisdomCrowds2024}
Dietrich, F., \& Spiekermann, K. (2024). Deliberation and the wisdom of crowds. \emph{Economic Theory}. \url{https://doi.org/10.1007/s00199-024-01595-4}

\bibitem[\citeproctext]{ref-einavThinkingThemselvesEffect2018}
Einav, S. (2018). Thinking for themselves? The effect of informant independence on children's endorsement of testimony from a consensus. \emph{Social Development}, \emph{27}(1), 73--86. \url{https://doi.org/10.1111/sode.12264}

\bibitem[\citeproctext]{ref-fusaroChildrenAssessInformant2008}
Fusaro, M., \& Harris, P. L. (2008). Children assess informant reliability using bystanders{'} non-verbal cues. \emph{Developmental Science}, \emph{11}(5), 771--777. \url{https://doi.org/10.1111/j.1467-7687.2008.00728.x}

\bibitem[\citeproctext]{ref-galtonVoxPopuli1907}
Galton, F. (1907). Vox Populi. \emph{Nature}, \emph{75}(1949), 450--451. \url{https://doi.org/10.1038/075450a0}

\bibitem[\citeproctext]{ref-hahnHowGoodYour2018}
Hahn, U., Merdes, C., \& Sydow, M. von. (2018). How Good Is Your Evidence and How Would You Know? \emph{Topics in Cognitive Science}, \emph{10}(4), 660--678. \url{https://doi.org/10.1111/tops.12374}

\bibitem[\citeproctext]{ref-harkinsInformationUtilityMultiple1987}
Harkins, S. G., \& Petty, R. E. (1987). Information utility and the multiple source effect. \emph{Journal of Personality and Social Psychology}, \emph{52}(2), 260.

\bibitem[\citeproctext]{ref-harriesCombiningAdviceWeight2004}
Harries, C., Yaniv, I., \& Harvey, N. (2004). Combining advice: the weight of a dissenting opinion in the consensus. \emph{Journal of Behavioral Decision Making}, \emph{17}(5), 333--348. \url{https://doi.org/10.1002/bdm.474}

\bibitem[\citeproctext]{ref-harveyTakingAdviceAccepting1997}
Harvey, N., \& Fischer, I. (1997). Taking Advice: Accepting Help, Improving Judgment, and Sharing Responsibility. \emph{Organizational Behavior and Human Decision Processes}, \emph{70}(2), 117--133. \url{https://doi.org/10.1006/obhd.1997.2697}

\bibitem[\citeproctext]{ref-hastieRobustBeautyMajority2005}
Hastie, R., \& Kameda, T. (2005). The Robust Beauty of Majority Rules in Group Decisions. \emph{Psychological Review}, \emph{112}(2), 494--508. \url{https://doi.org/10.1037/0033-295X.112.2.494}

\bibitem[\citeproctext]{ref-herrmannStickScriptEffect2013}
Herrmann, P. A., Legare, C. H., Harris, P. L., \& Whitehouse, H. (2013). Stick to the script: The effect of witnessing multiple actors on children{'}s imitation. \emph{Cognition}, \emph{129}(3), 536--543. \url{https://doi.org/10.1016/j.cognition.2013.08.010}

\bibitem[\citeproctext]{ref-hessPsychologicalAdaptationsAssessing2006}
Hess, N. H., \& Hagen, E. H. (2006). Psychological adaptations for assessing gossip veracity. \emph{Human Nature}, \emph{17}(3), 337--354. \url{https://doi.org/10.1007/s12110-006-1013-z}

\bibitem[\citeproctext]{ref-jaylesHowSocialInformation2017a}
Jayles, B., Kim, H., Escobedo, R., Cezera, S., Blanchet, A., Kameda, T., \ldots{} Theraulaz, G. (2017). How social information can improve estimation accuracy in human groups. \emph{Proceedings of the National Academy of Sciences}, \emph{114}(47), 12620--12625. \url{https://doi.org/10.1073/pnas.1703695114}

\bibitem[\citeproctext]{ref-kammerSystematicReviewEmpirical2023}
Kämmer, J. E., Choshen-Hillel, S., Müller-Trede, J., Black, S. L., \& Weibler, J. (2023). A systematic review of empirical studies on advice-based decisions in behavioral and organizational research. \emph{Decision}, \emph{10}(2), 107--137. \url{https://doi.org/10.1037/dec0000199}

\bibitem[\citeproctext]{ref-koenigTrustTestimonyChildren2004}
Koenig, M. A., Clément, F., \& Harris, P. L. (2004). Trust in Testimony: Children's Use of True and False Statements. \emph{Psychological Science}, \emph{15}(10), 694--698. \url{https://doi.org/10.1111/j.0956-7976.2004.00742.x}

\bibitem[\citeproctext]{ref-ladhaCondorcetJuryTheorem1992}
Ladha, K. K. (1992). The Condorcet Jury Theorem, Free Speech, and Correlated Votes. \emph{American Journal of Political Science}, \emph{36}(3), 617. \url{https://doi.org/10.2307/2111584}

\bibitem[\citeproctext]{ref-larrickIntuitionsCombiningOpinions2006}
Larrick, R. P., \& Soll, J. B. (2006). Intuitions About Combining Opinions: Misappreciation of the Averaging Principle. \emph{Management Science}, \emph{52}(1), 111--127. \url{https://doi.org/10.1287/mnsc.1050.0459}

\bibitem[\citeproctext]{ref-lopesSocialValidationEveryday2007}
Lopes, D., Vala, J., \& Garcia-Marques, L. (2007). Social validation of everyday knowledge: Heterogeneity and consensus functionality. \emph{Group Dynamics: Theory, Research, and Practice}, \emph{11}(3), 223--239. \url{https://doi.org/10.1037/1089-2699.11.3.223}

\bibitem[\citeproctext]{ref-mannesAreWeWise2009}
Mannes, A. E. (2009). Are We Wise About the Wisdom of Crowds? The Use of Group Judgments in Belief Revision. \emph{Management Science}, \emph{55}(8), 1267--1279. \url{https://doi.org/10.1287/mnsc.1090.1031}

\bibitem[\citeproctext]{ref-mercierArgumentativeTheoryPredictions2016a}
Mercier, H. (2016). The Argumentative Theory: Predictions and Empirical Evidence. \emph{Trends in Cognitive Sciences}, \emph{20}(9), 689--700. \url{https://doi.org/10.1016/j.tics.2016.07.001}

\bibitem[\citeproctext]{ref-mercierNotBornYesterday2020}
Mercier, Hugo. (2020). \emph{Not born yesterday: the science of who we trust and what we believe}. Retrieved from \url{https://doi.org/10.1515/9780691198842}

\bibitem[\citeproctext]{ref-mercierDoesDiscussionMake2022}
Mercier, H., \& Claidière, N. (2022). Does discussion make crowds any wiser? \emph{Cognition}, \emph{222}, 104912. \url{https://doi.org/10.1016/j.cognition.2021.104912}

\bibitem[\citeproctext]{ref-mercierUtilizingSimpleCues2019}
Mercier, Hugo, \& Miton, H. (2019). Utilizing simple cues to informational dependency. \emph{Evolution and Human Behavior}, \emph{40}(3), 301--314. \url{https://doi.org/10.1016/j.evolhumbehav.2019.01.001}

\bibitem[\citeproctext]{ref-mercierMajorityRulesHow2019}
Mercier, H., \& Morin, O. (2019). Majority rules: how good are we at aggregating convergent opinions? \emph{Evolutionary Human Sciences}, \emph{1}, e6. \url{https://doi.org/10.1017/ehs.2019.6}

\bibitem[\citeproctext]{ref-mollemanStrategiesIntegratingDisparate2020}
Molleman, L., Tump, A. N., Gradassi, A., Herzog, S., Jayles, B., Kurvers, R. H. J. M., \& Bos, W. van den. (2020). Strategies for integrating disparate social information. \emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{287}(1939), 20202413. \url{https://doi.org/10.1098/rspb.2020.2413}

\bibitem[\citeproctext]{ref-morganDevelopmentAdaptiveConformity2015}
Morgan, T. J. H., Laland, K. N., \& Harris, P. L. (2015). The development of adaptive conformity in young children: effects of uncertainty and consensus. \emph{Developmental Science}, \emph{18}(4), 511--524. \url{https://doi.org/10.1111/desc.12231}

\bibitem[\citeproctext]{ref-morganEvolutionaryBasisHuman2012}
Morgan, T. J. H., Rendell, L. E., Ehn, M., Hoppitt, W., \& Laland, K. N. (2012). The evolutionary basis of human social learning. \emph{Proceedings of the Royal Society B: Biological Sciences}, \emph{279}(1729), 653--662. \url{https://doi.org/10.1098/rspb.2011.1172}

\bibitem[\citeproctext]{ref-nguyenCognitiveIslandsRunaway2020}
Nguyen, C. T. (2020). Cognitive islands and runaway echo chambers: problems for epistemic dependence on experts. \emph{Synthese}, \emph{197}(7), 2803--2821. \url{https://doi.org/10.1007/s11229-018-1692-0}

\bibitem[\citeproctext]{ref-pillowEarlyUnderstandingPerception1989}
Pillow, B. H. (1989). Early understanding of perception as a source of knowledge. \emph{Journal of Experimental Child Psychology}, \emph{47}(1), 116--129. \url{https://doi.org/10.1016/0022-0965(89)90066-0}

\bibitem[\citeproctext]{ref-richardsonThinkingTakesTime2022}
Richardson, E., \& Keil, F. C. (2022). Thinking takes time: Children use agents' response times to infer the source, quality, and complexity of their knowledge. \emph{Cognition}, \emph{224}, 105073. \url{https://doi.org/10.1016/j.cognition.2022.105073}

\bibitem[\citeproctext]{ref-romeijnLearningJurorCompetence2011}
Romeijn, J.-W., \& Atkinson, D. (2011). Learning juror competence: a generalized Condorcet Jury Theorem. \emph{Politics, Philosophy \& Economics}, \emph{10}(3), 237--262. \url{https://doi.org/10.1177/1470594X10372317}

\bibitem[\citeproctext]{ref-smithAnimalSignals2003}
Smith, J. M., \& Harper, D. (2003). \emph{Animal signals}. Oxford University Press.

\bibitem[\citeproctext]{ref-sollStrategiesRevisingJudgment2009}
Soll, J. B., \& Larrick, R. P. (2009). Strategies for revising judgment: How (and how well) people use others{'} opinions. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{35}(3), 780--805. \url{https://doi.org/10.1037/a0015145}

\bibitem[\citeproctext]{ref-sperberEpistemicVigilance2010}
Sperber, D., Clément, F., Heintz, C., Mascaro, O., Mercier, H., Origgi, G., \& Wilson, D. (2010). Epistemic vigilance. \emph{Mind \& Language}, \emph{25}(4), 359--393.

\bibitem[\citeproctext]{ref-vanstekelenburgScientificConsensusCommunicationContested2022}
Van Stekelenburg, A., Schaap, G., Veling, H., Van 'T Riet, J., \& Buijzen, M. (2022). Scientific-Consensus Communication About Contested Science: A Preregistered Meta-Analysis. \emph{Psychological Science}, \emph{33}(12), 1989--2008. \url{https://doi.org/10.1177/09567976221083219}

\bibitem[\citeproctext]{ref-xieSensitivityEvidentialDependencies2022}
Xie, B., \& Hayes, B. (2022). Sensitivity to Evidential Dependencies in Judgments Under Uncertainty. \emph{Cognitive Science}, \emph{46}(5). \url{https://doi.org/10.1111/cogs.13144}

\bibitem[\citeproctext]{ref-yanivWeightingTrimmingHeuristics1997}
Yaniv, I. (1997). \emph{Weighting and Trimming: Heuristics for Aggregating Judgments under Uncertainty}. 13.

\bibitem[\citeproctext]{ref-yanivReceivingOtherPeople2004}
Yaniv, I. (2004). Receiving other people{'}s advice: Influence and benefit. \emph{Organizational Behavior and Human Decision Processes}, \emph{93}(1), 1--13. \url{https://doi.org/10.1016/j.obhdp.2003.08.002}

\bibitem[\citeproctext]{ref-yanivSpuriousConsensusOpinion2009}
Yaniv, I., Choshen-Hillel, S., \& Milyavsky, M. (2009). Spurious consensus and opinion revision: Why might people be more confident in their less accurate judgments? \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{35}(2), 558--563. \url{https://doi.org/10.1037/a0014589}

\bibitem[\citeproctext]{ref-yanivAdviceTakingDecision2000}
Yaniv, I., \& Kleinberger, E. (2000). Advice Taking in Decision Making: Egocentric Discounting and Reputation Formation. \emph{Organizational Behavior and Human Decision Processes}, \emph{83}(2), 260--281. \url{https://doi.org/10.1006/obhd.2000.2909}

\bibitem[\citeproctext]{ref-yinSmartlyFollowingOthers2024}
Yin, J., Xu, Z., Lin, J., Zhou, W., \& Guo, X. (2024). Smartly following others: Majority influence depends on how the majority behavior is formed. \emph{Journal of Experimental Social Psychology}, \emph{115}, 104644. \url{https://doi.org/10.1016/j.jesp.2024.104644}

\bibitem[\citeproctext]{ref-yousifIllusionConsensusFailure2019}
Yousif, S. R., Aboody, R., \& Keil, F. C. (2019). The Illusion of Consensus: A Failure to Distinguish Between True and False Consensus. \emph{Psychological Science}, \emph{30}(8), 1195--1204. \url{https://doi.org/10.1177/0956797619856844}

\end{CSLReferences}

\newpage

\appendix

\section{Analytical argument}\label{analytical}

\subsection{Categorical choice}\label{categorical-choice}

\subsubsection{Inferring population average competence}\label{inferring-population-average-competence}

There are \(n+1\) individuals, who choose among \(m+1\) answers. There is one good answer and \(m\) wrong ones. We denote \(\bar p\) the probability that a random individual is right. The probability to be wrong is \(1-\bar p\), and the probability to choose a particular wrong answer is \(\frac{1-\bar p}{m}\). We denote this last probability \(\bar q\) (it is common in probabilities to use \(q=1-p\), here note that it is not the case).

We assume that the population is quite large, so that it is very likely to be the biggest group of answers. We want to estimate the average competence \(\bar p\) from the size of the majority \(K\). To simplify, we concentrate on situations where either \(\bar p\) is quite large compared to \(\bar q\), or the population quite large. In this situation, it is almost certain that the biggest group is made of individuals who chose the right answer and a proportion \(\bar p\) of the population should, on average, choose this answer. The size \(\frac{K}{n+1}\) should thus be a good estimator of \(\bar p\).

Figure \ref{fig:biggestgroup} illustrates this logic. When the population is competent enough or large enough (e.g.~when \(n=100\)), the average size of the biggest group closely aligns with the average competence. When \(n=10\) and average competence is low, it is frequent that some individuals converge by chance on another answer, nd the size of the majority is is therefore on average larger than \(\bar p\), hence an overestimate of its members competence.

\begin{figure}

\includegraphics[width=0.75\linewidth]{figures/Benoit_biggest_group} \hfill{}

\caption{Expected size of the majority depending on average competence, with \(n=10\) (blue) and \(n=100\) (orange).}\label{fig:biggestgroup}
\end{figure}

\subsubsection{Inferring within population competence}\label{inferring-within-population-competence}

Here, we focus on one individual and we aim to infer his personal level of competence \(p\). We observe the number \(K\) of individuals who agree with him. Either the focus individual is right (which happens with probability \(p\)) and \(K\) follows a binomial distribution of parameters \(n\) and \(\bar p\); or the individual is wrong (which happens with probability \(1-p\)) and then \(K\) follows a binomial distribution of parameters \(n\) and \(\bar q\). Thus, \(K\) follows a mixture distribution of different parameters, with weights \(p\) and \(1-p\).

Using the mixture distributions property, we can compute the mean of \(K\):
\[
E(K) = n(p \bar p + (1-p)\bar q) = n(p (\bar p - \bar q) + \bar q)
\]

Clearly, \(E(K)\) increases with \(p\) if and only if \(\bar p - \bar q > 0\). In other words, if the population is better than chance, then we expect that the more competent an individual is, the more people agree with him.

But now, what exactly can we infer from \(p\) by observing \(K\)? To answer that, we need to assume a distribution of competence levels in the population, that we will also use as a prior. We assume that in the competence levels are distributed as a beta distribution, of parameters \(\alpha\) and \(\beta\). Their density is \(f(x) = x^\alpha(1-x)^\beta\). Then, \(\bar p = \frac{\alpha}{\alpha+\beta}\), the mean of our beta distribution.

Using Bayes formula, we can write the posterior distribution of \(p\):
\[
\Pi(x|K=k) \propto f(x)P(K=k|p=x) = f(x)(x\bar p^k(1-\bar p)^{n-k} + (1-x)\bar q^k (1-\bar q)^{n-k})
\]

\[
= x^{\alpha+1}(1-x)^\beta[\bar p^k(1-\bar p)^{n-k}] + x^{\alpha}(1-x)^{\beta+1} [\bar q^k (1-\bar q)^{n-k}]
\]
We recognise a mixture distribution between two beta distributions, with parameters \(\alpha+1\) and \(\beta\) for the first one, \(\alpha\) and \(\beta + 1\) for the second. The first one has a bigger mean than our original Beta(\(\alpha\),\(\beta\)) distribution, the second a lower one (see Fig. \ref{fig:posterior}). In a way, the bayesian updating shifts the distribution to the right or to the left, depending on \(K\).

\begin{figure}

\includegraphics[width=0.75\linewidth]{figures/Benoit_posterior} \hfill{}

\caption{Prior and boundary posterior distributions. These distributions are the limit distributions when the individual is either part of a consensus or a dissensus, and when \(\bar p \gg \bar q\) or \(n\) is large. Here, \(\alpha=\beta=2\) and we do not need to define the other parameters. As we explain below, the prior distribution determines the range between the possible posterior distributions.}\label{fig:posterior}
\end{figure}

The weightings (\(\bar p^k(1-\bar p)^{n-k}\) and \(\bar q^k (1-\bar q)^{n-k}\) correspond to the probability that \(k\) individuals choose the right answer, and a particular wrong answer, respectively (we took out the binomial coefficients, who are equal for both answer types). Since \(\bar p > \bar q\), the former is much stronger for large \(k\), and the latter much stronger for low \(k\). For \(k=n\) (i.e.~when the focus individual is part of a consensus), the ratio of weighting if \((p/q)^n\), which is very large when \(n\) is large or when \(\bar p \gg \bar q\). In this case, for high \(k\), the posterior distribution is well approximated by a Beta(\(\alpha+1\),\(\beta\)). Conversely, when \(k\) is close to zero (and when the other conditions are present), the posterior distribution is well approximated by a Beta(\(\alpha\),\(\beta+1\)).

We can plot the mean of the posterior distribution, which can be interpreted as the average estimation we can make of the individual competence, depending on \(K\) (Fig. \ref{fig:competence-k}). For low \(n\) and \(m\) (the main experiment used \(n=m=2\), the function is not too steep. For higher values, it gets sigmoidal, or even a jump function.

\begin{figure}

\includegraphics[width=0.75\linewidth]{figures/Benoit_competence_k} \hfill{}

\caption{Mean of the posterior competence distribution, depending on the number of individuals agreeing with the focus individual. Here, \(n=m=10\).}\label{fig:competence-k}
\end{figure}

How much we learn from the observation depends crucially on \(\alpha\) and \(\beta\), that is, on the prior distribution of competence in the population. The range of possible mean estimates is \(\frac{1}{\alpha+\beta + 1}\), it is larger when \(\alpha\) and \(\beta\) are low.

For instance, let us set \(\alpha=\beta\) (which implies \(\bar p = \frac 12\)). Now, we vary the two shape parameters from close to 0 (beta distributions need positive shape parameters) to a large number. This shifts the distribution from a bimodal one (with \(\alpha=\beta \approx 0\), the distribution is perfectly bimodal, with half of the population always right and the other half always right) to a sharp unimodal distribution (when \(\alpha\) and \(\beta\) are large, the whole population has basically a competence of .5). The more bimodal the distribution, the larger the range of possible estimates (Fig. \ref{fig:update}). In the limit, when \(\alpha\) and \(\beta\) are close to 0, we can infer that an individual is either perfectly competent or perfectly incompetent, depending on whether they are part of a consensus or alone in their opinion. On the contrary, when \(\alpha\) and \(\beta\) are large, we do not change our estimate at all as the distribution is concentrated around .5. Intuitively, this makes sense: If we assume everyone is equally (in)competent, with no exception, we wouldn't infer anything from observing a convergence.

\begin{figure}

\includegraphics[width=0.75\linewidth]{figures/Benoit_update} \hfill{}

\caption{Range of possible estimates (difference between the minimum and the maximum estimate), depending on the shape of the distribution (here, we assume that \(\alpha=\beta\)).}\label{fig:update}
\end{figure}

\subsubsection{Inferring accuracy}\label{inferring-accuracy}

Now, instead of inferring competence, we want to infer accuracy, i.e.~the probability that an individual answer is correct.

Again, using Bayes formula, we have:

\[
p(right|K=k) = \frac{p(K=k|right)p(right)}{p(K=k)} = \frac{\bar p (\bar p^{k-1} (1-\bar p)^{n-k})}{((\bar p^k (1-\bar p)^{n-k}) + (\bar q^k (1-\bar q)^{n-k}))} 
\]

\[
= \frac{1}{1 +  (\bar p/\bar q)^k \left( \frac{1-\bar q}{1-\bar p}\right)^{n-k} }
\]

Compared to competence, it is clear (Figure \ref{fig:accuracy}) that accuracy can be inferred much more strongly.

\begin{figure}

\includegraphics[width=0.75\linewidth]{figures/Benoit_accuracy} \hfill{}

\caption{Estimated competence and accuracy depending on the degree of consensus. Here, we assume that \(\alpha=1\), \(n=20\) and \(m=4\). \(\beta\) is inferred from \(\alpha\) and the majority size.}\label{fig:accuracy}
\end{figure}

\subsection{Continuous choices}\label{continuous-choices}

A group of \(N\) individuals makes a prediction. The right answer is \(\theta\), we assume their answers \(x_i, i \in 1,…,N\) are normally distributed around \(\theta\), with a variance \(\sigma^2\) reflecting the group incompetence (so the lower \(\sigma^2\), the more the group can be said to be competent).

We estimate the degree of divergence through the unbiased sample variance \(\hat \sigma^2\), defined as the average squared distance to the sample mean \(\bar x =\frac 1 n \sum_{i=1}^n x_i\):

\[\hat \sigma^2 = \frac 1 {n-1} \sum_{i=1}^n (x_i - \bar x)^2\]

It is well known that \(\hat \sigma^2\) (when coupled with the sample mean) is a ``sufficient'' estimator of \(\sigma^2\), that is, it contains all the available information about \(\sigma^2\). It follows, from the Lehmann-Scheffé theorem, that it is also the minimum-variance unbiased estimator, that is, the most precise way to estimate \(\sigma^2\) while staying unbiased.

\clearpage
\newpage

\section{Simulations}\label{simulations}

Are we justified in inferring competence and accuracy from convergence? To complement our analytical argument, we ran simulations, intended to mirror our experimental setup. We find that--under certain conditions--more convergent groups indeed tend to be more competent and accurate. In this appendix, we describe these simulations in detail.

\subsection{Numerical choice context}\label{numerical-choice-context}

When several people estimate a quantity (numeric scenario), their convergence can be measured for example by the empirical variance. The closer the estimates, i.e.~the smaller the empirical variance, the greater convergence. This measure is at the group level.

To provide a normative answer, we ran simulations for a scenario in which individuals provide an estimate on a scale from 1000 to 2000. In our simulations, we suppose that an individual's answer is drawn from a normal distribution. Each individual has their own normal distribution. All normal distributions are centered around the true answer - but they differ in their standard deviations. The value of that standard deviation is what we define as an individual's competence. The lower the standard deviation, the higher the competence, i.e.~the more likely a guess drawn from the normal distribution will be close to the true answer. We (arbitrarily) define a range of competence: we set the lowest competence equal to the range of possible values, i.e.the largest standard deviation (2000 - 1000) = 1000. We set the highest competence to 0.1\% of the range of possible values, i.e.the smallest standard deviation (0.001 x 1000) = 1 (see Fig. \ref{fig:example-competence-numeric}).

\begin{figure}
\includegraphics[width=0.75\linewidth]{output/figures/example-competence-numeric} \caption{Range of possible data generating functions for individuals.}\label{fig:example-competence-numeric}
\end{figure}

We suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution. This competence distribution can take vary different shapes, depending on the alpha and beta parameters that describe the distribution (see Fig. \ref{fig:simulations}).

We draw an estimate for each individual based on their respective competence distribution. For each individual, we then measure accuracy as the (squared) distance between their estimate and the true answer. Having a competence and an accuracy value for each individual, we randomly assign individuals to groups of three. For each group, we calculate the average of the three individuals' competence and accuracy. We measure the convergence of a group by calculating the standard deviation of the estimates. We run this simulation on a sample size of roughly 99900 (varying slightly as a function of sample size). We repeat this simulation process for various sample sizes and competence distributions. The results are displayed in Fig. \ref{fig:simulation-accuracy-numeric} for accuracy, and Fig. \ref{fig:simulation-competence-numeric} for competence. Across all underlying competence distributions, we find a positive correlation between convergence and accuracy, which tends towards 1 as sample size increases (see Fig. \ref{fig:simulation-numeric-correlations}). As for competence, we find a positive correlation between convergence and competence across all underlying competence distributions. However, these correlations are weaker than for accuracy, and do not increase with sample size (see Fig. \ref{fig:simulation-numeric-correlations}).

\begin{figure}
\includegraphics[width=0.75\linewidth]{output/figures/simulation-numeric-correlations} \caption{Correlation between the convergence of the answers of a group of agents and (left) the average accuracy of their answers; (right) the average competence of the agents, as a function of how many agents are in the group, and of the competence distribution of the agents (see Fig. \ref{fig:simulations}).}\label{fig:simulation-numeric-correlations}
\end{figure}

\subsection{Categorical choice context}\label{categorical-choice-context}

\begin{figure}
\includegraphics[width=1\linewidth]{output/figures/simulation-accuracy-sample-categorical} \caption{Accuracy as a function of convergence, for different population competence distributions and sample sizes. The number of choice options is three.}\label{fig:simulation-accuracy-sample-categorical}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{output/figures/simulation-competence-sample-categorical} \caption{Competence as a function of convergence (i.e.~vote share for an option), for different population competence distributions and sample sizes. The number of choice options is three.}\label{fig:simulation-competence-sample-categorical}
\end{figure}

When people make a choice based on several categories, their answers cannot be ranked by their nature (i.e.~they are nominal, not ordinal), and that there are fewer of them (e.g.~one of three possible products to choose, instead of an estimate between one and two thousand). In this case, convergence can be measured by the share of people agreeing on an answer. The larger the share of informants agreeing on an answer, the greater convergence. This measure is at the response level, nested within the group level.

As for the numeric scenario, we ran simulations to provide a normative answer as to whether it is justified to infer accuracy and competence from greater convergence. We, again, suppose that an individual's answer is drawn from an internal distribution - in this case, a multinomial distribution, that describes how likely the individual is to choose each available option. If there are m choice options, an individual has the probability p of picking the right one, and the probability of (1-p)/(m-1) to pick any other option. Each individual has their own multinomial distribution. We define competence as the probability of making the correct choice. The higher the competence, the greater the probability that an individual will choose the correct option. Competence values range from being at chance for selecting the correct option (p = 1/m) to certainly selecting the correct option (p = 1). As before in the numeric case, suppose that individual competence levels are drawn from a competence distribution, which can be expressed by a beta distribution (see Fig. \ref{fig:simulations}). Based on their competence level, we draw an estimate for each individual. We measure an individual's accuracy as a binary outcome, namely whether they picked the correct option, or not. We then randomly assign individuals to groups of informants (we vary the sample size from one simulation to another). Within these groups, we calculate the share of individuals voting for an answer option. For example, in a scenario in which three individuals pick among three options (A, B and C), two individuals might vote C and one B. In this case we obtain an average accuracy and an average competence value for a share of 2/3 (option C) and for a share of 1/3 (option B). We simulate this on a population of 99900 individuals. We repeat this procedure varying the underlying population competence distributions, and additionally varying either (a) the sample size of informants, or (b) the number of choice options. If we vary the sample size, we hold the number of choice options constant at n = 3, and vice versa when varying the number of choice options. Fig. \ref{fig:simulation-accuracy-sample-categorical} shows the average accuracy, and Fig. \ref{fig:simulation-competence-sample-categorical} the average competence value for each share of votes, for different competence levels and varying the sample size. Fig. \ref{fig:simulation-accuracy-options-categorical} and Fig. \ref{fig:simulation-competence-options-categorical} display the same relationship, but varying the number of choice options instead. The figures display that, across all sample sizes and competence levels, the larger the share of votes for an option, the more accurate the option is on average. That relationship appears to follow some sigmoid curve which switches from an average accuracy of 0 to an average accuracy of 1 before a share of 0.5 is attained, and which is steeper for larger sample sizes. For competence, we observe a similar sigmoid-like relationship, but of lesser amplitude and varying considerably as a function of the underlying population competence distributions.

\begin{figure}
\includegraphics[width=0.75\linewidth]{output/figures/simulation-accuracy-options-categorical} \caption{Accuracy as a function of vote share for an option, for different population competence distributions and number of choice options. Points represent averages across all simulations within the respective segment. The sample size is three.}\label{fig:simulation-accuracy-options-categorical}
\end{figure}

\begin{figure}
\includegraphics[width=0.75\linewidth]{output/figures/simulation-competence-options-categorical} \caption{Competence as a function of vote share for an option, for different population competence distributions and number of choice options. Points represent averages across all simulations within the respective segment. The sample size is three.}\label{fig:simulation-competence-options-categorical}
\end{figure}

In sum, given the set of specific assumptions we made, our simulations suggest that people are indeed justified in inferring accuracy and competence from convergence in both numeric and categorical choice settings.

\begin{figure}
\centering
\includegraphics{output/figures/simulation-accuracy-numeric.pdf}
\caption{\label{fig:simulation-accuracy-numeric}Simulation results showing the relationship between convergence and accuracy for different population competence distributions.}
\end{figure}

\begin{figure}
\centering
\includegraphics{output/figures/simulation-competence-numeric.pdf}
\caption{\label{fig:simulation-competence-numeric}Simulation results showing the relationship between convergence and competence for different population competence distributions.}
\end{figure}

\clearpage

\section{Experiment 1}\label{exp1}

\subsection{Stimuli}\label{stimuli}

\FloatBarrier

The means of the normal distributions that we draw our estimates from were distinct between sets of estimates. Considering our within-participant design, we wanted to ensure that participants understood each set of estimates as being the result of a different, unrelated game. In order to assure that random draws from the distributions will (most likely) appear on the response scale (1000 to 2000), we constrained the means of all normal distributions to lie between the first and third quartile of the response scale (i.e.~smallest possible mean was 1225 and largest 1775). We define a set of eight means--one for each set of estimates--that cover the range from first to third quartile of the predefined scale with an equal interval (1250, 1325, 1400, 1475, 1550, 1625, 1700, 1775). We randomly paired means with conditions when generating the stimuli. We then drew the set of estimates from the respective normal distributions given the assigned means and the condition constraints. We repeated this three times, resulting in three different series of eight sets of estimates. We randomly assign participants to one of these series. Additionally, for each participant, we randomize the order of appearance of the sets of estimates within the respective series. Images of all sets of estimates can be found on the OSF.

\subsection{Attention check}\label{attention-check}

Imagine you are playing video games with a friend and at some point your friend says: ``I don't want to play this game anymore! To make sure that you read the instructions, please write the three following words''I pay attention'' in the box below. I really dislike this game, it's the most overrated game ever.''

Do you agree with your friend? (Yes/No)

\subsection{Results}\label{results}

Figure \ref{fig:exp1-plot} visualizes the results and table \ref{tab:exp1-table} contains descriptive results. Figure \ref{fig:exp1-outcomes} visualizes score differences between the two outcomes, accuracy and competence.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:exp1-table}}

\begin{tabular}{llll}
\toprule
Convergence & \multicolumn{1}{c}{Number} & \multicolumn{1}{c}{Accuracy} & \multicolumn{1}{c}{Competence}\\
\midrule
divergent & small & 3.152 (sd = 1.495) & 3.57 (sd = 1.341)\\
divergent & large & 3.232 (sd = 1.282) & 3.465 (sd = 1.184)\\
convergent & small & 4.425 (sd = 1.461) & 4.695 (sd = 1.221)\\
convergent & large & 4.695 (sd = 1.424) & 4.805 (sd = 1.251)\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\begin{figure}
\centering
\includegraphics{output/figures/exp1-plot.pdf}
\caption{\label{fig:exp1-plot}Distributions of accuracy and competence by level of convergence.}
\end{figure}

\subsection{Research questions}\label{research-questions}

We had three additional research questions regarding the number of informants:

\textbf{\emph{RQ1: Do H1 and H2 hold for both a small {[}3{]} and a large {[}10{]} number of estimates?}}

\textbf{\emph{RQ2: When making a guess based on the opinions of (independent) informants, will participants be more confident about their guess when there is a larger number of estimates compared to when this number is smaller?}}

\textbf{\emph{RQ3: Is there an interaction effect between the number of estimates and convergence on perceived competence of informants?}}

Regarding RQ1, we ran the same mixed models as for the two hypotheses, but without number as a covariate. Instead, we ran separate models for the three and the ten informants conditions. We find that for both sub-samples, there is a positive effect of convergence on both accuracy (\(\hat{b}_{\text{three informants}}\) = 1.27 {[}1.112, 1.433{]}, p = \textless{} .001; \(\hat{b}_{\text{ten informants}}\) = 1.46 {[}1.285, 1.64{]}, p = \textless{} .001) and competence (\(\hat{b}_{\text{three informants}}\) = 1.12 {[}0.955, 1.295{]}, p = \textless{} .001; \(\hat{b}_{\text{ten informants}}\) = 1.34 {[}1.142, 1.538{]}, p = \textless{} .001).

To test the other two research questions, we ran the same mixed-models as for the hypotheses, but this time including an interaction between number of informants and convergence. We use deviation-coded versions of our convergent variable (divergent = -0.5, convergent = 0.5), allowing us to have a coefficient measuring the main effect of the number of informants in our model. Regarding RQ2, pooling across convergent and divergent conditions, we find a main effect of number, such that participants had more confidence in their estimate when they relied on ten informants compared to three informants (\(\hat{b}_{\text{Accuracy}}\) = 0.18 {[}0.076, 0.274{]}, p = \textless{} .001). Regarding RQ3, we find an interaction between number of informants and convergence on competence: the positive effect of convergence was stronger in scenarios involving ten informants compared to three informants (\(\hat{b}_{\text{Competence}}\) = 0.22 {[}0.028, 0.402{]}, p = 0.024). We do not find a statistically significant interaction on accuracy (\(\hat{b}_{\text{Accuracy}}\) = 0.19 {[}-0.007, 0.387{]}, p = 0.059).

\begin{figure}
\centering
\includegraphics{output/figures/exp1-outcomes.pdf}
\caption{\label{fig:exp1-outcomes}Differences between accuracy and competence ratings, by level of convergence.}
\end{figure}

\subsection{Minor deviation from preregistration}\label{minor-deviation-from-preregistration}

In our preregistration, we stated we would run paired t-tests and only conduct mixed models as a robustness check. However, it became clear to us later that mixed models are more appropriate in light of our data structure. In the main article, we therefore only report the results of the mixed models. Note, however, that using paired t-tests, we confirm our results (Accuracy: -1.37, t-statistic = -26.14, p \textless{} .001; Competence: -1.23, t-statistic = -21.82, p \textless{} .001)

\subsection{Sensitivity simulation}\label{sensitivity-simulation}

We had based our power calculations on a paired t-test (using the widely used tool G*Power), which does not correspond to the mixed-models we eventually ran. To address this point, here, we seek to identify the minimum effect size that our analysis with mixed models could have detected, given a 90\% power, an N of 200 participants, and other parameters derived from the estimates observed in the sample of Experiment 1 (see \ref{tab:exp1-regression-table}). By contrast to a standard power simulation (which varies sample size), a sensitivity simulation varies the effect size, while holding the sample size and desired power constant.

\begin{table}

\caption{\label{tab:exp1-regression-table}Model Results Experiment 1}
\centering
\begin{tabular}[t]{lcc}
\toprule
  & Accuracy & Competence\\
\midrule
(Intercept) & 3.105 & 3.516\\
 & (0.077) & (0.068)\\
convergenceconvergent & 1.368 & 1.232\\
 & (0.072) & (0.085)\\
numberlarge & 0.175 & 0.002\\
 & (0.050) & (0.048)\\
SD (Intercept ID) & 0.905 & 0.764\\
SD (convergenceconvergent ID) & 0.732 & 0.993\\
Cor (Intercept\textasciitilde{}convergenceconvergent ID) & -0.128 & -0.554\\
SD (Observations) & 1.006 & 0.955\\
\midrule
Num.Obs. & 1600 & 1600\\
R2 Marg. & 0.191 & 0.195\\
R2 Cond. & 0.594 & 0.532\\
AIC & 5131.0 & 4950.5\\
BIC & 5168.7 & 4988.1\\
ICC & 0.5 & 0.4\\
RMSE & 0.91 & 0.86\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{output/figures/sensitivity.pdf}
\caption{\label{fig:sensitivity}Results of the sensitivity analysis.}
\end{figure}

The results of this can be seen in Fig. \ref{fig:sensitivity}. In our sample, we found a statistically significant effect size of 1.37 for accuracy and an effect size of 1.23 for competence (both on their original scales from 1 to 7). The simulations show that, given our parameter assumptions (based on the observations in our sample) and a sample size of N = 200, we would have detected an effect as small as 0.4 with a power of at least 90\% (in fact nearly 100\%) for both accuracy and competence.

\clearpage

\section{Experiment 2}\label{exp2}

\subsection{Design}\label{design-6}

\FloatBarrier

We manipulated informational dependency. According to the condition, players read a different introduction before seeing a set of estimates (see Table \ref{tab:exp2-conditions}).

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:exp2-conditions}}

\begin{tabular}{m{3cm}m{12cm}}
\toprule
Condition & \multicolumn{1}{c}{Description}\\
\midrule
Independence & Players are asked to make completely independent decisions – they cannot see each other’s estimates, or talk with each other before giving their estimates.\\
Discussion & Players are asked to talk with each other about the game at length before giving their estimates.\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\subsection{Attention check}\label{attention-check-1}

Imagine you are playing video games with a friend and at some point your friend says: ``I don't want to play this game anymore! To make sure that you read the instructions, please write the three following words''I pay attention'' in the box below. I really dislike this game, it's the most overrated game ever.''

Do you agree with your friend? (Yes/No)

\subsection{Results}\label{results-1}

Figure \ref{fig:exp2-plot} visualizes the results and table \ref{tab:exp2-table} contains descriptive results.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:exp2-table}}

\begin{tabular}{lll}
\toprule
Independence & \multicolumn{1}{c}{Accuracy} & \multicolumn{1}{c}{Competence}\\
\midrule
dependent & 4.03 (sd = 1.389) & 4.48 (sd = 1.022)\\
independent & 3.775 (sd = 1.502) & 4.36 (sd = 0.967)\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\begin{figure}
\centering
\includegraphics{output/figures/exp2-plot.pdf}
\caption{\label{fig:exp2-plot}Distributions of accuracy and competence by level of informational dependency.}
\end{figure}

\clearpage

\section{Experiment 3}\label{exp3}

\subsection{Design}\label{design-7}

\FloatBarrier

We manipulated two factors: informational dependency (two levels, independence and conflict of interest, see Table \ref{tab:conditions-exp3}; between participants) and convergence (two levels, convergence and divergence; within participants). Participants saw four scenarios, one for each combination of the two factors convergence and informational dependency.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:conditions-exp3}}

\begin{tabular}{m{3cm}m{12cm}}
\toprule
Condition & \multicolumn{1}{c}{Description}\\
\midrule
Independence & Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.\\
Conflict of interest & All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future.\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\subsection{Attention check}\label{attention-check-2}

Imagine you are playing video games with a friend and at some point your friend says: ``I don't want to play this game anymore! To make sure that you read the instructions, please write the three following words''I pay attention'' in the box below. I really dislike this game, it's the most overrated game ever.''

Do you agree with your friend? (Yes/No)

\subsection{Stimuli}\label{stimuli-1}

Our design required that each participant sees four different sets of predictions--two convergent and two divergent ones. We generated new stimuli: As in experiment 1, estimates would appear on a scale from 1000 to 2000. By contrast with experiment 1, we generated the sets of estimates with random draws from uniform distributions (instead of normal distributions). We varied the range of these distributions according to convergence (60 for convergence, 600 for divergence). Switching from normal distributions to uniform distributions made `unlucky' draws in which conditions are visually not distinguishable less likely. And additional difference to experiment one was that all scenarios only involved groups of three informants (never ten).

Similar to the previous experiments, we also vary the value on the prediction scale (from 1000 to 2000) across which the range is centered. Considering our within-participant design, this, again, made it seem more likely that participants understand each set of predictions as being the result of a different stock, with a different true value. In order to assure all random draws from the distributions would appear on the response scale, we constrained the center of the uniform distributions to lie between 1300 and 1700. We define four center values -- one per set of predictions -- that divide this interval in (rounded) quartiles (1300, 1430, 1570, 1700). Given a center and a range, we then draw the predictions from uniform distributions. For example, in a draw for a divergent set of estimates with a center of 1700, each value within a range of 1400 and 2000 is equally likely to get selected. To avoid that single draws overlap too much within the same set, we defined a minimum space of 5 between the three predictions of a set.

To minimize confounding of convergence with a certain placement of the center value, we paired center values with conditions such that each condition appears once in each half of the scale and each condition gets one of the extreme values. For example, in one set the convergence condition would get assigned center values of 1300 and 1570, the divergent condition center values of 1430 and 1700). We generated four such series (all possible combinations) and randomly assigned participants to one of them. Additionally, for each participant, we randomized the order of appearance of the sets of predictions within the respective series.

For each set of predictions, we calculated the empirical mean based on the randomly drawn estimates. In the conflict of interest condition, this mean was inserted as the value that participants were be told experts would gain from. Consequently, the convergent predictions converge around what is said to be the incentivized value for the experts to choose.

\subsection{Results}\label{results-2}

Table \ref{tab:exp3-table} contains descriptive results.

\begin{table}

\caption{\label{tab:exp3-table}}
\centering
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Divergent & Convergent & Divergent & Convergent\\
\midrule
Conflict of interest & 3.566 (sd = 1.215) & 4.455 (sd = 1.409) & 3.732 (sd = 1.101) & 4.556 (sd = 1.19)\\
Independent & 3.4 (sd = 1.08) & 5.28 (sd = 1.052) & 3.61 (sd = 1.111) & 5.235 (sd = 0.992)\\
\bottomrule
\end{tabular}
\end{table}

\clearpage

\section{Experiment 4}\label{exp4}

\FloatBarrier

\subsection{Attention check}\label{attention-check-3}

Imagine you are playing video games with a friend and at some point your friend says: ``I don't want to play this game anymore! To make sure that you read the instructions, please write the three following words''I pay attention'' in the box below. I really dislike this game, it's the most overrated game ever.''

Do you agree with your friend? (Yes/No)

\subsection{Stimuli}\label{stimuli-2}

We manipulate convergence within participants. All participants see all four conditions, with two stimuli (i.e.~game results) per condition (see Table \ref{tab:exp4-stimuli}). Each participant therefore sees eight stimuli in total (4 convergence levels x 2 stimuli)

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1007}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4497}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4497}}@{}}
\caption{\label{tab:exp4-stimuli}All stimuli by levels of convergence}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
minority (0) & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/opp_majority_a.png} & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/opp_majority_b.png} \\
dissensus (1) & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/divergence_a.png} & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/divergence_b.png} \\
majority (2) & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/majority_a.png} & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/majority_b.png} \\
consensus (3) & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/consensus_a.png} & \includegraphics[width=0.6\textwidth,height=\textheight]{Experiment_4/figures/stimuli/consensus_b.png} \\
\end{longtable}

\subsection{Results}\label{results-3}

Table \ref{tab:exp4-table} contains descriptive results and Figure \ref{fig:exp4-plot} visualizes the results.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:exp4-table}}

\begin{tabular}{lll}
\toprule
Convergence & \multicolumn{1}{c}{Accuracy} & \multicolumn{1}{c}{Competence}\\
\midrule
minority (0) & 33.13 (sd = 18.267) & 3.49 (sd = 1.156)\\
divergence (1) & 38.525 (sd = 13.708) & 3.93 (sd = 0.726)\\
majority (2) & 64.065 (sd = 13.86) & 4.88 (sd = 0.713)\\
consensus (3) & 80.745 (sd = 18.633) & 5.45 (sd = 0.981)\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\begin{figure}
\centering
\includegraphics{output/figures/exp4-plot.pdf}
\caption{\label{fig:exp4-plot}Results of Experiment 4, showing the distributions of accuracy and competence by level of convergence.}
\end{figure}

\subsection{Comparison with simulation}\label{comparison-with-simulation}

We compared the accuracy and competence ratings of the participants to actual accuracy and competence values from simulated data. The data was generated using the model described in the main paper, and assuming a uniform competence distribution. Compared to this data generating model, participants underestimate the effect of convergence for both accuracy and competence, but more for accuracy. Table \ref{tab:model-vs-participant-table} compares the results of regressions on participant data and regressions on simulated data. Fig. \ref{fig:exp4-model} visualizes the results.

\begin{table}

\caption{\label{tab:model-vs-participant-table}Participants vs. Model (Exp. 4)}
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Participants} & \multicolumn{2}{c}{Model} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Accuracy & Competence & Accuracy  & Competence \\
\midrule
(Intercept) & 28.859*** & 3.413*** & 15.211*** & 3.233***\\
 & (1.702) & (0.098) & (0.221) & (0.008)\\
convergence & 16.838*** & 0.683*** & 28.471*** & 0.425***\\
 & (0.922) & (0.053) & (0.097) & (0.003)\\
SD (Intercept id) & 15.926 & 0.929 &  & \\
SD (convergence id) & 8.640 & 0.497 &  & \\
Cor (Intercept\textasciitilde{}convergence id) & -0.808 & -0.857 &  & \\
SD (Observations) & 10.159 & 0.553 &  & \\
\midrule
Num.Obs. & 800 & 800 & 333000 & 333000\\
R2 &  &  & 0.205 & 0.047\\
R2 Adj. &  &  & 0.205 & 0.047\\
R2 Marg. & 0.555 & 0.408 &  & \\
R2 Cond. & 0.839 & 0.786 &  & \\
AIC & 6402.6 & 1756.0 & 3301803.1 & 1051474.5\\
BIC & 6430.7 & 1784.1 & 3301835.2 & 1051506.7\\
ICC & 0.6 & 0.6 &  & \\
Log.Lik. &  &  & -1650898.534 & -525734.261\\
F &  &  & 85844.746 & 16470.902\\
RMSE & 8.98 & 0.49 & 34.42 & 1.17\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{output/figures/exp4-model.pdf}
\caption{\label{fig:exp4-model}Comparison of accuracy (top) and competence (bottom) ratings in participant data (left) to observed accuracy in simulated data (right), by level of convergence.}
\end{figure}

\clearpage

\section{Experiment 5}\label{exp5}

\subsection{Design}\label{design-8}

\FloatBarrier

We manipulated convergence within participants in the same way we did in experiment 4. In addition, between participants, we manipulated informational dependence, akin to experiment 3 (Table \ref{tab:exp5-condition}). In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is. In the independent condition, there was no such conflict of interest and experts were described as independent.

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:exp5-condition}}

\begin{tabular}{m{3cm}m{12cm}}
\toprule
Condition & \multicolumn{1}{c}{Description}\\
\midrule
Independence condition & The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.\\
Conflict of interest condition & The three advisors have already invested in one of the three options, the same option for all three. As a result, they have an incentive to push that option in their recommendations.\\
\bottomrule
\end{tabular}

\end{threeparttable}
\end{center}

\end{table}

\subsection{Attention check}\label{attention-check-4}

Imagine you are playing video games with a friend and at some point your friend says: ``I don't want to play this game anymore! To make sure that you read the instructions, please write the three following words''I pay attention'' in the box below. I really dislike this game, it's the most overrated game ever.''

Do you agree with your friend? (Yes/No)

\subsection{Results}\label{results-4}

Table \ref{tab:exp5-table} contains descriptive results.

\begin{table}

\caption{\label{tab:exp5-table}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Conflict of interest & Independent & Conflict of interest & Independent\\
\midrule
minority (0) & 39.733 (sd = 20.238) & 36.46 (sd = 16.793) & 3.802 (sd = 1.146) & 3.606 (sd = 1.006)\\
divergence (1) & 41.52 (sd = 13.344) & 42.51 (sd = 13.6) & 4.005 (sd = 0.878) & 4.045 (sd = 0.839)\\
majority (2) & 55.416 (sd = 16.637) & 59.515 (sd = 14.575) & 4.495 (sd = 0.921) & 4.737 (sd = 0.879)\\
consensus (3) & 66.198 (sd = 24.825) & 71.914 (sd = 23.457) & 4.96 (sd = 1.273) & 5.247 (sd = 1.264)\\
\bottomrule
\end{tabular}}
\end{table}

\clearpage

\section{Experiment 6}\label{exp6}

\subsection{Design}\label{design-9}

\FloatBarrier

The main focus of experiment 6 was on comparing 3-options conditions to 10-options conditions (see Table \ref{tab:stimuli-options}). Additionally, we varied between two versions of the 10-options conditions: one in which the range of the answers corresponded to the range of the three options condition (see Table \ref{tab:stimuli-10}), and another with increased range (see Table \ref{tab:stimuli-10-alt}).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.4965}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.5035}}@{}}
\caption{\label{tab:stimuli-options}Example of a consensus stimulus for the two `Number of option' conditions}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Number of options: 3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of options: 10
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Number of options: 3
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Number of options: 10
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\includegraphics[width=0.4\textwidth,height=\textheight]{stimuli/Experiment_6/stimuli_Exp6/consensus_3_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{stimuli/Experiment_6/stimuli_Exp6/consensus_10_a.png} \\
\end{longtable}

\subsection{Attention check}\label{attention-check-5}

Imagine you are playing video games with a friend and at some point your friend says: ``I don't want to play this game anymore! To make sure that you read the instructions, please write the three following words''I pay attention'' in the box below. I really dislike this game, it's the most overrated game ever.''

Do you agree with your friend? (Yes/No)

\subsection{Stimuli}\label{stimuli-3}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0993}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4503}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4503}}@{}}
\caption{\label{tab:stimuli-10}Stimuli for 10 options condition by levels of convergence}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
minority (0) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_b.png} \\
dissensus (1) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_b.png} \\
majority (2) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_b.png} \\
consensus (3) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_b.png} \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0943}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4528}}
  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4528}}@{}}
\caption{\label{tab:stimuli-10-alt}Alternative stimuli for 10 options condition by levels of convergence}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Level
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version a)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Version b)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
minority (0) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_b_alt.png} \\
dissensus (1) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_b_alt.png} \\
majority (2) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_b_alt.png} \\
consensus (3) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_b_alt.png} \\
\end{longtable}

\subsection{Results}\label{results-5}

Figure \ref{fig:exp5-plot} visualizes the results and table \ref{tab:exp5-table} contains descriptive results.

\begin{table}

\caption{\label{tab:exp6-table}}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lllll}
\toprule
\multicolumn{1}{c}{} & \multicolumn{2}{c}{Accuracy} & \multicolumn{2}{c}{Competence} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & 10 options & 3 options & 10 options & 3 options\\
\midrule
minority (0) & 32.333 (sd = 22.987) & 35.703 (sd = 20.829) & 3.61 (sd = 1.245) & 3.574 (sd = 1.227)\\
divergence (1) & 37.287 (sd = 19.569) & 40.861 (sd = 18.759) & 4.067 (sd = 0.831) & 4.01 (sd = 0.881)\\
majority (2) & 63.69 (sd = 21.826) & 64.301 (sd = 17.355) & 4.957 (sd = 0.885) & 4.834 (sd = 0.881)\\
consensus (3) & 79.967 (sd = 24.15) & 80.152 (sd = 20.774) & 5.647 (sd = 1.035) & 5.466 (sd = 1.051)\\
\bottomrule
\end{tabular}}
\end{table}

\begin{figure}
\centering
\includegraphics{output/figures/exp6-plot.pdf}
\caption{\label{fig:exp6-plot}Interaction of convergence and informational dependency.}
\end{figure}

\subsection{Comparison with simulation}\label{comparison-with-simulation-1}

Since we didn't find a difference between three and ten choice option scenarios in the participant data, we wanted to have a normative point of reference of what should have been the expected difference, given our model. We therefor compared participant ratings to observed values from simulated data. The data was generated using the model described in the main paper, and assuming a uniform competence distribution. Compared to this data generating model, participants considerably underestimated how the number of choice options alters the effect of convergence for both accuracy and competence. Table \ref{tab:model-vs-participant-table} compares the results of regressions on participant data and regressions on simulated data. Fig. \ref{fig:exp6-model} visualizes the results.

\begin{table}

\caption{\label{tab:unnamed-chunk-101}Participants vs. Model (Exp. 6)}
\centering
\begin{tabular}[t]{lcccc}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Participants} & \multicolumn{2}{c}{Model} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5}
  & Accuracy & Competence & Accuracy  & Competence \\
\midrule
(Intercept) & 29.830*** & 3.508*** & 11.407*** & 3.189***\\
 & (1.197) & (0.063) & (0.220) & (0.009)\\
convergence & 16.305*** & 0.675*** & 31.596*** & 0.526***\\
 & (0.600) & (0.032) & (0.130) & (0.005)\\
number\_options\_effect\_code & -3.812 & 0.024 & -13.794*** & -0.207***\\
 & (2.394) & (0.126) & (0.439) & (0.018)\\
convergence × number\_options\_effect\_code & 1.252 & 0.050 & 8.498*** & 0.251***\\
 & (1.200) & (0.065) & (0.260) & (0.011)\\
SD (Intercept id) & 19.383 & 1.031 &  & \\
SD (convergence id) & 9.625 & 0.527 &  & \\
Cor (Intercept\textasciitilde{}convergence id) & -0.699 & -0.829 &  & \\
SD (Observations) & 12.103 & 0.598 &  & \\
\midrule
Num.Obs. & 2384 & 2384 & 51003 & 51003\\
R2 &  &  & 0.539 & 0.165\\
R2 Adj. &  &  & 0.539 & 0.165\\
R2 Marg. & 0.423 & 0.355 &  & \\
R2 Cond. & 0.814 & 0.779 &  & \\
AIC & 19980.3 & 5624.5 & 497742.9 & 170192.2\\
BIC & 20026.5 & 5670.7 & 497787.1 & 170236.4\\
ICC & 0.7 & 0.7 &  & \\
Log.Lik. &  &  & -248866.472 & -85091.109\\
F &  &  & 19902.699 & 3370.232\\
RMSE & 10.67 & 0.53 & 31.83 & 1.28\\
\bottomrule
\multicolumn{5}{l}{\rule{0pt}{1em}+ p $<$ 0.1, * p $<$ 0.05, ** p $<$ 0.01, *** p $<$ 0.001}\\
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics{output/figures/exp6-model.pdf}
\caption{\label{fig:exp6-model}Comparison of accuracy (top) and competence (bottom) ratings in participant data (left) to observations in simulated data (right), by level of convergence.}
\end{figure}

\end{document}

### 2024-09-11 initial_manuscript.tex
add .tex file and original cover letter

--- 2024-09-11 paper.tex
@@
 
-We begin by reviewing existing studies showing that people infer that competent informants who converge in their opinions are likely to be accurate.
-The wisdom of crowds literature provides normative grounds for this inference.
-We then argue that both the experimental and theoretical literature have paid little attention to extending this inference to cases in which there is no information about the informants' competence, and to inferences about the competence of the informants.
-We first develop normative models, both analytically and with simulations, to show that inferences from convergence to accuracy and to competence are warranted under a wide range of parameters.
-Second, we present a series of experiments in which participants evaluate both the accuracy and competence of informants as a function of how much their answers converge on a given problem, in the absence of any priors about these individuals' competence, or what the correct answer is.
+We first review experiments testing whether people infer that individuals whose answers converge are more likely to be right, and to be competent.
+We then argue that little attention has been paid to the normative models that might explain when and why an inference from convergence to competence is warranted.
+We develop such models, both analytically and with simulations.
+Finally, we present a series of experiments in which participants evaluate the accuracy and competence of individuals as a function of how much their answers converge on a given answer, in the absence of any priors about these individuals' competence, or what the correct answer is.
 
@@
 
-In the continuous case, the most relevant evidence comes from the literature on `advice-taking' (for review, see, Kämmer, Choshen-Hillel, Müller-Trede, Black, \& Weibler, 2023).
-In these experiments, participants are called `judges' who need to make numerical estimates--sometimes on factual knowledge, e.g.~`What year was the Suez Canal opened first?' (Yaniv, 2004), sometimes on knowledge controlled by the experimenters, e.g.~`How many animals were on the screen you saw briefly?' (Molleman et al., 2020).
+In the continuous case, the most relevant evidence comes from the literature on `advice-taking.' In these experiments, participants are called `judges' who need to make numerical estimates--sometimes on factual knowledge, e.g.~`What year was the Suez Canal opened first?' (Yaniv, 2004), sometimes on knowledge controlled by the experimenters, e.g.~`How many animals were on the screen you saw briefly?' (Molleman et al., 2020).
 To help answer these questions, participants are given estimates from others, the `advisors'.
@@
 In categorical choice contexts, there is ample and long-standing (e.g.
-Crutchfield, 1955) evidence from experimental psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see H.
-Mercier \& Morin, 2019).
+Crutchfield, 1955) evidence from social psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see Mercier \& Morin, 2019).
 This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role (e.g.~because the answers are private, see Asch, 1956).
@@
 
-If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the informants' competence.
+If many studies have demonstrated that participants tend to infer that more convergent answers are more likely to be correct, few have examined whether participants thought that this convergence was indicative of the individuals' competence.
 One potential exception is a study with preschoolers in which the children were more likely to believe the opinion of an informant who had previously been the member of a majority over that of an informant who had dissented from the majority (Corriveau et al., 2009).
@@
 
-\section{What inferences from convergence should we expect people to draw?}\label{what-inferences-from-convergence-should-we-expect-people-to-draw}
-
-Should we expect that people be able to infer that more convergent answers likely indicate not only more accurate answers, but also that those who gave the answers were competent?
-In order to make the best of communicated information, humans have to be able to evaluate it, so as to discard inaccurate or harmful information, while accepting accurate and beneficial information (Smith \& Harper, 2003).
-It has been argued that a suite of cognitive mechanisms--mechanisms of epistemic vigilance--evolved to serve this function (Hugo Mercier, 2020; Sperber et al., 2010).
-Since the opinion of more than one individual is often available to us, there should be mechanisms of epistemic vigilance dedicated to processing such situations.
-It would be these mechanisms that lead us to put more weight on an opinion that is shared by a larger majority (in relative or absolute terms), and, in some cases at least, to discount majority opinion when the opinions haven't been formed independently of each other (Hugo Mercier \& Miton, 2019).
-Evidence suggests that these mechanisms rely on heuristics which become more refined with age (Morgan et al., 2015), and which are far from perfect (Hugo Mercier \& Miton, 2019; in particular, they ignore many cases of informational dependencies, see, e.g.
-Yousif, Aboody, \& Keil, 2019).
-As mentioned above, in the experiments evaluating how people process convergent information, the participants had grounds to believe that the information came from competent informants.
-However, the same mechanisms could lead people to perform the same inference when they do not have such information (especially since, as is shown presently, such an inference is warranted).
-
-Regarding the inference from convergence to competence, other cognitive mechanisms allow us to infer how competent people are, based on a variety of cues, from visual access (Pillow, 1989), to the time it takes to answer a question (Richardson \& Keil, 2022).
-One of the most basic of these mechanisms infers from the fact that someone was right, that they possess some underlying competence (e.g.
-Koenig, Clément, \& Harris, 2004).
-As a result, if participants infer that convergent opinions are more likely to be accurate, they should also infer that the informants who provided the opinions are competent.
-Before testing whether participants infer accuracy and competence from convergence, we show that these inferences are normatively warranted, both in the categorical and in the continuous case.
+\section{Is inferring competence from convergence justified?}\label{is-inferring-competence-from-convergence-justified}
+
+There is clear evidence that, both in the continuous and the categorical case, participants infer that convergent opinions are more likely to be correct.
+There is also tentative evidence that participants infer that individuals holding more convergent opinions are more likely to be competent.
+Is that second inference justified?
+More precisely, is it normatively warranted to claim that individuals holding convergent opinions are more likely to be right not only for this specific question, but also be right for other (related) questions?
+We first provide a general analytic answer, before presenting simulations for the continuous and categorical cases.
 
@@
 
-\caption{Results of the analytic argument.
-The figure shows the average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario.
+\caption{Average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario.
 Here, we assume that 20 individuals answer a 5-choice question.
@@
 
-A population of around 990000 agents (varying slightly as a function of group sizes) with different competence levels is generated.
+A population of around 990000 agents (varying slightly with as a function of group sizes) with different competence levels is generated.
 An answer is drawn for each agent, based on their respective competence distribution.
@@
 Competence values range from chance level (p = 1/m) to always selecting the correct option (p = 1).
-Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see Fig.
+Individual competence levels are drawn from the same population competence beta distributions as in the numerical case (see see Fig.
 \ref{fig:simulations} A).
@@
 Within these groups, we first calculate the share of individuals voting for each answer, allowing us to measure convergence.
-For example, in a scenario with three choice options and three individuals, two might vote for option A, and one for option C, resulting in two levels of convergence, \(2/3\) for A and \(1/3\) for C.
-For each level of convergence occurring within a group, we then compute (i) the accuracy (either 1 if the correct option or 0 else),, (ii) the average competence of agents.
-Across all groups, we then compute the averages of these values, for each level of convergence.
-
-We repeat this procedure varying population competence distributions and the size of informant groups, holding the number of choice options constant at n = 3 (for simulations with varying choice options, see ESM section \ref{simulations}).
+For each level of convergence occurring within a group, we then compute (i) the average accuracy, (ii) the average competence of agents.
+Across all groups, we then compute the averages of these within-group averages, for each level of convergence.
+
+We repeat this procedure varying population competence distributions, and the size of informant groups, holding the number of choice options constant at n = 3 (for simulations with varying choice options, see ESM section \ref{simulations}).
 Fig.
-\ref{fig:simulations} B shows the average accuracy (left), and the average competence (right) value as a function of convergence, across different underlying competence distributions and group sizes.
-
-The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Irrespective of group size (and number of choice options) and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate.
-\item
-  For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents.
+\ref{fig:simulations} B shows the average accuracy (left), and the average competence (right) value for each share of votes, across different underlying competence distributions and group sizes.
+
+The simulations for the numerical and categorical case demonstrate a similar pattern, which can be summarized as follows: (i) Irrespective of group size (and number of choice options) and of the competence distribution, there is a very strong relation between convergence and accuracy: more convergent answers tend to be more accurate.
+(ii) For any group size and any competence distribution, there is a relation between convergence and the competence of the agents: more convergent answers tend to stem from more competent agents.
 The strength of this relation is not much affected by the number of agents whose answers are converging, but, although it is always positive, it ranges from very weak to very strong depending on the population's competence distribution.
-\item
-  The relation between convergence and accuracy is always much stronger than the relation between convergence and competence of the agents.
-\end{enumerate}
+(iii) The relation between convergence and accuracy is always much stronger than the relation between convergence and average competence of the agents.
 
@@
 \includegraphics{output/figures/simulations.pdf}
-\caption{\label{fig:simulations}Results of the simulations.
-\textbf{A} Shows the different population competence distributions we considered in our simulations.
-In the continuous simulations, competence values of 0 correspond to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a very small standard deviation (1, on the same scale).
-In the categorical simulations, a competence value of 0 corresponds to chance (e.g.~in a 3-choice-options scenario, an individual picking the correct answer with a probability of \(1/3\)), while a competence value of 1 corresponds to definitely picking the correct answer.
+\caption{\label{fig:simulations}\textbf{A} Shows the different population competence distributions we considered in our simulations.
+In the continuous simulations, competence values of 0 correspond to a very large standard deviation (1000, with a mean of 1500, on a scale from 1000 to 2000), thereby practically taking the form of a uniform distribution, while competence of 1 corresponds to a standard deviation of 1.
+In the categorical simulations, a competence value of 0 corresponds to chance (e.g.~in a 3-choice-options scenario, an individual picking the correct answer with a probability of \(1/3\)), while competence of 1 corresponds to definitely picking the correct answer.
 \textbf{B} Shows the results of simulations in a categorical setting with three choice options.
-Points represent average accuracy (left)/competence (right) values by degree of convergence (measured by the share of votes for an option), for different population competence distributions (panels) and sample sizes (colors).
+Points represent average accuracy (left)/competence (right) values by degree of convergences (measured by the share of votes for an options), for different population competence distributions and sample sizes.
 \textbf{C} Shows the results in a continuous setting.
-Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to being further up on the y-axis) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).}
+Regression lines represent the correlation between accuracy (left; measured by squared distance to true mean and reversed such that greater accuracy corresponds to top of the figure) or competence (right), respectively, and convergence (reversed such that greater convergence corresponds to being more right on the x-axis).}
 \end{figure}
@@
 
-Our models indicate that groups of informants are more likely to have given accurate answers, and to be competent, when their answers converge.
-In a series of experiments, we test whether people draw these inferences both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6).
+We test whether people infer that informants are more likely to give accurate answers, and to be competent, when their answers converge, both in numerical tasks (Experiments 1, 2, 3), and in categorical tasks (Experiments 4, 5, 6).
 By contrast with previous studies, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be.
-There has recently been much interest in investigating whether participants are able to take informational dependencies into account when evaluating convergent information (Desai, Xie, \& Hayes, 2022; Hugo Mercier \& Miton, 2019; Xie \& Hayes, 2022; Yin, Xu, Lin, Zhou, \& Guo, 2024; Yousif et al., 2019).
-To test whether participants took into account dependencies between the informants in the present context, this factor was manipulated in two different ways (in Experiment 2, some informants had discussed with each other; in Experiments 3 and 5, some informants had an incentive to provide the same answer).
-We predicted that dependencies between the informants' answers should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.
+We also manipulated (Experiments 2, 3, and 5) whether the convergence of the answers could be explained by something else than accuracy, which should reduce participants' reliance on the convergence of the answer as a cue to accuracy and to competence.
 
 All experiments were preregistered.
-All documents, data and code can be found on Open Science Framework \href{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}{project page} (\url{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}).
+Preregistration documents, data and code can be found on Open Science Framework \href{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}{project page} (\url{https://osf.io/6abqy/?view_only=42632bccea604fd7928dbe58e087d23b}).
 All analyses were conducted in R (version 4.2.2) using R Studio.
@@
 In Experiment 1, participants were provided with a set of numerical estimates which were more or less convergent, and asked whether they thought the estimates were accurate, and whether the informants making the estimates were competent.
-Perceptions of accuracy were measured as the confidence in what the participants thought the correct answer was, on the basis of the numerical estimates provided: the more participants think they can confidently infer the correct answer, the more they must think the estimates accurate, on average (the results replicate with a more direct measure of accuracy, see Experiment 3).
+Perceptions of accuracy were measured as the confidence in what the participants thought the correct answer was, on the basis of the numerical estimates provided: the more participants think they can confidently infer the correct answer, the more they must think the estimates accurate, on average.
 Our hypotheses were:
@@
 
-We manipulated two experimental factors, with two levels each: the convergence of the estimates (how close they were to each other; levels: divergent/convergent), and the number of estimates (how many players there were; levels: three/ten).
+We manipulated two experimental factors, with two levels each: the convergence of the estimates (how close they were to each other; levels: divergent/convergent), and the number of estimates (how many players there were; levels: low/high).
 This latter factor was chiefly included to make our results more robust, and was not attached to specific hypotheses.
-We used a 2 (convergence: divergent/convergent) x 2 (number: three/ten) within-participant design, with each participant seeing all the conditions.
+We used a 2 (convergence: divergent/convergent) x 2 (number: low/high) within-participant design, with each participant seeing all the conditions.
 Participants saw two different sets of estimates per condition, for a total of eight sets of estimates per participant.
@@
 
-To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slopes of convergence for participants.
+To account for dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slopes of convergence for participants.
 In the models for our hypotheses, we control for the number of estimates provided to the participants (three or ten).
@@
 We do not find a statistically significant interaction that would indicate a difference of the effect of convergence (\(\hat{b}\) = 0.14 {[}-0.001, 0.271{]}, 0.052).
-Pooled across divergent and convergent conditions, we find that participants reported lower perceived accuracy than competence (\(\hat{b}\) = -0.26 {[}-0.359, -0.156{]}, \textless{} .001)
+We find that pooled across divergent and convergent conditions, participants reported lower perceived accuracy than competence (\(\hat{b}\) = -0.26 {[}-0.359, -0.156{]}, \textless{} .001)
 
 In summary, as predicted, when the informants' answers were more convergent, participants were more confident that their answers were correct, and they believed the informants to be more competent.
-This was true both when there were three informants and when there were ten informants.
+This was true whether there were three or ten estimates.
 
@@
 
-We have shown that it is rational to infer that convergent estimates are more likely to be accurate, and to have been made by competent individuals, only if these individuals were independent and unbiased.
-However, convergence could come about differently.
+It is rational to infer that convergent estimates are more likely to be accurate, and to have been made by competent individuals, only if the convergence is best explained by the accuracy of the estimates.
+However, convergence could also be the outcome of other factors.
 If the individuals do not make their estimates independently of each other, a single individual might exert a strong influence on the others, making their convergence a poor cue to their accuracy.
@@
 
-Note that these predictions only stem from past empirical results, and are not necessarily normatively justified (modeling the effects of discussion would be beyond the scope of this paper, but see Dietrich \& Spiekermann, 2024).
-
 \subsection{Methods}\label{methods-1}
@@
 We used the materials generated for the convergent condition of Experiment 1.
-By contrast to Experiment 1, participants saw only two stimuli in total (one set of estimates per condition), and we only used stimuli involving groups of three informants.
-Otherwise, we proceeded just as in Experiment 1: we randomly assigned individual participants to one of the three series of stimuli, and for each participant, we randomized the order of appearance of conditions.
+By contrast to Experiment 1, participants saw only two stimuli in total (one set of estimates per condition).
+Otherwise, we proceeded just as in experiment one: we randomly assigned individual participants to one of the three series of stimuli, and for each participant, we randomized the order of appearance of conditions.
 
@@
 
-Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently made estimates, to be less accurate, or produced by less competent individuals.
-This might stem from the fact that participants, in various situations, neglect informational dependencies (Yousif et al., 2019), or from the fact that discussing groups actually perform better than non-discussing groups in a range of tasks (for review, see, e.g., H.
-Mercier, 2016), including numerical estimates (e.g.
-H.
-Mercier \& Claidière, 2022).
+Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently, to be less accurate, or made by less competent individuals.
+This suggests that participants might not be sensitive to at least some forms of informational dependency between the individuals making the estimates.
+Although previous studies have found that participants sometimes discount the opinions of groups compared to those of independent individuals, the superior performance of groups over individuals for a range of tasks (for review, see, e.g., Mercier, 2016), including numerical estimates (e.g.
+Mercier \& Claidière, 2022) suggests that this discounting might have been misguided.
 As a result, the participants in the current experiment might have been behaving rationally when they did not discount the estimates made after discussion.
@@
 
-Experiment 3 tests whether participants are sensitive to another potential source of dependency between convergent estimates: when the individuals making the estimate share an incentive to bias their estimates and disregard accuracy.
+Experiment 3 tests whether participants are sensitive to another potential source of dependency between convergent estimates: when the individuals making the estimate share an incentive to bias their estimates and disregarde accuracy.
 Even though Experiment 3 is formally similar to Experiment 1, the setting is different, as participants were told that they would be looking at (fictional) predictions of experts for stock values, instead of the answers of individuals in abstract games.
 In the conflict of interest condition, the experts had an incentive to value the stock in a given way, while they had no such conflict of interest in the independence condition.
-We tested for an interaction, namely whether the positive effect of convergence is reduced when informants are systematically biased, compared to when they are not.
+Given the nature of this manipulation, participants might have discounted the opinions of the conflicted experts, irrespective of the degree of convergence of their estimates.
+As a result, we could not just condition on convergent answers, as we did in Experiment 2.
+Instead, we tested for an interaction, namely whether the positive effect of convergence is reduced when informants are systematically biased, compared to when they are not.
 On this basis, we formulate four hypotheses, two of which are identical to those of Experiment 1, and only apply in the independent condition, and two that bear on the interaction:
@@
 \textbf{\emph{H2b: The effect of convergence on competence (H1b) is more positive in a context where informants are independent compared to when they are in a conflict of interest.}}
-
-We have not conducted simulations to validate these predictions.
-However, given the operationalization chosen, in which convergence should provide little evidence of either accuracy or competence, we believe the predictions regarding the superiority of the independent informants stem naturally from the model of independent informants presented above.
 
@@
 Each participant rated four scenarios, two for each level of convergence.
-By contrast to Experiment 1, all scenarios only involved groups of three informants.
 
@@
 We find evidence for all four hypotheses.
-As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample of Experiment 3 to half of the participants, namely those who were assigned to the independence condition.
-On this reduced sample, we ran the exact same analyses as in Experiment 1 and replicated the results.
+As for the first set of hypotheses, to match the setting of experiment one, we reduced the sample of experiment three to half of the participants, namely those who were assigned to the independence condition.
+On this reduced sample, we ran the exact same analyses as in experiment 1 and replicated the results.
 As for accuracy, participants rated informants in convergent scenarios (mean = 5.28, sd = 1.05) as more accurate than in divergent ones (mean = 3.40, sd = 1.08; \(\hat{b}_{\text{Accuracy}}\) = 1.88 {[}1.658, 2.102{]}, p = \textless{} .001).
@@
 \includegraphics{output/figures/exp3-plot.pdf}
-\caption{\label{fig:exp3-plot}Results of Experiment 3, showing the distributions of accuracy and competence ratings by convergence and informational dependency.}
+\caption{\label{fig:exp3-plot}Distributions of accuracy and competence by convergence and informational dependency.}
 \end{figure}
@@
 The set-up is similar to that of Experiment 1, except that the outcomes seen by the participants are not numerical estimates, but choices made between a few options.
-An additional difference is that participants rate a focal informant, and not a group of informants.
-There were two reasons for this choice: First, accuracy is not on a continuum as in the first three experiments (an option was either correct or not), so forming an average across informants who chose different options was less sensible.
-Second, rating a focal individual allowed us to have a minority condition, which would not have been possible when providing an average rating for a group.
 Experiment 4 tests hypotheses that are analogous to those of Experiment 1:
@@
 
-To account for dependencies of observations due to our within-participant design, we ran mixed models, with a random intercept and a random slope for participants.
+To account for dependencies of observations due to our within-participant design, we ran mixed models, with random intercept and random slope for participants.
 
@@
 
-In the ESM (section \ref{exp4}), we show that compared to what the normative models would predict, participants underestimate the effect of convergence on both accuracy and competence, but especially on accuracy.
+In the ESM (section \ref{exp4}), we show participants underestimate the effect of convergence on both accuracy and competence, but especially on accuracy.
 
@@
 
-Experiment 5 is a conceptual replication of Experiment 3 in a categorical instead of a numerical case: are participants less likely to infer that more convergent estimates are more accurate, and the individuals who made them more competent, when the estimates are made by individuals with a conflict of interest pushing them to all provide a given answer, compared to when they are made by independent individuals?
+Experiment 5 is a conceptual replication of Experiment 3 in a categorical instead of a numerical case: are participants less likely to infer that more convergent estimates are more accurate, and the individuals who made them more competent, when the estimates are made by individuals with a conflict of interest pushing them to all provide a given answer, compared to when they are made by independent participants?
 The independence condition of Experiment 5 also serves as a replication of Experiment 4, leading to the following hypotheses:
@@
 The power simulation for accuracy suggested that for 80 participants, we would have a power of at least 90\% for the interaction effect.
-The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants would we also detect an effect of convergence.
+The simulation for competence suggested that with already 40 participants, we would detect an interaction, but only with 60 participants we would also detect an effect of convergence.
 Due to uncertainty about our assumptions and because resources were available for a larger sample, we recruited 200 participants, in the UK and via Prolific (99 female, 100, 1 non-identified; \(age_\text{mean}\): 41.88, \(age_\text{sd}\): 13.94, \(age_\text{median}\): 39).
@@
 Participants answered with a slider on a scale from 0 to 100.
-To assess perceptions of competence, we asked: ``How competent do you think advisor 1 is regarding such investment recommendations?'' Participants answered on a 7-point Likert scale (from (1)``not competent at all'' to (7)``extremely competent'').
+To assess perceptions of competence, we asked: ``How competent do you think advisor 1 is regarding such investment recommendations?'' Participants answered on a 7-point Likert scale (from (1)``not competent at all'' to (2)``extremely competent'').
 
@@
 In the conflict of interest condition, experts were introduced this way: ``The three advisors have already invested in one of the three options, the same option for all three.
-As a result, they have an incentive to push that option in their recommendations.'' Participants assigned to the independence condition read: ``The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.''
+As a result, they have an incentive to push that option in their recommendations.'' For the independence condition: ``The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.''
 
@@
 \ref{fig:exp5-plot}).
-To test H1a and H1b, we use the same analyses as in Experiment 4, restricted on the independence condition, and replicate the results.
+To test H1a and H1b, we use the same analyses as in Experiment 4, restricted on the independent condition, and replicate the results.
 We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 12.34 {[}10.362, 14.311{]}, p = \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.56 {[}0.459, 0.665{]}, p = \textless{} .001).
@@
 
-\section{Experiment 6}\label{experiment-6}
+\subsection{Experiment 6}\label{experiment-6}
 
@@
 The number of choice options was manipulated between participants.
-Participants were randomly assigned to either see stimuli with three options (as in Experiment 4), or stimuli with ten options.
+Participants were randomly assigned to either to see stimuli with three options (as in Experiment 4), or with ten options.
 Participants assigned to the ten options condition were divided into one of two distinct sub-conditions: one in which the range of the answers corresponds to the range of the three options condition, and another with increased range (see ESM, section \ref{exp6}).
@@
 
-We tested whether the effect of convergence is larger on accuracy than on competence in an exploratory, non-preregistered analysis.
+We tested whether the effect of convergence is larger on accuracy than on competence, in an exploratory, non-preregistered analysis.
 To this end, we first standardized both outcome scores to account for the different scales.
 We then regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for the number of choice options.
-In line with all previous experiments, we find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy (\(\hat{b}\) = -0.05 {[}-0.075, -0.023{]}, \textless{} .001).
+In line with all previous experiments, find a negative interaction, indicating that convergence had a smaller effect on competence than on accuracy (\(\hat{b}\) = -0.05 {[}-0.075, -0.023{]}, \textless{} .001).
 
@@
 
-Using both analytical arguments and simulations, we have shown that, under a wide range of parameters, more convergent answers tend to be more accurate, and to have been made by more competent informants.
+Using both analytical arguments and simulations, we showed that, under a wide range of parameters, more convergent answers tend to be more accurate, and to have been made by more competent informants.
 
@@
 
-We also observe that, in line with our simulations, participants draw stronger inferences from convergence to accuracy than to competence in five of the six experiments.
-
-On the whole, participants thus appear to draw normatively justified inferences: 1.
-They infer that more convergent answers are more accurate; 2.
-They infer that more convergent answers are coming from more competent informants; 3.
-The inference on accuracy tends to be stronger than the inference on competence ; 4.
-Both the inference on accuracy and on the inference on competence are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other.
+We also observe that, in line with the simulations, participants draw stronger inferences from convergence to accuracy than to competence in five of the six experiments.
+
+On the whole, participants thus appear to draw normatively justified inferences: (i) they infer that more convergent answers are more accurate; (ii) they infer that more convergent answers are coming from more competent informants; (iii) inference (i) tends to be stronger than inference (ii); (iv) both inference (i) and (ii) are weaker when the informants have a conflict of interest, but not when they are merely discussing with each other.
 The only exception appears to be that participants do not take into account the number of options in categorical choices scenarios.
@@
 This inference has been evidenced in several experiments, both for numerical estimates (Budescu \& Rantilla, 2000; Budescu et al., 2003; Budescu \& Yu, 2007; e.g.
-Molleman et al., 2020; Yaniv et al., 2009), and for categorical choices (e.g., Morgan et al., 2012; for review, see H.
-Mercier \& Morin, 2019).
-However, these experiments do not test whether this inference of accuracy of the information extends to an inference of competence of the informants.
-Moreover, by their design, participants arguably assumed a degree of competence among the informants.
+Molleman et al., 2020; Yaniv et al., 2009), and for categorical choices (e.g., Morgan et al., 2012; for review, see Mercier \& Morin, 2019).
+However, in these experiments, the participants arguably assumed a degree of competence among the individuals whose answers they saw.
 For instance, when children are confronted with several individuals who agree on how to name a novel object (e.g.
@@
 
-Here, we assumed no prior knowledge of individual competence, asking the question: if we see informants, whose competence is unknown, converge on an answer, is it rational to infer that this answer is more likely to be correct, and that the informants are likely to be competent?
-We have shown that the answer is yes on both counts--assuming there is no systematic bias among the informants.
-An analytical argument and a series of simulations revealed that, for both the numerical choice context and the categorical choice context, the more individuals agree on an answer, the more likely the answer is correct, and the more likely the individuals are competent, with the former effect being stronger than the latter.
+Here, we assume no prior knowledge of individual competence, asking the question: if we see a set of individuals, whose competence is unknown, converge on an answer, is it rational to infer that this answer is more likely to be correct, and that the individuals are likely to be competent?
+We show that the answer is yes on both counts--assuming there is no systematic bias among the individuals answering.
+An analytical argument and a series of simulations reveal that, for both the numerical choice context and the categorical choice context, the more individuals agree on an answer, the more likely the answer is to be correct, and the more likely the individuals are competent, with the former effect being stronger than the latter.
 Moreover, this is true for a wide range of assumed population distributions of competence.
-In a series of experiments, we have shown that participants (UK) draw these inferences, but that they do so less when there is reason to assume a bias among informants.
+
+In a series of experiments, we show that participants (UK) draw these inferences: when presented with more convergent answers, they tend to believe the answers are more likely to be correct, and that the individuals who made them are more likely to be competent.
+This is true for numerical estimates and for categorical answers.
+These beliefs are weakened when the individuals making the estimates are systematically biased by a shared conflict of interest (Experiments 3 and 5), but not by a potential source of dependence between the individuals (discussion, Experiment 2).
 
@@
 From that competence, it can be inferred that the individuals will tend to agree, and that their answers will tend to be accurate.
-Here, we have shown that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be a valid cue when determining who tends to be an expert.
+Here, we show that the reverse inference--from agreement to competence--is also warranted, and that it is warranted under a wide range of circumstances: If one does not suspect any systematic bias, convergence alone can be a valid cue when determining who tends to be an expert.
 This finding qualifies work suggesting that people need a certain degree of expertise themselves, in order to figure out who is an expert (Hahn, Merdes, \& Sydow, 2018; Nguyen, 2020).
 
-The present experiments show that people draw inferences from convergence to accuracy and to competence, but then do not precisely show how they do it.
-As suggested in the introduction, it is plausible that participants use the combination of two heuristics: one leading them from convergence to accuracy, and one from accuracy to competence.
-This two step process is coherent with the observation that the inference from convergence to accuracy (one step) is stronger than that from convergence to competence (two steps).
-However, our results are also compatible with participants performing two inferences from convergence, one to accuracy and one to competence, the latter being weaker because the inferences roughly follow the normative model.
-Our results are coherent with work on other mechanisms of epistemic vigilance that process the combination of several informants' opinions, showing that these mechanisms rely on sensible cues, but also systematically fail to take some subtle cues into account (H.
-Mercier \& Morin, 2019).
-
-A most prominent context in which the inferences uncovered here might play a role is that of science.
-Much of science is counterintuitive, and most people do not have the background knowledge to evaluate most scientific evidence.
-However, science is, arguably, the institution in which individuals end up converging the most in their opinions (on consensus being the defining trait of science by contrast with other intellectual enterprises, see Collins, 2002).
+People might draw these inferences from convergence in a variety of contexts, but the most prominent one might be science.
+Science is, arguably, the institution in which individuals end up converging the most in their opinions.
 For instance, scientists within many disciplines agree on things ranging from the distance between the solar system and the center of the galaxy to the atomic structure of DNA.
@@
 The current study has a number of limitations.
-In our simulations, we assume agents to be independent and unbiased.
+In our simulations, we assume agents to be independent.
 Following previous work generalizing the Condorcet Jury Theorem to cases of informational dependency (Ladha, 1992), more robust simulations would show that--while still assuming no systematic bias--our results hold even when agents influence each others' answers.
@@
 Although the main results replicate well across our experiments, and we can thus be reasonably certain of their robustness, it's not clear how much they can be generalized.
-Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from--here, UK citizens (Coppock, 2019).
+Experimental results with convenience samples can usually be generalized at least to the broader population the samples were drawn from (here, UK citizens) (Coppock, 2019).
 However, we do not know whether they would generalize to other cultures.
 
-These limitations could be overcome by replicating the present results in different cultures, using more ecologically valid stimuli.
+Future studies could overcome these limitations by replicating the present results in different cultures, using more ecologically valid stimuli.
 For instance, it would be interesting to test whether the inference described here, from convergence to competence, might be partly responsible for the fact that people tend to believe scientists to be competent (Cologna et al., 2024).
-Finally, future studies could also attempt to systematically model cases of informational dependencies more subtle than those used in Experiments 3 and 5, to derive and test normative predictions.
 
@@
 
-\bibitem[\citeproctext]{ref-collinsSociologyPhilosophiesGlobal2002}
-Collins, R.
-(2002).
-\emph{The sociology of philosophies: a global theory of intellectual change} (4.
-print., 1.
-Harvard Univ.
-Pr.
-paperback ed., 2000).
-Cambridge, Mass.
-London: Belknap Press of Harvard Univ.
-Press.
-
 \bibitem[\citeproctext]{ref-colognaTrustScientistsTheir2024}
@@
 
-\bibitem[\citeproctext]{ref-desaiGettingSourceIllusion2022}
-Desai, S. C., Xie, B., \& Hayes, B. K. (2022). Getting to the source of the illusion of consensus. \emph{Cognition}, \emph{223}, 105023. \url{https://doi.org/10.1016/j.cognition.2022.105023}
-
 \bibitem[\citeproctext]{ref-dietrichEpistemicDemocracyDefensible2013}
 Dietrich, F., \& Spiekermann, K. (2013). Epistemic Democracy with Defensible Premises. \emph{Economics and Philosophy}, \emph{29}(1), 87--120. \url{https://doi.org/10.1017/S0266267113000096}
-
-\bibitem[\citeproctext]{ref-dietrichDeliberationWisdomCrowds2024}
-Dietrich, F., \& Spiekermann, K. (2024). Deliberation and the wisdom of crowds. \emph{Economic Theory}. \url{https://doi.org/10.1007/s00199-024-01595-4}
 
@@
 
-\bibitem[\citeproctext]{ref-kammerSystematicReviewEmpirical2023}
-Kämmer, J.
-E., Choshen-Hillel, S., Müller-Trede, J., Black, S.
-L., \& Weibler, J.
-(2023).
-A systematic review of empirical studies on advice-based decisions in behavioral and organizational research.
-\emph{Decision}, \emph{10}(2), 107--137.
-\url{https://doi.org/10.1037/dec0000199}
-
-\bibitem[\citeproctext]{ref-koenigTrustTestimonyChildren2004}
-Koenig, M.
-A., Clément, F., \& Harris, P.
-L.
-(2004).
-Trust in Testimony: Children's Use of True and False Statements.
-\emph{Psychological Science}, \emph{15}(10), 694--698.
-\url{https://doi.org/10.1111/j.0956-7976.2004.00742.x}
-
 \bibitem[\citeproctext]{ref-ladhaCondorcetJuryTheorem1992}
@@
 
-\bibitem[\citeproctext]{ref-mercierNotBornYesterday2020}
-Mercier, Hugo. (2020). \emph{Not born yesterday: the science of who we trust and what we believe}. Retrieved from \url{https://doi.org/10.1515/9780691198842}
-
 \bibitem[\citeproctext]{ref-mercierDoesDiscussionMake2022}
 Mercier, H., \& Claidière, N. (2022). Does discussion make crowds any wiser? \emph{Cognition}, \emph{222}, 104912. \url{https://doi.org/10.1016/j.cognition.2021.104912}
-
-\bibitem[\citeproctext]{ref-mercierUtilizingSimpleCues2019}
-Mercier, Hugo, \& Miton, H.
-(2019).
-Utilizing simple cues to informational dependency.
-\emph{Evolution and Human Behavior}, \emph{40}(3), 301--314.
-\url{https://doi.org/10.1016/j.evolhumbehav.2019.01.001}
 
@@
 
-\bibitem[\citeproctext]{ref-pillowEarlyUnderstandingPerception1989}
-Pillow, B.
-H.
-(1989).
-Early understanding of perception as a source of knowledge.
-\emph{Journal of Experimental Child Psychology}, \emph{47}(1), 116--129.
-\url{https://doi.org/10.1016/0022-0965(89)90066-0}
-
-\bibitem[\citeproctext]{ref-richardsonThinkingTakesTime2022}
-Richardson, E., \& Keil, F.
-C.
-(2022).
-Thinking takes time: Children use agents' response times to infer the source, quality, and complexity of their knowledge.
-\emph{Cognition}, \emph{224}, 105073.
-\url{https://doi.org/10.1016/j.cognition.2022.105073}
-
 \bibitem[\citeproctext]{ref-romeijnLearningJurorCompetence2011}
@@
 \url{https://doi.org/10.1177/1470594X10372317}
-
-\bibitem[\citeproctext]{ref-smithAnimalSignals2003}
-Smith, J. M., \& Harper, D. (2003). \emph{Animal signals}. Oxford University Press.
 
@@
 
-\bibitem[\citeproctext]{ref-sperberEpistemicVigilance2010}
-Sperber, D., Clément, F., Heintz, C., Mascaro, O., Mercier, H., Origgi, G., \& Wilson, D. (2010). Epistemic vigilance. \emph{Mind \& Language}, \emph{25}(4), 359--393.
-
 \bibitem[\citeproctext]{ref-vanstekelenburgScientificConsensusCommunicationContested2022}
@@
 
-\bibitem[\citeproctext]{ref-xieSensitivityEvidentialDependencies2022}
-Xie, B., \& Hayes, B. (2022). Sensitivity to Evidential Dependencies in Judgments Under Uncertainty. \emph{Cognitive Science}, \emph{46}(5). \url{https://doi.org/10.1111/cogs.13144}
-
 \bibitem[\citeproctext]{ref-yanivWeightingTrimmingHeuristics1997}
@@
 \url{https://doi.org/10.1006/obhd.2000.2909}
-
-\bibitem[\citeproctext]{ref-yinSmartlyFollowingOthers2024}
-Yin, J., Xu, Z., Lin, J., Zhou, W., \& Guo, X.
-(2024).
-Smartly following others: Majority influence depends on how the majority behavior is formed.
-\emph{Journal of Experimental Social Psychology}, \emph{115}, 104644.
-\url{https://doi.org/10.1016/j.jesp.2024.104644}
-
-\bibitem[\citeproctext]{ref-yousifIllusionConsensusFailure2019}
-Yousif, S.
-R., Aboody, R., \& Keil, F.
-C.
-(2019).
-The Illusion of Consensus: A Failure to Distinguish Between True and False Consensus.
-\emph{Psychological Science}, \emph{30}(8), 1195--1204.
-\url{https://doi.org/10.1177/0956797619856844}
 
@@
 
-\subsection{Minor deviation from preregistration}\label{minor-deviation-from-preregistration}
-
-In our preregistration, we stated we would run paired t-tests and only conduct mixed models as a robustness check.
-However, it became clear to us later that mixed models are more appropriate in light of our data structure.
-In the main article, we therefore only report the results of the mixed models.
-Note, however, that using paired t-tests, we confirm our results (Accuracy: -1.37, t-statistic = -26.14, p \textless{} .001; Competence: -1.23, t-statistic = -21.82, p \textless{} .001)
-
-\subsection{Sensitivity simulation}\label{sensitivity-simulation}
-
-We had based our power calculations on a paired t-test (using the widely used tool G*Power), which does not correspond to the mixed-models we eventually ran.
-To address this point, here, we seek to identify the minimum effect size that our analysis with mixed models could have detected, given a 90\% power, an N of 200 participants, and other parameters derived from the estimates observed in the sample of Experiment 1 (see \ref{tab:exp1-regression-table}).
-By contrast to a standard power simulation (which varies sample size), a sensitivity simulation varies the effect size, while holding the sample size and desired power constant.
-
-\begin{table}
-
-\caption{\label{tab:exp1-regression-table}Model Results Experiment 1}
-\centering
-\begin{tabular}[t]{lcc}
-\toprule
-  & Accuracy & Competence\\
-\midrule
-(Intercept) & 3.105 & 3.516\\
- & (0.077) & (0.068)\\
-convergenceconvergent & 1.368 & 1.232\\
- & (0.072) & (0.085)\\
-numberlarge & 0.175 & 0.002\\
- & (0.050) & (0.048)\\
-SD (Intercept ID) & 0.905 & 0.764\\
-SD (convergenceconvergent ID) & 0.732 & 0.993\\
-Cor (Intercept\textasciitilde{}convergenceconvergent ID) & -0.128 & -0.554\\
-SD (Observations) & 1.006 & 0.955\\
-\midrule
-Num.Obs. & 1600 & 1600\\
-R2 Marg. & 0.191 & 0.195\\
-R2 Cond. & 0.594 & 0.532\\
-AIC & 5131.0 & 4950.5\\
-BIC & 5168.7 & 4988.1\\
-ICC & 0.5 & 0.4\\
-RMSE & 0.91 & 0.86\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{figure}
-\centering
-\includegraphics{output/figures/sensitivity.pdf}
-\caption{\label{fig:sensitivity}Results of the sensitivity analysis.}
-\end{figure}
-
-The results of this can be seen in Fig.
-\ref{fig:sensitivity}.
-In our sample, we found a statistically significant effect size of 1.37 for accuracy and an effect size of 1.23 for competence (both on their original scales from 1 to 7).
-The simulations show that, given our parameter assumptions (based on the observations in our sample) and a sample size of N = 200, we would have detected an effect as small as 0.4 with a power of at least 90\% (in fact nearly 100\%) for both accuracy and competence.
-
 \clearpage
@@
 Switching from normal distributions to uniform distributions made `unlucky' draws in which conditions are visually not distinguishable less likely.
-And additional difference to experiment one was that all scenarios only involved groups of three informants (never ten).
 
@@
 \includegraphics{output/figures/exp4-plot.pdf}
-\caption{\label{fig:exp4-plot}Results of Experiment 4, showing the distributions of accuracy and competence by level of convergence.}
+\caption{\label{fig:exp4-plot}Distributions of accuracy and competence by level of convergence.}
 \end{figure}
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0993}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4503}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4503}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1447}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4277}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4277}}@{}}
 \caption{\label{tab:stimuli-10}Stimuli for 10 options condition by levels of convergence}\tabularnewline
@@
 \endlastfoot
-minority (0) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_b.png} \\
+opposing majority (0) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_b.png} \\
 dissensus (1) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_b.png} \\
 majority (2) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_b.png} \\
-consensus (3) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_b.png} \\
+consensus (10) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_a.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_b.png} \\
 \end{longtable}
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.0943}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4528}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4528}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.1377}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4311}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.4311}}@{}}
 \caption{\label{tab:stimuli-10-alt}Alternative stimuli for 10 options condition by levels of convergence}\tabularnewline
@@
 \endlastfoot
-minority (0) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_b_alt.png} \\
+opposing majority (0) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/minority_10_b_alt.png} \\
 dissensus (1) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/divergence_10_b_alt.png} \\
 majority (2) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/majority_10_b_alt.png} \\
-consensus (3) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_b_alt.png} \\
+consensus (10) & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_a_alt.png} & \includegraphics[width=0.4\textwidth,height=\textheight]{Experiment_6/figures/stimuli/consensus_10_b_alt.png} \\
 \end{longtable}
@@
 
-\caption{\label{tab:unnamed-chunk-101}Participants vs. Model (Exp. 6)}
+\caption{\label{tab:unnamed-chunk-91}Participants vs. Model (Exp. 6)}
 \centering

### 2024-11-04 paper.Rmd
update paper revisions round 1
before proof, make compatible with up-to-date R version

@@
 
+author: 
+  - name          : "Jan Pfänder"
+    affiliation   : "1"
+
+  - name          : "Benoît De Courson"
+    affiliation   : "2"
+
+  - name          : "Hugo Mercier"
+    affiliation   : "1"
+    corresponding : yes    # Define only one corresponding author
+    email         : "hugo.mercier@gmail.com"
+    address: " "
+
 affiliation:
-  - id            : " "
-    institution   : " "
+  - id            : "1"
+    institution   : "Institut Jean Nicod, Département d’études cognitives, ENS, EHESS, PSL University, CNRS, France"
+  - id            : "2"
+    institution   : "Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany"
+
+authornote: |
+  We thank the anonymous reviewers for their very valuable feedback.
+HM received funding from the ANR (SCALUP, ANR-21-CE28-0016-01), and from the John Tempelton Foundation ("An Evolutionary and Cultural Perspective on Intellectual Humility via Intellectual Curiosity and Epistemic Deference").
+The ANR also supported this project through FrontCog (ANR-17-EURE-0017), and PSL (ANR-10-IDEX-0001-02).
 
@@
 
-keywords          :
+keywords          : consensus, convergence, majority rules, competence, trust in science
 
@@
 
+\newpage
+
 # Introduction
@@
 # RQ2 (as separate t-test):
-accuracy <- t.test(confidence ~ number, data = exp1, paired = TRUE)
-competence <- t.test(competence ~ number, data = exp1, paired = TRUE)
+# accuracy <- t.test(confidence ~ number, data = exp1, paired = TRUE) 
+# competence <- t.test(competence ~ number, data = exp1, paired = TRUE)
 
@@
     split(.$number), 
-  accuracy = clean_t_test(accuracy), 
-  competence = clean_t_test(competence),
+  # accuracy = clean_t_test(accuracy), 
+  # competence = clean_t_test(competence),
   # RQ2 and RQ3 in a single mixed model
@@
 Each participant rated four scenarios, two for each level of convergence.
-By contrast to experiment one, all scenarios only involved groups of three informants.
+By contrast to Experiment 1, all scenarios only involved groups of three informants.
 
@@
 
-```{=tex}
 \clearpage
 \newpage
-```
+
 ```{r child = "appendix_simulations.Rmd"}

### 2024-11-04 paper.tex
before proof, make compatible with up-to-date R version

@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 % definitions for citeproc citations
@@
   {\xapptocmd\section
-    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
+    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi: #1}}
     {}{\InnerPatchFailed}%
@@
 {}{\PatchFailed}
+\makeatother
+\keywords{consensus, convergence, majority rules, competence, trust in science}
 \usepackage{csquotes}
@@
   pdftitle={How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?},
+  pdfauthor={Jan Pfänder1, Benoît De Courson2, \& Hugo Mercier1},
   pdflang={en-EN},
+  pdfkeywords={consensus, convergence, majority rules, competence, trust in science},
   hidelinks,
@@
 \title{How wise is the crowd: Can we infer people are accurate and competent merely because they agree with each other?}
-\author{\phantom{0}}
+\author{Jan Pfänder\textsuperscript{1}, Benoît De Courson\textsuperscript{2}, \& Hugo Mercier\textsuperscript{1}}
 \date{}
@@
 
-\affiliation{\vspace{0.5cm}\textsuperscript{} }
+\authornote{
+
+We thank the anonymous reviewers for their very valuable feedback.
+HM received funding from the ANR (SCALUP, ANR-21-CE28-0016-01), and from the John Tempelton Foundation (``An Evolutionary and Cultural Perspective on Intellectual Humility via Intellectual Curiosity and Epistemic Deference'').
+The ANR also supported this project through FrontCog (ANR-17-EURE-0017), and PSL (ANR-10-IDEX-0001-02).
+
+Correspondence concerning this article should be addressed to Hugo Mercier, . E-mail: \href{mailto:hugo.mercier@gmail.com}{\nolinkurl{hugo.mercier@gmail.com}}
+
+}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Institut Jean Nicod, Département d'études cognitives, ENS, EHESS, PSL University, CNRS, France\\\textsuperscript{2} Max Planck Institute for the Study of Crime, Security and Law, Freiburg im Breisgau, Germany}
 
@@
 
+\newpage
+
 \section{Introduction}\label{introduction}
@@
 In categorical choice contexts, there is ample and long-standing (e.g.
-Crutchfield, 1955) evidence from experimental psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see H.
-Mercier \& Morin, 2019).
+Crutchfield, 1955) evidence from experimental psychology that participants are more likely to be influenced by majority opinions, and that this influence is stronger when the majority is larger, both in absolute and in relative terms (e.g., Morgan, Rendell, Ehn, Hoppitt, \& Laland, 2012; for review, see Mercier \& Morin, 2019).
 This is true even if normative conformity (when people follow the majority because of social pressure rather than a belief that the majority is correct) is unlikely to play an important role (e.g.~because the answers are private, see Asch, 1956).
@@
 In order to make the best of communicated information, humans have to be able to evaluate it, so as to discard inaccurate or harmful information, while accepting accurate and beneficial information (Smith \& Harper, 2003).
-It has been argued that a suite of cognitive mechanisms--mechanisms of epistemic vigilance--evolved to serve this function (Hugo Mercier, 2020; Sperber et al., 2010).
+It has been argued that a suite of cognitive mechanisms--mechanisms of epistemic vigilance--evolved to serve this function (Mercier, 2020; Sperber et al., 2010).
 Since the opinion of more than one individual is often available to us, there should be mechanisms of epistemic vigilance dedicated to processing such situations.
-It would be these mechanisms that lead us to put more weight on an opinion that is shared by a larger majority (in relative or absolute terms), and, in some cases at least, to discount majority opinion when the opinions haven't been formed independently of each other (Hugo Mercier \& Miton, 2019).
-Evidence suggests that these mechanisms rely on heuristics which become more refined with age (Morgan et al., 2015), and which are far from perfect (Hugo Mercier \& Miton, 2019; in particular, they ignore many cases of informational dependencies, see, e.g.
+It would be these mechanisms that lead us to put more weight on an opinion that is shared by a larger majority (in relative or absolute terms), and, in some cases at least, to discount majority opinion when the opinions haven't been formed independently of each other (Mercier \& Miton, 2019).
+Evidence suggests that these mechanisms rely on heuristics which become more refined with age (Morgan et al., 2015), and which are far from perfect (Mercier \& Miton, 2019; in particular, they ignore many cases of informational dependencies, see, e.g.
 Yousif, Aboody, \& Keil, 2019).
@@
 By contrast with previous studies, participants were not given any information about the tasks--how difficult they were--and the informants--how competent they might be.
-There has recently been much interest in investigating whether participants are able to take informational dependencies into account when evaluating convergent information (Desai, Xie, \& Hayes, 2022; Hugo Mercier \& Miton, 2019; Xie \& Hayes, 2022; Yin, Xu, Lin, Zhou, \& Guo, 2024; Yousif et al., 2019).
+There has recently been much interest in investigating whether participants are able to take informational dependencies into account when evaluating convergent information (Desai, Xie, \& Hayes, 2022; Mercier \& Miton, 2019; Xie \& Hayes, 2022; Yin, Xu, Lin, Zhou, \& Guo, 2024; Yousif et al., 2019).
 To test whether participants took into account dependencies between the informants in the present context, this factor was manipulated in two different ways (in Experiment 2, some informants had discussed with each other; in Experiments 3 and 5, some informants had an incentive to provide the same answer).
@@
 Contrary to the hypotheses, participants did not deem convergent estimates made after a discussion, compared to independently made estimates, to be less accurate, or produced by less competent individuals.
-This might stem from the fact that participants, in various situations, neglect informational dependencies (Yousif et al., 2019), or from the fact that discussing groups actually perform better than non-discussing groups in a range of tasks (for review, see, e.g., H.
-Mercier, 2016), including numerical estimates (e.g.
-H.
+This might stem from the fact that participants, in various situations, neglect informational dependencies (Yousif et al., 2019), or from the fact that discussing groups actually perform better than non-discussing groups in a range of tasks (for review, see, e.g., Mercier, 2016), including numerical estimates (e.g.
 Mercier \& Claidière, 2022).
@@
 This inference has been evidenced in several experiments, both for numerical estimates (Budescu \& Rantilla, 2000; Budescu et al., 2003; Budescu \& Yu, 2007; e.g.
-Molleman et al., 2020; Yaniv et al., 2009), and for categorical choices (e.g., Morgan et al., 2012; for review, see H.
-Mercier \& Morin, 2019).
+Molleman et al., 2020; Yaniv et al., 2009), and for categorical choices (e.g., Morgan et al., 2012; for review, see Mercier \& Morin, 2019).
 However, these experiments do not test whether this inference of accuracy of the information extends to an inference of competence of the informants.
@@
 However, our results are also compatible with participants performing two inferences from convergence, one to accuracy and one to competence, the latter being weaker because the inferences roughly follow the normative model.
-Our results are coherent with work on other mechanisms of epistemic vigilance that process the combination of several informants' opinions, showing that these mechanisms rely on sensible cues, but also systematically fail to take some subtle cues into account (H.
-Mercier \& Morin, 2019).
+Our results are coherent with work on other mechanisms of epistemic vigilance that process the combination of several informants' opinions, showing that these mechanisms rely on sensible cues, but also systematically fail to take some subtle cues into account (Mercier \& Morin, 2019).
 
@@
 \bibitem[\citeproctext]{ref-mercierNotBornYesterday2020}
-Mercier, Hugo. (2020). \emph{Not born yesterday: the science of who we trust and what we believe}. Retrieved from \url{https://doi.org/10.1515/9780691198842}
+Mercier, H. (2020). \emph{Not born yesterday: the science of who we trust and what we believe}. Retrieved from \url{https://doi.org/10.1515/9780691198842}
 
@@
 \bibitem[\citeproctext]{ref-mercierUtilizingSimpleCues2019}
-Mercier, Hugo, \& Miton, H.
+Mercier, H., \& Miton, H.
 (2019).
@@
 In the main article, we therefore only report the results of the mixed models.
-Note, however, that using paired t-tests, we confirm our results (Accuracy: -1.37, t-statistic = -26.14, p \textless{} .001; Competence: -1.23, t-statistic = -21.82, p \textless{} .001)
+Note, however, that using paired t-tests, we confirm our results (Accuracy: 1.37, t-statistic = 26.14, p \textless{} .001; Competence: 1.23, t-statistic = 21.82, p \textless{} .001)
 
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:exp1-regression-table}Model Results Experiment 1}
@@
 
-\subsection{Design}\label{design-6}
-
 \FloatBarrier
-
-We manipulated informational dependency. According to the condition, players read a different introduction before seeing a set of estimates (see Table \ref{tab:exp2-conditions}).
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp2-conditions}}
-
-\begin{tabular}{m{3cm}m{12cm}}
-\toprule
-Condition & \multicolumn{1}{c}{Description}\\
-\midrule
-Independence & Players are asked to make completely independent decisions – they cannot see each other’s estimates, or talk with each other before giving their estimates.\\
-Discussion & Players are asked to talk with each other about the game at length before giving their estimates.\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
 
@@
 
-\subsection{Design}\label{design-7}
-
 \FloatBarrier
-
-We manipulated two factors: informational dependency (two levels, independence and conflict of interest, see Table \ref{tab:conditions-exp3}; between participants) and convergence (two levels, convergence and divergence; within participants).
-Participants saw four scenarios, one for each combination of the two factors convergence and informational dependency.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:conditions-exp3}}
-
-\begin{tabular}{m{3cm}m{12cm}}
-\toprule
-Condition & \multicolumn{1}{c}{Description}\\
-\midrule
-Independence & Experts are independent of each other, and have no conflict of interest in predicting the stock value - they do not personally profit in any way from any future valuation of the stock.\\
-Conflict of interest & All three experts have invested in the specific stock whose value they are predicting, and they benefit if other people believe that the stock will be valued at [mean of respective distribution] in the future.\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
 
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:exp3-table}}
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:model-vs-participant-table}Participants vs. Model (Exp. 4)}
@@
 Log.Lik. &  &  & -1650898.534 & -525734.261\\
-F &  &  & 85844.746 & 16470.902\\
 RMSE & 8.98 & 0.49 & 34.42 & 1.17\\
@@
 
-\subsection{Design}\label{design-8}
-
 \FloatBarrier
-
-We manipulated convergence within participants in the same way we did in experiment 4.
-In addition, between participants, we manipulated informational dependence, akin to experiment 3 (Table \ref{tab:exp5-condition}).
-In the biased condition, experts were described to gain personally from recommending a certain investment option - but without specifying what that option is.
-In the independent condition, there was no such conflict of interest and experts were described as independent.
-
-\begin{table}[tbp]
-
-\begin{center}
-\begin{threeparttable}
-
-\caption{\label{tab:exp5-condition}}
-
-\begin{tabular}{m{3cm}m{12cm}}
-\toprule
-Condition & \multicolumn{1}{c}{Description}\\
-\midrule
-Independence condition & The three advisors are independent of each other, and have no conflict of interest in making investment recommendations.\\
-Conflict of interest condition & The three advisors have already invested in one of the three options, the same option for all three.
-As a result, they have an incentive to push that option in their recommendations.\\
-\bottomrule
-\end{tabular}
-
-\end{threeparttable}
-\end{center}
-
-\end{table}
 
@@
 \begin{table}
-
+\centering
 \caption{\label{tab:exp5-table}}
 \centering
-\resizebox{\linewidth}{!}{
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
 \begin{tabular}[t]{lllll}
@@
 
-\subsection{Design}\label{design-9}
+\subsection{Design}\label{design-6}
 
@@
 \begin{table}
-
+\centering\centering
 \caption{\label{tab:exp6-table}}
 \centering
-\resizebox{\linewidth}{!}{
+\resizebox{\ifdim\width>\linewidth\linewidth\else\width\fi}{!}{
 \begin{tabular}[t]{lllll}
@@
 \begin{table}
-
-\caption{\label{tab:unnamed-chunk-101}Participants vs. Model (Exp. 6)}
+\centering
+\caption{\label{tab:unnamed-chunk-98}Participants vs. Model (Exp. 6)}
 \centering
@@
 Log.Lik. &  &  & -248866.472 & -85091.109\\
-F &  &  & 19902.699 & 3370.232\\
 RMSE & 10.67 & 0.53 & 31.83 & 1.28\\

### 2024-11-24 paper.tex
update some p-value reporting after proofs, removing "=" when smaller than 0.001

@@
 \caption{Results of the analytic argument.
-The figure shows the average estimated competence and accuracy for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario.
-Here, we assume that 20 individuals answer a 5-choice question.
-For values higher than four, we assume that the focal individual is part of the majority.
-Below four, we assume that the focal individual is part of a minority, and that the majority size is four.
-Two situations are represented: one in which there is low variance in the distribution of competence in the population (orange line in the insert), and one in which there is high variance in this distribution (blue line in the insert).}\label{fig:figure-Benoit-2}
+The figure shows the average estimated competence for a focal individual, depending on the proportion of individuals agreeing with them, in a categorical choice scenario.
+Here, we assume that 20 individuals independently pick one out of 5 choice options.
+Two situations are represented: one in which there is low variance in the distribution of competence in the population (the bell-shaped curve; solid line), and one in which there is high variance in this distribution (the u-shaped curve; dotted line).}\label{fig:figure-Benoit-2}
 \end{figure}
@@
 Visualizations and descriptive statistics can be found in ESM, section \ref{exp1}.
-We find a positive effect of convergence on accuracy: Participants were more confident about their estimate in convergent scenarios (mean = 4.56, sd = 1.45) than in divergent ones (mean = 3.19, sd = 1.39; \(\hat{b}_{\text{Accuracy}}\) = 1.37 {[}1.225, 1.51{]}, p = \textless{} .001).
-We also find a positive effect of convergence on competence: participants rated players as more competent in convergent scenarios (mean = 4.75, sd = 1.24) than in divergent ones (mean = 3.52, sd = 1.27; \(\hat{b}_{\text{Competence}}\) = 1.23 {[}1.065, 1.4{]}, p = \textless{} .001).
+We find a positive effect of convergence on accuracy: Participants were more confident about their estimate in convergent scenarios (mean = 4.56, sd = 1.45) than in divergent ones (mean = 3.19, sd = 1.39; \(\hat{b}_{\text{Accuracy}}\) = 1.37 {[}1.225, 1.51{]}, p \textless{} .001).
+We also find a positive effect of convergence on competence: participants rated players as more competent in convergent scenarios (mean = 4.75, sd = 1.24) than in divergent ones (mean = 3.52, sd = 1.27; \(\hat{b}_{\text{Competence}}\) = 1.23 {[}1.065, 1.4{]}, p \textless{} .001).
 
@@
 To this end, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for the number of informants.
-We do not find a statistically significant interaction that would indicate a difference of the effect of convergence (\(\hat{b}\) = 0.14 {[}-0.001, 0.271{]}, 0.052).
+We do not find a statistically significant interaction that would indicate a difference of the effect of convergence (\(\hat{b}\) = 0.14 {[}-0.001, 0.271{]}, = 0.052).
 Pooled across divergent and convergent conditions, we find that participants reported lower perceived accuracy than competence (\(\hat{b}\) = -0.26 {[}-0.359, -0.156{]}, \textless{} .001)
@@
 On this reduced sample, we ran the exact same analyses as in Experiment 1 and replicated the results.
-As for accuracy, participants rated informants in convergent scenarios (mean = 5.28, sd = 1.05) as more accurate than in divergent ones (mean = 3.40, sd = 1.08; \(\hat{b}_{\text{Accuracy}}\) = 1.88 {[}1.658, 2.102{]}, p = \textless{} .001).
-As for competence, participants rated informants in convergent scenarios (mean = 5.24, sd = 0.99) as more competent than in divergent ones (mean = 3.61, sd = 1.11; \(\hat{b}_{\text{Competence}}\) = 1.62 {[}1.411, 1.839{]}, p = \textless{} .001).
+As for accuracy, participants rated informants in convergent scenarios (mean = 5.28, sd = 1.05) as more accurate than in divergent ones (mean = 3.40, sd = 1.08; \(\hat{b}_{\text{Accuracy}}\) = 1.88 {[}1.658, 2.102{]}, p \textless{} .001).
+As for competence, participants rated informants in convergent scenarios (mean = 5.24, sd = 0.99) as more competent than in divergent ones (mean = 3.61, sd = 1.11; \(\hat{b}_{\text{Competence}}\) = 1.62 {[}1.411, 1.839{]}, p \textless{} .001).
 
@@
 \ref{fig:exp3-plot}).
-In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 0.99 {[}0.634, 1.348{]}, p = \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.89 {[}0.636, 1.142{]}, p = \textless{} .001).
-Likewise the effect of convergence on competence is more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.80 {[}0.474, 1.13{]}, p = \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.82 {[}0.591, 1.056{]}, p = \textless{} .001).
+In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 0.99 {[}0.634, 1.348{]}, p \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.89 {[}0.636, 1.142{]}, p \textless{} .001).
+Likewise the effect of convergence on competence is more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.80 {[}0.474, 1.13{]}, p \textless{} .001) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.82 {[}0.591, 1.056{]}, p \textless{} .001).
 
@@
 To this end, we regressed the outcome score on convergence and its interaction with a binary variable indicating which outcome was asked for (accuracy or competence), while controlling for informational dependency.
-We find a negative interaction effect, indicating that pooled across independent and conflict of interest conditions, the effect of convergence had as smaller effect on competence than on accuracy (\(\hat{b}\) = -0.16 {[}-0.294, -0.028{]}, 0.018).
-Pooled across all conditions, participants reported higher perceived competence than accuracy (\(\hat{b}\) = 0.11 {[}0.025, 0.191{]}, 0.011).
+We find a negative interaction effect, indicating that pooled across independent and conflict of interest conditions, the effect of convergence had as smaller effect on competence than on accuracy (\(\hat{b}\) = -0.16 {[}-0.294, -0.028{]}, = 0.018).
+Pooled across all conditions, participants reported higher perceived competence than accuracy (\(\hat{b}\) = 0.11 {[}0.025, 0.191{]}, = 0.011).
 
@@
 
-As in the numerical setting, we found a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 16.84 {[}15.009, 18.668{]}, p = \textless{} .001; on a scale from 0 to 100) and competence (\(\hat{b}_{\text{Competence}}\) = 0.68 {[}0.578, 0.788{]}, p = \textless{} .001; on a scale from 1 to 7).
+As in the numerical setting, we found a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 16.84 {[}15.009, 18.668{]}, p \textless{} .001; on a scale from 0 to 100) and competence (\(\hat{b}_{\text{Competence}}\) = 0.68 {[}0.578, 0.788{]}, p \textless{} .001; on a scale from 1 to 7).
 
@@
 To test H1a and H1b, we use the same analyses as in Experiment 4, restricted on the independence condition, and replicate the results.
-We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 12.34 {[}10.362, 14.311{]}, p = \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.56 {[}0.459, 0.665{]}, p = \textless{} .001).
+We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 12.34 {[}10.362, 14.311{]}, p \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.56 {[}0.459, 0.665{]}, p \textless{} .001).
 
@@
 \ref{fig:exp5-plot}).
-In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 3.01 {[}0.027, 5.988{]}, p = 0.048) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) 9.33 {[}7.232, 11.426{]}, p = \textless{} .001).
-Likewise, the effect of convergence on competence was more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.16 {[}0.014, 0.316{]}, p = 0.032) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.40 {[}0.291, 0.503{]}, p = \textless{} .001).
+In the independence condition, the effect of convergence on accuracy was more positive (\(\hat{b}_{\text{interaction, Accuracy}}\) = 3.01 {[}0.027, 5.988{]}, p = 0.048) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) 9.33 {[}7.232, 11.426{]}, p \textless{} .001).
+Likewise, the effect of convergence on competence was more positive (\(\hat{b}_{\text{interaction, Competence}}\) = 0.16 {[}0.014, 0.316{]}, p = 0.032) than in the conflict of interest condition (\(\hat{b}_{\text{baseline}}\) = 0.40 {[}0.291, 0.503{]}, p \textless{} .001).
 
@@
 Due to budget constraints, we considered aiming for a sample of 300 participants as good enough, although slightly below our threshold.
-Due to two failed attention checks, our final sample consisted of 298 subjects, recruited, as in all experiments, in the UK and via Prolific (149 female, 149, 1 non-identified; \(age_\text{mean}\): 42.09, \(age_\text{sd}\): 13.06, \(age_\text{median}\): 40).
+Due to two failed attention checks, our final sample consisted of 298 subjects, recruited, as in all experiments, in the UK and via Prolific (149 female, 149; \(age_\text{mean}\): 42.09, \(age_\text{sd}\): 13.06, \(age_\text{median}\): 40).
 
@@
 On this reduced sample, we ran the exact same analyses as in experiment 4 and replicated the results.
-We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 15.68 {[}14.112, 17.246{]}, p = \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.65 {[}0.564, 0.736{]}, p = \textless{} .001).
+We find a positive effect of convergence on both accuracy (\(\hat{b}_{\text{Accuracy}}\) = 15.68 {[}14.112, 17.246{]}, p \textless{} .001) and competence (\(\hat{b}_{\text{Competence}}\) = 0.65 {[}0.564, 0.736{]}, p \textless{} .001).
 This finding also holds across the entire sample, pooling three and ten choice option observations (\(\hat{b}_{\text{Accuracy}}\) = 16.30 {[}15.124, 17.485{]}, \textless{} .001; \(\hat{b}_{\text{Competence}}\) = 0.68 {[}0.611, 0.739{]}, \textless{} .001).
 
-We do not find evidence of an interaction, i.e.~evidence that the number of choice options changes the effect of convergence (\(\hat{b}_{\text{interaction, Accuracy}}\) = 1.25 {[}-1.11, 3.613{]}, 0.298; \(\hat{b}_{\text{interaction, Competence}}\) = 0.05 {[}-0.078, 0.178{]}, 0.442).
+We do not find evidence of an interaction, i.e.~evidence that the number of choice options changes the effect of convergence (\(\hat{b}_{\text{interaction, Accuracy}}\) = 1.25 {[}-1.11, 3.613{]}, = 0.298; \(\hat{b}_{\text{interaction, Competence}}\) = 0.05 {[}-0.078, 0.178{]}, = 0.442).
 
@@
 Regarding RQ2, pooling across convergent and divergent conditions, we find a main effect of number, such that participants had more confidence in their estimate when they relied on ten informants compared to three informants (\(\hat{b}_{\text{Accuracy}}\) = 0.18 {[}0.076, 0.274{]}, p = \textless{} .001).
-Regarding RQ3, we find an interaction between number of informants and convergence on competence: the positive effect of convergence was stronger in scenarios involving ten informants compared to three informants (\(\hat{b}_{\text{Competence}}\) = 0.22 {[}0.028, 0.402{]}, p = 0.024).
-We do not find a statistically significant interaction on accuracy (\(\hat{b}_{\text{Accuracy}}\) = 0.19 {[}-0.007, 0.387{]}, p = 0.059).
+Regarding RQ3, we find an interaction between number of informants and convergence on competence: the positive effect of convergence was stronger in scenarios involving ten informants compared to three informants (\(\hat{b}_{\text{Competence}}\) = 0.22 {[}0.028, 0.402{]}, p = = 0.024).
+We do not find a statistically significant interaction on accuracy (\(\hat{b}_{\text{Accuracy}}\) = 0.19 {[}-0.007, 0.387{]}, p = = 0.059).
 

### 2024-11-24 README.qmd
update README

@@
 
-*Estimated run time*: Rendering the `anonymized.Rmd` document takes about 2mins on our machine (see [session info](#session-info))
+*Estimated run time*: Rendering the `paper.Rmd` document takes about 2mins on our machine (see [session info](#session-info))
 
@@
 The `/data` folders contain the raw data (`qualtrics.csv`) and cleaned data (`cleaned.csv`).
-The cleaning can be reproduced by running the `cleaning.Rmd`.
+The cleaning can be reproduced by running the `cleaning.Rmd` files.
 `Experiment_1` and `Experiment_4` also contain the code for the simulations (`model.Rmd`) and several data files with simulated data.
 - `functions/` folder contains functions that we rely upon in our analysis scripts. To keep the `paper.Rmd` document neat and readable, we wrote all lengthy code bits as external functions.
-- figures includes images used in the `paper.Rmd` file
+- figures are generated in the code of the `paper.Rmd` file
 - `howwise_0131_initial_submission.pdf` is the initial, and `howwise_0131.pdf` the final submission (after review) to the CogSci 2024 conference, based on the raw files but edited in [overleaf](https://www.overleaf.com/).
-- `output` folder lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.
+- running the `paper.Rmd` file will generate an `output` folder which lists all figures generated by the `paper.Rmd` file as `.png` and `.pdf` files.
 - `paper.pdf` is the rendered version of `paper.Rmd`
